### 2015-04-22 tex/main.tex

%%%%%%%%%%%%%%%%%%%%%%% file typeinst.tex %%%%%%%%%%%%%%%%%%%%%%%%%
%
% This is the LaTeX source for the instructions to authors using
% the LaTeX document class 'llncs.cls' for contributions to
% the Lecture Notes in Computer Sciences series.
% http://www.springer.com/lncs       Springer Heidelberg 2006/05/04
%
% It may be used as a template for your own input - copy it
% to a new file with a new name and use it as the basis
% for your article.
%
% NB: the document class 'llncs' has its own and detailed documentation, see
% ftp://ftp.springer.de/data/pubftp/pub/tex/latex/llncs/latex2e/llncsdoc.pdf
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\documentclass[runningheads,a4paper]{llncs}
\usepackage{times}
\usepackage{listings}
\usepackage{colonequals}
%\usepackage{mathpartir}
\usepackage{xspace}
\usepackage{amssymb}
\setcounter{tocdepth}{3}
\usepackage{graphicx}
\usepackage{stmaryrd}
\usepackage{cite}
\usepackage{hyperref}
\usepackage{framed}

% http://mirror.switch.ch/ftp/mirror/tex/macros/latex/contrib/xcolor/xcolor.pdf
\usepackage{xcolor}

%\usepackage{amsthm}
\usepackage{amsmath}    % need for subequations
\usepackage{booktabs}
%\usepackage{float}
%\usepackage{fullpage}
%\floatstyle{ruled}
%\newfloat{alg}{ht}{loa}
%\floatname{alg}{Algorithm}

\usepackage{url}
\urldef{\mailsa}\path|firstname.lastname@epfl.ch|
\newcommand{\keywords}[1]{\par\addvspace\baselineskip
\noindent\keywordname\enspace\ignorespaces#1}

\newcommand{\cvc}{\textsc{cvc}{\small 4}\xspace}
\newcommand{\cvciii}{\textsc{cvc}{\small 3}\xspace}
\newcommand{\ziii}{\textsc{z}{\small 3}\xspace}
\newcommand{\teq}{\approx}
\newcommand{\sorts}{\mathsf{sort}}
\newcommand{\funcs}{\mathsf{func}}
\newcommand{\cc}[1]{#1^*}
\newcommand{\terms}{\mathbf{T}}
\newcommand{\M}{\mathcal{M}}
\def\AIF{\qtab\keyword{if}\ }
\def\THEN{\ \keyword{then}\ }
\def\AELSE{\untab\qtab\keyword{else}\ }
\def\FI{\untab}
\def\RETURN{\keyword{return}\ }
\def\ENDPROC{\untab}
%\newtheorem{remark}{Remark}

%\input{scalalistings}
%\input{smtlib2listings}
\usepackage{program}

\begin{document}

\title{Finite Traces of Recursive Functions in SMT}

\author {Andrew Reynolds\inst{1} \and Jasmin Christian Blanchette\inst{2,3} \and more?}
\authorrunning {A. Reynolds \and J. C. Blanchette}
\institute{
\'Ecole Polytechnique F\'ed\'erale de Lausanne (EPFL), Switzerland
\and
Inria Nancy \& LORIA, Villers-l\`es-Nancy, France
\and
Max-Planck-Institut f\"ur Informatik, Saarbr\"ucken, Germany
}

\maketitle

\begin{abstract}
To do.
\end{abstract}

\section{Introduction}
\label{sec:introduction}
To do.

{
\bibliographystyle{abbrv}
\bibliography{bib}
}

\end{document}

### 2015-04-28 tex/main.tex
Initial work on encoding/preliminaries.

@@
 
+\newcommand{\con}[1]{\mathsf{#1}}
+
 \newcommand{\cvc}{\textsc{cvc}{\small 4}\xspace}
@@
 \newcommand{\teq}{\approx}
-\newcommand{\sorts}{\mathsf{sort}}
-\newcommand{\funcs}{\mathsf{func}}
 \newcommand{\cc}[1]{#1^*}
 \newcommand{\terms}{\mathbf{T}}
+\newcommand{\functerms}{\mathbf{DFT}}
 \newcommand{\M}{\mathcal{M}}
@@
 \def\ENDPROC{\untab}
+\newcommand{\conv}{\mathcal{C}}
+
+\newcommand{\ssorts}[1]{#1^\mathrm{s}}
+\newcommand{\sfuns}[1]{#1^\mathrm{f}}
+\newcommand{\sfundefs}[1]{#1^\mathrm{df}}
+\newcommand{\sfunndefs}[1]{#1^\mathrm{nf}}
+
+\newcommand{\pnone}{\con{none}}
+\newcommand{\ptrue}{\con{true}}
+\newcommand{\pfalse}{\con{false}}
+\newcommand{\pol}{\con{pol}}
+
 %\newtheorem{remark}{Remark}
+
+\newcommand{\Bool}{\con{Bool}}
+\newcommand{\ltrue}{\top}
+\newcommand{\lfalse}{\bot}
+\newcommand{\lite}{\con{ite}}
+
+\newcommand{\boolop}{\oplus}
+\newcommand{\forallf}[1]{\forall_{#1}}
+\newcommand{\fnull}{\emptyset}
+\newcommand{\farg}[1]{a_{#1}}
+\newcommand{\vecfarg}[1]{\vec{a}_{#1}}
+\newcommand{\fargsort}[1]{\upsilon_{#1}}
 
@@
 
-\author {Andrew Reynolds\inst{1} \and Jasmin Christian Blanchette\inst{2,3} \and more?}
-\authorrunning {A. Reynolds \and J. C. Blanchette}
+\author {Andrew Reynolds\inst{1} \and Jasmin Christian Blanchette\inst{2,3} \and Cesare Tinelli \inst{4}}
+\authorrunning {A. Reynolds \and J. C. Blanchette \and C. Tinelli}
 \institute{
@@
 Max-Planck-Institut f\"ur Informatik, Saarbr\"ucken, Germany
+\and
+Department of Computer Science, The University of Iowa 
 }
@@
 
+\section{Related Work}
+
+Leon, others?
+
+\subsection{Model Finding for Quantified Formulas in SMT}
+
+z3, cvc4
+
+\section{Approach}
+
+\subsection{Preliminaries}
+
+A \emph{signature} $\Sigma$ consists of 
+a set $\ssorts{\Sigma}$ of sort symbols and
+a set $\sfuns{\Sigma}$ of function symbols.
+We assume that signatures always include a Boolean sort $\Bool$ and constants 
+$\ltrue$ and $\lfalse$ of that sort (respectively, for true and false), 
+an (infix) equality predicate $\teq$ of type $\sigma \times \sigma$ for each $\sigma \in \ssorts{\Sigma}$,
+standard boolean connectives ($\wedge$, $\vee$, $\Leftrightarrow$, etc.),
+and a predicate $\lite$ of type $\Bool \times \sigma \times \sigma$ for each $\sigma \in \ssorts{\Sigma}$.
+
+In this paper, we associate quantified formulas with the uninterpreted functions whose semantics they define.
+We will write $\forallf{f} \vec x. \varphi$ to denote a quantified formula that is defines the semantics of uninterpreted function $f$.
+For consistency, we consider only quantified formulas of this form, 
+and allow $f$ to be a distinguished constant $\fnull$, 
+denoting that the quantified formula does not define the semantics of an uninterpreted function.
+When $f$ is not $\fnull$, we will refer to $\forallf{f} \vec x. \varphi$ as a \emph{function definition} (for $f$).
+We will write $\forall \vec x. \varphi$ as shorthand for $\forallf{\fnull} \vec x. \varphi$,
+and $\exists \vec x. \varphi$ as shorthand for $\neg \forall \vec x. \neg \varphi$.
+
+\begin{definition}
+A formula $\varphi$ is well-defined with respect to functions $S$ if it is of the form:
+\begin{equation*}
+(\forallf{f_1} \vec x_1. \varphi_1) \wedge \ldots \wedge (\forallf{f_n} \vec x_n. \varphi_n) \wedge \psi
+\end{equation*}
+where $f_1, \ldots, f_n$ are distinct, 
+$S = \{ f_1, \ldots, f_n \}$,
+and $\psi$ contains no function definitions.
+\end{definition}
+
+Given a signature $\Sigma$, 
+we parition its function symbols $\sfuns{\Sigma}$ into $\sfundefs{\Sigma} \mathrel{\uplus} \sfunndefs{\Sigma}$,
+where $\sfundefs{\Sigma}$ is the set of \emph{defined} function symbols,
+and $\sfunndefs{\Sigma}$ are the others.
+In this paper, we will consider $\Sigma$-formulas that are well-defined for $\sfundefs{\Sigma}$.
+
+We write $\terms( t )$ to denote the set of subterms in $t$.
+We write $\functerms^\Sigma( t )$ to denote the set of $f$-applications in $t$, where $f \in \sfundefs{\Sigma}$.
+
+\subsection{Encoding Well-Defined Assumption for Recursive Functions}
+
+In the remainder of the section, we fix a signature $\Sigma$.
+
+The main idea behind this translation is to abstract the set of tuples that a recursive function $f$ is invoked on as a (fresh) uninterpreted sort $\fargsort{f}$.
+We extend our signature $\Sigma$ as follows.
+For each $f \in\sfundefs{\Sigma}$ of type $\tau_1 \times \ldots \tau_n \rightarrow \tau$, we associate:
+\begin{enumerate}
+\item[-] an uninterpreted sort $\fargsort{f}$, and
+\item[-] a vector of $n$ uninterpreted functions $\farg{f}^1$, $\ldots$, $\farg{f}^n$ of type $(\fargsort{f} \rightarrow \tau_1)$, $\ldots$, $(\fargsort{f} \rightarrow \tau_n)$.
+\end{enumerate}
+
+The role of the uninterpreted functions $\farg{f}^1$, $\ldots$, $\farg{f}^n$ will be discussed more in the following.
+
+\begin{figure}[t]
+\begin{enumerate}
+%\begin{framed}
+\item[\ ] 
+$\conv( t^\tau, p )$ : 
+ \begin{itemize}
+  \item[] $\mathsf{if}$ $t \equiv \forallf{f} \vec x. t_1$
+    \begin{itemize}
+      \item[] $\mathsf{let} ( r_1, D_1 ) = \conv( t_1, p )$
+      \item[] $\mathsf{if}$ $f \equiv \fnull$
+      \begin{itemize}
+        \item[] $\mathsf{return}$ $( \forall \vec x. r_1, \forall \vec x. D_1 )$
+      \end{itemize}
+      \item[] else
+      \begin{itemize}
+        \item[] $\mathsf{assert}( D_1 = \emptyset )$;
+        \item[] $\mathsf{return}$ $( \forall y^{\fargsort{f}}. ( r_1 [ \vecfarg{f}( y ) / \vec x ] ), \emptyset )$
+      \end{itemize}
+    \end{itemize}
+  \item[] $\mathsf{else}$ $\mathsf{if}$ $\tau \equiv \Bool$ and $t \equiv t_1 \boolop \ldots \boolop t_n$ :
+    \begin{itemize}
+      \item[] $\mathsf{let}$ $( r_i, D_i ) = \conv( t_i, \pol( \boolop, i, p ) )$ for $i = 1, \ldots, n$
+      \item[] $\mathsf{if}$ $p = \ptrue$
+      \begin{itemize}
+        \item[] $\mathsf{return}$ $( ( r_1 \boolop \ldots \boolop r_n ) \wedge D_1 \wedge \ldots \wedge D_n, \emptyset )$
+      \end{itemize}
+      \item[] $\mathsf{else}$ $\mathsf{if}$ $p = \pfalse$
+      \begin{itemize}
+        \item[] $\mathsf{return}$ $( ( r_1 \boolop \ldots \boolop r_n ) \vee \neg D_1 \vee \ldots \vee \neg D_n, \emptyset )$
+      \end{itemize}
+      \item[] $\mathsf{else}$
+      \begin{itemize}
+        \item[] $\mathsf{return}$ $( r_1 \boolop \ldots \boolop r_n, D_1 \cup \ldots \cup D_n )$
+      \end{itemize}
+    \end{itemize}
+   \item[] $\mathsf{else}$
+   \begin{itemize}
+     \item[] $\mathsf{return}$ $( t, \{ \exists z^{\fargsort{f}}. ( \vecfarg{f}( z ) \teq \vec t ) \mid f( \vec t ) \in \functerms^\Sigma( t ) \} )$
+   \end{itemize}
+ \end{itemize}
+\end{enumerate}
+\vspace{-2ex}
+\caption{A translation procedure $\conv$ for a $\Sigma$-formula that is well-defined with respect to function definitions.
+In the procedure, $\pol( \boolop, i, p )$ returns the polarity of the $i^{th}$ child of a $\boolop$-application having polarity $p$.
+}
+\label{fig:encoding}
+\end{figure}
+
+Figure~\ref{fig:encoding} gives a procedure $\conv$ for translating a formula $t$ containing function definitions.
+The procedure takes two arguments: the term $t^\tau$ to translate, and a \emph{polarity} p (either $\ptrue$, $\pfalse$, or $\pnone$).
+If function definition quantifiers $\forallf{f}$ in $t$ are such that $f$ is well-defined and $\conv( t, \ptrue )$ returns $( r, \emptyset )$,
+then $r$ is equivalent to $t$.
+
+\begin{example}
+Consider the formula
+\end{example}
+
+\section{Experiments}
+
+\section{Conclusion}
+\label{sec:conclusion}
+
 {

### 2015-04-28 tex/main.tex
More work, add example.

@@
 \newcommand{\sfunndefs}[1]{#1^\mathrm{nf}}
+\newcommand{\sortint}{\mathrm{Int}}
 
@@
 
+\newcommand{\Sigmalia}{\Sigma_{LIA}}
+\newcommand{\extendsig}[1]{\mathcal{E}( #1 )}
+
+\newcommand{\rem}[1]{\textcolor{red}{[#1]}}
+\newcommand{\ajr}[1]{\rem{#1 --ajr}}
+
+\newcommand{\vthinspace}{\kern+0.083333em}
+\newcommand{\typ}[1]{^{\vthinspace #1}}
+
 %\input{scalalistings}
@@
 
-\title{Finite Traces of Recursive Functions in SMT}
+\title{Model Finding for Finite Function Traces in SMT}
 
@@
 
-Leon, others?
+Leon, Nitpick, others?
 
@@
 an (infix) equality predicate $\teq$ of type $\sigma \times \sigma$ for each $\sigma \in \ssorts{\Sigma}$,
-standard boolean connectives ($\wedge$, $\vee$, $\Leftrightarrow$, etc.),
-and a predicate $\lite$ of type $\Bool \times \sigma \times \sigma$ for each $\sigma \in \ssorts{\Sigma}$.
+and standard boolean connectives ($\neg$, $\wedge$, $\vee$, $\lite$, etc.).
+%and a function $\lite$ of type $\Bool \times \sigma \times \sigma$ for each $\sigma \in \ssorts{\Sigma}$.
+We will write $t\typ{\tau}$ to denote a term $t$ of type~$\tau$.
 
@@
 
+\ajr{theories, models, etc.}
+
 \begin{definition}
-A formula $\varphi$ is well-defined with respect to functions $S$ if it is of the form:
+A formula $\varphi$ \emph{well-defined with respect to functions} $S$ if it is of the form:
 \begin{equation*}
@@
 Given a signature $\Sigma$, 
-we parition its function symbols $\sfuns{\Sigma}$ into $\sfundefs{\Sigma} \mathrel{\uplus} \sfunndefs{\Sigma}$,
+we will commonly parition its function symbols $\sfuns{\Sigma}$ into $\sfundefs{\Sigma} \mathrel{\uplus} \sfunndefs{\Sigma}$,
 where $\sfundefs{\Sigma}$ is the set of \emph{defined} function symbols,
-and $\sfunndefs{\Sigma}$ are the others.
-In this paper, we will consider $\Sigma$-formulas that are well-defined for $\sfundefs{\Sigma}$.
+and $\sfunndefs{\Sigma}$ are the others,
+and consider $\Sigma$-formulas that are well-defined with respect to functions $\sfundefs{\Sigma}$.
 
 We write $\terms( t )$ to denote the set of subterms in $t$.
-We write $\functerms^\Sigma( t )$ to denote the set of $f$-applications in $t$, where $f \in \sfundefs{\Sigma}$.
+We write $\functerms^\Sigma( t )$ to denote the set of $f$-applications in $t$ such that $f \in \sfundefs{\Sigma}$.
 
 \subsection{Encoding Well-Defined Assumption for Recursive Functions}
-
+$\Leftrightarrow$, 
 In the remainder of the section, we fix a signature $\Sigma$.
 
-The main idea behind this translation is to abstract the set of tuples that a recursive function $f$ is invoked on as a (fresh) uninterpreted sort $\fargsort{f}$.
-We extend our signature $\Sigma$ as follows.
-For each $f \in\sfundefs{\Sigma}$ of type $\tau_1 \times \ldots \tau_n \rightarrow \tau$, we associate:
-\begin{enumerate}
-\item[-] an uninterpreted sort $\fargsort{f}$, and
-\item[-] a vector of $n$ uninterpreted functions $\farg{f}^1$, $\ldots$, $\farg{f}^n$ of type $(\fargsort{f} \rightarrow \tau_1)$, $\ldots$, $(\fargsort{f} \rightarrow \tau_n)$.
-\end{enumerate}
-
-The role of the uninterpreted functions $\farg{f}^1$, $\ldots$, $\farg{f}^n$ will be discussed more in the following.
+Say that $\Sigma$-formula $\varphi$ is well-defined with respect to functions $\sfundefs{\Sigma}$,
+and the definitions in $\sfundefs{\Sigma}$ are logically consistent \ajr{formalize}.
+In this section, we present a method which constructs an extended signature $\Sigma'$,
+and translates $\varphi$ to equisatisfiable $\Sigma'$-formula $\psi$ on which known model-finding procedures~\cite{GeDeM-CAV-09, ReyEtAl-1-RR-13} are applicable.
 
@@
 \item[\ ] 
-$\conv( t^\tau, p )$ : 
+$\conv( t\typ{\tau}, p )$ : 
  \begin{itemize}
-  \item[] $\mathsf{if}$ $t \equiv \forallf{f} \vec x. t_1$
-    \begin{itemize}
-      \item[] $\mathsf{let} ( r_1, D_1 ) = \conv( t_1, p )$
-      \item[] $\mathsf{if}$ $f \equiv \fnull$
-      \begin{itemize}
-        \item[] $\mathsf{return}$ $( \forall \vec x. r_1, \forall \vec x. D_1 )$
-      \end{itemize}
-      \item[] else
-      \begin{itemize}
-        \item[] $\mathsf{assert}( D_1 = \emptyset )$;
-        \item[] $\mathsf{return}$ $( \forall y^{\fargsort{f}}. ( r_1 [ \vecfarg{f}( y ) / \vec x ] ), \emptyset )$
-      \end{itemize}
-    \end{itemize}
-  \item[] $\mathsf{else}$ $\mathsf{if}$ $\tau \equiv \Bool$ and $t \equiv t_1 \boolop \ldots \boolop t_n$ :
+   \item[] $\mathsf{if}$ $\tau \equiv \Bool$ and $t \equiv t_1 \boolop \ldots \boolop t_n$ :
     \begin{itemize}
@@
     \end{itemize}
+  \item[] $\mathsf{else}$ $\mathsf{if}$ $t \equiv \forallf{f} \vec x. t_1$
+    \begin{itemize}
+      \item[] $\mathsf{let} ( r_1, D_1 ) = \conv( t_1, p )$
+      \item[] $\mathsf{if}$ $f \equiv \fnull$
+      \begin{itemize}
+        \item[] $\mathsf{return}$ $( \forall \vec x. r_1, \forall \vec x. D_1 )$
+      \end{itemize}
+      \item[] else
+      \begin{itemize}
+        \item[] $\mathsf{assert}( D_1 = \emptyset )$;
+        \item[] $\mathsf{return}$ $( \forall y\typ{\fargsort{f}}. ( r_1 [ \vecfarg{f}( y ) / \vec x ] ), \emptyset )$
+      \end{itemize}
+    \end{itemize}
    \item[] $\mathsf{else}$
    \begin{itemize}
-     \item[] $\mathsf{return}$ $( t, \{ \exists z^{\fargsort{f}}. ( \vecfarg{f}( z ) \teq \vec t ) \mid f( \vec t ) \in \functerms^\Sigma( t ) \} )$
+     \item[] $\mathsf{return}$ $( t, \{ \exists z\typ{\fargsort{f}}. ( \vecfarg{f}( z ) \teq \vec t ) \mid f( \vec t ) \in \functerms^\Sigma( t ) \} )$
    \end{itemize}
@@
 \caption{A translation procedure $\conv$ for a $\Sigma$-formula that is well-defined with respect to function definitions.
-In the procedure, $\pol( \boolop, i, p )$ returns the polarity of the $i^{th}$ child of a $\boolop$-application having polarity $p$.
+In the procedure, $\pol( \boolop, i, p )$ returns the polarity of the $i^{th}$ child of a $\boolop$-application having polarity $p$,
+where $\boolop$ is an (interpreted) predicate.
 }
@@
 
-Figure~\ref{fig:encoding} gives a procedure $\conv$ for translating a formula $t$ containing function definitions.
-The procedure takes two arguments: the term $t^\tau$ to translate, and a \emph{polarity} p (either $\ptrue$, $\pfalse$, or $\pnone$).
-If function definition quantifiers $\forallf{f}$ in $t$ are such that $f$ is well-defined and $\conv( t, \ptrue )$ returns $( r, \emptyset )$,
-then $r$ is equivalent to $t$.
+The intuition behind this translation is use an uninterpreted sort $\fargsort{f}$ to abstract the set of tuples that each recursive function $f$ is invoked on,
+and restrict the quantification of the function definition for $f$ to a variable of sort $\fargsort{f}$.
+In more detail, we construct an extended signature of $\Sigma$, which we will write as $\extendsig{\Sigma}$, as follows.
+For each $f \in \sfundefs{\Sigma}$ of type $\tau_1 \times \ldots \tau_n \rightarrow \tau$, 
+the signature $\extendsig{\Sigma}$ includes:
+\begin{enumerate}
+\item[-] an uninterpreted sort $\fargsort{f}$, and
+\item[-] a vector of $n$ uninterpreted functions $\farg{f}^1$, $\ldots$, $\farg{f}^n$ of type $(\fargsort{f} \rightarrow \tau_1)$, $\ldots$, $(\fargsort{f} \rightarrow \tau_n)$.
+\end{enumerate}
+%The interpretation of uninterpreted sort $\fargsort{f}$ will denote the elements (tuples) on which the function $f$ is applied.
+%The role of the uninterpreted functions $\farg{f}^1$, $\ldots$, $\farg{f}^n$ will be discussed more in the following.
+We then run the procedure $\conv$ in Figure~\ref{fig:encoding} on $\varphi$.
+The procedure $\conv$ takes two arguments: the term $t$ to translate, and a \emph{polarity} $p$ (either $\ptrue$, $\pfalse$, or $\pnone$).
+It returns a pair of the form $( r, D )$, where $r$ is a term of type $\tau$, and $D$ is a set of formulas, which we will call its \emph{domain constaints}.
+The role of $D$ is to ensure that the (restricted) function definition for $f$ includes certain tuples in its domain, which we explain more in the following.
+
+In the case that $t$ is an application of (interpreted) predicate $\boolop$ (e.g. a boolean connective, or equality),
+we recursively call $\conv$ on its subchildren $t_i$ and polarity $\pol( \boolop, i, p )$, where $\pol$ is defined as\footnote{The negation $\neg p$ of polarity $p$ is 
+$\pfalse$ if $p$ is $\ptrue$,
+$\ptrue$ if $p$ is $\pfalse$, and
+$\pnone$ if $p$ is $\pnone$.}:
+\begin{equation*}
+\pol( \boolop, i, p ) = \begin{cases}
+                         \pnone & \boolop \equiv \teq \text{ or } \boolop \equiv \lite \wedge i=0 \\
+                         \neg p & \boolop \equiv \neg \\
+                         p & \text{otherwise}
+                         \end{cases}
+\end{equation*}
+If $t$ has positive polarity,
+we construct a conjunction of the result of the recursive calls with domain constraints $D_1, \ldots, D_n$.
+Similarly, if $t$ has negative polarity,
+we construct a disjunction of the result of the recursive calls with the negation of domain constraints $D_1, \ldots, D_n$.
+It $t$ has no polarity,
+then we return the union of the domain constraints $D_1 \cup \ldots \cup D_n$.
+
+In the case that $t$ is a quantified formula $\forallf{f} \vec x. t_1$, it recursively calls $\conv$ on its body with the same polarity.
+In the case that $t$ is not a function definition, 
+the resulting quantified formula is reconstructed with the term $r_1$ returned by the recursive call,
+and a quantified prefix is appended to the domain constraints $D_1$ returned by the recursive call.
+In the case that $t$ is a function definition,
+we instead construct a quantified formula over a single variable $y$ of type $\fargsort{f}$,
+and replace all occurrences of $\vec x$ in $r_1$ with $\vecfarg{f}( y )$.
+In this case, since function definitions are a top-level conjunct, by case analysis on the return values of $\conv$, we know that $D_1$ is empty.
+
+We demonstrate this translation with a simple example.
 
 \begin{example}
-Consider the formula
+Consider the (combined) signature $\Sigma = \Sigmalia \cup \Sigma_u$,
+where $\ssorts{\Sigma_u} = \{ \sortint \}$, and 
+$\sfuns{\Sigma_u} = \{ s\typ{ \sortint \rightarrow \sortint }, c\typ{ \sortint } \}$,
+and the $\Sigma$-formula $\varphi$:
+\begin{equation*}
+\forall_{s} x\typ{\sortint}. ite( x \leq 0, s(x) \teq 0, s( x ) \teq x + s( x-1 ) ) \wedge s( c ) > 100
+\end{equation*}
+Here, $s$ is a function that returns the sum of positive numbers between $0$ and its argument $x$.
+The formula $\varphi$ is well-defined with respect to functions $\sfundefs{\Sigma_u}$,
+and states that the sum of all positive numbers between $0$ and $c$ is greater than $100$.
+The signature $\extendsig{\Sigma}$ includes the sort $\fargsort{s}$,
+and the uninterpreted function $\farg{s}\typ{ \fargsort{s} \rightarrow \sortint }$.
+The result of $\conv( \varphi, \ptrue )$ is the pair $( \psi, \emptyset )$, where after simplification, $\psi$ is the $\extendsig{\Sigma}$-formula:
+\begin{equation*}
+\begin{split}
+\forall y\typ{\fargsort{s}}. ite( & \farg{s}( y ) \leq 0,  \\
+ & s(\farg{s}( y )) \teq 0, \\
+ & s(\farg{s}( y )) \teq \farg{s}( y ) + s( \farg{s}( y )-1 ) ) \\
+ & \wedge \exists z\typ{\fargsort{s}}. \farg{s}( z ) \teq \farg{s}( y )-1 )
+\end{split}  
+\wedge ( s( a ) > 100 \wedge \exists z\typ{\fargsort{s}}. \farg{s}( z ) \teq a )
+\end{equation*}
+Note that the universal quantification in this example (the left disjunct) is $\ldots$
+$\square$
 \end{example}
+
+We claim that if $\varphi$ is well-defined with respect to functions $\sfundefs{\Sigma}$,
+function definitions in $\sfundefs{\Sigma}$ are logically consistent \ajr{formalize},
+and $\conv( \varphi, \ptrue )$ returns $( \psi, \emptyset )$,
+then $\psi$ is equisatisfiable to $\varphi$.
+
+\begin{lemma}
+Let $\varphi$ be a $\Sigma$-formula that is well-defined with respect to functions $\sfundefs{\Sigma}$,
+and let $( \psi, D )$ be the result of $\conv( \varphi, \ptrue )$.
+Then,
+(i) $D = \emptyset$, and
+(ii) for each model of $\varphi$, $\ldots$. \ajr{todo}
+\end{lemma}
 

### 2015-04-28 tex/main.tex
Minor fix
Work on results section, minor changes.

@@
 
-\title{Model Finding for Finite Function Traces in SMT}
+\title{Model Finding for Recursive Functions in SMT}
 
@@
 \subsection{Preliminaries}
+\label{sec:prelim}
 
 A \emph{signature} $\Sigma$ consists of 
-a set $\ssorts{\Sigma}$ of sort symbols and
+a set $\ssorts{\Sigma}$ of type symbols and
 a set $\sfuns{\Sigma}$ of function symbols.
-We assume that signatures always include a Boolean sort $\Bool$ and constants 
-$\ltrue$ and $\lfalse$ of that sort (respectively, for true and false), 
+We assume that signatures always include a Boolean type $\Bool$ and constants 
+$\ltrue$ and $\lfalse$ of that type (respectively, for true and false), 
 an (infix) equality predicate $\teq$ of type $\sigma \times \sigma$ for each $\sigma \in \ssorts{\Sigma}$,
@@
 
-\ajr{theories, models, etc.}
+\ajr{theories, models, equisatisfiable, etc.}
 
@@
 A formula $\varphi$ \emph{well-defined with respect to functions} $S$ if it is of the form:
-\begin{equation*}
+\begin{eqnarray} \label{eq:wdf}
 (\forallf{f_1} \vec x_1. \varphi_1) \wedge \ldots \wedge (\forallf{f_n} \vec x_n. \varphi_n) \wedge \psi
-\end{equation*}
+\end{eqnarray}
 where $f_1, \ldots, f_n$ are distinct, 
@@
 \subsection{Encoding Well-Defined Assumption for Recursive Functions}
-$\Leftrightarrow$, 
+\label{sec:encoding}
+
 In the remainder of the section, we fix a signature $\Sigma$.
@@
 \item[\ ] 
-$\conv( t\typ{\tau}, p )$ : 
+$\conv( \varphi )$ : 
+ \begin{itemize}
+   \item[] $\mathsf{let}$ $( \psi, D ) = \conv( \varphi, \ptrue )$;
+   %\item[] $\mathsf{assert}( D = \emptyset )$;
+   \item[] $\mathsf{return}$ $\psi$
+  \end{itemize}
+\end{enumerate}
+\begin{enumerate}
+%\begin{framed}
+\item[\ ] 
+$\conv_0( t\typ{\tau}, p )$ : 
  \begin{itemize}
@@
     \begin{itemize}
-      \item[] $\mathsf{let}$ $( r_i, D_i ) = \conv( t_i, \pol( \boolop, i, p ) )$ for $i = 1, \ldots, n$
+      \item[] $\mathsf{let}$ $( r_i, D_i ) = \conv_0( t_i, \pol( \boolop, i, p ) )$ for $i = 1, \ldots, n$;
       \item[] $\mathsf{if}$ $p = \ptrue$
@@
     \begin{itemize}
-      \item[] $\mathsf{let} ( r_1, D_1 ) = \conv( t_1, p )$
+      \item[] $\mathsf{let} ( r_1, D_1 ) = \conv_0( t_1, p )$;
       \item[] $\mathsf{if}$ $f \equiv \fnull$
@@
       \begin{itemize}
-        \item[] $\mathsf{assert}( D_1 = \emptyset )$;
+        %\item[] $\mathsf{assert}( D_1 = \emptyset )$;
         \item[] $\mathsf{return}$ $( \forall y\typ{\fargsort{f}}. ( r_1 [ \vecfarg{f}( y ) / \vec x ] ), \emptyset )$
@@
 \caption{A translation procedure $\conv$ for a $\Sigma$-formula that is well-defined with respect to function definitions.
-In the procedure, $\pol( \boolop, i, p )$ returns the polarity of the $i^{th}$ child of a $\boolop$-application having polarity $p$,
+In the subprocedure $\conv_0$, the procedure $\pol( \boolop, i, p )$ returns the polarity of the $i^{th}$ child of a $\boolop$-application having polarity $p$,
 where $\boolop$ is an (interpreted) predicate.
@@
 
-The intuition behind this translation is use an uninterpreted sort $\fargsort{f}$ to abstract the set of tuples that each recursive function $f$ is invoked on,
-and restrict the quantification of the function definition for $f$ to a variable of sort $\fargsort{f}$.
+The intuition behind this translation is use an uninterpreted type $\fargsort{f}$ to abstract the set of tuples that each recursive function $f$ is invoked on,
+and restrict the quantification of the function definition for $f$ to a variable of type $\fargsort{f}$.
 In more detail, we construct an extended signature of $\Sigma$, which we will write as $\extendsig{\Sigma}$, as follows.
@@
 \begin{enumerate}
-\item[-] an uninterpreted sort $\fargsort{f}$, and
+\item[-] an uninterpreted type $\fargsort{f}$, and
 \item[-] a vector of $n$ uninterpreted functions $\farg{f}^1$, $\ldots$, $\farg{f}^n$ of type $(\fargsort{f} \rightarrow \tau_1)$, $\ldots$, $(\fargsort{f} \rightarrow \tau_n)$.
@@
 We then run the procedure $\conv$ in Figure~\ref{fig:encoding} on $\varphi$.
-The procedure $\conv$ takes two arguments: the term $t$ to translate, and a \emph{polarity} $p$ (either $\ptrue$, $\pfalse$, or $\pnone$).
+This calls the subprocedure $\conv_0$ on $\varphi$ and the \emph{polarity} $\ptrue$.
+The procedure $\conv_0$ takes two arguments: the term $t$ to translate, and a polarity $p$ (either $\ptrue$, $\pfalse$, or $\pnone$).
 It returns a pair of the form $( r, D )$, where $r$ is a term of type $\tau$, and $D$ is a set of formulas, which we will call its \emph{domain constaints}.
@@
 \pol( \boolop, i, p ) = \begin{cases}
-                         \pnone & \boolop \equiv \teq \text{ or } \boolop \equiv \lite \wedge i=0 \\
+                         \pnone & \boolop \equiv \teq \text{ or } ( \boolop \equiv \lite \text{ and } i=0 ) \\
                          \neg p & \boolop \equiv \neg \\
@@
 Consider the (combined) signature $\Sigma = \Sigmalia \cup \Sigma_u$,
-where $\ssorts{\Sigma_u} = \{ \sortint \}$, and 
-$\sfuns{\Sigma_u} = \{ s\typ{ \sortint \rightarrow \sortint }, c\typ{ \sortint } \}$,
-and the $\Sigma$-formula $\varphi$:
+where $\ssorts{\Sigma_u} = \{ \sortint \}$ and 
+$\sfuns{\Sigma_u} = \{ s\typ{ \sortint \rightarrow \sortint }, c\typ{ \sortint } \}$.
+Consider the $\Sigma$-formula $\varphi$:
 \begin{equation*}
@@
 The signature $\extendsig{\Sigma}$ includes the sort $\fargsort{s}$,
-and the uninterpreted function $\farg{s}\typ{ \fargsort{s} \rightarrow \sortint }$.
+and the uninterpreted function $\farg{s}$ of sort $\fargsort{s} \rightarrow \sortint$.
 The result of $\conv( \varphi, \ptrue )$ is the pair $( \psi, \emptyset )$, where after simplification, $\psi$ is the $\extendsig{\Sigma}$-formula:
@@
 \end{split}  
-\wedge ( s( a ) > 100 \wedge \exists z\typ{\fargsort{s}}. \farg{s}( z ) \teq a )
+\wedge ( s( c ) > 100 \wedge \exists z\typ{\fargsort{s}}. \farg{s}( z ) \teq c )
 \end{equation*}
-Note that the universal quantification in this example (the left disjunct) is $\ldots$
+The universal quantification in the resulting formula (the left disjunct) is over an uninterpreted sort $\fargsort{s}$, 
+thus enabling finite model finding techniques~\cite{ReyEtAl-1-RR-13} to find a model for $\psi$ if a finite interpretation exists for $\fargsort{s}$.
+Moreover, all occurrences of $y$ in this formula are beneath applications of the uninterpreted function $\farg{s}$,
+implying that the formula is in the essentially uninterpreted fragment of~\cite{GeDeM-CAV-09}, for which there exists a complete instantiation procedure.
 $\square$
@@
 
-We claim that if $\varphi$ is well-defined with respect to functions $\sfundefs{\Sigma}$,
+It remains to show that our translation produces an equisatisfiable formulae.
+Namely,
+we claim that if $\varphi$ is well-defined with respect to functions $\sfundefs{\Sigma}$,
 function definitions in $\sfundefs{\Sigma}$ are logically consistent \ajr{formalize},
-and $\conv( \varphi, \ptrue )$ returns $( \psi, \emptyset )$,
-then $\psi$ is equisatisfiable to $\varphi$.
+then $\conv(\varphi)$ is equisatisfiable to $\varphi$.
 
 \begin{lemma}
-Let $\varphi$ be a $\Sigma$-formula that is well-defined with respect to functions $\sfundefs{\Sigma}$,
-and let $( \psi, D )$ be the result of $\conv( \varphi, \ptrue )$.
-Then,
-(i) $D = \emptyset$, and
-(ii) for each model of $\varphi$, $\ldots$. \ajr{todo}
+Let $\varphi$ be a $\Sigma$-formula that is well-defined with respect to functions $\sfundefs{\Sigma}$.
+\ajr{todo}
 \end{lemma}
 
-\section{Experiments}
+\section{Evaluation}
+
+In this section, we evaluate both 
+the overall impact of the encoding introduced in the previous section, and
+the performance of individual SMT techniques for benchmarks in the encoding.
+
+For benchmarks, we gathered a set of 245 benchmarks from two sources, which we will refer to as {\bf isa} and {\bf leon}.
+The first ({\bf isa}) consists of the 79 benchmarks from the Isaplanner benchmark suite~\cite{DBLP:conf/itp/JohanssonDB10} that did not contain higher-order functions.
+The benchmarks have been used recently as challenge problems for a variety of inductive theorem provers.
+These benchmarks heavily involve recursive functions, and are limited to the combined theory of uninterpreted functions and inductive datatypes.
+The second ({\bf leon}) consists of 166 benchmarks from the Leon repository~\footnote{Available at \url{https://github.com/epfl-lara/leon}.},
+which were constructed from verification conditions from simple Scala programs.
+These benchmarks also heavily involve recursive functions, 
+and cover a wide variety of theories, including bitvectors, arrays, linear and non-linear arithmetic, and inductive datatypes.
+Each of the 245 benchmarks are of the form described in Formula (~\ref{eq:wdf}) from Section~\ref{sec:prelim},
+namely they each are a conjoined list of function definitions followed by a (negated) conjecture.
+A majority of these benchmarks are unsatisfiable, although a handful of the benchmarks in the {\bf leon} set are satisfiable.
+
+For each of these 245 benchmarks, we considered (up to) three randomly selected mutated forms of its conjecture $\psi$.
+In particular, we considered unique conjectures that are obtained as a result of swapping a subterm of $\psi$ at one position
+with another of the same type at another position.
+In total, we considered 213 mutated forms of conjectures from {\bf isa}, and 427 mutated forms of conjectures from {\bf leon}.
+We will call these sets {\bf isa+m} and {\bf leon+m} respectively.
+
+In total, our benchmark set consisted of 885 benchmarks (640 mutants and 245 originals).
+We considered these 885 benchmarks both before and after the translation method $\conv$ as described in Section~\ref{sec:encoding}.
+We will denote the latter encoding using the suffix {\bf w}, for instance {\bf isa+w} or {\bf isa+mw}.
+
+For solvers, we considered the SMT solver \ziii~\cite{de-moura-bjoerner-2008}, 
+which runs both heuristic methods for quantifier instantiation,
+as well as methods for finding models for quantified formulas~\cite{GeDeM-CAV-09}.
+We also considered three configurations of \cvc~\cite{barrett-et-al-2011} which we will refer to as {\bf cvc4}, {\bf cvc4+f} and {\bf cvc4+i}.
+The default configuration {\bf cvc4} runs heuristic quantifier instantiation, but does not include techniques for finite model finding.
+The configuration {\bf cvc4+f} runs a finite model finding procedure~\ref{ReyEtAl-1-RR-13}.
+The last configuration {\bf cvc4+i} incorporates techniques for automating inductive reasoning in SMT~\ref{reynolds-kuncak-2015}.
+
+The results are shown in Figure~\ref{fig:results}.
 
@@
 \bibliographystyle{abbrv}
-\bibliography{bib}
+\bibliography{main}
 }

### 2015-04-28 tex/main.tex
Minor fixes

@@
 The default configuration {\bf cvc4} runs heuristic quantifier instantiation, but does not include techniques for finite model finding.
-The configuration {\bf cvc4+f} runs a finite model finding procedure~\ref{ReyEtAl-1-RR-13}.
-The last configuration {\bf cvc4+i} incorporates techniques for automating inductive reasoning in SMT~\ref{reynolds-kuncak-2015}.
+The configuration {\bf cvc4+f} runs a finite model finding procedure~\cite{ReyEtAl-1-RR-13}.
+The last configuration {\bf cvc4+i} incorporates techniques for automating inductive reasoning in SMT~\cite{reynolds-kuncak-2015}.
 

### 2015-04-29 tex/main.tex
uniformized affiliation
more

@@
 \and
-Department of Computer Science, The University of Iowa 
+Department of Computer Science, University of Iowa, USA
 }
@@
 \label{sec:introduction}
-To do.
+
+  * why model finding?
+  * but SMT solvers better at proving
+    * a single universal quantifier in an axiom is enough to put it off track
+      (``maybe unsat'')
+
+  * quantifiers in recursive fun defs
+
+  * intuitively, finite quantification suffices:
+    * if we make sure all needed values are there
+    * if well-founded (assumed!)
+
+  * nasty example (not well-founded)
+    * f x = f x + 1
+
+  * how we do it
+
+  * show that it works also for example with weird recursion schemes
 
@@
 
-Leon, Nitpick, others?
+Alloy (and Alloy idiom by Kuncak \& Jackson)
+The model finder for KIV based on Alloy
+Nitpick
+Leon
 

### 2015-04-29 tex/main.tex
more

@@
 
-Alloy (and Alloy idiom by Kuncak \& Jackson)
+first-order model finders: SEM, MACE/Mace, Paradox, iProver
+  either direct search or reduction to SAT
+
+Alloy (and Alloy idiom by Kuncak \& Jackson), Kodkod
 The model finder for KIV based on Alloy

### 2015-04-29 tex/main.tex
more

@@
 
+  * corecursive functions???
+    * productive
+
 \section{Related Work}
@@
 Alloy (and Alloy idiom by Kuncak \& Jackson), Kodkod
+
 The model finder for KIV based on Alloy
+
+Refute
+  * unsound handling of datatypes
+
 Nitpick
+  * inductive / coinductive predicates
+  * quotient types
+
 Leon
@@
 
+  * future work
+    * (co)inductive predicates
+
+{%\footnotesize
+\def\ackname{Acknowledgment}
+\paragraph{%\footnotesize
+\ackname.}
+
+This research is partially supported by the Inria technological development
+action \relax{Contrexemples Utilisables par Isabelle et Coq} (CUIC).
+
+}
+
 {

### 2015-04-29 tex/main.tex
more
Most of the evaluation section.

@@
 \def\ENDPROC{\untab}
-\newcommand{\conv}{\mathcal{C}}
+\newcommand{\conv}{\mathcal{W}}
 
@@
 
+\ajr{Demand for recursive functions in SMT (new draft includes define-funs-rec).  Usually not interested in showing logical consistency, there exist other tools that check e.g. termination. 
+Thus, want SMT solver to check satisfiability the assumption that recursive functions are consistent (where this needs to be defined precisely).
+}
+
+\paragraph{Contributions}
+
+\ajr{(1) Encoding, (2) Evaluation, (3) Front-end support for define-funs-rec in \cvc}
+
 \section{Related Work}
@@
 and translates $\varphi$ to equisatisfiable $\Sigma'$-formula $\psi$ on which known model-finding procedures~\cite{GeDeM-CAV-09, ReyEtAl-1-RR-13} are applicable.
+The intuition behind this translation is use an uninterpreted type to abstract the set of tuples that each recursive function $f$ is invoked on,
+and restrict the quantification of the function definition for $f$ to a variable of this type.
 
@@
 
-The intuition behind this translation is use an uninterpreted type $\fargsort{f}$ to abstract the set of tuples that each recursive function $f$ is invoked on,
-and restrict the quantification of the function definition for $f$ to a variable of type $\fargsort{f}$.
-In more detail, we construct an extended signature of $\Sigma$, which we will write as $\extendsig{\Sigma}$, as follows.
+In more detail, 
+say our signature $\Sigma$ contains a set of defined function symbols $\sfundefs{\Sigma} \subseteq \sfuns{\Sigma}$.
+We extend the signature of $\Sigma$ to a signature $\extendsig{\Sigma}$ containing the following additional definitions.
 For each $f \in \sfundefs{\Sigma}$ of type $\tau_1 \times \ldots \tau_n \rightarrow \tau$, 
@@
 %The role of the uninterpreted functions $\farg{f}^1$, $\ldots$, $\farg{f}^n$ will be discussed more in the following.
-We then run the procedure $\conv$ in Figure~\ref{fig:encoding} on $\varphi$.
-This calls the subprocedure $\conv_0$ on $\varphi$ and the \emph{polarity} $\ptrue$.
-The procedure $\conv_0$ takes two arguments: the term $t$ to translate, and a polarity $p$ (either $\ptrue$, $\pfalse$, or $\pnone$).
-It returns a pair of the form $( r, D )$, where $r$ is a term of type $\tau$, and $D$ is a set of formulas, which we will call its \emph{domain constaints}.
-The role of $D$ is to ensure that the (restricted) function definition for $f$ includes certain tuples in its domain, which we explain more in the following.
+
+Given the extended signature, we run the procedure $\conv$ in Figure~\ref{fig:encoding} on $\varphi$.
+This calls the subprocedure $\conv_0$, which takes two arguments: the term $t$ to translate, and a \emph{polarity} $p$ (either $\ptrue$, $\pfalse$, or $\pnone$).
+It returns a pair of the form $( r, D )$, where $r$ is a term of type $\tau$, and $D$ is a set of formulas.
+%The role of $D$ is to ensure that the (restricted) function definition for $f$ includes certain tuples in its domain, which we explain more in the following.
+At a high level, the translation modifies $\varphi$ in two ways.
+First, it restricts the quantification on function definitions for $f$ to the corresponding uninterpreted sort $\fargsort{f}$.
+Second, it augments $\varphi$ with additional constraints of the form $\exists z\typ{\fargsort{f}}. ( \vecfarg{f}( z ) \teq \vec t )$
+which ensure that this restriction for the defintion $f$ covers all relevant tuples of terms,
+namely the ones for which an $f$-application exists and is relevant to the satisfiability of the current branch of $\varphi$.
 
@@
 If $t$ has positive polarity,
-we construct a conjunction of the result of the recursive calls with domain constraints $D_1, \ldots, D_n$.
+we construct a conjunction of the result of the recursive calls with formulas $D_1, \ldots, D_n$.
 Similarly, if $t$ has negative polarity,
-we construct a disjunction of the result of the recursive calls with the negation of domain constraints $D_1, \ldots, D_n$.
+we construct a disjunction of the result of the recursive calls with the negation of formulas $D_1, \ldots, D_n$.
 It $t$ has no polarity,
-then we return the union of the domain constraints $D_1 \cup \ldots \cup D_n$.
+then we return the union of $D_1 \cup \ldots \cup D_n$.
 
@@
 the resulting quantified formula is reconstructed with the term $r_1$ returned by the recursive call,
-and a quantified prefix is appended to the domain constraints $D_1$ returned by the recursive call.
+and a quantified prefix is appended to the formula $D_1$ returned by the recursive call.
 In the case that $t$ is a function definition,
@@
 and states that the sum of all positive numbers between $0$ and $c$ is greater than $100$.
+This formula is satisfiable with a model that interprets $c$ as an integer greater than or equal to $14$.
+Due to the universal quantification in the left disjunct,
+state-of-the-art SMT techniques~\cite{GeDeM-CAV-09, ReyEtAl-1-RR-13} are unable to find a model for this formula.
 The signature $\extendsig{\Sigma}$ includes the sort $\fargsort{s}$,
@@
 \end{equation*}
-The universal quantification in the resulting formula (the left disjunct) is over an uninterpreted sort $\fargsort{s}$, 
-thus enabling finite model finding techniques~\cite{ReyEtAl-1-RR-13} to find a model for $\psi$ if a finite interpretation exists for $\fargsort{s}$.
+The universal quantification in the resulting formula is over an uninterpreted sort $\fargsort{s}$, 
+thus enabling finite model finding techniques~\cite{ReyEtAl-1-RR-13} to search for a finite interpretation for $\fargsort{s}$.
 Moreover, all occurrences of $y$ in this formula are beneath applications of the uninterpreted function $\farg{s}$,
@@
 which were constructed from verification conditions from simple Scala programs.
-These benchmarks also heavily involve recursive functions, 
-and cover a wide variety of theories, including bitvectors, arrays, linear and non-linear arithmetic, and inductive datatypes.
+These benchmarks also heavily involve recursive functions over inductive datatypes, 
+and also cover a wide variety of theories, including bitvectors, arrays, and linear and non-linear arithmetic.
 Each of the 245 benchmarks are of the form described in Formula (~\ref{eq:wdf}) from Section~\ref{sec:prelim},
@@
 with another of the same type at another position.
+Note that benchmarks created in this way have a high likelihood of having small and easy-to-find counterexamples.
 In total, we considered 213 mutated forms of conjectures from {\bf isa}, and 427 mutated forms of conjectures from {\bf leon}.
@@
 
-In total, our benchmark set consisted of 885 benchmarks (640 mutants and 245 originals).
-We considered these 885 benchmarks both before and after the translation method $\conv$ as described in Section~\ref{sec:encoding}.
-We will denote the latter encoding using the suffix {\bf w}, for instance {\bf isa+w} or {\bf isa+mw}.
+In total, our benchmark set consisted of 885 benchmarks (640 mutants plus the original 245).
+We considered these 885 benchmarks both before and after the translation $\conv$ as described in Section~\ref{sec:encoding}.
+We will denote the latter encoding using the suffix {\bf w}.
+For instance {\bf isa+w}, contains 79 benchmarks that are the results of applying $\conv$ to each of the 79 benchmarks in {\bf isa}.
 
 For solvers, we considered the SMT solver \ziii~\cite{de-moura-bjoerner-2008}, 
-which runs both heuristic methods for quantifier instantiation,
+which runs both heuristic methods for quantifier instantiation~\cite{DBLP:conf/cade/MouraB07},
 as well as methods for finding models for quantified formulas~\cite{GeDeM-CAV-09}.
 We also considered three configurations of \cvc~\cite{barrett-et-al-2011} which we will refer to as {\bf cvc4}, {\bf cvc4+f} and {\bf cvc4+i}.
-The default configuration {\bf cvc4} runs heuristic quantifier instantiation, but does not include techniques for finite model finding.
-The configuration {\bf cvc4+f} runs a finite model finding procedure~\cite{ReyEtAl-1-RR-13}.
+The default configuration {\bf cvc4} runs heuristic quantifier instantiation, 
+but does not include techniques for finding models.
+The configuration {\bf cvc4+f} runs a finite model finding procedure~\cite{ReyEtAl-1-RR-13, reynolds-et-al-2013}.
 The last configuration {\bf cvc4+i} incorporates techniques for automating inductive reasoning in SMT~\cite{reynolds-kuncak-2015}.
+%All configurations of \cvc incorporate techniques for conflict-based instantiation~\cite{} ..?
+
+\begin{figure}
+\centering
+{
+\begin{tabular}{|l|cc|cc|cc|cc|l|cc|cc|cc|cc|}
+\hline                                                                
+  & \multicolumn{2}{c|}{{\bf z3}}     & \multicolumn{2}{c|}{{\bf cvc4}}     & \multicolumn{2}{c|}{{\bf cvc4+f}}     & \multicolumn{2}{c|}{{\bf cvc4+i}}     & & \multicolumn{2}{c|}{{\bf z3}}     & \multicolumn{2}{c|}{{\bf cvc4}}     & \multicolumn{2}{c|}{{\bf cvc4+f}}     & \multicolumn{2}{c|}{{\bf cvc4+i}}
+\\                                                                    
+  & u & s & u & s & u & s & u & s & & u & s & u & s & u & s & u & s 
+\\                                                                    
+\hline                                                                    
+{\bf isa} & 14  & 0 & 15  & 0 & 15  & 0 & {\bf 61}  & 0 & {\bf isa+w} & 15  & 0 & 15  & 0 & 15  & 0 & 14  & 0
+\\                                                                  
+{\bf leon}  & 73  & 0 & 80  & 0 & 80  & 0 & {\bf 96}  & 0 & {\bf leon+w}  & 78  & 2 & 80  & 0 & 76  & {\bf 9} & 78  & 0
+\\                                                                   
+{\bf isa+m} & 17  & 0 & 18  & 0 & 18  & 0 & {\bf 44}  & 0 & {\bf isa+mw}  & 18  & 35  & 18  & 0 & 18  & {\bf 153} & 17  & 0
+\\                                                                  
+{\bf leon+m}  & 83  & 11  & 103 & 6 & 104 & 6 & {\bf 117} & 6 & {\bf leon+mw} & 98  & 75  & 98  & 6 & 95  & {\bf 169} & 98  & 6
+\\                                                                  
+\hline                                                             
+{\bf total} & 187 & 11  & 216 & 6 & 217 & 6 & {\bf 318} & 6 & {\bf total} & 209 & 112 & 211 & 6 & 204 & {\bf 331} & 207 & 6
+\\                                                                                            
+\hline                                                                                            
+\end{tabular}
+\\
+}
+\caption{Number of unsat and sat responses for of all configurations on benchmarks 
+before and after the translation $\conv$.}
+\label{fig:results}
+\end{figure}
 
 The results are shown in Figure~\ref{fig:results}.
+The results are divided into benchmarks before and after the translation $\conv$.
+First, we note that in no cases did a configuration answer differently across the two encodings.
+In other words, there were no cases in which \ziii or \cvc answered e.g. ``unsatisfiable" on a benchmark $\varphi$ in {\bf isa} 
+and ``satisfiable" on its corresponding benchmark $\conv( \varphi )$ in {\bf isa+w}.
+This provides confirming evidence that the translation $\conv$ is \emph{sound},
+as we did not find a case where $\conv( \varphi )$ was unsatisfiable when $\varphi$ was satisfiable.
+It also provides confirming evidence that the translation is \emph{complete} for this set of benchmarks,
+as we did not find a case where $\conv( \varphi )$ was satisfiable when $\varphi$ was unsatisfiable.
+This is to be expected, since the functions in our benchmark sets are intended to be logically consistent.
+
+Examining the first encoding in isolation (the first nine columns),
+the number of satisfiable responses from all configurations is quite low.
+In total, {\bf z3} manages to find only 11 benchmarks to be satisfiable,
+whereas each configuration of \cvc only finds 6.
+This confirms the shortcomings of existing SMT techniques for finding counterexamples for benchmarks containing recursive functions.
+In terms of number of unsatisfiable responses,
+the configuration {\bf cvc4+i} is the clear winner, finding 318 total benchmarks to be unsatisfiable.
+
+Now, examining the second encoding (the latter nine columns), 
+we see that using the translation $\conv$ drastically improves the ability of configurations for determining satisfiability.
+In total, {\bf cvc4+f} finds 331 of the 885 benchmarks to be satisfiable, including 9 benchmarks in the original {\bf leon} benchmark set.
+The performance of {\bf z3} for counterexamples also improves drastically, as it solves 112 satisfiable,
+including 10 that were unsolved by {\bf cvc4+f}.
+Moreover, the translation $\conv$ helps \ziii for unsatisfiable responses as well,
+as we see \ziii solves a total of 209 whereas it solves only 187 before the translation.
+This number is 2 fewer than {\bf cvc4}, which solves 211 unsatisfiable.
+Thus, in most cases the translation $\conv$ does not seriously degrade performance for unsatisfiable benchmark,
+and in some cases may actually aid the solver for determining unsatisfiability
+\footnote{The exception here is {\bf cvc4+i}, which makes use of type information of quantified formulas to make inductive arugments.}.
 
@@
   * future work
-    * (co)inductive predicates
+    * (co)inductive predicates?
+    * quotient types??
 

### 2015-04-29 tex/main.tex
Minor

@@
 \newcommand{\terms}{\mathbf{T}}
-\newcommand{\functerms}{\mathbf{DFT}}
+%\newcommand{\functerms}{\mathbf{DFT}}
 \newcommand{\M}{\mathcal{M}}
@@
 We write $\terms( t )$ to denote the set of subterms in $t$.
-We write $\functerms^\Sigma( t )$ to denote the set of $f$-applications in $t$ such that $f \in \sfundefs{\Sigma}$.
+%We write $\functerms^\Sigma( t )$ to denote the set of $f$-applications in $t$ such that $f \in \sfundefs{\Sigma}$.
 
@@
    \begin{itemize}
-     \item[] $\mathsf{return}$ $( t, \{ \exists z\typ{\fargsort{f}}. ( \vecfarg{f}( z ) \teq \vec t ) \mid f( \vec t ) \in \functerms^\Sigma( t ) \} )$
+     \item[] $\mathsf{return}$ $( t, \{ \exists z\typ{\fargsort{f}}. ( \vecfarg{f}( z ) \teq \vec t ) \mid f( \vec t ) \in \terms( t ), f \in \sfundefs{\Sigma} \} )$
    \end{itemize}
@@
 Due to the universal quantification in the left disjunct,
-state-of-the-art SMT techniques~\cite{GeDeM-CAV-09, ReyEtAl-1-RR-13} are unable to find a model for this formula.
+current SMT techniques~\cite{GeDeM-CAV-09, ReyEtAl-1-RR-13} are unable to find a model for this formula.
 The signature $\extendsig{\Sigma}$ includes the sort $\fargsort{s}$,
@@
 Now, examining the second encoding (the latter nine columns), 
-we see that using the translation $\conv$ drastically improves the ability of configurations for determining satisfiability.
+we see that using the translation $\conv$ drastically improves the ability of configurations for answering satisfiable.
 In total, {\bf cvc4+f} finds 331 of the 885 benchmarks to be satisfiable, including 9 benchmarks in the original {\bf leon} benchmark set.

### 2015-04-29 tex/main.tex
Rearranging, add intro text to main section, more additions.

@@
 
-\section{Related Work}
-
-first-order model finders: SEM, MACE/Mace, Paradox, iProver
-  either direct search or reduction to SAT
-
-Alloy (and Alloy idiom by Kuncak \& Jackson), Kodkod
-
-The model finder for KIV based on Alloy
-
-Refute
-  * unsound handling of datatypes
-
-Nitpick
-  * inductive / coinductive predicates
-  * quotient types
-
-Leon
-
-\subsection{Model Finding for Quantified Formulas in SMT}
-
-z3, cvc4
-
-\section{Approach}
-
-\subsection{Preliminaries}
+\section{Preliminaries}
 \label{sec:prelim}
@@
 
-\subsection{Encoding Well-Defined Assumption for Recursive Functions}
+\section{Related Work}
+
+first-order model finders: SEM, MACE/Mace, Paradox, iProver
+  either direct search or reduction to SAT
+
+Alloy (and Alloy idiom by Kuncak \& Jackson), Kodkod
+
+The model finder for KIV based on Alloy
+
+Refute
+  * unsound handling of datatypes
+
+Nitpick
+  * inductive / coinductive predicates
+  * quotient types
+
+Leon
+
+\section{Model Finding for Recursive Functions in SMT}
+
+Recent techniques~\cite{GeDeM-CAV-09, ReyEtAl-1-RR-13} 
+have focused on finding models for quantified formulas in SMT.
+A instatiation-based procedure is shown in~\cite{GeDeM-CAV-09} to be complete for formulas in the \emph{essentially uninterpreted} fragment.
+This fragment is limited to universally quantified formulas where all occurrences of variables occur as direct subchildren of uninterpreted functions.
+For instance, the formula $\forall x\typ{\sortint}. f( x ) \teq g( x ) + 5$ 
+is in the essentially uninterpreted fragment since all occurrences of $x$ occur beneath uninterpreted functions $f$ and $g$.
+Other syntactic criteria are identified in this paper that extend this fragment slightly, including cases when variables occur as arguments of arithmetic predicates.
+The techniques in~\cite{ReyEtAl-1-RR-13} can be used for finding finite models for quantified formulas over uninterpreted types as well as types having a 
+fixed finite interpretation (such as fixed-width bitvectors, datatype enumerations, etc.).
+For instance, 
+finite model finding techniques can find a model for
+the formula $\forall x\typ{\tau} y\typ{\tau} z\typ{\tau}. x \teq y \vee y \teq z \vee x \teq z$
+where the uninterpreted type $\tau$ is interpreted as a set of (at most) two elements.
+
+Unfortunately, these fragments do not coincide with the vast majority of quantified formulas that correspond to recursive function definitions.
+In particular, the fragments in~\cite{GeDeM-CAV-09} are not applicable when the argument of a recursive function is used in various ways (e.g. within an arithmetic expression), 
+while the fragment in~\cite{ReyEtAl-1-RR-13} is not applicable for functions over infinite domains such as the integers.
+So, for instance, the quantified formula $\forall x\typ{\sortint}. \lite( p \leq 0, p( x ) \teq 0, p( x ) \teq 2 * p( x - 1 ) )$ 
+is not in the essentially uninterpreted fragment (since $x$ occurs as a child of $-$),
+nor does finite model finding apply (since $x$ is of type $\sortint$).
+
+Here, we advocate an approach that translates formulas involving function definitions (such as the one mentioned in the previous paragraph)
+into equisatisfiable formulas where the techniques~\cite{GeDeM-CAV-09, ReyEtAl-1-RR-13} are applicable.
+
+\subsection{A Translation for Function Definitions}
 \label{sec:encoding}
@@
 In the remainder of the section, we fix a signature $\Sigma$.
-
 Say that $\Sigma$-formula $\varphi$ is well-defined with respect to functions $\sfundefs{\Sigma}$, 
-and the definitions in $\sfundefs{\Sigma}$ are logically consistent \ajr{formalize}.
-In this section, we present a method which constructs an extended signature $\Sigma'$,
-and translates $\varphi$ to equisatisfiable $\Sigma'$-formula $\psi$ on which known model-finding procedures~\cite{GeDeM-CAV-09, ReyEtAl-1-RR-13} are applicable.
-The intuition behind this translation is use an uninterpreted type to abstract the set of tuples that each recursive function $f$ is invoked on,
+and the definitions in $\sfundefs{\Sigma}$ are logically consistent.
+We may translate $\varphi$ into a equisatisfiable formula $\psi$ for which known model-finding procedures~\cite{GeDeM-CAV-09, ReyEtAl-1-RR-13} are applicable.
+In particular, this section presents a method for constructing an extended signature $\extendsig( \Sigma )$,
+and a $\extendsig( \Sigma )$-formula $\psi$ such that $\varphi$ and $\psi$ are equisatisfiable.
+The intuition behind this translation is use an uninterpreted type to abstract the set of \emph{relevant} tuples for each recursive function $f$,
 and restrict the quantification of the function definition for $f$ to a variable of this type.
+Informally, the relevant tuples $\vec t$ of a function $f$ are the ones for which the interpretation of $f( \vec t )$ is relevant to the satisfiability of $\varphi$.
 
@@
 At a high level, the translation modifies $\varphi$ in two ways.
-First, it restricts the quantification on function definitions for $f$ to the corresponding uninterpreted sort $\fargsort{f}$.
+First, it restricts the quantification on function definitions for $f$ to the corresponding uninterpreted type $\fargsort{f}$.
 Second, it augments $\varphi$ with additional constraints of the form $\exists z\typ{\fargsort{f}}. ( \vecfarg{f}( z ) \teq \vec t )$
@@
 
-We demonstrate this translation with a simple example.
+We demonstrate this translation with an example.
 
@@
 Consider the $\Sigma$-formula $\varphi$:
-\begin{equation*}
+\begin{eqnarray} \label{eq:ex-before}
 \forall_{s} x\typ{\sortint}. ite( x \leq 0, s(x) \teq 0, s( x ) \teq x + s( x-1 ) ) \wedge s( c ) > 100
-\end{equation*}
+\end{eqnarray} 
 Here, $s$ is a function that returns the sum of positive numbers between $0$ and its argument $x$.
@@
 This formula is satisfiable with a model that interprets $c$ as an integer greater than or equal to $14$.
-Due to the universal quantification in the left disjunct,
+Due to the universally quantified formula in the left disjunct,
 current SMT techniques~\cite{GeDeM-CAV-09, ReyEtAl-1-RR-13} are unable to find a model for this formula.
-The signature $\extendsig{\Sigma}$ includes the sort $\fargsort{s}$,
-and the uninterpreted function $\farg{s}$ of sort $\fargsort{s} \rightarrow \sortint$.
+The signature $\extendsig{\Sigma}$ includes the type $\fargsort{s}$,
+and the uninterpreted function $\farg{s}$ of type $\fargsort{s} \rightarrow \sortint$.
 The result of $\conv( \varphi, \ptrue )$ is the pair $( \psi, \emptyset )$, where after simplification, $\psi$ is the $\extendsig{\Sigma}$-formula:
-\begin{equation*}
+\begin{eqnarray} \label{eq:ex-after}
 \begin{split}
@@
 \wedge ( s( c ) > 100 \wedge \exists z\typ{\fargsort{s}}. \farg{s}( z ) \teq c )
-\end{equation*}
-The universal quantification in the resulting formula is over an uninterpreted sort $\fargsort{s}$, 
+\end{eqnarray} 
+The universal quantification in the resulting formula is over an uninterpreted type $\fargsort{s}$, 
 thus enabling finite model finding techniques~\cite{ReyEtAl-1-RR-13} to search for a finite interpretation for $\fargsort{s}$.
@@
 implying that the formula is in the essentially uninterpreted fragment of~\cite{GeDeM-CAV-09}, for which there exists a complete instantiation procedure.
+We observe that both \cvc and \ziii run indefinitely on formula~(\ref{eq:ex-before}),
+and both produce a model for formula~(\ref{eq:ex-after}) in $<0.1$ second.
 $\square$
@@
 
-It remains to show that our translation produces an equisatisfiable formulae.
-Namely,
-we claim that if $\varphi$ is well-defined with respect to functions $\sfundefs{\Sigma}$,
-function definitions in $\sfundefs{\Sigma}$ are logically consistent \ajr{formalize},
-then $\conv(\varphi)$ is equisatisfiable to $\varphi$.
+It is important to note that the translation $\conv$ does not preserve the models of $\varphi$.
+One model $\M$ for formula~(\ref{eq:ex-after}) in the previous example interprets 
+$\fargsort{s}$ as a finite set of elements $\{ u_0, \ldots, u_{14} \}$,
+$\farg{s}$ as a finite map mapping containing $u_i \mapsto i$ for $i = 0, \ldots, 14$,
+$c$ as $14$, 
+and $s$ as the (almost constant) function:
+\begin{eqnarray}
+\lambda x\typ{Int}. \lite( x \teq 0, 0, \lite( x \teq 1, 1, \lite( x \teq 2, 3, \ldots, \lite( x \teq 13, 91, 105 ) \ldots )
+\end{eqnarray}
+In other words, $s$ is a function mapping $x$ to the sum of all positive integers between $0$ and $x$ when $0 \leq x \leq 13$,
+and $105$ otherwise.
+Notice that $\M$ is \emph{not} a model for the original formula~(\ref{eq:ex-before}),
+since $\M$ interprets e.g. $s( -1 )$ and $s( 15 )$ as $105$.
+However, under the assumption that function definitions in $\sfundefs{\Sigma}$ are logically consistent, 
+we claim that $\conv(\varphi)$ is equisatisfiable to $\varphi$.
+The intuition is that the interpretation of a term such as $s( -1 )$ is not relevant to the satisfiability of~(\ref{eq:ex-before}).
+\ajr{add invariants, proofs}
 
-\begin{lemma}
-Let $\varphi$ be a $\Sigma$-formula that is well-defined with respect to functions $\sfundefs{\Sigma}$.
-\ajr{todo}
-\end{lemma}
+\begin{theorem}
+If $\varphi$ be a $\Sigma$-formula that is well-defined with respect to functions $\sfundefs{\Sigma}$,
+and function definitions in $\sfundefs{\Sigma}$ are logically consistent,
+then $\varphi$ and $\conv( \varphi )$ are equisatisfiable.
+\end{theorem}
 
@@
 and also cover a wide variety of theories, including bitvectors, arrays, and linear and non-linear arithmetic.
-Each of the 245 benchmarks are of the form described in Formula (~\ref{eq:wdf}) from Section~\ref{sec:prelim},
-namely they each are a conjoined list of function definitions followed by a (negated) conjecture.
+Each of the 245 benchmarks are of the form described in formula~(\ref{eq:wdf}) from Section~\ref{sec:prelim},
+in particular they each are a conjoined list of function definitions followed by a (negated) conjecture.
 A majority of these benchmarks are unsatisfiable, although a handful of the benchmarks in the {\bf leon} set are satisfiable.
@@
 but does not include techniques for finding models.
-The configuration {\bf cvc4+f} runs a finite model finding procedure~\cite{ReyEtAl-1-RR-13, reynolds-et-al-2013}.
+The configuration {\bf cvc4+f} runs heuristic instantiation and the finite model finding procedure described in~\cite{ReyEtAl-1-RR-13, reynolds-et-al-2013}.
 The last configuration {\bf cvc4+i} incorporates techniques for automating inductive reasoning in SMT~\cite{reynolds-kuncak-2015}.
@@
 First, we note that in no cases did a configuration answer differently across the two encodings.
-In other words, there were no cases in which \ziii or \cvc answered e.g. ``unsatisfiable" on a benchmark $\varphi$ in {\bf isa} 
-and ``satisfiable" on its corresponding benchmark $\conv( \varphi )$ in {\bf isa+w}.
+In other words, there were no cases in which \ziii or \cvc answered e.g. ``unsatisfiable" on a benchmark $\varphi$
+and ``satisfiable" on its corresponding benchmark $\conv( \varphi )$, or vice versa.
 This provides confirming evidence that the translation $\conv$ is \emph{sound},
@@
 and in some cases may actually aid the solver for determining unsatisfiability
-\footnote{The exception here is {\bf cvc4+i}, which makes use of type information of quantified formulas to make inductive arugments.}.
+\footnote{The exception here is {\bf cvc4+i}, which makes use of type information associated with quantified formulas for inductive arguments.}.
+
+We conclude from this evaluation that the translation $\conv$ enables SMT solvers to find counterexamples 
+for conjectures involving recursive functions whose definitions are logically consistent.
+Both {\bf z3} and {\bf cvc4+f} were able to find a large number of counterexamples after the translation,
+with the latter configuration finding significantly more instances (331 to 112).
+
+\section{Front-End Support for Recursive Function Definitions}
+\label{sec:front-end}
+
+We have seen a satisfiability-preserving translation $\conv$ that significantly improves the effectiveness of known SMT techniques
+for finding counterexamples for formulas involving recursive function definitions.
+For convienience, the translation $\conv$ has been implemented in the latest development version of \cvc~\cite{ReyEtAl-1-RR-13}.
 

### 2015-04-29 tex/main.tex
Mention front end.

@@
 
+\newcommand{\definefunreccmd}{\con{define}\text{-}\con{fun}\text{-}\con{rec}}
+\newcommand{\definefunsreccmd}{\con{define}\text{-}\con{funs}\text{-}\con{rec}}
+
 %\input{scalalistings}
@@
 
+\section{Related Work}
+
+first-order model finders: SEM, MACE/Mace, Paradox, iProver
+  either direct search or reduction to SAT
+
+Alloy (and Alloy idiom by Kuncak \& Jackson), Kodkod
+
+The model finder for KIV based on Alloy
+
+Refute
+  * unsound handling of datatypes
+
+Nitpick
+  * inductive / coinductive predicates
+  * quotient types
+
+Leon
+
+in smt : z3, cvc4 (briefly, text in main section gives detail)
+
 \section{Preliminaries}
@@
 
-\section{Related Work}
-
-first-order model finders: SEM, MACE/Mace, Paradox, iProver
-  either direct search or reduction to SAT
-
-Alloy (and Alloy idiom by Kuncak \& Jackson), Kodkod
-
-The model finder for KIV based on Alloy
-
-Refute
-  * unsound handling of datatypes
-
-Nitpick
-  * inductive / coinductive predicates
-  * quotient types
-
-Leon
-
 \section{Model Finding for Recursive Functions in SMT}
+\label{sec:model-finding-rec-fun}
 
@@
 \begin{example}
+\label{ex:translation}
 Consider the (combined) signature $\Sigma = \Sigmalia \cup \Sigma_u$,
@@
 \section{Evaluation}
+\label{sec:evaluation}
 
@@
 
-\begin{figure}
+\begin{figure}[t]
 \centering
@@
 For convienience, the translation $\conv$ has been implemented in the latest development version of \cvc~\cite{ReyEtAl-1-RR-13}.
+Function definitions $\forallf{f} \vec x. \varphi$ can be written in using the $\definefunreccmd$ command, 
+proposed in the draft of SMT version 2.5~\cite{}.
+For instance, the input from Example~\ref{ex:translation} could be encoded as:
+
+\begin{verbatim}
+(define-fun-rec s ((x Int)) Int (ite (<= x 0) 0 (s (- x 1))))
+(assert (> (s x) 100))
+(check-sat)
+\end{verbatim}
+
+When reading this input, 
+\cvc adds the (annotated) quantified formula $\forallf{s} x. s( x ) \teq \lite( x \leq 0, 0, s( x-1 ) )$ to its list of assertions,
+which after rewriting becomes $\forallf{s} x. \lite( x \leq 0, s( x ) \teq 0, s( x ) \teq s( x-1 ) )$.
+If \cvc is in finite model finding mode for recursive functions is enabled (using the command line parameter ``--fmf-fun"),
+then it will replace its list of known assertions based on the conversion $\conv$ before checking for satisfiability.
 
@@
 \ackname.}
-
+We would like to thank Damien Busato-Gaston and Emmanouil Koukoutos for providing the initial set of benchmarks used in the evaluation.
 This research is partially supported by the Inria technological development

### 2015-04-29 tex/main.tex
Fixes, more in prelim.

@@
 
+\newcommand{\Mo}{\mathbf{I}}
+
+\newcommand{\euf}{\ensuremath{\mathrm{E}}\xspace}
+\newcommand{\ari}{\ensuremath{\mathrm{A}}\xspace}
+
 %\input{scalalistings}
@@
 We will write $t\typ{\tau}$ to denote a term $t$ of type~$\tau$.
+We write $\terms( t )$ to denote the set of subterms in $t$.
+%We write $\functerms^\Sigma( t )$ to denote the set of $f$-applications in $t$ such that $f \in \sfundefs{\Sigma}$.
+
+A \emph{theory} is a pair $T = (\Sigma, \Mo)$ where 
+$\Sigma$ is a signature and  $\Mo$ is a class of $\Sigma$-interpretations,
+the \emph{models} of $T$.
+A $\Sigma$-formula $\varphi$ is 
+\emph{$T$-satisfiable} (resp., \emph{$T$-unsatisfiable}) 
+if it is satisfied by some (resp., no) interpretation in $\Mo$.
+Formulas $\varphi$ and $\psi$ are \emph{equisatisfiable} if
+$\varphi$ is $T$-satisfiable if and only if $\psi$ is $T$-satisfiable.
+%For a given signature $\Sigma$
+%the \emph{theory of equality with uninterpreted functions} \euf
+%consists of the set of all $\Sigma$-interpretations.
+%Informally, we refer to the sort and function symbols in this theory 
+%as \emph{uninterpreted}.
 
@@
 and $\exists \vec x. \varphi$ as shorthand for $\neg \forall \vec x. \neg \varphi$.
-
-\ajr{theories, models, equisatisfiable, etc.}
 
@@
 and consider $\Sigma$-formulas that are well-defined with respect to functions $\sfundefs{\Sigma}$.
-
-We write $\terms( t )$ to denote the set of subterms in $t$.
-%We write $\functerms^\Sigma( t )$ to denote the set of $f$-applications in $t$ such that $f \in \sfundefs{\Sigma}$.
 
@@
 We may translate $\varphi$ into a equisatisfiable formula $\psi$ for which known model-finding procedures~\cite{GeDeM-CAV-09, ReyEtAl-1-RR-13} are applicable.
-In particular, this section presents a method for constructing an extended signature $\extendsig( \Sigma )$,
-and a $\extendsig( \Sigma )$-formula $\psi$ such that $\varphi$ and $\psi$ are equisatisfiable.
+In particular, this section presents a method for constructing an extended signature $\extendsig{ \Sigma }$,
+and a $\extendsig{ \Sigma }$-formula $\psi$ such that $\varphi$ and $\psi$ are equisatisfiable.
 The intuition behind this translation is use an uninterpreted type to abstract the set of \emph{relevant} tuples for each recursive function $f$,
@@
 However, under the assumption that function definitions in $\sfundefs{\Sigma}$ are logically consistent, 
-we claim that $\conv(\varphi)$ is equisatisfiable to $\varphi$.
-The intuition is that the interpretation of a term such as $s( -1 )$ is not relevant to the satisfiability of~(\ref{eq:ex-before}).
+we claim that $\conv(\varphi)$ is equisatisfiable to $\varphi$, for all $\varphi$.
+The intuition is that the interpretation of a term such as $s( -1 )$ in the previous example is not relevant to the satisfiability of formula~(\ref{eq:ex-before}).
 \ajr{add invariants, proofs}
@@
 For benchmarks, we gathered a set of 245 benchmarks from two sources, which we will refer to as {\bf isa} and {\bf leon}.
-The first ({\bf isa}) consists of the 79 benchmarks from the Isaplanner benchmark suite~\cite{DBLP:conf/itp/JohanssonDB10} that did not contain higher-order functions.
+The first ({\bf isa}) consists of the 79 benchmarks from the Isaplanner benchmark suite~\cite{DBLP:conf/itp/JohanssonDB10} that do not contain higher-order functions.
 The benchmarks have been used recently as challenge problems for a variety of inductive theorem provers.
@@
 This number is 2 fewer than {\bf cvc4}, which solves 211 unsatisfiable.
-Thus, in most cases the translation $\conv$ does not seriously degrade performance for unsatisfiable benchmark,
+Thus, in most cases the translation $\conv$ does not significantly degrade performance for unsatisfiable benchmarks,
 and in some cases may actually aid the solver for determining unsatisfiability
@@
 
-\section{Front-End Support for Recursive Function Definitions}
+\section{Front-End Support for Recursive Function Definitions in CVC4}
 \label{sec:front-end}
@@
 proposed in the draft of SMT version 2.5~\cite{}.
-For instance, the input from Example~\ref{ex:translation} could be encoded as:
+For instance, the input from Example~\ref{ex:translation} can be written as:
 

### 2015-04-30 tex/main.tex
More text, contributions.

@@
 
-\ajr{(1) Encoding, (2) Evaluation, (3) Front-end support for define-funs-rec in \cvc}
+%\ajr{(1) Encoding, (2) Evaluation, (3) Front-end support for define-funs-rec in \cvc}
+
+In this paper, we make the following contributions:
+\begin{itemize}
+\item[-] We define a satisfiability-preserving translation method for formulas containing recursive function definitions
+that produces formulas for which existing model finding techniques~\cite{GeDeM-CAV-09, ReyEtAl-1-RR-13} are applicable.
+\item[-] We provide experimental evidence that this translation improves the effectiveness of SMT solvers \cvc and \ziii for finding counterexamples to verification conditions.
+\item[-] We detail front-end support for recursive function definitions and this translation in the SMT solver \cvc.
+\end{itemize}
 
@@
 However, under the assumption that function definitions in $\sfundefs{\Sigma}$ are logically consistent, 
-we claim that $\conv(\varphi)$ is equisatisfiable to $\varphi$, for all $\varphi$.
+we claim that $\conv(\varphi)$ is equisatisfiable to $\varphi$ for any input $\varphi$.
 The intuition is that the interpretation of a term such as $s( -1 )$ in the previous example is not relevant to the satisfiability of formula~(\ref{eq:ex-before}).
+Thus, a satisfiable (resp. unsatisfiable) response from an SMT solver on input $\conv( \varphi )$ implies the existence (resp. non-existence) of a model for $\varphi$.
+Moreover, information in models for $\conv( \varphi )$ contains relevant information regarding the models of $\varphi$.
+For instance, the model $\M$ for formula~(\ref{eq:ex-after}) described above interprets $c$ as $14$.
+In fact, we claim that there exists a model of $\varphi$ that coincides with $\M$ on its interpretation of all functions in $\sfuns{\Sigma}$ 
+other than those in $\sfundefs{\Sigma}$, i.e. $\{ s \}$ in the previous example.
+From a practical point of view, this is in fact a non-issue, as $s$ is the very function that was explicitly defined by the user,
+and hence already has an intended interpretation.
+
 \ajr{add invariants, proofs}
@@
 for finding counterexamples for formulas involving recursive function definitions.
-For convienience, the translation $\conv$ has been implemented in the latest development version of \cvc~\cite{ReyEtAl-1-RR-13}.
+For convienience, the translation $\conv$ has been implemented in the latest development version of \cvc~\cite{ReyEtAl-1-RR-13} (version 1.5 pre-release).
 Function definitions $\forallf{f} \vec x. \varphi$ can be written in using the $\definefunreccmd$ command, 

### 2015-04-30 tex/main.tex
Minor

@@
 Moreover, information in models for $\conv( \varphi )$ contains relevant information regarding the models of $\varphi$.
-For instance, the model $\M$ for formula~(\ref{eq:ex-after}) described above interprets $c$ as $14$.
-In fact, we claim that there exists a model of $\varphi$ that coincides with $\M$ on its interpretation of all functions in $\sfuns{\Sigma}$ 
+For instance, the model $\M$ for formula~(\ref{eq:ex-after}) described above interprets $c$ as $14$, 
+while there exists a model of formula~(\ref{eq:ex-before}) also interprets $c$ as $14$.
+In fact, we claim that there exists a model of $\varphi$ that coincides with a model of $\conv( \varphi )$ on its interpretation of all functions in $\sfuns{\Sigma}$ 
 other than those in $\sfundefs{\Sigma}$, i.e. $\{ s \}$ in the previous example.

### 2015-04-30 tex/main.tex
Minor

@@
 and the definitions in $\sfundefs{\Sigma}$ are logically consistent.
-We may translate $\varphi$ into a equisatisfiable formula $\psi$ for which known model-finding procedures~\cite{GeDeM-CAV-09, ReyEtAl-1-RR-13} are applicable.
-In particular, this section presents a method for constructing an extended signature $\extendsig{ \Sigma }$,
+%We may translate $\varphi$ into a equisatisfiable formula $\psi$ for which known model-finding procedures~\cite{GeDeM-CAV-09, ReyEtAl-1-RR-13} are applicable.
+This section presents a method for constructing an extended signature $\extendsig{ \Sigma }$,
 and a $\extendsig{ \Sigma }$-formula $\psi$ such that $\varphi$ and $\psi$ are equisatisfiable.

### 2015-04-30 tex/main.tex
tweaked title
Fix terminology

@@
 
-\title{Model Finding for Recursive Functions in SMT}
+%%% I added (Co) -- hope you like it --JB
+\title{Model Finding for (Co)recursive Functions in SMT}
 
@@
 and standard boolean connectives ($\neg$, $\wedge$, $\vee$, $\lite$, etc.).
-%and a function $\lite$ of type $\Bool \times \sigma \times \sigma$ for each $\sigma \in \ssorts{\Sigma}$.
+and a function $\lite$ of type $\Bool \times \sigma \times \sigma \rightarrow \lite$ for each $\sigma \in \ssorts{\Sigma}$.
 We will write $t\typ{\tau}$ to denote a term $t$ of type~$\tau$.
@@
 $\varphi$ is $T$-satisfiable if and only if $\psi$ is $T$-satisfiable.
-%For a given signature $\Sigma$
-%the \emph{theory of equality with uninterpreted functions} \euf
-%consists of the set of all $\Sigma$-interpretations.
-%Informally, we refer to the sort and function symbols in this theory 
-%as \emph{uninterpreted}.
+For a given signature $\Sigma$
+the \emph{theory of equality with uninterpreted functions} \euf
+consists of the set of all $\Sigma$-interpretations.
+We refer to the sort and function symbols in this theory as \emph{uninterpreted}.
 
@@
 
+A set of function definitions $\{ \varphi_1, \ldots, \varphi_n \}$ 
+is \emph{well-specified (in $T$)} if $\varphi_1 \wedge \ldots \wedge \varphi_n$ is $T$-satisfiable.
+For instance, $\{ \forallf{f} \vec x. f( x ) \teq 0 \}$ is well-specified,
+whereas $\{ \forallf{f} x. f( x ) \teq f( x ) + 1 \}$ is not. 
+Note that a well-specified function definition need not specify a terminating function, e.g.
+$\{ \forallf{f} x^\sortint. f( x ) \teq f( x - 1 ) \}$ or even $\{ \forallf{f} x. f( x ) \teq f( x ) \}$ are well-specified.
+\ajr{possibly revise}
+
 \section{Model Finding for Recursive Functions in SMT}
@@
 finite model finding techniques can find a model for
-the formula $\forall x\typ{\tau} y\typ{\tau} z\typ{\tau}. x \teq y \vee y \teq z \vee x \teq z$
-where the uninterpreted type $\tau$ is interpreted as a set of (at most) two elements.
+the formula $\forall x\typ{\tau} y\typ{\tau}. \neg x \teq y \vee f( x ) \teq f( y )$
+where the uninterpreted type $\tau$ is interpreted as a finite set, and $f$ as injective finite map.
 
@@
 
-\subsection{A Translation for Function Definitions}
+\subsection{A Translation for Well-Specified Function Definitions}
 \label{sec:encoding}
@@
 Say that $\Sigma$-formula $\varphi$ is well-defined with respect to functions $\sfundefs{\Sigma}$, 
-and the definitions in $\sfundefs{\Sigma}$ are logically consistent.
+and the set of function definitions corresponding to $\sfundefs{\Sigma}$ are well-specified.
 %We may translate $\varphi$ into a equisatisfiable formula $\psi$ for which known model-finding procedures~\cite{GeDeM-CAV-09, ReyEtAl-1-RR-13} are applicable.
@@
 \vspace{-2ex}
-\caption{A translation procedure $\conv$ for a $\Sigma$-formula that is well-defined with respect to function definitions.
+\caption{A translation procedure $\conv$ for a $\Sigma$-formula that is well-defined with respect to a set of well-specified function definitions $\sfundefs{\Sigma}$.
 In the subprocedure $\conv_0$, the procedure $\pol( \boolop, i, p )$ returns the polarity of the $i^{th}$ child of a $\boolop$-application having polarity $p$,
@@
 since $\M$ interprets e.g. $s( -1 )$ and $s( 15 )$ as $105$.
-However, under the assumption that function definitions in $\sfundefs{\Sigma}$ are logically consistent, 
+However, under the assumption that function definitions in $\sfundefs{\Sigma}$ are well-specified, 
 we claim that $\conv(\varphi)$ is equisatisfiable to $\varphi$ for any input $\varphi$.
@@
 If $\varphi$ be a $\Sigma$-formula that is well-defined with respect to functions $\sfundefs{\Sigma}$,
-and function definitions in $\sfundefs{\Sigma}$ are logically consistent,
+and the set of function definitions corresponding to functions in $\sfundefs{\Sigma}$ is well-specified,
 then $\varphi$ and $\conv( \varphi )$ are equisatisfiable.
@@
 and ``satisfiable" on its corresponding benchmark $\conv( \varphi )$, or vice versa.
-This provides confirming evidence that the translation $\conv$ is \emph{sound},
+This provides confirming evidence that the translation $\conv$ is \emph{refutation-sound},
 as we did not find a case where $\conv( \varphi )$ was unsatisfiable when $\varphi$ was satisfiable.
-It also provides confirming evidence that the translation is \emph{complete} for this set of benchmarks,
+It also provides confirming evidence that the translation is \emph{model-sound} for this set of benchmarks,
 as we did not find a case where $\conv( \varphi )$ was satisfiable when $\varphi$ was unsatisfiable.
-This is to be expected, since the functions in our benchmark sets are intended to be logically consistent.
+This is to be expected, since the functions in our benchmark sets are expected to be well-specified (and in fact, terminating).
 
@@
 We conclude from this evaluation that the translation $\conv$ enables SMT solvers to find counterexamples 
-for conjectures involving recursive functions whose definitions are logically consistent.
+for conjectures involving recursive functions whose definitions are well-specified.
 Both {\bf z3} and {\bf cvc4+f} were able to find a large number of counterexamples after the translation,
@@
 then it will replace its list of known assertions based on the conversion $\conv$ before checking for satisfiability.
+We stress that this option is model-sound only if the functions provided in the input are well-specified.
 

### 2015-04-30 tex/main.tex
Revisions, refutation-soundness argument.

@@
 %\newcommand{\functerms}{\mathbf{DFT}}
+\newcommand{\vars}{\mathbf{V}}
 \newcommand{\M}{\mathcal{M}}
+\newcommand{\I}{\mathcal{I}}
+\newcommand{\J}{\mathcal{J}}
 \def\AIF{\qtab\keyword{if}\ }
@@
 \begin{itemize}
-\item[-] We define a satisfiability-preserving translation method for formulas containing recursive function definitions
+\item[-] We define a satisfiability-preserving translation method for a class of formulas containing recursive function definitions
 that produces formulas for which existing model finding techniques~\cite{GeDeM-CAV-09, ReyEtAl-1-RR-13} are applicable.
@@
 
-In this paper, we associate quantified formulas with the uninterpreted functions whose semantics they define.
-We will write $\forallf{f} \vec x. \varphi$ to denote a quantified formula that is defines the semantics of uninterpreted function $f$.
-For consistency, we consider only quantified formulas of this form, 
-and allow $f$ to be a distinguished constant $\fnull$, 
-denoting that the quantified formula does not define the semantics of an uninterpreted function.
+%In this paper, we associate quantified formulas with uninterpreted functions (informally, those whose semantics they define).
+In this paper, we consider \emph{annotated quantifed formulas} of the form $\forallf{f} \vec x. \varphi$, 
+where $f$ is an uninterpreted function from $\sfuns{\Sigma}$, or the distinguished constant $\fnull$.
 When $f$ is not $\fnull$, we will refer to $\forallf{f} \vec x. \varphi$ as a \emph{function definition} (for $f$).
@@
 where $\sfundefs{\Sigma}$ is the set of \emph{defined} function symbols,
-and $\sfunndefs{\Sigma}$ are the others,
 and consider $\Sigma$-formulas that are well-defined with respect to functions $\sfundefs{\Sigma}$.
@@
       \item[] $\mathsf{let} ( r_1, D_1 ) = \conv_0( t_1, p )$;
-      \item[] $\mathsf{if}$ $f \equiv \fnull$
+      \item[] $\mathsf{if}$ $f \in \sfundefs{\Sigma}$
       \begin{itemize}
-        \item[] $\mathsf{return}$ $( \forall \vec x. r_1, \forall \vec x. D_1 )$
+        %\item[] $\mathsf{assert}( D_1 = \emptyset )$;
+        \item[] $\mathsf{return}$ $( \forall y\typ{\fargsort{f}}. ( r_1 [ \vecfarg{f}( y ) / \vec x ] ), \emptyset )$
       \end{itemize}
@@
       \begin{itemize}
-        %\item[] $\mathsf{assert}( D_1 = \emptyset )$;
-        \item[] $\mathsf{return}$ $( \forall y\typ{\fargsort{f}}. ( r_1 [ \vecfarg{f}( y ) / \vec x ] ), \emptyset )$
+        \item[] $\mathsf{return}$ $( \forall \vec x. r_1, \forall \vec x. D_1 )$
       \end{itemize}
@@
 In the subprocedure $\conv_0$, the procedure $\pol( \boolop, i, p )$ returns the polarity of the $i^{th}$ child of a $\boolop$-application having polarity $p$,
-where $\boolop$ is an (interpreted) predicate.
+where $\boolop$ is an interpreted predicate.
 }
@@
 We extend the signature of $\Sigma$ to a signature $\extendsig{\Sigma}$ containing the following additional definitions.
-For each $f \in \sfundefs{\Sigma}$ of type $\tau_1 \times \ldots \tau_n \rightarrow \tau$, 
+For each $f \in \sfundefs{\Sigma}$ of type $\tau_1 \times \ldots \times \tau_n \rightarrow \tau$, 
 the signature $\extendsig{\Sigma}$ includes:
@@
 In the case that $t$ is a quantified formula $\forallf{f} \vec x. t_1$, it recursively calls $\conv$ on its body with the same polarity.
+In the case that $t$ is a function definition of a function in $\sfundefs{\Sigma}$,
+we instead construct a quantified formula over a single variable $y$ of type $\fargsort{f}$,
+and replace all occurrences of $\vec x$ in $r_1$ with $\vecfarg{f}( y )$.
+In this case, since function definitions are a top-level conjunct, by case analysis on the return values of $\conv$, we know that $D_1$ is empty.
 In the case that $t$ is not a function definition, 
@@
 and a quantified prefix is appended to the formula $D_1$ returned by the recursive call.
-In the case that $t$ is a function definition,
-we instead construct a quantified formula over a single variable $y$ of type $\fargsort{f}$,
-and replace all occurrences of $\vec x$ in $r_1$ with $\vecfarg{f}( y )$.
-In this case, since function definitions are a top-level conjunct, by case analysis on the return values of $\conv$, we know that $D_1$ is empty.
+
+Otherwise, $t$ is either an application of an uninterpreted predicate, or a term of a type other than $\Bool$.
+In this case, the second component of our return value contains a constraint of the form $\exists z\typ{\fargsort{f}}. ( \vecfarg{f}( z ) \teq \vec t )$
+for each subterm $f( \vec t )$ of $t$ such that $f$ is a function in $\sfundefs{\Sigma}$.
+When such constraints are asserted positively, this indicates that the tuple of uninterpreted functions $\vecfarg{f}$ must include $\vec t$ in its range.
 
@@
 \label{ex:translation}
-Consider the (combined) signature $\Sigma = \Sigmalia \cup \Sigma_u$,
-where $\ssorts{\Sigma_u} = \{ \sortint \}$ and 
-$\sfuns{\Sigma_u} = \{ s\typ{ \sortint \rightarrow \sortint }, c\typ{ \sortint } \}$.
+Consider the (combined) signature of linear arithmetic and uninterpreted functions $\Sigma = \Sigmalia \cup \Sigma_{\euf}$,
+where $\ssorts{\Sigma_{\euf}} = \{ \sortint \}$ and 
+$\sfuns{\Sigma_{\euf}} = \{ s\typ{ \sortint \rightarrow \sortint }, c\typ{ \sortint } \}$.
 Consider the $\Sigma$-formula $\varphi$:
@@
 It is important to note that the translation $\conv$ does not preserve the models of $\varphi$.
-One model $\M$ for formula~(\ref{eq:ex-after}) in the previous example interprets 
+One model $\I$ for formula~(\ref{eq:ex-after}) in the previous example interprets 
 $\fargsort{s}$ as a finite set of elements $\{ u_0, \ldots, u_{14} \}$,
@@
 and $105$ otherwise.
-Notice that $\M$ is \emph{not} a model for the original formula~(\ref{eq:ex-before}),
-since $\M$ interprets e.g. $s( -1 )$ and $s( 15 )$ as $105$.
+Notice that $\I$ is \emph{not} a model for the original formula~(\ref{eq:ex-before}),
+since $\I$ interprets e.g. $s( -1 )$ and $s( 15 )$ as $105$.
 However, under the assumption that function definitions in $\sfundefs{\Sigma}$ are well-specified, 
@@
 Moreover, information in models for $\conv( \varphi )$ contains relevant information regarding the models of $\varphi$.
-For instance, the model $\M$ for formula~(\ref{eq:ex-after}) described above interprets $c$ as $14$, 
+For instance, the model $\I$ for formula~(\ref{eq:ex-after}) described above interprets $c$ as $14$, 
 while there exists a model of formula~(\ref{eq:ex-before}) also interprets $c$ as $14$.
-In fact, we claim that there exists a model of $\varphi$ that coincides with a model of $\conv( \varphi )$ on its interpretation of all functions in $\sfuns{\Sigma}$ 
+In fact, we claim that there exists a model of $\varphi$ that coincides with each model of $\conv( \varphi )$ on its interpretation of all functions in $\sfuns{\Sigma}$ 
 other than those in $\sfundefs{\Sigma}$, i.e. $\{ s \}$ in the previous example.
@@
 
+%\begin{lemma}
+%Let $\M$ be a model satisfying $\conv( \varphi )$, and
+%%and let $\conv( \varphi )^\ast$ be the quantifier-free formula obtained by 
+%let $M$ be a set of literals such that $\M \models M$ and $M \models_p \conv( \varphi )$.
+%For each assignment $\sigma$ over $\vars{ M }$,
+%each $f( \vec t ) \in \terms( M \sigma )$ where $f \in \sfundefs{\Sigma}$, 
+%there exists a formula of the form $\exists z\typ{\fargsort{f}}. ( \vecfarg{f}( z ) \teq \vec t )$ in $M \sigma$.
+%\end{lemma}
+
 \begin{theorem}
@@
 \end{theorem}
+\begin{proof}
+(Sketch)  First, we show that if $\varphi$ has a model $\I$, then $\conv( \varphi )$ has a model $\J$.
+Let $\J$ be the $\extendsig{\Sigma}$-interpretation that interprets all sorts $\sigma \in \ssorts{\Sigma}$ as $\sigma^\I$,
+all functions $f \in \sfuns{\Sigma}$ as $f^\I$,
+and for each function $f$ of type $\tau_1 \times \ldots \tau_n \rightarrow \tau$ in $\sfundefs{\Sigma}$, 
+interprets $\fargsort{f}$ as a tuple of values of type $\tau_1^\I \times \ldots \times \tau_n^\I$,
+and each $\farg{f}^i$ as the $i^{th}$ projection on such tuples for each $i = 1, \ldots, n$.
+Note that $\J$ satisfies \emph{every} constraint of the form $\exists z\typ{\fargsort{f}}. ( \vecfarg{f}( z ) \teq \vec t )$,
+since by our construction of $\J$ there exists a $z$ such that $z^\J = \vec t^\J$.
+Since $\I$ satisfies $\varphi$ and due to our construction of $\conv( \varphi )$, we conclude $\J$ is a model of $\conv( \varphi )$.
+
+Second, we show that if $\conv( \varphi )$ has a model $\I$, then $\varphi$ has a model $\J$.
+
+\end{proof}
 
@@
 then it will replace its list of known assertions based on the conversion $\conv$ before checking for satisfiability.
-We stress that this option is model-sound only if the functions provided in the input are well-specified.
+Note that this option is model-sound only if the functions provided in the input are well-specified.
 

### 2015-04-30 tex/main.tex
Clarify proof a bit more.

@@
 since by our construction of $\J$ there exists a $z$ such that $z^\J = \vec t^\J$.
-Since $\I$ satisfies $\varphi$ and due to our construction of $\conv( \varphi )$, we conclude $\J$ is a model of $\conv( \varphi )$.
+By case analysis it follows that $\conv_0$ maintains the invariant that
+if $\conv_0( t\typ{\Bool}, \ptrue )$ returns $( r\typ{\Bool}, D )$,
+then $\I$ satisfies $t$ if and only if $\J$ satisfies $r$.
+Thus, we conclude $\J$ is a model of $\conv( \varphi )$.
 

### 2015-05-01 tex/main.tex
More work on prelim

@@
 
-%\ajr{(1) Encoding, (2) Evaluation, (3) Front-end support for define-funs-rec in \cvc}
-
 In this paper, we make the following contributions:
@@
 $\ltrue$ and $\lfalse$ of that type (respectively, for true and false), 
-an (infix) equality predicate $\teq$ of type $\sigma \times \sigma$ for each $\sigma \in \ssorts{\Sigma}$,
+an (infix) equality predicate $\teq$ of type $\tau \times \tau$ for each $\tau \in \ssorts{\Sigma}$,
 and standard boolean connectives ($\neg$, $\wedge$, $\vee$, $\lite$, etc.).
-and a function $\lite$ of type $\Bool \times \sigma \times \sigma \rightarrow \lite$ for each $\sigma \in \ssorts{\Sigma}$.
+and a function $\lite$ of type $\Bool \times \tau \times \tau \rightarrow \tau$ for each $\tau \in \ssorts{\Sigma}$.
 We will write $t\typ{\tau}$ to denote a term $t$ of type~$\tau$.
@@
 
+A \emph{$\Sigma$-interpretation $\I$} %is a mathematical structure that maps
+each $\tau \in \ssorts{\Sigma}$ to a non-empty set $\tau^\I$,
+the \emph{domain} of $\tau$ in $\I$;
+each $f$ of type $\tau_1 \cdots \tau_n \rightarrow \tau$ in $\sfuns{\Sigma}$ 
+to a total function $f^\I$ of type $\tau_1^\I \times \cdots \times \tau_n^\I \rightarrow \tau^\I$.
 A \emph{theory} is a pair $T = (\Sigma, \Mo)$ where 
@@
 where $f$ is an uninterpreted function from $\sfuns{\Sigma}$, or the distinguished constant $\fnull$.
-When $f$ is not $\fnull$, we will refer to $\forallf{f} \vec x. \varphi$ as a \emph{function definition} (for $f$).
+An annotated quantified formula $\forallf{f} \vec x. \varphi$ is a \emph{function definition (for $f$)} if 
+$f$ is a non-$\fnull$ function of type $\tau_1 \times \ldots \times \tau_n \rightarrow \tau$, 
+$\vec x$ is a tuple of $n$ variables of type $\tau_1, \ldots, \tau_n$
+and $\forallf{f} \vec x. \varphi$ is eqvuivalent to $\forall \vec x. f( \vec x ) \teq t$ for some $t$.
 We will write $\forall \vec x. \varphi$ as shorthand for $\forallf{\fnull} \vec x. \varphi$,
@@
 \begin{definition}
-A formula $\varphi$ \emph{well-defined with respect to functions} $S$ if it is of the form:
+A formula $\varphi$ is \emph{well-defined with respect to functions} $\{ f_1, \ldots, f_n \}$ if it is of the form:
 \begin{eqnarray} \label{eq:wdf}
@@
 \end{eqnarray}
-where $f_1, \ldots, f_n$ are distinct, 
-$S = \{ f_1, \ldots, f_n \}$,
+where $f_1, \ldots, f_n$ are not $\fnull$ and distinct, 
 and $\psi$ contains no function definitions.
+We will commonly call $\psi$ the \emph{conjecture} of $\varphi$.
 \end{definition}
@@
 $\{ \forallf{f} x^\sortint. f( x ) \teq f( x - 1 ) \}$ or even $\{ \forallf{f} x. f( x ) \teq f( x ) \}$ are well-specified.
-\ajr{possibly revise}
+\ajr{revise}
 

### 2015-05-03 tex/main.tex
Minor fixes and comments

@@
 \newcommand{\ajr}[1]{\rem{#1 --ajr}}
+\newcommand{\jb}[1]{\rem{#1 --jb}}
+\newcommand{\ct}[1]{\rem{#1 --ct}}
 
@@
 %%% I added (Co) -- hope you like it --JB
-\title{Model Finding for (Co)recursive Functions in SMT}
+\title{Model Finding for (Co)recursive Functions in SMT\thanks{
+This research is partially supported by the Inria technological development
+action \relax{Contrexemples Utilisables par Isabelle et Coq} (CUIC).
+}
+}
 
@@
 
-first-order model finders: SEM, MACE/Mace, Paradox, iProver
+first-order model finders: SEM, MACE/Mace, Paradox, FMDarwin, iProver
   either direct search or reduction to SAT
@@
 an (infix) equality predicate $\teq$ of type $\tau \times \tau$ for each $\tau \in \ssorts{\Sigma}$,
-and standard boolean connectives ($\neg$, $\wedge$, $\vee$, $\lite$, etc.).
+and standard boolean connectives ($\neg$, $\wedge$, $\vee$, etc.).
 and a function $\lite$ of type $\Bool \times \tau \times \tau \rightarrow \tau$ for each $\tau \in \ssorts{\Sigma}$.
@@
 $\vec x$ is a tuple of $n$ variables of type $\tau_1, \ldots, \tau_n$
-and $\forallf{f} \vec x. \varphi$ is eqvuivalent to $\forall \vec x. f( \vec x ) \teq t$ for some $t$.
+and $\forallf{f} \vec x. \varphi$ is equivalent to $\forall \vec x. f( \vec x ) \teq t$ for some $t$.
 We will write $\forall \vec x. \varphi$ as shorthand for $\forallf{\fnull} \vec x. \varphi$,
@@
 We will commonly call $\psi$ the \emph{conjecture} of $\varphi$.
+\ct{Don't we need additional restrictions on the $\varphi_i$'s here?
+(E.g., they are quantifier-free, or perhaps existential or containing only let binders.)}
 \end{definition}
@@
 have focused on finding models for quantified formulas in SMT.
-A instatiation-based procedure is shown in~\cite{GeDeM-CAV-09} to be complete for formulas in the \emph{essentially uninterpreted} fragment.
+A instantiation-based procedure is shown in~\cite{GeDeM-CAV-09} to be complete for formulas in the \emph{essentially uninterpreted} fragment.
 This fragment is limited to universally quantified formulas where all occurrences of variables occur as direct subchildren of uninterpreted functions.
@@
 Here, we advocate an approach that translates formulas involving function definitions (such as the one mentioned in the previous paragraph)
-into equisatisfiable formulas where the techniques~\cite{GeDeM-CAV-09, ReyEtAl-1-RR-13} are applicable.
+into equisatisfiable \ct{to be defined} formulas where the techniques~\cite{GeDeM-CAV-09, ReyEtAl-1-RR-13} are applicable.
 
@@
 and a $\extendsig{ \Sigma }$-formula $\psi$ such that $\varphi$ and $\psi$ are equisatisfiable.
-The intuition behind this translation is use an uninterpreted type to abstract the set of \emph{relevant} tuples for each recursive function $f$,
+The intuition behind this translation is to use an uninterpreted type \ct{to be defined} to abstract the set of \emph{relevant} tuples for each recursive function $f$,
 and restrict the quantification of the function definition for $f$ to a variable of this type.
@@
 Function definitions $\forallf{f} \vec x. \varphi$ can be written in using the $\definefunreccmd$ command, 
-proposed in the draft of SMT version 2.5~\cite{}.
-For instance, the input from Example~\ref{ex:translation} can be written as:
+from Cersion 2.5~\cite{} of the SMT-LIB standard.
+For instance, the formula $\varphi$ in Example~\ref{ex:translation} 
+can be written as:
 
+{\small
 \begin{verbatim}
@@
 \end{verbatim}
+}
 
@@
 which after rewriting becomes $\forallf{s} x. \lite( x \leq 0, s( x ) \teq 0, s( x ) \teq s( x-1 ) )$.
-If \cvc is in finite model finding mode for recursive functions is enabled (using the command line parameter ``--fmf-fun"),
+If \cvc's finite model finding mode for recursive functions is enabled (using the command line parameter ``--fmf-fun"),
 then it will replace its list of known assertions based on the conversion $\conv$ before checking for satisfiability.
-Note that this option is model-sound only if the functions provided in the input are well-specified.
+Note that this option is model-sound \ct{define?} only if the functions provided in the input are well-specified.
 
@@
 We would like to thank Damien Busato-Gaston and Emmanouil Koukoutos for providing the initial set of benchmarks used in the evaluation.
-This research is partially supported by the Inria technological development
-action \relax{Contrexemples Utilisables par Isabelle et Coq} (CUIC).
-
 }

### 2015-05-05 tex/main.tex
Minor

@@
 (Sketch)  First, we show that if $\varphi$ has a model $\I$, then $\conv( \varphi )$ has a model $\J$.
-Let $\J$ be the $\extendsig{\Sigma}$-interpretation that interprets all sorts $\sigma \in \ssorts{\Sigma}$ as $\sigma^\I$,
+Let $\J$ be the $\extendsig{\Sigma}$-interpretation that interprets all types $\tau \in \ssorts{\Sigma}$ as $\tau^\I$,
 all functions $f \in \sfuns{\Sigma}$ as $f^\I$,

### 2015-05-05 tex/main.tex
mild edits -- revert anything you don't like

@@
 
+\let\oldSigma=\Sigma
+\def\Sigma{\mathrm{\oldSigma}}
+
 \newcommand{\cvc}{\textsc{cvc}{\small 4}\xspace}
@@
 \newcommand{\cc}[1]{#1^*}
-\newcommand{\terms}{\mathbf{T}}
+
+%%% I'm trying to avoid bold whenever it can be avoided (except for headings
+%%% etc.) --JB
+%\newcommand{\terms}{\mathbf{T}}
+\newcommand{\terms}{\mathcal{T}}
+
 %\newcommand{\functerms}{\mathbf{DFT}}
@@
 \newcommand{\M}{\mathcal{M}}
-\newcommand{\I}{\mathcal{I}}
+\newcommand{\I}{\mathcal{J}} %%% there's no difference, right? --JB
 \newcommand{\J}{\mathcal{J}}
@@
 
-\newcommand{\ssorts}[1]{#1^\mathrm{s}}
+%\newcommand{\ssorts}[1]{#1^\mathrm{s}}
+\newcommand{\ssorts}[1]{#1^\mathrm{y}}
+
 \newcommand{\sfuns}[1]{#1^\mathrm{f}}
@@
 %%% I added (Co) -- hope you like it --JB
-\title{Model Finding for (Co)recursive Functions in SMT\thanks{
+\title{Model Finding for (Co)recursive Functions in SMT\thanks{%
 This research is partially supported by the Inria technological development
-action \relax{Contrexemples Utilisables par Isabelle et Coq} (CUIC).
+action ``Contre-exemples Utilisables par Isabelle et Coq'' (CUIC).
 }
@@
 A \emph{signature} $\Sigma$ consists of 
-a set $\ssorts{\Sigma}$ of type symbols and
+a set $\ssorts{\Sigma}$ of type symbols (sorts) and
 a set $\sfuns{\Sigma}$ of function symbols.
 We assume that signatures always include a Boolean type $\Bool$ and constants 
-$\ltrue$ and $\lfalse$ of that type (respectively, for true and false), 
-an (infix) equality predicate $\teq$ of type $\tau \times \tau$ for each $\tau \in \ssorts{\Sigma}$,
-and standard boolean connectives ($\neg$, $\wedge$, $\vee$, etc.).
+$\ltrue$ (truth) and $\lfalse$ (falsity) of that type,
+an infix equality predicate $\teq$ of type $\tau \times \tau \to \Bool$
+for each $\tau \in \ssorts{\Sigma}$,
+standard boolean connectives ($\neg$, $\wedge$, $\vee$, etc.),
 and a function $\lite$ of type $\Bool \times \tau \times \tau \rightarrow \tau$ for each $\tau \in \ssorts{\Sigma}$.
-We will write $t\typ{\tau}$ to denote a term $t$ of type~$\tau$.
+We write $t\typ{\tau}$ to denote a term $t$ of type~$\tau$.
 We write $\terms( t )$ to denote the set of subterms in $t$.

### 2015-05-05 tex/main.tex
more

@@
 \documentclass[runningheads,a4paper]{llncs}
-\usepackage{times}
+\usepackage{mathptmx}
 \usepackage{listings}
@@
 
+\DeclareFontFamily{OT1}{pzc}{}
+\DeclareFontShape{OT1}{pzc}{m}{it}{<-> s * [1.10] pzcmi7t}{}
+\DeclareMathAlphabet{\mathcalx}{OT1}{pzc}{m}{it}
+
 \newcommand{\con}[1]{\mathsf{#1}}
@@
 %\newcommand{\terms}{\mathbf{T}}
-\newcommand{\terms}{\mathcal{T}}
+\newcommand{\terms}{\mathcalx{T}}
 
@@
 \newcommand{\vars}{\mathbf{V}}
-\newcommand{\M}{\mathcal{M}}
-\newcommand{\I}{\mathcal{J}} %%% there's no difference, right? --JB
-\newcommand{\J}{\mathcal{J}}
+\newcommand{\M}{\mathcalx{M}}
+\newcommand{\I}{\mathcalx{J}} %%% there's no difference, right? --JB
+%% (I find \mathcalx{I} ugly, with its two horizontal bars, hence the J)
+\newcommand{\J}{\mathcalx{J}}
 \def\AIF{\qtab\keyword{if}\ }
@@
 \def\ENDPROC{\untab}
-\newcommand{\conv}{\mathcal{W}}
+\newcommand{\conv}{\mathcalx{W}}
 
@@
 \newcommand{\Sigmalia}{\Sigma_{LIA}}
-\newcommand{\extendsig}[1]{\mathcal{E}( #1 )}
+\newcommand{\extendsig}[1]{\mathcalx{E}( #1 )}
 
@@
 
-\newcommand{\Mo}{\mathbf{I}}
+\newcommand{\Mo}{{\mathcal{\!J\!}}}
 
@@
 A \emph{signature} $\Sigma$ consists of 
-a set $\ssorts{\Sigma}$ of type symbols (sorts) and
+a set $\ssorts{\Sigma}$ of type symbols (or sorts) and
 a set $\sfuns{\Sigma}$ of function symbols.
@@
 and a function $\lite$ of type $\Bool \times \tau \times \tau \rightarrow \tau$ for each $\tau \in \ssorts{\Sigma}$.
-We write $t\typ{\tau}$ to denote a term $t$ of type~$\tau$.
-We write $\terms( t )$ to denote the set of subterms in $t$.
+We write $t\typ{\tau}$ to denote a term $t$ of type~$\tau$ and
+$\terms( t )$ to denote the set of subterms in $t$.
 %We write $\functerms^\Sigma( t )$ to denote the set of $f$-applications in $t$ such that $f \in \sfundefs{\Sigma}$.
 
-A \emph{$\Sigma$-interpretation $\I$} %is a mathematical structure that maps
-each $\tau \in \ssorts{\Sigma}$ to a non-empty set $\tau^\I$,
-the \emph{domain} of $\tau$ in $\I$;
-each $f$ of type $\tau_1 \cdots \tau_n \rightarrow \tau$ in $\sfuns{\Sigma}$ 
+A \emph{$\Sigma$-interpretation $\I$} %is a mathematical structure that
+maps each $\tau \in \ssorts{\Sigma}$ to a nonempty set $\tau^\I$,
+the \emph{domain} of~$\tau$ in $\I$,
+and each $\con{f}$ of type $\tau_1 \times \cdots \times \tau_n \rightarrow \tau$ in
+$\sfuns{\Sigma}$ 
 to a total function $f^\I$ of type $\tau_1^\I \times \cdots \times \tau_n^\I \rightarrow \tau^\I$.
@@
 The intuition is that the interpretation of a term such as $s( -1 )$ in the previous example is not relevant to the satisfiability of formula~(\ref{eq:ex-before}).
-Thus, a satisfiable (resp. unsatisfiable) response from an SMT solver on input $\conv( \varphi )$ implies the existence (resp. non-existence) of a model for $\varphi$.
+Thus, a satisfiable (resp. unsatisfiable) response from an SMT solver on input $\conv( \varphi )$ implies the existence (resp.\ nonexistence) of a model for $\varphi$.
 Moreover, information in models for $\conv( \varphi )$ contains relevant information regarding the models of $\varphi$.
@@
 other than those in $\sfundefs{\Sigma}$, i.e. $\{ s \}$ in the previous example.
-From a practical point of view, this is in fact a non-issue, as $s$ is the very function that was explicitly defined by the user,
+From a practical point of view, this is not an issue, because
+$s$ is the very function that was explicitly defined by the user,
 and hence already has an intended interpretation.

### 2015-05-05 tex/main.tex
tuned typesetting

@@
-%%%%%%%%%%%%%%%%%%%%%%% file typeinst.tex %%%%%%%%%%%%%%%%%%%%%%%%%
-%
-% This is the LaTeX source for the instructions to authors using
-% the LaTeX document class 'llncs.cls' for contributions to
-% the Lecture Notes in Computer Sciences series.
-% http://www.springer.com/lncs       Springer Heidelberg 2006/05/04
-%
-% It may be used as a template for your own input - copy it
-% to a new file with a new name and use it as the basis
-% for your article.
-%
-% NB: the document class 'llncs' has its own and detailed documentation, see
-% ftp://ftp.springer.de/data/pubftp/pub/tex/latex/llncs/latex2e/llncsdoc.pdf
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
 \documentclass[runningheads,a4paper]{llncs}
+\usepackage[T1]{fontenc}
+\usepackage[scaled=.82]{beramono}
+\usepackage[scaled=.86]{helvet}
 \usepackage{mathptmx}
@@
 
+\renewcommand\vec[1]{\overline{#1}}
+
 \let\oldSigma=\Sigma
 \def\Sigma{\mathrm{\oldSigma}}
+
+\let\oldneg=\neg
+\def\neg{\oldneg\:}
 
@@
 \newcommand{\boolop}{\oplus}
-\newcommand{\forallf}[1]{\forall_{#1}}
+\newcommand{\forallf}[1]{\forall_{\!#1\:}}
 \newcommand{\fnull}{\emptyset}
@@
 
-\newcommand{\euf}{\ensuremath{\mathrm{E}}\xspace}
-\newcommand{\ari}{\ensuremath{\mathrm{A}}\xspace}
+%\newcommand{\euf}{\ensuremath{\mathrm{E}}\xspace}
+\newcommand{\euf}{\ensuremath{\mathcalx{E}}\xspace}
+%\newcommand{\ari}{\ensuremath{\mathrm{A}}\xspace}
+\newcommand{\ari}{\ensuremath{\mathcalx{A}}\xspace}
 
@@
 $\ltrue$ (truth) and $\lfalse$ (falsity) of that type,
-an infix equality predicate $\teq$ of type $\tau \times \tau \to \Bool$
+an infix equality predicate ${\teq} : \tau \times \tau \to \Bool$
 for each $\tau \in \ssorts{\Sigma}$,
 standard boolean connectives ($\neg$, $\wedge$, $\vee$, etc.),
-and a function $\lite$ of type $\Bool \times \tau \times \tau \rightarrow \tau$ for each $\tau \in \ssorts{\Sigma}$.
+and a function $\lite : \Bool \times \tau \times \tau \rightarrow \tau$ for each $\tau \in \ssorts{\Sigma}$.
 We write $t\typ{\tau}$ to denote a term $t$ of type~$\tau$ and
@@
 the \emph{domain} of~$\tau$ in $\I$,
-and each $\con{f}$ of type $\tau_1 \times \cdots \times \tau_n \rightarrow \tau$ in
+and each $\con{f} : \tau_1 \times \cdots \times \tau_n \rightarrow \tau$ in
 $\sfuns{\Sigma}$ 
-to a total function $f^\I$ of type $\tau_1^\I \times \cdots \times \tau_n^\I \rightarrow \tau^\I$.
+to a total function $\con{f}^\I : \tau_1^\I \times \cdots \times \tau_n^\I \rightarrow \tau^\I$.
 A \emph{theory} is a pair $T = (\Sigma, \Mo)$ where 
@@
 A $\Sigma$-formula $\varphi$ is 
-\emph{$T$-satisfiable} (resp., \emph{$T$-unsatisfiable}) 
-if it is satisfied by some (resp., no) interpretation in $\Mo$.
+\emph{$T$-satisfiable}
+if it is satisfied by some interpretation in $\Mo$;
+otherwise, it is \emph{$T$-unsatisfiable}.
 Formulas $\varphi$ and $\psi$ are \emph{equisatisfiable} if
 $\varphi$ is $T$-satisfiable if and only if $\psi$ is $T$-satisfiable.
-For a given signature $\Sigma$
+Given a signature $\Sigma$,
 the \emph{theory of equality with uninterpreted functions} \euf
 consists of the set of all $\Sigma$-interpretations.
-We refer to the sort and function symbols in this theory as \emph{uninterpreted}.
+We refer to the type and function symbols in this theory as \emph{uninterpreted}.
 
 %In this paper, we associate quantified formulas with uninterpreted functions (informally, those whose semantics they define).
-In this paper, we consider \emph{annotated quantifed formulas} of the form $\forallf{f} \vec x. \varphi$, 
-where $f$ is an uninterpreted function from $\sfuns{\Sigma}$, or the distinguished constant $\fnull$.
-An annotated quantified formula $\forallf{f} \vec x. \varphi$ is a \emph{function definition (for $f$)} if 
-$f$ is a non-$\fnull$ function of type $\tau_1 \times \ldots \times \tau_n \rightarrow \tau$, 
-$\vec x$ is a tuple of $n$ variables of type $\tau_1, \ldots, \tau_n$
-and $\forallf{f} \vec x. \varphi$ is equivalent to $\forall \vec x. f( \vec x ) \teq t$ for some $t$.
-We will write $\forall \vec x. \varphi$ as shorthand for $\forallf{\fnull} \vec x. \varphi$,
-and $\exists \vec x. \varphi$ as shorthand for $\neg \forall \vec x. \neg \varphi$.
+In this paper, we consider \emph{annotated quantifed formulas} of the form
+$\forallf{a} \vec x.\; \varphi$, where the annotation~$a$ is either
+an uninterpreted function $\con{f} \in \sfuns{\Sigma}$ or the special symbol
+$\fnull$.
+Given $\con{f} : \tau_1 \times \ldots \times \tau_n \rightarrow \tau$,
+an annotated
+quantified formula $\forallf{\con{f}} \vec x.\; \varphi$ is a \emph{function definition}
+(\,\emph{for $\con{f}$}\vthinspace) if $\vec x$ is a tuple of variables of type
+$\tau_1, \ldots, \tau_n$ and $\forallf{\con{f}} \vec x.\; \varphi$ is equivalent to
+$\forall \vec x.\;\, \con{f}( \vec x ) \teq t$ for some $t$. We will write $\forall \vec
+x. \varphi$ as shorthand for $\forallf{\fnull} \vec x.\; \varphi$, and $\exists
+\vec x. \varphi$ as shorthand for $\neg \forall \vec x.\; \neg \varphi$.
 
@@
 \begin{eqnarray} \label{eq:wdf}
-(\forallf{f_1} \vec x_1. \varphi_1) \wedge \ldots \wedge (\forallf{f_n} \vec x_n. \varphi_n) \wedge \psi
+(\forallf{f_1} \vec x_1.\; \varphi_1) \wedge \ldots \wedge (\forallf{f_n} \vec x_n.\; \varphi_n) \wedge \psi
 \end{eqnarray}
@@
 is \emph{well-specified (in $T$)} if $\varphi_1 \wedge \ldots \wedge \varphi_n$ is $T$-satisfiable.
-For instance, $\{ \forallf{f} \vec x. f( x ) \teq 0 \}$ is well-specified,
-whereas $\{ \forallf{f} x. f( x ) \teq f( x ) + 1 \}$ is not. 
+For instance, $\{ \forallf{\con{f}} \vec x.\;\, \con{f}( x ) \teq 0 \}$ is well-specified,
+whereas $\{ \forallf{\con{f}} x.\;\, \con{f}( x ) \teq \con{f}( x ) + 1 \}$ is not. 
 Note that a well-specified function definition need not specify a terminating function, e.g.
-$\{ \forallf{f} x^\sortint. f( x ) \teq f( x - 1 ) \}$ or even $\{ \forallf{f} x. f( x ) \teq f( x ) \}$ are well-specified.
+$\{ \forallf{\con{f}} x^\sortint.\;\, \con{f}( x ) \teq \con{f}( x - 1 ) \}$ or even $\{ \forallf{\con{f}} x.\;\, \con{f}( x ) \teq \con{f}( x ) \}$ are well-specified.
 \ajr{revise}
@@
 This fragment is limited to universally quantified formulas where all occurrences of variables occur as direct subchildren of uninterpreted functions.
-For instance, the formula $\forall x\typ{\sortint}. f( x ) \teq g( x ) + 5$ 
-is in the essentially uninterpreted fragment since all occurrences of $x$ occur beneath uninterpreted functions $f$ and $g$.
+For instance, the formula $\forall x\typ{\sortint}.\;\, \con{f}( x ) \teq g( x ) + 5$ 
+is in the essentially uninterpreted fragment since all occurrences of $x$ occur beneath uninterpreted functions $\con{f}$ and $g$.
 Other syntactic criteria are identified in this paper that extend this fragment slightly, including cases when variables occur as arguments of arithmetic predicates.
@@
 finite model finding techniques can find a model for
-the formula $\forall x\typ{\tau} y\typ{\tau}. \neg x \teq y \vee f( x ) \teq f( y )$
-where the uninterpreted type $\tau$ is interpreted as a finite set, and $f$ as injective finite map.
+the formula $\forall x\typ{\tau} y\typ{\tau}.\;\, \neg x \teq y \vee \con{f}( x ) \teq \con{f}( y )$
+where the uninterpreted type $\tau$ is interpreted as a finite set, and $\con{f}$ as injective finite map.
 
@@
 while the fragment in~\cite{ReyEtAl-1-RR-13} is not applicable for functions over infinite domains such as the integers.
-So, for instance, the quantified formula $\forall x\typ{\sortint}. \lite( p \leq 0, p( x ) \teq 0, p( x ) \teq 2 * p( x - 1 ) )$ 
+So, for instance, the quantified formula $\forall x\typ{\sortint}.\;\, \lite( p \leq 0, p( x ) \teq 0, p( x ) \teq 2 * p( x - 1 ) )$ 
 is not in the essentially uninterpreted fragment (since $x$ occurs as a child of $-$),
@@
 and a $\extendsig{ \Sigma }$-formula $\psi$ such that $\varphi$ and $\psi$ are equisatisfiable.
-The intuition behind this translation is to use an uninterpreted type \ct{to be defined} to abstract the set of \emph{relevant} tuples for each recursive function $f$,
-and restrict the quantification of the function definition for $f$ to a variable of this type.
-Informally, the relevant tuples $\vec t$ of a function $f$ are the ones for which the interpretation of $f( \vec t )$ is relevant to the satisfiability of $\varphi$.
+The intuition behind this translation is to use an uninterpreted type \ct{to be defined} to abstract the set of \emph{relevant} tuples for each recursive function $\con{f}$,
+and restrict the quantification of the function definition for $\con{f}$ to a variable of this type.
+Informally, the relevant tuples $\vec t$ of a function $\con{f}$ are the ones for which the interpretation of $\con{f}( \vec t )$ is relevant to the satisfiability of $\varphi$.
 
@@
     \end{itemize}
-  \item[] $\mathsf{else}$ $\mathsf{if}$ $t \equiv \forallf{f} \vec x. t_1$
+  \item[] $\mathsf{else}$ $\mathsf{if}$ $t \equiv \forallf{\con{f}} \vec x.\;\, t_1$
     \begin{itemize}
       \item[] $\mathsf{let} ( r_1, D_1 ) = \conv_0( t_1, p )$;
-      \item[] $\mathsf{if}$ $f \in \sfundefs{\Sigma}$
+      \item[] $\mathsf{if}$ $\con{f} \in \sfundefs{\Sigma}$
       \begin{itemize}
         %\item[] $\mathsf{assert}( D_1 = \emptyset )$;
-        \item[] $\mathsf{return}$ $( \forall y\typ{\fargsort{f}}. ( r_1 [ \vecfarg{f}( y ) / \vec x ] ), \emptyset )$
+        \item[] $\mathsf{return}$ $( \forall y\typ{\fargsort{\con{f}}}.\;\, ( r_1 [ \vecfarg{\con{f}}( y ) / \vec x ] ), \emptyset )$
       \end{itemize}
@@
       \begin{itemize}
-        \item[] $\mathsf{return}$ $( \forall \vec x. r_1, \forall \vec x. D_1 )$
+        \item[] $\mathsf{return}$ $( \forall \vec x.\; r_1, \forall \vec x.\; D_1 )$
       \end{itemize}
@@
    \begin{itemize}
-     \item[] $\mathsf{return}$ $( t, \{ \exists z\typ{\fargsort{f}}. ( \vecfarg{f}( z ) \teq \vec t ) \mid f( \vec t ) \in \terms( t ), f \in \sfundefs{\Sigma} \} )$
+     \item[] $\mathsf{return}$ $( t, \{ \exists z\typ{\fargsort{\con{f}}}. ( \vecfarg{f}( z ) \teq \vec t ) \mid f( \vec t ) \in \terms( t ), f \in \sfundefs{\Sigma} \} )$
    \end{itemize}
@@
 We extend the signature of $\Sigma$ to a signature $\extendsig{\Sigma}$ containing the following additional definitions.
-For each $f \in \sfundefs{\Sigma}$ of type $\tau_1 \times \ldots \times \tau_n \rightarrow \tau$, 
+For each $\con{f} \in \sfundefs{\Sigma}$ of type $\tau_1 \times \ldots \times \tau_n \rightarrow \tau$, 
 the signature $\extendsig{\Sigma}$ includes:
 \begin{enumerate}
-\item[-] an uninterpreted type $\fargsort{f}$, and
-\item[-] a vector of $n$ uninterpreted functions $\farg{f}^1$, $\ldots$, $\farg{f}^n$ of type $(\fargsort{f} \rightarrow \tau_1)$, $\ldots$, $(\fargsort{f} \rightarrow \tau_n)$.
+\item[-] an uninterpreted type $\fargsort{\con{f}}$, and
+\item[-] a vector of $n$ uninterpreted functions $\farg{\con{f}}^1$, $\ldots$, $\farg{\con{f}}^n$ of type $(\fargsort{\con{f}} \rightarrow \tau_1)$, $\ldots$, $(\fargsort{\con{f}} \rightarrow \tau_n)$.
 \end{enumerate}
-%The interpretation of uninterpreted sort $\fargsort{f}$ will denote the elements (tuples) on which the function $f$ is applied.
-%The role of the uninterpreted functions $\farg{f}^1$, $\ldots$, $\farg{f}^n$ will be discussed more in the following.
+%The interpretation of uninterpreted sort $\fargsort{\con{f}}$ will denote the elements (tuples) on which the function $\con{f}$ is applied.
+%The role of the uninterpreted functions $\farg{\con{f}}^1$, $\ldots$, $\farg{\con{f}}^n$ will be discussed more in the following.
 
@@
 It returns a pair of the form $( r, D )$, where $r$ is a term of type $\tau$, and $D$ is a set of formulas.
-%The role of $D$ is to ensure that the (restricted) function definition for $f$ includes certain tuples in its domain, which we explain more in the following.
+%The role of $D$ is to ensure that the (restricted) function definition for $\con{f}$ includes certain tuples in its domain, which we explain more in the following.
 At a high level, the translation modifies $\varphi$ in two ways.
-First, it restricts the quantification on function definitions for $f$ to the corresponding uninterpreted type $\fargsort{f}$.
-Second, it augments $\varphi$ with additional constraints of the form $\exists z\typ{\fargsort{f}}. ( \vecfarg{f}( z ) \teq \vec t )$
-which ensure that this restriction for the defintion $f$ covers all relevant tuples of terms,
-namely the ones for which an $f$-application exists and is relevant to the satisfiability of the current branch of $\varphi$.
+First, it restricts the quantification on function definitions for $\con{f}$ to the corresponding uninterpreted type $\fargsort{\con{f}}$.
+Second, it augments $\varphi$ with additional constraints of the form $\exists z\typ{\fargsort{\con{f}}}. ( \vecfarg{\con{f}}( z ) \teq \vec t )$
+which ensure that this restriction for the defintion $\con{f}$ covers all relevant tuples of terms,
+namely the ones for which an $\con{f}$-application exists and is relevant to the satisfiability of the current branch of $\varphi$.
 
-In the case that $t$ is an application of (interpreted) predicate $\boolop$ (e.g. a boolean connective, or equality),
+In the case that $t$ is an application of (interpreted) predicate $\boolop$ (e.g., a boolean connective, or equality),
 we recursively call $\conv$ on its subchildren $t_i$ and polarity $\pol( \boolop, i, p )$, where $\pol$ is defined as\footnote{The negation $\neg p$ of polarity $p$ is 
@@
 
-In the case that $t$ is a quantified formula $\forallf{f} \vec x. t_1$, it recursively calls $\conv$ on its body with the same polarity.
+In the case that $t$ is a quantified formula $\forallf{\con{f}} \vec x.\; t_1$, it recursively calls $\conv$ on its body with the same polarity.
 In the case that $t$ is a function definition of a function in $\sfundefs{\Sigma}$,
-we instead construct a quantified formula over a single variable $y$ of type $\fargsort{f}$,
-and replace all occurrences of $\vec x$ in $r_1$ with $\vecfarg{f}( y )$.
+we instead construct a quantified formula over a single variable $y$ of type $\fargsort{\con{f}}$,
+and replace all occurrences of $\vec x$ in $r_1$ with $\vecfarg{\con{f}}( y )$.
 In this case, since function definitions are a top-level conjunct, by case analysis on the return values of $\conv$, we know that $D_1$ is empty.
@@
 Otherwise, $t$ is either an application of an uninterpreted predicate, or a term of a type other than $\Bool$.
-In this case, the second component of our return value contains a constraint of the form $\exists z\typ{\fargsort{f}}. ( \vecfarg{f}( z ) \teq \vec t )$
-for each subterm $f( \vec t )$ of $t$ such that $f$ is a function in $\sfundefs{\Sigma}$.
-When such constraints are asserted positively, this indicates that the tuple of uninterpreted functions $\vecfarg{f}$ must include $\vec t$ in its range.
+In this case, the second component of our return value contains a constraint of the form $\exists z\typ{\fargsort{\con{f}}}. ( \vecfarg{\con{f}}( z ) \teq \vec t )$
+for each subterm $\con{f}( \vec t )$ of $t$ such that $\con{f}$ is a function in $\sfundefs{\Sigma}$.
+When such constraints are asserted positively, this indicates that the tuple of uninterpreted functions $\vecfarg{\con{f}}$ must include $\vec t$ in its range.
 
@@
 \begin{eqnarray} \label{eq:ex-before}
-\forall_{s} x\typ{\sortint}. ite( x \leq 0, s(x) \teq 0, s( x ) \teq x + s( x-1 ) ) \wedge s( c ) > 100
+\forall_{s} x\typ{\sortint}.\; \lite( x \leq 0, s(x) \teq 0, s( x ) \teq x + s( x-1 ) ) \wedge s( c ) > 100
 \end{eqnarray} 
@@
 \begin{split}
-\forall y\typ{\fargsort{s}}. ite( & \farg{s}( y ) \leq 0,  \\
+\forall y\typ{\fargsort{s}}.\;\, \lite( & \farg{s}( y ) \leq 0,  \\
  & s(\farg{s}( y )) \teq 0, \\
@@
 Notice that $\I$ is \emph{not} a model for the original formula~(\ref{eq:ex-before}),
-since $\I$ interprets e.g. $s( -1 )$ and $s( 15 )$ as $105$.
+since $\I$ interprets e.g.\ $s( -1 )$ and $s( 15 )$ as $105$.
 However, under the assumption that function definitions in $\sfundefs{\Sigma}$ are well-specified, 
@@
 %For each assignment $\sigma$ over $\vars{ M }$,
-%each $f( \vec t ) \in \terms( M \sigma )$ where $f \in \sfundefs{\Sigma}$, 
-%there exists a formula of the form $\exists z\typ{\fargsort{f}}. ( \vecfarg{f}( z ) \teq \vec t )$ in $M \sigma$.
+%each $\con{f}( \vec t ) \in \terms( M \sigma )$ where $\con{f} \in \sfundefs{\Sigma}$, 
+%there exists a formula of the form $\exists z\typ{\fargsort{\con{f}}}. ( \vecfarg{\con{f}}( z ) \teq \vec t )$ in $M \sigma$.
 %\end{lemma}
@@
 Let $\J$ be the $\extendsig{\Sigma}$-interpretation that interprets all types $\tau \in \ssorts{\Sigma}$ as $\tau^\I$,
-all functions $f \in \sfuns{\Sigma}$ as $f^\I$,
-and for each function $f$ of type $\tau_1 \times \ldots \tau_n \rightarrow \tau$ in $\sfundefs{\Sigma}$, 
-interprets $\fargsort{f}$ as a tuple of values of type $\tau_1^\I \times \ldots \times \tau_n^\I$,
-and each $\farg{f}^i$ as the $i^{th}$ projection on such tuples for each $i = 1, \ldots, n$.
-Note that $\J$ satisfies \emph{every} constraint of the form $\exists z\typ{\fargsort{f}}. ( \vecfarg{f}( z ) \teq \vec t )$,
+all functions $\con{f} \in \sfuns{\Sigma}$ as $\con{f}^\I$,
+and for each function $\con{f}$ of type $\tau_1 \times \ldots \tau_n \rightarrow \tau$ in $\sfundefs{\Sigma}$, 
+interprets $\fargsort{\con{f}}$ as a tuple of values of type $\tau_1^\I \times \ldots \times \tau_n^\I$,
+and each $\farg{\con{f}}^i$ as the $i^{th}$ projection on such tuples for each $i = 1, \ldots, n$.
+Note that $\J$ satisfies \emph{every} constraint of the form $\exists z\typ{\fargsort{\con{f}}}. ( \vecfarg{\con{f}}( z ) \teq \vec t )$,
 since by our construction of $\J$ there exists a $z$ such that $z^\J = \vec t^\J$.
@@
 First, we note that in no cases did a configuration answer differently across the two encodings.
-In other words, there were no cases in which \ziii or \cvc answered e.g. ``unsatisfiable" on a benchmark $\varphi$
+In other words, there were no cases in which \ziii or \cvc answered e.g.\ ``unsatisfiable" on a benchmark $\varphi$
 and ``satisfiable" on its corresponding benchmark $\conv( \varphi )$, or vice versa.
@@
 Thus, in most cases the translation $\conv$ does not significantly degrade performance for unsatisfiable benchmarks,
-and in some cases may actually aid the solver for determining unsatisfiability
-\footnote{The exception here is {\bf cvc4+i}, which makes use of type information associated with quantified formulas for inductive arguments.}.
+and in some cases may actually aid the solver for determining unsatisfiability.%
+\footnote{The exception here is {\bf cvc4+i}, which makes use of type
+information associated with quantified formulas for inductive arguments.}
 
@@
 For convienience, the translation $\conv$ has been implemented in the latest development version of \cvc~\cite{ReyEtAl-1-RR-13} (version 1.5 pre-release).
-Function definitions $\forallf{f} \vec x. \varphi$ can be written in using the $\definefunreccmd$ command, 
+Function definitions $\forallf{\con{f}} \vec x.\; \varphi$ can be written in using the $\definefunreccmd$ command, 
 from Cersion 2.5~\cite{} of the SMT-LIB standard.
@@
 When reading this input, 
-\cvc adds the (annotated) quantified formula $\forallf{s} x. s( x ) \teq \lite( x \leq 0, 0, s( x-1 ) )$ to its list of assertions,
-which after rewriting becomes $\forallf{s} x. \lite( x \leq 0, s( x ) \teq 0, s( x ) \teq s( x-1 ) )$.
+\cvc adds the (annotated) quantified formula $\forallf{s} x.\;\, s( x ) \teq \lite( x \leq 0, 0, s( x-1 ) )$ to its list of assertions,
+which after rewriting becomes $\forallf{s} x.\;\, \lite( x \leq 0, s( x ) \teq 0, s( x ) \teq s( x-1 ) )$.
 If \cvc's finite model finding mode for recursive functions is enabled (using the command line parameter ``--fmf-fun"),

### 2015-05-05 tex/main.tex
used 'admissible' terminology

@@
 $\tau_1, \ldots, \tau_n$ and $\forallf{\con{f}} \vec x.\; \varphi$ is equivalent to
-$\forall \vec x.\;\, \con{f}( \vec x ) \teq t$ for some $t$. We will write $\forall \vec
-x. \varphi$ as shorthand for $\forallf{\fnull} \vec x.\; \varphi$, and $\exists
+$\forall \vec x.\;\, \con{f}( \vec x ) \teq t$ for some $t$. We write $\forall \vec
+x.\; \varphi$ as shorthand for $\forallf{\fnull} \vec x.\; \varphi$, and $\exists
 \vec x. \varphi$ as shorthand for $\neg \forall \vec x.\; \neg \varphi$.
 
-\begin{definition}
-A formula $\varphi$ is \emph{well-defined with respect to functions} $\{ f_1, \ldots, f_n \}$ if it is of the form:
-\begin{eqnarray} \label{eq:wdf}
-(\forallf{f_1} \vec x_1.\; \varphi_1) \wedge \ldots \wedge (\forallf{f_n} \vec x_n.\; \varphi_n) \wedge \psi
-\end{eqnarray}
-where $f_1, \ldots, f_n$ are not $\fnull$ and distinct, 
+\begin{definition}%
+\rm
+A formula $\varphi$ is in \emph{definitional form with respect to} functions $\{
+\con{f}_1, \ldots, \con{f}_n \}$ if it is of the form
+%
+%\kern-\abovedisplayskip
+%\kern+\abovedisplayshortskip
+%
+\begin{equation} \label{eq:wdf}
+(\forallf{\con{f}_1} \vec x_1.\; \varphi_1) \wedge \ldots \wedge (\forallf{\con{f}_n} \vec x_n.\; \varphi_n) \wedge \psi
+\end{equation}
+%
+where $\con{f}_1, \ldots, \con{f}_n$ are not $\fnull$ and distinct, 
 and $\psi$ contains no function definitions.
-We will commonly call $\psi$ the \emph{conjecture} of $\varphi$.
+We call $\psi$ the \emph{conjecture} of $\varphi$.
 \ct{Don't we need additional restrictions on the $\varphi_i$'s here?
@@
 Given a signature $\Sigma$,
-we will commonly parition its function symbols $\sfuns{\Sigma}$ into $\sfundefs{\Sigma} \mathrel{\uplus} \sfunndefs{\Sigma}$,
+we partition its function symbols $\sfuns{\Sigma}$ into $\sfundefs{\Sigma} \mathrel{\uplus} \sfunndefs{\Sigma}$,
 where $\sfundefs{\Sigma}$ is the set of \emph{defined} function symbols,
-and consider $\Sigma$-formulas that are well-defined with respect to functions $\sfundefs{\Sigma}$.
+and consider $\Sigma$-formulas that are in definitional form with respect to functions $\sfundefs{\Sigma}$.
 
 A set of function definitions $\{ \varphi_1, \ldots, \varphi_n \}$ 
-is \emph{well-specified (in $T$)} if $\varphi_1 \wedge \ldots \wedge \varphi_n$ is $T$-satisfiable.
-For instance, $\{ \forallf{\con{f}} \vec x.\;\, \con{f}( x ) \teq 0 \}$ is well-specified,
+is \emph{admissible} (\emph{in $T$}) if $\varphi_1 \wedge \ldots \wedge \varphi_n$ is $T$-satisfiable.
+For example, $\{ \forallf{\con{f}} \vec x.\;\, \con{f}( x ) \teq 0 \}$ is admissible,
 whereas $\{ \forallf{\con{f}} x.\;\, \con{f}( x ) \teq \con{f}( x ) + 1 \}$ is not. 
-Note that a well-specified function definition need not specify a terminating function, e.g.
-$\{ \forallf{\con{f}} x^\sortint.\;\, \con{f}( x ) \teq \con{f}( x - 1 ) \}$ or even $\{ \forallf{\con{f}} x.\;\, \con{f}( x ) \teq \con{f}( x ) \}$ are well-specified.
+Note that an admissible function definition need not specify a terminating function, e.g.
+$\{ \forallf{\con{f}} x^\sortint.\;\, \con{f}( x ) \teq \con{f}( x - 1 ) \}$ or even $\{ \forallf{\con{f}} x.\;\, \con{f}( x ) \teq \con{f}( x ) \}$ are admissible.
 \ajr{revise}
@@
 This fragment is limited to universally quantified formulas where all occurrences of variables occur as direct subchildren of uninterpreted functions.
-For instance, the formula $\forall x\typ{\sortint}.\;\, \con{f}( x ) \teq g( x ) + 5$ 
+For example, the formula $\forall x\typ{\sortint}.\;\, \con{f}( x ) \teq g( x ) + 5$ 
 is in the essentially uninterpreted fragment since all occurrences of $x$ occur beneath uninterpreted functions $\con{f}$ and $g$.
@@
 fixed finite interpretation (such as fixed-width bitvectors, datatype enumerations, etc.).
-For instance, 
-finite model finding techniques can find a model for
+For example, finite model finding techniques can find a model for
 the formula $\forall x\typ{\tau} y\typ{\tau}.\;\, \neg x \teq y \vee \con{f}( x ) \teq \con{f}( y )$
@@
 In the remainder of the section, we fix a signature $\Sigma$.
-Say that $\Sigma$-formula $\varphi$ is well-defined with respect to functions $\sfundefs{\Sigma}$, 
-and the set of function definitions corresponding to $\sfundefs{\Sigma}$ are well-specified.
+Say that $\Sigma$-formula $\varphi$ is in definitional form with respect to functions $\sfundefs{\Sigma}$, 
+and the set of function definitions corresponding to $\sfundefs{\Sigma}$ are admissible.
 %We may translate $\varphi$ into a equisatisfiable formula $\psi$ for which known model-finding procedures~\cite{GeDeM-CAV-09, ReyEtAl-1-RR-13} are applicable.
@@
 \vspace{-2ex}
-\caption{A translation procedure $\conv$ for a $\Sigma$-formula that is well-defined with respect to a set of well-specified function definitions $\sfundefs{\Sigma}$.
+\caption{A translation procedure $\conv$ for a $\Sigma$-formula that is in definitional form with respect to a set of admissible function definitions $\sfundefs{\Sigma}$.
 In the subprocedure $\conv_0$, the procedure $\pol( \boolop, i, p )$ returns the polarity of the $i^{th}$ child of a $\boolop$-application having polarity $p$,
@@
 Consider the $\Sigma$-formula $\varphi$:
-\begin{eqnarray} \label{eq:ex-before}
+\begin{equation} \label{eq:ex-before}
 \forall_{s} x\typ{\sortint}.\; \lite( x \leq 0, s(x) \teq 0, s( x ) \teq x + s( x-1 ) ) \wedge s( c ) > 100
-\end{eqnarray} 
+\end{equation} 
 Here, $s$ is a function that returns the sum of positive numbers between $0$ and its argument $x$.
-The formula $\varphi$ is well-defined with respect to functions $\sfundefs{\Sigma_u}$,
+The formula $\varphi$ is in definitional form with respect to functions $\sfundefs{\Sigma_u}$,
 and states that the sum of all positive numbers between $0$ and $c$ is greater than $100$.
@@
 and $s$ as the (almost constant) function:
-\begin{eqnarray}
+\begin{equation}
 \lambda x\typ{Int}. \lite( x \teq 0, 0, \lite( x \teq 1, 1, \lite( x \teq 2, 3, \ldots, \lite( x \teq 13, 91, 105 ) \ldots )
-\end{eqnarray}
+\end{equation}
 In other words, $s$ is a function mapping $x$ to the sum of all positive integers between $0$ and $x$ when $0 \leq x \leq 13$,
@@
 since $\I$ interprets e.g.\ $s( -1 )$ and $s( 15 )$ as $105$.
-However, under the assumption that function definitions in $\sfundefs{\Sigma}$ are well-specified, 
+However, under the assumption that function definitions in $\sfundefs{\Sigma}$ are admissible, 
 we claim that $\conv(\varphi)$ is equisatisfiable to $\varphi$ for any input $\varphi$.
@@
 Moreover, information in models for $\conv( \varphi )$ contains relevant information regarding the models of $\varphi$.
-For instance, the model $\I$ for formula~(\ref{eq:ex-after}) described above interprets $c$ as $14$, 
+For example, the model $\I$ for formula~(\ref{eq:ex-after}) described above interprets $c$ as $14$, 
 while there exists a model of formula~(\ref{eq:ex-before}) also interprets $c$ as $14$.
@@
 \begin{theorem}
-If $\varphi$ be a $\Sigma$-formula that is well-defined with respect to functions $\sfundefs{\Sigma}$,
-and the set of function definitions corresponding to functions in $\sfundefs{\Sigma}$ is well-specified,
+If $\varphi$ be a $\Sigma$-formula that is in definitional form with respect to functions $\sfundefs{\Sigma}$,
+and the set of function definitions corresponding to functions in $\sfundefs{\Sigma}$ is admissible,
 then $\varphi$ and $\conv( \varphi )$ are equisatisfiable.
@@
 We will denote the latter encoding using the suffix {\bf w}.
-For instance {\bf isa+w}, contains 79 benchmarks that are the results of applying $\conv$ to each of the 79 benchmarks in {\bf isa}.
+For example, {\bf isa+w} contains 79 benchmarks that are the results of
+applying $\conv$ to each of the 79 benchmarks in {\bf isa}.
 
@@
 as we did not find a case where $\conv( \varphi )$ was satisfiable when $\varphi$ was unsatisfiable.
-This is to be expected, since the functions in our benchmark sets are expected to be well-specified (and in fact, terminating).
+This is to be expected, since the functions in our benchmark sets are expected to be admissible (and in fact, terminating).
 
@@
 We conclude from this evaluation that the translation $\conv$ enables SMT solvers to find counterexamples 
-for conjectures involving recursive functions whose definitions are well-specified.
+for conjectures involving recursive functions whose definitions are admissible.
 Both {\bf z3} and {\bf cvc4+f} were able to find a large number of counterexamples after the translation,
@@
 Function definitions $\forallf{\con{f}} \vec x.\; \varphi$ can be written in using the $\definefunreccmd$ command, 
-from Cersion 2.5~\cite{} of the SMT-LIB standard.
-For instance, the formula $\varphi$ in Example~\ref{ex:translation} 
+from Version 2.5~\cite{} of the SMT-LIB standard.
+For example, the formula $\varphi$ in Example~\ref{ex:translation} 
 can be written as:
@@
 then it will replace its list of known assertions based on the conversion $\conv$ before checking for satisfiability.
-Note that this option is model-sound \ct{define?} only if the functions provided in the input are well-specified.
+Note that this option is model-sound \ct{define?} only if the functions provided in the input are admissible.
 

### 2015-05-05 tex/main.tex
help tab completion ;)

@@
 is \emph{admissible} (\emph{in $T$}) if $\varphi_1 \wedge \ldots \wedge \varphi_n$ is $T$-satisfiable.
-For example, $\{ \forallf{\con{f}} \vec x.\;\, \con{f}( x ) \teq 0 \}$ is admissible,
+For example, $\{ \forallf{\con{f}} x.\;\, \con{f}( x ) \teq 0 \}$
+and $\{ \forallf{\con{f}} x.\;\, \con{f}( x ) \teq \con{f}( x ) \}$ are admissible,
 whereas $\{ \forallf{\con{f}} x.\;\, \con{f}( x ) \teq \con{f}( x ) + 1 \}$ is not.
-Note that an admissible function definition need not specify a terminating function, e.g.
-$\{ \forallf{\con{f}} x^\sortint.\;\, \con{f}( x ) \teq \con{f}( x - 1 ) \}$ or even $\{ \forallf{\con{f}} x.\;\, \con{f}( x ) \teq \con{f}( x ) \}$ are admissible.
 \ajr{revise}
 
-\section{Model Finding for Recursive Functions in SMT}
+\jb{TODO: Jasmin: Identify fragments: terminating recursive functions, etc.}
+
+%Terminating function specifications are always admissible, but also some
+%nonterminating
+%An admissible function definition need not specify a terminating function, e.g.
+%$\{ \forallf{\con{f}} x^\sortint.\;\, \con{f}( x ) \teq \con{f}( x - 1 ) \}$ or even $\{ \forallf{\con{f}} x.\;\, \con{f}( x ) \teq \con{f}( x ) \}$ are admissible.
+
+%\section{Model Finding for Recursive Functions in SMT}
+\section{Model Finding for (Co)recursive Functions in SMT}
 \label{sec:model-finding-rec-fun}
+
+\jb{Break up big section into several? Or at least rename it to avoid repeating the
+title. (What other subsections are planned?)}
 
@@
 have focused on finding models for quantified formulas in SMT.
-A instantiation-based procedure is shown in~\cite{GeDeM-CAV-09} to be complete for formulas in the \emph{essentially uninterpreted} fragment.
+\cite{GeDeM-CAV-09}
+A instantiation-based procedure is shown in~ to be complete for formulas in the \emph{essentially uninterpreted} fragment.
 This fragment is limited to universally quantified formulas where all occurrences of variables occur as direct subchildren of uninterpreted functions.
@@
 \bibliographystyle{abbrv}
-\bibliography{main}
+\bibliography{bib}
 }

### 2015-05-05 tex/main.tex
more

@@
 have focused on finding models for quantified formulas in SMT.
-\cite{GeDeM-CAV-09}
-A instantiation-based procedure is shown in~ to be complete for formulas in the \emph{essentially uninterpreted} fragment.
-This fragment is limited to universally quantified formulas where all occurrences of variables occur as direct subchildren of uninterpreted functions.
-For example, the formula $\forall x\typ{\sortint}.\;\, \con{f}( x ) \teq g( x ) + 5$ 
-is in the essentially uninterpreted fragment since all occurrences of $x$ occur beneath uninterpreted functions $\con{f}$ and $g$.
-Other syntactic criteria are identified in this paper that extend this fragment slightly, including cases when variables occur as arguments of arithmetic predicates.
-The techniques in~\cite{ReyEtAl-1-RR-13} can be used for finding finite models for quantified formulas over uninterpreted types as well as types having a 
-fixed finite interpretation (such as fixed-width bitvectors, datatype enumerations, etc.).
-For example, finite model finding techniques can find a model for
-the formula $\forall x\typ{\tau} y\typ{\tau}.\;\, \neg x \teq y \vee \con{f}( x ) \teq \con{f}( y )$
-where the uninterpreted type $\tau$ is interpreted as a finite set, and $\con{f}$ as injective finite map.
+Ge and de Moura \cite{GeDeM-CAV-09} presented a complete instantiation-based
+procedure for formulas in the \emph{essentially uninterpreted} fragment.
+This fragment is limited to universally quantified formulas where all
+occurrences of variables occur as direct subterms of uninterpreted
+functions. For example, the formula $\forall x%\typ{\sortint} --- the sort is not important, right?
+.\;\, \con{f}( x )
+\teq \con{g}( x ) + 5$ is in the essentially uninterpreted fragment.
+% since all
+% occurrences of $x$ occur beneath uninterpreted functions $\con{f}$ and
+% $\con{g}$.
+Other syntactic criteria are identified in this paper that extend
+this fragment slightly, including cases when variables occur as arguments of
+arithmetic predicates. The techniques in~Reynolds et al.\
+\cite{ReyEtAl-1-RR-13} can be used for finding finite models for quantified
+formulas over uninterpreted types as well as types having a fixed finite
+interpretation, such as fixed-width bitvectors and enumerated datatypes.
+Finite model finding techniques can
+find a model for the formula $\forall x\typ{\tau} y\typ{\tau}.\;\, \neg x \teq
+y \vee \con{f}( x ) \teq \con{f}( y )$ where the uninterpreted type $\tau$ is
+interpreted as a finite set and $\con{f}$ as injective finite map.
+\jb{Weird example. This formula is a tautology of EUF! Maybe you put the
+negation at the right place?}
 
@@
 \ackname.}
-We would like to thank Damien Busato-Gaston and Emmanouil Koukoutos for providing the initial set of benchmarks used in the evaluation.
+We would like to thank Damien Busato-Gaston and Emmanouil Koukoutos for
+providing the initial set of benchmarks used in the evaluation.
 }

### 2015-05-05 tex/main.tex
tuned
more

@@
 %\newcommand{\euf}{\ensuremath{\mathrm{E}}\xspace}
-\newcommand{\euf}{\ensuremath{\mathcalx{E}}\xspace}
+\newcommand{\euf}{\ensuremath{\mathcalx{UF}}\xspace}
 %\newcommand{\ari}{\ensuremath{\mathrm{A}}\xspace}
@@
 
+Recent techniques~\cite{GeDeM-CAV-09, ReyEtAl-1-RR-13} 
+have focused on finding models for quantified formulas in SMT.
+Ge and de Moura \cite{GeDeM-CAV-09} presented a complete instantiation-based
+procedure for formulas in the \emph{essentially uninterpreted} fragment.
+This fragment is limited to universally quantified formulas where all
+occurrences of variables occur as direct subterms of uninterpreted
+functions. For example, the formula $\forall x%\typ{\sortint} --- the sort is not important, right?
+.\;\, \con{f}( x )
+\teq \con{g}( x ) + 5$ is in the essentially uninterpreted fragment.
+% since all
+% occurrences of $x$ occur beneath uninterpreted functions $\con{f}$ and
+% $\con{g}$.
+Other syntactic criteria are identified in this paper that extend
+this fragment slightly, including cases when variables occur as arguments of
+arithmetic predicates. The techniques in~Reynolds et al.\
+\cite{ReyEtAl-1-RR-13} can be used for finding finite models for quantified
+formulas over uninterpreted types as well as types having a fixed finite
+interpretation, such as fixed-width bitvectors and enumerated datatypes.
+Finite model finding techniques can
+find a model for the formula $\forall x\typ{\tau} y\typ{\tau}.\;\, \neg x \teq
+y \vee \con{f}( x ) \teq \con{f}( y )$ where the uninterpreted type $\tau$ is
+interpreted as a finite set and $\con{f}$ as injective finite map.
+\jb{Weird example. This formula is a tautology of EUF! Maybe you put the
+negation at the right place?}
+
+Unfortunately, these fragments cannot accommodate the vast majority of
+quantified formulas that correspond to recursive function definitions.
+In particular, the fragments in~\cite{GeDeM-CAV-09} are not applicable when the argument of a recursive function is used in various ways (e.g. within an arithmetic expression), 
+while the fragment in~\cite{ReyEtAl-1-RR-13} is not applicable for functions over infinite domains such as the integers.
+So, for instance, the quantified formula $\forall x\typ{\sortint}.\;\, \lite( p \leq 0, p( x ) \teq 0, p( x ) \teq 2 * p( x - 1 ) )$ 
+is not in the essentially uninterpreted fragment (since $x$ occurs as a child of $-$),
+nor does finite model finding apply (since $x$ is of type $\sortint$).
+
+Here, we advocate an approach that translates formulas involving function
+definitions (such as the one mentioned in the previous paragraph)
+into equisatisfiable \ct{to be defined} formulas where the
+techniques~\cite{GeDeM-CAV-09, ReyEtAl-1-RR-13} are applicable.
+
 \paragraph{Contributions}
@@
 \end{itemize}
-
-\section{Related Work}
-
-first-order model finders: SEM, MACE/Mace, Paradox, FMDarwin, iProver
-  either direct search or reduction to SAT
-
-Alloy (and Alloy idiom by Kuncak \& Jackson), Kodkod
-
-The model finder for KIV based on Alloy
-
-Refute
-  * unsound handling of datatypes
-
-Nitpick
-  * inductive / coinductive predicates
-  * quotient types
-
-Leon
-
-in smt : z3, cvc4 (briefly, text in main section gives detail)
 
@@
 Given a signature $\Sigma$,
-we partition its function symbols $\sfuns{\Sigma}$ into $\sfundefs{\Sigma} \mathrel{\uplus} \sfunndefs{\Sigma}$,
-where $\sfundefs{\Sigma}$ is the set of \emph{defined} function symbols,
-and consider $\Sigma$-formulas that are in definitional form with respect to functions $\sfundefs{\Sigma}$.
+we distinguish the subset of \emph{defined} function symbols $\sfundefs{\Sigma}
+\subseteq \sfuns{\Sigma}$.
+We consider $\Sigma$-formulas that are in definitional form with respect to
+functions $\sfundefs{\Sigma}$.
 
@@
 
-%\section{Model Finding for Recursive Functions in SMT}
-\section{Model Finding for (Co)recursive Functions in SMT}
-\label{sec:model-finding-rec-fun}
-
-\jb{Break up big section into several? Or at least rename it to avoid repeating the
-title. (What other subsections are planned?)}
-
-Recent techniques~\cite{GeDeM-CAV-09, ReyEtAl-1-RR-13} 
-have focused on finding models for quantified formulas in SMT.
-Ge and de Moura \cite{GeDeM-CAV-09} presented a complete instantiation-based
-procedure for formulas in the \emph{essentially uninterpreted} fragment.
-This fragment is limited to universally quantified formulas where all
-occurrences of variables occur as direct subterms of uninterpreted
-functions. For example, the formula $\forall x%\typ{\sortint} --- the sort is not important, right?
-.\;\, \con{f}( x )
-\teq \con{g}( x ) + 5$ is in the essentially uninterpreted fragment.
-% since all
-% occurrences of $x$ occur beneath uninterpreted functions $\con{f}$ and
-% $\con{g}$.
-Other syntactic criteria are identified in this paper that extend
-this fragment slightly, including cases when variables occur as arguments of
-arithmetic predicates. The techniques in~Reynolds et al.\
-\cite{ReyEtAl-1-RR-13} can be used for finding finite models for quantified
-formulas over uninterpreted types as well as types having a fixed finite
-interpretation, such as fixed-width bitvectors and enumerated datatypes.
-Finite model finding techniques can
-find a model for the formula $\forall x\typ{\tau} y\typ{\tau}.\;\, \neg x \teq
-y \vee \con{f}( x ) \teq \con{f}( y )$ where the uninterpreted type $\tau$ is
-interpreted as a finite set and $\con{f}$ as injective finite map.
-\jb{Weird example. This formula is a tautology of EUF! Maybe you put the
-negation at the right place?}
-
-Unfortunately, these fragments do not coincide with the vast majority of quantified formulas that correspond to recursive function definitions.
-In particular, the fragments in~\cite{GeDeM-CAV-09} are not applicable when the argument of a recursive function is used in various ways (e.g. within an arithmetic expression), 
-while the fragment in~\cite{ReyEtAl-1-RR-13} is not applicable for functions over infinite domains such as the integers.
-So, for instance, the quantified formula $\forall x\typ{\sortint}.\;\, \lite( p \leq 0, p( x ) \teq 0, p( x ) \teq 2 * p( x - 1 ) )$ 
-is not in the essentially uninterpreted fragment (since $x$ occurs as a child of $-$),
-nor does finite model finding apply (since $x$ is of type $\sortint$).
-
-Here, we advocate an approach that translates formulas involving function definitions (such as the one mentioned in the previous paragraph)
-into equisatisfiable \ct{to be defined} formulas where the techniques~\cite{GeDeM-CAV-09, ReyEtAl-1-RR-13} are applicable.
-
-\subsection{A Translation for Well-Specified Function Definitions}
+%\section{A Translation for Admissible Function Definitions}
+%\section{A Translation for Function Definitions}
+\section{The Translation}
 \label{sec:encoding}
 
-In the remainder of the section, we fix a signature $\Sigma$.
-Say that $\Sigma$-formula $\varphi$ is in definitional form with respect to functions $\sfundefs{\Sigma}$, 
-and the set of function definitions corresponding to $\sfundefs{\Sigma}$ are admissible.
+Let us fix a signature $\Sigma$, a $\Sigma$-formula
+$\varphi$ in definitional form with respect to functions $\sfundefs{\Sigma}$,
+whose definitions are admissible.
 %We may translate $\varphi$ into a equisatisfiable formula $\psi$ for which known model-finding procedures~\cite{GeDeM-CAV-09, ReyEtAl-1-RR-13} are applicable.
-This section presents a method for constructing an extended signature $\extendsig{ \Sigma }$,
-and a $\extendsig{ \Sigma }$-formula $\psi$ such that $\varphi$ and $\psi$ are equisatisfiable.
-The intuition behind this translation is to use an uninterpreted type \ct{to be defined} to abstract the set of \emph{relevant} tuples for each recursive function $\con{f}$,
-and restrict the quantification of the function definition for $\con{f}$ to a variable of this type.
-Informally, the relevant tuples $\vec t$ of a function $\con{f}$ are the ones for which the interpretation of $\con{f}( \vec t )$ is relevant to the satisfiability of $\varphi$.
+This section presents a method for constructing an extended signature
+$\extendsig{ \Sigma }$ and a $\extendsig{ \Sigma }$-formula $\psi$ such that
+$\varphi$ and $\psi$ are equisatisfiable.
+\jb{Equisatisfiability as defined in Sect.\ 2 requires the same $\Sigma$ and
+the same $T$ for both formulas, which isn't the case here.}
+The intuition behind this translation
+is to use an uninterpreted type \ct{to be defined} to abstract the set of
+\emph{relevant} tuples for each recursive function $\con{f}$ and restrict the
+quantification of the function definition for $\con{f}$ to a variable of this
+type. Informally, the relevant tuples $\vec t$ of a function $\con{f}$ are the
+ones for which the interpretation of $\con{f}( \vec t )$ is relevant to the
+satisfiability of $\varphi$.
 
@@
 
+\section{Related Work}
+
+first-order model finders: SEM, MACE/Mace, Paradox, FMDarwin, iProver
+  either direct search or reduction to SAT
+
+Alloy (and Alloy idiom by Kuncak \& Jackson), Kodkod
+
+The model finder for KIV based on Alloy
+
+Refute
+  * unsound handling of datatypes
+
+Nitpick
+  * inductive / coinductive predicates
+  * quotient types
+
+Leon
+
+in smt : z3, cvc4 (briefly, text in main section gives detail)
+
 \section{Conclusion}

### 2015-05-05 tex/main.tex
Work on well-specified.

@@
 
-%In this paper, we associate quantified formulas with uninterpreted functions (informally, those whose semantics they define).
 In this paper, we consider \emph{annotated quantifed formulas} of the form
@@
 \vec x. \varphi$ as shorthand for $\neg \forall \vec x.\; \neg \varphi$.
+For a formula $\varphi$, 
+we write $\varphi[\vec t/\vec x]$ to denote the result of replacing all occurrences of $\vec x$ with $\vec t$.
 
@@
 
-A set of function definitions $\{ \varphi_1, \ldots, \varphi_n \}$ 
-is \emph{admissible} (\emph{in $T$}) if $\varphi_1 \wedge \ldots \wedge \varphi_n$ is $T$-satisfiable.
-For example, $\{ \forallf{\con{f}} x.\;\, \con{f}( x ) \teq 0 \}$
-and $\{ \forallf{\con{f}} x.\;\, \con{f}( x ) \teq \con{f}( x ) \}$ are admissible,
-whereas $\{ \forallf{\con{f}} x.\;\, \con{f}( x ) \teq \con{f}( x ) + 1 \}$ is not.
-\ajr{revise}
+%A set of function definitions $\{ \varphi_1, \ldots, \varphi_n \}$ 
+%is \emph{well-specified (in $T$)} if $\varphi_1 \wedge \ldots \wedge \varphi_n$ is $T$-satisfiable.
+%For instance, $\{ \forallf{f} \vec x. f( x ) \teq 0 \}$ is well-specified,
+%whereas $\{ \forallf{f} x. f( x ) \teq f( x ) + 1 \}$ is not. 
+%Note that a well-specified function definition need not specify a terminating function, e.g.
+%$\{ \forallf{f} x^\sortint. f( x ) \teq f( x - 1 ) \}$ or even $\{ \forallf{f} x. f( x ) \teq f( x ) \}$ are well-specified.
+%\ajr{revise}
+
+\begin{definition}
+Given a set of function definitions $\mathcal{D} = \{ \forallf{f_1} \vec x. \varphi_1, \ldots, \forallf{f_n} \vec x. \varphi_n \}$,
+a formula $\varphi$ is \emph{closed under function expansion with respect to $\mathcal{D}$} if
+\begin{eqnarray}
+\varphi \text{ entails } \displaystyle\bigwedge\limits_{i=1}^n \{ \varphi_i[ \vec t / \vec x ] \mid f_i( \vec t ) \in \terms( \varphi ) \}
+\end{eqnarray}
+The set $\mathcal{D}$ is \emph{admissible} if whenever there exists a model for $\varphi$ that is closed under function expansion,
+there exists a model for $\varphi \wedge \forallf{f_1} \vec x. \varphi_1 \wedge \ldots \wedge \forallf{f_n} \vec x. \varphi_n$.
+\end{definition}
+
+For instance, the set $\{ \forallf{f} \vec x. f( x ) \teq 0 \}$ is well-specified.
+A set of well-specified function definitions need be terminating, e.g.
+$\{ \forallf{f} x^\sortint. f( x ) \teq f( x - 1 ) \}$ or even $\{ \forallf{f} x. f( x ) \teq f( x ) \}$ are well-specified.
+On the other hand, the set $\{ \forallf{f} x. f( x ) \teq f( x ) + 1 \}$ is not well-specified, 
+since the empty set of formulas is closed under function expansion with respect to this set,
+and there is no model satisfying $\forallf{f} x. f( x ) \teq f( x ) + 1$.
+Consider the set $\{ \forallf{f} x. f( x ) \teq f( x ), \forallf{g}. \lite( \neg f( x ) \teq 5, g( x ) \teq g( x ) + 1, g( x ) \teq f( x ) ) \}$.
+While this set has a model where $f$ and $g$ are interpreted as the constant $5$, it is not well-specified
+since the formula $f( 0 ) \teq 4$ is closed under function expansion with respect to this set,
+but there exists no interpretation satisfying both $f( 0 ) \teq 4$ and $\forallf{g}. \lite( \neg f( x ) \teq 5, g( x ) \teq g( x ) + 1, g( x ) \teq f( x ) )$.
 
@@
 and $s$ as the (almost constant) function:
-\begin{equation}
+\begin{equation} \label{eq:approx-interp}
 \lambda x\typ{Int}. \lite( x \teq 0, 0, \lite( x \teq 1, 1, \lite( x \teq 2, 3, \ldots, \lite( x \teq 13, 91, 105 ) \ldots )
@@
 \begin{theorem}
-If $\varphi$ be a $\Sigma$-formula that is in definitional form with respect to functions $\sfundefs{\Sigma}$,
-and the set of function definitions corresponding to functions in $\sfundefs{\Sigma}$ is admissible,
+If $\varphi$ is a $\Sigma$-formula that is in definitional form with respect to functions $\sfundefs{\Sigma}$,
+and the set of function definitions $\mathcal{D}$ corresponding to functions in $\sfundefs{\Sigma}$ is admissible,
 then $\varphi$ and $\conv( \varphi )$ are equisatisfiable.
@@
 \begin{proof}
-(Sketch)  First, we show that if $\varphi$ has a model $\I$, then $\conv( \varphi )$ has a model $\J$.
+(Sketch)  First, we show that if $\varphi$ is satisfied by $\Sigma$-interpretation $\I$, then $\conv( \varphi )$ is satisfied by a $\extendsig{\Sigma}$-interpretation $\J$.
 Let $\J$ be the $\extendsig{\Sigma}$-interpretation that interprets all types $\tau \in \ssorts{\Sigma}$ as $\tau^\I$,
@@
 
-Second, we show that if $\conv( \varphi )$ has a model $\I$, then $\varphi$ has a model $\J$.
+Second, we show that if $\conv( \varphi )$ is satisfied by a $\extendsig{\Sigma}$-interpretation $\I$, then $\varphi$ is satisfied by $\Sigma$-interpretation $\J$.
+Since $\varphi$ is well-defined with respect to functions defined by $\mathcal{D}$, we know $\varphi$ is of the form $\mathcal{D} \wedge \varphi_0$.
+We first argue that $\I$ satisfies a $\Sigma$-formula $\psi$ that is closed under function expansion with respect to $\mathcal{D}$,
+and where $\psi$ entails $\varphi_0$.
 
+Since there exists a model for $\psi$ that is closed under function expansion with respect to $\mathcal{D}$, 
+then since $\mathcal{D}$ is well-specified,
+by definition there exists a model for $\psi \wedge \mathcal{D}$, which entails $\mathcal{D} \wedge \varphi_0$, which is $\varphi$.
 \end{proof}
+
+Let us revisit the translated formula~(\ref{eq:ex-after}) from Example~\ref{ex:translation}.
+This formula is unsatisfiable if and only if formula~(\ref{eq:ex-before}) is unsatisfiable since $\fargsort{f}$ can be interpreted as $\sortint$,
+and $\farg{f}$ can be interpreted as the identity function.
+Conversely, we claim formula~(\ref{eq:ex-after}) is satisfiable if and only if formula~(\ref{eq:ex-before}) is satisfiable,
+noting that the set $\{ \forall_{s} x\typ{\sortint}. ite( x \leq 0, s(x) \teq 0, s( x ) \teq x + s( x-1 ) ) \}$ is well-specified.
+As stated in the proof of the previous theorem, 
+the reason is that any model of formula~(\ref{eq:ex-after}) must satisfy a formula that is closed under function expansion,
+hence implying that formula~(\ref{eq:ex-before}) has a model.
+In more detail, let $\varphi_s$ be the formula $\forall_{s} x. ite( x \leq 0, s(x) \teq 0, s( x ) \teq x + s( x-1 ) )$.
+The rightmost conjunct 
+$\exists z\typ{\fargsort{s}}. \farg{s}( z ) \teq c$ ensures 
+that any model of formula~(\ref{eq:ex-after}) satisfying $s( c ) > 100$ also satisfies $\varphi_s[c/x]$.
+Similarly, the nested existential $\exists z. \farg{s}( z ) \teq \farg{s}( y )-1 )$
+ensures that this constraint is enforced recursively.
+%Given a model $\I$ of formula~(\ref{eq:ex-after}) that interprets $c$ as $14$, it is easy to see that $\I$ also satisfies the formula:
+%\begin{eqnarray}
+%\begin{split}
+%s( c ) > 100 \wedge & \displaystyle\bigwedge\limits_{i=0}^{13} ( \neg c-i \leq 0 \wedge s( c-i ) \teq i + s( c-i-1 ) ) \\
+%                    & \wedge c-14 \leq 0 \wedge s( c-14 ) \teq 0
+%\end{split}
+%\end{eqnarray}
 
@@
 then it will replace its list of known assertions based on the conversion $\conv$ before checking for satisfiability.
-Note that this option is model-sound \ct{define?} only if the functions provided in the input are admissible.
+%If the functions provided in the input are well-specified, 
+%then a ``satisfiable" response from the solver implies that a model exists for the original formula.
+Accordingly, the solver will output an approximation of the interpretation for recursive function definitions,
+that is, it may output the interpretation~(\ref{eq:approx-interp}) from Section~\ref{sec:encoding} for $s$.
+
+Notice the option ``--fmf-fun" effectively instructs the SMT solver to assume that the functions provided by the user are admissible.
+If this option is used for a set of functions definitions that is \emph{not} admissible, then the solver is not model-sound, that is, 
+it may return ``satisfiable" for unsatisfiable inputs.
+As such,
+this option targets users who wish to determine the satisfiability of inputs that involving function definitions that are known to be consistent,
+whereas it does not target users who wish to check the consistency of a set of function definitions.
 

### 2015-05-05 tex/main.tex
more

--- 2015-05-05 tex/main.tex
@@
 \newcommand{\pnone}{\con{none}}
-\newcommand{\ptrue}{\con{true}}
-\newcommand{\pfalse}{\con{false}}
+\newcommand{\ptrue}{\con{pos}}
+\newcommand{\pfalse}{\con{neg}}
 \newcommand{\pol}{\con{pol}}
@@
 
+%\newcommand\concret{\con{a}}
+%\newcommand\concret{\con{c}}
+\newcommand\concret{\gamma} %%% from the abstract interpretation literature
+
 \newcommand{\boolop}{\oplus}
@@
 \newcommand{\fnull}{\emptyset}
-\newcommand{\farg}[1]{a_{#1}}
-\newcommand{\vecfarg}[1]{\vec{a}_{#1}}
-\newcommand{\fargsort}[1]{\upsilon_{#1}}
+\newcommand{\vecfarg}[1]{\vec{\concret}_{\vthinspace#1}}
+\newcommand{\farg}[1]{\concret_{\vthinspace#1}}
+\newcommand{\fargx}[2]{\concret^{\vthinspace#2}_{\vthinspace#1}}
+\newcommand{\fargsort}[1]{\upsilon_{\vthinspace#1}}
 
@@
 is to use an uninterpreted type \ct{to be defined} to abstract the set of
-\emph{relevant} tuples for each recursive function $\con{f}$ and restrict the
+\relax{relevant} tuples for each recursive function $\con{f}$ and restrict the
 quantification of the function definition for $\con{f}$ to a variable of this
@@
 In more detail, 
-say our signature $\Sigma$ contains a set of defined function symbols $\sfundefs{\Sigma} \subseteq \sfuns{\Sigma}$.
+suppose our signature $\Sigma$ contains a set of defined function symbols $\sfundefs{\Sigma} \subseteq \sfuns{\Sigma}$.
 We extend the signature of $\Sigma$ to a signature $\extendsig{\Sigma}$ containing the following additional definitions.
-For each $\con{f} \in \sfundefs{\Sigma}$ of type $\tau_1 \times \ldots \times \tau_n \rightarrow \tau$, 
-the signature $\extendsig{\Sigma}$ includes:
-\begin{enumerate}
-\item[-] an uninterpreted type $\fargsort{\con{f}}$, and
-\item[-] a vector of $n$ uninterpreted functions $\farg{\con{f}}^1$, $\ldots$, $\farg{\con{f}}^n$ of type $(\fargsort{\con{f}} \rightarrow \tau_1)$, $\ldots$, $(\fargsort{\con{f}} \rightarrow \tau_n)$.
-\end{enumerate}
+For each $\con{f} : \tau_1 \times \ldots \times \tau_n \rightarrow \tau \in \sfundefs{\Sigma}$,
+the signature $\extendsig{\Sigma}$ includes
+\begin{itemize}
+\item an uninterpreted \emph{abstract type} $\fargsort{\con{f}}$;
+\item $n$ uninterpreted \emph{concretization functions} $\fargx{\con{f}}{1} : \fargsort{\con{f}} \rightarrow \tau_1$, \ldots, $\fargx{\con{f}}{n} : \fargsort{\con{f}} \rightarrow \tau_n$.
+\end{itemize}
 %The interpretation of uninterpreted sort $\fargsort{\con{f}}$ will denote the elements (tuples) on which the function $\con{f}$ is applied.
@@
 
-Given the extended signature, we run the procedure $\conv$ in Figure~\ref{fig:encoding} on $\varphi$.
-This calls the subprocedure $\conv_0$, which takes two arguments: the term $t$ to translate, and a \emph{polarity} $p$ (either $\ptrue$, $\pfalse$, or $\pnone$).
-It returns a pair of the form $( r, D )$, where $r$ is a term of type $\tau$, and $D$ is a set of formulas.
-%The role of $D$ is to ensure that the (restricted) function definition for $\con{f}$ includes certain tuples in its domain, which we explain more in the following.
-At a high level, the translation modifies $\varphi$ in two ways.
-First, it restricts the quantification on function definitions for $\con{f}$ to the corresponding uninterpreted type $\fargsort{\con{f}}$.
-Second, it augments $\varphi$ with additional constraints of the form $\exists z\typ{\fargsort{\con{f}}}. ( \vecfarg{\con{f}}( z ) \teq \vec t )$
-which ensure that this restriction for the defintion $\con{f}$ covers all relevant tuples of terms,
-namely the ones for which an $\con{f}$-application exists and is relevant to the satisfiability of the current branch of $\varphi$.
+The procedure $\conv$ in Figure~\ref{fig:encoding} translates the $\Sigma$-formula
+$\varphi$ into an $\extendsig{\Sigma}$-formula.
+This calls the subprocedure $\conv_0$, which takes two arguments:\ the term $t$
+to translate and a \emph{polarity} $p$ (either $\ptrue$, $\pfalse$, or
+$\pnone$). It returns a pair $( r,\, D )$, where $r$ is a term of
+type $\tau$ and $D$ is a set of formulas.
+\jb{Explain r and D here briefly.}
+%The role of $D$ is to ensure that
+%the (restricted) function definition for $\con{f}$ includes certain tuples in
+%its domain, which we explain more in the following.
+
+The translation modifies $\varphi$ in two main ways. First, it restricts the
+quantification on function definitions for $\con{f}$ to the corresponding
+uninterpreted type $\fargsort{\con{f}}$ and inserts appropriate . Second, it augments $\varphi$ with
+additional constraints of the form $\exists z\typ{\fargsort{\con{f}}}. (
+\vecfarg{\con{f}}( z ) \teq \vec t )$ which ensure that this restriction for
+the defintion $\con{f}$ covers all relevant tuples of terms, namely the ones
+for which an $\con{f}$-application exists and is relevant to the satisfiability
+of the current branch of $\varphi$.
 
@@
 interprets $\fargsort{\con{f}}$ as a tuple of values of type $\tau_1^\I \times \ldots \times \tau_n^\I$,
-and each $\farg{\con{f}}^i$ as the $i^{th}$ projection on such tuples for each $i = 1, \ldots, n$.
+and each $\fargx{\con{f}}{i}$ as the $i^{th}$ projection on such tuples for each $i = 1, \ldots, n$.
 Note that $\J$ satisfies \emph{every} constraint of the form $\exists z\typ{\fargsort{\con{f}}}. ( \vecfarg{\con{f}}( z ) \teq \vec t )$,

### 2015-05-05 tex/main.tex
more

@@
 
+\newcommand{\negvthinspace}{\kern-0.083333em}
 \newcommand{\vthinspace}{\kern+0.083333em}
@@
 to translate and a \emph{polarity} $p$ (either $\ptrue$, $\pfalse$, or
-$\pnone$). It returns a pair $( r,\, D )$, where $r$ is a term of
-type $\tau$ and $D$ is a set of formulas.
-\jb{Explain r and D here briefly.}
+$\pnone$). It returns a pair $( t'\negvthinspace,\, D )$, where $t'$ is a term of
+type $\tau$ and $D$ is a set of formulas encoding additional constraints.
 %The role of $D$ is to ensure that
@@
 quantification on function definitions for $\con{f}$ to the corresponding
-uninterpreted type $\fargsort{\con{f}}$ and inserts appropriate . Second, it augments $\varphi$ with
-additional constraints of the form $\exists z\typ{\fargsort{\con{f}}}. (
-\vecfarg{\con{f}}( z ) \teq \vec t )$ which ensure that this restriction for
-the defintion $\con{f}$ covers all relevant tuples of terms, namely the ones
-for which an $\con{f}$-application exists and is relevant to the satisfiability
-of the current branch of $\varphi$.
+uninterpreted type $\fargsort{\con{f}}$ and inserts appropriate coercions
+$\fargx{\con{f}}{i}$. Second, it augments $\varphi$ with additional constraints
+$\exists z : \fargsort{\con{f}}.\;\, \vecfarg{\con{f}}( z ) \teq \vec t$; these
+ensure that the restricted definition for $\con{f}$ covers all relevant tuples
+of terms, namely the ones for which an $\con{f}$ application exists and is
+relevant to the satisfiability of the current branch of $\varphi$.
+\jb{what does ``current branch of'' mean? can we drop it / rephrase it / vague
+it up?}
 

### 2015-05-05 tex/main.tex
more

@@
 This calls the subprocedure $\conv_0$, which takes two arguments:\ the term $t$
-to translate and a \emph{polarity} $p$ (either $\ptrue$, $\pfalse$, or
-$\pnone$). It returns a pair $( t'\negvthinspace,\, D )$, where $t'$ is a term of
+to translate and a \emph{polarity} $p$, which is either $\ptrue$, $\pfalse$, or
+$\pnone$. It returns a pair $( t'\negvthinspace,\, D )$, where $t'$ is a term of
 type $\tau$ and $D$ is a set of formulas encoding additional constraints.
@@
 
-In the case that $t$ is an application of (interpreted) predicate $\boolop$ (e.g., a boolean connective, or equality),
-we recursively call $\conv$ on its subchildren $t_i$ and polarity $\pol( \boolop, i, p )$, where $\pol$ is defined as\footnote{The negation $\neg p$ of polarity $p$ is 
-$\pfalse$ if $p$ is $\ptrue$,
-$\ptrue$ if $p$ is $\pfalse$, and
-$\pnone$ if $p$ is $\pnone$.}:
+If $t$ is an application of an interpreted predicate $\boolop
+\in \{{\neg}, {\land}, {\lor}, {\teq}, {\lite}\}$,
+we recursively call $\conv$ on the arguments $t_i$ and polarity $\pol( \boolop,
+i, p )$, where $\pol$ is defined as
 \begin{equation*}
-\pol( \boolop, i, p ) = \begin{cases}
-                         \pnone & \boolop \equiv \teq \text{ or } ( \boolop \equiv \lite \text{ and } i=0 ) \\
-                         \neg p & \boolop \equiv \neg \\
+\pol( \boolop, i, p ) =
+\begin{cases}
+\pnone & \text{if $\boolop = {\teq}$ or $(\boolop, i) = (\lite, 0)$} \\
+\neg p & \text{if $\boolop = {\neg}$} \\
 p & \text{otherwise}
@@
 \end{equation*}
-If $t$ has positive polarity,
-we construct a conjunction of the result of the recursive calls with formulas $D_1, \ldots, D_n$.
-Similarly, if $t$ has negative polarity,
-we construct a disjunction of the result of the recursive calls with the negation of formulas $D_1, \ldots, D_n$.
-It $t$ has no polarity,
+The negation $\neg p$ of polarity $p$ is $\pfalse$ if $p$ is $\ptrue$, $\ptrue$
+if $p$ is $\pfalse$, and $\pnone$ if $p$ is $\pnone$.
+
+If $t$ has positive polarity, we construct a conjunction of the result of the
+recursive calls with formulas $D_1, \ldots, D_n$. Similarly, if $t$ has
+negative polarity, we construct a disjunction of the result of the recursive
+calls with the negation of formulas $D_1, \ldots, D_n$. It $t$ has no polarity,
 then we return the union of $D_1 \cup \ldots \cup D_n$.
 
-In the case that $t$ is a quantified formula $\forallf{\con{f}} \vec x.\; t_1$, it recursively calls $\conv$ on its body with the same polarity.
-In the case that $t$ is a function definition of a function in $\sfundefs{\Sigma}$,
-we instead construct a quantified formula over a single variable $y$ of type $\fargsort{\con{f}}$,
+If $t$ is a quantified formula $\forallf{\con{f}} \vec x.\; t_1$, it
+recursively calls $\conv$ on its body with the same polarity. If $t$ is a
+function definition of a function in $\sfundefs{\Sigma}$, we instead construct
+a quantified formula over a single variable $y$ of type $\fargsort{\con{f}}$,
 and replace all occurrences of $\vec x$ in $r_1$ with $\vecfarg{\con{f}}( y )$.
-In this case, since function definitions are a top-level conjunct, by case analysis on the return values of $\conv$, we know that $D_1$ is empty.
-In the case that $t$ is not a function definition, 
-the resulting quantified formula is reconstructed with the term $r_1$ returned by the recursive call,
-and a quantified prefix is appended to the formula $D_1$ returned by the recursive call.
+In this case, since function definitions are a top-level conjunct, by case
+analysis on the return values of $\conv$, we know that $D_1$ is empty. In the
+case that $t$ is not a function definition, the resulting quantified formula is
+reconstructed with the term $r_1$ returned by the recursive call, and a
+quantified prefix is appended to the formula $D_1$ returned by the recursive
+call.
 
-Otherwise, $t$ is either an application of an uninterpreted predicate, or a term of a type other than $\Bool$.
-In this case, the second component of our return value contains a constraint of the form $\exists z\typ{\fargsort{\con{f}}}. ( \vecfarg{\con{f}}( z ) \teq \vec t )$
-for each subterm $\con{f}( \vec t )$ of $t$ such that $\con{f}$ is a function in $\sfundefs{\Sigma}$.
-When such constraints are asserted positively, this indicates that the tuple of uninterpreted functions $\vecfarg{\con{f}}$ must include $\vec t$ in its range.
+Otherwise, $t$ is either an application of an uninterpreted predicate, or a
+term of a type other than $\Bool$. In this case, the second component of our
+return value contains a constraint of the form $\exists
+z\typ{\fargsort{\con{f}}}. ( \vecfarg{\con{f}}( z ) \teq \vec t )$ for each
+subterm $\con{f}( \vec t )$ of $t$ such that $\con{f}$ is a function in
+$\sfundefs{\Sigma}$. When such constraints are asserted positively, this
+indicates that the tuple of uninterpreted functions $\vecfarg{\con{f}}$ must
+include $\vec t$ in its range.
 
@@
 \end{equation} 
-Here, $s$ is a function that returns the sum of positive numbers between $0$ and its argument $x$.
-The formula $\varphi$ is in definitional form with respect to functions $\sfundefs{\Sigma_u}$,
-and states that the sum of all positive numbers between $0$ and $c$ is greater than $100$.
-This formula is satisfiable with a model that interprets $c$ as an integer greater than or equal to $14$.
-Due to the universally quantified formula in the left disjunct,
-current SMT techniques~\cite{GeDeM-CAV-09, ReyEtAl-1-RR-13} are unable to find a model for this formula.
-The signature $\extendsig{\Sigma}$ includes the type $\fargsort{s}$,
-and the uninterpreted function $\farg{s}$ of type $\fargsort{s} \rightarrow \sortint$.
-The result of $\conv( \varphi, \ptrue )$ is the pair $( \psi, \emptyset )$, where after simplification, $\psi$ is the $\extendsig{\Sigma}$-formula:
+%
+Here, $s$ is a function that returns the sum of positive numbers between $0$
+and its argument $x$. The formula $\varphi$ is in definitional form with
+respect to functions $\sfundefs{\Sigma_u}$, and states that the sum of all
+positive numbers between $0$ and $c$ is greater than $100$. This formula is
+satisfiable with a model that interprets $c$ as an integer greater than or
+equal to $14$. Due to the universally quantified formula in the left disjunct,
+current SMT techniques~\cite{GeDeM-CAV-09, ReyEtAl-1-RR-13} are unable to find
+a model for this formula. The signature $\extendsig{\Sigma}$ includes the type
+$\fargsort{s}$, and the uninterpreted function $\farg{s}$ of type $\fargsort{s}
+\rightarrow \sortint$. The result of $\conv( \varphi, \ptrue )$ is the pair $(
+\psi, \emptyset )$, where after simplification, $\psi$ is the
+$\extendsig{\Sigma}$-formula
+%
 \begin{eqnarray} \label{eq:ex-after}
@@
 \end{eqnarray} 
-The universal quantification in the resulting formula is over an uninterpreted type $\fargsort{s}$, 
-thus enabling finite model finding techniques~\cite{ReyEtAl-1-RR-13} to search for a finite interpretation for $\fargsort{s}$.
-Moreover, all occurrences of $y$ in this formula are beneath applications of the uninterpreted function $\farg{s}$,
-implying that the formula is in the essentially uninterpreted fragment of~\cite{GeDeM-CAV-09}, for which there exists a complete instantiation procedure.
-We observe that both \cvc and \ziii run indefinitely on formula~(\ref{eq:ex-before}),
-and both produce a model for formula~(\ref{eq:ex-after}) in $<0.1$ second.
-$\square$
+%
+The universal quantification in the resulting formula is over an uninterpreted
+type $\fargsort{s}$, thus enabling finite model finding
+techniques~\cite{ReyEtAl-1-RR-13} to search for a finite interpretation for
+$\fargsort{s}$. Moreover, all occurrences of $y$ in this formula are beneath
+applications of the uninterpreted function $\farg{s}$, implying that the
+formula is in the essentially uninterpreted fragment of~\cite{GeDeM-CAV-09},
+for which there exists a complete instantiation procedure. We observe that both
+\cvc and \ziii run indefinitely on formula~(\ref{eq:ex-before}), and both
+produce a model for formula~(\ref{eq:ex-after}) in $<0.1$ second. $\square$
 \end{example}

### 2015-05-05 tex/main.tex
Bug fixes, converge on terminologies.

--- 2015-05-05 tex/main.tex
@@
 \newcommand{\pnone}{\con{none}}
-\newcommand{\ptrue}{\con{pos}}
-\newcommand{\pfalse}{\con{neg}}
+\newcommand{\ptrue}{\con{true}}
+\newcommand{\pfalse}{\con{false}}
 \newcommand{\pol}{\con{pol}}
@@
 
-%\newcommand\concret{\con{a}}
-%\newcommand\concret{\con{c}}
-\newcommand\concret{\gamma} %%% from the abstract interpretation literature
-
 \newcommand{\boolop}{\oplus}
@@
 \newcommand{\fnull}{\emptyset}
-\newcommand{\vecfarg}[1]{\vec{\concret}_{\vthinspace#1}}
-\newcommand{\farg}[1]{\concret_{\vthinspace#1}}
-\newcommand{\fargx}[2]{\concret^{\vthinspace#2}_{\vthinspace#1}}
-\newcommand{\fargsort}[1]{\upsilon_{\vthinspace#1}}
+\newcommand{\farg}[1]{a_{#1}}
+\newcommand{\vecfarg}[1]{\vec{a}_{#1}}
+\newcommand{\fargsort}[1]{\upsilon_{#1}}
 
@@
 Finite model finding techniques can
-find a model for the formula $\forall x\typ{\tau} y\typ{\tau}.\;\, \neg x \teq
-y \vee \con{f}( x ) \teq \con{f}( y )$ where the uninterpreted type $\tau$ is
+find a model for the formula $\forall x\typ{\tau} y\typ{\tau}.\;\, x \teq
+y \vee \neg \con{f}( x ) \teq \con{f}( y )$ where the uninterpreted type $\tau$ is
 interpreted as a finite set and $\con{f}$ as injective finite map.
-\jb{Weird example. This formula is a tautology of EUF! Maybe you put the
-negation at the right place?}
 
@@
 (\,\emph{for $\con{f}$}\vthinspace) if $\vec x$ is a tuple of variables of type
-$\tau_1, \ldots, \tau_n$ and $\forallf{\con{f}} \vec x.\; \varphi$ is equivalent to
-$\forall \vec x.\;\, \con{f}( \vec x ) \teq t$ for some $t$. We write $\forall \vec
+$\tau_1, \ldots, \tau_n$ and $\varphi$ is a (quantifier-free) formula equivalent to
+$\con{f}( \vec x ) \teq t$ for some $t$. We write $\forall \vec
 x.\; \varphi$ as shorthand for $\forallf{\fnull} \vec x.\; \varphi$, and $\exists
 \vec x. \varphi$ as shorthand for $\neg \forall \vec x.\; \neg \varphi$.
+For a formula $\varphi$, 
+we write $\varphi[\vec t/\vec x]$ to denote the result of replacing all occurrences of $\vec x$ with $\vec t$.
 
@@
 We call $\psi$ the \emph{conjecture} of $\varphi$.
-\ct{Don't we need additional restrictions on the $\varphi_i$'s here?
-(E.g., they are quantifier-free, or perhaps existential or containing only let binders.)}
+%\ct{Don't we need additional restrictions on the $\varphi_i$'s here?
+%(E.g., they are quantifier-free, or perhaps existential or containing only let binders.)}
+% Now fixed above -ajr
 \end{definition}
@@
 
-A set of function definitions $\{ \varphi_1, \ldots, \varphi_n \}$ 
-is \emph{admissible} (\emph{in $T$}) if $\varphi_1 \wedge \ldots \wedge \varphi_n$ is $T$-satisfiable.
-For example, $\{ \forallf{\con{f}} x.\;\, \con{f}( x ) \teq 0 \}$
-and $\{ \forallf{\con{f}} x.\;\, \con{f}( x ) \teq \con{f}( x ) \}$ are admissible,
-whereas $\{ \forallf{\con{f}} x.\;\, \con{f}( x ) \teq \con{f}( x ) + 1 \}$ is not.
-\ajr{revise}
+\begin{definition}
+Given a set of function definitions $\mathcal{D} = \{ \forallf{f_1} \vec x. \varphi_1, \ldots, \forallf{f_n} \vec x. \varphi_n \}$,
+a formula $\varphi$ is \emph{closed under function expansion with respect to $\mathcal{D}$} if
+\begin{eqnarray}
+\varphi \text{ entails } \displaystyle\bigwedge\limits_{i=1}^n \{ \varphi_i[ \vec t / \vec x ] \mid f_i( \vec t ) \in \terms( \varphi ) \}
+\end{eqnarray}
+The set $\mathcal{D}$ is \emph{admissible} if whenever there exists a model for $\varphi$ that is closed under function expansion,
+there exists a model for $\varphi \wedge \forallf{f_1} \vec x. \varphi_1 \wedge \ldots \wedge \forallf{f_n} \vec x. \varphi_n$.
+\end{definition}
+
+For instance, the set $\{ \forallf{f} \vec x. f( x ) \teq 0 \}$ is admissible.
+A set of admissible function definitions can include definitions of non-terminating functions, e.g.
+$\{ \forallf{f} x^\sortint. f( x ) \teq f( x - 1 ) \}$ or even $\{ \forallf{f} x. f( x ) \teq f( x ) \}$ are admissible.
+On the other hand, the set $\{ \forallf{f} x. f( x ) \teq f( x ) + 1 \}$ is not admissible, 
+since the empty set of formulas is closed under function expansion with respect to this set,
+and there is no model satisfying $\forallf{f} x. f( x ) \teq f( x ) + 1$.
+Consider the set $\{ \forallf{f} x. f( x ) \teq f( x ), \forallf{g}. \lite( \neg f( x ) \teq 5, g( x ) \teq g( x ) + 1, g( x ) \teq f( x ) ) \}$.
+While this set has a model where $f$ and $g$ are interpreted as the constant $5$, it is not admissible
+since the formula $f( 0 ) \teq 4$ is closed under function expansion with respect to this set,
+but there exists no interpretation satisfying both $f( 0 ) \teq 4$ and $\forallf{g}. \lite( \neg f( x ) \teq 5, g( x ) \teq g( x ) + 1, g( x ) \teq f( x ) )$.
 
@@
 is to use an uninterpreted type \ct{to be defined} to abstract the set of
-\relax{relevant} tuples for each recursive function $\con{f}$ and restrict the
+\emph{relevant} tuples for each recursive function $\con{f}$ and restrict the
 quantification of the function definition for $\con{f}$ to a variable of this
@@
 In more detail, 
-suppose our signature $\Sigma$ contains a set of defined function symbols $\sfundefs{\Sigma} \subseteq \sfuns{\Sigma}$.
+say our signature $\Sigma$ contains a set of defined function symbols $\sfundefs{\Sigma} \subseteq \sfuns{\Sigma}$.
 We extend the signature of $\Sigma$ to a signature $\extendsig{\Sigma}$ containing the following additional definitions.
-For each $\con{f} : \tau_1 \times \ldots \times \tau_n \rightarrow \tau \in \sfundefs{\Sigma}$,
-the signature $\extendsig{\Sigma}$ includes
-\begin{itemize}
-\item an uninterpreted \emph{abstract type} $\fargsort{\con{f}}$;
-\item $n$ uninterpreted \emph{concretization functions} $\fargx{\con{f}}{1} : \fargsort{\con{f}} \rightarrow \tau_1$, \ldots, $\fargx{\con{f}}{n} : \fargsort{\con{f}} \rightarrow \tau_n$.
-\end{itemize}
+For each $\con{f} \in \sfundefs{\Sigma}$ of type $\tau_1 \times \ldots \times \tau_n \rightarrow \tau$, 
+the signature $\extendsig{\Sigma}$ includes:
+\begin{enumerate}
+\item[-] an uninterpreted type $\fargsort{\con{f}}$, and
+\item[-] a vector of $n$ uninterpreted functions $\farg{\con{f}}^1$, $\ldots$, $\farg{\con{f}}^n$ of type $(\fargsort{\con{f}} \rightarrow \tau_1)$, $\ldots$, $(\fargsort{\con{f}} \rightarrow \tau_n)$.
+\end{enumerate}
 %The interpretation of uninterpreted sort $\fargsort{\con{f}}$ will denote the elements (tuples) on which the function $\con{f}$ is applied.
@@
 
-The procedure $\conv$ in Figure~\ref{fig:encoding} translates the $\Sigma$-formula
-$\varphi$ into an $\extendsig{\Sigma}$-formula.
-This calls the subprocedure $\conv_0$, which takes two arguments:\ the term $t$
-to translate and a \emph{polarity} $p$ (either $\ptrue$, $\pfalse$, or
-$\pnone$). It returns a pair $( r,\, D )$, where $r$ is a term of
-type $\tau$ and $D$ is a set of formulas.
-\jb{Explain r and D here briefly.}
-%The role of $D$ is to ensure that
-%the (restricted) function definition for $\con{f}$ includes certain tuples in
-%its domain, which we explain more in the following.
-
-The translation modifies $\varphi$ in two main ways. First, it restricts the
-quantification on function definitions for $\con{f}$ to the corresponding
-uninterpreted type $\fargsort{\con{f}}$ and inserts appropriate . Second, it augments $\varphi$ with
-additional constraints of the form $\exists z\typ{\fargsort{\con{f}}}. (
-\vecfarg{\con{f}}( z ) \teq \vec t )$ which ensure that this restriction for
-the defintion $\con{f}$ covers all relevant tuples of terms, namely the ones
-for which an $\con{f}$-application exists and is relevant to the satisfiability
-of the current branch of $\varphi$.
+Given the extended signature, we run the procedure $\conv$ in Figure~\ref{fig:encoding} on $\varphi$.
+This calls the subprocedure $\conv_0$, which takes two arguments: the term $t$ to translate, and a \emph{polarity} $p$ (either $\ptrue$, $\pfalse$, or $\pnone$).
+It returns a pair of the form $( r, D )$, where $r$ is a term of type $\tau$, and $D$ is a set of formulas.
+%The role of $D$ is to ensure that the (restricted) function definition for $\con{f}$ includes certain tuples in its domain, which we explain more in the following.
+At a high level, the translation modifies $\varphi$ in two ways.
+First, it restricts the quantification on function definitions for $\con{f}$ to the corresponding uninterpreted type $\fargsort{\con{f}}$.
+Second, it augments $\varphi$ with additional constraints of the form $\exists z\typ{\fargsort{\con{f}}}. ( \vecfarg{\con{f}}( z ) \teq \vec t )$
+which ensure that this restriction for the defintion $\con{f}$ covers all relevant tuples of terms,
+namely the ones for which an $\con{f}$-application exists and is relevant to the satisfiability of the current branch of $\varphi$.
 
@@
 interprets $\fargsort{\con{f}}$ as a tuple of values of type $\tau_1^\I \times \ldots \times \tau_n^\I$,
-and each $\fargx{\con{f}}{i}$ as the $i^{th}$ projection on such tuples for each $i = 1, \ldots, n$.
+and each $\farg{\con{f}}^i$ as the $i^{th}$ projection on such tuples for each $i = 1, \ldots, n$.
 Note that $\J$ satisfies \emph{every} constraint of the form $\exists z\typ{\fargsort{\con{f}}}. ( \vecfarg{\con{f}}( z ) \teq \vec t )$,
@@
 
-Second, we show that if $\conv( \varphi )$ has a model $\I$, then $\varphi$ has a model $\J$.
+Second, we show that if $\conv( \varphi )$ is satisfied by a $\extendsig{\Sigma}$-interpretation $\I$, then $\varphi$ is satisfied by $\Sigma$-interpretation $\J$.
+Since $\varphi$ is in definitional form with respect to functions defined by $\mathcal{D}$, we know $\varphi$ is of the form $\mathcal{D} \wedge \varphi_0$.
+We first argue that $\I$ satisfies a $\Sigma$-formula $\psi$ that is closed under function expansion with respect to $\mathcal{D}$,
+and where $\psi$ entails $\varphi_0$.
 
+Since there exists a model for $\psi$ that is closed under function expansion with respect to $\mathcal{D}$, 
+then since $\mathcal{D}$ is admissible,
+by definition there exists a model for $\psi \wedge \mathcal{D}$, which entails $\mathcal{D} \wedge \varphi_0$, which is $\varphi$.
 \end{proof}
+
+Let us revisit the translated formula~(\ref{eq:ex-after}) from Example~\ref{ex:translation}.
+This formula is unsatisfiable if and only if formula~(\ref{eq:ex-before}) is unsatisfiable since $\fargsort{f}$ can be interpreted as $\sortint$,
+and $\farg{f}$ can be interpreted as the identity function.
+Conversely, we claim formula~(\ref{eq:ex-after}) is satisfiable if and only if formula~(\ref{eq:ex-before}) is satisfiable,
+noting that the set $\{ \forall_{s} x\typ{\sortint}. ite( x \leq 0, s(x) \teq 0, s( x ) \teq x + s( x-1 ) ) \}$ is admissble.
+As stated in the proof of the previous theorem, 
+the reason is that any model of formula~(\ref{eq:ex-after}) must satisfy a formula that is closed under function expansion,
+hence implying that formula~(\ref{eq:ex-before}) has a model.
+In more detail, let $\varphi_s$ be the formula $\forall_{s} x. ite( x \leq 0, s(x) \teq 0, s( x ) \teq x + s( x-1 ) )$.
+The rightmost conjunct 
+$\exists z\typ{\fargsort{s}}. \farg{s}( z ) \teq c$ ensures 
+that any model of formula~(\ref{eq:ex-after}) satisfying $s( c ) > 100$ also satisfies $\varphi_s[c/x]$.
+Similarly, the nested existential $\exists z. \farg{s}( z ) \teq \farg{s}( y )-1 )$
+ensures that this constraint is enforced recursively.
+%Given a model $\I$ of formula~(\ref{eq:ex-after}) that interprets $c$ as $14$, it is easy to see that $\I$ also satisfies the formula:
+\begin{eqnarray}
+\begin{split}
+s( c ) > 100 \wedge & \displaystyle\bigwedge\limits_{i=0}^{13} ( \neg c-i \leq 0 \wedge s( c-i ) \teq i + s( c-i-1 ) ) \\
+                    & \wedge c-14 \leq 0 \wedge s( c-14 ) \teq 0
+\end{split}
+\end{eqnarray}
 
@@
 then it will replace its list of known assertions based on the conversion $\conv$ before checking for satisfiability.
-Note that this option is model-sound \ct{define?} only if the functions provided in the input are admissible.
+%If the functions provided in the input are admissible, 
+%then a ``satisfiable" response from the solver implies that a model exists for the original formula.
+Accordingly, the solver will output an approximation of the interpretation for recursive function definitions,
+that is, it may output the interpretation~(\ref{eq:approx-interp}) from Section~\ref{sec:encoding} for $s$.
+
+Notice the option ``--fmf-fun" effectively instructs the SMT solver to assume that the functions provided by the user are admissible.
+If this option is used for a set of functions definitions that is \emph{not} admissible, then the solver is not model-sound, that is, 
+it may return ``satisfiable" for unsatisfiable inputs.
+As such,
+this option targets users who wish to determine the satisfiability of inputs that involving function definitions that are known to be consistent,
+whereas it does not target users who wish to check the consistency of a set of function definitions.
 

### 2015-05-05 tex/main.tex
started making constraints type correct

--- 2015-05-05 tex/main.tex
@@
 
+%\newcommand\pneg{\neg}
+\newcommand\pneg{-}
+
 \newcommand{\con}[1]{\mathsf{#1}}
@@
 \def\Sigma{\mathrm{\oldSigma}}
+
+\def\Chi{\mathrm{X}}
 
@@
 
-\newcommand{\boolop}{\oplus}
+\newcommand{\boolop}{\con{b}}
 \newcommand{\forallf}[1]{\forall_{\!#1\:}}
@@
 
-\begin{figure}[t]
-\begin{enumerate}
-%\begin{framed}
-\item[\ ] 
-$\conv( \varphi )$ : 
- \begin{itemize}
-   \item[] $\mathsf{let}$ $( \psi, D ) = \conv( \varphi, \ptrue )$;
-   %\item[] $\mathsf{assert}( D = \emptyset )$;
-   \item[] $\mathsf{return}$ $\psi$
-  \end{itemize}
-\end{enumerate}
-\begin{enumerate}
-%\begin{framed}
-\item[\ ] 
-$\conv_0( t\typ{\tau}, p )$ : 
- \begin{itemize}
-   \item[] $\mathsf{if}$ $\tau \equiv \Bool$ and $t \equiv t_1 \boolop \ldots \boolop t_n$ :
-    \begin{itemize}
-      \item[] $\mathsf{let}$ $( r_i, D_i ) = \conv_0( t_i, \pol( \boolop, i, p ) )$ for $i = 1, \ldots, n$;
-      \item[] $\mathsf{if}$ $p = \ptrue$
-      \begin{itemize}
-        \item[] $\mathsf{return}$ $( ( r_1 \boolop \ldots \boolop r_n ) \wedge D_1 \wedge \ldots \wedge D_n, \emptyset )$
-      \end{itemize}
-      \item[] $\mathsf{else}$ $\mathsf{if}$ $p = \pfalse$
-      \begin{itemize}
-        \item[] $\mathsf{return}$ $( ( r_1 \boolop \ldots \boolop r_n ) \vee \neg D_1 \vee \ldots \vee \neg D_n, \emptyset )$
-      \end{itemize}
-      \item[] $\mathsf{else}$
-      \begin{itemize}
-        \item[] $\mathsf{return}$ $( r_1 \boolop \ldots \boolop r_n, D_1 \cup \ldots \cup D_n )$
-      \end{itemize}
-    \end{itemize}
-  \item[] $\mathsf{else}$ $\mathsf{if}$ $t \equiv \forallf{\con{f}} \vec x.\;\, t_1$
-    \begin{itemize}
-      \item[] $\mathsf{let} ( r_1, D_1 ) = \conv_0( t_1, p )$;
-      \item[] $\mathsf{if}$ $\con{f} \in \sfundefs{\Sigma}$
-      \begin{itemize}
-        %\item[] $\mathsf{assert}( D_1 = \emptyset )$;
-        \item[] $\mathsf{return}$ $( \forall y\typ{\fargsort{\con{f}}}.\;\, ( r_1 [ \vecfarg{\con{f}}( y ) / \vec x ] ), \emptyset )$
-      \end{itemize}
-      \item[] else
-      \begin{itemize}
-        \item[] $\mathsf{return}$ $( \forall \vec x.\; r_1, \forall \vec x.\; D_1 )$
-      \end{itemize}
-    \end{itemize}
-   \item[] $\mathsf{else}$
-   \begin{itemize}
-     \item[] $\mathsf{return}$ $( t, \{ \exists z\typ{\fargsort{\con{f}}}. ( \vecfarg{f}( z ) \teq \vec t ) \mid f( \vec t ) \in \terms( t ), f \in \sfundefs{\Sigma} \} )$
-   \end{itemize}
- \end{itemize}
-\end{enumerate}
-\vspace{-2ex}
-\caption{A translation procedure $\conv$ for a $\Sigma$-formula that is in definitional form with respect to a set of admissible function definitions $\sfundefs{\Sigma}$.
-In the subprocedure $\conv_0$, the procedure $\pol( \boolop, i, p )$ returns the polarity of the $i^{th}$ child of a $\boolop$-application having polarity $p$,
-where $\boolop$ is an interpreted predicate.
-}
-\label{fig:encoding}
-\end{figure}
-
 In more detail, 
@@
 
-The procedure $\conv$ in Figure~\ref{fig:encoding} translates the $\Sigma$-formula
-$\varphi$ into an $\extendsig{\Sigma}$-formula.
-This calls the subprocedure $\conv_0$, which takes two arguments:\ the term $t$
+The procedure $\conv$ translates the $\Sigma$-formula
+$\varphi$ into an $\extendsig{\Sigma}$-formula. It relies
+on the subprocedure $\conv_0$, which takes two arguments:\ the term $t$
 to translate and a \emph{polarity} $p$, which is either $\ptrue$, $\pfalse$, or
-$\pnone$. It returns a pair $( t'\negvthinspace,\, D )$, where $t'$ is a term of
-type $\tau$ and $D$ is a set of formulas encoding additional constraints.
+$\pnone$. It returns a pair $( t'\negvthinspace,\, \Chi )$, where $t'$ is a term of
+type $\tau$ and $\Chi$ is a set of formulas encoding additional constraints.
+These constraints are necessary if the polarity is $\pnone$; otherwise, $\Chi$
+is always the empty set.
+
+\jb{Merge para above with para below}
+
+A translation procedure $\conv$ for a $\Sigma$-formula that is in definitional
+form with respect to a set of admissible function definitions
+$\sfundefs{\Sigma}$. In the auxiliary procedure $\conv_0$, the procedure $\pol(
+\boolop, i, p )$ returns the polarity of the $i${th} argument of a
+$\boolop$ application having polarity $p$, where $\boolop$ is an interpreted
+predicate.
+
+\begin{enumerate}
+%\begin{framed}
+\item[\ ] 
+$\conv_0( t\typ{\tau},\, p )$ : 
+ \begin{itemize}
+   \item[] $\mathsf{if}$ $\tau = \Bool$ and $t = \boolop(t_1,\ldots,t_n)$:
+    \begin{itemize}
+      \item[] $\mathsf{let}$ $( r_i,\, \Chi_i ) = \conv_0( t_i,\, \pol( \boolop,\, i,\, p ) )$ for $i = 1, \ldots, n$;
+      \item[] $\mathsf{let}$ $\Chi = \Chi_1 \mathrel\cup \cdots \mathrel\cup \Chi_n$;
+      \item[] $\mathsf{let}$ $\chi = \bigwedge \Chi$;
+      \item[] $\mathsf{if}$ $p = \ptrue$
+      \begin{itemize}
+        \item[] $\mathsf{return}$ $( ( r_1 \boolop \ldots \boolop r_n ) \wedge \chi,\, \emptyset )$
+      \end{itemize}
+      \item[] $\mathsf{else}$ $\mathsf{if}$ $p = \pfalse$
+      \begin{itemize}
+        \item[] $\mathsf{return}$ $( ( r_1 \boolop \ldots \boolop r_n ) \vee \neg \chi,\, \emptyset )$
+      \end{itemize}
+      \item[] $\mathsf{else}$
+      \begin{itemize}
+        \item[] $\mathsf{return}$ $( r_1 \boolop \ldots \boolop r_n,\, \Chi )$
+      \end{itemize}
+    \end{itemize}
+  \item[] $\mathsf{else}$ $\mathsf{if}$ $t \equiv \forallf{\con{f}} \vec x.\;\, t_1$
+    \begin{itemize}
+      \item[] $\mathsf{let}$ $( r_1,\, D_1 ) = \conv_0( t_1,\, p )$;
+      \item[] $\mathsf{if}$ $\con{f} \in \sfundefs{\Sigma}$
+      \begin{itemize}
+        %\item[] $\mathsf{assert}( D_1 = \emptyset )$;
+        \item[] $\mathsf{return}$ $( \forall y\typ{\fargsort{\con{f}}}.\;\, ( r_1 [ \vecfarg{\con{f}}( y ) / \vec x ] ),\, \emptyset )$
+      \end{itemize}
+      \item[] else
+      \begin{itemize}
+        \item[] $\mathsf{return}$ $( \forall \vec x.\; r_1,\, \forall \vec x.\; D_1 )$
+      \end{itemize}
+    \end{itemize}
+   \item[] $\mathsf{else}$
+   \begin{itemize}
+     \item[] $\mathsf{return}$ $( t,\, \{ \exists z\typ{\fargsort{\con{f}}}. ( \vecfarg{f}( z ) \teq \vec t ) \mid f( \vec t ) \in \terms( t ),\, f \in \sfundefs{\Sigma} \} )$
+   \end{itemize}
+ \end{itemize}
+\end{enumerate}
+\begin{enumerate}
+%\begin{framed}
+\item[\ ] 
+$\conv( \varphi )$ : 
+ \begin{itemize}
+   \item[] $\mathsf{let}$ $( \psi,\, \_ ) = \conv_0( \varphi,\, \ptrue )$;
+   %\item[] $\mathsf{assert}( D = \emptyset )$;
+   \item[] $\mathsf{return}$ $\psi$
+  \end{itemize}
+\end{enumerate}
+
 %The role of $D$ is to ensure that
@@
 \begin{cases}
-\pnone & \text{if $\boolop = {\teq}$ or $(\boolop, i) = (\lite, 0)$} \\
-\neg p & \text{if $\boolop = {\neg}$} \\
-p & \text{otherwise}
+p & \text{if either $\boolop \in \{{\land}, {\lor}\}$ or $\boolop = \lite$ and $i \in \{2, 3\}$} \\[-\jot]
+\pneg p & \text{if $\boolop = {\neg}$} \\[-\jot]
+\pnone & \text{otherwise}
 \end{cases}
 \end{equation*}
-The negation $\neg p$ of polarity $p$ is $\pfalse$ if $p$ is $\ptrue$, $\ptrue$
+The negation $\pneg p$ of a polarity $p$ is $\pfalse$ if $p$ is $\ptrue$, $\ptrue$
 if $p$ is $\pfalse$, and $\pnone$ if $p$ is $\pnone$.
@@
 If $t$ has positive polarity, we construct a conjunction of the result of the
-recursive calls with formulas $D_1, \ldots, D_n$. Similarly, if $t$ has
-negative polarity, we construct a disjunction of the result of the recursive
-calls with the negation of formulas $D_1, \ldots, D_n$. It $t$ has no polarity,
-then we return the union of $D_1 \cup \ldots \cup D_n$.
+recursive calls with the formulas in $D = D_1 \mathrel\cup \cdots \mathrel\cup
+D_n$. Similarly, if $t$ has negative polarity, we construct a disjunction of
+the result of the recursive calls with the negation of formulas $D_1, \ldots,
+D_n$. It $t$ has no polarity, then we return the union of $D_1 \cup \ldots \cup
+D_n$.
 

### 2015-05-05 tex/main.tex
tuned algo

@@
     \begin{itemize}
-      \item[] $\mathsf{let}$ $( r_i,\, \Chi_i ) = \conv_0( t_i,\, \pol( \boolop,\, i,\, p ) )$ for $i = 1, \ldots, n$;
+      \item[] $\mathsf{let}$ $( t'_i,\, \Chi_i ) = \conv_0( t_i,\, \pol( \boolop,\, i,\, p ) )$ for $i = 1, \ldots, n$;
       \item[] $\mathsf{let}$ $\Chi = \Chi_1 \mathrel\cup \cdots \mathrel\cup \Chi_n$;
@@
       \begin{itemize}
-        \item[] $\mathsf{return}$ $( ( r_1 \boolop \ldots \boolop r_n ) \wedge \chi,\, \emptyset )$
+        \item[] $\mathsf{return}$ $( \boolop(t'_1, \ldots, t'_n) ) \wedge \chi,\, \emptyset )$
       \end{itemize}
@@
       \begin{itemize}
-        \item[] $\mathsf{return}$ $( ( r_1 \boolop \ldots \boolop r_n ) \vee \neg \chi,\, \emptyset )$
+        \item[] $\mathsf{return}$ $( \boolop(t'_1, \ldots, t'_n) \vee \neg \chi,\, \emptyset )$
       \end{itemize}
@@
       \begin{itemize}
-        \item[] $\mathsf{return}$ $( r_1 \boolop \ldots \boolop r_n,\, \Chi )$
+        \item[] $\mathsf{return}$ $( \boolop(t'_1, \ldots, t'_n),\, \Chi )$
       \end{itemize}
@@
     \begin{itemize}
-      \item[] $\mathsf{let}$ $( r_1,\, D_1 ) = \conv_0( t_1,\, p )$;
+      \item[] $\mathsf{let}$ $( t'_1,\, D_1 ) = \conv_0( t_1,\, p )$;
       \item[] $\mathsf{if}$ $\con{f} \in \sfundefs{\Sigma}$
@@
         %\item[] $\mathsf{assert}( D_1 = \emptyset )$;
-        \item[] $\mathsf{return}$ $( \forall y\typ{\fargsort{\con{f}}}.\;\, ( r_1 [ \vecfarg{\con{f}}( y ) / \vec x ] ),\, \emptyset )$
+        \item[] $\mathsf{return}$ $( \forall y\typ{\fargsort{\con{f}}}.\;\, ( t'_1 [ \vecfarg{\con{f}}( y ) / \vec x ] ),\, \emptyset )$
       \end{itemize}
@@
       \begin{itemize}
-        \item[] $\mathsf{return}$ $( \forall \vec x.\; r_1,\, \forall \vec x.\; D_1 )$
+        \item[] $\mathsf{return}$ $( \forall \vec x.\; t'_1,\, \forall \vec x.\; D_1 )$
       \end{itemize}
@@
 a quantified formula over a single variable $y$ of type $\fargsort{\con{f}}$,
-and replace all occurrences of $\vec x$ in $r_1$ with $\vecfarg{\con{f}}( y )$.
+and replace all occurrences of $\vec x$ in $t'_1$ with $\vecfarg{\con{f}}( y )$.
 In this case, since function definitions are a top-level conjunct, by case
@@
 case that $t$ is not a function definition, the resulting quantified formula is
-reconstructed with the term $r_1$ returned by the recursive call, and a
+reconstructed with the term $t'_1$ returned by the recursive call, and a
 quantified prefix is appended to the formula $D_1$ returned by the recursive

### 2015-05-05 tex/main.tex
more

@@
 to translate and a \emph{polarity} $p$, which is either $\ptrue$, $\pfalse$, or
-$\pnone$. It returns a pair $( t'\negvthinspace,\, \Chi )$, where $t'$ is a term of
-type $\tau$ and $\Chi$ is a set of formulas encoding additional constraints.
-These constraints are necessary if the polarity is $\pnone$; otherwise, $\Chi$
-is always the empty set.
+$\pnone$. It returns a pair $( t'\negvthinspace,\, \chi )$, where $t'$ is a term of
+type $\tau$ and $\chi$ is an additional constraint.
+This constraint is necessary if the polarity is $\pnone$; otherwise, any necessary
+constraint is encoded directly in $t'$ and $\chi$ is simply $\true$.
 
@@
 
+\jb{what happens with quantifiers other than $\forallf{f}$?}
+\jb{transform ``procedure'' into equation / functional program?}
+
 \begin{enumerate}
@@
  \begin{itemize}
-   \item[] $\mathsf{if}$ $\tau = \Bool$ and $t = \boolop(t_1,\ldots,t_n)$:
+   \item[] $\mathsf{if}$ $\tau = \Bool$ and $t = \boolop(t_1,\ldots,t_n)$ $\mathsf{then}$
     \begin{itemize}
-      \item[] $\mathsf{let}$ $( t'_i,\, \Chi_i ) = \conv_0( t_i,\, \pol( \boolop,\, i,\, p ) )$ for $i = 1, \ldots, n$;
-      \item[] $\mathsf{let}$ $\Chi = \Chi_1 \mathrel\cup \cdots \mathrel\cup \Chi_n$;
-      \item[] $\mathsf{let}$ $\chi = \bigwedge \Chi$;
-      \item[] $\mathsf{if}$ $p = \ptrue$
+      \item[] $\mathsf{let}$ $( t'_i,\, \chi_i ) = \conv_0( t_i,\, \pol( \boolop,\, i,\, p ) )$ for $i = 1, \ldots, n$;
+      \item[] $\mathsf{let}$ $\chi = \chi_1 \mathrel\wedge \cdots \mathrel\wedge \chi_n$;
+      \item[] $\mathsf{if}$ $p = \ptrue$ $\mathsf{then}$
       \begin{itemize}
-        \item[] $\mathsf{return}$ $( \boolop(t'_1, \ldots, t'_n) ) \wedge \chi,\, \emptyset )$
+        \item[] $\mathsf{return}$ $( \boolop(t'_1, \ldots, t'_n) ) \wedge \chi,\, \true )$
       \end{itemize}
-      \item[] $\mathsf{else}$ $\mathsf{if}$ $p = \pfalse$
+      \item[] $\mathsf{else}$ $\mathsf{if}$ $p = \pfalse$ $\mathsf{then}$
       \begin{itemize}
-        \item[] $\mathsf{return}$ $( \boolop(t'_1, \ldots, t'_n) \vee \neg \chi,\, \emptyset )$
+        \item[] $\mathsf{return}$ $( \boolop(t'_1, \ldots, t'_n) \vee \neg \chi,\, \true )$
       \end{itemize}
@@
       \begin{itemize}
-        \item[] $\mathsf{return}$ $( \boolop(t'_1, \ldots, t'_n),\, \Chi )$
+        \item[] $\mathsf{return}$ $( \boolop(t'_1, \ldots, t'_n),\, \chi )$
       \end{itemize}
     \end{itemize}
-  \item[] $\mathsf{else}$ $\mathsf{if}$ $t \equiv \forallf{\con{f}} \vec x.\;\, t_1$
+  \item[] $\mathsf{else}$ $\mathsf{if}$ $t = \forallf{\con{f}} \vec x.\;\, u$ $\mathsf{then}$
     \begin{itemize}
-      \item[] $\mathsf{let}$ $( t'_1,\, D_1 ) = \conv_0( t_1,\, p )$;
-      \item[] $\mathsf{if}$ $\con{f} \in \sfundefs{\Sigma}$
+      \item[] $\mathsf{let}$ $( u',\, \chi ) = \conv_0( u,\, p )$;
+      \item[] $\mathsf{if}$ $\con{f} \in \sfundefs{\Sigma}$ $\mathsf{then}$
       \begin{itemize}
-        %\item[] $\mathsf{assert}( D_1 = \emptyset )$;
-        \item[] $\mathsf{return}$ $( \forall y\typ{\fargsort{\con{f}}}.\;\, ( t'_1 [ \vecfarg{\con{f}}( y ) / \vec x ] ),\, \emptyset )$
+        %\item[] $\mathsf{assert}( \chi = \true )$;
+        \item[] $\mathsf{return}$ $( \forall y\typ{\fargsort{\con{f}}}.\; u' [ \vecfarg{\con{f}}( y ) / \vec x ],\, \true )$
+\jb{looks fishy -- I need to look at it again}
       \end{itemize}
-      \item[] else
+      \item[] $\mathsf{else}$
       \begin{itemize}
-        \item[] $\mathsf{return}$ $( \forall \vec x.\; t'_1,\, \forall \vec x.\; D_1 )$
+        \item[] $\mathsf{return}$ $( \forall \vec x.\; u',\, \forall \vec x.\; \chi )$
       \end{itemize}
@@
    \begin{itemize}
-     \item[] $\mathsf{return}$ $( t,\, \{ \exists z\typ{\fargsort{\con{f}}}. ( \vecfarg{f}( z ) \teq \vec t ) \mid f( \vec t ) \in \terms( t ),\, f \in \sfundefs{\Sigma} \} )$
+     \item[] $\mathsf{return}$ $( t,\, \bigwedge\, \{ \exists z : \fargsort{\con{f}}.\; \vecfarg{\con{f}}( z ) \teq \vec t \mid \con{f}( \vec t ) \in \terms( t ),\, \con{f} \in \sfundefs{\Sigma} \} )$
    \end{itemize}

### 2015-05-05 tex/main.tex
simplified logic
continued tuning the algo
more

@@
 \noindent\keywordname\enspace\ignorespaces#1}
+
+%\newcommand\return{$\mathsf{return}$\xspace}
+\newcommand\return{}
+
+%\newcommand\bigtuple[1]{$(${#1}$)$}
+\newcommand\bigtuple[1]{$\bigl<${#1}$\bigr>$}
 
@@
 
-The procedure $\conv$ translates the $\Sigma$-formula
+The function $\conv$ translates the $\Sigma$-formula
 $\varphi$ into an $\extendsig{\Sigma}$-formula. It relies
-on the subprocedure $\conv_0$, which takes two arguments:\ the term $t$
+on the auxiliary function $\conv_0$, which takes two arguments:\ the term $t$
 to translate and a \emph{polarity} $p$, which is either $\ptrue$, $\pfalse$, or
@@
 
-A translation procedure $\conv$ for a $\Sigma$-formula that is in definitional
+A translation function $\conv$ for a $\Sigma$-formula that is in definitional
 form with respect to a set of admissible function definitions
-$\sfundefs{\Sigma}$. In the auxiliary procedure $\conv_0$, the procedure $\pol(
+$\sfundefs{\Sigma}$. In the auxiliary function $\conv_0$, the function $\pol(
 \boolop, i, p )$ returns the polarity of the $i${th} argument of a
@@
 
-\jb{what happens with quantifiers other than $\forallf{f}$?}
-\jb{transform ``procedure'' into equation / functional program?}
-
 \begin{enumerate}
@@
 \item[\ ] 
-$\conv_0( t\typ{\tau},\, p )$ : 
+$\conv_0( t\typ{\tau},\, p )$ $=$
  \begin{itemize}
@@
     \begin{itemize}
-      \item[] $\mathsf{let}$ $( t'_i,\, \chi_i ) = \conv_0( t_i,\, \pol( \boolop,\, i,\, p ) )$ for $i = 1, \ldots, n$;
-      \item[] $\mathsf{let}$ $\chi = \chi_1 \mathrel\wedge \cdots \mathrel\wedge \chi_n$;
+      \item[] $\mathsf{let}$ $( t'_i,\, \chi_i ) = \conv_0( t_i,\, \pol( \boolop,\, i,\, p ) )$ $\mathsf{for}$ $i = 1, \ldots, n$ $\mathsf{in}$
+      \item[] $\mathsf{let}$ $\chi = \chi_1 \mathrel\wedge \cdots \mathrel\wedge \chi_n$ $\mathsf{in}$
       \item[] $\mathsf{if}$ $p = \ptrue$ $\mathsf{then}$
       \begin{itemize}
-        \item[] $\mathsf{return}$ $( \boolop(t'_1, \ldots, t'_n) ) \wedge \chi,\, \true )$
+        \item[] \return \bigtuple{$\boolop(t'_1, \ldots, t'_n) ) \wedge \chi,\, \true$}
       \end{itemize}
@@
       \begin{itemize}
-        \item[] $\mathsf{return}$ $( \boolop(t'_1, \ldots, t'_n) \vee \neg \chi,\, \true )$
+        \item[] \return \bigtuple{$\boolop(t'_1, \ldots, t'_n) \vee \neg \chi,\, \true$}
       \end{itemize}
@@
       \begin{itemize}
-        \item[] $\mathsf{return}$ $( \boolop(t'_1, \ldots, t'_n),\, \chi )$
+        \item[] \return \bigtuple{$\boolop(t'_1, \ldots, t'_n),\, \chi$}
       \end{itemize}
@@
     \begin{itemize}
-      \item[] $\mathsf{let}$ $( u',\, \chi ) = \conv_0( u,\, p )$;
-      \item[] $\mathsf{if}$ $\con{f} \in \sfundefs{\Sigma}$ $\mathsf{then}$
+      \item[] $\mathsf{let}$ $( u',\, \chi ) = \conv_0( u,\, p )$ $\mathsf{in}$
+      %\item[] $\mathsf{assert}( \chi = \true )$;
+      \item[] \return \bigtuple{$\forall y\typ{\fargsort{\con{f}}}.\; u' [ \vecfarg{\con{f}}( y ) / \vec x ],\, \true$}
+    \end{itemize}
+  \item[] $\mathsf{else}$ $\mathsf{if}$ $t = \forall \vec x.\;\, u$ $\mathsf{then}$
     \begin{itemize}
-        %\item[] $\mathsf{assert}( \chi = \true )$;
-        \item[] $\mathsf{return}$ $( \forall y\typ{\fargsort{\con{f}}}.\; u' [ \vecfarg{\con{f}}( y ) / \vec x ],\, \true )$
-\jb{looks fishy -- I need to look at it again}
+      \item[] $\mathsf{let}$ $( u',\, \chi ) = \conv_0( u,\, p )$ $\mathsf{in}$
+      \item[] \return \bigtuple{$\forall \vec x.\; u',\, \forall \vec x.\; \chi$}
     \end{itemize}
@@
    \begin{itemize}
-        \item[] $\mathsf{return}$ $( \forall \vec x.\; u',\, \forall \vec x.\; \chi )$
-      \end{itemize}
-    \end{itemize}
-   \item[] $\mathsf{else}$
-   \begin{itemize}
-     \item[] $\mathsf{return}$ $( t,\, \bigwedge\, \{ \exists z : \fargsort{\con{f}}.\; \vecfarg{\con{f}}( z ) \teq \vec t \mid \con{f}( \vec t ) \in \terms( t ),\, \con{f} \in \sfundefs{\Sigma} \} )$
+     \item[] \return \bigtuple{$t,\, \bigwedge\, \{ \exists z : \fargsort{\con{f}}.\; \vecfarg{\con{f}}( z ) \teq \vec t \mid \con{f}( \vec t ) \in \terms( t ),\, \con{f} \in \sfundefs{\Sigma} \}$}
    \end{itemize}
@@
 \item[\ ] 
-$\conv( \varphi )$ : 
+$\conv( \varphi )$ $=$
  \begin{itemize}
-   \item[] $\mathsf{let}$ $( \psi,\, \_ ) = \conv_0( \varphi,\, \ptrue )$;
+   \item[] $\mathsf{let}$ $( \psi,\, \_ ) = \conv_0( \varphi,\, \ptrue )$ $\mathsf{in}$
    %\item[] $\mathsf{assert}( D = \emptyset )$;
-   \item[] $\mathsf{return}$ $\psi$
+   \item[] \return $\psi$
   \end{itemize}

### 2015-05-05 tex/main.tex
Example giving intuition of proof.

@@
 
-\ajr{add invariants, proofs}
-
-%\begin{lemma}
-%Let $\M$ be a model satisfying $\conv( \varphi )$, and
-%%and let $\conv( \varphi )^\ast$ be the quantifier-free formula obtained by 
-%let $M$ be a set of literals such that $\M \models M$ and $M \models_p \conv( \varphi )$.
-%For each assignment $\sigma$ over $\vars{ M }$,
-%each $\con{f}( \vec t ) \in \terms( M \sigma )$ where $\con{f} \in \sfundefs{\Sigma}$, 
-%there exists a formula of the form $\exists z\typ{\fargsort{\con{f}}}. ( \vecfarg{\con{f}}( z ) \teq \vec t )$ in $M \sigma$.
-%\end{lemma}
-
 \begin{theorem}
-If $\varphi$ is a $\Sigma$-formula that is in definitional form with respect to functions $\sfundefs{\Sigma}$,
+If $\varphi$ is a $\Sigma$-formula that is in definitional form with respect to $\sfundefs{\Sigma}$,
 and the set of function definitions $\mathcal{D}$ corresponding to functions in $\sfundefs{\Sigma}$ is admissible,
@@
 Second, we show that if $\conv( \varphi )$ is satisfied by a $\extendsig{\Sigma}$-interpretation $\I$, then $\varphi$ is satisfied by $\Sigma$-interpretation $\J$.
-Since $\varphi$ is in definitional form with respect to functions defined by $\mathcal{D}$, we know $\varphi$ is of the form $\mathcal{D} \wedge \varphi_0$.
+Since $\varphi$ is in definitional form with respect to the functions defined by $\mathcal{D}$, we know $\varphi$ is of the form $\mathcal{D} \wedge \varphi_0$.
 We first argue that $\I$ satisfies a $\Sigma$-formula $\psi$ that is closed under function expansion with respect to $\mathcal{D}$,
@@
 ensures that this constraint is enforced recursively.
-%Given a model $\I$ of formula~(\ref{eq:ex-after}) that interprets $c$ as $14$, it is easy to see that $\I$ also satisfies the formula:
-\begin{eqnarray}
+Given a model $\I$ of formula~(\ref{eq:ex-after}) that interprets $c$ as $14$, 
+it is easy to see that $\I$ must satisfy the formula:
+\begin{eqnarray} \label{eq:fun-closure}
 \begin{split}
-s( c ) > 100 \wedge & \displaystyle\bigwedge\limits_{i=0}^{13} ( \neg c-i \leq 0 \wedge s( c-i ) \teq i + s( c-i-1 ) ) \\
+\psi := s( c ) > 100 \wedge & \displaystyle\bigwedge\limits_{i=0}^{13} ( \neg (c-i \leq 0) \wedge s( c-i ) \teq i + s( c-i-1 ) ) \\
                             & \wedge c-14 \leq 0 \wedge s( c-14 ) \teq 0
@@
 \end{eqnarray}
+Notice that $\psi$ is closed under function expansion, 
+since it contains the $s$-applications that are equivalent to $s( c-i )$ for $i = 0, \ldots, 14$, and
+entails $\varphi_s[c-i/x]$ for $i = 0, \ldots, 14$.
+Since $\varphi_s$ is an admissible definition,
+this implies there exists a model for $\psi \wedge \varphi_s$,
+which implies there exists a model for formula~(\ref{eq:ex-before}).
 

### 2015-05-05 tex/main.tex
edits

@@
 
-%%% I added (Co) -- hope you like it --JB
-\title{Model Finding for (Co)recursive Functions in SMT\thanks{%
+\title{Model Finding for Recursive Functions in SMT\thanks{%
 This research is partially supported by the Inria technological development
@@
 \author {Andrew Reynolds\inst{1} \and Jasmin Christian Blanchette\inst{2,3} \and Cesare Tinelli \inst{4}}
-\authorrunning {A. Reynolds \and J. C. Blanchette \and C. Tinelli}
+\authorrunning {A. Reynolds, J. C. Blanchette, C. Tinelli}
 \institute{
@@
 Conversely, we claim formula~(\ref{eq:ex-after}) is satisfiable if and only if formula~(\ref{eq:ex-before}) is satisfiable,
-noting that the set $\{ \forall_{s} x\typ{\sortint}. ite( x \leq 0, s(x) \teq 0, s( x ) \teq x + s( x-1 ) ) \}$ is admissble.
+noting that the set $\{ \forall_{s} x\typ{\sortint}.\;\, \lite( x \leq 0, s(x) \teq 0, s( x ) \teq x + s( x-1 ) ) \}$ is admissble.
 As stated in the proof of the previous theorem, 
@@
 
-\section{Front-End Support for Recursive Function Definitions in CVC4}
+\section{Front-End Support %for Recursive Function Definitions
+  in CVC4}
 \label{sec:front-end}

### 2015-05-05 tex/main.tex
Formalize entails.
Minor

@@
 otherwise, it is \emph{$T$-unsatisfiable}.
+A formula $\varphi$ entails $\psi$, written $\varphi \models_T \psi$,
+if all interpretations in $\Mo$ that satisfy $\varphi$ also satisfy $\psi$.
 Formulas $\varphi$ and $\psi$ are \emph{equisatisfiable} if
@@
 \begin{eqnarray}
-\varphi \text{ entails } \displaystyle\bigwedge\limits_{i=1}^n \{ \varphi_i[ \vec t / \vec x ] \mid f_i( \vec t ) \in \terms( \varphi ) \}
+\varphi \models_T \displaystyle\bigwedge\limits_{i=1}^n \{ \varphi_i[ \vec t / \vec x ] \mid f_i( \vec t ) \in \terms( \varphi ) \}
 \end{eqnarray}
@@
 that any model of formula~(\ref{eq:ex-after}) satisfying $s( c ) > 100$ also satisfies $\varphi_s[c/x]$.
-Similarly, the nested existential $\exists z. \farg{s}( z ) \teq \farg{s}( y )-1 )$
+Similarly, the nested existential $\exists z. \farg{s}( z ) \teq \farg{s}( y )-1$
 ensures that this constraint is enforced recursively.

### 2015-05-05 tex/main.tex
more edits

--- 2015-05-05 tex/main.tex
@@
 
-If $t$ has positive polarity, we construct a conjunction of the result of the
-recursive calls with the formulas in $D = D_1 \mathrel\cup \cdots \mathrel\cup
-D_n$. Similarly, if $t$ has negative polarity, we construct a disjunction of
-the result of the recursive calls with the negation of formulas $D_1, \ldots,
-D_n$. It $t$ has no polarity, then we return the union of $D_1 \cup \ldots \cup
-D_n$.
+If $t$ has positive polarity, $\conv_0$ adds the constraint $\chi$ derived from
+the subterms as a conjunction. Dually, if $t$ has negative polarity, the
+function constructs a disjunction with the negated constraint. It $t$ has no
+known polarity, the constrained $\chi$ is simply returned to the caller.
 
-If $t$ is a quantified formula $\forallf{\con{f}} \vec x.\; t_1$, it
-recursively calls $\conv$ on its body with the same polarity. If $t$ is a
-function definition of a function in $\sfundefs{\Sigma}$, we instead construct
-a quantified formula over a single variable $y$ of type $\fargsort{\con{f}}$,
-and replace all occurrences of $\vec x$ in $t'_1$ with $\vecfarg{\con{f}}( y )$.
-In this case, since function definitions are a top-level conjunct, by case
-analysis on the return values of $\conv$, we know that $D_1$ is empty. In the
-case that $t$ is not a function definition, the resulting quantified formula is
-reconstructed with the term $t'_1$ returned by the recursive call, and a
-quantified prefix is appended to the formula $D_1$ returned by the recursive
-call.
+If $t$ is a quantified formula, $\conv_0$ recursively calls itself on the body
+with the same polarity. If $t$ is a function definition of a function in
+$\sfundefs{\Sigma}$, $\conv_0$ constructs a quantified formula over a single variable
+$y$ of type $\fargsort{\con{f}}$ and replaces all occurrences of $\vec x$ in
+$t'_1$ with $\vecfarg{\con{f}}( y )$.Since function definitions
+are a top-level conjunct, %by case analysis on the return values of $\conv$,
+we know that $\chi$ is simply $\true$.
+If $t$ is not a function definition, the resulting
+quantified formula is reconstructed with the body $u'$ returned by the
+recursive call, and a quantifier is prefixed to the constraint $\chi$
+returned by the call.
 
-Otherwise, $t$ is either an application of an uninterpreted predicate, or a
-term of a type other than $\Bool$. In this case, the second component of our
-return value contains a constraint of the form $\exists
-z\typ{\fargsort{\con{f}}}. ( \vecfarg{\con{f}}( z ) \teq \vec t )$ for each
-subterm $\con{f}( \vec t )$ of $t$ such that $\con{f}$ is a function in
-$\sfundefs{\Sigma}$. When such constraints are asserted positively, this
-indicates that the tuple of uninterpreted functions $\vecfarg{\con{f}}$ must
-include $\vec t$ in its range.
+Otherwise, $t$ is either an application of an uninterpreted predicate or a term
+of a type other than $\Bool$. The returned constraint is a conjunction of
+formulas of the form $\exists z\typ{\fargsort{\con{f}}}.\;\, \vecfarg{\con{f}}(
+z ) \teq \vec t$ for each subterm $\con{f}( \vec t )$ of $t$ such that $\con{f}
+\in \sfundefs{\Sigma}$. Such constraints, when asserted positively, ensure that
+the tuple of uninterpreted functions $\vecfarg{\con{f}}$ must include $\vec t$
+in its range.
 
-We demonstrate this translation with an example.
+%We demonstrate this translation with an example.
 

### 2015-05-05 tex/main.tex
more

@@
 \newcommand{\con}[1]{\mathsf{#1}}
+\let\const=\con
 
@@
 \def\neg{\oldneg\:}
+
+\let\oldvee=\vee
+\def\vee{\mathrel{\oldvee}}
+\let\oldwedge=\wedge
+\def\wedge{\mathrel{\oldwedge}}
 
@@
 \newcommand{\sfunndefs}[1]{#1^\mathrm{nf}}
-\newcommand{\sortint}{\mathrm{Int}}
+\newcommand{\sortint}{\ty{Int}}
 
@@
 
-\newcommand{\Bool}{\con{Bool}}
+\newcommand\ty[1]{\con{#1}}
+\newcommand{\Bool}{\ty{Bool}}
 \newcommand{\ltrue}{\top}
@@
 
-\newcommand{\Sigmalia}{\Sigma_{LIA}}
+\newcommand{\Sigmalia}{\Sigma_{\mathcalx{LIA}}}
 \newcommand{\extendsig}[1]{\mathcalx{E}( #1 )}
@@
 while the fragment in~\cite{ReyEtAl-1-RR-13} is not applicable for functions over infinite domains such as the integers.
-So, for instance, the quantified formula $\forall x\typ{\sortint}.\;\, \lite( p \leq 0, p( x ) \teq 0, p( x ) \teq 2 * p( x - 1 ) )$ 
+So, for instance, the quantified formula $\forall x : {\sortint}.\;\, \lite( p \leq 0, p( x ) \teq 0, p( x ) \teq 2 * p( x - 1 ) )$ 
 is not in the essentially uninterpreted fragment (since $x$ occurs as a child of $-$),
@@
 A set of admissible function definitions can include definitions of non-terminating functions, e.g.
-$\{ \forallf{f} x^\sortint. f( x ) \teq f( x - 1 ) \}$ or even $\{ \forallf{f} x. f( x ) \teq f( x ) \}$ are admissible.
+$\{ \forallf{f} x : \sortint.\;\, f( x ) \teq f( x - 1 ) \}$ or even $\{ \forallf{f} x. f( x ) \teq f( x ) \}$ are admissible.
 On the other hand, the set $\{ \forallf{f} x. f( x ) \teq f( x ) + 1 \}$ is not admissible, 
@@
       \item[] $\mathsf{let}$ $( t'_i,\, \chi_i ) = \conv_0( t_i,\, \pol( \boolop,\, i,\, p ) )$ $\mathsf{for}$ $i = 1, \ldots, n$ $\mathsf{in}$
-      \item[] $\mathsf{let}$ $\chi = \chi_1 \mathrel\wedge \cdots \mathrel\wedge \chi_n$ $\mathsf{in}$
+      \item[] $\mathsf{let}$ $\chi = \chi_1 \wedge \cdots \wedge \chi_n$ $\mathsf{in}$
       \item[] $\mathsf{if}$ $p = \ptrue$ $\mathsf{then}$
@@
 If $t$ is an application of an interpreted predicate $\boolop
-\in \{{\neg}, {\land}, {\lor}, {\teq}, {\lite}\}$,
+\in \{{\neg}, {\wedge}, {\vee}, {\teq}, {\lite}\}$,
 we recursively call $\conv$ on the arguments $t_i$ and polarity $\pol( \boolop,
@@
 \begin{cases}
-p & \text{if either $\boolop \in \{{\land}, {\lor}\}$ or $\boolop = \lite$ and $i \in \{2, 3\}$} \\[-\jot]
+p & \text{if either $\boolop \in \{{\wedge}, {\vee}\}$ or $\boolop = \lite$ and $i \in \{2, 3\}$} \\[-\jot]
 \pneg p & \text{if $\boolop = {\neg}$} \\[-\jot]
@@
 \label{ex:translation}
-Consider the (combined) signature of linear arithmetic and uninterpreted functions $\Sigma = \Sigmalia \cup \Sigma_{\euf}$,
-where $\ssorts{\Sigma_{\euf}} = \{ \sortint \}$ and 
-$\sfuns{\Sigma_{\euf}} = \{ s\typ{ \sortint \rightarrow \sortint }, c\typ{ \sortint } \}$.
-Consider the $\Sigma$-formula $\varphi$:
+Consider the combined signature of linear arithmetic and uninterpreted
+functions $\Sigma = \Sigmalia \cup \Sigma_{\euf}$, where
+$\ssorts{\Sigma_{\euf}} = \{ \sortint \}$ and
+$\sfuns{\Sigma_{\euf}} = \{ \con{s} : { \sortint \rightarrow \sortint },\, \const{c} : { \sortint } \}$.
+Let $\varphi$ be the $\Sigma$-formula
 \begin{equation} \label{eq:ex-before}
-\forall_{s} x\typ{\sortint}.\; \lite( x \leq 0, s(x) \teq 0, s( x ) \teq x + s( x-1 ) ) \wedge s( c ) > 100
+\forallf{\con{s}} x : {\sortint}.\;\, \lite( x \leq 0,\, s(x) \teq 0,\, s( x ) \teq x + s( x-1 ) ) \wedge s( c ) > 100
 \end{equation} 

### 2015-05-05 tex/main.tex
more

@@
 
+\newcommand\xend{{\hfill$\scriptstyle\blacksquare$}}
+
 %\newcommand\bigtuple[1]{$(${#1}$)$}
@@
 
+%%% for \mathcalx
 \DeclareFontFamily{OT1}{pzc}{}
@@
 \DeclareMathAlphabet{\mathcalx}{OT1}{pzc}{m}{it}
+
+%%% for correctly sized Greek
+\DeclareSymbolFont{letters}{OML}{txmi}{m}{it}
 
@@
 \newcommand{\fnull}{\emptyset}
-\newcommand{\vecfarg}[1]{\vec{\concret}_{\vthinspace#1}}
-\newcommand{\farg}[1]{\concret_{\vthinspace#1}}
-\newcommand{\fargx}[2]{\concret^{\vthinspace#2}_{\vthinspace#1}}
-\newcommand{\fargsort}[1]{\upsilon_{\vthinspace#1}}
+\newcommand{\vecfarg}[1]{\vec{\concret}_{#1}}
+\newcommand{\farg}[1]{\concret_{#1}}
+\newcommand{\fargx}[2]{\concret^{\vthinspace#2}_{#1}}
+\newcommand{\fargsort}[1]{\alpha_{#1}}
 
@@
 while the fragment in~\cite{ReyEtAl-1-RR-13} is not applicable for functions over infinite domains such as the integers.
-So, for instance, the quantified formula $\forall x : {\sortint}.\;\, \lite( p \leq 0, p( x ) \teq 0, p( x ) \teq 2 * p( x - 1 ) )$ 
+So, for instance, the quantified formula $\forall x : {\sortint}.\;\, \lite\bigl( p \leq 0, p( x ) \teq 0, p( x ) \teq 2 * p( x - 1 ) \bigr)$ 
 is not in the essentially uninterpreted fragment (since $x$ occurs as a child of $-$),
@@
 \begin{definition}
-Given a set of function definitions $\mathcal{D} = \{ \forallf{f_1} \vec x. \varphi_1, \ldots, \forallf{f_n} \vec x. \varphi_n \}$,
+Given a set of function definitions $\mathcal{D} = \{ \forallf{f_1} \vec x.\; \varphi_1, \ldots, \forallf{f_n} \vec x.\; \varphi_n \}$,
 a formula $\varphi$ is \emph{closed under function expansion with respect to $\mathcal{D}$} if
@@
 The set $\mathcal{D}$ is \emph{admissible} if whenever there exists a model for $\varphi$ that is closed under function expansion,
-there exists a model for $\varphi \wedge \forallf{f_1} \vec x. \varphi_1 \wedge \ldots \wedge \forallf{f_n} \vec x. \varphi_n$.
+there exists a model for $\varphi \wedge \forallf{f_1} \vec x.\; \varphi_1 \wedge \ldots \wedge \forallf{f_n} \vec x.\; \varphi_n$.
 \end{definition}
 
-For instance, the set $\{ \forallf{f} \vec x. f( x ) \teq 0 \}$ is admissible.
+For instance, the set $\{ \forallf{\con{f}} \vec x.\; \con{f}( x ) \teq 0 \}$ is admissible.
 A set of admissible function definitions can include definitions of non-terminating functions, e.g.
-$\{ \forallf{f} x : \sortint.\;\, f( x ) \teq f( x - 1 ) \}$ or even $\{ \forallf{f} x. f( x ) \teq f( x ) \}$ are admissible.
-On the other hand, the set $\{ \forallf{f} x. f( x ) \teq f( x ) + 1 \}$ is not admissible, 
+$\{ \forallf{\con{f}} x : \sortint.\;\, \con{f}( x ) \teq \con{f}( x - 1 ) \}$ or even $\{ \forallf{\con{f}} x.\;\, \con{f}( x ) \teq \con{f}( x ) \}$ are admissible.
+On the other hand, the set $\{ \forallf{\con{f}} x.\;\, \con{f}( x ) \teq \con{f}( x ) + 1 \}$ is not admissible, 
 since the empty set of formulas is closed under function expansion with respect to this set,
-and there is no model satisfying $\forallf{f} x. f( x ) \teq f( x ) + 1$.
-Consider the set $\{ \forallf{f} x. f( x ) \teq f( x ), \forallf{g}. \lite( \neg f( x ) \teq 5, g( x ) \teq g( x ) + 1, g( x ) \teq f( x ) ) \}$.
+and there is no model satisfying $\forallf{\con{f}} x.\;\, \con{f}( x ) \teq \con{f}( x ) + 1$.
+Consider the set $\{ \forallf{\con{f}} x.\;\, \con{f}( x ) \teq \con{f}( x ), \forallf{\con{g}}.\;\, \lite\bigl( \neg \con{f}( x ) \teq 5, \con{g}( x ) \teq \con{g}( x ) + 1, \con{g}( x ) \teq \con{f}( x ) \bigr) \}$.
 While this set has a model where $f$ and $g$ are interpreted as the constant $5$, it is not admissible
-since the formula $f( 0 ) \teq 4$ is closed under function expansion with respect to this set,
-but there exists no interpretation satisfying both $f( 0 ) \teq 4$ and $\forallf{g}. \lite( \neg f( x ) \teq 5, g( x ) \teq g( x ) + 1, g( x ) \teq f( x ) )$.
+since the formula $\con{f}( 0 ) \teq 4$ is closed under function expansion with respect to this set,
+but there exists no interpretation satisfying both $\con{f}( 0 ) \teq 4$ and $\forallf{\con{g}}.\;\, \lite( \neg \con{f}( x ) \teq 5, \con{g}( x ) \teq \con{g}( x ) + 1, \con{g}( x ) \teq \con{f}( x ) )$.
 
@@
 Let us fix a signature $\Sigma$, a $\Sigma$-formula
-$\varphi$ in definitional form with respect to functions $\sfundefs{\Sigma}$,
+$\varphi$ in definitional form with respect to $\sfundefs{\Sigma}$,
 whose definitions are admissible.
@@
 Consider the combined signature of linear arithmetic and uninterpreted
-functions $\Sigma = \Sigmalia \cup \Sigma_{\euf}$, where
+functions $\Sigma = \Sigmalia \mathrel{\cup} \Sigma_{\euf}$, where
 $\ssorts{\Sigma_{\euf}} = \{ \sortint \}$ and
@@
 \begin{equation} \label{eq:ex-before}
-\forallf{\con{s}} x : {\sortint}.\;\, \lite( x \leq 0,\, s(x) \teq 0,\, s( x ) \teq x + s( x-1 ) ) \wedge s( c ) > 100
+\forallf{\con{s}} n : {\sortint}.\;\, \lite\bigl( n \leq 0,\; \con{s}(n) \teq 0,\;
+  \con{s}( n ) \teq n + \con{s}( n - 1 ) \bigr) \wedge \con{s}( \con{c} ) > 100
 \end{equation} 
 %
-Here, $s$ is a function that returns the sum of positive numbers between $0$
-and its argument $x$. The formula $\varphi$ is in definitional form with
-respect to functions $\sfundefs{\Sigma_u}$, and states that the sum of all
-positive numbers between $0$ and $c$ is greater than $100$. This formula is
-satisfiable with a model that interprets $c$ as an integer greater than or
-equal to $14$. Due to the universally quantified formula in the left disjunct,
+Here, $\const{s}$ is a function that returns the sum of the positive integers
+up to $n$. The formula $\varphi$ is in definitional form with
+respect to $\sfundefs{\Sigma_u}$ and states that the sum of all
+positive numbers between $0$ and $\const{c}$ is greater than $100$. This formula is
+satisfiable with a model that interprets $\const{c}$ as an integer greater than or
+equal to $14$. Due to the universal quantifier,
 current SMT techniques~\cite{GeDeM-CAV-09, ReyEtAl-1-RR-13} are unable to find
 a model for this formula. The signature $\extendsig{\Sigma}$ includes the type
-$\fargsort{s}$, and the uninterpreted function $\farg{s}$ of type $\fargsort{s}
-\rightarrow \sortint$. The result of $\conv( \varphi, \ptrue )$ is the pair $(
-\psi, \emptyset )$, where after simplification, $\psi$ is the
+$\fargsort{s}$ and the uninterpreted function $\farg{\con{s}}$ of type $\fargsort{s}
+\rightarrow \sortint$. The result of $\conv( \varphi,\, \ptrue )$ is the pair $(
+\psi,\, \true )$, where (after simplification) $\psi$ is the
 $\extendsig{\Sigma}$-formula
 %
-\begin{eqnarray} \label{eq:ex-after}
-\begin{split}
-\forall y\typ{\fargsort{s}}.\;\, \lite( & \farg{s}( y ) \leq 0,  \\
- & s(\farg{s}( y )) \teq 0, \\
- & s(\farg{s}( y )) \teq \farg{s}( y ) + s( \farg{s}( y )-1 ) ) \\
- & \wedge \exists z\typ{\fargsort{s}}. \farg{s}( z ) \teq \farg{s}( y )-1 )
-\end{split}  
-\wedge ( s( c ) > 100 \wedge \exists z\typ{\fargsort{s}}. \farg{s}( z ) \teq c )
-\end{eqnarray} 
+\begin{equation} \label{eq:ex-after}
+\!\begin{aligned}[t]
+  & \bigl(\forall a : \fargsort{s}.\;\, \lite\bigl( \farg{\con{s}}( a ) \leq 0,\;
+     \con{s}(\farg{\con{s}}( a )) \teq 0,\;
+     \con{s}(\farg{\con{s}}( a )) \teq \farg{\con{s}}( a ) + \con{s}( \farg{\con{s}}( a )-1 ) \bigr) \\[-\jot]
+ & {} \wedge \exists z\typ{\fargsort{s}}. \farg{\con{s}}( z ) \teq \farg{\con{s}}( a )-1 ) \\[-\jot]
+ & {} \wedge ( \con{s}( c ) > 100 \wedge \exists z\typ{\fargsort{s}}. \farg{\con{s}}( z ) \teq c )
+\end{aligned}
+\end{equation}
 %
@@
 techniques~\cite{ReyEtAl-1-RR-13} to search for a finite interpretation for
-$\fargsort{s}$. Moreover, all occurrences of $y$ in this formula are beneath
-applications of the uninterpreted function $\farg{s}$, implying that the
+$\fargsort{s}$. Moreover, all occurrences of $a$ in this formula are beneath
+applications of the uninterpreted function $\farg{\con{s}}$, implying that the
 formula is in the essentially uninterpreted fragment of~\cite{GeDeM-CAV-09},
@@
 \cvc and \ziii run indefinitely on formula~(\ref{eq:ex-before}), and both
-produce a model for formula~(\ref{eq:ex-after}) in $<0.1$ second. $\square$
+produce a model for formula~(\ref{eq:ex-after}) in $<0.1$ second.\xend
 \end{example}
@@
 $\fargsort{s}$ as a finite set of elements $\{ u_0, \ldots, u_{14} \}$,
-$\farg{s}$ as a finite map mapping containing $u_i \mapsto i$ for $i = 0, \ldots, 14$,
+$\farg{\con{s}}$ as a finite map mapping containing $u_i \mapsto i$ for $i = 0, \ldots, 14$,
 $c$ as $14$, 
@@
 \begin{equation} \label{eq:approx-interp}
-\lambda x\typ{Int}. \lite( x \teq 0, 0, \lite( x \teq 1, 1, \lite( x \teq 2, 3, \ldots, \lite( x \teq 13, 91, 105 ) \ldots )
+\lambda x\typ{Int}. \lite\bigl( x \teq 0, 0, \lite( x \teq 1, 1, \lite( x \teq 2, 3, \ldots, \lite( x \teq 13, 91, 105 ) \ldots )\bigr)
 \end{equation}
@@
 Notice that $\I$ is \emph{not} a model for the original formula~(\ref{eq:ex-before}),
-since $\I$ interprets e.g.\ $s( -1 )$ and $s( 15 )$ as $105$.
+since $\I$ interprets e.g.\ $\con{s}( -1 )$ and $\con{s}( 15 )$ as $105$.
 However, under the assumption that function definitions in $\sfundefs{\Sigma}$ are admissible, 
 we claim that $\conv(\varphi)$ is equisatisfiable to $\varphi$ for any input $\varphi$.
-The intuition is that the interpretation of a term such as $s( -1 )$ in the previous example is not relevant to the satisfiability of formula~(\ref{eq:ex-before}).
+The intuition is that the interpretation of a term such as $\con{s}( -1 )$ in the previous example is not relevant to the satisfiability of formula~(\ref{eq:ex-before}).
 Thus, a satisfiable (resp. unsatisfiable) response from an SMT solver on input $\conv( \varphi )$ implies the existence (resp.\ nonexistence) of a model for $\varphi$.
@@
 Conversely, we claim formula~(\ref{eq:ex-after}) is satisfiable if and only if formula~(\ref{eq:ex-before}) is satisfiable,
-noting that the set $\{ \forall_{s} x\typ{\sortint}.\;\, \lite( x \leq 0, s(x) \teq 0, s( x ) \teq x + s( x-1 ) ) \}$ is admissble.
+noting that the set $\{ \forallf{\con{s}} x\typ{\sortint}.\;\, \lite\bigl( x \leq 0, \con{s}(x) \teq 0, \con{s}( x ) \teq x + \con{s}( x-1 ) \bigr) \}$ is admissble.
 As stated in the proof of the previous theorem, 
@@
 hence implying that formula~(\ref{eq:ex-before}) has a model.
-In more detail, let $\varphi_s$ be the formula $\forall_{s} x. ite( x \leq 0, s(x) \teq 0, s( x ) \teq x + s( x-1 ) )$.
+In more detail, let $\varphi_s$ be the formula $\forallf{\con{s}} x.\;\, \lite\bigl( x \leq 0, \con{s}(x) \teq 0, \con{s}( x ) \teq x + \con{s}( x-1 ) \bigr)$.
 The rightmost conjunct 
-$\exists z\typ{\fargsort{s}}. \farg{s}( z ) \teq c$ ensures 
-that any model of formula~(\ref{eq:ex-after}) satisfying $s( c ) > 100$ also satisfies $\varphi_s[c/x]$.
-Similarly, the nested existential $\exists z. \farg{s}( z ) \teq \farg{s}( y )-1 )$
+$\exists z\typ{\fargsort{s}}. \farg{\con{s}}( z ) \teq c$ ensures 
+that any model of formula~(\ref{eq:ex-after}) satisfying $\con{s}( c ) > 100$ also satisfies $\varphi_s[c/x]$.
+Similarly, the nested existential $\exists z. \farg{\con{s}}( z ) \teq \farg{\con{s}}( y )-1 )$
 ensures that this constraint is enforced recursively.
@@
 \begin{split}
-\psi := s( c ) > 100 \wedge & \displaystyle\bigwedge\limits_{i=0}^{13} ( \neg (c-i \leq 0) \wedge s( c-i ) \teq i + s( c-i-1 ) ) \\
-                            & \wedge c-14 \leq 0 \wedge s( c-14 ) \teq 0
+\psi := \con{s}( c ) > 100 \wedge & \displaystyle\bigwedge\limits_{i=0}^{13} ( \neg (c-i \leq 0) \wedge \con{s}( c-i ) \teq i + \con{s}( c-i-1 ) ) \\
+                            & \wedge c-14 \leq 0 \wedge \con{s}( c-14 ) \teq 0
 \end{split}
@@
 Notice that $\psi$ is closed under function expansion, 
-since it contains the $s$-applications that are equivalent to $s( c-i )$ for $i = 0, \ldots, 14$, and
+since it contains the $s$-applications that are equivalent to $\con{s}( c-i )$ for $i = 0, \ldots, 14$, and
 entails $\varphi_s[c-i/x]$ for $i = 0, \ldots, 14$.
@@
 When reading this input, 
-\cvc adds the (annotated) quantified formula $\forallf{s} x.\;\, s( x ) \teq \lite( x \leq 0, 0, s( x-1 ) )$ to its list of assertions,
-which after rewriting becomes $\forallf{s} x.\;\, \lite( x \leq 0, s( x ) \teq 0, s( x ) \teq s( x-1 ) )$.
+\cvc adds the (annotated) quantified formula $\forallf{\con{s}} x.\;\, \con{s}( x ) \teq \lite\bigl( x \leq 0, 0, \con{s}( x-1 )\bigr)$ to its list of assertions,
+which after rewriting becomes $\forallf{\con{s}} x.\;\, \lite\bigl( x \leq 0, \con{s}( x ) \teq 0, \con{s}( x ) \teq \con{s}( x-1 ) \bigr)$.
 If \cvc's finite model finding mode for recursive functions is enabled (using the command line parameter ``--fmf-fun"),

### 2015-05-05 tex/main.tex

@@
 The rightmost conjunct 
-$\exists z\typ{\fargsort{s}}. \farg{\con{s}}( z ) \teq c$ ensures 
-that any model of formula~(\ref{eq:ex-after}) satisfying $\con{s}( c ) > 100$ also satisfies $\varphi_s[c/x]$.
-Similarly, the nested existential $\exists z. \farg{\con{s}}( z ) \teq \farg{\con{s}}( y )-1 )$
+$\exists z : {\fargsort{\con{s}}}. \farg{\con{s}}( z ) \teq \con{c}$ ensures 
+that any model of formula~(\ref{eq:ex-after}) satisfying $\con{s}( \con{c} ) > 100$ also satisfies $\varphi_{\con{s}}[\con{c}/x]$.
+Similarly, the nested existential $\exists z.\;\, \farg{\con{s}}( z ) \teq \farg{\con{s}}( y )-1$
 ensures that this constraint is enforced recursively.

### 2015-05-05 tex/main.tex
renamed z to a, so that it fits with the type \alpha (abstract)

@@
    \begin{itemize}
-     \item[] \return \bigtuple{$t,\, \bigwedge\, \{ \exists z : \fargsort{\con{f}}.\; \vecfarg{\con{f}}( z ) \teq \vec t \mid \con{f}( \vec t ) \in \terms( t ),\, \con{f} \in \sfundefs{\Sigma} \}$}
+     \item[] \return \bigtuple{$t,\, \bigwedge\, \{ \exists a : \fargsort{\con{f}}.\; \vecfarg{\con{f}}( a ) \teq \vec t \mid \con{f}( \vec t ) \in \terms( t ),\, \con{f} \in \sfundefs{\Sigma} \}$}
    \end{itemize}
@@
 $\fargx{\con{f}}{i}$. Second, it augments $\varphi$ with additional constraints
-$\exists z : \fargsort{\con{f}}.\;\, \vecfarg{\con{f}}( z ) \teq \vec t$; these
+$\exists a : \fargsort{\con{f}}.\;\, \vecfarg{\con{f}}( a ) \teq \vec t$; these
 ensure that the restricted definition for $\con{f}$ covers all relevant tuples
@@
 of a type other than $\Bool$. The returned constraint is a conjunction of
-formulas of the form $\exists z\typ{\fargsort{\con{f}}}.\;\, \vecfarg{\con{f}}(
-z ) \teq \vec t$ for each subterm $\con{f}( \vec t )$ of $t$ such that $\con{f}
+formulas of the form $\exists a : {\fargsort{\con{f}}}.\;\, \vecfarg{\con{f}}(
+a ) \teq \vec t$ for each subterm $\con{f}( \vec t )$ of $t$ such that $\con{f}
 \in \sfundefs{\Sigma}$. Such constraints, when asserted positively, ensure that
@@
 \begin{equation} \label{eq:ex-after}
+\!\begin{aligned}[c]
+  & \phantom{{\wedge}\; }\bigl(
     \!\begin{aligned}[t]
-  & \bigl(\forall a : \fargsort{s}.\;\, \lite\bigl( \farg{\con{s}}( a ) \leq 0,\;
+      \forall a : \fargsort{s}.\; & \phantom{{\wedge}\;}\lite\bigl( \farg{\con{s}}( a ) \leq 0,\;
         \con{s}(\farg{\con{s}}( a )) \teq 0,\;
-     \con{s}(\farg{\con{s}}( a )) \teq \farg{\con{s}}( a ) + \con{s}( \farg{\con{s}}( a )-1 ) \bigr) \\[-\jot]
- & {} \wedge \exists z\typ{\fargsort{s}}. \farg{\con{s}}( z ) \teq \farg{\con{s}}( a )-1 ) \\[-\jot]
- & {} \wedge ( \con{s}( c ) > 100 \wedge \exists z\typ{\fargsort{s}}. \farg{\con{s}}( z ) \teq c )
+        \con{s}(\farg{\con{s}}( a )) \teq \farg{\con{s}}( a ) + \con{s}( \farg{\con{s}}( a )-1 ) \bigr)
+\\[-\jot]
+      {\wedge}\; & \exists a : {\fargsort{s}}.\;\, \farg{\con{s}}( a ) \teq \farg{\con{s}}( a )-1 \bigr)
+\end{aligned}
+\\[-\jot]
+ & {\wedge}\; \bigl( \con{s}( c ) > 100 \wedge \exists a : {\fargsort{s}}.\;\, \farg{\con{s}}( a ) \teq \con{c} \bigr)
 \end{aligned}
@@
 and each $\fargx{\con{f}}{i}$ as the $i^{th}$ projection on such tuples for each $i = 1, \ldots, n$.
-Note that $\J$ satisfies \emph{every} constraint of the form $\exists z\typ{\fargsort{\con{f}}}. ( \vecfarg{\con{f}}( z ) \teq \vec t )$,
-since by our construction of $\J$ there exists a $z$ such that $z^\J = \vec t^\J$.
+Note that $\J$ satisfies \emph{every} constraint of the form $\exists a : {\fargsort{\con{f}}}. ( \vecfarg{\con{f}}( a ) \teq \vec t )$,
+since by our construction of $\J$ there exists an $a$ such that $a^\J = \vec t^\J$.
 By case analysis it follows that $\conv_0$ maintains the invariant that
@@
 The rightmost conjunct 
-$\exists z : {\fargsort{\con{s}}}. \farg{\con{s}}( z ) \teq \con{c}$ ensures 
+$\exists a : {\fargsort{\con{s}}}. \farg{\con{s}}( a ) \teq \con{c}$ ensures 
 that any model of formula~(\ref{eq:ex-after}) satisfying $\con{s}( \con{c} ) > 100$ also satisfies $\varphi_{\con{s}}[\con{c}/x]$.
-Similarly, the nested existential $\exists z.\;\, \farg{\con{s}}( z ) \teq \farg{\con{s}}( y )-1$
+Similarly, the nested existential $\exists a.\;\, \farg{\con{s}}( a ) \teq \farg{\con{s}}( y )-1$
 ensures that this constraint is enforced recursively.

### 2015-05-05 tex/main.tex
more

@@
 %\newcommand\pneg{\neg}
-\newcommand\pneg{-}
+\newcommand\pnegate{-}
 
@@
 
-\newcommand{\pnone}{\con{none}}
-\newcommand{\ptrue}{\con{pos}}
-\newcommand{\pfalse}{\con{neg}}
+\newcommand{\pnone}{\con{apolar}}
+\newcommand{\ppos}{\con{pos}}
+\newcommand{\pneg}{\con{neg}}
 \newcommand{\pol}{\con{pol}}
@@
 on the auxiliary function $\conv_0$, which takes two arguments:\ the term $t$
-to translate and a \emph{polarity} $p$, which is either $\ptrue$, $\pfalse$, or
+to translate and a \emph{polarity} $p$, which is either $\ppos$, $\pneg$, or
 $\pnone$. It returns a pair $( t'\negvthinspace,\, \chi )$, where $t'$ is a term of
@@
 This constraint is necessary if the polarity is $\pnone$; otherwise, any necessary
-constraint is encoded directly in $t'$ and $\chi$ is simply $\true$.
+constraint is encoded directly in $t'$ and $\chi$ is simply $\ltrue$.
 
@@
       \item[] $\mathsf{let}$ $\chi = \chi_1 \wedge \cdots \wedge \chi_n$ $\mathsf{in}$
-      \item[] $\mathsf{if}$ $p = \ptrue$ $\mathsf{then}$
+      \item[] $\mathsf{if}$ $p = \ppos$ $\mathsf{then}$
       \begin{itemize}
-        \item[] \return \bigtuple{$\boolop(t'_1, \ldots, t'_n) ) \wedge \chi,\, \true$}
+        \item[] \return \bigtuple{$\boolop(t'_1, \ldots, t'_n) ) \wedge \chi,\, \ltrue$}
       \end{itemize}
-      \item[] $\mathsf{else}$ $\mathsf{if}$ $p = \pfalse$ $\mathsf{then}$
+      \item[] $\mathsf{else}$ $\mathsf{if}$ $p = \pneg$ $\mathsf{then}$
       \begin{itemize}
-        \item[] \return \bigtuple{$\boolop(t'_1, \ldots, t'_n) \vee \neg \chi,\, \true$}
+        \item[] \return \bigtuple{$\boolop(t'_1, \ldots, t'_n) \vee \neg \chi,\, \ltrue$}
       \end{itemize}
@@
       \item[] $\mathsf{let}$ $( u',\, \chi ) = \conv_0( u,\, p )$ $\mathsf{in}$
-      %\item[] $\mathsf{assert}( \chi = \true )$;
-      \item[] \return \bigtuple{$\forall y\typ{\fargsort{\con{f}}}.\; u' [ \vecfarg{\con{f}}( y ) / \vec x ],\, \true$}
+      %\item[] $\mathsf{assert}( \chi = \ltrue )$;
+      \item[] \return \bigtuple{$\forall y\typ{\fargsort{\con{f}}}.\; u' [ \vecfarg{\con{f}}( y ) / \vec x ],\, \ltrue$}
     \end{itemize}
@@
  \begin{itemize}
-   \item[] $\mathsf{let}$ $( \psi,\, \_ ) = \conv_0( \varphi,\, \ptrue )$ $\mathsf{in}$
+   \item[] $\mathsf{let}$ $( \psi,\, \_ ) = \conv_0( \varphi,\, \ppos )$ $\mathsf{in}$
    %\item[] $\mathsf{assert}( D = \emptyset )$;
@@
 p & \text{if either $\boolop \in \{{\wedge}, {\vee}\}$ or $\boolop = \lite$ and $i \in \{2, 3\}$} \\[-\jot]
-\pneg p & \text{if $\boolop = {\neg}$} \\[-\jot]
+\pnegate p & \text{if $\boolop = {\neg}$} \\[-\jot]
 \pnone & \text{otherwise}
@@
 \end{equation*}
-The negation $\pneg p$ of a polarity $p$ is $\pfalse$ if $p$ is $\ptrue$, $\ptrue$
-if $p$ is $\pfalse$, and $\pnone$ if $p$ is $\pnone$.
+The negation $\pnegate p$ of a polarity $p$ is $\pneg$ if $p$ is $\ppos$, $\ppos$
+if $p$ is $\pneg$, and $\pnone$ if $p$ is $\pnone$.
 
@@
 are a top-level conjunct, %by case analysis on the return values of $\conv$,
-we know that $\chi$ is simply $\true$.
+we know that $\chi$ is simply $\ltrue$.
 If $t$ is not a function definition, the resulting
@@
 $\fargsort{s}$ and the uninterpreted function $\farg{\con{s}}$ of type $\fargsort{s}
-\rightarrow \sortint$. The result of $\conv( \varphi,\, \ptrue )$ is the pair $(
-\psi,\, \true )$, where (after simplification) $\psi$ is the
+\rightarrow \sortint$. The result of $\conv( \varphi,\, \ppos )$ is the pair $(
+\psi,\, \ltrue )$, where (after simplification) $\psi$ is the
 $\extendsig{\Sigma}$-formula
@@
 By case analysis it follows that $\conv_0$ maintains the invariant that
-if $\conv_0( t\typ{\Bool}, \ptrue )$ returns $( r\typ{\Bool}, D )$,
+if $\conv_0( t\typ{\Bool}, \ppos )$ returns $( r\typ{\Bool}, D )$,
 then $\I$ satisfies $t$ if and only if $\J$ satisfies $r$.

### 2015-05-05 tex/main.tex
more

@@
 
-\newcommand{\cvc}{\textsc{cvc}{\small 4}\xspace}
-\newcommand{\cvciii}{\textsc{cvc}{\small 3}\xspace}
-\newcommand{\ziii}{\textsc{z}{\small 3}\xspace}
+%\newcommand{\cvc}{\textsc{cvc}{\small 4}\xspace}
+%\newcommand{\cvciii}{\textsc{cvc}{\small 3}\xspace}
+%\newcommand{\ziii}{\textsc{z}{\small 3}\xspace}
+
+%%% @ANDY, @CESARE: I hope you'll forgive me these. I find the above fonts
+%%% distracting. --JB
+\newcommand\cvc{CVC4\xspace}
+\newcommand\cvciii{CVC3\xspace}
+\newcommand\ziii{Z3\xspace}
+
 \newcommand{\teq}{\approx}
@@
 a model for this formula. The signature $\extendsig{\Sigma}$ includes the type
-$\fargsort{s}$ and the uninterpreted function $\farg{\con{s}}$ of type $\fargsort{s}
+$\fargsort{\con{s}}$ and the uninterpreted function $\farg{\con{s}}$ of type $\fargsort{\con{s}}
 \rightarrow \sortint$. The result of $\conv( \varphi,\, \ppos )$ is the pair $(
@@
     \!\begin{aligned}[t]
-      \forall a : \fargsort{s}.\; & \phantom{{\wedge}\;}\lite\bigl( \farg{\con{s}}( a ) \leq 0,\;
+      \forall a : \fargsort{\con{s}}.\; & \phantom{{\wedge}\;}\lite\bigl( \farg{\con{s}}( a ) \leq 0,\;
         \con{s}(\farg{\con{s}}( a )) \teq 0,\;
@@
 \\[-\jot]
-      {\wedge}\; & \exists a : {\fargsort{s}}.\;\, \farg{\con{s}}( a ) \teq \farg{\con{s}}( a )-1 \bigr)
+      & {\wedge}\; \exists a : {\fargsort{\con{s}}}.\;\, \farg{\con{s}}( a ) \teq \farg{\con{s}}( a )-1 \bigr)
 \end{aligned}
 \\[-\jot]
- & {\wedge}\; \bigl( \con{s}( c ) > 100 \wedge \exists a : {\fargsort{s}}.\;\, \farg{\con{s}}( a ) \teq \con{c} \bigr)
+ & {\wedge}\; \bigl( \con{s}( c ) > 100 \wedge \exists a : {\fargsort{\con{s}}}.\;\, \farg{\con{s}}( a ) \teq \con{c} \bigr)
 \end{aligned}
@@
 The universal quantification in the resulting formula is over an uninterpreted
-type $\fargsort{s}$, thus enabling finite model finding
-techniques~\cite{ReyEtAl-1-RR-13} to search for a finite interpretation for
-$\fargsort{s}$. Moreover, all occurrences of $a$ in this formula are beneath
+type $\fargsort{\con{s}}$, thus enabling the finite model finding
+techniques by Reynolds et al.\ \cite{ReyEtAl-1-RR-13} to search for a finite interpretation for
+$\fargsort{\con{s}}$. Moreover, all occurrences of $a$ in this formula are beneath
 applications of the uninterpreted function $\farg{\con{s}}$, implying that the
-formula is in the essentially uninterpreted fragment of~\cite{GeDeM-CAV-09},
-for which there exists a complete instantiation procedure. We observe that both
+formula is in the essentially uninterpreted fragment,
+for which Ge and de Moura \cite{GeDeM-CAV-09} provide a complete instantiation
+procedure. We observe that both
 \cvc and \ziii run indefinitely on formula~(\ref{eq:ex-before}), and both
@@
 One model $\I$ for formula~(\ref{eq:ex-after}) in the previous example interprets 
-$\fargsort{s}$ as a finite set of elements $\{ u_0, \ldots, u_{14} \}$,
+$\fargsort{\con{s}}$ as a finite set of elements $\{ u_0, \ldots, u_{14} \}$,
 $\farg{\con{s}}$ as a finite map mapping containing $u_i \mapsto i$ for $i = 0, \ldots, 14$,

### 2015-05-05 tex/main.tex
more

@@
 if all interpretations in $\Mo$ that satisfy $\varphi$ also satisfy $\psi$.
-Formulas $\varphi$ and $\psi$ are \emph{equisatisfiable} if
-$\varphi$ is $T$-satisfiable if and only if $\psi$ is $T$-satisfiable.
 Given a signature $\Sigma$,
@@
 $\extendsig{ \Sigma }$ and a $\extendsig{ \Sigma }$-formula $\psi$ such that
-$\varphi$ and $\psi$ are equisatisfiable.
-\jb{Equisatisfiability as defined in Sect.\ 2 requires the same $\Sigma$ and
-the same $T$ for both formulas, which isn't the case here.}
+$\psi$ is satisfiable if and only if $\varphi$ is satisfiable---i.e.,
+$\varphi$ and $\psi$ are \emph{equisatisfiable}.
 The intuition behind this translation
@@
 for which Ge and de Moura \cite{GeDeM-CAV-09} provide a complete instantiation
-procedure. We observe that both
-\cvc and \ziii run indefinitely on formula~(\ref{eq:ex-before}), and both
-produce a model for formula~(\ref{eq:ex-after}) in $<0.1$ second.\xend
+procedure. Both
+%%% Added "apparently". Or are we sure?
+\cvc and \ziii apparently run indefinitely on formula~(\ref{eq:ex-before}), and both
+produce a model for formula~(\ref{eq:ex-after}) within 100 milliseconds.\xend
 \end{example}
 
-It is important to note that the translation $\conv$ does not preserve the models of $\varphi$.
+The translation $\conv$ does not preserve the models of $\varphi$.
 One model $\I$ for formula~(\ref{eq:ex-after}) in the previous example interprets 
 $\fargsort{\con{s}}$ as a finite set of elements $\{ u_0, \ldots, u_{14} \}$,
-$\farg{\con{s}}$ as a finite map mapping containing $u_i \mapsto i$ for $i = 0, \ldots, 14$,
-$c$ as $14$, 
-and $s$ as the (almost constant) function:
+$\farg{\con{s}}$ as a finite map $u_i \mapsto i$ for $i = 0, \ldots, 14$,
+$\con{c}$ as $14$, 
+and $\con{s}$ as the (almost constant) function:
+%
+%% TYPESETTING: \mathbin and to make it fit nicely
 \begin{equation} \label{eq:approx-interp}
-\lambda x\typ{Int}. \lite\bigl( x \teq 0, 0, \lite( x \teq 1, 1, \lite( x \teq 2, 3, \ldots, \lite( x \teq 13, 91, 105 ) \ldots )\bigr)
+\lambda x : \sortint.\,\, \lite( x \mathbin{\teq} 0,\, 0,\,
+  \lite( x \mathbin{\teq} 1,\, 1,\,
+    \lite( x \mathbin{\teq} 2,\, 3,\,
+      \lite(\ldots,\; \lite( x \mathbin{\teq} 13,\, 91,\, 105 )\negvthinspace \ldots ))))
 \end{equation}
-In other words, $s$ is a function mapping $x$ to the sum of all positive integers between $0$ and $x$ when $0 \leq x \leq 13$,
-and $105$ otherwise.
-Notice that $\I$ is \emph{not} a model for the original formula~(\ref{eq:ex-before}),
-since $\I$ interprets e.g.\ $\con{s}( -1 )$ and $\con{s}( 15 )$ as $105$.
-However, under the assumption that function definitions in $\sfundefs{\Sigma}$ are admissible, 
-we claim that $\conv(\varphi)$ is equisatisfiable to $\varphi$ for any input $\varphi$.
-The intuition is that the interpretation of a term such as $\con{s}( -1 )$ in the previous example is not relevant to the satisfiability of formula~(\ref{eq:ex-before}).
-Thus, a satisfiable (resp. unsatisfiable) response from an SMT solver on input $\conv( \varphi )$ implies the existence (resp.\ nonexistence) of a model for $\varphi$.
-Moreover, information in models for $\conv( \varphi )$ contains relevant information regarding the models of $\varphi$.
-For example, the model $\I$ for formula~(\ref{eq:ex-after}) described above interprets $c$ as $14$, 
-while there exists a model of formula~(\ref{eq:ex-before}) also interprets $c$ as $14$.
-In fact, we claim that there exists a model of $\varphi$ that coincides with each model of $\conv( \varphi )$ on its interpretation of all functions in $\sfuns{\Sigma}$ 
-other than those in $\sfundefs{\Sigma}$, i.e. $\{ s \}$ in the previous example.
-From a practical point of view, this is not an issue, because
-$s$ is the very function that was explicitly defined by the user,
-and hence already has an intended interpretation.
+%
+In other words, $\const{s}$ is interpreted as a function mapping $x$ to the sum
+of all positive integers between $0$ and $x$ when $0 \leq x \leq 13$, and $105$
+otherwise.
+The interpretation $\I$, restricted to $\Sigma$, is {not} a model for the original formula~(\ref{eq:ex-before}),
+since $\I$ wrongly interprets $\con{s}( x )$ as $105$ when $x < 0$ or $x > 14$.
+
+However, under the assumption that the function definitions in
+$\sfundefs{\Sigma}$ are admissible, we claim that $\conv(\varphi)$ is
+equisatisfiable to $\varphi$ for any input $\varphi$. The intuition is that the
+interpretation of a term such as $\con{s}( -1 )$ in the previous example is not
+relevant to the satisfiability of formula~(\ref{eq:ex-before}). Thus, a
+satisfiable (resp. unsatisfiable) response from an SMT solver on input $\conv(
+\varphi )$ implies the existence (resp.\ nonexistence) of a model for
+$\varphi$. Moreover, information in models for $\conv( \varphi )$ contains
+relevant information regarding the models of $\varphi$. For example, the model
+$\I$ for formula~(\ref{eq:ex-after}) described above interprets $c$ as $14$,
+while there exists a model of formula~(\ref{eq:ex-before}) also interprets $c$
+as $14$. In fact, we claim that there exists a model of $\varphi$ that
+coincides with each model of $\conv( \varphi )$ on its interpretation of all
+functions in $\sfuns{\Sigma}$ other than those in $\sfundefs{\Sigma}$, i.e. $\{
+s \}$ in the previous example. From a practical point of view, this is not an
+issue, because $s$ is the very function that was explicitly defined by the
+user, and hence already has an intended interpretation.
 

### 2015-05-05 tex/main.tex
more

@@
 \def\ENDPROC{\untab}
-\newcommand{\conv}{\mathcalx{W}}
+%%% @ANDY: I'm trying "A" like "abstraction", again in sync with the abstract--concrete
+%%% terminology from abstract interpretation
+\newcommand{\conv}{\mathcalx{A}}
 
@@
 However, under the assumption that the function definitions in
-$\sfundefs{\Sigma}$ are admissible, we claim that $\conv(\varphi)$ is
-equisatisfiable to $\varphi$ for any input $\varphi$. The intuition is that the
+$\sfundefs{\Sigma}$ are admissible, %we claim that
+$\conv(\varphi)$ is
+equisatisfiable with $\varphi$ for any input $\varphi$. The intuition is that the
 interpretation of a term such as $\con{s}( -1 )$ in the previous example is not
 relevant to the satisfiability of formula~(\ref{eq:ex-before}). Thus, a
-satisfiable (resp. unsatisfiable) response from an SMT solver on input $\conv(
-\varphi )$ implies the existence (resp.\ nonexistence) of a model for
+satisfiable or unsatisfiable response from an SMT solver on input $\conv(
+\varphi )$ implies the existence or nonexistence of a model for
 $\varphi$. Moreover, information in models for $\conv( \varphi )$ contains
-relevant information regarding the models of $\varphi$. For example, the model
-$\I$ for formula~(\ref{eq:ex-after}) described above interprets $c$ as $14$,
-while there exists a model of formula~(\ref{eq:ex-before}) also interprets $c$
-as $14$. In fact, we claim that there exists a model of $\varphi$ that
+pertinent information regarding the models of $\varphi$. For example, the model
+$\I$ for formula~(\ref{eq:ex-after}) described above interprets $\con{c}$ as $14$,
+and there exists a model of formula~(\ref{eq:ex-before}) that also interprets $\con{c}$
+as $14$. In general, there exists a model of $\varphi$ that
 coincides with each model of $\conv( \varphi )$ on its interpretation of all
-functions in $\sfuns{\Sigma}$ other than those in $\sfundefs{\Sigma}$, i.e. $\{
-s \}$ in the previous example. From a practical point of view, this is not an
-issue, because $s$ is the very function that was explicitly defined by the
-user, and hence already has an intended interpretation.
+functions in $\sfuns{\Sigma} - \sfundefs{\Sigma}$, i.e., $\{
+\con{s} \}$ in the example.
+
+%%% @ANDY: I don't get the following comment --jb
+%
+%From a practical point of view, this is not an
+%issue, because $s$ is the very function that was explicitly defined by the
+%user, and hence already has an intended interpretation.
 
 \begin{theorem}
-If $\varphi$ is a $\Sigma$-formula that is in definitional form with respect to $\sfundefs{\Sigma}$,
-and the set of function definitions $\mathcal{D}$ corresponding to functions in $\sfundefs{\Sigma}$ is admissible,
-then $\varphi$ and $\conv( \varphi )$ are equisatisfiable.
+If $\varphi$ is a\/ $\Sigma$-formula that is in definitional form with respect to\/
+$\sfundefs{\Sigma}$
+and the set of function definitions $\mathcal{D}$ corresponding to functions in
+$\sfundefs{\Sigma}$ is admissible, then $\varphi$ and $\conv( \varphi )$ are
+equisatisfiable.
 \end{theorem}
@@
 
+  * abstract interpretation
+
 Leon

### 2015-05-05 tex/main.tex
more

@@
 
+%\newcommand{\DDD}{\mathcalx{D}}
+\newcommand\DDD{\Delta}
+
 %\newcommand\return{$\mathsf{return}$\xspace}
@@
 \def\Sigma{\mathrm{\oldSigma}}
+
+\let\oldDelta=\Delta
+\def\Delta{\mathrm{\oldDelta}}
 
@@
 \newcommand{\sfuns}[1]{#1^\mathrm{f}}
-\newcommand{\sfundefs}[1]{#1^\mathrm{df}}
-\newcommand{\sfunndefs}[1]{#1^\mathrm{nf}}
+\newcommand{\sfundefs}[1]{#1^\mathrm{def}}
+%\newcommand{\sfunndefs}[1]{#1^\mathrm{nf}}
 \newcommand{\sortint}{\ty{Int}}
@@
 \begin{definition}
-Given a set of function definitions $\mathcal{D} = \{ \forallf{f_1} \vec x.\; \varphi_1, \ldots, \forallf{f_n} \vec x.\; \varphi_n \}$,
-a formula $\varphi$ is \emph{closed under function expansion with respect to $\mathcal{D}$} if
+Given a set of function definitions $\DDD = \{ \forallf{f_1} \vec x.\; \varphi_1, \ldots, \forallf{f_n} \vec x.\; \varphi_n \}$,
+a formula $\varphi$ is \emph{closed under function expansion with respect to $\DDD$} if
 \begin{eqnarray}
@@
 \end{eqnarray}
-The set $\mathcal{D}$ is \emph{admissible} if whenever there exists a model for $\varphi$ that is closed under function expansion,
+The set $\DDD$ is \emph{admissible} if whenever there exists a model for $\varphi$ that is closed under function expansion,
 there exists a model for $\varphi \wedge \forallf{f_1} \vec x.\; \varphi_1 \wedge \ldots \wedge \forallf{f_n} \vec x.\; \varphi_n$.
@@
 $\sfundefs{\Sigma}$
-and the set of function definitions $\mathcal{D}$ corresponding to functions in
-$\sfundefs{\Sigma}$ is admissible, then $\varphi$ and $\conv( \varphi )$ are
+and the set of function definitions\/ $\DDD$ corresponding to functions in
+$\sfundefs{\Sigma}$ is admissible, then\/ $\varphi$ and\/ $\conv( \varphi )$ are
 equisatisfiable.
@@
 Second, we show that if $\conv( \varphi )$ is satisfied by a $\extendsig{\Sigma}$-interpretation $\I$, then $\varphi$ is satisfied by $\Sigma$-interpretation $\J$.
-Since $\varphi$ is in definitional form with respect to the functions defined by $\mathcal{D}$, we know $\varphi$ is of the form $\mathcal{D} \wedge \varphi_0$.
-We first argue that $\I$ satisfies a $\Sigma$-formula $\psi$ that is closed under function expansion with respect to $\mathcal{D}$,
+Since $\varphi$ is in definitional form with respect to the functions defined by $\DDD$, we know $\varphi$ is of the form $\DDD \wedge \varphi_0$.
+We first argue that $\I$ satisfies a $\Sigma$-formula $\psi$ that is closed under function expansion with respect to $\DDD$,
 and where $\psi$ entails $\varphi_0$.
 
-Since there exists a model for $\psi$ that is closed under function expansion with respect to $\mathcal{D}$, 
-then since $\mathcal{D}$ is admissible,
-by definition there exists a model for $\psi \wedge \mathcal{D}$, which entails $\mathcal{D} \wedge \varphi_0$, which is $\varphi$.
+Since there exists a model for $\psi$ that is closed under function expansion with respect to $\DDD$, 
+then since $\DDD$ is admissible,
+by definition there exists a model for $\psi \wedge \DDD$, which entails $\DDD \wedge \varphi_0$, which is $\varphi$.
 \end{proof}

### 2015-05-05 tex/main.tex
tuned whitespaces
more

@@
 \usepackage[T1]{fontenc}
+\usepackage{amsmath}
+\usepackage{amssymb}
 \usepackage[scaled=.82]{beramono}
@@
 \usepackage{xspace}
-\usepackage{amssymb}
 \setcounter{tocdepth}{3}
@@
 \usepackage{cite}
-\usepackage{hyperref}
+\usepackage[
+   a4paper,
+   pdftex,
+   pdftitle={Model Finding for Recursive Functions in SMT},
+   pdfauthor={Andrew Reynolds, Jasmin Christian Blanchette, and Cesare Tinelli},
+   pdfkeywords={},
+   pdfborder={0 0 0},
+   draft=false,
+   bookmarksnumbered,
+   bookmarks,
+   bookmarksdepth=2,
+   bookmarksopenlevel=2,
+   bookmarksopen]{hyperref}
 \usepackage{framed}
@@
 
+\spnewtheorem*{proofsketch}{Proof sketch}{\itshape}{\rmfamily}
+
 \usepackage{url}
@@
 \newcommand\return{}
-
-\newcommand\xend{{\hfill$\scriptstyle\blacksquare$}}
 
@@
 
+\newcommand\xend{{\hfill$\scriptstyle\blacksquare$}}
+\renewcommand\qed{{\hfill$\squareforqed$}}
+
 \begin{document}
@@
 
-\ajr{Demand for recursive functions in SMT (new draft includes define-funs-rec).  Usually not interested in showing logical consistency, there exist other tools that check e.g. termination. 
-Thus, want SMT solver to check satisfiability the assumption that recursive functions are consistent (where this needs to be defined precisely).
+\ajr{Demand for recursive functions in SMT (new draft includes
+define-funs-rec). Usually not interested in showing logical consistency, there
+exist other tools that check e.g. termination.
+Thus, want SMT solver to check satisfiability the assumption that recursive
+functions are consistent (where this needs to be defined precisely).
 }
@@
 $\fnull$.
-Given $\con{f} : \tau_1 \times \ldots \times \tau_n \rightarrow \tau$,
+Given $\con{f} : \tau_1 \times \cdots \times \tau_n \rightarrow \tau$,
 an annotated
@@
 \begin{eqnarray}
-\varphi \models_T \displaystyle\bigwedge\limits_{i=1}^n \{ \varphi_i[ \vec t / \vec x ] \mid f_i( \vec t ) \in \terms( \varphi ) \}
+\varphi \models_T \bigwedge_{i=1}^n \{ \varphi_i[ \vec t / \vec x ] \mid f_i( \vec t ) \in \terms( \varphi ) \}
 \end{eqnarray}
@@
 We extend the signature of $\Sigma$ to a signature $\extendsig{\Sigma}$ containing the following additional definitions.
-For each $\con{f} : \tau_1 \times \ldots \times \tau_n \rightarrow \tau \in \sfundefs{\Sigma}$,
+For each $\con{f} : \tau_1 \times \cdots \times \tau_n \rightarrow \tau \in \sfundefs{\Sigma}$,
 the signature $\extendsig{\Sigma}$ includes
@@
 
-\begin{theorem}
+\begin{theorem}\label{thm:equi}
 If $\varphi$ is a\/ $\Sigma$-formula that is in definitional form with respect to\/
@@
 \end{theorem}
-\begin{proof}
-(Sketch)  First, we show that if $\varphi$ is satisfied by $\Sigma$-interpretation $\I$, then $\conv( \varphi )$ is satisfied by a $\extendsig{\Sigma}$-interpretation $\J$.
+\begin{proofsketch}
+First, we show that if $\varphi$ is satisfied by $\Sigma$-interpretation $\I$, then $\conv( \varphi )$ is satisfied by a $\extendsig{\Sigma}$-interpretation $\J$.
 Let $\J$ be the $\extendsig{\Sigma}$-interpretation that interprets all types $\tau \in \ssorts{\Sigma}$ as $\tau^\I$,
 all functions $\con{f} \in \sfuns{\Sigma}$ as $\con{f}^\I$,
-and for each function $\con{f}$ of type $\tau_1 \times \ldots \tau_n \rightarrow \tau$ in $\sfundefs{\Sigma}$, 
-interprets $\fargsort{\con{f}}$ as a tuple of values of type $\tau_1^\I \times \ldots \times \tau_n^\I$,
+and for each function $\con{f}$ of type $\tau_1 \times \cdots \tau_n \rightarrow \tau$ in $\sfundefs{\Sigma}$,
+interprets $\fargsort{\con{f}}$ as a tuple of values of type $\tau_1^\I \times \cdots \times \tau_n^\I$,
 and each $\fargx{\con{f}}{i}$ as the $i^{th}$ projection on such tuples for each $i = 1, \ldots, n$.
@@
 
-Second, we show that if $\conv( \varphi )$ is satisfied by a $\extendsig{\Sigma}$-interpretation $\I$, then $\varphi$ is satisfied by $\Sigma$-interpretation $\J$.
-Since $\varphi$ is in definitional form with respect to the functions defined by $\DDD$, we know $\varphi$ is of the form $\DDD \wedge \varphi_0$.
-We first argue that $\I$ satisfies a $\Sigma$-formula $\psi$ that is closed under function expansion with respect to $\DDD$,
-and where $\psi$ entails $\varphi_0$.
+Second, we show that if $\conv( \varphi )$ is satisfied by a
+$\extendsig{\Sigma}$-interpretation $\I$, then $\varphi$ is satisfied by
+$\Sigma$-interpretation $\J$. Since $\varphi$ is in definitional form with
+respect to the functions defined by $\DDD$, we know $\varphi$ is of the form
+$\DDD \wedge \varphi_0$. We first argue that $\I$ satisfies a $\Sigma$-formula
+$\psi$ that is closed under function expansion with respect to $\DDD$, and
+where $\psi$ entails $\varphi_0$.
 
-Since there exists a model for $\psi$ that is closed under function expansion with respect to $\DDD$, 
-then since $\DDD$ is admissible,
-by definition there exists a model for $\psi \wedge \DDD$, which entails $\DDD \wedge \varphi_0$, which is $\varphi$.
-\end{proof}
+Since there exists a model for $\psi$ that is closed under function expansion
+with respect to $\DDD$, then since $\DDD$ is admissible, by definition there
+exists a model for $\psi \wedge \DDD$, which entails $\DDD \wedge \varphi_0$,
+which is $\varphi$.
+\qed
+\end{proofsketch}
 
-Let us revisit the translated formula~(\ref{eq:ex-after}) from Example~\ref{ex:translation}.
-This formula is unsatisfiable if and only if formula~(\ref{eq:ex-before}) is unsatisfiable since $\fargsort{f}$ can be interpreted as $\sortint$,
-and $\farg{f}$ can be interpreted as the identity function.
-Conversely, we claim formula~(\ref{eq:ex-after}) is satisfiable if and only if formula~(\ref{eq:ex-before}) is satisfiable,
-noting that the set $\{ \forallf{\con{s}} x\typ{\sortint}.\;\, \lite\bigl( x \leq 0, \con{s}(x) \teq 0, \con{s}( x ) \teq x + \con{s}( x-1 ) \bigr) \}$ is admissble.
-As stated in the proof of the previous theorem, 
-the reason is that any model of formula~(\ref{eq:ex-after}) must satisfy a formula that is closed under function expansion,
-hence implying that formula~(\ref{eq:ex-before}) has a model.
-In more detail, let $\varphi_s$ be the formula $\forallf{\con{s}} x.\;\, \lite\bigl( x \leq 0, \con{s}(x) \teq 0, \con{s}( x ) \teq x + \con{s}( x-1 ) \bigr)$.
-The rightmost conjunct 
-$\exists a : {\fargsort{\con{s}}}. \farg{\con{s}}( a ) \teq \con{c}$ ensures 
-that any model of formula~(\ref{eq:ex-after}) satisfying $\con{s}( \con{c} ) > 100$ also satisfies $\varphi_{\con{s}}[\con{c}/x]$.
-Similarly, the nested existential $\exists a.\;\, \farg{\con{s}}( a ) \teq \farg{\con{s}}( y )-1$
-ensures that this constraint is enforced recursively.
-Given a model $\I$ of formula~(\ref{eq:ex-after}) that interprets $c$ as $14$, 
-it is easy to see that $\I$ must satisfy the formula:
-\begin{eqnarray} \label{eq:fun-closure}
+\begin{example}
+Let us revisit the translated formula~(\ref{eq:ex-after}) from
+Example~\ref{ex:translation}. This formula is unsatisfiable if and only if
+formula~(\ref{eq:ex-before}) is unsatisfiable since $\fargsort{f}$ can be
+interpreted as $\sortint$, and $\farg{f}$ can be interpreted as the identity
+function. Conversely, we claim formula~(\ref{eq:ex-after}) is satisfiable if
+and only if formula~(\ref{eq:ex-before}) is satisfiable, noting that the set
+$\{ \forallf{\con{s}} x\typ{\sortint}.\;\, \lite\bigl( x \leq 0, \con{s}(x)
+\teq 0, \con{s}( x ) \teq x + \con{s}( x-1 ) \bigr) \}$ is admissble.
+As stated
+in the proof of Theorem~\ref{thm:equi}, the reason is that any model of
+formula~(\ref{eq:ex-after}) must satisfy a formula that is closed under
+function expansion, hence implying that formula~(\ref{eq:ex-before}) has a
+model. In more detail, let $\varphi_\con{s}$ be the formula $\forallf{\con{s}} x.\;\,
+\lite\bigl( x \leq 0, \con{s}(x) \teq 0, \con{s}( x ) \teq x + \con{s}( x-1 )
+\bigr)$. The rightmost conjunct $\exists a : {\fargsort{\con{s}}}.
+\farg{\con{s}}( a ) \teq \con{c}$ ensures that any model of
+formula~(\ref{eq:ex-after}) satisfying $\con{s}( \con{c} ) > 100$ also
+satisfies $\varphi_{\con{s}}[\con{c}/x]$. Similarly, the nested existential
+$\exists a.\;\, \farg{\con{s}}( a ) \teq \farg{\con{s}}( y )-1$ ensures that
+this constraint is enforced recursively. Given a model $\I$ of
+formula~(\ref{eq:ex-after}) that interprets $\con{c}$ as $14$, it is easy to see that
+$\I$ must satisfy the following formula $\psi$:
+%
+\begin{equation} \label{eq:fun-closure}
 \begin{split}
-\psi := \con{s}( c ) > 100 \wedge & \displaystyle\bigwedge\limits_{i=0}^{13} ( \neg (c-i \leq 0) \wedge \con{s}( c-i ) \teq i + \con{s}( c-i-1 ) ) \\
-                            & \wedge c-14 \leq 0 \wedge \con{s}( c-14 ) \teq 0
+& \con{s}( \con{c} ) > 100 \wedge %\displaystyle
+  \bigwedge\nolimits_{\vthinspace i=0}^{\!13} \bigl( \neg (\con{c}-i \leq 0) \wedge \con{s}( \con{c}-i ) \teq i + \con{s}( \con{c}-i-1 ) \bigr) \\[-\jot]
+{\wedge}\; & \con{c}-14 \leq 0 \wedge \con{s}( \con{c}-14 ) \teq 0
 \end{split}
-\end{eqnarray}
-Notice that $\psi$ is closed under function expansion, 
-since it contains the $s$-applications that are equivalent to $\con{s}( c-i )$ for $i = 0, \ldots, 14$, and
-entails $\varphi_s[c-i/x]$ for $i = 0, \ldots, 14$.
-Since $\varphi_s$ is an admissible definition,
-this implies there exists a model for $\psi \wedge \varphi_s$,
-which implies there exists a model for formula~(\ref{eq:ex-before}).
+\end{equation}
+%
+This formula is closed under function expansion,
+since it contains the $\con{s}$ applications corresponding to
+$\con{s}( \con{c}-i )$ for $i = 0, \ldots, 14$. The formula also
+entails $\varphi_\con{s}[\con{c}-i/x]$ for $i = 0, \ldots, 14$.
+Since $\varphi_\con{s}$ is an admissible definition,
+there must exist a model for $\psi \wedge \varphi_\con{s}$,
+which means that there exists a model for formula~(\ref{eq:ex-before}).\xend
+\end{example}
 

### 2015-05-05 tex/main.tex
more

@@
 %\newcommand\bigtuple[1]{$(${#1}$)$}
-\newcommand\bigtuple[1]{$\bigl<${#1}$\bigr>$}
+\newcommand\bigtuple[1]{$\bigl(${#1}$\bigr)$}
 
@@
    \begin{itemize}
-     \item[] \return \bigtuple{$t,\, \bigwedge\, \{ \exists a : \fargsort{\con{f}}.\; \vecfarg{\con{f}}( a ) \teq \vec t \mid \con{f}( \vec t ) \in \terms( t ),\, \con{f} \in \sfundefs{\Sigma} \}$}
+     \item[] \return \bigtuple{$t,\; \bigwedge\, \{ \exists a : \fargsort{\con{f}}.\; \vecfarg{\con{f}}( a ) \teq \vec t \mid \con{f}( \vec t ) \in \terms( t ),\, \con{f} \in \sfundefs{\Sigma} \}$}
    \end{itemize}
@@
 By case analysis it follows that $\conv_0$ maintains the invariant that
-if $\conv_0( t\typ{\Bool}, \ppos )$ returns $( r\typ{\Bool}, D )$,
-then $\I$ satisfies $t$ if and only if $\J$ satisfies $r$.
+if $\conv_0( t\typ{\Bool}, \ppos )$ returns $( {t'}, \chi )$,
+then $\I$ satisfies $t$ if and only if $\J$ satisfies $t'$.
 Thus, we conclude $\J$ is a model of $\conv( \varphi )$.

### 2015-05-05 tex/main.tex
more

@@
 $\con{c}$ as $14$,
-and $\con{s}$ as the (almost constant) function:
+and $\con{s}$ as the (almost constant) function
 %
@@
 \begin{equation} \label{eq:approx-interp}
-\lambda x : \sortint.\,\, \lite( x \mathbin{\teq} 0,\, 0,\,
+\lambda x : \sortint.\;\, \lite( x \mathbin{\teq} 0,\, 0,\,
   \lite( x \mathbin{\teq} 1,\, 1,\,

### 2015-05-05 tex/main.tex
more

@@
 \def\wedge{\mathrel{\oldwedge}}
+
+\newcommand\win{\bf}
+\newcommand\bench{\ttfamily}
 
@@
 
-For benchmarks, we gathered a set of 245 benchmarks from two sources, which we will refer to as {\bf isa} and {\bf leon}.
-The first ({\bf isa}) consists of the 79 benchmarks from the Isaplanner benchmark suite~\cite{DBLP:conf/itp/JohanssonDB10} that do not contain higher-order functions.
+For benchmarks, we gathered a set of 245 benchmarks from two sources, which we will refer to as {\bench isa} and {\bench leon}.
+The first ({\bench isa}) consists of the 79 benchmarks from the Isaplanner benchmark suite~\cite{DBLP:conf/itp/JohanssonDB10} that do not contain higher-order functions.
 The benchmarks have been used recently as challenge problems for a variety of inductive theorem provers.
 These benchmarks heavily involve recursive functions, and are limited to the combined theory of uninterpreted functions and inductive datatypes.
-The second ({\bf leon}) consists of 166 benchmarks from the Leon repository~\footnote{Available at \url{https://github.com/epfl-lara/leon}.},
+The second ({\bench leon}) consists of 166 benchmarks from the Leon repository~\footnote{Available at \url{https://github.com/epfl-lara/leon}.},
 which were constructed from verification conditions from simple Scala programs.
@@
 in particular they each are a conjoined list of function definitions followed by a (negated) conjecture.
-A majority of these benchmarks are unsatisfiable, although a handful of the benchmarks in the {\bf leon} set are satisfiable.
+A majority of these benchmarks are unsatisfiable, although a handful of the benchmarks in the {\bench leon} set are satisfiable.
 
@@
 Note that benchmarks created in this way have a high likelihood of having small and easy-to-find counterexamples.
-In total, we considered 213 mutated forms of conjectures from {\bf isa}, and 427 mutated forms of conjectures from {\bf leon}.
-We will call these sets {\bf isa+m} and {\bf leon+m} respectively.
+In total, we considered 213 mutated forms of conjectures from {\bench isa}, and 427 mutated forms of conjectures from {\bench leon}.
+We will call these sets {\bench isa+m} and {\bench leon+m} respectively.
 
@@
 We considered these 885 benchmarks both before and after the translation $\conv$ as described in Section~\ref{sec:encoding}.
-We will denote the latter encoding using the suffix {\bf w}.
-For example, {\bf isa+w} contains 79 benchmarks that are the results of
-applying $\conv$ to each of the 79 benchmarks in {\bf isa}.
+We will denote the latter encoding using the suffix {\bench w}.
+For example, {\bench isa+w} contains 79 benchmarks that are the results of
+applying $\conv$ to each of the 79 benchmarks in {\bench isa}.
 
@@
 as well as methods for finding models for quantified formulas~\cite{GeDeM-CAV-09}.
-We also considered three configurations of \cvc~\cite{barrett-et-al-2011} which we will refer to as {\bf cvc4}, {\bf cvc4+f} and {\bf cvc4+i}.
-The default configuration {\bf cvc4} runs heuristic quantifier instantiation,
+We also considered three configurations of \cvc~\cite{barrett-et-al-2011} which we will refer to as {\bench cvc4}, {\bench cvc4+f} and {\bench cvc4+i}.
+The default configuration {\bench cvc4} runs heuristic quantifier instantiation,
 but does not include techniques for finding models.
-The configuration {\bf cvc4+f} runs heuristic instantiation and the finite model finding procedure described in~\cite{ReyEtAl-1-RR-13, reynolds-et-al-2013}.
-The last configuration {\bf cvc4+i} incorporates techniques for automating inductive reasoning in SMT~\cite{reynolds-kuncak-2015}.
+The configuration {\bench cvc4+f} runs heuristic instantiation and the finite model finding procedure described in~\cite{ReyEtAl-1-RR-13, reynolds-et-al-2013}.
+The last configuration {\bench cvc4+i} incorporates techniques for automating inductive reasoning in SMT~\cite{reynolds-kuncak-2015}.
 %All configurations of \cvc incorporate techniques for conflict-based instantiation~\cite{} ..?
@@
 \hline
-  & \multicolumn{2}{c|}{{\bf z3}}     & \multicolumn{2}{c|}{{\bf cvc4}}     & \multicolumn{2}{c|}{{\bf cvc4+f}}     & \multicolumn{2}{c|}{{\bf cvc4+i}}     & & \multicolumn{2}{c|}{{\bf z3}}     & \multicolumn{2}{c|}{{\bf cvc4}}     & \multicolumn{2}{c|}{{\bf cvc4+f}}     & \multicolumn{2}{c|}{{\bf cvc4+i}}
+  & \multicolumn{2}{c|}{{\bench z3}}     & \multicolumn{2}{c|}{{\bench cvc4}}     & \multicolumn{2}{c|}{{\bench cvc4+f}}     & \multicolumn{2}{c|}{{\bench cvc4+i}}     & & \multicolumn{2}{c|}{{\bench z3}}     & \multicolumn{2}{c|}{{\bench cvc4}}     & \multicolumn{2}{c|}{{\bench cvc4+f}}     & \multicolumn{2}{c|}{{\bench cvc4+i}}
 \\
@@
 \hline
-{\bf isa} & 14  & 0 & 15  & 0 & 15  & 0 & {\bf 61}  & 0 & {\bf isa+w} & 15  & 0 & 15  & 0 & 15  & 0 & 14  & 0
+{\bench isa} & 14  & 0 & 15  & 0 & 15  & 0 & {\win 61}  & 0 & {\bench isa+w} & 15  & 0 & 15  & 0 & 15  & 0 & 14  & 0
 \\
-{\bf leon}  & 73  & 0 & 80  & 0 & 80  & 0 & {\bf 96}  & 0 & {\bf leon+w}  & 78  & 2 & 80  & 0 & 76  & {\bf 9} & 78  & 0
+{\bench leon}  & 73  & 0 & 80  & 0 & 80  & 0 & {\win 96}  & 0 & {\bench leon+w}  & 78  & 2 & 80  & 0 & 76  & {\win 9} & 78  & 0
 \\
-{\bf isa+m} & 17  & 0 & 18  & 0 & 18  & 0 & {\bf 44}  & 0 & {\bf isa+mw}  & 18  & 35  & 18  & 0 & 18  & {\bf 153} & 17  & 0
+{\bench isa+m} & 17  & 0 & 18  & 0 & 18  & 0 & {\win 44}  & 0 & {\bench isa+mw}  & 18  & 35  & 18  & 0 & 18  & {\win 153} & 17  & 0
 \\
-{\bf leon+m}  & 83  & 11  & 103 & 6 & 104 & 6 & {\bf 117} & 6 & {\bf leon+mw} & 98  & 75  & 98  & 6 & 95  & {\bf 169} & 98  & 6
+{\bench leon+m}  & 83  & 11  & 103 & 6 & 104 & 6 & {\win 117} & 6 & {\bench leon+mw} & 98  & 75  & 98  & 6 & 95  & {\win 169} & 98  & 6
 \\
 \hline
-{\bf total} & 187 & 11  & 216 & 6 & 217 & 6 & {\bf 318} & 6 & {\bf total} & 209 & 112 & 211 & 6 & 204 & {\bf 331} & 207 & 6
+{\bench total} & 187 & 11  & 216 & 6 & 217 & 6 & {\win 318} & 6 & {\bench total} & 209 & 112 & 211 & 6 & 204 & {\win 331} & 207 & 6
 \\
@@
 the number of satisfiable responses from all configurations is quite low.
-In total, {\bf z3} manages to find only 11 benchmarks to be satisfiable,
+In total, {\bench z3} manages to find only 11 benchmarks to be satisfiable,
 whereas each configuration of \cvc only finds 6.
@@
 In terms of number of unsatisfiable responses,
-the configuration {\bf cvc4+i} is the clear winner, finding 318 total benchmarks to be unsatisfiable.
+the configuration {\bench cvc4+i} is the clear winner, finding 318 total benchmarks to be unsatisfiable.
 
@@
 we see that using the translation $\conv$ drastically improves the ability of configurations for answering satisfiable.
-In total, {\bf cvc4+f} finds 331 of the 885 benchmarks to be satisfiable, including 9 benchmarks in the original {\bf leon} benchmark set.
-The performance of {\bf z3} for counterexamples also improves drastically, as it solves 112 satisfiable,
-including 10 that were unsolved by {\bf cvc4+f}.
+In total, {\bench cvc4+f} finds 331 of the 885 benchmarks to be satisfiable, including 9 benchmarks in the original {\bench leon} benchmark set.
+The performance of {\bench z3} for counterexamples also improves drastically, as it solves 112 satisfiable,
+including 10 that were unsolved by {\bench cvc4+f}.
 Moreover, the translation $\conv$ helps \ziii for unsatisfiable responses as well,
 as we see \ziii solves a total of 209 whereas it solves only 187 before the translation.
-This number is 2 fewer than {\bf cvc4}, which solves 211 unsatisfiable.
+This number is 2 fewer than {\bench cvc4}, which solves 211 unsatisfiable.
 Thus, in most cases the translation $\conv$ does not significantly degrade performance for unsatisfiable benchmarks,
 and in some cases may actually aid the solver for determining unsatisfiability.%
-\footnote{The exception here is {\bf cvc4+i}, which makes use of type
+\footnote{The exception here is {\bench cvc4+i}, which makes use of type
 information associated with quantified formulas for inductive arguments.}
@@
 for conjectures involving recursive functions whose definitions are admissible.
-Both {\bf z3} and {\bf cvc4+f} were able to find a large number of counterexamples after the translation,
+Both {\bench z3} and {\bench cvc4+f} were able to find a large number of counterexamples after the translation,
 with the latter configuration finding significantly more instances (331 to 112).

### 2015-05-05 tex/main.tex
more

@@
 
-In this section, we evaluate both
-the overall impact of the encoding introduced in the previous section, and
-the performance of individual SMT techniques for benchmarks in the encoding.
+In this section, we evaluate both the overall impact of the encoding introduced
+in the previous section, and the performance of individual SMT techniques for
+benchmarks in the encoding.
 
-For benchmarks, we gathered a set of 245 benchmarks from two sources, which we will refer to as {\bench isa} and {\bench leon}.
-The first ({\bench isa}) consists of the 79 benchmarks from the Isaplanner benchmark suite~\cite{DBLP:conf/itp/JohanssonDB10} that do not contain higher-order functions.
-The benchmarks have been used recently as challenge problems for a variety of inductive theorem provers.
-These benchmarks heavily involve recursive functions, and are limited to the combined theory of uninterpreted functions and inductive datatypes.
-The second ({\bench leon}) consists of 166 benchmarks from the Leon repository~\footnote{Available at \url{https://github.com/epfl-lara/leon}.},
-which were constructed from verification conditions from simple Scala programs.
-These benchmarks also heavily involve recursive functions over inductive datatypes,
-and also cover a wide variety of theories, including bitvectors, arrays, and linear and non-linear arithmetic.
-Each of the 245 benchmarks are of the form described in formula~(\ref{eq:wdf}) from Section~\ref{sec:prelim},
-in particular they each are a conjoined list of function definitions followed by a (negated) conjecture.
-A majority of these benchmarks are unsatisfiable, although a handful of the benchmarks in the {\bench leon} set are satisfiable.
+We gathered a set of 245 benchmarks from two sources, which we will refer to as
+{\bench isa} and {\bench leon}. The first ({\bench isa}) consists of the 79
+benchmarks from the Isaplanner benchmark
+suite~\cite{DBLP:conf/itp/JohanssonDB10} that do not contain higher-order
+functions. The benchmarks have been used recently as challenge problems for a
+variety of inductive theorem provers. These benchmarks heavily involve
+recursive functions, and are limited to the combined theory of uninterpreted
+functions and inductive datatypes. The second ({\bench leon}) consists of 166
+benchmarks from the Leon repository,\footnote{%Available at
+\url{https://github.com/epfl-lara/leon}} which were constructed from
+verification conditions from simple Scala programs. These benchmarks also
+heavily involve recursive functions over inductive datatypes, and also cover a
+wide variety of theories, including bitvectors, arrays, and linear and
+non-linear arithmetic. Each of the 245 benchmarks are of the form described in
+formula~(\ref{eq:wdf}) from Section~\ref{sec:prelim}.
+% in particular they each
+%are a conjoined list of function definitions followed by a (negated)
+%conjecture.
+A majority of these benchmarks are unsatisfiable, although a
+handful of the benchmarks in the {\bench leon} set are satisfiable.
 
@@
 }
-\caption{Number of unsat and sat responses for of all configurations on benchmarks
-before and after the translation $\conv$.}
+\caption{\,Number of unsat (u) and sat (s) responses for of all configurations on benchmarks
+before and after the translation $\conv$}
 \label{fig:results}

### 2015-05-05 tex/main.tex
more

@@
 {
-\begin{tabular}{|l|cc|cc|cc|cc|l|cc|cc|cc|cc|}
-\hline
-  & \multicolumn{2}{c|}{{\bench z3}}     & \multicolumn{2}{c|}{{\bench cvc4}}     & \multicolumn{2}{c|}{{\bench cvc4+f}}     & \multicolumn{2}{c|}{{\bench cvc4+i}}     & & \multicolumn{2}{c|}{{\bench z3}}     & \multicolumn{2}{c|}{{\bench cvc4}}     & \multicolumn{2}{c|}{{\bench cvc4+f}}     & \multicolumn{2}{c|}{{\bench cvc4+i}}
-\\
+\begin{tabular}{l@{\kern0.5em}rr@{\kern0.5em}rr@{\kern0.5em}rr@{\kern0.5em}rr@{\kern0.5em}l@{\kern0.5em}rr@{\kern0.5em}rr@{\kern0.5em}rr@{\kern0.5em}rr}
+  & \multicolumn{2}{c@{\kern0.5em}}{{\bench z3}}     & \multicolumn{2}{c@{\kern0.5em}}{{\bench cvc4}}     & \multicolumn{2}{c@{\kern0.5em}}{{\bench cvc4+f}}     & \multicolumn{2}{c@{\kern0.5em}}{{\bench cvc4+i}}     & & \multicolumn{2}{c@{\kern0.5em}}{{\bench z3}}     & \multicolumn{2}{c@{\kern0.5em}}{{\bench cvc4}}     & \multicolumn{2}{c@{\kern0.5em}}{{\bench cvc4+f}}     & \multicolumn{2}{c@{\kern0.5em}}{{\bench cvc4+i}}
+\\[-.25\jot]
   & u & s & u & s & u & s & u & s & & u & s & u & s & u & s & u & s
 \\
-\hline
+\midrule
 {\bench isa} & 14  & 0 & 15  & 0 & 15  & 0 & {\win 61}  & 0 & {\bench isa+w} & 15  & 0 & 15  & 0 & 15  & 0 & 14  & 0
@@
 \\
-\hline
-{\bench total} & 187 & 11  & 216 & 6 & 217 & 6 & {\win 318} & 6 & {\bench total} & 209 & 112 & 211 & 6 & 204 & {\win 331} & 207 & 6
-\\
-\hline
+Total & 187 & 11  & 216 & 6 & 217 & 6 & {\win 318} & 6 & Total & 209 & 112 & 211 & 6 & 204 & {\win 331} & 207 & 6
 \end{tabular}

### 2015-05-05 tex/main.tex
more

--- 2015-05-05 tex/main.tex
@@
 \begin{figure}[t]
-\centering
-{
-\begin{tabular}{|l|cc|cc|cc|cc|l|cc|cc|cc|cc|}
-\hline
-  & \multicolumn{2}{c|}{{\bench z3}}     & \multicolumn{2}{c|}{{\bench cvc4}}     & \multicolumn{2}{c|}{{\bench cvc4+f}}     & \multicolumn{2}{c|}{{\bench cvc4+i}}     & & \multicolumn{2}{c|}{{\bench z3}}     & \multicolumn{2}{c|}{{\bench cvc4}}     & \multicolumn{2}{c|}{{\bench cvc4+f}}     & \multicolumn{2}{c|}{{\bench cvc4+i}}
+\noindent\hbox{}{\hfill
+\begin{tabular}{l@{\kern0.5em}rr@{\kern0.5em}rr@{\kern0.5em}rr@{\kern0.5em}rr}
+  & \multicolumn{2}{c@{\kern0.5em}}{{\bench z3}}     & \multicolumn{2}{c@{\kern0.5em}}{{\bench cvc4}}     & \multicolumn{2}{c@{\kern0.5em}}{{\bench cvc4+f}}     & \multicolumn{2}{c@{\kern0.5em}}{{\bench cvc4+i}}
+\\[-.25\jot]
+  & \multicolumn{1}{c}{u} & s
+  & u & s
+  & u & s
+  & u & s
 \\
-  & u & s & u & s & u & s & u & s & & u & s & u & s & u & s & u & s
+\midrule
+{\bench isa} & 14  & 0 & 15  & 0 & 15  & 0 & {\win 61}  & 0
 \\
-\hline
-{\bench isa} & 14  & 0 & 15  & 0 & 15  & 0 & {\win 61}  & 0 & {\bench isa+w} & 15  & 0 & 15  & 0 & 15  & 0 & 14  & 0
+{\bench leon}  & 73  & 0 & 80  & 0 & 80  & 0 & {\win 96}  & 0
 \\
-{\bench leon}  & 73  & 0 & 80  & 0 & 80  & 0 & {\win 96}  & 0 & {\bench leon+w}  & 78  & 2 & 80  & 0 & 76  & {\win 9} & 78  & 0
+{\bench isa+m} & 17  & 0 & 18  & 0 & 18  & 0 & {\win 44}  & 0
 \\
-{\bench isa+m} & 17  & 0 & 18  & 0 & 18  & 0 & {\win 44}  & 0 & {\bench isa+mw}  & 18  & 35  & 18  & 0 & 18  & {\win 153} & 17  & 0
+{\bench leon+m}  & 83  & 11  & 103 & 6 & 104 & 6 & {\win 117} & 6
 \\
-{\bench leon+m}  & 83  & 11  & 103 & 6 & 104 & 6 & {\win 117} & 6 & {\bench leon+mw} & 98  & 75  & 98  & 6 & 95  & {\win 169} & 98  & 6
+Total & 187 & 11  & 216 & 6 & 217 & 6 & {\win 318} & 6
+\end{tabular}\hfill\hfill
+\begin{tabular}{l@{\kern0.5em}rr@{\kern0.5em}rr@{\kern0.5em}rr@{\kern0.5em}rr}
+  & \multicolumn{2}{c@{\kern0.5em}}{{\bench z3}}     & \multicolumn{2}{c@{\kern0.5em}}{{\bench cvc4}}     & \multicolumn{2}{c@{\kern0.5em}}{{\bench cvc4+f}}     & \multicolumn{2}{c@{\kern0.5em}}{{\bench cvc4+i}}
+\\[-.25\jot]
+  & u & s & u & s & u & s & u & s
 \\
-\hline
-{\bench total} & 187 & 11  & 216 & 6 & 217 & 6 & {\win 318} & 6 & {\bench total} & 209 & 112 & 211 & 6 & 204 & {\win 331} & 207 & 6
+\midrule
+{\bench isa+w} & 15  & 0 & 15  & 0 & 15  & 0 & 14  & 0
 \\
-\hline
-\end{tabular}
+{\bench leon+w}  & 78  & 2 & 80  & 0 & 76  & {\win 9} & 78  & 0
 \\
+{\bench isa+mw}  & 18  & 35  & 18  & 0 & 18  & {\win 153} & 17  & 0
+\\
+{\bench leon+mw} & 98  & 75  & 98  & 6 & 95  & {\win 169} & 98  & 6
+\\
+Total & 209 & 112 & 211 & 6 & 204 & {\win 331} & 207 & 6
+\end{tabular}\hfill\hbox{}
+
 }

### 2015-05-05 tex/main.tex
work on tables

@@
 \begin{figure}[t]
-\noindent\hbox{}{\hfill
-\begin{tabular}{l@{\kern0.5em}rr@{\kern0.5em}rr@{\kern0.5em}rr@{\kern0.5em}rr}
-  & \multicolumn{2}{c@{\kern0.5em}}{{\bench z3}}     & \multicolumn{2}{c@{\kern0.5em}}{{\bench cvc4}}     & \multicolumn{2}{c@{\kern0.5em}}{{\bench cvc4+f}}     & \multicolumn{2}{c@{\kern0.5em}}{{\bench cvc4+i}}
+\normalsize
+\centering
+\begin{tabular}{l@{\kern1.5em}r@{\kern0.375em}r@{\kern1.5em}r@{\kern0.375em}r@{\kern1.5em}r@{\kern0.375em}r@{\kern1.5em}r@{\kern0.375em}r}
+  & \multicolumn{2}{c@{\kern1.5em}}{{\bench z3}}     & \multicolumn{2}{c@{\kern1.5em}}{{\bench cvc4}}     & \multicolumn{2}{c@{\kern1.5em}}{{\bench cvc4+f}}     & \multicolumn{2}{c@{\kern1.5em}}{{\bench cvc4+i}}
 \\[-.25\jot]
-  & \multicolumn{1}{c}{u} & s
-  & u & s
-  & u & s
-  & u & s
+  & \hfill u \hfill & \hfill s \hfill
+  & \hfill u \hfill & \hfill s \hfill
+  & \hfill u \hfill & \hfill s \hfill
+  & \hfill u \hfill & \hfill s \hfill
 \\
@@
 \\
-{\bench leon+m}  & 83  & 11  & 103 & 6 & 104 & 6 & {\win 117} & 6
-\\
-Total & 187 & 11  & 216 & 6 & 217 & 6 & {\win 318} & 6
-\end{tabular}\hfill\hfill
-\begin{tabular}{l@{\kern0.5em}rr@{\kern0.5em}rr@{\kern0.5em}rr@{\kern0.5em}rr}
-  & \multicolumn{2}{c@{\kern0.5em}}{{\bench z3}}     & \multicolumn{2}{c@{\kern0.5em}}{{\bench cvc4}}     & \multicolumn{2}{c@{\kern0.5em}}{{\bench cvc4+f}}     & \multicolumn{2}{c@{\kern0.5em}}{{\bench cvc4+i}}
+{\bench leon+m\phantom{w}}  & 83  & {\win 11}  & 103 & 6 & 104 & 6 & {\win 117} & 6
+\\[\jot]
+Total & 187 & {\win \phantom{0}11}  & 216 & \phantom{00}6 & 217 & \phantom{00}6 & {\win 318} & \phantom{00}6
+\end{tabular}
+\caption{\,Number of unsat (u) and sat (s) responses on original benchmarks}
+\label{fig:results}
+\end{figure}
+
+\begin{figure}[t]
+\normalsize
+{
+\centering
+\begin{tabular}{l@{\kern1.5em}r@{\kern0.375em}r@{\kern1.5em}r@{\kern0.375em}r@{\kern1.5em}r@{\kern0.375em}r@{\kern1.5em}r@{\kern0.375em}r}
+  & \multicolumn{2}{c@{\kern0.375em}}{{\bench z3}}     & \multicolumn{2}{c@{\kern0.375em}}{{\bench cvc4}}     & \multicolumn{2}{c@{\kern0.375em}}{{\bench cvc4+f}}     & \multicolumn{2}{c@{\kern0.375em}}{{\bench cvc4+i}}
 \\[-.25\jot]
-  & u & s & u & s & u & s & u & s
+  & \hfill u \hfill & \hfill s \hfill
+  & \hfill u \hfill & \hfill s \hfill
+  & \hfill u \hfill & \hfill s \hfill
+  & \hfill u \hfill & \hfill s \hfill
 \\
@@
 {\bench leon+mw} & 98  & 75  & 98  & 6 & 95  & {\win 169} & 98  & 6
-\\
-Total & 209 & 112 & 211 & 6 & 204 & {\win 331} & 207 & 6
-\end{tabular}\hfill\hbox{}
+\\[\jot]
+Total & 209 & 112 & 211 & \phantom{00}6 & 204 & {\win 331} & 207 & \phantom{00}6
+\end{tabular}
 
 }
-\caption{\,Number of unsat (u) and sat (s) responses for of all configurations on benchmarks
-before and after the translation $\conv$}
+\caption{\,Number of unsat (u) and sat (s) responses on translated benchmarks}
 \label{fig:results}

### 2015-05-05 tex/main.tex
more

@@
   & \multicolumn{2}{c@{\kern1.5em}}{{\bench z3}}     & \multicolumn{2}{c@{\kern1.5em}}{{\bench cvc4}}     & \multicolumn{2}{c@{\kern1.5em}}{{\bench cvc4+f}}     & \multicolumn{2}{c@{\kern1.5em}}{{\bench cvc4+i}}
-\\[-.25\jot]
-  & \hfill u \hfill & \hfill s \hfill
-  & \hfill u \hfill & \hfill s \hfill
-  & \hfill u \hfill & \hfill s \hfill
-  & \hfill u \hfill & \hfill s \hfill
+\\[-1pt]
+  & \hfill $\varphi$ \hfill & $\conv(\varphi)$\!\!
+  & \hfill $\varphi$ \hfill & $\conv(\varphi)$\!\!
+  & \hfill $\varphi$ \hfill & $\conv(\varphi)$\!\!
+  & \hfill $\varphi$ \hfill & $\conv(\varphi)$\!\!
 \\
 \midrule
-{\bench isa} & 14  & 0 & 15  & 0 & 15  & 0 & {\win 61}  & 0
+{\bench isa} & 14  & 15 & 15  & 15 & 15  & 15 & {\win 61}  & 14
 \\
-{\bench leon}  & 73  & 0 & 80  & 0 & 80  & 0 & {\win 96}  & 0
+{\bench leon}  & 73  & 78 & 80  & 80 & 80  & 76 & {\win 96}  & 78
 \\
-{\bench isa+m} & 17  & 0 & 18  & 0 & 18  & 0 & {\win 44}  & 0
+{\bench isa+m} & 17  & 18 & 18  & 18 & 18  & 18 & {\win 44}  & 17
 \\
-{\bench leon+m\phantom{w}}  & 83  & {\win 11}  & 103 & 6 & 104 & 6 & {\win 117} & 6
+{\bench leon+m\phantom{w}}  & 83  & 98 & 103 & 98 & 104 & 95 & {\win 117} & 98
 \\[\jot]
-Total & 187 & {\win \phantom{0}11}  & 216 & \phantom{00}6 & 217 & \phantom{00}6 & {\win 318} & \phantom{00}6
+Total & 187 & 209 & 216 & 211 & 217 & 204 & {\win 318} & 217
 \end{tabular}
-\caption{\,Number of unsat (u) and sat (s) responses on original benchmarks}
+\caption{\,Number of ``unsatisfiable'' responses on benchmarks without and with translation with $\conv$}
 \label{fig:results}
@@
   & \multicolumn{2}{c@{\kern0.375em}}{{\bench z3}}     & \multicolumn{2}{c@{\kern0.375em}}{{\bench cvc4}}     & \multicolumn{2}{c@{\kern0.375em}}{{\bench cvc4+f}}     & \multicolumn{2}{c@{\kern0.375em}}{{\bench cvc4+i}}
-\\[-.25\jot]
-  & \hfill u \hfill & \hfill s \hfill
-  & \hfill u \hfill & \hfill s \hfill
-  & \hfill u \hfill & \hfill s \hfill
-  & \hfill u \hfill & \hfill s \hfill
+\\[-1pt]
+  & \hfill $\varphi$ \hfill & $\conv(\varphi)$\!\!
+  & \hfill $\varphi$ \hfill & $\conv(\varphi)$\!\!
+  & \hfill $\varphi$ \hfill & $\conv(\varphi)$\!\!
+  & \hfill $\varphi$ \hfill & $\conv(\varphi)$\!\!
 \\
 \midrule
-{\bench isa+w} & 15  & 0 & 15  & 0 & 15  & 0 & 14  & 0
+{\bench isa} & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0
 \\
-{\bench leon+w}  & 78  & 2 & 80  & 0 & 76  & {\win 9} & 78  & 0
+{\bench leon}  & 0 & 2 & 0 & 0 & 0 & {\win 9} & 0 & 0
 \\
-{\bench isa+mw}  & 18  & 35  & 18  & 0 & 18  & {\win 153} & 17  & 0
+{\bench isa+m} & 0 & 35 & 0 & 0 & 0 & {\win 153} & 0 & 0
 \\
-{\bench leon+mw} & 98  & 75  & 98  & 6 & 95  & {\win 169} & 98  & 6
+{\bench leon+m}  & 11  & 75 & 6 & 6 & 6 & {\win 169} & 6 & 6
 \\[\jot]
-Total & 209 & 112 & 211 & \phantom{00}6 & 204 & {\win 331} & 207 & \phantom{00}6
+Total & \phantom{0}11  & 112 & \phantom{00}6 & \phantom{00}6 & \phantom{00}6 & {\win 331} & \phantom{00}6 & \phantom{00}6
 \end{tabular}
@@
 }
-\caption{\,Number of unsat (u) and sat (s) responses on translated benchmarks}
+\caption{\,Number of ``satisfiable'' responses on benchmarks without and with translation with $\conv$}
 \label{fig:results}

### 2015-05-05 tex/main.tex
rearranged tables

@@
 \begin{tabular}{l@{\kern1.5em}r@{\kern0.375em}r@{\kern1.5em}r@{\kern0.375em}r@{\kern1.5em}r@{\kern0.375em}r@{\kern1.5em}r@{\kern0.375em}r}
+  & \multicolumn{2}{c@{\kern0.375em}}{{\bench z3}}     & \multicolumn{2}{c@{\kern0.375em}}{{\bench cvc4}}     & \multicolumn{2}{c@{\kern0.375em}}{{\bench cvc4+f}}     & \multicolumn{2}{c@{\kern0.375em}}{{\bench cvc4+i}}
+\\%[-1pt]
+  & \hfill $\varphi$ \hfill & $\conv(\varphi)$\!\!
+  & \hfill $\varphi$ \hfill & $\conv(\varphi)$\!\!
+  & \hfill $\varphi$ \hfill & $\conv(\varphi)$\!\!
+  & \hfill $\varphi$ \hfill & $\conv(\varphi)$\!\!
+\\
+\midrule
+{\bench isa} & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0
+\\
+{\bench leon}  & 0 & 2 & 0 & 0 & 0 & {\win 9} & 0 & 0
+\\
+{\bench isa+m} & 0 & 35 & 0 & 0 & 0 & {\win 153} & 0 & 0
+\\
+{\bench leon+m}  & 11  & 75 & 6 & 6 & 6 & {\win 169} & 6 & 6
+\\[\jot]
+Total & \phantom{0}11  & 112 & \phantom{00}6 & \phantom{00}6 & \phantom{00}6 & {\win 331} & \phantom{00}6 & \phantom{00}6
+\end{tabular}
+\caption{\,Number of ``satisfiable'' responses on benchmarks without and with translation with $\conv$}
+\label{fig:sat}
+\end{figure}
+
+\begin{figure}[t]
+\normalsize
+\centering
+\begin{tabular}{l@{\kern1.5em}r@{\kern0.375em}r@{\kern1.5em}r@{\kern0.375em}r@{\kern1.5em}r@{\kern0.375em}r@{\kern1.5em}r@{\kern0.375em}r}
   & \multicolumn{2}{c@{\kern1.5em}}{{\bench z3}}     & \multicolumn{2}{c@{\kern1.5em}}{{\bench cvc4}}     & \multicolumn{2}{c@{\kern1.5em}}{{\bench cvc4+f}}     & \multicolumn{2}{c@{\kern1.5em}}{{\bench cvc4+i}}
-\\[-1pt]
+\\%[-1pt]
   & \hfill $\varphi$ \hfill & $\conv(\varphi)$\!\!
@@
 \caption{\,Number of ``unsatisfiable'' responses on benchmarks without and with translation with $\conv$}
-\label{fig:results}
+\label{fig:unsat}
 \end{figure}
 
-\begin{figure}[t]
-\normalsize
-{
-\centering
-\begin{tabular}{l@{\kern1.5em}r@{\kern0.375em}r@{\kern1.5em}r@{\kern0.375em}r@{\kern1.5em}r@{\kern0.375em}r@{\kern1.5em}r@{\kern0.375em}r}
-  & \multicolumn{2}{c@{\kern0.375em}}{{\bench z3}}     & \multicolumn{2}{c@{\kern0.375em}}{{\bench cvc4}}     & \multicolumn{2}{c@{\kern0.375em}}{{\bench cvc4+f}}     & \multicolumn{2}{c@{\kern0.375em}}{{\bench cvc4+i}}
-\\[-1pt]
-  & \hfill $\varphi$ \hfill & $\conv(\varphi)$\!\!
-  & \hfill $\varphi$ \hfill & $\conv(\varphi)$\!\!
-  & \hfill $\varphi$ \hfill & $\conv(\varphi)$\!\!
-  & \hfill $\varphi$ \hfill & $\conv(\varphi)$\!\!
-\\
-\midrule
-{\bench isa} & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0
-\\
-{\bench leon}  & 0 & 2 & 0 & 0 & 0 & {\win 9} & 0 & 0
-\\
-{\bench isa+m} & 0 & 35 & 0 & 0 & 0 & {\win 153} & 0 & 0
-\\
-{\bench leon+m}  & 11  & 75 & 6 & 6 & 6 & {\win 169} & 6 & 6
-\\[\jot]
-Total & \phantom{0}11  & 112 & \phantom{00}6 & \phantom{00}6 & \phantom{00}6 & {\win 331} & \phantom{00}6 & \phantom{00}6
-\end{tabular}
-
-}
-\caption{\,Number of ``satisfiable'' responses on benchmarks without and with translation with $\conv$}
-\label{fig:results}
-\end{figure}
-
-The results are shown in Figure~\ref{fig:results}.
-The results are divided into benchmarks before and after the translation $\conv$.
+The results are shown in Figures \ref{fig:sat} and \ref{fig:sat}.
+The results are divided into benchmarks triggering ``unsatisfiable'' and ``satisfiable''
+responses and further into benchmarks before and after the translation $\conv$.
+\jb{rephrase}
 First, we note that in no cases did a configuration answer differently across the two encodings.
-In other words, there were no cases in which \ziii or \cvc answered e.g.\ ``unsatisfiable" on a benchmark $\varphi$
-and ``satisfiable" on its corresponding benchmark $\conv( \varphi )$, or vice versa.
+In other words, there were no cases in which \ziii or \cvc answered e.g.\ ``unsatisfiable'' on a benchmark $\varphi$
+and ``satisfiable'' on its corresponding benchmark $\conv( \varphi )$, or vice versa.
 This provides confirming evidence that the translation $\conv$ is \emph{refutation-sound},

### 2015-05-05 tex/main.tex
more

@@
 \centering
-\begin{tabular}{l@{\kern1.5em}r@{\kern0.375em}r@{\kern1.5em}r@{\kern0.375em}r@{\kern1.5em}r@{\kern0.375em}r@{\kern1.5em}r@{\kern0.375em}r}
+\begin{tabular}{l@{\kern1.5em}r@{\kern0.375em}r@{\kern1.5em}r@{\kern0.375em}r@{\kern1.5em}r@{\kern0.375em}r@{\kern1.5em}r@{\kern0.375em}r@{\,\,}}
   & \multicolumn{2}{c@{\kern0.375em}}{{\bench z3}}     & \multicolumn{2}{c@{\kern0.375em}}{{\bench cvc4}}     & \multicolumn{2}{c@{\kern0.375em}}{{\bench cvc4+f}}     & \multicolumn{2}{c@{\kern0.375em}}{{\bench cvc4+i}}
@@
 \centering
-\begin{tabular}{l@{\kern1.5em}r@{\kern0.375em}r@{\kern1.5em}r@{\kern0.375em}r@{\kern1.5em}r@{\kern0.375em}r@{\kern1.5em}r@{\kern0.375em}r}
-  & \multicolumn{2}{c@{\kern1.5em}}{{\bench z3}}     & \multicolumn{2}{c@{\kern1.5em}}{{\bench cvc4}}     & \multicolumn{2}{c@{\kern1.5em}}{{\bench cvc4+f}}     & \multicolumn{2}{c@{\kern1.5em}}{{\bench cvc4+i}}
+\begin{tabular}{l@{\kern1.5em}r@{\kern0.375em}r@{\kern1.5em}r@{\kern0.375em}r@{\kern1.5em}r@{\kern0.375em}r@{\kern1.5em}r@{\kern0.375em}r@{\,\,}}
+  & \multicolumn{2}{c@{\kern0.375em}}{{\bench z3}}     & \multicolumn{2}{c@{\kern0.375em}}{{\bench cvc4}}     & \multicolumn{2}{c@{\kern0.375em}}{{\bench cvc4+f}}     & \multicolumn{2}{c@{\kern0.375em}}{{\bench cvc4+i}}
 \\%[-1pt]
@@
 \\
-{\bench leon+m\phantom{w}}  & 83  & 98 & 103 & 98 & 104 & 95 & {\win 117} & 98
+{\bench leon+m}  & 83  & 98 & 103 & 98 & 104 & 95 & {\win 117} & 98
 \\[\jot]

### 2015-05-05 tex/main.tex
more

@@
 
+\begin{verbatim}
 first-order model finders: SEM, MACE/Mace, Paradox, FMDarwin, iProver
@@
 Nitpick
+  * nice: three-valued logic (e.g. local overflow)
+    * solving a different problem:
+      * no built-in notion, what's approximated is an idealized notion, no
+        injections $\concret$
+  * not so nice:
+    one abstract domain per type, not per function argument
+  * guards + one-domain per arg are a good combi, guards + one-single-domain is bad because of
+    odd schemes, give unsat example
+
+  * other features:
     * inductive / coinductive predicates
@@
 in smt : z3, cvc4 (briefly, text in main section gives detail)
+\end{verbatim}
 
@@
 
+\begin{verbatim}
   * future work
-    * (co)inductive predicates?
-    * quotient types??
+    * more eval
+      * against other non-SMT-solvers (notably Leon and Nitpick)
+      * problems exercising codatatypes (which CVC4 supports \cite{xxx}) and
+        corecursive functions
+    * more features
+      * e.g. encode (co)inductive predicates as (co)recursive functions
+\end{verbatim}
 

### 2015-05-06 tex/main.tex
Minor fixes/comments.

@@
 find a model for the formula $\forall x\typ{\tau} y\typ{\tau}.\;\, x \teq
-y \vee \neg \con{f}( x ) \teq \con{f}( y )$ where the uninterpreted type $\tau$ is
+y \vee \neg \con{f}( x ) \teq \con{f}( y )$ 
+\ajr{Need to settle on notation for types (currently we are using both superscript and colon).}
+where the uninterpreted type $\tau$ is
 interpreted as a finite set and $\con{f}$ as injective finite map.
@@
 
-Here, we advocate an approach that translates formulas involving function
+In this paper, we advocate a method for translating formulas involving function
 definitions (such as the one mentioned in the previous paragraph) 
-into equisatisfiable \ct{to be defined} formulas where the
-techniques~\cite{GeDeM-CAV-09, ReyEtAl-1-RR-13} are applicable.
+into ones where the techniques~\cite{GeDeM-CAV-09, ReyEtAl-1-RR-13} are applicable.
 
@@
 The intuition behind this translation
-is to use an uninterpreted type \ct{to be defined} to abstract the set of
+is to use an uninterpreted type to abstract the set of
 \relax{relevant} tuples for each recursive function $\con{f}$ and restrict the
@@
       %\item[] $\mathsf{assert}( \chi = \ltrue )$;
-      \item[] \return \bigtuple{$\forall y\typ{\fargsort{\con{f}}}.\; u' [ \vecfarg{\con{f}}( y ) / \vec x ],\, \ltrue$}
+      \item[] \return \bigtuple{$\forall y : \fargsort{\con{f}}.\; u' [ \vecfarg{\con{f}}( y ) / \vec x ],\, \ltrue$}
     \end{itemize}
@@
     \!\begin{aligned}[t]
-      \forall a : \fargsort{\con{s}}.\; & \phantom{{\wedge}\;}\lite\bigl( \farg{\con{s}}( a ) \leq 0,\;
+      \forall a : \fargsort{\con{s}}.\; \lite\bigl( \farg{\con{s}}( a ) \leq 0,\;
         \con{s}(\farg{\con{s}}( a )) \teq 0,\;
-        \con{s}(\farg{\con{s}}( a )) \teq \farg{\con{s}}( a ) + \con{s}( \farg{\con{s}}( a )-1 ) \bigr)
+        & \phantom{{\wedge}\;} \con{s}(\farg{\con{s}}( a )) \teq \farg{\con{s}}( a ) + \con{s}( \farg{\con{s}}( a )-1 )
 \\[-\jot]
-      & {\wedge}\; \exists a : {\fargsort{\con{s}}}.\;\, \farg{\con{s}}( a ) \teq \farg{\con{s}}( a )-1 \bigr)
+      & {\wedge}\; \exists b : {\fargsort{\con{s}}}.\;\, \farg{\con{s}}( b ) \teq \farg{\con{s}}( a )-1 \bigr) \bigr)
 \end{aligned}
@@
 %%% Added "apparently". Or are we sure?  
-\cvc and \ziii apparently run indefinitely on formula~(\ref{eq:ex-before}), and both
+% Yes, I'm planning to post the 2 versions of the benchmark on the evaluation page (when we add it) --ajr
+\cvc and \ziii run indefinitely on formula~(\ref{eq:ex-before}), and both
 produce a model for formula~(\ref{eq:ex-after}) within 100 milliseconds.\xend
@@
 %%% @ANDY: I don't get the following comment --jb
-%
+%   @JASMIN : The point of this comment is that the user is not interested in the SMT solver inferring any new information regarding the interpretation of s,
+%             since s is a "defined" function, the user already has an intended interpretation in their head of what s is (here, the sum of positive integers between 0...x).
+%             Thus, it isn't a big deal that the SMT solver "approximates" the interpretation of s during a check, since we already know what s is.
 %From a practical point of view, this is not an
@@
 formula~(\ref{eq:ex-after}) satisfying $\con{s}( \con{c} ) > 100$ also
-satisfies $\varphi_{\con{s}}[\con{c}/x]$. Similarly, the nested existential
-$\exists a.\;\, \farg{\con{s}}( a ) \teq \farg{\con{s}}( y )-1$ ensures that
-this constraint is enforced recursively. Given a model $\I$ of
+satisfies $\varphi_{\con{s}}[\con{c}/x]$. The nested existential
+$\exists b.\;\, \farg{\con{s}}( b ) \teq \farg{\con{s}}( a )-1$ ensures that
+a similar constraint is enforced recursively: any model satisfying $\varphi_{\con{s}}[t/x]$
+must also satisfy $\varphi_{\con{s}}[t-1/x]$ when $t > 0$. Given a model $\I$ of
 formula~(\ref{eq:ex-after}) that interprets $\con{c}$ as $14$, it is easy to see that
+due to the structure of formula~(\ref{eq:ex-after}),
 $\I$ must satisfy the following formula $\psi$:
@@
 there must exist a model for $\psi \wedge \varphi_\con{s}$,
-which means that there exists a model for formula~(\ref{eq:ex-before}).\xend
+which means there exists a model for formula~(\ref{eq:ex-before}).\xend
 \end{example}

### 2015-05-06 tex/main.tex
merge

@@
 
+\newcommand\cvcd{CVC4d\xspace}
+\newcommand\cvcf{CVC4f\xspace}
+\newcommand\cvci{CVC4i\xspace}
+\newcommand\ziiib{\ziii\xspace}
+
 %\newcommand{\DDD}{\mathcalx{D}}
@@
 \let\oldneg=\neg
-\def\neg{\oldneg\:}
+\def\neg{\oldneg\;}
 
@@
 
-Recent techniques~\cite{GeDeM-CAV-09, ReyEtAl-1-RR-13}
+Recent techniques~\cite{GeDeM-CAV-09, ReyEtAl-1-RR-13,reynolds-et-al-2013}
 have focused on finding models for quantified formulas in SMT.
@@
 arithmetic predicates. The techniques in~Reynolds et al.\
-\cite{ReyEtAl-1-RR-13} can be used for finding finite models for quantified
+\cite{ReyEtAl-1-RR-13,reynolds-et-al-2013} can be used for finding finite
+models for quantified
 formulas over uninterpreted types as well as types having a fixed finite
@@
 Finite model finding techniques can
-find a model for the formula $\forall x\typ{\tau} y\typ{\tau}.\;\, x \teq
-y \vee \neg \con{f}( x ) \teq \con{f}( y )$ 
-\ajr{Need to settle on notation for types (currently we are using both superscript and colon).}
-where the uninterpreted type $\tau$ is
+find a model for the formula $\forall x,\, y : \tau.\;\, x \teq
+y \vee \neg \con{f}( x ) \teq \con{f}( y )$ where the uninterpreted type $\tau$ is
 interpreted as a finite set and $\con{f}$ as injective finite map.
@@
 In particular, the fragments in~\cite{GeDeM-CAV-09} are not applicable when the argument of a recursive function is used in various ways (e.g. within an arithmetic expression),
-while the fragment in~\cite{ReyEtAl-1-RR-13} is not applicable for functions over infinite domains such as the integers.
+while the fragment in~\cite{ReyEtAl-1-RR-13,reynolds-et-al-2013} is not applicable for functions over infinite domains such as the integers.
 So, for instance, the quantified formula $\forall x : {\sortint}.\;\, \lite\bigl( p \leq 0, p( x ) \teq 0, p( x ) \teq 2 * p( x - 1 ) \bigr)$
@@
 \item[-] We define a satisfiability-preserving translation method for a class of formulas containing recursive function definitions
-that produces formulas for which existing model finding techniques~\cite{GeDeM-CAV-09, ReyEtAl-1-RR-13} are applicable.
+that produces formulas for which existing model finding techniques~\cite{GeDeM-CAV-09, ReyEtAl-1-RR-13,reynolds-et-al-2013} are applicable.
 \item[-] We provide experimental evidence that this translation improves the effectiveness of SMT solvers \cvc and \ziii for finding counterexamples to verification conditions.
@@
 equal to $14$. Due to the universal quantifier,
-current SMT techniques~\cite{GeDeM-CAV-09, ReyEtAl-1-RR-13} are unable to find
+current SMT techniques~\cite{GeDeM-CAV-09, ReyEtAl-1-RR-13,reynolds-et-al-2013} are unable to find
 a model for this formula. The signature $\extendsig{\Sigma}$ includes the type
@@
 type $\fargsort{\con{s}}$, thus enabling the finite model finding
-techniques by Reynolds et al.\ \cite{ReyEtAl-1-RR-13} to search for a finite interpretation for
+techniques by Reynolds et al.\ \cite{ReyEtAl-1-RR-13,reynolds-et-al-2013} to search for a finite interpretation for
 $\fargsort{\con{s}}$. Moreover, all occurrences of $a$ in this formula are beneath
@@
 We considered these 885 benchmarks both before and after the translation $\conv$ as described in Section~\ref{sec:encoding}.
-We will denote the latter encoding using the suffix {\bench w}.
-For example, {\bench isa+w} contains 79 benchmarks that are the results of
-applying $\conv$ to each of the 79 benchmarks in {\bench isa}.
+For example, {\bench isa} contains 79 original benchmarks $\varphi$
+and 79 translated benchmarks $\conv(\varphi)$.
 
@@
 as well as methods for finding models for quantified formulas~\cite{GeDeM-CAV-09}.
-We also considered three configurations of \cvc~\cite{barrett-et-al-2011} which we will refer to as {\bench cvc4}, {\bench cvc4+f} and {\bench cvc4+i}.
-The default configuration {\bench cvc4} runs heuristic quantifier instantiation,
+We also considered three configurations of \cvc~\cite{barrett-et-al-2011} which we will refer to as \cvcd, \cvcf, and \cvci.
+The default configuration \cvcd runs heuristic quantifier instantiation,
 but does not include techniques for finding models.
-The configuration {\bench cvc4+f} runs heuristic instantiation and the finite model finding procedure described in~\cite{ReyEtAl-1-RR-13, reynolds-et-al-2013}.
-The last configuration {\bench cvc4+i} incorporates techniques for automating inductive reasoning in SMT~\cite{reynolds-kuncak-2015}.
+The configuration \cvcf runs heuristic instantiation and the finite model finding procedure described in~\cite{ReyEtAl-1-RR-13,reynolds-et-al-2013}.
+The last configuration \cvci incorporates techniques for automating inductive reasoning in SMT~\cite{reynolds-kuncak-2015}.
 %All configurations of \cvc incorporate techniques for conflict-based instantiation~\cite{} ..?
@@
 \begin{tabular}{l@{\kern1.5em}r@{\kern0.375em}r@{\kern1.5em}r@{\kern0.375em}r@{\kern1.5em}r@{\kern0.375em}r@{\kern1.5em}r@{\kern0.375em}r@{\,\,}}
-  & \multicolumn{2}{c@{\kern0.375em}}{{\bench z3}}     & \multicolumn{2}{c@{\kern0.375em}}{{\bench cvc4}}     & \multicolumn{2}{c@{\kern0.375em}}{{\bench cvc4+f}}     & \multicolumn{2}{c@{\kern0.375em}}{{\bench cvc4+i}}
+  & \multicolumn{2}{c@{\kern1.5em}}{\phantom{0}\ziiib}     & \multicolumn{2}{c@{\kern1.5em}}{\phantom{0}\cvcd}     & \multicolumn{2}{c@{\kern1.5em}}{\phantom{0}\cvcf}     & \multicolumn{2}{c@{\,\,}}{\phantom{0}\cvci}
 \\%[-1pt]
@@
 
-The results are shown in Figures \ref{fig:sat} and \ref{fig:sat}.
+The results are shown in Figures \ref{fig:sat} and \ref{fig:unsat}.
 The results are divided into benchmarks triggering ``unsatisfiable'' and ``satisfiable''
@@
 the number of satisfiable responses from all configurations is quite low.
-In total, {\bench z3} manages to find only 11 benchmarks to be satisfiable,
+In total, \ziiib manages to find only 11 benchmarks to be satisfiable,
 whereas each configuration of \cvc only finds 6.
@@
 In terms of number of unsatisfiable responses,
-the configuration {\bench cvc4+i} is the clear winner, finding 318 total benchmarks to be unsatisfiable.
+the configuration \cvci is the clear winner, finding 318 total benchmarks to be unsatisfiable.
 
@@
 we see that using the translation $\conv$ drastically improves the ability of configurations for answering satisfiable.
-In total, {\bench cvc4+f} finds 331 of the 885 benchmarks to be satisfiable, including 9 benchmarks in the original {\bench leon} benchmark set.
-The performance of {\bench z3} for counterexamples also improves drastically, as it solves 112 satisfiable,
-including 10 that were unsolved by {\bench cvc4+f}.
+In total, \cvcf finds 331 of the 885 benchmarks to be satisfiable, including 9 benchmarks in the original {\bench leon} benchmark set.
+The performance of \ziiib for counterexamples also improves drastically, as it solves 112 satisfiable,
+including 10 that were unsolved by \cvcf.
 Moreover, the translation $\conv$ helps \ziii for unsatisfiable responses as well,
 as we see \ziii solves a total of 209 whereas it solves only 187 before the translation.
-This number is 2 fewer than {\bench cvc4}, which solves 211 unsatisfiable.
+This number is 2 fewer than \cvcd, which solves 211 unsatisfiable.
 Thus, in most cases the translation $\conv$ does not significantly degrade performance for unsatisfiable benchmarks,
 and in some cases may actually aid the solver for determining unsatisfiability.%
-\footnote{The exception here is {\bench cvc4+i}, which makes use of type
+\footnote{The exception here is \cvci, which makes use of type
 information associated with quantified formulas for inductive arguments.}
@@
 for conjectures involving recursive functions whose definitions are admissible.
-Both {\bench z3} and {\bench cvc4+f} were able to find a large number of counterexamples after the translation,
+Both \ziiib and \cvcf were able to find a large number of counterexamples after the translation,
 with the latter configuration finding significantly more instances (331 to 112).
@@
 for finding counterexamples for formulas involving recursive function definitions.
-For convienience, the translation $\conv$ has been implemented in the latest development version of \cvc~\cite{ReyEtAl-1-RR-13} (version 1.5 pre-release).
+For convienience, the translation $\conv$ has been implemented in the latest development version of \cvc %~\cite{ReyEtAl-1-RR-13}
+(version 1.5 pre-release).
 Function definitions $\forallf{\con{f}} \vec x.\; \varphi$ can be written in using the $\definefunreccmd$ command,

### 2015-05-06 tex/main.tex
more

@@
 \newcommand\cvci{CVC4i\xspace}
-\newcommand\ziiib{\ziii\xspace}
+\newcommand\ziiib{\ziii}
 
@@
 formulas over uninterpreted types as well as types having a fixed finite
-interpretation, such as fixed-width bitvectors and enumerated datatypes.
+interpretation, such as fixed-width bit vectors and enumerated datatypes.
 Finite model finding techniques can
@@
 and a function $\lite : \Bool \times \tau \times \tau \rightarrow \tau$ for each $\tau \in \ssorts{\Sigma}$.
-We write $t\typ{\tau}$ to denote a term $t$ of type~$\tau$ and
-$\terms( t )$ to denote the set of subterms in $t$.
+We write $t\typ{\tau}$ to denote a metavariable ranging over terms of
+type~$\tau$ and $\terms( t )$ to denote the set of subterms in $t$.
 %We write $\functerms^\Sigma( t )$ to denote the set of $f$-applications in $t$ such that $f \in \sfundefs{\Sigma}$.
@@
 heavily involve recursive functions over inductive datatypes, and also cover a
-wide variety of theories, including bitvectors, arrays, and linear and
-non-linear arithmetic. Each of the 245 benchmarks are of the form described in
-formula~(\ref{eq:wdf}) from Section~\ref{sec:prelim}.
+wide variety of theories, including bit vectors, arrays, and linear and
+nonlinear arithmetic. Each benchmark is in definitional form with respect to a
+set of terminating functions.
+%
 % in particular they each
@@
 \jb{rephrase}
-First, we note that in no cases did a configuration answer differently across the two encodings.
-In other words, there were no cases in which \ziii or \cvc answered e.g.\ ``unsatisfiable'' on a benchmark $\varphi$
-and ``satisfiable'' on its corresponding benchmark $\conv( \varphi )$, or vice versa.
-This provides confirming evidence that the translation $\conv$ is \emph{refutation-sound},
-as we did not find a case where $\conv( \varphi )$ was unsatisfiable when $\varphi$ was satisfiable.
-It also provides confirming evidence that the translation is \emph{model-sound} for this set of benchmarks,
-as we did not find a case where $\conv( \varphi )$ was satisfiable when $\varphi$ was unsatisfiable.
-This is to be expected, since the functions in our benchmark sets are expected to be admissible (and in fact, terminating).
+%
+The raw evaluation data reveals no cases in which a solver answered
+``unsatisfiable'' on a benchmark $\varphi$ and ``satisfiable'' on its
+corresponding benchmark $\conv( \varphi )$, or vice versa, thereby
+corroborating Theorem~\ref{thm:equi}.
 
-Examining the first encoding in isolation (the first nine columns),
-the number of satisfiable responses from all configurations is quite low.
-In total, \ziiib manages to find only 11 benchmarks to be satisfiable,
-whereas each configuration of \cvc only finds 6.
-This confirms the shortcomings of existing SMT techniques for finding counterexamples for benchmarks containing recursive functions.
-In terms of number of unsatisfiable responses,
-the configuration \cvci is the clear winner, finding 318 total benchmarks to be unsatisfiable.
+Figure~\ref{fig:sat} shows that for untranslated benchmarks (the ``$\varphi$''
+columns), the number of satisfiable responses is very low, across all
+configurations. This confirms the shortcomings of existing SMT techniques for
+finding counterexamples for benchmarks containing recursive functions.
+%
+The translation $\conv$ (the ``$\conv(\varphi)$'' columns) has a major
+impact. \cvcf finds 331 of the 885 benchmarks to be satisfiable,
+including 9~benchmarks in the unmutated {\bench leon} benchmark set. The
+performance of \ziiib for counterexamples also improves dramatically, as it
+solves 112 satisfiable problems, including 10 that were unsolved by \cvcf.
+%
+We conclude that the translation $\conv$ enables SMT
+solvers to find counterexamples for conjectures involving recursive functions
+whose definitions are admissible.
+%Both \ziiib and \cvcf were able to find a
+%large number of counterexamples after the translation, with the latter
+%configuration finding significantly more instances (331 to 112).
 
-Now, examining the second encoding (the latter nine columns),
-we see that using the translation $\conv$ drastically improves the ability of configurations for answering satisfiable.
-In total, \cvcf finds 331 of the 885 benchmarks to be satisfiable, including 9 benchmarks in the original {\bench leon} benchmark set.
-The performance of \ziiib for counterexamples also improves drastically, as it solves 112 satisfiable,
-including 10 that were unsolved by \cvcf.
-Moreover, the translation $\conv$ helps \ziii for unsatisfiable responses as well,
-as we see \ziii solves a total of 209 whereas it solves only 187 before the translation.
-This number is 2 fewer than \cvcd, which solves 211 unsatisfiable.
-Thus, in most cases the translation $\conv$ does not significantly degrade performance for unsatisfiable benchmarks,
-and in some cases may actually aid the solver for determining unsatisfiability.%
-\footnote{The exception here is \cvci, which makes use of type
-information associated with quantified formulas for inductive arguments.}
-
-We conclude from this evaluation that the translation $\conv$ enables SMT solvers to find counterexamples
-for conjectures involving recursive functions whose definitions are admissible.
-Both \ziiib and \cvcf were able to find a large number of counterexamples after the translation,
-with the latter configuration finding significantly more instances (331 to 112).
+In terms of number of unsatisfiable responses, the configuration \cvci is the
+clear winner, finding 318 total benchmarks to be unsatisfiable. Moreover, the
+translation $\conv$ helps \ziii for ``unsatisfiable'' responses as well, as we
+see \ziii solves a total of 209 whereas it solves only 187 before the
+translation.
+%
+This number is 2 fewer than \cvcd, which solves 211 unsatisfiable. Thus, in
+most cases the translation $\conv$ does not significantly degrade performance
+for unsatisfiable benchmarks, and in some cases may actually aid the solver for
+determining unsatisfiability. The exception here is \cvci, whose inductive
+arguments are guided by the types of quantified variables.
 

### 2015-05-06 tex/main.tex
Minor changes/addition to front-end section.

@@
 of terms, namely the ones for which an $\con{f}$ application exists and is
-relevant to the satisfiability of the current branch of $\varphi$.
-\jb{what does ``current branch of'' mean? can we drop it / rephrase it / vague
-it up?}
+relevant to the satisfiability of $\varphi$.
+%\jb{what does ``current branch of'' mean? can we drop it / rephrase it / vague
+%it up?}
 
@@
 \begin{verbatim}
-(define-fun-rec s ((x Int)) Int (ite (<= x 0) 0 (s (- x 1))))
-(assert (> (s x) 100))
+(define-fun-rec s ((x Int)) Int (ite (<= x 0) 0 (+ x (s (- x 1)))))
+(declare-fun c () Int)
+(assert (> (s c) 100))
 (check-sat)
@@
 %then a ``satisfiable" response from the solver implies that a model exists for the original formula.
-Accordingly, the solver will output an approximation of the interpretation for recursive function definitions,
-that is, it may output the interpretation~(\ref{eq:approx-interp}) from Section~\ref{sec:encoding} for $s$.
+Accordingly, the solver will output the approximation of the interpretation it used for recursive function definitions,
+that is, it will output the following model.
+\ajr{prefer having the full output like this here?}
+%it may output the interpretation~(\ref{eq:approx-interp}) from Section~\ref{sec:encoding} for $s$.
+
+{\small
+\begin{verbatim}
+(model
+(define-fun s (($x1 Int)) Int 
+               (ite (= $x1 14) 105 (ite (= $x1 13) 91 (ite (= $x1 12) 78 
+               (ite (= $x1 11) 66 (ite (= $x1 10) 55 (ite (= $x1 4) 10 
+               (ite (= $x1 9) 45 (ite (= $x1 8) 36 (ite (= $x1 7) 28 
+               (ite (= $x1 6) 21 (ite (= $x1 3) 6 (ite (= $x1 5) 15 
+               (ite (= $x1 2) 3 (ite (= $x1 1) 1 0)))))))))))))))
+(define-fun c () Int 14)
+)
+\end{verbatim}
+}
 

### 2015-05-06 tex/main.tex
tuning

--- 2015-05-06 tex/main.tex
@@
   & \phantom{{\wedge}\; }\bigl(
+      \forall a : \fargsort{\con{s}}.\; \lite\bigl(
         \!\begin{aligned}[t]
-      \forall a : \fargsort{\con{s}}.\; \lite\bigl( \farg{\con{s}}( a ) \leq 0,\;
+         &  \farg{\con{s}}( a ) \leq 0,\;
           \con{s}(\farg{\con{s}}( a )) \teq 0,\;
-        & \phantom{{\wedge}\;} \con{s}(\farg{\con{s}}( a )) \teq \farg{\con{s}}( a ) + \con{s}( \farg{\con{s}}( a )-1 )
 \\[-\jot]
-      & {\wedge}\; \exists b : {\fargsort{\con{s}}}.\;\, \farg{\con{s}}( b ) \teq \farg{\con{s}}( a )-1 \bigr) \bigr)
+  & \con{s}(\farg{\con{s}}( a )) \teq \farg{\con{s}}( a ) + \con{s}( \farg{\con{s}}( a )-1 )
+    \wedge \exists b : {\fargsort{\con{s}}}.\;\, \farg{\con{s}}( b ) \teq \farg{\con{s}}( a )-1 \bigr) \bigr)
 \end{aligned}
@@
 
-We have seen a satisfiability-preserving translation $\conv$ that significantly improves the effectiveness of known SMT techniques
-for finding counterexamples for formulas involving recursive function definitions.
-For convienience, the translation $\conv$ has been implemented in the latest development version of \cvc %~\cite{ReyEtAl-1-RR-13}
-(version 1.5 pre-release).
+The translation $\conv$ significantly improves the effectiveness of known SMT
+techniques for finding counterexamples for formulas involving recursive
+function definitions.
+For the convenience of end users, it
+has been implemented in the latest development version of \cvc %~\cite{ReyEtAl-1-RR-13}
+(version~1.5 prerelease).
 Function definitions $\forallf{\con{f}} \vec x.\; \varphi$ can be written in using the $\definefunreccmd$ command,
-from Version 2.5~\cite{} of the SMT-LIB standard.
-For example, the formula $\varphi$ in Example~\ref{ex:translation}
-can be written as:
-
-{\small
+from the SMT-LIB 2.5 draft standard \cite{smtlib25}.
+For example, formula~(\ref{eq:ex-before}) from Example~\ref{ex:translation}
+can be specified as
+%
+\begin{quote}
 \begin{verbatim}
@@
 \end{verbatim}
-}
-
+\end{quote}
+%
 When reading this input,
-\cvc adds the (annotated) quantified formula $\forallf{\con{s}} x.\;\, \con{s}( x ) \teq \lite\bigl( x \leq 0, 0, \con{s}( x-1 )\bigr)$ to its list of assertions,
-which after rewriting becomes $\forallf{\con{s}} x.\;\, \lite\bigl( x \leq 0, \con{s}( x ) \teq 0, \con{s}( x ) \teq \con{s}( x-1 ) \bigr)$.
+\cvc adds the annotated quantified formula
+$$\forallf{\con{s}} x.\;\, \con{s}( x ) \teq \lite\bigl( x \leq 0,\, 0,\, \con{s}( x-1 )\bigr)$$
+to its list of assertions,
+which after rewriting becomes
+$$\forallf{\con{s}} x.\;\, \lite\bigl( x \leq 0, \con{s}( x ) \teq 0, \con{s}( x ) \teq \con{s}( x-1 ) \bigr)$$
 If \cvc's finite model finding mode for recursive functions is enabled (using the command line parameter ``--fmf-fun"),

### 2015-05-06 tex/main.tex
tuned

@@
 \\[-\jot]
- & {\wedge}\; \bigl( \con{s}( c ) > 100 \wedge \exists a : {\fargsort{\con{s}}}.\;\, \farg{\con{s}}( a ) \teq \con{c} \bigr)
+ & {\wedge}\; \bigl( \con{s}( \con{c} ) > 100 \wedge \exists a : {\fargsort{\con{s}}}.\;\, \farg{\con{s}}( a ) \teq \con{c} \bigr)
 \end{aligned}

### 2015-05-06 tex/main.tex
more

@@
 \newcommand{\sfuns}[1]{#1^\mathrm{f}}
-\newcommand{\sfundefs}[1]{#1^\mathrm{def}}
+\newcommand{\sfundefs}[1]{#1^\mathrm{dfn}}
 %\newcommand{\sfunndefs}[1]{#1^\mathrm{nf}}
@@
     * productive
+
+\begin{verbatim}
+this option targets users who wish to determine the satisfiability of inputs
+that involving function definitions that are known to be consistent, whereas it
+does not target users who wish to check the consistency of a set of function
+definitions.
+\end{verbatim}
 
@@
 responses and further into benchmarks before and after the translation $\conv$.
-\jb{rephrase}
-%
 The raw evaluation data reveals no cases in which a solver answered
@@
 \begin{verbatim}
-(define-fun-rec s ((x Int)) Int (ite (<= x 0) 0 (+ x (s (- x 1)))))
+(define-fun-rec s ((n Int)) Int (ite (<= n 0) 0 (+ n (s (- n 1)))))
 (declare-fun c () Int)
@@
 \cvc adds the annotated quantified formula
-$$\forallf{\con{s}} x.\;\, \con{s}( x ) \teq \lite\bigl( x \leq 0,\, 0,\, \con{s}( x-1 )\bigr)$$
+$$\forallf{\con{s}} x.\;\, \con{s}( x ) \teq \lite\bigl( x \leq 0,\; 0,\; \con{s}( x-1 )\bigr)$$
 to its list of assertions,
 which after rewriting becomes
-$$\forallf{\con{s}} x.\;\, \lite\bigl( x \leq 0, \con{s}( x ) \teq 0, \con{s}( x ) \teq \con{s}( x-1 ) \bigr)$$
-If \cvc's finite model finding mode for recursive functions is enabled (using the command line parameter ``--fmf-fun"),
-then it will replace its list of known assertions based on the conversion $\conv$ before checking for satisfiability.
+$$\forallf{\con{s}} x.\;\, \lite\bigl( x \leq 0,\; \con{s}( x ) \teq 0,\; \con{s}( x ) \teq \con{s}( x-1 ) \bigr)$$
+If \cvc's finite model finding mode for recursive functions is enabled (using
+the command-line option \texttt{--fmf-fun}), it will replace its list of known
+assertions based on the conversion $\conv$ before checking for satisfiability.
 %If the functions provided in the input are admissible,
 %then a ``satisfiable" response from the solver implies that a model exists for the original formula.
-Accordingly, the solver will output the approximation of the interpretation it used for recursive function definitions,
-that is, it will output the following model.
+Accordingly, the solver will output the approximation of the interpretation it
+used for recursive function definitions. For our example, it will output the following
+model:
 \ajr{prefer having the full output like this here?}
+\jb{I like it the way you did it now, but I took the freedom to reformat the output slightly
+(to avoid the small font---I generally try to avoid ad hoc font mixtures---and to show
+the nesting more clearly).}
 %it may output the interpretation~(\ref{eq:approx-interp}) from Section~\ref{sec:encoding} for $s$.
-
-{\small
+%
+\begin{quote}
 \begin{verbatim}
@@
 \end{verbatim}
-}
-
-Notice the option ``--fmf-fun" effectively instructs the SMT solver to assume that the functions provided by the user are admissible.
-If this option is used for a set of functions definitions that is \emph{not} admissible, then the solver is not model-sound, that is,
-it may return ``satisfiable" for unsatisfiable inputs.
-As such,
-this option targets users who wish to determine the satisfiability of inputs that involving function definitions that are known to be consistent,
-whereas it does not target users who wish to check the consistency of a set of function definitions.
+\end{quote}
+%
+The \texttt{--fmf-fun} option instructs CVC4 to assume that the
+functions provided by the user are admissible.
+If this condition is not met, the solver may answer ``satisfiable'' on
+an unsatisfiable problem. For example, if we add the inconsistent definition
+%
+\begin{quote}
+\begin{verbatim}
+(define-fun-rec e ((n Int)) Int (+ (e n) 1))
+\end{verbatim}
+\end{quote}
+%
+to the above problem, CVC4 will still find the same ``model''
+if \texttt{--fmf-fun} is specified.
 

### 2015-05-06 tex/main.tex
sketched intro

@@
 \begin{abstract}
-To do.
+
+  * finite model finding techniques for satisfiability modulo theories (SMT)
+    have emerged in recent years
+  * suffer from the major limitation that they cannot handle universal
+    quantifiers ranging over infinite types
+  * this paper introduces a simple translation to
+    produces an equisatifiable problem that avoids infinite quantification
+    for a large fragment of recursive functions,
+    including terminating, tail-recursive, and productive functions.
+  * empirical evaluation confirms that the approach works
+  * implemented as a preprocessor in CVC4
 \end{abstract}
@@
 \end{quote}
-%
+
 The \texttt{--fmf-fun} option instructs CVC4 to assume that the

### 2015-05-06 tex/main.tex
I meant abstract not intro

--- 2015-05-06 tex/main.tex
@@
 \begin{abstract}
-To do.
+SMT solvers have recently been extended with finite model finding techniques.
+However, these cannot cope with universal quantifiers ranging over infinite
+types. This paper introduces a translation that eliminates infinite quantifiers
+in axioms specifying well-behaved recursive functions, including terminating,
+tail-recursive, and productive functions. An empirical evaluation confirms that
+the approach works well on benchmarks from various sources. The translation is
+implemented as a preprocessor in the solver CVC4.
 \end{abstract}

### 2015-05-06 tex/main.tex
Start proof.
more

@@
 types. This paper introduces a translation that eliminates infinite quantifiers
-in axioms specifying well-behaved recursive functions, including terminating,
-tail-recursive, and productive functions. An empirical evaluation confirms that
-the approach works well on benchmarks from various sources. The translation is
-implemented as a preprocessor in the solver CVC4.
+in axioms specifying a large class of recursive functions, including
+terminating, tail-recursive, and productive functions. An empirical evaluation
+confirms that the approach works well on benchmarks from various sources. The
+translation is implemented as a preprocessor in the solver CVC4.
 \end{abstract}
@@
       (``maybe unsat'')
-
-  * quantifiers in recursive fun defs
-
-  * intuitively, finite quantification suffices:
-    * if we make sure all needed values are there
-    * if well-founded (assumed!)
-
-  * nasty example (not well-founded)
-    * f x = f x + 1
-
-  * how we do it
-
-  * show that it works also for example with weird recursion schemes
-
-  * corecursive functions???
-    * productive
-
-\begin{verbatim}
-this option targets users who wish to determine the satisfiability of inputs
-that involving function definitions that are known to be consistent, whereas it
-does not target users who wish to check the consistency of a set of function
-definitions.
-\end{verbatim}
-
-\ajr{Demand for recursive functions in SMT (new draft includes
-define-funs-rec). Usually not interested in showing logical consistency, there
-exist other tools that check e.g. termination.
-Thus, want SMT solver to check satisfiability the assumption that recursive
-functions are consistent (where this needs to be defined precisely).
-}
 
@@
 
-\paragraph{Contributions}
-
-In this paper, we make the following contributions:
-\begin{itemize}
-\item[-] We define a satisfiability-preserving translation method for a class of formulas containing recursive function definitions
-that produces formulas for which existing model finding techniques~\cite{GeDeM-CAV-09, ReyEtAl-1-RR-13,reynolds-et-al-2013} are applicable.
-\item[-] We provide experimental evidence that this translation improves the effectiveness of SMT solvers \cvc and \ziii for finding counterexamples to verification conditions.
-\item[-] We detail front-end support for recursive function definitions and this translation in the SMT solver \cvc.
-\end{itemize}
+The recursive functions must meet a precise semantic criterion to be admissible
+(Section~\ref{sec:prelim}). This criterion is met by terminating recursive
+function definitions, but also by tail-recursive functions and even by
+productive corecursive functions. The criterion is not met by inconsistent
+definitions such as $\forall n : {\sortint}.\;\, \con{f}(n) \teq \con{f}(n) +
+1$, but also by some consistent definitions.
+
+Demand for recursive functions in SMT (new draft includes
+define-funs-rec). Usually not interested in showing logical consistency, there
+exist other tools that check e.g. termination.
+Thus, want SMT solver to check satisfiability the assumption that recursive
+functions are consistent (where this needs to be defined precisely).
+
+We define an (un)satisfiability-preserving translation method for a class of
+formulas containing recursive function definitions (Section~\ref{sec:encoding})
+that produces formulas for which existing model finding techniques are
+applicable. The idea
+
+replace a definition $\forall x : \tau.\; \const{f}(x) \teq t$ by $\forall a :
+\fargsort{\tau}.\; \const{f}(\farg{\con{f}}(a)) \teq t[\concret(a)/x]$, where
+$\fargsort{\tau}$ is an uninterpreted ``abstract'' type and $\farg{\con{f}} :
+\fargsort{\tau} \to \tau$ a coercion from the abstract type to the concrete
+type. For admissible definitions, this preserves satisfiability and
+unsatisfiability, while making it possible to find finite models using known
+techniques.
+
+We provide experimental evidence that this translation improves the
+effectiveness of SMT solvers \cvc and \ziii for finding counterexamples to
+verification conditions (Section~\ref{sec:evaluation}).
+
+We detail front-end support for recursive function definitions and this
+translation in the SMT solver \cvc (Section~\ref{sec:front-end}).
+
+\newpage
+
+  * quantifiers in recursive fun defs
+
+  * intuitively, finite quantification suffices:
+    * if we make sure all needed values are there
+    * if well-founded (assumed!)
+
+  * nasty example (not well-founded)
+    * f x = f x + 1
+
+  * how we do it
+
+  * corecursive functions???
+    * productive
+
+\begin{verbatim}
+this option targets users who wish to determine the satisfiability of inputs
+that involving function definitions that are known to be consistent, whereas it
+does not target users who wish to check the consistency of a set of function
+definitions.
+\end{verbatim}
 
@@
     odd schemes, give unsat example
+    [* show that it works also for example with weird recursion schemes]
 

### 2015-05-06 tex/main.tex
work on intro

@@
 
-  * why model finding?
-  * but SMT solvers better at proving
-    * a single universal quantifier in an axiom is enough to put it off track
-      (``maybe unsat'')
-
-Recent techniques~\cite{GeDeM-CAV-09, ReyEtAl-1-RR-13,reynolds-et-al-2013}
-have focused on finding models for quantified formulas in SMT.
-Ge and de Moura \cite{GeDeM-CAV-09} presented a complete instantiation-based
-procedure for formulas in the \emph{essentially uninterpreted} fragment.
+Many SMT solvers can reason about quantified formulas using incomplete
+instantiation-based methods \cite{MouraBjoerner07,ReynoldsTinelliMoura14}.
+These methods work well in the context of proving (i.e., showing
+unsatisfiability), but they generally cannot be used to find models (i.e.,
+showing satisfiability). Typically, a single universal quantifier in one of the
+axioms of a problem is enough to prevent the discovery of models.
+
+In the past few years, techniques have been developed to finding models for
+quantified formulas in SMT.
+Ge and de Moura \cite{GeDeM-CAV-09} introduced a complete instantiation-based
+procedure for formulas in the \relax{essentially uninterpreted} fragment.
 This fragment is limited to universally quantified formulas where all
 occurrences of variables occur as direct subterms of uninterpreted
-functions. For example, the formula $\forall x%\typ{\sortint} --- the sort is not important, right?
+functions---e.g., $\forall x%\typ{\sortint} --- the sort is not important, right?
 .\;\, \con{f}( x )
-\teq \con{g}( x ) + 5$ is in the essentially uninterpreted fragment.
+\teq \con{g}( x ) + 5$.
 % since all
@@
 % $\con{g}$.
-Other syntactic criteria are identified in this paper that extend
+(Other syntactic criteria extend
 this fragment slightly, including cases when variables occur as arguments of
-arithmetic predicates. The techniques in~Reynolds et al.\
-\cite{ReyEtAl-1-RR-13,reynolds-et-al-2013} can be used for finding finite
+arithmetic predicates.) Reynolds et al.\
+\cite{ReyEtAl-1-RR-13,reynolds-et-al-2013} introduced techniques for finding finite
 models for quantified
@@
 interpretation, such as fixed-width bit vectors and enumerated datatypes.
-Finite model finding techniques can
+These techniques can
 find a model for the formula $\forall x,\, y : \tau.\;\, x \teq
-y \vee \neg \con{f}( x ) \teq \con{f}( y )$ where the uninterpreted type $\tau$ is
-interpreted as a finite set and $\con{f}$ as injective finite map.
+y \vee \neg \con{f}( x ) \teq \con{f}( y )$.
+% where the uninterpreted type $\tau$ is
+% interpreted as a finite set and $\con{f}$ as injective finite map.
 
 Unfortunately, these fragments cannot accommodate the vast majority of
-quantified formulas that correspond to recursive function definitions.
-In particular, the fragments in~\cite{GeDeM-CAV-09} are not applicable when the argument of a recursive function is used in various ways (e.g. within an arithmetic expression),
-while the fragment in~\cite{ReyEtAl-1-RR-13,reynolds-et-al-2013} is not applicable for functions over infinite domains such as the integers.
-So, for instance, the quantified formula $\forall x : {\sortint}.\;\, \lite\bigl( p \leq 0, p( x ) \teq 0, p( x ) \teq 2 * p( x - 1 ) \bigr)$
-is not in the essentially uninterpreted fragment (since $x$ occurs as a child of $-$),
-nor does finite model finding apply (since $x$ is of type $\sortint$).
+quantified formulas that correspond to recursive function definitions: The
+essentially uninterpreted fragment does not allow the argument of a
+recursive function to be used deep inside a complex term on the right-hand side,
+whereas the finite model finding techniques of Reynolds et al.\ are not
+applicable for functions over infinite domains such as the integers. A simple
+example where both approaches fail is $\forall n : {\sortint}.\;\, \lite\bigl(
+\con{p} \leq 0,\allowbreak\; \con{p}( n ) \teq 1,\allowbreak\; \con{p}( n ) \teq 2 * \con{p}( n - 1 ) \bigr)$.
 
@@
 where $\psi$ entails $\varphi_0$.
+Since $\I$ is a model of $\varphi_0$,
+there exists a set of literals $\psi_0$ over the atoms occurring in $\varphi_0$
+such that $\I$ satisfies $\psi_0$ and $\psi_0$ entails $\varphi_0$.
+\ajr{todo}
 

### 2015-05-06 tex/main.tex
more

@@
 terminating, tail-recursive, and productive functions. An empirical evaluation
-confirms that the approach works well on benchmarks from various sources. The
+confirms that the approach works well on benchmarks from two sources. The
 translation is implemented as a preprocessor in the solver CVC4.
@@
 
-Many SMT solvers can reason about quantified formulas using incomplete
-instantiation-based methods \cite{MouraBjoerner07,ReynoldsTinelliMoura14}.
+Many solvers based on SMT (satisfiability modulo theories) can reason about
+quantified formulas using incomplete instantiation-based methods
+\cite{MouraBjoerner07,ReynoldsTinelliMoura14}.
 These methods work well in the context of proving (i.e., showing
@@
 applicable for functions over infinite domains such as the integers. A simple
-example where both approaches fail is $\forall n : {\sortint}.\;\, \lite\bigl(
-\con{p} \leq 0,\allowbreak\; \con{p}( n ) \teq 1,\allowbreak\; \con{p}( n ) \teq 2 * \con{p}( n - 1 ) \bigr)$.
-
-In this paper, we advocate a method for translating formulas involving function
-definitions (such as the one mentioned in the previous paragraph) 
-into ones where the techniques~\cite{GeDeM-CAV-09, ReyEtAl-1-RR-13} are applicable.
-
+example where both approaches fail is
+$$\forall n : {\sortint}.\;\, \lite\bigl(
+\con{p} \leq 0,\allowbreak\; \con{p}( n ) \teq 1,\allowbreak\; \con{p}( n ) \teq 2 * \con{p}( n - 1 ) \bigr)$$
+This state of affairs is highly unsatisfactory, given the frequence of
+recursive definitions in practice and the impending addition of a dedicated
+command, \texttt{define-funs-rec}, to the SMT-LIB standard \cite{smtlib25}.
+
+In this paper, we present a method for translating formulas involving function
+definitions into formulas where finite model finding techniques can be applied.
 The recursive functions must meet a precise semantic criterion to be admissible
@@
 function definitions, but also by tail-recursive functions and even by
-productive corecursive functions. The criterion is not met by inconsistent
+productive corecursive functions. It is, however, not met by inconsistent
 definitions such as $\forall n : {\sortint}.\;\, \con{f}(n) \teq \con{f}(n) +
-1$, but also by some consistent definitions.
-
-Demand for recursive functions in SMT (new draft includes
-define-funs-rec). Usually not interested in showing logical consistency, there
+1$, and it is possible to construct consistent examples that do not meet it.
+
+We define an (un)satisfiability-preserving translation method for a class of
+formulas containing admissible recursive function definitions
+(Section~\ref{sec:encoding}) that produces formulas for which existing model
+finding techniques are applicable. The main insight is that a
+recursive definition $\forall x : \tau.\; \const{f}(x) \teq t$
+can be translated to
+$\forall a :
+\fargsort{\tau}.\; \const{f}(\farg{\con{f}}(a)) \teq t[\concret(a)/x]$, where
+$\fargsort{\tau}$ is an uninterpreted ``abstract'' type and $\farg{\con{f}} :
+\fargsort{\tau} \to \tau$ a coercion from the abstract type to the concrete
+type. The translation preserves satisfiability and, for admissible definitions,
+unsatisfiability, and it makes finite model finding possible.
+
+Our empirical evaluation on benchmarks from provides evidence that this translation improves the
+effectiveness of SMT solvers \cvc and \ziii for finding counterexamples to
+verification conditions (Section~\ref{sec:evaluation}).
+
+We detail front-end support for recursive function definitions and this
+translation in the SMT solver \cvc (Section~\ref{sec:front-end}).
+
+Admissibility must be checked separately---for example, using a
+syntactic criterion or a termination prover.
+
+Usually not interested in showing logical consistency, there
 exist other tools that check e.g. termination.
@@
 functions are consistent (where this needs to be defined precisely).
-
-We define an (un)satisfiability-preserving translation method for a class of
-formulas containing recursive function definitions (Section~\ref{sec:encoding})
-that produces formulas for which existing model finding techniques are
-applicable. The idea
-
-replace a definition $\forall x : \tau.\; \const{f}(x) \teq t$ by $\forall a :
-\fargsort{\tau}.\; \const{f}(\farg{\con{f}}(a)) \teq t[\concret(a)/x]$, where
-$\fargsort{\tau}$ is an uninterpreted ``abstract'' type and $\farg{\con{f}} :
-\fargsort{\tau} \to \tau$ a coercion from the abstract type to the concrete
-type. For admissible definitions, this preserves satisfiability and
-unsatisfiability, while making it possible to find finite models using known
-techniques.
-
-We provide experimental evidence that this translation improves the
-effectiveness of SMT solvers \cvc and \ziii for finding counterexamples to
-verification conditions (Section~\ref{sec:evaluation}).
-
-We detail front-end support for recursive function definitions and this
-translation in the SMT solver \cvc (Section~\ref{sec:front-end}).
 

### 2015-05-06 tex/main.tex
Work on proof.

@@
 respect to the functions defined by $\DDD$, we know $\varphi$ is of the form
-$\DDD \wedge \varphi_0$. We first argue that $\I$ satisfies a $\Sigma$-formula
-$\psi$ that is closed under function expansion with respect to $\DDD$, and
-where $\psi$ entails $\varphi_0$.
-Since $\I$ is a model of $\varphi_0$,
-there exists a set of literals $\psi_0$ over the atoms occurring in $\varphi_0$
-such that $\I$ satisfies $\psi_0$ and $\psi_0$ entails $\varphi_0$.
-\ajr{todo}
-
+$\DDD \wedge \varphi_0$. 
+%We first argue that $\I$ satisfies a $\Sigma$-formula
+%$\psi$ that is closed under function expansion with respect to $\DDD$, and
+%where $\psi$ entails $\varphi_0$.
+To do so, 
+we first construct a sequence of $\extendsig{\Sigma}$-literals $M_0, M_1, \ldots$ such that $\I$ satisfies each $M_i$,
+and a sequence of mappings $C^{\con f}_0, C^{\con f}_1, \ldots$ from values $\fargsort{f}^\I$ 
+to tuples of terms of type $\tau_1 \times \cdots \times \tau_n$ 
+for each function $\con{f}$ of type $\tau_1 \times \cdots \times \tau_n \rightarrow \tau$ in $\sfundefs{\Sigma}$.
+Let $M_0$ be a set of literals over the atoms occurring in $\varphi_0$
+such that $\I$ satisfies each literal in $M_0$ and $\bigwedge M_0$ entails $\conv( \varphi_0 )$.
+For each $\con{f} \in \sfundefs{\Sigma}$, let $C^{\con{f}}_0$ be the empty mapping.
+For each $i \geq 0$, 
+let $X_i$ be the set of constraints of the form $\exists a : {\fargsort{\con{f}}}. ( \vecfarg{\con{f}}( a ) \teq \vec t )$ occurring in $M_i$
+and let $L_i$ be the others in $M_i$.
+For each $\con{f} \in \sfundefs{\Sigma}$, let $C^{\con{f}}_{i+1}$ be a mapping containing $C^{\con{f}}_i$ and a minimal number of entries
+such that for all $\exists a : {\fargsort{\con{f}}}. ( \vecfarg{\con{f}}( a ) \teq \vec t ) \in X_i$,
+there exists a $u \mapsto \vec s \in C^{\con{f}}_{i+1}$ such that $\vecfarg{\con{f}}^\I( u ) = \vec s^\I = \vec t^\I$.
+Let $M_{i+1}$ be the set containing $M_i$ as well as
+
+Let $M^\ast$ be the limit of this sequence, which we partition into sets $X^\ast$ and $L^\ast$ in the same way as above.
+Let $\psi$ be $\bigwedge L^\ast$.
+We know $\psi$ is a $\Sigma$-formula since \ajr{blah}
+We know $\psi$ is closed under function expansion with respect to $\DDD$ since \ajr{blah}
+Finally, we know $\psi$ entails $\varphi_0$ since \ajr{blah}
 Since there exists a model for $\psi$ that is closed under function expansion
@@
 a similar constraint is enforced recursively: any model satisfying $\varphi_{\con{s}}[t/x]$
-must also satisfy $\varphi_{\con{s}}[t-1/x]$ when $t > 0$. Given a model $\I$ of
+must also satisfy $\varphi_{\con{s}}[t-1/x]$ when $t > 0$ \ajr{revise this sentence, not quite accurate}. Given a model $\I$ of
 formula~(\ref{eq:ex-after}) that interprets $\con{c}$ as $14$, it is easy to see that
@@
 Since $\varphi_\con{s}$ is an admissible definition,
-there must exist a model for $\psi \wedge \varphi_\con{s}$,
-which means there exists a model for formula~(\ref{eq:ex-before}).\xend
+there must exist a $\Sigma$-interpretation satisfying $\psi \wedge \varphi_\con{s}$,
+which means there exists a $\Sigma$-interpretation satisfying formula~(\ref{eq:ex-before}).\xend
 \end{example}

### 2015-05-06 tex/main.tex
fix

--- 2015-05-06 tex/main.tex
@@
 $$\forall n : {\sortint}.\;\, \lite\bigl(
-\con{p} \leq 0,\allowbreak\; \con{p}( n ) \teq 1,\allowbreak\; \con{p}( n ) \teq 2 * \con{p}( n - 1 ) \bigr)$$
+n \leq 0,\allowbreak\; \con{p}( n ) \teq 1,\allowbreak\; \con{p}( n ) \teq 2 * \con{p}( n - 1 ) \bigr)$$
 This state of affairs is highly unsatisfactory, given the frequence of

### 2015-05-06 tex/main.tex
intro

@@
 
-\newcommand\cvcd{CVC4d\xspace}
-\newcommand\cvcf{CVC4f\xspace}
-\newcommand\cvci{CVC4i\xspace}
+\newcommand\cvcd{{\cvc}d\xspace}
+\newcommand\cvcf{{\cvc}f\xspace}
+\newcommand\cvci{{\cvc}i\xspace}
 \newcommand\ziiib{\ziii}
@@
 SMT solvers have recently been extended with finite model finding techniques.
-However, these cannot cope with universal quantifiers ranging over infinite
-types. This paper introduces a translation that eliminates infinite quantifiers
-in axioms specifying a large class of recursive functions, including
-terminating, tail-recursive, and productive functions. An empirical evaluation
-confirms that the approach works well on benchmarks from two sources. The
-translation is implemented as a preprocessor in the solver CVC4.
+However, these generally cannot cope with universal quantifiers ranging over
+infinite types. This paper introduces a translation that eliminates infinite
+quantifiers in axioms specifying a large class of recursive functions,
+including terminating, tail-recursive, and productive functions. An empirical
+evaluation confirms that the approach works well on benchmarks from two
+independent sources. The translation is implemented as a preprocessor in the
+solver CVC4.
 \end{abstract}
@@
 % $\con{g}$.
-(Other syntactic criteria extend
+Other syntactic criteria extend
 this fragment slightly, including cases when variables occur as arguments of
-arithmetic predicates.) Reynolds et al.\
+arithmetic predicates. More recently, Reynolds et al.\
 \cite{ReyEtAl-1-RR-13,reynolds-et-al-2013} introduced techniques for finding finite
@@
 find a model for the formula $\forall x,\, y : \tau.\;\, x \teq
-y \vee \neg \con{f}( x ) \teq \con{f}( y )$.
+y \vee \allowbreak \neg \con{f}( x ) \teq \con{f}( y )$, where $\tau$ is uninterpreted.
 % where the uninterpreted type $\tau$ is
@@
 
-Unfortunately, these fragments cannot accommodate the vast majority of
+Unfortunately, neither of these fragments can accommodate the vast majority of
 quantified formulas that correspond to recursive function definitions: The
@@
 recursive definitions in practice and the impending addition of a dedicated
-command, \texttt{define-funs-rec}, to the SMT-LIB standard \cite{smtlib25}.
+command, \texttt{define-fun-rec}, to the SMT-LIB standard \cite{smtlib25}.
 
@@
 
-We define an (un)satisfiability-preserving translation method for a class of
+We define a translation for a class of
 formulas containing admissible recursive function definitions
@@
 type. The translation preserves satisfiability and, for admissible definitions,
-unsatisfiability, and it makes finite model finding possible.
-
-Our empirical evaluation on benchmarks from provides evidence that this translation improves the
-effectiveness of SMT solvers \cvc and \ziii for finding counterexamples to
-verification conditions (Section~\ref{sec:evaluation}).
-
-We detail front-end support for recursive function definitions and this
-translation in the SMT solver \cvc (Section~\ref{sec:front-end}).
-
-Admissibility must be checked separately---for example, using a
-syntactic criterion or a termination prover.
-
-Usually not interested in showing logical consistency, there
-exist other tools that check e.g. termination.
-Thus, want SMT solver to check satisfiability the assumption that recursive
-functions are consistent (where this needs to be defined precisely).
-
-\newpage
-
-  * quantifiers in recursive fun defs
-
-  * intuitively, finite quantification suffices:
-    * if we make sure all needed values are there
-    * if well-founded (assumed!)
-
-  * nasty example (not well-founded)
-    * f x = f x + 1
-
-  * how we do it
-
-  * corecursive functions???
-    * productive
-
-\begin{verbatim}
-this option targets users who wish to determine the satisfiability of inputs
-that involving function definitions that are known to be consistent, whereas it
-does not target users who wish to check the consistency of a set of function
-definitions.
-\end{verbatim}
+unsatisfiability, and it makes finite model finding possible in practice.
+
+Our empirical evaluation on benchmarks from IsaPlanner and Leon provides
+evidence that this translation improves the effectiveness of the SMT solvers
+\cvc and \ziii for finding counterexamples to verification conditions
+(Section~\ref{sec:evaluation}). The approach is implemented as a preprocessor
+in \cvc. When invoked with the appropriate command-line option,
+the solver assumes functions introduced using the \texttt{define-fun-rec} command are
+admissible and translates them accordingly (Section~\ref{sec:front-end}).
+
+%this option targets users who wish to determine the satisfiability of inputs
+%that involving function definitions that are known to be consistent, whereas it
+%does not target users who wish to check the consistency of a set of function
+%definitions.
 
@@
 {\bench isa} and {\bench leon}. The first ({\bench isa}) consists of the 79
-benchmarks from the Isaplanner benchmark
+benchmarks from the IsaPlanner
 suite~\cite{DBLP:conf/itp/JohanssonDB10} that do not contain higher-order
@@
 
-\section{Front-End Support %for Recursive Function Definitions
-  in CVC4}
+\section{Front-End Support in \cvc}
 \label{sec:front-end}
@@
 
-The \texttt{--fmf-fun} option instructs CVC4 to assume that the
+The \texttt{--fmf-fun} option instructs \cvc to assume that the
 functions provided by the user are admissible.
-If this condition is not met, the solver may answer ``satisfiable'' on
+Admissibility must be checked separately---for example, using a
+syntactic criterion or a termination prover.
+If this criterion is not met, the solver may answer ``satisfiable'' on
 an unsatisfiable problem. For example, if we add the inconsistent definition
@@
 %
-to the above problem, CVC4 will still find the same ``model''
+to the above problem, \cvc will still find the same ``model''
 if \texttt{--fmf-fun} is specified.
+
+%Usually not interested in showing logical consistency, there
+%exist other tools that check e.g. termination.
+%Thus, want SMT solver to check satisfiability the assumption that recursive
+%functions are consistent (where this needs to be defined precisely).
 
@@
       * against other non-SMT-solvers (notably Leon and Nitpick)
-      * problems exercising codatatypes (which CVC4 supports \cite{xxx}) and
+      * problems exercising codatatypes (which \cvc supports \cite{xxx}) and
         corecursive functions

### 2015-05-06 tex/main.tex
more

@@
 \newcommand\win{\bf}
-\newcommand\bench{\ttfamily}
+%\newcommand\bench{\ttfamily}
+
+\newcommand\isa{Isa\xspace}
+\newcommand\isam{Isa-Mut\xspace}
+\newcommand\leon{Leon\xspace}
+\newcommand\leonm{Leon-Mut\xspace}
 
@@
 \end{abstract}
+
+%% The institutions above shouldn't count as footnotes
+\setcounter{footnote}{0}
 
@@
 in \cvc. When invoked with the appropriate command-line option,
-the solver assumes functions introduced using the \texttt{define-fun-rec} command are
-admissible and translates them accordingly (Section~\ref{sec:front-end}).
+the solver assumes functions introduced using the \texttt{define-fun-rec}
+command are admissible and translates them accordingly
+(Section~\ref{sec:front-end}).
 
@@
 
+\jb{Add link to our evaluation data, like we did in CADE paper}
+
 In this section, we evaluate both the overall impact of the encoding introduced
-in the previous section, and the performance of individual SMT techniques for
+in the previous section and the performance of individual SMT techniques for
 benchmarks in the encoding.
@@
 We gathered a set of 245 benchmarks from two sources, which we will refer to as
-{\bench isa} and {\bench leon}. The first ({\bench isa}) consists of the 79
+\isa and \leon. The first source consists of the 79
 benchmarks from the IsaPlanner
@@
 recursive functions, and are limited to the combined theory of uninterpreted
-functions and inductive datatypes. The second ({\bench leon}) consists of 166
+functions and inductive datatypes. The second source consists of 166
 benchmarks from the Leon repository,\footnote{%Available at
@@
 A majority of these benchmarks are unsatisfiable, although a
-handful of the benchmarks in the {\bench leon} set are satisfiable.
+handful of the benchmarks in the \leon set are satisfiable.
 
@@
 Note that benchmarks created in this way have a high likelihood of having small and easy-to-find counterexamples.
-In total, we considered 213 mutated forms of conjectures from {\bench isa}, and 427 mutated forms of conjectures from {\bench leon}.
-We will call these sets {\bench isa+m} and {\bench leon+m} respectively.
+In total, we considered 213 mutated forms of conjectures from \isa, and 427 mutated forms of conjectures from \leon.
+We will call these sets \isam and \leonm respectively.
 
@@
 We considered these 885 benchmarks both before and after the translation $\conv$ as described in Section~\ref{sec:encoding}.
-For example, {\bench isa} contains 79 original benchmarks $\varphi$
+For example, \isa contains 79 original benchmarks $\varphi$
 and 79 translated benchmarks $\conv(\varphi)$.
@@
 \midrule
-{\bench isa} & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0
+\isa & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0
 \\
-{\bench leon}  & 0 & 2 & 0 & 0 & 0 & {\win 9} & 0 & 0
+\leon  & 0 & 2 & 0 & 0 & 0 & {\win 9} & 0 & 0
 \\
-{\bench isa+m} & 0 & 35 & 0 & 0 & 0 & {\win 153} & 0 & 0
+\isam & 0 & 35 & 0 & 0 & 0 & {\win 153} & 0 & 0
 \\
-{\bench leon+m}  & 11  & 75 & 6 & 6 & 6 & {\win 169} & 6 & 6
+\leonm  & 11  & 75 & 6 & 6 & 6 & {\win 169} & 6 & 6
 \\[\jot]
@@
 \midrule
-{\bench isa} & 14  & 15 & 15  & 15 & 15  & 15 & {\win 61}  & 14
+\isa & 14  & 15 & 15  & 15 & 15  & 15 & {\win 61}  & 14
 \\
-{\bench leon}  & 73  & 78 & 80  & 80 & 80  & 76 & {\win 96}  & 78
+\leon  & 73  & 78 & 80  & 80 & 80  & 76 & {\win 96}  & 78
 \\
-{\bench isa+m} & 17  & 18 & 18  & 18 & 18  & 18 & {\win 44}  & 17
+\isam & 17  & 18 & 18  & 18 & 18  & 18 & {\win 44}  & 17
 \\
-{\bench leon+m}  & 83  & 98 & 103 & 98 & 104 & 95 & {\win 117} & 98
+\leonm  & 83  & 98 & 103 & 98 & 104 & 95 & {\win 117} & 98
 \\[\jot]
@@
 impact. \cvcf finds 331 of the 885 benchmarks to be satisfiable,
-including 9~benchmarks in the unmutated {\bench leon} benchmark set. The
+including 9~benchmarks in the unmutated \leon benchmark set. The
 performance of \ziiib for counterexamples also improves dramatically, as it

### 2015-05-06 tex/main.tex
more

@@
 However, these generally cannot cope with universal quantifiers ranging over
-infinite types. This paper introduces a translation that eliminates infinite
+infinite types, such as integers and recursive datatypes.
+This paper introduces a translation that eliminates infinite
 quantifiers in axioms specifying a large class of recursive functions,
@@
 whereas the finite model finding techniques of Reynolds et al.\ are not
-applicable for functions over infinite domains such as the integers. A simple
-example where both approaches fail is
+applicable for functions over infinite domains such as the integers or
+recursive datatypes. A simple example where both approaches fail is
 $$\forall n : {\sortint}.\;\, \lite\bigl(
@@
 and only if formula~(\ref{eq:ex-before}) is satisfiable, noting that the set
-$\{ \forallf{\con{s}} x\typ{\sortint}.\;\, \lite\bigl( x \leq 0, \con{s}(x)
-\teq 0, \con{s}( x ) \teq x + \con{s}( x-1 ) \bigr) \}$ is admissble.
+$\{ \forallf{\con{s}} x\typ{\sortint}.\;\, \lite\bigl( x \leq 0,\allowbreak\; \con{s}(x)
+\teq 0,\; \con{s}( x ) \teq x + \con{s}( x-1 ) \bigr) \}$ is admissble.
 As stated
@@
 
+We presented a translation scheme that makes more SMT problems amenable to
+finite model findings in SMT. In conjunction with the finite model finding
+techniques introduced by Reynolds et al.\
+\cite{ReyEtAl-1-RR-13,reynolds-et-al-2013}, it makes it possible to
+find models of quantified formulas over infinite types, such as arithmetic
+types and recursive datatypes.
+
+in definitional form with respect to admissible function definitions
+
+Universal quantifiers ranging over infinite types
+
 \begin{verbatim}

### 2015-05-06 tex/main.tex
related work

@@
 uninterpreted type $\fargsort{\con{f}}$ and inserts appropriate coercions
-$\fargx{\con{f}}{i}$. Second, it augments $\varphi$ with additional constraints
+$\fargx{\con{f}}{i}$. Second, it augments $\varphi$ with additional exisential
+constraints
 $\exists a : \fargsort{\con{f}}.\;\, \vecfarg{\con{f}}( a ) \teq \vec t$; these
@@
 
-\begin{verbatim}
-first-order model finders: SEM, MACE/Mace, Paradox, FMDarwin, iProver
-  either direct search or reduction to SAT
-
-Alloy (and Alloy idiom by Kuncak \& Jackson), Kodkod
-
-The model finder for KIV based on Alloy
-
-Refute
-  * unsound handling of datatypes
-
-Nitpick
-  * nice: three-valued logic (e.g. local overflow)
-    * solving a different problem:
-      * no built-in notion, what's approximated is an idealized notion, no
-        injections $\concret$
-  * not so nice:
-    one abstract domain per type, not per function argument
-  * guards + one-domain per arg are a good combi, guards + one-single-domain is bad because of
-    odd schemes, give unsat example
-    [* show that it works also for example with weird recursion schemes]
-
-  * other features:
-    * inductive / coinductive predicates
-    * quotient types
-
-  * abstract interpretation
-
-Leon
-
-in smt : z3, cvc4 (briefly, text in main section gives detail)
-\end{verbatim}
+We described the most closely related work, by Ge and de Moura
+\cite{GeDeM-CAV-09} and by Reynolds et al.\
+\cite{ReyEtAl-1-RR-13,reynolds-et-al-2013}, in the introduction already.
+Outside the world of SMT, model finding and counterexample generation have been
+studied many times. The text below is partly adapted from Blanchette's Ph.D.\
+thesis \cite{blanchette-2012-phd}.
+
+Some finite model finders are based on a reduction to Boolean satisfiability
+(SAT). They translate the input problem to propositional logic and pass it to a
+SAT solver. The translation is parameterized by upper or exact finite bounds on
+the cardinalities of the atomic types. This procedure was pioneered by McCune
+in the earlier versions of Mace (also capitalized MACE) \cite{mccune-1994}.
+Other SAT-based finders are Paradox \cite{claessen-sorensson-2003}, FMDarwin
+\cite{xxxx}, and iProver \cite{xxxx} for first-order logic with equality; the
+Alloy Analyzer and its back-end Kodkod \cite{torlak-jackson-2007} for
+first-order relational logic; and Refute \cite{weber-2008} and Nitpick
+\cite{blanchette-nipkow-2010} for higher-order logic.
+
+An alternative to translating the problem to propositional logic is to perform
+an exhaustive model search directly on the original problem. Given fixed
+cardinalities, the search space is represented in memory as multidimensional
+arrays. The procedure tries different values in the function and predicate
+table entries, checking each time if the problem is satisfied and backtracking
+if necessary. This approach was pioneered by SEM \cite{zhang-zhang-1995} and
+FINDER \cite{slaney-1994} and serves as the basis of many more model finders,
+notably the Alloy Analyzer's precursor \cite{jackson-1996} and the later
+versions of~Mace \cite{mccune-prover9-mace4}.
+
+Most of the above tool cannot cope with recursive datatypes or other infinite
+types. Kuncak and Jackson \cite{kuncak-and-jackson-2005} presented an idiom for
+encoding datatypes and recursive functions in Alloy, by approximating datatypes
+by finite subterm-closed substructures. The approach produces sound (fragments
+of) models for formulas in the \relax{existential--bounded-universal} fragment
+(i.e., formulas whose prenex normal forms contain no unbounded universal
+quantifiers ranging over datatypes). This idiom was further developed by Dunets
+et al.\ \cite{dunets-et-al-2010}, who showed how to handle primitive recursion.
+Their definedness guards play a similar role to the existential constraints
+generated by the $\conv$ translations.
+
+The higher-order model finder Nitpick for the Isabelle/HOL proof assistant uses
+another variant of Kuncak and Jackson's approach inside a Kleene-style
+three-valued logic. It was also the first tool of its kind to corecursion and
+codatatypes \cite{blanchette-2010-tap}. The three-valued logic approach
+enriches each type with an unknown value ($\bot$), which is propagated by
+function application. This approach works reasonably well in Nitpick, because
+it builds on Kodkod and its relational logic, but our initial experiments with
+\cvc convinced us that it is more efficient to avoid unknown values by adding
+constraints.
+
+The Leon system \cite{leon} implements a procedure that can both produces
+proofs and counterexamples for properties of terminating functions written in a
+subset of Scala. Leon is based on an SMT solver. It avoids quantifiers
+altogether by unfolding recursive definitions up to a certain depth, which is
+increased iteratively as necessary.
+\jb{Andy: Can you add some more here? You're the Leon pro!}
+
+%Unlike an SMT solver, Nitpick cannot rely on a built-in notion of
+%a datatype; it axiomatizes finite subterm-closed substructures.
+
+%\begin{verbatim}
+%Nitpick
+%  * nice: three-valued logic (e.g. local overflow)
+%    * solving a different problem:
+%      * no built-in notion, what's approximated is an idealized notion, no
+%        injections $\concret$
+%  * not so nice:
+%    one abstract domain per type, not per function argument
+%  * guards + one-domain per arg are a good combi, guards + one-single-domain is bad because of
+%    odd schemes, give unsat example
+%    [* show that it works also for example with weird recursion schemes]
+%
+%  * other features:
+%    * inductive / coinductive predicates
+%    * quotient types
+%
+%  * abstract interpretation
+%
+%Leon
+%\end{verbatim}
 
@@
 \cite{ReyEtAl-1-RR-13,reynolds-et-al-2013}, it makes it possible to
-find models of quantified formulas over infinite types, such as arithmetic
-types and recursive datatypes.
-
-in definitional form with respect to admissible function definitions
-
-Universal quantifiers ranging over infinite types
-
-\begin{verbatim}
-  * future work
-    * more eval
-      * against other non-SMT-solvers (notably Leon and Nitpick)
-      * problems exercising codatatypes (which \cvc supports \cite{xxx}) and
-        corecursive functions
-    * more features
-      * e.g. encode (co)inductive predicates as (co)recursive functions
-\end{verbatim}
+find models of quantified formulas over infinite types, such as integers
+and recursive datatypes.
+
+In future work, it would be interesting to evaluate the approach against other
+model finders, notably Leon and Nitpick, and enrich the benchmark suite with
+problems exercising \cvc's support for codatatypes
+\cite{reynolds-blanchette-2015-codata}. We also plan to integrate \cvc as a
+counterexample generator in proof assistants.
 

### 2015-05-06 tex/main.tex
related

@@
 \cite{ReyEtAl-1-RR-13,reynolds-et-al-2013}, in the introduction already.
-Outside the world of SMT, model finding and counterexample generation have been
+Outside the world of SMT, model finding has been
 studied many times. The text below is partly adapted from Blanchette's Ph.D.\
@@
 \jb{Andy: Can you add some more here? You're the Leon pro!}
+
+Model finding is concerned with satisfying arbitrary logical constraints. Some
+tools are optimized to work with problems that correspond to total functional
+programs. The QuickCheck tool for Haskell \cite{xxx} is an early example, based
+on random testing. Exhaustive testing and narrowing are other popular
+strategies \cite{xxx,yyy}. These tool are often much faster than model finders,
+but they typically cannot cope with underspecification and nonexecutable
+functions.
+
+% others
+%are Agsy for Agda \cite{xxx}, Quickcheck for Isabelle/HOL \cite{xxx}, and
+% QuickChick for Coq \cite{xxx}.
 

### 2015-05-06 tex/main.tex
more text

@@
 table entries, checking each time if the problem is satisfied and backtracking
-if necessary. This approach was pioneered by SEM \cite{zhang-zhang-1995} and
-FINDER \cite{slaney-1994} and serves as the basis of many more model finders,
+if necessary. This approach was pioneered by FINDER \cite{slaney-1994} and
+SEM \cite{zhang-zhang-1995} and serves as the basis of many more model finders,
 notably the Alloy Analyzer's precursor \cite{jackson-1996} and the later
@@
 Most of the above tool cannot cope with recursive datatypes or other infinite
-types. Kuncak and Jackson \cite{kuncak-and-jackson-2005} presented an idiom for
+types. Kuncak and Jackson \cite{kuncak-jackson-2005} presented an idiom for
 encoding datatypes and recursive functions in Alloy, by approximating datatypes
@@
 another variant of Kuncak and Jackson's approach inside a Kleene-style
-three-valued logic. It was also the first tool of its kind to corecursion and
-codatatypes \cite{blanchette-2010-tap}. The three-valued logic approach
-enriches each type with an unknown value ($\bot$), which is propagated by
-function application. This approach works reasonably well in Nitpick, because
-it builds on Kodkod and its relational logic, but our initial experiments with
-\cvc convinced us that it is more efficient to avoid unknown values by adding
-constraints.
-
-The Leon system \cite{leon} implements a procedure that can both produces
-proofs and counterexamples for properties of terminating functions written in a
-subset of Scala. Leon is based on an SMT solver. It avoids quantifiers
-altogether by unfolding recursive definitions up to a certain depth, which is
-increased iteratively as necessary.
-\jb{Andy: Can you add some more here? You're the Leon pro!}
+three-valued logic, inspired by abstract interpretation. It was also the first
+tool of its kind to corecursion and codatatypes
+\cite{blanchette-2013-relational}. The three-valued logic approach extends each
+approximated type with an unknown value, which is propagated by function
+application. This approach works reasonably well in Nitpick, because it builds
+on Kodkod's relational logic, but our initial experiments with \cvc convinced
+us that it is more efficient to avoid unknown values by adding constraints.
+
+The Leon system \cite{suter-et-al-2011} implements a procedure that can both
+produces proofs and counterexamples for properties of terminating functions
+written in a subset of Scala. Leon is based on an SMT solver. It avoids
+quantifiers altogether by unfolding recursive definitions up to a certain
+depth, which is increased iteratively as necessary.
+\jb{Andy: Can you add some more here? You're the Leon pro! E.g. why/when do we think
+that our approach might be better?}
 
@@
 tools are optimized to work with problems that correspond to total functional
-programs. The QuickCheck tool for Haskell \cite{xxx} is an early example, based
-on random testing. Exhaustive testing and narrowing are other popular
-strategies \cite{xxx,yyy}. These tool are often much faster than model finders,
-but they typically cannot cope with underspecification and nonexecutable
-functions.
+programs. The QuickCheck tool for Haskell \cite{claessen-hughes-2000} is an early example, based
+on random testing. Bounded exhaustive testing \cite{runciman-et-al-2008} and
+narrowing \cite{lindblad-2008-testing} are other popular strategies. These tool
+are often much faster than model finders, but they typically cannot cope with
+underspecification and nonexecutable functions.
 

### 2015-05-06 tex/main.tex
more

@@
 
-In this paper, we consider \emph{annotated quantifed formulas} of the form
+In this paper, we consider \emph{annotated quantified formulas} of the form
 $\forallf{a} \vec x.\; \varphi$, where the annotation~$a$ is either
@@
 \rm
-A formula $\varphi$ is in \emph{definitional form with respect to} functions $\{
-\con{f}_1, \ldots, \con{f}_n \}$ if it is of the form
+A formula $\varphi$ is in \emph{definitional form with respect to} $n$ distinct function symbols
+$\{ \con{f}_1, \ldots, \con{f}_n \} \subseteq \sfuns{\Sigma}$ if it is of the form
 %
@@
 \begin{equation} \label{eq:wdf}
-(\forallf{\con{f}_1} \vec x_1.\; \varphi_1) \wedge \ldots \wedge (\forallf{\con{f}_n} \vec x_n.\; \varphi_n) \wedge \psi
+(\forallf{\con{f}_1} \vec x_1.\; \varphi_1) \wedge \cdots \wedge
+(\forallf{\con{f}_n} \vec x_n.\; \varphi_n) \wedge \psi
 \end{equation}
 %
-where $\con{f}_1, \ldots, \con{f}_n$ are not $\fnull$ and distinct,
-and $\psi$ contains no function definitions.
+where $\psi$ contains no function definitions.
 We call $\psi$ the \emph{conjecture} of $\varphi$.
@@
 \begin{definition}
-Given a set of function definitions $\DDD = \{ \forallf{f_1} \vec x.\; \varphi_1, \ldots, \forallf{f_n} \vec x.\; \varphi_n \}$,
-a formula $\varphi$ is \emph{closed under function expansion with respect to $\DDD$} if
+Given a set of function definitions $\DDD = \{ \forallf{f_1} \vec x.\;
+\varphi_1, \ldots, \forallf{f_n} \vec x.\; \varphi_n \}$, a formula $\varphi$
+is \emph{closed under function expansion with respect to $\DDD$} if
 \begin{eqnarray}
@@
 The set $\DDD$ is \emph{admissible} if whenever there exists a model for $\varphi$ that is closed under function expansion,
-there exists a model for $\varphi \wedge \forallf{f_1} \vec x.\; \varphi_1 \wedge \ldots \wedge \forallf{f_n} \vec x.\; \varphi_n$.
+there exists a model for $\varphi \wedge \forallf{f_1} \vec x.\; \varphi_1 \wedge \cdots \wedge \forallf{f_n} \vec x.\; \varphi_n$.
 \end{definition}
 
-For instance, the set $\{ \forallf{\con{f}} \vec x.\; \con{f}( x ) \teq 0 \}$ is admissible.
+For example, the set $\{ \forallf{\con{f}} \vec x.\; \con{f}( x ) \teq 0 \}$ is admissible.
 A set of admissible function definitions can include definitions of non-terminating functions, e.g.
@@
 
-For each of these 245 benchmarks, we considered (up to) three randomly selected mutated forms of its conjecture $\psi$.
-In particular, we considered unique conjectures that are obtained as a result of swapping a subterm of $\psi$ at one position
-with another of the same type at another position.
-Note that benchmarks created in this way have a high likelihood of having small and easy-to-find counterexamples.
-In total, we considered 213 mutated forms of conjectures from \isa, and 427 mutated forms of conjectures from \leon.
-We will call these sets \isam and \leonm respectively.
-
-In total, our benchmark set consisted of 885 benchmarks (640 mutants plus the original 245).
-We considered these 885 benchmarks both before and after the translation $\conv$ as described in Section~\ref{sec:encoding}.
-For example, \isa contains 79 original benchmarks $\varphi$
-and 79 translated benchmarks $\conv(\varphi)$.
+For each of these 245 benchmarks, we considered (up to) three randomly selected
+mutated forms of its conjecture $\psi$. In particular, we considered unique
+conjectures that are obtained as a result of swapping a subterm of $\psi$ at
+one position with another of the same type at another position.
+Note that benchmarks created in this way have a high likelihood of having small
+and easy-to-find counterexamples. In total, we considered 213 mutated forms of
+conjectures from \isa, and 427 mutated forms of conjectures from \leon. We will
+call these sets \isam and \leonm respectively.
+
+In total, our benchmark set consisted of 885 benchmarks (640 mutants plus the
+original 245). We considered these 885 benchmarks both before and after the
+translation $\conv$ as described in Section~\ref{sec:encoding}. For example,
+\isa contains 79 original benchmarks $\varphi$ and 79 translated benchmarks
+$\conv(\varphi)$.
 
@@
 as well as methods for finding models for quantified formulas~\cite{GeDeM-CAV-09}.
-We also considered three configurations of \cvc~\cite{barrett-et-al-2011} which we will refer to as \cvcd, \cvcf, and \cvci.
+We also considered three configurations of \cvc~\cite{barrett-et-al-2011} which
+we will refer to as \cvcd, \cvcf, and \cvci.
 The default configuration \cvcd runs heuristic quantifier instantiation,
 but does not include techniques for finding models.
-The configuration \cvcf runs heuristic instantiation and the finite model finding procedure described in~\cite{ReyEtAl-1-RR-13,reynolds-et-al-2013}.
-The last configuration \cvci incorporates techniques for automating inductive reasoning in SMT~\cite{reynolds-kuncak-2015}.
+The configuration \cvcf runs heuristic instantiation and the finite model
+finding procedure described in~\cite{ReyEtAl-1-RR-13,reynolds-et-al-2013}. The
+last configuration \cvci incorporates techniques for automating inductive
+reasoning in SMT~\cite{reynolds-kuncak-2015}.
 %All configurations of \cvc incorporate techniques for conflict-based instantiation~\cite{} ..?

### 2015-05-06 tex/main.tex
tuned

@@
 
-\begin{definition}%
-\rm
+\begin{definition}\rm
 A formula $\varphi$ is in \emph{definitional form with respect to} $n$ distinct function symbols
@@
 
-\begin{definition}
+\begin{definition}\rm
 Given a set of function definitions $\DDD = \{ \forallf{f_1} \vec x.\;
@@
 is \emph{closed under function expansion with respect to $\DDD$} if
-\begin{eqnarray}
-\varphi \models_T \bigwedge_{i=1}^n \{ \varphi_i[ \vec t / \vec x ] \mid f_i( \vec t ) \in \terms( \varphi ) \}
-\end{eqnarray}
+\[\varphi \models_T \bigwedge_{i=1}^n \{ \varphi_i[ \vec t / \vec x ] \mid f_i( \vec t ) \in \terms( \varphi ) \}\]
 The set $\DDD$ is \emph{admissible} if whenever there exists a model for $\varphi$ that is closed under function expansion,

### 2015-05-06 tex/main.tex
renamed macros

@@
 
-%\newcommand{\ssorts}[1]{#1^\mathrm{s}}
-\newcommand{\ssorts}[1]{#1^\mathrm{y}}
+%\newcommand{\stypes}[1]{#1^\mathrm{s}}
+\newcommand{\stypes}[1]{#1^\mathrm{y}}
 
@@
 %\newcommand{\sfunndefs}[1]{#1^\mathrm{nf}}
-\newcommand{\sortint}{\ty{Int}}
+\newcommand{\typeint}{\ty{Int}}
 
@@
 \newcommand{\fargx}[2]{\concret^{\vthinspace#2}_{#1}}
-\newcommand{\fargsort}[1]{\alpha_{#1}}
+\newcommand{\fargtype}[1]{\alpha_{#1}}
 
@@
 However, these generally cannot cope with universal quantifiers ranging over
-infinite types, such as integers and recursive datatypes.
-This paper introduces a translation that eliminates infinite
-quantifiers in axioms specifying a large class of recursive functions,
-including terminating, tail-recursive, and productive functions. An empirical
-evaluation confirms that the approach works well on benchmarks from two
-independent sources. The translation is implemented as a preprocessor in the
-solver CVC4.
+infinite types, such as integers and recursive datatypes. This paper introduces
+a translation that eliminates infinite quantifiers in axioms specifying a large
+class of recursive functions, including terminating, tail-recursive, and
+productive functions. An empirical evaluation confirms that the approach works
+well on benchmarks from two sources. The translation is implemented as a
+preprocessor in the solver CVC4.
 \end{abstract}
@@
 \cite{MouraBjoerner07,ReynoldsTinelliMoura14}.
-These methods work well in the context of proving (i.e., showing
-unsatisfiability), but they generally cannot be used to find models (i.e.,
-showing satisfiability). Typically, a single universal quantifier in one of the
+These methods work well in the context of proving (i..e, showing
+unsatisfiability), but they are of little help for finding models (i.e.,
+showing satisfiability). Often, a single universal quantifier in one of the
 axioms of a problem is enough to prevent the discovery of models.
@@
 occurrences of variables occur as direct subterms of uninterpreted
-functions---e.g., $\forall x%\typ{\sortint} --- the sort is not important, right?
+functions---e.g., $\forall x%\typ{\typeint} --- the type is not important, right?
 .\;\, \con{f}( x )
@@
 this fragment slightly, including cases when variables occur as arguments of
-arithmetic predicates. More recently, Reynolds et al.\
+arithmetic predicates. Subsequently, Reynolds et al.\
 \cite{ReyEtAl-1-RR-13,reynolds-et-al-2013} introduced techniques for finding finite
@@
 These techniques can
-find a model for the formula $\forall x,\, y : \tau.\;\, x \teq
+find a model for a formula such as $\forall x,\, y : \tau.\;\, x \teq
 y \vee \allowbreak \neg \con{f}( x ) \teq \con{f}( y )$, where $\tau$ is uninterpreted.
@@
 recursive datatypes. A simple example where both approaches fail is
-$$\forall n : {\sortint}.\;\, \lite\bigl(
+$$\forall n : {\typeint}.\;\, \lite\bigl(
 n \leq 0,\allowbreak\; \con{p}( n ) \teq 1,\allowbreak\; \con{p}( n ) \teq 2 * \con{p}( n - 1 ) \bigr)$$
@@
 productive corecursive functions. It is, however, not met by inconsistent
-definitions such as $\forall n : {\sortint}.\;\, \con{f}(n) \teq \con{f}(n) +
+definitions such as $\forall n : {\typeint}.\;\, \con{f}(n) \teq \con{f}(n) +
 1$, and it is possible to construct consistent examples that do not meet it.
@@
 $\forall a :
-\fargsort{\tau}.\; \const{f}(\farg{\con{f}}(a)) \teq t[\concret(a)/x]$, where
-$\fargsort{\tau}$ is an uninterpreted ``abstract'' type and $\farg{\con{f}} :
-\fargsort{\tau} \to \tau$ a coercion from the abstract type to the concrete
+\fargtype{\tau}.\; \const{f}(\farg{\con{f}}(a)) \teq t[\concret(a)/x]$, where
+$\fargtype{\tau}$ is an uninterpreted ``abstract'' type and $\farg{\con{f}} :
+\fargtype{\tau} \to \tau$ a coercion from the abstract type to the concrete
 type. The translation preserves satisfiability and, for admissible definitions,
@@
 A \emph{signature} $\Sigma$ consists of
-a set $\ssorts{\Sigma}$ of type symbols (or sorts) and
+a set $\stypes{\Sigma}$ of type symbols (or sorts) and
 a set $\sfuns{\Sigma}$ of function symbols.
@@
 an infix equality predicate ${\teq} : \tau \times \tau \to \Bool$
-for each $\tau \in \ssorts{\Sigma}$,
+for each $\tau \in \stypes{\Sigma}$,
 standard boolean connectives ($\neg$, $\wedge$, $\vee$, etc.),
-and a function $\lite : \Bool \times \tau \times \tau \rightarrow \tau$ for each $\tau \in \ssorts{\Sigma}$.
+and a function $\lite : \Bool \times \tau \times \tau \rightarrow \tau$ for each $\tau \in \stypes{\Sigma}$.
 We write $t\typ{\tau}$ to denote a metavariable ranging over terms of
@@
 A \emph{$\Sigma$-interpretation $\I$} %is a mathematical structure that
-maps each $\tau \in \ssorts{\Sigma}$ to a nonempty set $\tau^\I$,
+maps each $\tau \in \stypes{\Sigma}$ to a nonempty set $\tau^\I$,
 the \emph{domain} of~$\tau$ in $\I$,
@@
 %
-\begin{equation} \label{eq:wdf}
-(\forallf{\con{f}_1} \vec x_1.\; \varphi_1) \wedge \cdots \wedge
-(\forallf{\con{f}_n} \vec x_n.\; \varphi_n) \wedge \psi
-\end{equation}
+\[(\forallf{\con{f}_1} \vec x_1.\; \varphi_1) \wedge \cdots \wedge
+(\forallf{\con{f}_n} \vec x_n.\; \varphi_n) \wedge \psi\]
 %
@@
 Given a signature $\Sigma$,
-we distinguish the subset of \emph{defined} function symbols $\sfundefs{\Sigma}
-\subseteq \sfuns{\Sigma}$.
+we distinguish the subset of (\emph{recursively}) \emph{defined} function
+symbols $\sfundefs{\Sigma} \subseteq \sfuns{\Sigma}$.
 We consider $\Sigma$-formulas that are in definitional form with respect to
@@
 is \emph{closed under function expansion with respect to $\DDD$} if
-\[\varphi \models_T \bigwedge_{i=1}^n \{ \varphi_i[ \vec t / \vec x ] \mid f_i( \vec t ) \in \terms( \varphi ) \}\]
+\[\varphi \,\models_T\, \bigwedge\nolimits_{\,i=1}^{n} \{ \varphi_i[ \vec t / \vec x ] \mid f_i( \vec t ) \in \terms( \varphi ) \}\]
 The set $\DDD$ is \emph{admissible} if whenever there exists a model for $\varphi$ that is closed under function expansion,
@@
 A set of admissible function definitions can include definitions of non-terminating functions, e.g.
-$\{ \forallf{\con{f}} x : \sortint.\;\, \con{f}( x ) \teq \con{f}( x - 1 ) \}$ or even $\{ \forallf{\con{f}} x.\;\, \con{f}( x ) \teq \con{f}( x ) \}$ are admissible.
+$\{ \forallf{\con{f}} x : \typeint.\;\, \con{f}( x ) \teq \con{f}( x - 1 ) \}$ or even $\{ \forallf{\con{f}} x.\;\, \con{f}( x ) \teq \con{f}( x ) \}$ are admissible.
 On the other hand, the set $\{ \forallf{\con{f}} x.\;\, \con{f}( x ) \teq \con{f}( x ) + 1 \}$ is not admissible,
@@
 %An admissible function definition need not specify a terminating function, e.g.
-%$\{ \forallf{\con{f}} x^\sortint.\;\, \con{f}( x ) \teq \con{f}( x - 1 ) \}$ or even $\{ \forallf{\con{f}} x.\;\, \con{f}( x ) \teq \con{f}( x ) \}$ are admissible.
+%$\{ \forallf{\con{f}} x^\typeint.\;\, \con{f}( x ) \teq \con{f}( x - 1 ) \}$ or even $\{ \forallf{\con{f}} x.\;\, \con{f}( x ) \teq \con{f}( x ) \}$ are admissible.
 
@@
 \begin{itemize}
-\item an uninterpreted \emph{abstract type} $\fargsort{\con{f}}$;
-\item $n$ uninterpreted \emph{concretization functions} $\fargx{\con{f}}{1} : \fargsort{\con{f}} \rightarrow \tau_1$, \ldots, $\fargx{\con{f}}{n} : \fargsort{\con{f}} \rightarrow \tau_n$.
+\item an uninterpreted \emph{abstract type} $\fargtype{\con{f}}$;
+\item $n$ uninterpreted \emph{concretization functions} $\fargx{\con{f}}{1} : \fargtype{\con{f}} \rightarrow \tau_1$, \ldots, $\fargx{\con{f}}{n} : \fargtype{\con{f}} \rightarrow \tau_n$.
 \end{itemize}
-%The interpretation of uninterpreted sort $\fargsort{\con{f}}$ will denote the elements (tuples) on which the function $\con{f}$ is applied.
+%The interpretation of uninterpreted type $\fargtype{\con{f}}$ will denote the elements (tuples) on which the function $\con{f}$ is applied.
 %The role of the uninterpreted functions $\farg{\con{f}}^1$, $\ldots$, $\farg{\con{f}}^n$ will be discussed more in the following.
@@
       %\item[] $\mathsf{assert}( \chi = \ltrue )$;
-      \item[] \return \bigtuple{$\forall y : \fargsort{\con{f}}.\; u' [ \vecfarg{\con{f}}( y ) / \vec x ],\, \ltrue$}
+      \item[] \return \bigtuple{$\forall y : \fargtype{\con{f}}.\; u' [ \vecfarg{\con{f}}( y ) / \vec x ],\, \ltrue$}
     \end{itemize}
@@
    \begin{itemize}
-     \item[] \return \bigtuple{$t,\; \bigwedge\, \{ \exists a : \fargsort{\con{f}}.\; \vecfarg{\con{f}}( a ) \teq \vec t \mid \con{f}( \vec t ) \in \terms( t ),\, \con{f} \in \sfundefs{\Sigma} \}$}
+     \item[] \return \bigtuple{$t,\; \bigwedge\, \{ \exists a : \fargtype{\con{f}}.\; \vecfarg{\con{f}}( a ) \teq \vec t \mid \con{f}( \vec t ) \in \terms( t ),\, \con{f} \in \sfundefs{\Sigma} \}$}
    \end{itemize}
@@
 quantification on function definitions for $\con{f}$ to the corresponding
-uninterpreted type $\fargsort{\con{f}}$ and inserts appropriate coercions
+uninterpreted type $\fargtype{\con{f}}$ and inserts appropriate coercions
 $\fargx{\con{f}}{i}$. Second, it augments $\varphi$ with additional exisential
 constraints
-$\exists a : \fargsort{\con{f}}.\;\, \vecfarg{\con{f}}( a ) \teq \vec t$; these
+$\exists a : \fargtype{\con{f}}.\;\, \vecfarg{\con{f}}( a ) \teq \vec t$; these
 ensure that the restricted definition for $\con{f}$ covers all relevant tuples
@@
 $\sfundefs{\Sigma}$, $\conv_0$ constructs a quantified formula over a single variable
-$y$ of type $\fargsort{\con{f}}$ and replaces all occurrences of $\vec x$ in
+$y$ of type $\fargtype{\con{f}}$ and replaces all occurrences of $\vec x$ in
 $t'_1$ with $\vecfarg{\con{f}}( y )$.Since function definitions
@@
 of a type other than $\Bool$. The returned constraint is a conjunction of
-formulas of the form $\exists a : {\fargsort{\con{f}}}.\;\, \vecfarg{\con{f}}(
+formulas of the form $\exists a : {\fargtype{\con{f}}}.\;\, \vecfarg{\con{f}}(
 a ) \teq \vec t$ for each subterm $\con{f}( \vec t )$ of $t$ such that $\con{f}
@@
 functions $\Sigma = \Sigmalia \mathrel{\cup} \Sigma_{\euf}$, where
-$\ssorts{\Sigma_{\euf}} = \{ \sortint \}$ and
-$\sfuns{\Sigma_{\euf}} = \{ \con{s} : { \sortint \rightarrow \sortint },\, \const{c} : { \sortint } \}$.
+$\stypes{\Sigma_{\euf}} = \{ \typeint \}$ and
+$\sfuns{\Sigma_{\euf}} = \{ \con{s} : { \typeint \rightarrow \typeint },\, \const{c} : { \typeint } \}$.
 Let $\varphi$ be the $\Sigma$-formula
 \begin{equation} \label{eq:ex-before}
-\forallf{\con{s}} n : {\sortint}.\;\, \lite\bigl( n \leq 0,\; \con{s}(n) \teq 0,\;
+\forallf{\con{s}} n : {\typeint}.\;\, \lite\bigl( n \leq 0,\; \con{s}(n) \teq 0,\;
   \con{s}( n ) \teq n + \con{s}( n - 1 ) \bigr) \wedge \con{s}( \con{c} ) > 100
@@
 a model for this formula. The signature $\extendsig{\Sigma}$ includes the type
-$\fargsort{\con{s}}$ and the uninterpreted function $\farg{\con{s}}$ of type $\fargsort{\con{s}}
-\rightarrow \sortint$. The result of $\conv( \varphi,\, \ppos )$ is the pair $(
+$\fargtype{\con{s}}$ and the uninterpreted function $\farg{\con{s}}$ of type $\fargtype{\con{s}}
+\rightarrow \typeint$. The result of $\conv( \varphi,\, \ppos )$ is the pair $(
 \psi,\, \ltrue )$, where (after simplification) $\psi$ is the
@@
   & \phantom{{\wedge}\; }\bigl(
-      \forall a : \fargsort{\con{s}}.\; \lite\bigl(
+      \forall a : \fargtype{\con{s}}.\; \lite\bigl(
         \!\begin{aligned}[t]
@@
   & \con{s}(\farg{\con{s}}( a )) \teq \farg{\con{s}}( a ) + \con{s}( \farg{\con{s}}( a )-1 )
-    \wedge \exists b : {\fargsort{\con{s}}}.\;\, \farg{\con{s}}( b ) \teq \farg{\con{s}}( a )-1 \bigr) \bigr)
+    \wedge \exists b : {\fargtype{\con{s}}}.\;\, \farg{\con{s}}( b ) \teq \farg{\con{s}}( a )-1 \bigr) \bigr)
 \end{aligned}
 \\[-\jot]
- & {\wedge}\; \bigl( \con{s}( \con{c} ) > 100 \wedge \exists a : {\fargsort{\con{s}}}.\;\, \farg{\con{s}}( a ) \teq \con{c} \bigr)
+ & {\wedge}\; \bigl( \con{s}( \con{c} ) > 100 \wedge \exists a : {\fargtype{\con{s}}}.\;\, \farg{\con{s}}( a ) \teq \con{c} \bigr)
 \end{aligned}
@@
 The universal quantification in the resulting formula is over an uninterpreted
-type $\fargsort{\con{s}}$, thus enabling the finite model finding
+type $\fargtype{\con{s}}$, thus enabling the finite model finding
 techniques by Reynolds et al.\ \cite{ReyEtAl-1-RR-13,reynolds-et-al-2013} to search for a finite interpretation for
-$\fargsort{\con{s}}$. Moreover, all occurrences of $a$ in this formula are beneath
+$\fargtype{\con{s}}$. Moreover, all occurrences of $a$ in this formula are beneath
 applications of the uninterpreted function $\farg{\con{s}}$, implying that the
@@
 One model $\I$ for formula~(\ref{eq:ex-after}) in the previous example interprets
-$\fargsort{\con{s}}$ as a finite set of elements $\{ u_0, \ldots, u_{14} \}$,
+$\fargtype{\con{s}}$ as a finite set of elements $\{ u_0, \ldots, u_{14} \}$,
 $\farg{\con{s}}$ as a finite map $u_i \mapsto i$ for $i = 0, \ldots, 14$,
@@
 \begin{equation} \label{eq:approx-interp}
-\lambda x : \sortint.\;\, \lite( x \mathbin{\teq} 0,\, 0,\,
+\lambda x : \typeint.\;\, \lite( x \mathbin{\teq} 0,\, 0,\,
   \lite( x \mathbin{\teq} 1,\, 1,\,
@@
 First, we show that if $\varphi$ is satisfied by $\Sigma$-interpretation $\I$, then $\conv( \varphi )$ is satisfied by a $\extendsig{\Sigma}$-interpretation $\J$.
-Let $\J$ be the $\extendsig{\Sigma}$-interpretation that interprets all types $\tau \in \ssorts{\Sigma}$ as $\tau^\I$,
+Let $\J$ be the $\extendsig{\Sigma}$-interpretation that interprets all types $\tau \in \stypes{\Sigma}$ as $\tau^\I$,
 all functions $\con{f} \in \sfuns{\Sigma}$ as $\con{f}^\I$,
 and for each function $\con{f}$ of type $\tau_1 \times \cdots \tau_n \rightarrow \tau$ in $\sfundefs{\Sigma}$,
-interprets $\fargsort{\con{f}}$ as a tuple of values of type $\tau_1^\I \times \cdots \times \tau_n^\I$,
+interprets $\fargtype{\con{f}}$ as a tuple of values of type $\tau_1^\I \times \cdots \times \tau_n^\I$,
 and each $\fargx{\con{f}}{i}$ as the $i^{th}$ projection on such tuples for each $i = 1, \ldots, n$.
-Note that $\J$ satisfies \emph{every} constraint of the form $\exists a : {\fargsort{\con{f}}}. ( \vecfarg{\con{f}}( a ) \teq \vec t )$,
+Note that $\J$ satisfies \emph{every} constraint of the form $\exists a : {\fargtype{\con{f}}}. ( \vecfarg{\con{f}}( a ) \teq \vec t )$,
 since by our construction of $\J$ there exists an $a$ such that $a^\J = \vec t^\J$.
@@
 we first construct a sequence of $\extendsig{\Sigma}$-literals $M_0, M_1, \ldots$ such that $\I$ satisfies each $M_i$,
-and a sequence of mappings $C^{\con f}_0, C^{\con f}_1, \ldots$ from values $\fargsort{f}^\I$ 
+and a sequence of mappings $C^{\con f}_0, C^{\con f}_1, \ldots$ from values $\fargtype{f}^\I$ 
 to tuples of terms of type $\tau_1 \times \cdots \times \tau_n$ 
@@
 For each $i \geq 0$, 
-let $X_i$ be the set of constraints of the form $\exists a : {\fargsort{\con{f}}}. ( \vecfarg{\con{f}}( a ) \teq \vec t )$ occurring in $M_i$
+let $X_i$ be the set of constraints of the form $\exists a : {\fargtype{\con{f}}}. ( \vecfarg{\con{f}}( a ) \teq \vec t )$ occurring in $M_i$
 and let $L_i$ be the others in $M_i$.
 For each $\con{f} \in \sfundefs{\Sigma}$, let $C^{\con{f}}_{i+1}$ be a mapping containing $C^{\con{f}}_i$ and a minimal number of entries
-such that for all $\exists a : {\fargsort{\con{f}}}. ( \vecfarg{\con{f}}( a ) \teq \vec t ) \in X_i$,
+such that for all $\exists a : {\fargtype{\con{f}}}. ( \vecfarg{\con{f}}( a ) \teq \vec t ) \in X_i$,
 there exists a $u \mapsto \vec s \in C^{\con{f}}_{i+1}$ such that $\vecfarg{\con{f}}^\I( u ) = \vec s^\I = \vec t^\I$.
@@
 Example~\ref{ex:translation}. This formula is unsatisfiable if and only if
-formula~(\ref{eq:ex-before}) is unsatisfiable since $\fargsort{f}$ can be
-interpreted as $\sortint$, and $\farg{f}$ can be interpreted as the identity
+formula~(\ref{eq:ex-before}) is unsatisfiable since $\fargtype{f}$ can be
+interpreted as $\typeint$, and $\farg{f}$ can be interpreted as the identity
 function. Conversely, we claim formula~(\ref{eq:ex-after}) is satisfiable if
 and only if formula~(\ref{eq:ex-before}) is satisfiable, noting that the set
-$\{ \forallf{\con{s}} x\typ{\sortint}.\;\, \lite\bigl( x \leq 0,\allowbreak\; \con{s}(x)
+$\{ \forallf{\con{s}} x\typ{\typeint}.\;\, \lite\bigl( x \leq 0,\allowbreak\; \con{s}(x)
 \teq 0,\; \con{s}( x ) \teq x + \con{s}( x-1 ) \bigr) \}$ is admissble.
@@
 \lite\bigl( x \leq 0, \con{s}(x) \teq 0, \con{s}( x ) \teq x + \con{s}( x-1 )
-\bigr)$. The rightmost conjunct $\exists a : {\fargsort{\con{s}}}.
+\bigr)$. The rightmost conjunct $\exists a : {\fargtype{\con{s}}}.
 \farg{\con{s}}( a ) \teq \con{c}$ ensures that any model of

### 2015-05-06 tex/main.tex
more

@@
 
+Our setting is a monomorphic (or many-sorted) first-order logic, such as
+the one defined by SMT-LIB \cite{smtlib25}.
 A \emph{signature} $\Sigma$ consists of
@@
 We assume that signatures always include a Boolean type $\Bool$ and constants
-$\ltrue$ (truth) and $\lfalse$ (falsity) of that type,
+$\ltrue, \lfalse : \Bool$ for truth and falsity,
 an infix equality predicate ${\teq} : \tau \times \tau \to \Bool$
 for each $\tau \in \stypes{\Sigma}$,
-standard boolean connectives ($\neg$, $\wedge$, $\vee$, etc.),
-and a function $\lite : \Bool \times \tau \times \tau \rightarrow \tau$ for each $\tau \in \stypes{\Sigma}$.
+standard Boolean connectives ($\neg$, $\wedge$, $\vee$, etc.),
+and a function symbol $\lite : \Bool \times \tau \times \tau \rightarrow \tau$ (``if--then--else'')
+for each $\tau \in \stypes{\Sigma}$.
 We write $t\typ{\tau}$ to denote a metavariable ranging over terms of
@@
 A \emph{$\Sigma$-interpretation $\I$} %is a mathematical structure that
-maps each $\tau \in \stypes{\Sigma}$ to a nonempty set $\tau^\I$,
-the \emph{domain} of~$\tau$ in $\I$,
-and each $\con{f} : \tau_1 \times \cdots \times \tau_n \rightarrow \tau$ in
+maps each type $\tau \in \stypes{\Sigma}$ to a nonempty set $\tau^\I$,
+the \emph{domain} of~$\tau$ in~$\I$,
+and each function symbol $\con{f} : \tau_1 \times \cdots \times \tau_n \rightarrow \tau$ in
 $\sfuns{\Sigma}$
@@
 For example, the set $\{ \forallf{\con{f}} \vec x.\; \con{f}( x ) \teq 0 \}$ is admissible.
-A set of admissible function definitions can include definitions of non-terminating functions, e.g.
+A set of admissible function definitions can include definitions of nonterminating functions, e.g.
 $\{ \forallf{\con{f}} x : \typeint.\;\, \con{f}( x ) \teq \con{f}( x - 1 ) \}$ or even $\{ \forallf{\con{f}} x.\;\, \con{f}( x ) \teq \con{f}( x ) \}$ are admissible.

### 2015-05-06 tex/main.tex
more

@@
 
+\renewcommand\models{\mathrel{\vDash}}
+
 %\newcommand{\DDD}{\mathcalx{D}}
@@
 We define a translation for a class of
-formulas containing admissible recursive function definitions
-(Section~\ref{sec:encoding}) that produces formulas for which existing model
-finding techniques are applicable. The main insight is that a
+formulas involving admissible recursive function definitions
+(Section~\ref{sec:encoding}). The main insight is that a
 recursive definition $\forall x : \tau.\; \const{f}(x) \teq t$
@@
 (Section~\ref{sec:evaluation}). The approach is implemented as a preprocessor
-in \cvc. When invoked with the appropriate command-line option,
-the solver assumes functions introduced using the \texttt{define-fun-rec}
-command are admissible and translates them accordingly
-(Section~\ref{sec:front-end}).
+in \cvc (Section~\ref{sec:front-end}). When invoked with the appropriate
+command-line option, the solver assumes that functions introduced using the
+\texttt{define-fun-rec} command are admissible and translates them accordingly.
 
@@
 type~$\tau$ and $\terms( t )$ to denote the set of subterms in $t$.
+Formulas are terms of type $\Bool$.
+Given a term $u$, we write $u[\vec t/\vec x]$ to denote the result of replacing
+all occurrences of $\vec x$ with $\vec t$ in $u$.
+
 %We write $\functerms^\Sigma( t )$ to denote the set of $f$-applications in $t$ such that $f \in \sfundefs{\Sigma}$.
@@
 otherwise, it is \emph{$T$-unsatisfiable}.
-A formula $\varphi$ entails $\psi$, written $\varphi \models_T \psi$,
+A formula $\varphi$ \emph{entails} $\psi$, written $\varphi \models_T \psi$,
 if all interpretations in $\Mo$ that satisfy $\varphi$ also satisfy $\psi$.
@@
 In this paper, we consider \emph{annotated quantified formulas} of the form
-$\forallf{a} \vec x.\; \varphi$, where the annotation~$a$ is either
-an uninterpreted function $\con{f} \in \sfuns{\Sigma}$ or the special symbol
-$\fnull$.
+$\forallf{\const{f}} \vec x.\; \varphi$, where $\con{f} \in \sfuns{\Sigma}$ is
+an uninterpreted function symbol, in addition to standard quantified formulas
+$\forall \vec x.\; \varphi$.
 Given $\con{f} : \tau_1 \times \cdots \times \tau_n \rightarrow \tau$,
@@
 $\tau_1, \ldots, \tau_n$ and $\varphi$ is a (quantifier-free) formula equivalent to
-$\con{f}( \vec x ) \teq t$ for some $t$. We write $\forall \vec
-x.\; \varphi$ as shorthand for $\forallf{\fnull} \vec x.\; \varphi$, and $\exists
-\vec x. \varphi$ as shorthand for $\neg \forall \vec x.\; \neg \varphi$.
-For a formula $\varphi$,
-we write $\varphi[\vec t/\vec x]$ to denote the result of replacing all occurrences of $\vec x$ with $\vec t$.
+$\con{f}( \vec x ) \teq t$ for some $t$. We write $\exists
+\vec x.\; \varphi$ as shorthand for $\neg \forall \vec x.\; \neg \varphi$.
 
 \begin{definition}\rm
-A formula $\varphi$ is in \emph{definitional form with respect to} $n$ distinct function symbols
-$\{ \con{f}_1, \ldots, \con{f}_n \} \subseteq \sfuns{\Sigma}$ if it is of the form
+A formula $\varphi$ is in \emph{definitional form with respect to}
+$\{ \con{f}_1, \ldots, \con{f}_n \}$ if it is of the form
 %
@@
 %
-where $\psi$ contains no function definitions.
+where $\con{f}_1, \ldots, \con{f}_n \in \sfuns{\Sigma}$ are distinct
+and $\psi$ contains no function definitions.
 We call $\psi$ the \emph{conjecture} of $\varphi$.
@@
 We consider $\Sigma$-formulas that are in definitional form with respect to
-functions $\sfundefs{\Sigma}$.
+$\sfundefs{\Sigma}$.
 
 \begin{definition}\rm
-Given a set of function definitions $\DDD = \{ \forallf{f_1} \vec x.\;
-\varphi_1, \ldots, \forallf{f_n} \vec x.\; \varphi_n \}$, a formula $\varphi$
+Given a set of function definitions $\DDD = \{ \forallf{\con{f}_1} \vec x.\;
+\varphi_1, \ldots, \forallf{\con{f}_n} \vec x.\; \varphi_n \}$, a formula $\varphi$
 is \emph{closed under function expansion with respect to $\DDD$} if
-\[\varphi \,\models_T\, \bigwedge\nolimits_{\,i=1}^{n} \{ \varphi_i[ \vec t / \vec x ] \mid f_i( \vec t ) \in \terms( \varphi ) \}\]
+\[\varphi \,\models_T\, \bigwedge\nolimits_{\,i=1}^{n} \{ \varphi_i[ \vec t / \vec x ] \mid \con{f}_i( \vec t ) \in \terms( \varphi ) \}\]
 The set $\DDD$ is \emph{admissible} if whenever there exists a model for $\varphi$ that is closed under function expansion,
-there exists a model for $\varphi \wedge \forallf{f_1} \vec x.\; \varphi_1 \wedge \cdots \wedge \forallf{f_n} \vec x.\; \varphi_n$.
+there exists a model for $\varphi \wedge \forallf{\con{f}_1} \vec x.\; \varphi_1 \wedge \cdots \wedge \forallf{\con{f}_n} \vec x.\; \varphi_n$.
 \end{definition}
 
-For example, the set $\{ \forallf{\con{f}} \vec x.\; \con{f}( x ) \teq 0 \}$ is admissible.
-A set of admissible function definitions can include definitions of nonterminating functions, e.g.
-$\{ \forallf{\con{f}} x : \typeint.\;\, \con{f}( x ) \teq \con{f}( x - 1 ) \}$ or even $\{ \forallf{\con{f}} x.\;\, \con{f}( x ) \teq \con{f}( x ) \}$ are admissible.
+For example, the set $\{ \forallf{\con{f}} \vec x.\;\, \con{f}( x ) \teq 0 \}$ is admissible.
+An admissible set can include definitions of nonterminating functions, e.g.
+$\{ \forallf{\con{f}} x : \typeint.\;\, \con{f}( x ) \teq \con{f}( x - 1 ) \}$ or even $\{ \forallf{\con{f}} x.\allowbreak\;\, \con{f}( x ) \teq \con{f}( x ) \}$ are admissible.
 On the other hand, the set $\{ \forallf{\con{f}} x.\;\, \con{f}( x ) \teq \con{f}( x ) + 1 \}$ is not admissible,

### 2015-05-06 tex/main.tex
more

@@
 
+\let\oldchi=\chi
+\def\chi{{\vthinspace\oldchi}}
+
 \let\oldSigma=\Sigma
@@
 
-\newcommand{\pnone}{\con{apolar}}
+\newcommand{\pnone}{\con{none}}
 \newcommand{\ppos}{\con{pos}}
@@
 \newcommand{\vthinspace}{\kern+0.083333em}
+\newcommand{\vvthinspace}{\kern+0.0416666em}
 \newcommand{\typ}[1]{^{\vthinspace #1}}
@@
 
-Let us fix a signature $\Sigma$, a $\Sigma$-formula
+First, we fix a signature $\Sigma$ and a $\Sigma$-formula
 $\varphi$ in definitional form with respect to $\sfundefs{\Sigma}$,
@@
 This section presents a method for constructing an extended signature
-$\extendsig{ \Sigma }$ and a $\extendsig{ \Sigma }$-formula $\psi$ such that
+$\extendsig{ \Sigma }$ and an $\extendsig{ \Sigma }$-formula $\psi$ such that
 $\psi$ is satisfiable if and only if $\varphi$ is satisfiable---i.e.,
-$\varphi$ and $\psi$ are \emph{equisatisfiable}.
+$\varphi$ and $\psi$ are \relax{equisatisfiable}.
 The intuition behind this translation
@@
 
-In more detail,
+More precisely,
 suppose our signature $\Sigma$ contains a set of defined function symbols $\sfundefs{\Sigma} \subseteq \sfuns{\Sigma}$.
@@
 on the auxiliary function $\conv_0$, which takes two arguments:\ the term $t$
-to translate and a \emph{polarity} $p$, which is either $\ppos$, $\pneg$, or
-$\pnone$. It returns a pair $( t'\negvthinspace,\, \chi )$, where $t'$ is a term of
+to translate and a \relax{polarity}~$p$, which is either $\ppos$, $\pneg$, or
+$\pnone$. \vthinspace$\conv_0$ returns a pair $( t'\negvthinspace,\, \chi )$, where $t'$ is a term of
 type $\tau$ and $\chi$ is an additional constraint.
-This constraint is necessary if the polarity is $\pnone$; otherwise, any necessary
-constraint is encoded directly in $t'$ and $\chi$ is simply $\ltrue$.
-
-\jb{Merge para above with para below}
-
-A translation function $\conv$ for a $\Sigma$-formula that is in definitional
-form with respect to a set of admissible function definitions
-$\sfundefs{\Sigma}$. In the auxiliary function $\conv_0$, the function $\pol(
-\boolop, i, p )$ returns the polarity of the $i${th} argument of a
+This constraint is needed if the polarity is $\pnone$; otherwise, it is encoded
+directly in $t'$ and $\chi$ is simply $\ltrue$.
+In the auxiliary function $\conv_0$, the function $\pol(
+\boolop, i, p )$ returns the polarity of the $i$\vvthinspace{th} argument of a
 $\boolop$ application having polarity $p$, where $\boolop$ is an interpreted
@@
       %\item[] $\mathsf{assert}( \chi = \ltrue )$;
-      \item[] \return \bigtuple{$\forall y : \fargtype{\con{f}}.\; u' [ \vecfarg{\con{f}}( y ) / \vec x ],\, \ltrue$}
+      \item[] \return \bigtuple{$\forall a : \fargtype{\con{f}}.\; u' [ \vecfarg{\con{f}}( a ) / \vec x ],\, \ltrue$}
     \end{itemize}
@@
  \begin{itemize}
-   \item[] $\mathsf{let}$ $( \psi,\, \_ ) = \conv_0( \varphi,\, \ppos )$ $\mathsf{in}$
+   \item[] $\mathsf{let}$ $( \varphi',\, \_ ) = \conv_0( \varphi,\, \ppos )$ $\mathsf{in}$
    %\item[] $\mathsf{assert}( D = \emptyset )$;
-   \item[] \return $\psi$
+   \item[] \return $\varphi'$
   \end{itemize}
@@
 
-The translation modifies $\varphi$ in two main ways. First, it restricts the
+The translation alters $\varphi$ in two main ways. First, it restricts the
 quantification on function definitions for $\con{f}$ to the corresponding
@@
 $\sfundefs{\Sigma}$, $\conv_0$ constructs a quantified formula over a single variable
-$y$ of type $\fargtype{\con{f}}$ and replaces all occurrences of $\vec x$ in
-$t'_1$ with $\vecfarg{\con{f}}( y )$.Since function definitions
+$a$ of type $\fargtype{\con{f}}$ and replaces all occurrences of $\vec x$ in
+$t'_1$ with $\vecfarg{\con{f}}( a )$. Since function definitions
 are a top-level conjunct, %by case analysis on the return values of $\conv$,

### 2015-05-06 tex/main.tex
some vertical space
more

@@
 
-\newcommand{\pnone}{\con{none}}
-\newcommand{\ppos}{\con{pos}}
-\newcommand{\pneg}{\con{neg}}
+%\newcommand{\pnone}{\con{none}}
+%\newcommand{\ppos}{\con{pos}}
+%\newcommand{\pneg}{\con{neg}}
+%%% @ANDY: as usual, revert if you don't like
+\newcommand{\pnone}{{\pm}}
+\newcommand{\ppos}{{+}}
+\newcommand{\pneg}{{-}}
 \newcommand{\pol}{\con{pol}}
@@
 In the auxiliary function $\conv_0$, the function $\pol(
-\boolop, i, p )$ returns the polarity of the $i$\vvthinspace{th} argument of a
-$\boolop$ application having polarity $p$, where $\boolop$ is an interpreted
+\boolop, i, p )$ returns the polarity of the $i$\vvthinspace{th} argument of an
+application of $\boolop$ having polarity $p$, where $\boolop$ is an interpreted
 predicate.
@@
 $\conv_0( t\typ{\tau},\, p )$ $=$
+\\[-.8\baselineskip]
  \begin{itemize}
@@
         \item[] \return \bigtuple{$\boolop(t'_1, \ldots, t'_n),\, \chi$}
+\\[-.8\baselineskip]
       \end{itemize}
@@
       \item[] \return \bigtuple{$\forall a : \fargtype{\con{f}}.\; u' [ \vecfarg{\con{f}}( a ) / \vec x ],\, \ltrue$}
+\\[-.8\baselineskip]
     \end{itemize}
@@
       \item[] \return \bigtuple{$\forall \vec x.\; u',\, \forall \vec x.\; \chi$}
+\\[-.8\baselineskip]
     \end{itemize}
@@
 $\conv( \varphi )$ $=$
+\\[-.8\baselineskip]
  \begin{itemize}
@@
 
-If $t$ has positive polarity, $\conv_0$ adds the constraint $\chi$ derived from
-the subterms as a conjunction. Dually, if $t$ has negative polarity, the
+If $t$ is positive, $\conv_0$ adds the constraint $\chi$ derived from
+the subterms as a conjunction. Dually, if $t$ is negative, the
 function constructs a disjunction with the negated constraint. It $t$ has no

### 2015-05-06 tex/main.tex
minor
More work on proof, almost ready to finish a draft of it.

@@
 function constructs a disjunction with the negated constraint. It $t$ has no
-known polarity, the constrained $\chi$ is simply returned to the caller.
+polarity, the constrained $\chi$ is simply returned to the caller.
 
@@
 %user, and hence already has an intended interpretation.
+
+\begin{lemma}\label{lem:conv}
+If $\varphi$ is a formula not containing function definitions,
+and $\I$ is a $\extendsig{\Sigma}$-interpretation satisfying $\conv( \varphi )$,
+then $\I$ satisfies a set of literals $L \cup X$ such that
+$L$ is a set of $\Sigma$-literals that entail $\varphi$,
+and $X$ is the set of literals
+$\{ \exists a : {\fargtype{\con{f}}}. ( \vecfarg{\con{f}}( a ) \teq \vec t ) \mid \con{f}( \vec t ) \in \terms( L ), \con{f} \in \sfundefs{\Sigma} \}$.
+\end{lemma}
 
@@
 $\DDD \wedge \varphi_0$. 
-%We first argue that $\I$ satisfies a $\Sigma$-formula
-%$\psi$ that is closed under function expansion with respect to $\DDD$, and
-%where $\psi$ entails $\varphi_0$.
-To do so, 
-we first construct a sequence of $\extendsig{\Sigma}$-literals $M_0, M_1, \ldots$ such that $\I$ satisfies each $M_i$,
-and a sequence of mappings $C^{\con f}_0, C^{\con f}_1, \ldots$ from values $\fargtype{f}^\I$ 
-to tuples of terms of type $\tau_1 \times \cdots \times \tau_n$ 
-for each function $\con{f}$ of type $\tau_1 \times \cdots \times \tau_n \rightarrow \tau$ in $\sfundefs{\Sigma}$.
-Let $M_0$ be a set of literals over the atoms occurring in $\varphi_0$
+We argue that $\I$ satisfies a $\Sigma$-formula
+$\psi$ that is closed under function expansion with respect to $\DDD$, and
+where $\psi$ entails $\varphi_0$.
+To construct $\psi$, we define a sequence of sets of $\extendsig{\Sigma}$-literals $M_0 \subseteq M_1 \subseteq \ldots$ 
+such that each $M_i$ is 
+satisfied by $\I$ and
+can be partitioned into two sets $L_i$ and $X_i$,
+where $L_i$ are $\Sigma$-literals,
+and $X_i$ consists of literals of the form $\exists a : {\fargtype{\con{f}}}. ( \vecfarg{\con{f}}( a ) \teq \vec t )$
+for each term $\con{f}( \vec t ) \in \terms( L_i )$ where $\con{f} \in \sfundefs{\Sigma}$.
+Let $M_0$ be 
+a set of literals over atoms occurring in $\conv{ \varphi_0 }$
 such that $\I$ satisfies each literal in $M_0$ and $\bigwedge M_0$ entails $\conv( \varphi_0 )$.
-For each $\con{f} \in \sfundefs{\Sigma}$, let $C^{\con{f}}_0$ be the empty mapping.
+By case analysis on the return values of $\conv_0$ 
+it can be shown that $M_0$ can be partitioned into $L_0$ and $X_0$ in the way described above.
 For each $i \geq 0$,
-let $X_i$ be the set of constraints of the form $\exists a : {\fargtype{\con{f}}}. ( \vecfarg{\con{f}}( a ) \teq \vec t )$ occurring in $M_i$
-and let $L_i$ be the others in $M_i$.
-For each $\con{f} \in \sfundefs{\Sigma}$, let $C^{\con{f}}_{i+1}$ be a mapping containing $C^{\con{f}}_i$ and a minimal number of entries
-such that for all $\exists a : {\fargtype{\con{f}}}. ( \vecfarg{\con{f}}( a ) \teq \vec t ) \in X_i$,
-there exists a $u \mapsto \vec s \in C^{\con{f}}_{i+1}$ such that $\vecfarg{\con{f}}^\I( u ) = \vec s^\I = \vec t^\I$.
-Let $M_{i+1}$ be the set containing $M_i$ as well as
-
-Let $M^\ast$ be the limit of this sequence, which we partition into sets $X^\ast$ and $L^\ast$ in the same way as above.
-Let $\psi$ be $\bigwedge L^\ast$.
-We know $\psi$ is a $\Sigma$-formula since \ajr{blah}
-We know $\psi$ is closed under function expansion with respect to $\DDD$ since \ajr{blah}
-Finally, we know $\psi$ entails $\varphi_0$ since \ajr{blah}
+let $\psi'$ be $\bigwedge \{ \conv( \varphi_f[ \vec t / \vec x ] ) \mid \con{f}( \vec t ) \in \terms( L_i ), f \in \sfundefs{\Sigma} \}$,
+where $\varphi_f$ denotes the function definition of $f$ in $\DDD$.
+Note that $\I$ also satisfies $\psi'$ since $\I$ satisfies $X_i$.
+Let $M_{i+1}$ be the union of $M_0$ and
+a set of literals $M$ over atoms occurring in $\psi'$
+such that $\I$ satisfies each literal in $M$ and $\bigwedge M$ entails $\psi'$,
+where we again may partition $M_{i+1}$ into $L_{i+1}$ and $X_{i+1}$.
+Let $M^\ast$ be the limit of this sequence, 
+which we may partition into (possibly infinite) sets $X^\ast$ and $L^\ast$ in the same way as above.
+Let $\psi$ be the $\Sigma$-formula $\bigwedge L^\ast$.
+We know $\psi$ is closed under function expansion with respect to $\DDD$ due to its construction.
 Since there exists a model for $\psi$ that is closed under function expansion
@@
 Since $\varphi_\con{s}$ is an admissible definition,
-there must exist a $\Sigma$-interpretation satisfying $\psi \wedge \varphi_\con{s}$,
-which means there exists a $\Sigma$-interpretation satisfying formula~(\ref{eq:ex-before}).\xend
+there must exist a $\Sigma$-interpretation satisfying $\psi \wedge \varphi_\con{s}$
+which entails formula~(\ref{eq:ex-before}).\xend
 \end{example}

### 2015-05-06 tex/main.tex
Minor changes throughout.

@@
 \begin{abstract}
-SMT solvers have recently been extended with finite model finding techniques.
+SMT solvers have recently been extended with techniques for finding models
+in presence of quantified formulas.
 However, these generally cannot cope with universal quantifiers ranging over
@@
 and there is no model satisfying $\forallf{\con{f}} x.\;\, \con{f}( x ) \teq \con{f}( x ) + 1$.
-Consider the set $\{ \forallf{\con{f}} x.\;\, \con{f}( x ) \teq \con{f}( x ), \forallf{\con{g}}.\;\, \lite\bigl( \neg \con{f}( x ) \teq 5, \con{g}( x ) \teq \con{g}( x ) + 1, \con{g}( x ) \teq \con{f}( x ) \bigr) \}$.
+Consider the set $\{ \forallf{\con{f}} x.\;\, \con{f}( x ) \teq \con{f}( x ), \forallf{\con{g}} x.\;\, \lite\bigl( \neg \con{f}( x ) \teq 5, \con{g}( x ) \teq \con{g}( x ) + 1, \con{g}( x ) \teq \con{f}( x ) \bigr) \}$.
 While this set has a model where $f$ and $g$ are interpreted as the constant $5$, it is not admissible
 since the formula $\con{f}( 0 ) \teq 4$ is closed under function expansion with respect to this set,
-but there exists no interpretation satisfying both $\con{f}( 0 ) \teq 4$ and $\forallf{\con{g}}.\;\, \lite( \neg \con{f}( x ) \teq 5, \con{g}( x ) \teq \con{g}( x ) + 1, \con{g}( x ) \teq \con{f}( x ) )$.
+but there exists no interpretation satisfying both $\con{f}( 0 ) \teq 4$ and $\forallf{\con{g}} x.\;\, \lite( \neg \con{f}( x ) \teq 5, \con{g}( x ) \teq \con{g}( x ) + 1, \con{g}( x ) \teq \con{f}( x ) )$.
 
@@
 $\exists b.\;\, \farg{\con{s}}( b ) \teq \farg{\con{s}}( a )-1$ ensures that
-a similar constraint is enforced recursively: any model satisfying $\varphi_{\con{s}}[t/x]$
-must also satisfy $\varphi_{\con{s}}[t-1/x]$ when $t > 0$ \ajr{revise this sentence, not quite accurate}. Given a model $\I$ of
+a similar constraint is enforced recursively: when $\varphi_{\con{s}}[t/x]$ is used in part to evaluate $s( c )$,
+any model of $\varphi_{\con{s}}[t/x]$ must also satisfy $\varphi_{\con{s}}[t-1/x]$ when $t > 0$. Given a model $\I$ of
 formula~(\ref{eq:ex-after}) that interprets $\con{c}$ as $14$, it is easy to see that
@@
 we will refer to as \cvcd, \cvcf, and \cvci.
-The default configuration \cvcd runs heuristic quantifier instantiation,
+The default configuration \cvcd runs heuristic and conflict-based techniques for quantifier instantiation~\cite{ReynoldsTinelliMoura14},
 but does not include techniques for finding models.
@@
 reasoning in SMT~\cite{reynolds-kuncak-2015}.
-%All configurations of \cvc incorporate techniques for conflict-based instantiation~\cite{} ..?
 
@@
 Accordingly, the solver will output the approximation of the interpretation it
-used for recursive function definitions. For our example, it will output the following
-model:
-\ajr{prefer having the full output like this here?}
-\jb{I like it the way you did it now, but I took the freedom to reformat the output slightly
-(to avoid the small font---I generally try to avoid ad hoc font mixtures---and to show
-the nesting more clearly).}
+used for recursive function definitions. For our example, it will produce the following output:
+%\ajr{prefer having the full output like this here?}
+%\jb{I like it the way you did it now, but I took the freedom to reformat the output slightly
+%(to avoid the small font---I generally try to avoid ad hoc font mixtures---and to show
+%the nesting more clearly).}
 %it may output the interpretation~(\ref{eq:approx-interp}) from Section~\ref{sec:encoding} for $s$.
@@
 %
-to the above problem, \cvc will still find the same ``model''
+to the above problem, \cvc will still produce the same output
 if \texttt{--fmf-fun} is specified.
@@
 \cite{GeDeM-CAV-09} and by Reynolds et al.\
-\cite{ReyEtAl-1-RR-13,reynolds-et-al-2013}, in the introduction already.
+\cite{ReyEtAl-1-RR-13,reynolds-et-al-2013}, in the text already.
 Outside the world of SMT, model finding has been

### 2015-05-06 tex/main.tex
More minor fixes and some questions

@@
 This fragment is limited to universally quantified formulas where all
-occurrences of variables occur as direct subterms of uninterpreted
+%ct occurrences of 
+variables occur as direct subterms of uninterpreted
 functions---e.g., $\forall x%\typ{\typeint} --- the type is not important, right?
@@
 find a model for a formula such as $\forall x,\, y : \tau.\;\, x \teq
-y \vee \allowbreak \neg \con{f}( x ) \teq \con{f}( y )$, where $\tau$ is uninterpreted.
+y \vee \allowbreak \neg \con{f}( x ) \teq \con{f}( y )$, where $\tau$ is an uninterpreted type.
 % where the uninterpreted type $\tau$ is
@@
 applicable for functions over infinite domains such as the integers or
-recursive datatypes. A simple example where both approaches fail is
+algebraic datatypes. A simple example where both approaches fail is
 $$\forall n : {\typeint}.\;\, \lite\bigl(
@@
 
-In this paper, we present a method for translating formulas involving function
+In this paper, we present a method for translating formulas involving recursice function
 definitions into formulas where finite model finding techniques can be applied.
@@
 function definitions, but also by tail-recursive functions and even by
-productive corecursive functions. It is, however, not met by inconsistent
+\emph{productive} co-recursive functions. It is, however, not met by inconsistent
 definitions such as $\forall n : {\typeint}.\;\, \con{f}(n) \teq \con{f}(n) +
-1$, and it is possible to construct consistent examples that do not meet it.
+1$, %ct and it is possible to construct consistent examples that do not meet it.
+as well as other example of consistent definition.
 
@@
 \fargtype{\tau}.\; \const{f}(\farg{\con{f}}(a)) \teq t[\concret(a)/x]$, where
-$\fargtype{\tau}$ is an uninterpreted ``abstract'' type and $\farg{\con{f}} :
+$\fargtype{\tau}$ is an uninterpreted \emph{abstract} type and $\farg{\con{f}} :
 \fargtype{\tau} \to \tau$ a coercion from the abstract type to the concrete
 type. The translation preserves satisfiability and, for admissible definitions,
-unsatisfiability, and it makes finite model finding possible in practice.
-
-Our empirical evaluation on benchmarks from IsaPlanner and Leon provides
+unsatisfiability, and makes finite model finding possible in practice
+for problems in this class.
+
+Our empirical evaluation on benchmarks from the IsaPlanner proof planner~\cite{} and Leon verifier~\cite{} provides
 evidence that this translation improves the effectiveness of the SMT solvers
@@
 
-Our setting is a monomorphic (or many-sorted) first-order logic, such as
-the one defined by SMT-LIB \cite{smtlib25}.
+Our setting is a (monomorphic) many-sorted first-order logic
+like the one defined by SMT-LIB \cite{smtlib25}.
 A \emph{signature} $\Sigma$ consists of
-a set $\stypes{\Sigma}$ of type symbols (or sorts) and
-a set $\sfuns{\Sigma}$ of function symbols.
+a set $\stypes{\Sigma}$ of %ct type symbols
+fist-order types 
+%
+(or sorts) and a set $\sfuns{\Sigma}$ of function symbols over these types.
 We assume that signatures always include a Boolean type $\Bool$ and constants
@@
 standard Boolean connectives ($\neg$, $\wedge$, $\vee$, etc.),
-and a function symbol $\lite : \Bool \times \tau \times \tau \rightarrow \tau$ (``if--then--else'')
+and an if-then-else function symbol $\lite : \Bool \times \tau \times \tau \rightarrow \tau$
 for each $\tau \in \stypes{\Sigma}$.
+Formulas are terms of type $\Bool$.
 We write $t\typ{\tau}$ to denote a metavariable ranging over terms of
 type~$\tau$ and $\terms( t )$ to denote the set of subterms in $t$.
-Formulas are terms of type $\Bool$.
 Given a term $u$, we write $u[\vec t/\vec x]$ to denote the result of replacing
@@
 otherwise, it is \emph{$T$-unsatisfiable}.
-A formula $\varphi$ \emph{entails} $\psi$, written $\varphi \models_T \psi$,
+A formula $\varphi$ \emph{$T$-entails} $\psi$, written $\varphi \models_T \psi$,
 if all interpretations in $\Mo$ that satisfy $\varphi$ also satisfy $\psi$.
+Two formulas $\varphi$ and $\psi$ are \emph{$T$-equivalent} 
+if each $T$-entails the other.
 Given a signature $\Sigma$,
@@
 We refer to the type and function symbols in this theory as \emph{uninterpreted}.
-
-In this paper, we consider \emph{annotated quantified formulas} of the form
+\ct{This is both ambiguous and insufficient.
+Since this is SMT we should fix some background theory $T$,
+then extend its signature with additional symbols and call \emph{those} uninterpreted.
+}
+
+In addition to standard quantified formulas $\forall \vec x.\; \varphi$,
+we consider \emph{annotated quantified formulas} of the form
 $\forallf{\const{f}} \vec x.\; \varphi$, where $\con{f} \in \sfuns{\Sigma}$ is
-an uninterpreted function symbol, in addition to standard quantified formulas
-$\forall \vec x.\; \varphi$.
+an uninterpreted function symbol, 
+whose semantics is the same as standard quantified formulas.
 Given $\con{f} : \tau_1 \times \cdots \times \tau_n \rightarrow \tau$,
@@
 quantified formula $\forallf{\con{f}} \vec x.\; \varphi$ is a \emph{function definition}
-(\,\emph{for $\con{f}$}\vthinspace) if $\vec x$ is a tuple of variables of type
-$\tau_1, \ldots, \tau_n$ and $\varphi$ is a (quantifier-free) formula equivalent to
-$\con{f}( \vec x ) \teq t$ for some $t$. We write $\exists
-\vec x.\; \varphi$ as shorthand for $\neg \forall \vec x.\; \neg \varphi$.
+(\,\emph{for $\con{f}$}\vthinspace) if $\vec x$ is a tuple of variables
+$x_1, \ldots, x_n$ of respective type
+$\tau_1, \ldots, \tau_n$ and $\varphi$ is a quantifier-free formula 
+$T$-equivalent to $\con{f}( \vec x ) \teq t$ for some term $t$ of type $\tau$.
+\ct{Why can't $\varphi$ be existential instead of just quantifier free?} 
+We write $\exists
+\vec x.\; \varphi$ as a shorthand for $\neg \forall \vec x.\; \neg \varphi$.
 
@@
 %
-where $\con{f}_1, \ldots, \con{f}_n \in \sfuns{\Sigma}$ are distinct
+where $\con{f}_1, \ldots, \con{f}_n \in \sfuns{\Sigma}$ are distinct function symbols,
+%ct added
+$\forallf{\con{f}_i} \vec x_i.\; \varphi_i$ is a function definition
+for $i = 1, \ldots, n$,
+%
+\ct{I think we need more conditions here on how a symbol $\con{f}_i$ can occur in $\varphi_j$.}
 and $\psi$ contains no function definitions.

### 2015-05-07 tex/main.tex
Several minor edits and several minor and not so minor comments and questions.

@@
 However, these generally cannot cope with universal quantifiers ranging over
-infinite types, such as integers and recursive datatypes. This paper introduces
+infinite types, such as integers and algebraic datatypes. This paper introduces
 a translation that eliminates infinite quantifiers in axioms specifying a large
@@
 \label{sec:introduction}
+
+\ct{The paper used "recursive datatype" in some places and "inductive datatype" in others. I have replaced both by "algebraic datatype" which is more accurate for \cvc.
+("recursive datatype" is a more general notion and "inductive datatype" has a stronger meaning in higher-order logics.)}
 
@@
 a set $\stypes{\Sigma}$ of %ct type symbols
-fist-order types 
-%
+first-order types 
+%
+\ct{What the rationale for using the letter $y$ in $\stypes{\Sigma}$?}
 (or sorts) and a set $\sfuns{\Sigma}$ of function symbols over these types.
@@
 \begin{definition}\rm
-Given a set of function definitions $\DDD = \{ \forallf{\con{f}_1} \vec x.\;
-\varphi_1, \ldots, \forallf{\con{f}_n} \vec x.\; \varphi_n \}$, a formula $\varphi$
-is \emph{closed under function expansion with respect to $\DDD$} if
-\[\varphi \,\models_T\, \bigwedge\nolimits_{\,i=1}^{n} \{ \varphi_i[ \vec t / \vec x ] \mid \con{f}_i( \vec t ) \in \terms( \varphi ) \}\]
-The set $\DDD$ is \emph{admissible} if whenever there exists a model for $\varphi$ that is closed under function expansion,
-there exists a model for $\varphi \wedge \forallf{\con{f}_1} \vec x.\; \varphi_1 \wedge \cdots \wedge \forallf{\con{f}_n} \vec x.\; \varphi_n$.
+Given a set of function definitions 
+$\DDD = \{ \forallf{\con{f}_1} \vec x.\; \varphi_1, \ldots, \forallf{\con{f}_n} \vec x.\; \varphi_n \}$, 
+a formula $\varphi$ 
+\ct{I suppose $\varphi$ is meant to be ground.}
+is \emph{closed under function expansion 
+with respect to $\DDD$} if 
+\[
+ \varphi \,\models_T\, \bigwedge\nolimits_{\,i=1}^{n} \{ \varphi_i[ \vec t / \vec x ] \mid \con{f}_i( \vec t ) \in \terms( \varphi ) \} .
+\]
+The set $\DDD$ is \emph{admissible} if for every $T$-satisfiable formula 
+$\varphi$ closed under function expansion wrt $\DDD$
+the formula
+$\varphi \wedge \forallf{\con{f}_1} \vec x.\; \varphi_1 \wedge \cdots \wedge \forallf{\con{f}_n} \vec x.\; \varphi_n$ is also $T$-satisfiable.
+\ct{I}
 \end{definition}
 
-For example, the set $\{ \forallf{\con{f}} \vec x.\;\, \con{f}( x ) \teq 0 \}$ is admissible.
-An admissible set can include definitions of nonterminating functions, e.g.
+For example, if $T$ is the theory of integer arithmetic,
+the set $\{ \forallf{\con{f}} \vec x:\typeint.\;\, \con{f}( x ) \teq 0 \}$ is admissible.
+An admissible set can include definitions of nonterminating functions,
+\ct{There is no notion of termination in FOL. We either define this properly or we just replace "definitions of nonterminating functions, e.g.," with "definitions like"}
+e.g.,
 $\{ \forallf{\con{f}} x : \typeint.\;\, \con{f}( x ) \teq \con{f}( x - 1 ) \}$ or even $\{ \forallf{\con{f}} x.\allowbreak\;\, \con{f}( x ) \teq \con{f}( x ) \}$ are admissible.
 On the other hand, the set $\{ \forallf{\con{f}} x.\;\, \con{f}( x ) \teq \con{f}( x ) + 1 \}$ is not admissible,
-since the empty set of formulas is closed under function expansion with respect to this set,
-and there is no model satisfying $\forallf{\con{f}} x.\;\, \con{f}( x ) \teq \con{f}( x ) + 1$.
-Consider the set $\{ \forallf{\con{f}} x.\;\, \con{f}( x ) \teq \con{f}( x ), \forallf{\con{g}} x.\;\, \lite\bigl( \neg \con{f}( x ) \teq 5, \con{g}( x ) \teq \con{g}( x ) + 1, \con{g}( x ) \teq \con{f}( x ) \bigr) \}$.
-While this set has a model where $f$ and $g$ are interpreted as the constant $5$, it is not admissible
+since the formula $\ltrue$ is (trivially) closed under function expansion with respect to this set,
+and there is no model of $T$ satisfying $\forallf{\con{f}} x.\;\, \con{f}( x ) \teq \con{f}( x ) + 1$.
+Consider the set 
+$\{ 
+ \forallf{\con{f}} x.\;\, \con{f}( x ) \teq \con{f}( x ),\;
+ \forallf{\con{g}} x.\;\, \lite\bigl( \con{f}( x ) \teq 5,\;
+   \con{g}( x ) \teq \con{f}( x ),\,
+   \con{g}( x ) \teq \con{g}( x ) + 1)
+\}$.
+While this set has a model where $\con f$ and $\con g$ are interpreted as the constant $5$, it is not admissible
 since the formula $\con{f}( 0 ) \teq 4$ is closed under function expansion with respect to this set,
-but there exists no interpretation satisfying both $\con{f}( 0 ) \teq 4$ and $\forallf{\con{g}} x.\;\, \lite( \neg \con{f}( x ) \teq 5, \con{g}( x ) \teq \con{g}( x ) + 1, \con{g}( x ) \teq \con{f}( x ) )$.
+but there exists no interpretation satisfying both $\con{f}( 0 ) \teq 4$ and $\forallf{\con{g}} x.\;\, \lite( \con{f}( x ) \teq 5,\, \con{g}( x ) \teq \con{f}( x ),\, \con{g}( x ) \teq \con{g}( x ) + 1 )$.
 
@@
 
-First, we fix a signature $\Sigma$ and a $\Sigma$-formula
-$\varphi$ in definitional form with respect to $\sfundefs{\Sigma}$,
-whose definitions are admissible.
-%We may translate $\varphi$ into a equisatisfiable formula $\psi$ for which known model-finding procedures~\cite{GeDeM-CAV-09, ReyEtAl-1-RR-13} are applicable.
-This section presents a method for constructing an extended signature
-$\extendsig{ \Sigma }$ and an $\extendsig{ \Sigma }$-formula $\psi$ such that
-$\psi$ is satisfiable if and only if $\varphi$ is satisfiable---i.e.,
-$\varphi$ and $\psi$ are \relax{equisatisfiable}.
-The intuition behind this translation
-is to use an uninterpreted type to abstract the set of
-\relax{relevant} tuples for each recursive function $\con{f}$ and restrict the
-quantification of the function definition for $\con{f}$ to a variable of this
-type. Informally, the relevant tuples $\vec t$ of a function $\con{f}$ are the
-ones for which the interpretation of $\con{f}( \vec t )$ is relevant to the
-satisfiability of $\varphi$.
-
-More precisely,
-suppose our signature $\Sigma$ contains a set of defined function symbols $\sfundefs{\Sigma} \subseteq \sfuns{\Sigma}$.
-We extend the signature of $\Sigma$ to a signature $\extendsig{\Sigma}$ containing the following additional definitions.
-For each $\con{f} : \tau_1 \times \cdots \times \tau_n \rightarrow \tau \in \sfundefs{\Sigma}$,
-the signature $\extendsig{\Sigma}$ includes
-\begin{itemize}
-\item an uninterpreted \emph{abstract type} $\fargtype{\con{f}}$;
-\item $n$ uninterpreted \emph{concretization functions} $\fargx{\con{f}}{1} : \fargtype{\con{f}} \rightarrow \tau_1$, \ldots, $\fargx{\con{f}}{n} : \fargtype{\con{f}} \rightarrow \tau_n$.
-\end{itemize}
-%The interpretation of uninterpreted type $\fargtype{\con{f}}$ will denote the elements (tuples) on which the function $\con{f}$ is applied.
-%The role of the uninterpreted functions $\farg{\con{f}}^1$, $\ldots$, $\farg{\con{f}}^n$ will be discussed more in the following.
-
-The function $\conv$ translates the $\Sigma$-formula
-$\varphi$ into an $\extendsig{\Sigma}$-formula. It relies
-on the auxiliary function $\conv_0$, which takes two arguments:\ the term $t$
-to translate and a \relax{polarity}~$p$, which is either $\ppos$, $\pneg$, or
-$\pnone$. \vthinspace$\conv_0$ returns a pair $( t'\negvthinspace,\, \chi )$, where $t'$ is a term of
-type $\tau$ and $\chi$ is an additional constraint.
-This constraint is needed if the polarity is $\pnone$; otherwise, it is encoded
-directly in $t'$ and $\chi$ is simply $\ltrue$.
-In the auxiliary function $\conv_0$, the function $\pol(
-\boolop, i, p )$ returns the polarity of the $i$\vvthinspace{th} argument of an
-application of $\boolop$ having polarity $p$, where $\boolop$ is an interpreted
-predicate.
-
+\begin{figure}[t]
 \begin{enumerate}
@@
 \end{enumerate}
+\caption{Definition of function $\conv$.}
+\label{fig:A}
+\end{figure}
+
+First, we fix a signature $\Sigma$ and a $\Sigma$-formula
+$\varphi$ in definitional form with respect to $\sfundefs{\Sigma}$,
+whose definitions are admissible.
+%We may translate $\varphi$ into a equisatisfiable formula $\psi$ for which known model-finding procedures~\cite{GeDeM-CAV-09, ReyEtAl-1-RR-13} are applicable.
+This section presents a method for constructing an extended signature
+$\extendsig{ \Sigma }$ and an $\extendsig{ \Sigma }$-formula $\psi$ such that
+$\psi$ is satisfiable if and only if $\varphi$ is satisfiable---i.e.,
+$\varphi$ and $\psi$ are \relax{equisatisfiable}.
+The intuition behind this translation
+is to use an uninterpreted type to abstract the set of
+\relax{relevant} tuples for each recursive function $\con{f}$ and restrict the
+quantification of the function definition for $\con{f}$ to a variable of this
+type. Informally, the relevant tuples $\vec t$ of a function $\con{f}$ are the
+ones for which the interpretation of $\con{f}( \vec t )$ is relevant to the
+satisfiability of $\varphi$.
+
+More precisely,
+suppose our signature $\Sigma$ contains a set of defined function symbols $\sfundefs{\Sigma} \subseteq \sfuns{\Sigma}$.
+We extend the signature of $\Sigma$ to a signature $\extendsig{\Sigma}$ containing the following additional definitions.
+For each $\con{f} : \tau_1 \times \cdots \times \tau_n \rightarrow \tau \in \sfundefs{\Sigma}$,
+the signature $\extendsig{\Sigma}$ includes
+\begin{itemize}
+\item an uninterpreted \emph{abstract type} $\fargtype{\con{f}}$;
+\item $n$ uninterpreted \emph{concretization functions} $\fargx{\con{f}}{1} : \fargtype{\con{f}} \rightarrow \tau_1$, \ldots, $\fargx{\con{f}}{n} : \fargtype{\con{f}} \rightarrow \tau_n$.
+\end{itemize}
+%The interpretation of uninterpreted type $\fargtype{\con{f}}$ will denote the elements (tuples) on which the function $\con{f}$ is applied.
+%The role of the uninterpreted functions $\farg{\con{f}}^1$, $\ldots$, $\farg{\con{f}}^n$ will be discussed more in the following.
+
+The function $\conv$ defined in Figure~\ref{fig:A} translates the $\Sigma$-formula
+$\varphi$ into an $\extendsig{\Sigma}$-formula. It relies
+on the auxiliary function $\conv_0$, which takes two arguments:\ the term $t$
+to translate and a \relax{polarity}~$p$, which is either $\ppos$, $\pneg$, or
+$\pnone$. \vthinspace$\conv_0$ returns a pair $( t'\negvthinspace,\, \chi )$, where $t'$ is a term of
+type $\tau$ and $\chi$ is an additional constraint.
+This constraint is needed if the polarity is $\pnone$; otherwise, it is encoded
+directly in $t'$ and $\chi$ is simply $\ltrue$.
+In the auxiliary function $\conv_0$, the function $\pol(
+\boolop, i, p )$ returns the polarity of the $i$\vvthinspace{th} argument of an
+application of $\boolop$ having polarity $p$, where $\boolop$ is an interpreted
+predicate.
 
@@
 uninterpreted type $\fargtype{\con{f}}$ and inserts appropriate coercions
-$\fargx{\con{f}}{i}$. Second, it augments $\varphi$ with additional exisential
+$\fargx{\con{f}}{i}$. Second, it augments $\varphi$ with additional existential
 constraints
@@
 
-If $t$ is an application of an interpreted predicate $\boolop
+If $t$ is an application of a logical predicate $\boolop
 \in \{{\neg}, {\wedge}, {\vee}, {\teq}, {\lite}\}$,
@@
 functions $\Sigma = \Sigmalia \mathrel{\cup} \Sigma_{\euf}$, where
-$\stypes{\Sigma_{\euf}} = \{ \typeint \}$ and
+$\stypes{\Sigma_{\euf}} = \{ \typeint \}$ 
+\ct{shouldn't this be $\stypes{\Sigmalia} = \{ \typeint \}$?}
+and
 $\sfuns{\Sigma_{\euf}} = \{ \con{s} : { \typeint \rightarrow \typeint },\, \const{c} : { \typeint } \}$.
@@
 type $\fargtype{\con{s}}$, thus enabling the finite model finding
-techniques by Reynolds et al.\ \cite{ReyEtAl-1-RR-13,reynolds-et-al-2013} to search for a finite interpretation for
-$\fargtype{\con{s}}$. Moreover, all occurrences of $a$ in this formula are beneath
-applications of the uninterpreted function $\farg{\con{s}}$, implying that the
-formula is in the essentially uninterpreted fragment,
-for which Ge and de Moura \cite{GeDeM-CAV-09} provide a complete instantiation
-procedure. Both
+techniques by Reynolds et al.\ \cite{ReyEtAl-1-RR-13,reynolds-et-al-2013},
+implemented in \cvc, to search for a finite interpretation for $\fargtype{\con{s}}$. 
+Since all occurrences of the quantified variable $a$ in this formula are 
+beneath applications of the uninterpreted function $\farg{\con{s}}$, 
+the formula is in the essentially uninterpreted fragment,
+for which Ge and de Moura \cite{GeDeM-CAV-09} provide 
+a complete instantiation procedure, implemented in \ziii. 
+Both
 %%% Added "apparently". Or are we sure?  
 % Yes, I'm planning to post the 2 versions of the benchmark on the evaluation page (when we add it) --ajr
-\cvc and \ziii run indefinitely on formula~(\ref{eq:ex-before}), and both
-produce a model for formula~(\ref{eq:ex-after}) within 100 milliseconds.\xend
+\cvc and \ziii run indefinitely on formula~(\ref{eq:ex-before}), 
+while they both produce a model for formula~(\ref{eq:ex-after}) 
+within 100 milliseconds.\xend
 \end{example}
@@
 The translation $\conv$ does not preserve the models of $\varphi$.
+\ct{"Preserves models" means that it does not drop models, not that it adds models.}
 One model $\I$ for formula~(\ref{eq:ex-after}) in the previous example interprets
@@
 $\sfundefs{\Sigma}$ are admissible, %we claim that
-$\conv(\varphi)$ is
-equisatisfiable with $\varphi$ for any input $\varphi$. The intuition is that the
+$\conv(\varphi)$ is equisatisfiable with $\varphi$ for any input $\varphi$.
+\ct{Must define "equisatisfiable".
+What we mean here is a weak form of satisfiability, we could instead say that
+$\conv(\varphi)$ is $T$-satisfiable iff $\varphi$ is $T$-satisfiable.
+} 
+The intuition is that the
 interpretation of a term such as $\con{s}( -1 )$ in the previous example is not
@@
 $\{ \forallf{\con{s}} x\typ{\typeint}.\;\, \lite\bigl( x \leq 0,\allowbreak\; \con{s}(x)
-\teq 0,\; \con{s}( x ) \teq x + \con{s}( x-1 ) \bigr) \}$ is admissble.
+\teq 0,\; \con{s}( x ) \teq x + \con{s}( x-1 ) \bigr) \}$ is admissible.
+\ct{
+You cannot possibly mean "if and only if" in the two sentences above.
+Once I show that (1) is unsat iff (2) is unsat, I do not need to argue that
+(1) is sat iff (2) is sat. That is immediate.
+}
 As stated
@@
 variety of inductive theorem provers. These benchmarks heavily involve
-recursive functions, and are limited to the combined theory of uninterpreted
-functions and inductive datatypes. The second source consists of 166
+recursively-defined functions, and are limited to the combined theory of uninterpreted
+functions and algebraic datatypes. The second source consists of 166
 benchmarks from the Leon repository,\footnote{%Available at
@@
 verification conditions from simple Scala programs. These benchmarks also
-heavily involve recursive functions over inductive datatypes, and also cover a
-wide variety of theories, including bit vectors, arrays, and linear and
+heavily involve recursively-defined functions over algebraic datatypes, 
+but cover a wide variety of theories, including bit vectors, arrays, and linear and
 nonlinear arithmetic. Each benchmark is in definitional form with respect to a
@@
 A majority of these benchmarks are unsatisfiable, although a
-handful of the benchmarks in the \leon set are satisfiable.
+handful of benchmarks in the \leon set are satisfiable.
 
@@
 Note that benchmarks created in this way have a high likelihood of having small
-and easy-to-find counterexamples. In total, we considered 213 mutated forms of
+and easy-to-find counterexamples \ct{contermodels?}. In total, we considered 213 mutated forms of
 conjectures from \isa, and 427 mutated forms of conjectures from \leon. We will
@@
 \end{tabular}
-\caption{\,Number of ``satisfiable'' responses on benchmarks without and with translation with $\conv$}
+\caption{\,Number of \emph{sat} responses on benchmarks without and with translation with $\conv$}
 \label{fig:sat}
@@
 \end{tabular}
-\caption{\,Number of ``unsatisfiable'' responses on benchmarks without and with translation with $\conv$}
+\caption{\,Number of \emph{unsat} responses on benchmarks without and with translation with $\conv$}
 \label{fig:unsat}
@@
 The results are shown in Figures \ref{fig:sat} and \ref{fig:unsat}.
-The results are divided into benchmarks triggering ``unsatisfiable'' and ``satisfiable''
+They are divided into benchmarks triggering \emph{unsat} and \emph{sat}
 responses and further into benchmarks before and after the translation $\conv$.
 The raw evaluation data reveals no cases in which a solver answered
-``unsatisfiable'' on a benchmark $\varphi$ and ``satisfiable'' on its
-corresponding benchmark $\conv( \varphi )$, or vice versa, thereby
-corroborating Theorem~\ref{thm:equi}.
+\emph{unsat} on a benchmark $\varphi$ and \emph{sat} on its
+corresponding benchmark $\conv( \varphi )$, or vice versa, 
+thereby corroborating Theorem~\ref{thm:equi}.
+\ct{?! \emph{Theorems} do not need corroboration! 
+Maybe we should say "consistently with Theorem~\ref{thm:equi}"?}
 
 Figure~\ref{fig:sat} shows that for untranslated benchmarks (the ``$\varphi$''
-columns), the number of satisfiable responses is very low, across all
+columns), the number of \emph{sat} responses is very low, across all
 configurations. This confirms the shortcomings of existing SMT techniques for
-finding counterexamples for benchmarks containing recursive functions.
+finding counterexamples for benchmarks containing recursively-defined functions.
 %
@@
 We conclude that the translation $\conv$ enables SMT
-solvers to find counterexamples for conjectures involving recursive functions
+solvers to find counterexamples for conjectures involving recursively-defined functions
 whose definitions are admissible.
@@
 
-In terms of number of unsatisfiable responses, the configuration \cvci is the
+In terms of number of \emph{unsat} responses, the configuration \cvci is the
 clear winner, finding 318 total benchmarks to be unsatisfiable. Moreover, the
-translation $\conv$ helps \ziii for ``unsatisfiable'' responses as well, as we
+translation $\conv$ helps \ziii for \emph{unsat} responses as well, as we
 see \ziii solves a total of 209 whereas it solves only 187 before the
@@
 %
-This number is 2 fewer than \cvcd, which solves 211 unsatisfiable. Thus, in
+This number is 2 fewer than \cvcd, which finds 211 benchmarks unsatisfiable. Thus, in
 most cases the translation $\conv$ does not significantly degrade performance
@@
 (version~1.5 prerelease).
-Function definitions $\forallf{\con{f}} \vec x.\; \varphi$ can be written in using the $\definefunreccmd$ command,
-from the SMT-LIB 2.5 draft standard \cite{smtlib25}.
+Function definitions $\forallf{\con{f}} \vec x.\; \varphi$ can be written using the $\definefunreccmd$ command,
+from the upcoming SMT-LIB 2.5 standard \cite{smtlib25}.
 For example, formula~(\ref{eq:ex-before}) from Example~\ref{ex:translation}
@@
 \begin{quote}
+\small
 \begin{verbatim}
@@
 functions provided by the user are admissible.
-Admissibility must be checked separately---for example, using a
+Admissibility must be checked separately \emph{by the user}---for example, using a
 syntactic criterion or a termination prover.
-If this criterion is not met, the solver may answer ``satisfiable'' on
+If this criterion is not met, \cvc may answer \emph{sat} on
 an unsatisfiable problem. For example, if we add the inconsistent definition
@@
 Outside the world of SMT, model finding has been
-studied many times. The text below is partly adapted from Blanchette's Ph.D.\
+studied many times. 
+The text below is partly adapted from Blanchette's Ph.D.\
 thesis \cite{blanchette-2012-phd}.
+\ct{The last sentence is strange. I would remove it since Jasmin is one of the authors.}
 
@@
 Other SAT-based finders are Paradox \cite{claessen-sorensson-2003}, FMDarwin
-\cite{xxxx}, and iProver \cite{xxxx} for first-order logic with equality; the
+\cite{xxxx}, and iProver \cite{xxxx} for first-order logic with equality;
+\ct{FMDarwin is not SAT-based although it is Mace-like.
+Calling iProver SAT-based is a stretch since it is based 
+on a natively first-order calculus even it uses an auxiliary SAT solver heavily.
+}
 Alloy Analyzer and its back-end Kodkod \cite{torlak-jackson-2007} for
@@
 
-Most of the above tool cannot cope with recursive datatypes or other infinite
+Most of the above tool cannot cope with algebraic datatypes or other infinite
 types. Kuncak and Jackson \cite{kuncak-jackson-2005} presented an idiom for
@@
 that our approach might be better?}
+
+\ct{About Leon, the reviewers and readers may wonder why we do not do the same, that is, just progressively unfold the recursive definitions until a model is found.
+We should probably say something in this regard.
+}
 
@@
 find models of quantified formulas over infinite types, such as integers
-and recursive datatypes.
+and algebraic datatypes.
 
 In future work, it would be interesting to evaluate the approach against other
-model finders, notably Leon and Nitpick, and enrich the benchmark suite with
+model finders, notably Leon \ct{?? Leon is a model finder? Alloy?} and Nitpick, and enrich the benchmark suite with
 problems exercising \cvc's support for codatatypes
@@
 counterexample generator in proof assistants.
+\ct{Further work could also include identifying simple and/or syntactic sufficient conditions for admissibility. }
 

### 2015-05-07 tex/main.tex
Minor fixes

@@
 \newcommand{\M}{\mathcalx{M}}
-\newcommand{\I}{\mathcalx{J}} %%% there's no difference, right? --JB
+\newcommand{\I}{\mathcalx{I}} %%% there's no difference, right? --JB  need to simultaneously refer to 2 different interpretations --ajr.
 %% (I find \mathcalx{I} ugly, with its two horizontal bars, hence the J)
@@
 \newcommand{\farg}[1]{\concret_{#1}}
-\newcommand{\fargx}[2]{\concret^{\vthinspace#2}_{#1}}
+\newcommand{\fargx}[2]{\concret_{#1,\vthinspace#2}}
 \newcommand{\fargtype}[1]{\alpha_{#1}}
@@
 and $\I$ is a $\extendsig{\Sigma}$-interpretation satisfying $\conv( \varphi )$,
-then $\I$ satisfies a set of literals $L \cup X$ such that
-$L$ is a set of $\Sigma$-literals that entail $\varphi$,
-and $X$ is the set of literals
+then $\I$ satisfies a set of $\Sigma$-literals $L$ that entail $\varphi$,
+and $\I$ satisfies
 $\{ \exists a : {\fargtype{\con{f}}}. ( \vecfarg{\con{f}}( a ) \teq \vec t ) \mid \con{f}( \vec t ) \in \terms( L ), \con{f} \in \sfundefs{\Sigma} \}$.
@@
 $\exists b.\;\, \farg{\con{s}}( b ) \teq \farg{\con{s}}( a )-1$ ensures that
-a similar constraint is enforced recursively: when $\varphi_{\con{s}}[t/x]$ is used in part to evaluate $s( c )$,
+a similar constraint is enforced recursively: when $\varphi_{\con{s}}[t/x]$ is used in part to evaluate $\con{s}( \con{c} )$,
 any model of $\varphi_{\con{s}}[t/x]$ must also satisfy $\varphi_{\con{s}}[t-1/x]$ when $t > 0$. Given a model $\I$ of
@@
 %
-to the above problem, \cvc will still produce the same output
+to the above problem, \cvc will still return ``satisfiable"
 if \texttt{--fmf-fun} is specified.

### 2015-05-07 tex/main.tex
Address some of Cesare's comments.

@@
 $T$-equivalent to $\con{f}( \vec x ) \teq t$ for some term $t$ of type $\tau$.
-\ct{Why can't $\varphi$ be existential instead of just quantifier free?} 
+%\ct{Why can't $\varphi$ be existential instead of just quantifier free?} 
+%I don't think it makes much sense to have any quantifiers within function bodies, 
+%since they aren't operational, that's why I put quantifier-free. --ajr
 We write $\exists
@@
 $\DDD = \{ \forallf{\con{f}_1} \vec x.\; \varphi_1, \ldots, \forallf{\con{f}_n} \vec x.\; \varphi_n \}$, 
-a formula $\varphi$ 
-\ct{I suppose $\varphi$ is meant to be ground.}
-is \emph{closed under function expansion 
-with respect to $\DDD$} if 
+a ground formula $\varphi$ 
+%\ct{I suppose $\varphi$ is meant to be ground.} yes --ajr
+is \emph{closed under function expansion with respect to $\DDD$} if 
 \[
@@
 \ct{There is no notion of termination in FOL. We either define this properly or we just replace "definitions of nonterminating functions, e.g.," with "definitions like"}
+\ajr{I agree. Jasmin, can you fix this?  We need to be careful in the same way when we talk about the "terminating" functions fragment we handle.}
 e.g.,
@@
 functions $\Sigma = \Sigmalia \mathrel{\cup} \Sigma_{\euf}$, where
-$\stypes{\Sigma_{\euf}} = \{ \typeint \}$ 
-\ct{shouldn't this be $\stypes{\Sigmalia} = \{ \typeint \}$?}
-and
+%$\stypes{\Sigma_{\euf}} = \{ \typeint \}$ 
+%\ct{shouldn't this be $\stypes{\Sigmalia} = \{ \typeint \}$?}
+%and
+% Something went wrong, let's just remove it --ajr
 $\sfuns{\Sigma_{\euf}} = \{ \con{s} : { \typeint \rightarrow \typeint },\, \const{c} : { \typeint } \}$.
@@
 
-The translation $\conv$ does not preserve the models of $\varphi$.
-\ct{"Preserves models" means that it does not drop models, not that it adds models.}
+The translation $\conv$ results in formulas whose models are different than those of $\varphi$.
+%\ct{"Preserves models" means that it does not drop models, not that it adds models.}
 One model $\I$ for formula~(\ref{eq:ex-after}) in the previous example interprets
@@
 Let us revisit the translated formula~(\ref{eq:ex-after}) from
-Example~\ref{ex:translation}. This formula is unsatisfiable if and only if
+Example~\ref{ex:translation}. This formula is unsatisfiable only if
 formula~(\ref{eq:ex-before}) is unsatisfiable since $\fargtype{f}$ can be
 interpreted as $\typeint$, and $\farg{f}$ can be interpreted as the identity
-function. Conversely, we claim formula~(\ref{eq:ex-after}) is satisfiable if
-and only if formula~(\ref{eq:ex-before}) is satisfiable, noting that the set
+function. Conversely, we claim formula~(\ref{eq:ex-after}) is satisfiable
+only if formula~(\ref{eq:ex-before}) is satisfiable, noting that the set
 $\{ \forallf{\con{s}} x\typ{\typeint}.\;\, \lite\bigl( x \leq 0,\allowbreak\; \con{s}(x)
 \teq 0,\; \con{s}( x ) \teq x + \con{s}( x-1 ) \bigr) \}$ is admissible.
-\ct{
-You cannot possibly mean "if and only if" in the two sentences above.
-Once I show that (1) is unsat iff (2) is unsat, I do not need to argue that
-(1) is sat iff (2) is sat. That is immediate.
-}
+%\ct{
+%You cannot possibly mean "if and only if" in the two sentences above.
+%Once I show that (1) is unsat iff (2) is unsat, I do not need to argue that
+%(1) is sat iff (2) is sat. That is immediate.
+%}
+% You're right, changed both to "only if" --ajr
 As stated
@@
 
-\jb{Add link to our evaluation data, like we did in CADE paper}
-
 In this section, we evaluate both the overall impact of the encoding introduced
@@
 benchmarks in the encoding.
+The experimental data is publicly available.%
+\footnote{\url{http://lara.epfl.ch/~reynolds/SMT2015-recfun/}}
 
@@
 Note that benchmarks created in this way have a high likelihood of having small
-and easy-to-find counterexamples \ct{contermodels?}. In total, we considered 213 mutated forms of
+and easy-to-find countermodels. In total, we considered 213 mutated forms of
 conjectures from \isa, and 427 mutated forms of conjectures from \leon. We will
@@
 For solvers, we considered the SMT solver \ziii~\cite{de-moura-bjoerner-2008},
-which runs both heuristic methods for quantifier instantiation~\cite{DBLP:conf/cade/MouraB07},
+which runs both heuristic methods for quantifier instantiation~\cite{MouraBjoerner07},
 as well as methods for finding models for quantified formulas~\cite{GeDeM-CAV-09}.
@@
 studied many times.
-The text below is partly adapted from Blanchette's Ph.D.\
-thesis \cite{blanchette-2012-phd}.
-\ct{The last sentence is strange. I would remove it since Jasmin is one of the authors.}
 
@@
 that our approach might be better?}
-
 \ct{About Leon, the reviewers and readers may wonder why we do not do the same, that is, just progressively unfold the recursive definitions until a model is found.

### 2015-05-07 tex/main.tex
Proof of hard direction in relatively good shape.

@@
 
+\newcommand{\absconstraints}{X}
+
+\newtheorem{cor}{Corollary}
+
 %\input{scalalistings}
@@
 SMT solvers have recently been extended with techniques for finding models
-in presence of quantified formulas.
+in presence of (universally) quantified formulas.
 However, these generally cannot cope with universal quantifiers ranging over
@@
 
+In the following, for a ground formula $\varphi$, 
+we write $\absconstraints( \varphi )$ to denote the set of constraints 
+that force our concretization functions to include the necessary elements in their range
+for determining the satisfiability of $\varphi$ with respect to function definitions in our translation.
+Formally, we define $\absconstraints( \varphi )$ as:
+\[
+\absconstraints( \varphi ) = \{ \exists a : {\fargtype{\con{f}}}. ( \vecfarg{\con{f}}( a ) \teq \vec t ) \mid \con{f}( \vec t ) \in \terms( \varphi ), \con{f} \in \sfundefs{\Sigma} \}
+\]
+The following lemma states the central invariant behind our conversion function $\conv$.
+
 \begin{lemma}\label{lem:conv}
-If $\varphi$ is a formula not containing function definitions,
-and $\I$ is a $\extendsig{\Sigma}$-interpretation satisfying $\conv( \varphi )$,
-then $\I$ satisfies a set of $\Sigma$-literals $L$ that entail $\varphi$,
-and $\I$ satisfies
-$\{ \exists a : {\fargtype{\con{f}}}. ( \vecfarg{\con{f}}( a ) \teq \vec t ) \mid \con{f}( \vec t ) \in \terms( L ), \con{f} \in \sfundefs{\Sigma} \}$.
+Let $\varphi$ be a formula not containing function definitions,
+and $\I$ be a $\extendsig{\Sigma}$-interpretation satisfying $\conv( \varphi )$.
+Then, $\I$ satisfies a set of literals $L \cup \absconstraints( L )$ such that $L$ is a set of $\Sigma$-literals that entail $\varphi$.
+%There exists a set of ground $\Sigma$-literals $L$ such that $L$ entails $\varphi$, 
+%and $\I$ satisfies $L \cup \absconstraints( L )$.
 \end{lemma}
+
+\begin{cor}\label{cor:conv}
+If $\varphi$ is a formula not containing function definitions, then $\conv( \varphi )$ entails $\varphi$.
+\end{cor}
 
@@
 $\DDD \wedge \varphi_0$. 
-We argue that $\I$ satisfies a $\Sigma$-formula
-$\psi$ that is closed under function expansion with respect to $\DDD$, and
-where $\psi$ entails $\varphi_0$.
-To construct $\psi$, we define a sequence of sets of $\extendsig{\Sigma}$-literals $M_0 \subseteq M_1 \subseteq \ldots$ 
-such that each $M_i$ is 
-satisfied by $\I$ and
-can be partitioned into two sets $L_i$ and $X_i$,
-where $L_i$ are $\Sigma$-literals,
-and $X_i$ consists of literals of the form $\exists a : {\fargtype{\con{f}}}. ( \vecfarg{\con{f}}( a ) \teq \vec t )$
-for each term $\con{f}( \vec t ) \in \terms( L_i )$ where $\con{f} \in \sfundefs{\Sigma}$.
-Let $M_0$ be 
-a set of literals over atoms occurring in $\conv{ \varphi_0 }$
-such that $\I$ satisfies each literal in $M_0$ and $\bigwedge M_0$ entails $\conv( \varphi_0 )$.
-By case analysis on the return values of $\conv_0$ 
-it can be shown that $M_0$ can be partitioned into $L_0$ and $X_0$ in the way described above.
+%We argue that $\I$ satisfies a $\Sigma$-formula
+%$\psi$ that is closed under function expansion with respect to $\DDD$, and
+%where $\psi$ entails $\varphi_0$.
+%To construct $\psi$, 
+First, we define a sequence of $\Sigma$-literals sets $L_0 \subseteq L_1 \subseteq \ldots$
+such that $\I$ satisfies $L_i \cup \absconstraints( L_i )$ for each $i = 0, 1, \ldots$.
+Since $\I$ satisfies $\conv( \varphi_0 )$,
+by Lemma~\ref{lem:conv}, 
+$\I$ satisfies a set of literals $L \cup \absconstraints( L )$ where $L$ is a set of $\Sigma$-literals that entail $\varphi_0$.
+Let $L_0 = L$.
 For each $i \geq 0$,
-let $\psi'$ be $\bigwedge \{ \conv( \varphi_f[ \vec t / \vec x ] ) \mid \con{f}( \vec t ) \in \terms( L_i ), f \in \sfundefs{\Sigma} \}$,
-where $\varphi_f$ denotes the function definition of $f$ in $\DDD$.
-Note that $\I$ also satisfies $\psi'$ since $\I$ satisfies $X_i$.
-Let $M_{i+1}$ be the union of $M_0$ and
-a set of literals $M$ over atoms occurring in $\psi'$
-such that $\I$ satisfies each literal in $M$ and $\bigwedge M$ entails $\psi'$,
-where we again may partition $M_{i+1}$ into $L_{i+1}$ and $X_{i+1}$.
-Let $M^\ast$ be the limit of this sequence, 
-which we may partition into (possibly infinite) sets $X^\ast$ and $L^\ast$ in the same way as above.
-Let $\psi$ be the $\Sigma$-formula $\bigwedge L^\ast$.
-We know $\psi$ is closed under function expansion with respect to $\DDD$ due to its construction.
-Since there exists a model for $\psi$ that is closed under function expansion
-with respect to $\DDD$, then since $\DDD$ is admissible, by definition there
-exists a model for $\psi \wedge \DDD$, which entails $\DDD \wedge \varphi_0$,
+let $\psi_i$ be the formula $\bigwedge \{ \conv( \varphi_{\con{f}}[ \vec t / \vec x ] ) \mid \con{f}( \vec t ) \in \terms( L_i ), f \in \sfundefs{\Sigma} \}$,
+where the function definition of $\con{f}$ in $\DDD$ is $\forallf{f} \vec x. \varphi_{\con{f}}$.
+Since $\I$ satisfies $\conv( \forallf{f} \vec x. \varphi_{\con{f}} )$ and $\absconstraints( L_i )$,
+we know that $\I$ also satisfies $\psi_i$.
+Thus by Lemma~\ref{lem:conv},
+$\I$ satisfies a set of literals $L \cup \absconstraints( L )$ where $L$ is a set of $\Sigma$-literals that entail $\psi_i$.
+Let $L_{i+1} = L_0 \cup L$.
+Let $L_n$ be the limit of this sequence,
+and let $\psi$ be the $\Sigma$-formula $\bigwedge L_n$.
+To show $\psi$ is closed under function expansion with respect to $\DDD$,
+first note that by construction $\psi$ entails $\psi_n$.
+For any function $\con{f}$ and terms $\vec t$, since $\varphi_{\con{f}}[ \vec t / \vec x ]$ does not contain function definitions,
+by Corollary~\ref{cor:conv},
+$\conv( \varphi_{\con{f}}[ \vec t / \vec x ] )$ entails $\varphi_{\con{f}}[ \vec t / \vec x ]$.
+Thus, $\psi$ entails $\{ \varphi_{\con{f}}[ \vec t / \vec x ] \mid \con{f}( \vec t ) \in \terms( \psi ), f \in \sfundefs{\Sigma} \}$,
+which by definition implies that $\psi$ is closed under function expansion with respect to $\DDD$.
+Furthermore, $\psi$ entails $\varphi_0$ since $L_0 \subseteq L_n$.
+Since $\psi$ is a $T$-satisfiable formula that is closed under function expansion with respect to $\DDD$ and $\DDD$ is admissible, 
+by definition there exists a $\Sigma$-interpretation $\J$ satisfying $\psi \wedge \DDD$, which entails $\DDD \wedge \varphi_0$,
 which is $\varphi$.
+\ajr{Sketchy referring to "limit" of sequence, if it can be infinite. TODO: clarify/restrict.}
 \qed
 \end{proofsketch}
+
+The intuition of the above proof is as follows.
+
+For clarity, we demonstrate this in the context of a concrete example.
 
@@
 any model of $\varphi_{\con{s}}[t/x]$ must also satisfy $\varphi_{\con{s}}[t-1/x]$ when $t > 0$. Given a model $\I$ of
-formula~(\ref{eq:ex-after}) that interprets $\con{c}$ as $14$, it is easy to see that
-due to the structure of formula~(\ref{eq:ex-after}),
-$\I$ must satisfy the following formula $\psi$:
+formula~(\ref{eq:ex-after}) that interprets $\con{c}$ as $14$,
+due to the structure of formula~(\ref{eq:ex-after}), it can be shown that $\I$ must satisfy the following formula $\psi$:
 %

### 2015-05-07 tex/main.tex
tuning

--- 2015-05-07 tex/main.tex
@@
 
-\ct{The paper used "recursive datatype" in some places and "inductive datatype" in others. I have replaced both by "algebraic datatype" which is more accurate for \cvc.
-("recursive datatype" is a more general notion and "inductive datatype" has a stronger meaning in higher-order logics.)}
+%\ct{The paper used "recursive datatype" in some places and "inductive datatype" in others. I have replaced both by "algebraic datatype" which is more accurate for \cvc.
+%("recursive datatype" is a more general notion and "inductive datatype" has a stronger meaning in higher-order logics.)}  
+%% I agree --JB.
 
@@
 models for quantified
-formulas over uninterpreted types as well as types having a fixed finite
-interpretation, such as fixed-width bit vectors and enumerated datatypes.
+formulas over uninterpreted types and types having a fixed finite
+interpretation. %, such as fixed-width bit vectors and enumerated datatypes.
 These techniques can

### 2015-05-07 tex/main.tex
more

@@
 algebraic datatypes. A simple example where both approaches fail is
-$$\forall n : {\typeint}.\;\, \lite\bigl(
-n \leq 0,\allowbreak\; \con{p}( n ) \teq 1,\allowbreak\; \con{p}( n ) \teq 2 * \con{p}( n - 1 ) \bigr)$$
+$\forall n : {\typeint}.\;\, \lite\bigl(
+n \leq 0,\allowbreak\; \con{p}( n ) \teq 1,\allowbreak\; \con{p}( n ) \teq 2 * \con{p}( n - 1 ) \bigr)$.
 This state of affairs is highly unsatisfactory, given the frequence of

### 2015-05-07 tex/main.tex
more

@@
 %\newcommand{\stypes}[1]{#1^\mathrm{s}}
-\newcommand{\stypes}[1]{#1^\mathrm{y}}
+\newcommand{\stypes}[1]{#1^\mathrm{ty}}
 
@@
 definitions into formulas where finite model finding techniques can be applied.
-The recursive functions must meet a precise semantic criterion to be admissible
+The recursive functions must meet a semantic criterion to be admissible
 (Section~\ref{sec:prelim}). This criterion is met by terminating recursive
 function definitions, but also by tail-recursive functions and even by
-\emph{productive} co-recursive functions. It is, however, not met by inconsistent
+\relax{productive} corecursive functions. It is, however, not met by inconsistent
 definitions such as $\forall n : {\typeint}.\;\, \con{f}(n) \teq \con{f}(n) +
@@
 (Section~\ref{sec:encoding}). The main insight is that a
-recursive definition $\forall x : \tau.\; \const{f}(x) \teq t$
+recursive definition $\forall x : \tau.\;\, \const{f}(x) \teq t$
 can be translated to
 $\forall a :
-\fargtype{\tau}.\; \const{f}(\farg{\con{f}}(a)) \teq t[\concret(a)/x]$, where
+\fargtype{\tau}.\;\, \const{f}(\farg{\con{f}}(a)) \teq t[\concret(a)/x]$, where
 $\fargtype{\tau}$ is an uninterpreted \emph{abstract} type and $\farg{\con{f}} :
-\fargtype{\tau} \to \tau$ a coercion from the abstract type to the concrete
+\fargtype{\tau} \to \tau$ is a coercion from the abstract type to the concrete
 type. The translation preserves satisfiability and, for admissible definitions,
@@
 (Section~\ref{sec:evaluation}). The approach is implemented as a preprocessor
-in \cvc (Section~\ref{sec:front-end}). When invoked with the appropriate
+in \cvc (Section~\ref{sec:front-end}). When invoked with a
 command-line option, the solver assumes that functions introduced using the
@@
 \ct{What the rationale for using the letter $y$ in $\stypes{\Sigma}$?}
+\jb{y as in t\textbf{y}pe. changed to `ty'}
 (or sorts) and a set $\sfuns{\Sigma}$ of function symbols over these types.
@@
 standard Boolean connectives ($\neg$, $\wedge$, $\vee$, etc.),
-and an if-then-else function symbol $\lite : \Bool \times \tau \times \tau \rightarrow \tau$
+and an if--then--else function symbol
+$\lite : \Bool \times \tau \times \tau \rightarrow \tau$
 for each $\tau \in \stypes{\Sigma}$.

### 2015-05-07 tex/main.tex
more

@@
 (\,\emph{for $\con{f}$}\vthinspace) if $\vec x$ is a tuple of variables
-$x_1, \ldots, x_n$ of respective type
-$\tau_1, \ldots, \tau_n$ and $\varphi$ is a quantifier-free formula 
+$x_1 : \tau_1$, $\ldots,$ $x_n : \tau_n$
+and $\varphi$ is a quantifier-free formula 
 $T$-equivalent to $\con{f}( \vec x ) \teq t$ for some term $t$ of type $\tau$.
@@
 A formula $\varphi$ is in \emph{definitional form with respect to}
-$\{ \con{f}_1, \ldots, \con{f}_n \}$ if it is of the form
+$\{ \con{f}_1, \ldots, \con{f}_n \} \subseteq \sfuns{\Sigma}$ if it is of the
+form
 %
@@
 %
-where $\con{f}_1, \ldots, \con{f}_n \in \sfuns{\Sigma}$ are distinct function symbols,
+where $\con{f}_1, \ldots, \con{f}_n$ are distinct function symbols,
 %ct added
@@
 \ct{I think we need more conditions here on how a symbol $\con{f}_i$ can occur in $\varphi_j$.}
+\jb{do we?}
 and $\psi$ contains no function definitions.
@@
 
-Given a signature $\Sigma$,
-we distinguish the subset of (\emph{recursively}) \emph{defined} function
-symbols $\sfundefs{\Sigma} \subseteq \sfuns{\Sigma}$.
+Given a signature $\Sigma$, we distinguish the subset of \emph{defined}
+function symbols $\sfundefs{\Sigma} \subseteq \sfuns{\Sigma}$.
 We consider $\Sigma$-formulas that are in definitional form with respect to
@@
 The set $\DDD$ is \emph{admissible} if for every $T$-satisfiable formula 
-$\varphi$ closed under function expansion wrt $\DDD$
+$\varphi$ closed under function expansion with respect to $\DDD$,
 the formula
-$\varphi \wedge \forallf{\con{f}_1} \vec x.\; \varphi_1 \wedge \cdots \wedge \forallf{\con{f}_n} \vec x.\; \varphi_n$ is also $T$-satisfiable.
+$\varphi \wedge (\forallf{\con{f}_1} \vec x.\; \varphi_1) \wedge \cdots \wedge (\forallf{\con{f}_n} \vec x.\; \varphi_n)$ is also $T$-satisfiable.
 \ct{I}

### 2015-05-07 tex/main.tex
more

@@
 \ct{I think we need more conditions here on how a symbol $\con{f}_i$ can occur in $\varphi_j$.}
-\jb{do we?}
+\jb{do we? I'd think that the admissibility criterion below is enough.}
 and $\psi$ contains no function definitions.
@@
 \ajr{I agree. Jasmin, can you fix this?  We need to be careful in the same way when we talk about the "terminating" functions fragment we handle.}
+\jb{Yes, I need to rework and expand this whole part. Today.}
 e.g.,

### 2015-05-07 tex/main.tex
more

@@
 \newcommand{\farg}[1]{\concret_{#1}}
-\newcommand{\fargx}[2]{\concret_{#1,\vthinspace#2}}
+\newcommand{\fargx}[2]{\concret_{#1,#2}}
 \newcommand{\fargtype}[1]{\alpha_{#1}}
@@
 \begin{figure}[t]
+\normalsize
 \begin{enumerate}
@@
 \end{enumerate}
-\caption{Definition of function $\conv$.}
+\caption{\,Definition of function $\conv$}
 \label{fig:A}
@@
 
-First, we fix a signature $\Sigma$ and a $\Sigma$-formula
+We fix a signature $\Sigma$ and a $\Sigma$-formula
 $\varphi$ in definitional form with respect to $\sfundefs{\Sigma}$,
 whose definitions are admissible.
-%We may translate $\varphi$ into a equisatisfiable formula $\psi$ for which known model-finding procedures~\cite{GeDeM-CAV-09, ReyEtAl-1-RR-13} are applicable.
+%We may translate $\varphi$ into a equisatisfiable formula $\phi'$ for which known model-finding procedures~\cite{GeDeM-CAV-09, ReyEtAl-1-RR-13} are applicable.
 This section presents a method for constructing an extended signature
-$\extendsig{ \Sigma }$ and an $\extendsig{ \Sigma }$-formula $\psi$ such that
-$\psi$ is satisfiable if and only if $\varphi$ is satisfiable---i.e.,
-$\varphi$ and $\psi$ are \relax{equisatisfiable}.
+$\extendsig{ \Sigma }$ and an $\extendsig{ \Sigma }$-formula $\varphi'$ such that
+$\varphi'$ is satisfiable if and only if $\varphi$ is satisfiable---i.e.,
+$\varphi$ and $\varphi'$ are \relax{equisatisfiable}.
 The intuition behind this translation
@@
 suppose our signature $\Sigma$ contains a set of defined function symbols $\sfundefs{\Sigma} \subseteq \sfuns{\Sigma}$.
-We extend the signature of $\Sigma$ to a signature $\extendsig{\Sigma}$ containing the following additional definitions.
 For each $\con{f} : \tau_1 \times \cdots \times \tau_n \rightarrow \tau \in \sfundefs{\Sigma}$,
-the signature $\extendsig{\Sigma}$ includes
+the extended signature $\extendsig{\Sigma}$ includes
 \begin{itemize}
@@
 \rightarrow \typeint$. The result of $\conv( \varphi,\, \ppos )$ is the pair $(
-\psi,\, \ltrue )$, where (after simplification) $\psi$ is the
+\varphi',\, \ltrue )$, where (after simplification) $\varphi'$ is the
 $\extendsig{\Sigma}$-formula
@@
 %We argue that $\I$ satisfies a $\Sigma$-formula
-%$\psi$ that is closed under function expansion with respect to $\DDD$, and
-%where $\psi$ entails $\varphi_0$.
-%To construct $\psi$, 
+%$\varphi'$ that is closed under function expansion with respect to $\DDD$, and
+%where $\varphi'$ entails $\varphi_0$.
+%To construct $\varphi'$, 
 First, we define a sequence of $\Sigma$-literals sets $L_0 \subseteq L_1 \subseteq \ldots$

### 2015-05-07 tex/main.tex
minor edits

@@
 \newcommand\pnegate{-}
+%\newcommand\pnegate{{\sim}}
 
@@
 \fargtype{\tau}.\;\, \const{f}(\farg{\con{f}}(a)) \teq t[\concret(a)/x]$, where
-$\fargtype{\tau}$ is an uninterpreted \emph{abstract} type and $\farg{\con{f}} :
-\fargtype{\tau} \to \tau$ is a coercion from the abstract type to the concrete
+$\fargtype{\tau}$ is an uninterpreted \relax{abstract} type and $\farg{\con{f}} :
+\fargtype{\tau} \to \tau$ converts the abstract type to the concrete
 type. The translation preserves satisfiability and, for admissible definitions,
@@
 to translate and a \relax{polarity}~$p$, which is either $\ppos$, $\pneg$, or
-$\pnone$. \vthinspace$\conv_0$ returns a pair $( t'\negvthinspace,\, \chi )$, where $t'$ is a term of
+$\pnone$ (none). \vthinspace$\conv_0$ returns a pair $( t'\negvthinspace,\, \chi )$, where $t'$ is a term of
 type $\tau$ and $\chi$ is an additional constraint.
@@
 \boolop, i, p )$ returns the polarity of the $i$\vvthinspace{th} argument of an
-application of $\boolop$ having polarity $p$, where $\boolop$ is an interpreted
+application of~$\boolop$ having polarity $p$, where $\boolop$ is an interpreted
 predicate.
@@
 quantification on function definitions for $\con{f}$ to the corresponding
-uninterpreted type $\fargtype{\con{f}}$ and inserts appropriate coercions
+uninterpreted type $\fargtype{\con{f}}$ and inserts appropriate selectors
 $\fargx{\con{f}}{i}$. Second, it augments $\varphi$ with additional existential
@@
 \in \{{\neg}, {\wedge}, {\vee}, {\teq}, {\lite}\}$,
-we recursively call $\conv$ on the arguments $t_i$ and polarity $\pol( \boolop,
-i, p )$, where $\pol$ is defined as
+$\conv_0$ calls itself recursively on the arguments $t_i$ and polarity $\pol(
+\boolop, i, p )$, where $\pol$ is defined as
 \begin{equation*}
@@
 If $t$ is positive, $\conv_0$ adds the constraint $\chi$ derived from
-the subterms as a conjunction. Dually, if $t$ is negative, the
-function constructs a disjunction with the negated constraint. It $t$ has no
-polarity, the constrained $\chi$ is simply returned to the caller.
+the subterms as a conjunction. Dually, if $t$ is negative, it constructs a
+disjunction with the negated constraint. It $t$ has no polarity, the
+constraint $\chi$ is simply returned to the caller.
 

### 2015-05-07 tex/main.tex
Improvements to explanation of Example 2.

@@
 constraints
-$\exists a : \fargtype{\con{f}}.\;\, \vecfarg{\con{f}}( a ) \teq \vec t$; these
+$\exists a : \fargtype{\con{f}}.\;\, (\vecfarg{\con{f}}( a ) \teq \vec t)$; these
 ensure that the restricted definition for $\con{f}$ covers all relevant tuples
@@
 and $\I$ be a $\extendsig{\Sigma}$-interpretation satisfying $\conv( \varphi )$.
-Then, $\I$ satisfies a set of literals $L \cup \absconstraints( L )$ such that $L$ is a set of $\Sigma$-literals that entail $\varphi$.
+Then, $\I$ satisfies $L \cup \absconstraints( L )$, where $L$ is a set of ground $\Sigma$-literals that entail $\varphi$.
 %There exists a set of ground $\Sigma$-literals $L$ such that $L$ entails $\varphi$, 
@@
 The intuition of the above proof is as follows.
-
-For clarity, we demonstrate this in the context of a concrete example.
+First, $\conv( \varphi )$ cannot be unsatisfiable when $\varphi$ is satisfiable
+since any model of $\varphi$ can be extended in a straightforward way to satisfy $\conv( \varphi )$.
+Second, if a model is found for $\conv( \varphi )$,
+then the constraints of the form $\exists a : \fargtype{\con{f}}.\;\, (\vecfarg{\con{f}}( a ) \teq \vec t)$
+occurring in $\conv( \varphi )$ ensure that this model also satisfies
+a $\Sigma$-formula that is closed under function expansion that entails the conjecture of $\varphi$.
+Hence when $\DDD$ is admissible, this implies the existence of a model for $\varphi$.
+For clarity, we demonstrate the consequences of this theorem in the context of a concrete example.
 
@@
 Let us revisit the translated formula~(\ref{eq:ex-after}) from
-Example~\ref{ex:translation}. This formula is unsatisfiable only if
+Example~\ref{ex:translation}. 
+This formula is unsatisfiable only if
 formula~(\ref{eq:ex-before}) is unsatisfiable since $\fargtype{f}$ can be
@@
 only if formula~(\ref{eq:ex-before}) is satisfiable, noting that the set
-$\{ \forallf{\con{s}} x\typ{\typeint}.\;\, \lite\bigl( x \leq 0,\allowbreak\; \con{s}(x)
-\teq 0,\; \con{s}( x ) \teq x + \con{s}( x-1 ) \bigr) \}$ is admissible.
+$\{ \forallf{\con{s}} x\typ{\typeint}.\;\, \varphi_\con{s} \}$ is admissible,
+where $\varphi_\con{s}$ is the formula $\lite\bigl( x \leq 0, \con{s}(x) \teq 0, \con{s}( x ) \teq x + \con{s}( x-1 ) \bigr)$.
 %\ct{
@@
 % You're right, changed both to "only if" --ajr
-As stated
-in the proof of Theorem~\ref{thm:equi}, the reason is that any model of
-formula~(\ref{eq:ex-after}) must satisfy a formula that is closed under
-function expansion, hence implying that formula~(\ref{eq:ex-before}) has a
-model. In more detail, let $\varphi_\con{s}$ be the formula $\forallf{\con{s}} x.\;\,
-\lite\bigl( x \leq 0, \con{s}(x) \teq 0, \con{s}( x ) \teq x + \con{s}( x-1 )
-\bigr)$. The rightmost conjunct $\exists a : {\fargtype{\con{s}}}.
-\farg{\con{s}}( a ) \teq \con{c}$ ensures that any model of
-formula~(\ref{eq:ex-after}) satisfying $\con{s}( \con{c} ) > 100$ also
-satisfies $\varphi_{\con{s}}[\con{c}/x]$. The nested existential
-$\exists b.\;\, \farg{\con{s}}( b ) \teq \farg{\con{s}}( a )-1$ ensures that
-a similar constraint is enforced recursively: when $\varphi_{\con{s}}[t/x]$ is used in part to evaluate $\con{s}( \con{c} )$,
-any model of $\varphi_{\con{s}}[t/x]$ must also satisfy $\varphi_{\con{s}}[t-1/x]$ when $t > 0$. Given a model $\I$ of
-formula~(\ref{eq:ex-after}) that interprets $\con{c}$ as $14$,
-due to the structure of formula~(\ref{eq:ex-after}), it can be shown that $\I$ must satisfy the following formula $\psi$:
+%As stated
+%in the proof of Theorem~\ref{thm:equi}, the reason is that any model of
+%formula~(\ref{eq:ex-after}) must satisfy a formula that is closed under
+%function expansion, hence implying that formula~(\ref{eq:ex-before}) has a
+%model.
+%The rightmost conjunct $\exists a : {\fargtype{\con{s}}}. (\farg{\con{s}}( a ) \teq \con{c})$ 
+%ensures that if $\con{s}( \con{c} ) > 100$ is a constraint that 
+%formula~(\ref{eq:ex-after}) satisfying $\con{s}( \con{c} ) > 100$ also
+%satisfies $\varphi_{\con{s}}[\con{c}/x]$. 
+%The nested existential
+%$\exists b.\;\, \farg{\con{s}}( b ) \teq \farg{\con{s}}( a )-1$ ensures that
+%a similar constraint is enforced recursively: when $\varphi_{\con{s}}[t/x]$ is used in part to evaluate $\con{s}( \con{c} )$,
+%any model of $\varphi_{\con{s}}[t/x]$ must also satisfy $\varphi_{\con{s}}[t-1/x]$ when $t > 0$. 
+It is easy to see that any interpretation $\I$ satisfying formula~(\ref{eq:ex-after}) must satisfy $\{ \con{s}( \con{c} ) > 100 \}$ 
+(call this set $L_0$),
+and moreover must satisfy the set $\absconstraints( L_0 ) = \{ \exists a : {\fargtype{\con{s}}}. (\farg{\con{s}}( a ) \teq \con{c}) \}$.
+The latter ensures that $\I$ satisfies a formula that entails both $\con{s}( \con{c} ) > 100$ and $\varphi_{\con{s}}[\con{c}/x]$.
+Since $\I$ also satisfies both the translated function definition for $\con{s}$ (e.g. the first conjunct of formula~(\ref{eq:ex-after})) 
+and $\absconstraints( L_0 )$, it must also satisfy the formula $\psi_1$:
+\[
+\lite\bigl( \con{c} \leq 0,\; 
+            \con{s}(\con{c}) \teq 0,\;
+            \con{s}(\con{c}) \teq \con{c} + \con{s}( \con{c}-1 )
+            \wedge \exists b : {\fargtype{\con{s}}}.\;\, (\farg{\con{s}}( b ) \teq \con{c}-1) \bigr) \bigr)
+\]
+
+The existential $\exists b : {\fargtype{\con{s}}}.\;\, (\farg{\con{s}}( b ) \teq \con{c}-1)$ ensures that a similar constraint is enforced recursively
+when $\I$ satisfies $\neg \con{c} \leq 0$.
+Hence, given a model of formula~(\ref{eq:ex-after}) that interprets $\con{c}$ as $14$,
+this model must satisfy the following formula $\psi$:
 %
@@
 & \con{s}( \con{c} ) > 100 \wedge %\displaystyle
-  \bigwedge\nolimits_{\vthinspace i=0}^{\!13} \bigl( \neg (\con{c}-i \leq 0) \wedge \con{s}( \con{c}-i ) \teq i + \con{s}( \con{c}-i-1 ) \bigr) \\[-\jot]
+  \bigwedge\nolimits_{\vthinspace i=0}^{\!13} \bigl( \neg (\con{c}-i \leq 0) \wedge \con{s}( \con{c}-i ) \teq (\con{c}-i) + \con{s}( (\con{c}-i)-1 ) \bigr) \\[-\jot]
 {\wedge}\; & \con{c}-14 \leq 0 \wedge \con{s}( \con{c}-14 ) \teq 0

### 2015-05-07 tex/main.tex
more

--- 2015-05-07 tex/main.tex
@@
 
-%\section{A Translation for Admissible Function Definitions}
-%\section{A Translation for Function Definitions}
 \section{The Translation}
 \label{sec:encoding}
+
+%%% TYPESETTING: approximated, to get centering
+\newcommand{\itemx}{\itemindent5em\item}
 
@@
 %\begin{framed}
-\item[\ ]
+\itemx[\ ]
 $\conv_0( t\typ{\tau},\, p )$ $=$
@@
  \begin{itemize}
-   \item[] $\mathsf{if}$ $\tau = \Bool$ and $t = \boolop(t_1,\ldots,t_n)$ $\mathsf{then}$
+   \itemx[] $\mathsf{if}$ $\tau = \Bool$ and $t = \boolop(t_1,\ldots,t_n)$ $\mathsf{then}$
     \begin{itemize}
-      \item[] $\mathsf{let}$ $( t'_i,\, \chi_i ) = \conv_0( t_i,\, \pol( \boolop,\, i,\, p ) )$ $\mathsf{for}$ $i = 1, \ldots, n$ $\mathsf{in}$
-      \item[] $\mathsf{let}$ $\chi = \chi_1 \wedge \cdots \wedge \chi_n$ $\mathsf{in}$
-      \item[] $\mathsf{if}$ $p = \ppos$ $\mathsf{then}$
+      \itemx[] $\mathsf{let}$ $( t'_i,\, \chi_i ) = \conv_0( t_i,\, \pol( \boolop,\, i,\, p ) )$ $\mathsf{for}$ $i = 1, \ldots, n$ $\mathsf{in}$%\kern5em$\mid$
+      \itemx[] $\mathsf{let}$ $\chi = \chi_1 \wedge \cdots \wedge \chi_n$ $\mathsf{in}$
+      \itemx[] $\mathsf{if}$ $p = \ppos$ $\mathsf{then}$
       \begin{itemize}
-        \item[] \return \bigtuple{$\boolop(t'_1, \ldots, t'_n) ) \wedge \chi,\, \ltrue$}
+        \itemx[] \return \bigtuple{$\boolop(t'_1, \ldots, t'_n) ) \wedge \chi,\, \ltrue$}
       \end{itemize}
-      \item[] $\mathsf{else}$ $\mathsf{if}$ $p = \pneg$ $\mathsf{then}$
+      \itemx[] $\mathsf{else}$ $\mathsf{if}$ $p = \pneg$ $\mathsf{then}$
       \begin{itemize}
-        \item[] \return \bigtuple{$\boolop(t'_1, \ldots, t'_n) \vee \neg \chi,\, \ltrue$}
+        \itemx[] \return \bigtuple{$\boolop(t'_1, \ldots, t'_n) \vee \neg \chi,\, \ltrue$}
       \end{itemize}
-      \item[] $\mathsf{else}$
+      \itemx[] $\mathsf{else}$
       \begin{itemize}
-        \item[] \return \bigtuple{$\boolop(t'_1, \ldots, t'_n),\, \chi$}
+        \itemx[] \return \bigtuple{$\boolop(t'_1, \ldots, t'_n),\, \chi$}
 \\[-.8\baselineskip]
@@
     \end{itemize}
-  \item[] $\mathsf{else}$ $\mathsf{if}$ $t = \forallf{\con{f}} \vec x.\;\, u$ $\mathsf{then}$
+  \itemx[] $\mathsf{else}$ $\mathsf{if}$ $t = \forallf{\con{f}} \vec x.\;\, u$ $\mathsf{then}$
     \begin{itemize}
-      \item[] $\mathsf{let}$ $( u',\, \chi ) = \conv_0( u,\, p )$ $\mathsf{in}$
-      %\item[] $\mathsf{assert}( \chi = \ltrue )$;
-      \item[] \return \bigtuple{$\forall a : \fargtype{\con{f}}.\; u' [ \vecfarg{\con{f}}( a ) / \vec x ],\, \ltrue$}
+      \itemx[] $\mathsf{let}$ $( u',\, \chi ) = \conv_0( u,\, p )$ $\mathsf{in}$
+      %\itemx[] $\mathsf{assert}( \chi = \ltrue )$;
+      \itemx[] \return \bigtuple{$\forall a : \fargtype{\con{f}}.\; u' [ \vecfarg{\con{f}}( a ) / \vec x ],\, \ltrue$}
 \\[-.8\baselineskip]
     \end{itemize}
-  \item[] $\mathsf{else}$ $\mathsf{if}$ $t = \forall \vec x.\;\, u$ $\mathsf{then}$
+  \itemx[] $\mathsf{else}$ $\mathsf{if}$ $t = \forall \vec x.\;\, u$ $\mathsf{then}$
     \begin{itemize}
-      \item[] $\mathsf{let}$ $( u',\, \chi ) = \conv_0( u,\, p )$ $\mathsf{in}$
-      \item[] \return \bigtuple{$\forall \vec x.\; u',\, \forall \vec x.\; \chi$}
+      \itemx[] $\mathsf{let}$ $( u',\, \chi ) = \conv_0( u,\, p )$ $\mathsf{in}$
+      \itemx[] \return \bigtuple{$\forall \vec x.\; u',\, \forall \vec x.\; \chi$}
 \\[-.8\baselineskip]
     \end{itemize}
-   \item[] $\mathsf{else}$
+   \itemx[] $\mathsf{else}$
    \begin{itemize}
-     \item[] \return \bigtuple{$t,\; \bigwedge\, \{ \exists a : \fargtype{\con{f}}.\; \vecfarg{\con{f}}( a ) \teq \vec t \mid \con{f}( \vec t ) \in \terms( t ),\, \con{f} \in \sfundefs{\Sigma} \}$}
+     \itemx[] \return \bigtuple{$t,\; \bigwedge\, \{ \exists a : \fargtype{\con{f}}.\; \vecfarg{\con{f}}( a ) \teq \vec t \mid \con{f}( \vec t ) \in \terms( t ),\, \con{f} \in \sfundefs{\Sigma} \}$}
    \end{itemize}
@@
 %\begin{framed}
-\item[\ ]
+\itemx[\ ]
 $\conv( \varphi )$ $=$
@@
  \begin{itemize}
-   \item[] $\mathsf{let}$ $( \varphi',\, \_ ) = \conv_0( \varphi,\, \ppos )$ $\mathsf{in}$
-   %\item[] $\mathsf{assert}( D = \emptyset )$;
-   \item[] \return $\varphi'$
+   \itemx[] $\mathsf{let}$ $( \varphi',\, \_ ) = \conv_0( \varphi,\, \ppos )$ $\mathsf{in}$
+   %\itemx[] $\mathsf{assert}( D = \emptyset )$;
+   \itemx[] \return $\varphi'$
   \end{itemize}
@@
 
-If $t$ is a quantified formula, $\conv_0$ recursively calls itself on the body
-with the same polarity. If $t$ is a function definition of a function in
+If $t$ is a function definition of a function in
 $\sfundefs{\Sigma}$, $\conv_0$ constructs a quantified formula over a single variable
@@
 are a top-level conjunct, %by case analysis on the return values of $\conv$,
-we know that $\chi$ is simply $\ltrue$.
-If $t$ is not a function definition, the resulting
-quantified formula is reconstructed with the body $u'$ returned by the
-recursive call, and a quantifier is prefixed to the constraint $\chi$
-returned by the call.
+$\chi$ must be $\ltrue$.
+
+If $t$ is an unannotated quantified formula, $\conv_0$ recursively calls itself
+on the body with the same polarity. A quantifier is prefixed to the
+quantified formula and constraint returned by the recursive call.
 

### 2015-05-07 tex/main.tex
more
More improvements to ex 2.

@@
 \in \sfundefs{\Sigma}$. Such constraints, when asserted positively, ensure that
-the tuple of uninterpreted functions $\vecfarg{\con{f}}$ must include $\vec t$
+the tuple of uninterpreted functions $\vecfarg{\con{f}}$ has $\vec t$
 in its range.
@@
 \begin{equation} \label{eq:ex-before}
-\forallf{\con{s}} n : {\typeint}.\;\, \lite\bigl( n \leq 0,\; \con{s}(n) \teq 0,\;
-  \con{s}( n ) \teq n + \con{s}( n - 1 ) \bigr) \wedge \con{s}( \con{c} ) > 100
+\forallf{\con{s}} x : {\typeint}.\;\, \lite\bigl( x \leq 0,\; \con{s}(x) \teq 0,\;
+  \con{s}( x ) \teq x + \con{s}( x - 1 ) \bigr) \wedge \con{s}( \con{c} ) > 100
 \end{equation}
@@
 Here, $\const{s}$ is a function that returns the sum of the positive integers
-up to $n$. The formula $\varphi$ is in definitional form with
+up to $x$. The formula $\varphi$ is in definitional form with
 respect to $\sfundefs{\Sigma_u}$ and states that the sum of all
@@
 $\fargtype{\con{s}}$ and the uninterpreted function $\farg{\con{s}}$ of type $\fargtype{\con{s}}
-\rightarrow \typeint$. The result of $\conv( \varphi,\, \ppos )$ is the pair $(
-\varphi',\, \ltrue )$, where (after simplification) $\varphi'$ is the
-$\extendsig{\Sigma}$-formula
+\rightarrow \typeint$. The result of $\conv( \varphi )$ 
+is $\varphi'$ which (after simplification) is the $\extendsig{\Sigma}$-formula
 %
@@
 Let $\varphi$ be a formula not containing function definitions,
-and $\I$ be a $\extendsig{\Sigma}$-interpretation satisfying $\conv( \varphi )$.
+and let $\I$ be a $\extendsig{\Sigma}$-interpretation satisfying $\conv( \varphi )$.
 Then, $\I$ satisfies $L \cup \absconstraints( L )$, where $L$ is a set of ground $\Sigma$-literals that entail $\varphi$.
@@
 only if formula~(\ref{eq:ex-before}) is satisfiable, noting that the set
-$\{ \forallf{\con{s}} x\typ{\typeint}.\;\, \varphi_\con{s} \}$ is admissible,
+$\{ \forallf{\con{s}} x.\;\, \varphi_\con{s} \}$ is admissible,
 where $\varphi_\con{s}$ is the formula $\lite\bigl( x \leq 0, \con{s}(x) \teq 0, \con{s}( x ) \teq x + \con{s}( x-1 ) \bigr)$.
@@
 %any model of $\varphi_{\con{s}}[t/x]$ must also satisfy $\varphi_{\con{s}}[t-1/x]$ when $t > 0$. 
-It is easy to see that any interpretation $\I$ satisfying formula~(\ref{eq:ex-after}) must satisfy $\{ \con{s}( \con{c} ) > 100 \}$ 
-(call this set $L_0$),
-and moreover must satisfy the set $\absconstraints( L_0 ) = \{ \exists a : {\fargtype{\con{s}}}. (\farg{\con{s}}( a ) \teq \con{c}) \}$.
-The latter ensures that $\I$ satisfies a formula that entails both $\con{s}( \con{c} ) > 100$ and $\varphi_{\con{s}}[\con{c}/x]$.
+It is easy to see that any interpretation $\I$ satisfying formula~(\ref{eq:ex-after}) satisfies $L_0 \cup \absconstraints( L_0 )$,
+where $L_0 = \{ \con{s}( \con{c} ) > 100 \}$, 
+where $\absconstraints( L_0 )$ contains the single constraint $\exists a : {\fargtype{\con{s}}}. (\farg{\con{s}}( a ) \teq \con{c})$.
+%The latter ensures that $\I$ satisfies a formula that entails both $\con{s}( \con{c} ) > 100$ and $\varphi_{\con{s}}[\con{c}/x]$.
 Since $\I$ also satisfies both the translated function definition for $\con{s}$ (e.g. the first conjunct of formula~(\ref{eq:ex-after})) 
@@
 
-The existential $\exists b : {\fargtype{\con{s}}}.\;\, (\farg{\con{s}}( b ) \teq \con{c}-1)$ ensures that a similar constraint is enforced recursively
-when $\I$ satisfies $\neg \con{c} \leq 0$.
-Hence, given a model of formula~(\ref{eq:ex-after}) that interprets $\con{c}$ as $14$,
-this model must satisfy the following formula $\psi$:
+The resulting existential $\exists b : {\fargtype{\con{s}}}.\;\, (\farg{\con{s}}( b ) \teq \con{c}-1)$ 
+ensures that whenever $\I$ satisfies the set $L_1 = L_0 \cup \{ \neg \con{c} \leq 0, \con{s}(\con{c}) \teq \con{c} + \con{s}( \con{c}-1 ) \}$,
+$\I$ satisfies $\absconstraints( L_1 )$ as well.
+Hence, a model of formula~(\ref{eq:ex-after}) that interprets $\con{c}$ as $n$ must also satisfy the following formula $\psi$:
 %
@@
 & \con{s}( \con{c} ) > 100 \wedge %\displaystyle
-  \bigwedge\nolimits_{\vthinspace i=0}^{\!13} \bigl( \neg (\con{c}-i \leq 0) \wedge \con{s}( \con{c}-i ) \teq (\con{c}-i) + \con{s}( (\con{c}-i)-1 ) \bigr) \\[-\jot]
-{\wedge}\; & \con{c}-14 \leq 0 \wedge \con{s}( \con{c}-14 ) \teq 0
+  \bigwedge\nolimits_{\vthinspace i=0}^{\!n-1} \bigl( \neg (\con{c}-i \leq 0) \wedge \con{s}( \con{c}-i ) \teq (\con{c}-i) + \con{s}( (\con{c}-i)-1 ) \bigr) \\[-\jot]
+{\wedge}\; & \con{c}-n \leq 0 \wedge \con{s}( \con{c}-n ) \teq 0
 \end{split}
@@
 since it contains the $\con{s}$ applications corresponding to
-$\con{s}( \con{c}-i )$ for $i = 0, \ldots, 14$. The formula also
-entails $\varphi_\con{s}[\con{c}-i/x]$ for $i = 0, \ldots, 14$.
+$\con{s}( \con{c}-i )$ for $i = 0, \ldots, n$. The formula also
+entails $\varphi_\con{s}[\con{c}-i/x]$ for $i = 0, \ldots, n$.
 Since $\varphi_\con{s}$ is an admissible definition,
-there must exist a $\Sigma$-interpretation satisfying $\psi \wedge \varphi_\con{s}$
+there must exist a $\Sigma$-interpretation satisfying $\psi \wedge \forallf{\con{s}} x.\;\, \varphi_\con{s}$
 which entails formula~(\ref{eq:ex-before}).\xend

### 2015-05-07 tex/main.tex
more

--- 2015-05-07 tex/main.tex
@@
 Consider the combined signature of linear arithmetic and uninterpreted
-functions $\Sigma = \Sigmalia \mathrel{\cup} \Sigma_{\euf}$, where
+functions $\Sigma = \Sigmalia \mathrel{\uplus} \Sigma_{\euf}$, where
 %$\stypes{\Sigma_{\euf}} = \{ \typeint \}$ 
@@
 %
-Here, $\const{s}$ is a function that returns the sum of the positive integers
-up to $n$. The formula $\varphi$ is in definitional form with
-respect to $\sfundefs{\Sigma_u}$ and states that the sum of all
+The definition of $\const{s}$ specifies that it returns the sum of the
+positive integers up to $n$. The formula $\varphi$ is in definitional form with
+respect to $\sfundefs{\Sigma_{\euf}}$ \jb{correct? It said $\Sigma_u$, which was
+definitely wrong} and states that the sum of all
 positive numbers between $0$ and $\const{c}$ is greater than $100$. This formula is

### 2015-05-07 tex/main.tex

--- 2015-05-07 tex/main.tex
@@
 %
-Here, $\const{s}$ is a function that returns the sum of the positive integers
-up to $x$. The formula $\varphi$ is in definitional form with
-respect to $\sfundefs{\Sigma_u}$ and states that the sum of all
+\jb{I think $x$ is a great name for a metavariable standing for a variable, but
+why insist on using it for concrete variables of type int, when literature
+normally uses $i, \ldots, n$ (and keeps $x$ for real numbers and such)?}
+The definition of $\const{s}$ specifies that it returns the sum of the
+positive integers up to $x$. The formula $\varphi$ is in definitional form with
+respect to $\sfundefs{\Sigma_{\euf}}$ \jb{correct? It said $\Sigma_u$, which was
+definitely wrong} and states that the sum of all
 positive numbers between $0$ and $\const{c}$ is greater than $100$. This formula is

### 2015-05-07 tex/main.tex
more

@@
 definitely wrong} and states that the sum of all
-positive numbers between $0$ and $\const{c}$ is greater than $100$. This formula is
-satisfiable with a model that interprets $\const{c}$ as an integer greater than or
-equal to $14$. Due to the universal quantifier,
-current SMT techniques~\cite{GeDeM-CAV-09, ReyEtAl-1-RR-13,reynolds-et-al-2013} are unable to find
+positive numbers up to $\const{c}$ is greater than $100$. This formula is
+satisfiable with a model that interprets $\const{c}$ as $14$ or more.
+Due to the universal quantifier,
+current SMT techniques
+%~\cite{GeDeM-CAV-09, ReyEtAl-1-RR-13,reynolds-et-al-2013}
+%%% We described them in the intro. When I see references like that appear with
+%%% no description, as a reader, I get all paranoid and wonder if it's the old
+%%% stuff from the intro (Ge, Moura, Renoylds et al.) or if it's new stuff ---
+%%% and I cannot remember numbers like [9,18,19] across sections or even
+%%% paragraphs. --jb
+are unable to find
 a model for this formula. The signature $\extendsig{\Sigma}$ includes the type

### 2015-05-07 tex/main.tex
Minor

@@
 
-The resulting existential $\exists b : {\fargtype{\con{s}}}.\;\, (\farg{\con{s}}( b ) \teq \con{c}-1)$ 
+The existential $\exists b : {\fargtype{\con{s}}}.\;\, (\farg{\con{s}}( b ) \teq \con{c}-1)$ above
 ensures that whenever $\I$ satisfies the set $L_1 = L_0 \cup \{ \neg \con{c} \leq 0, \con{s}(\con{c}) \teq \con{c} + \con{s}( \con{c}-1 ) \}$,
 $\I$ satisfies $\absconstraints( L_1 )$ as well.
-Hence, a model of formula~(\ref{eq:ex-after}) that interprets $\con{c}$ as $n$ must also satisfy the following formula $\psi$:
+Hence, a model of formula~(\ref{eq:ex-after}) that interprets $\con{c}$ as $n$ must also satisfy the formula $\psi$:
 %

### 2015-05-07 tex/main.tex
tuned

@@
 $\varphi'$ is satisfiable if and only if $\varphi$ is satisfiable---i.e.,
-$\varphi$ and $\varphi'$ are \relax{equisatisfiable}.
+$\varphi$ and $\varphi'$ are \emph{equisatisfiable}.
 The intuition behind this translation

### 2015-05-07 tex/main.tex
tuned
Strengthen lemma, main Theorem is now in good shape.

@@
 \newcommand{\M}{\mathcalx{M}}
-\newcommand{\I}{\mathcalx{I}} %%% there's no difference, right? --JB  need to simultaneously refer to 2 different interpretations --ajr.
-%% (I find \mathcalx{I} ugly, with its two horizontal bars, hence the J)
+\newcommand{\I}{\mathcalx{I}}
 \newcommand{\J}{\mathcalx{J}}
@@
 
-\newcommand{\Mo}{{\mathcal{\!J\!}}}
+\newcommand{\Mo}{{\mathcal{I}}}
 
@@
 positive integers up to $x$. The formula $\varphi$ is in definitional form with
-respect to $\sfundefs{\Sigma_{\euf}}$ \jb{correct? It said $\Sigma_u$, which was
-definitely wrong} and states that the sum of all
+respect to $\sfundefs{\Sigma}$ 
+%\jb{correct? It said $\Sigma_u$, which was definitely wrong} 
+and states that the sum of all
 positive numbers up to $\const{c}$ is greater than $100$. This formula is
@@
 Let $\varphi$ be a formula not containing function definitions,
-and let $\I$ be a $\extendsig{\Sigma}$-interpretation satisfying $\conv( \varphi )$.
-Then, $\I$ satisfies $L \cup \absconstraints( L )$, where $L$ is a set of ground $\Sigma$-literals that entail $\varphi$.
+and let $\I$ be a $\extendsig{\Sigma}$-interpretation.
+Then, $\I$ satisfies $\conv( \varphi )$ if and only if
+$\I$ satisfies $L \cup \absconstraints( L )$, where $L$ is a set of ground $\Sigma$-literals that entail $\varphi$.
 %There exists a set of ground $\Sigma$-literals $L$ such that $L$ entails $\varphi$, 
@@
 since by our construction of $\J$ there exists an $a$ such that $a^\J = \vec t^\J$.
-By case analysis it follows that $\conv_0$ maintains the invariant that
-if $\conv_0( t\typ{\Bool}, \ppos )$ returns $( {t'}, \chi )$,
-then $\I$ satisfies $t$ if and only if $\J$ satisfies $t'$.
-Thus, we conclude $\J$ is a model of $\conv( \varphi )$.
+%By case analysis it follows that $\conv_0$ maintains the invariant that
+%if $\conv_0( t\typ{\Bool}, \ppos )$ returns $( {t'}, \chi )$,
+%then $\I$ satisfies $t$ if and only if $\J$ satisfies $t'$.
+Since $\J$ satisfies $\varphi$, it satisfies a set of ground literals $L$ that entail $\varphi$.
+Thus, $\J$ satisfies $L \cup \absconstraints( L )$, and by Lemma~\ref{lem:conv} we conclude $\J$ satisfies $\conv( \varphi )$.
 

### 2015-05-07 tex/main.tex
more

--- 2015-05-07 tex/main.tex
@@
 a translation that eliminates infinite quantifiers in axioms specifying a large
-class of recursive functions, including terminating, tail-recursive, and
+class of recursive functions, including terminating and
 productive functions. An empirical evaluation confirms that the approach works
@@
 (Section~\ref{sec:prelim}). This criterion is met by terminating recursive
-function definitions, but also by tail-recursive functions and even by
+function definitions, but also by
 \relax{productive} corecursive functions. It is, however, not met by inconsistent
@@
 \end{definition}
+
+Admissibility is a semantic criterion that must be satisfied for each function
+definition before applying the translation described in
+Section~\ref{sec:encoding}. This criterion is general enough to accommodate
+\emph{well-founded} function definitions, also called ``terminating'' definitions in a
+slight abuse of terminology. In such definitions, all recursive calls are
+decreasing with respect to a well-founded relation $r$, which must be supplied
+by the user or inferred automatically using a termination prover.
+
+First-order logic has built-in notion of computation or termination. To ensure
+that a function specification is well-founded, it is sufficient to require that
+it would terminate when seen as a functional program for \relax{some}
+evaluation order. For example,
+$\{\forall n : {\typeint}.\;\, \lite\bigl(
+n \leq 0,\allowbreak\; \con{p}( n ) \teq 1,\allowbreak\; \con{p}( n ) \teq 2 * \con{p}( n - 1 ) \bigr)\}$
+can be shown well-founded under an evaluation strategy that consides one of the
+two $\lite$ branches, based on the result of evaluating the condition.
+Logically, such dependencies can be captured by congruence rules. Krauss
+develops these ideas systematically in the more general context of higher-order
+logic \cite[Section 2]{krauss-2009-phd}.
+
+\begin{theorem}
+Well-founded f
+\end{theorem}
+
+\newpage
 

### 2015-05-07 tex/main.tex
more

@@
 definition before applying the translation described in
-Section~\ref{sec:encoding}. This criterion is general enough to accommodate
-\emph{well-founded} function definitions, also called ``terminating'' definitions in a
-slight abuse of terminology. In such definitions, all recursive calls are
-decreasing with respect to a well-founded relation $r$, which must be supplied
-by the user or inferred automatically using a termination prover.
+Section~\ref{sec:encoding}. It is interesting it connect it with the more
+standard notion of \emph{well-founded} function definitions, often called
+``terminating'' definitions in a slight abuse of terminology. In such
+definitions, all recursive calls are decreasing with respect to a well-founded
+relation $r$, which must be supplied by the user or inferred automatically
+using a termination prover. This ensures that the function is uniquely defined
+at all points.
 
@@
 evaluation order. For example,
-$\{\forall n : {\typeint}.\;\, \lite\bigl(
+$\{\forallf{\con{f}} n : {\typeint}.\;\, \lite\bigl(
 n \leq 0,\allowbreak\; \con{p}( n ) \teq 1,\allowbreak\; \con{p}( n ) \teq 2 * \con{p}( n - 1 ) \bigr)\}$
@@
 
+Another interesting fragment is the class of \emph{productive} corecursive
+functions. Corecursive functions are functions to a coalgebraic datatype.
+These functions can call themselves forever, without their being inconsistent.
+Productive corecursive functions are functions that progressively reveal parts
+of their (potentially infinite)
+output \cite{turner-1995,abbott-et-al-2005,mcbride-productive}.
+For example, given a type of infinite streams constructed by $\con{scons}$,
+the definition
+$\{\forallf{\const{e}} n.\;\, \const{e}(n) \teq \con{scons}(n,\; \const{e}(n + 1))\}$
+falls within this fragment, since each call to $\const{e}$ produces one
+constructor before entering the nested call. Like terminating recursion,
+productive corecursion uniquely specifies a function. The following theorem
+connects these notion with admissibility.
+
 \begin{theorem}
-Well-founded f
+Let 
 \end{theorem}
@@
 \begin{verbatim}
-(define-fun-rec e ((n Int)) Int (+ (e n) 1))
+(define-fun-rec h ((n Int)) Int (+ (h n) 1))
 \end{verbatim}

### 2015-05-07 tex/main.tex
Changes to abstract, minor.

@@
 in presence of (universally) quantified formulas.
-However, these generally cannot cope with universal quantifiers ranging over
-infinite types, such as integers and algebraic datatypes. This paper introduces
-a translation that eliminates infinite quantifiers in axioms specifying a large
-class of recursive functions, including terminating and
-productive functions. An empirical evaluation confirms that the approach works
-well on benchmarks from two sources. The translation is implemented as a
-preprocessor in the solver CVC4.
+These techniques are generally limited to universally quantified formulas in some restricted fragment.
+This paper introduces a translation which reduces axioms specifying a large
+class of recursive functions, including terminating, tail-recursive, and
+productive functions to universally quantified formulas for which known SMT solving techniques are applicable.
+An empirical evaluation confirms that the approach improves 
+the performance of existing solvers on benchmarks from two sources. 
+The translation is implemented as a preprocessor in the solver CVC4.
 \end{abstract}
@@
 $\varphi \wedge (\forallf{\con{f}_1} \vec x.\; \varphi_1) \wedge \cdots \wedge (\forallf{\con{f}_n} \vec x.\; \varphi_n)$ is also $T$-satisfiable.
-\ct{I}
+%\ct{I}
 \end{definition}
@@
 standard notion of \emph{well-founded} function definitions, often called
-``terminating'' definitions in a slight abuse of terminology. In such
+\emph{terminating} definitions in a slight abuse of terminology. In such
 definitions, all recursive calls are decreasing with respect to a well-founded
@@
 constructor before entering the nested call. Like terminating recursion,
-productive corecursion uniquely specifies a function. The following theorem
-connects these notion with admissibility.
+productive corecursion totally specify the functions it defines. The
+following theorem connects these notion with admissibility.
 
@@
 Let
+$\DDD = \{ \forallf{\con{f}_1} \vec x.\; \varphi_1, \ldots, \forallf{\con{f}_n} \vec x.\; \varphi_n \}$
+be a set of function definitions such that
+%
+Then the set $\DDD$ is admissible.
 \end{theorem}
+\begin{proofsketch}
+\qed
+\end{proofsketch}
 
@@
 $\I$ satisfies $\absconstraints( L_1 )$ as well.
-Hence, a model of formula~(\ref{eq:ex-after}) that interprets $\con{c}$ as $n$ must also satisfy the formula $\psi$:
+Hence, by repeated applications of this reasoning, it follows that
+a model of formula~(\ref{eq:ex-after}) that interprets $\con{c}$ as $n$ must also satisfy the formula $\psi$:
 %

### 2015-05-07 tex/main.tex
more
minor

--- 2015-05-07 tex/main.tex
@@
 of their (potentially infinite)
-output \cite{turner-1995,abbott-et-al-2005,mcbride-productive}.
+output \cite{turner-1995,mcbride-productive}.
 For example, given a type of infinite streams constructed by $\con{scons}$,
@@
 Let
-$\DDD = \{ \forallf{\con{f}_1} \vec x.\; \varphi_1, \ldots, \forallf{\con{f}_n} \vec x.\; \varphi_n \}$
-be a set of function definitions such that
-%
-Then the set $\DDD$ is admissible.
+$\DDD = \{ \forallf{\con{f}_1} \vec x.\; \varphi_1, \ldots, \forallf{\con{f}_n}
+\vec x.\; \varphi_n \}
+$ be a set of function definitions such that for all
+$(\stypes{\Sigma},\, \sfuns{\Sigma} - \sfundefs{\Sigma})$-interpretations
+$\I$, there exists a unique
+$\Sigma$-interpretation $\I'$ that extends $\I$
+and for which $\I' \models \DDD$.
+Then the set\/ $\DDD$ is admissible.
 \end{theorem}
 \begin{proofsketch}
+Let $\varphi$ be a $T$-satisfiable formula closed under function expansion with
+respect to $\DDD$.
+
+We must show that formula
+$\varphi \wedge (\forallf{\con{f}_1} \vec x.\; \varphi_1) \wedge \cdots \wedge (\forallf{\con{f}_n} \vec x.\; \varphi_n)$ is also $T$-satisfiable.
+
+We have
+$\varphi \,\models_T\, \bigwedge\nolimits_{\,i=1}^{n} \{ \varphi_i[ \vec t / \vec x ] \mid \con{f}_i( \vec t ) \in \terms( \varphi ) \}$.
+
 \qed

### 2015-05-07 tex/main.tex
more

@@
 $\DDD = \{ \forallf{\con{f}_1} \vec x.\; \varphi_1, \ldots, \forallf{\con{f}_n} \vec x.\; \varphi_n \}$, 
-a ground formula $\varphi$ 
-%\ct{I suppose $\varphi$ is meant to be ground.} yes --ajr
+a ground formula $\psi$ 
+%\ct{I suppose $\psi$ is meant to be ground.} yes --ajr
 is \emph{closed under function expansion with respect to $\DDD$} if 
 \[
- \varphi \,\models_T\, \bigwedge\nolimits_{\,i=1}^{n} \{ \varphi_i[ \vec t / \vec x ] \mid \con{f}_i( \vec t ) \in \terms( \varphi ) \} .
+ \psi \,\models_T\, \bigwedge\nolimits_{\,i=1}^{n} \{ \psi_i[ \vec t / \vec x ] \mid \con{f}_i( \vec t ) \in \terms( \psi ) \} .
 \]
 The set $\DDD$ is \emph{admissible} if for every $T$-satisfiable formula 
-$\varphi$ closed under function expansion with respect to $\DDD$,
+$\psi$ closed under function expansion with respect to $\DDD$,
 the formula
-$\varphi \wedge (\forallf{\con{f}_1} \vec x.\; \varphi_1) \wedge \cdots \wedge (\forallf{\con{f}_n} \vec x.\; \varphi_n)$ is also $T$-satisfiable.
+$\psi \wedge \bigwedge \Delta$ is also $T$-satisfiable.
 %\ct{I}
@@
 \vec x.\; \varphi_n \}
-$ be a set of function definitions such that for all
+$ be a set of function definitions such that for every
 $(\stypes{\Sigma},\, \sfuns{\Sigma} - \sfundefs{\Sigma})$-interpretations
-$\I$, there exists a unique
-$\Sigma$-interpretation $\I'$ that extends $\I$
-and for which $\I' \models \DDD$.
+$\I_0$, there exists a unique
+$\Sigma$-interpretation $\I$ that extends $\I_0$
+and for which $\I \models \DDD$.
 Then the set\/ $\DDD$ is admissible.
@@
 
-We must show that formula
-$\varphi \wedge (\forallf{\con{f}_1} \vec x.\; \varphi_1) \wedge \cdots \wedge (\forallf{\con{f}_n} \vec x.\; \varphi_n)$ is also $T$-satisfiable.
-
-We have
-$\varphi \,\models_T\, \bigwedge\nolimits_{\,i=1}^{n} \{ \varphi_i[ \vec t / \vec x ] \mid \con{f}_i( \vec t ) \in \terms( \varphi ) \}$.
+We must show that the formula
+$\varphi \wedge \bigwedge \Delta$ is also $T$-satisfiable.
+
+Let $\J$ be a model of $\varphi$.
+
+Let $\I_0$ be the restriction of $\J$ to the function symbols in $\sfuns{\Sigma} -
+\sfundefs{\Sigma}$. By hypothesis, there exists a $\Sigma$-interpretation
+$\I$ such that $\I \models \Delta$.
+
+We must show that $\I$ coincides with $\J$ at the points of the functions
+$\con{f}_i$ that they care about:
+
+I.e. $\I$ is also a model of $\varphi$.
+And by assumoption $I \models \Delta$, hence
+
+%We have
+%$\varphi \,\models_T\, \bigwedge\nolimits_{\,i=1}^{n} \{ \varphi_i[ \vec t / \vec x ] \mid \con{f}_i( \vec t ) \in \terms( \varphi ) \}$.
 

### 2015-05-07 tex/main.tex
shorten and name

@@
 \begin{abstract}
-SMT solvers have recently been extended with techniques for finding models
-in presence of (universally) quantified formulas.
-These techniques are generally limited to universally quantified formulas in some restricted fragment.
+Model finding in SMT solvers is
+generally limited to universally quantified formulas in some restricted fragment.
 This paper introduces a translation which reduces axioms specifying a large
-class of recursive functions, including terminating, tail-recursive, and
-productive functions to universally quantified formulas for which known SMT solving techniques are applicable.
+class of recursive functions, including terminating and
+productive functions to universally quantified formulas for which known
+techniques are applicable.
 An empirical evaluation confirms that the approach improves 
@@
 algebraic datatypes. A simple example where both approaches fail is
-$\forall n : {\typeint}.\;\, \lite\bigl(
-n \leq 0,\allowbreak\; \con{p}( n ) \teq 1,\allowbreak\; \con{p}( n ) \teq 2 * \con{p}( n - 1 ) \bigr)$.
+\begin{equation}\label{eq:pii}
+\forall n : {\typeint}.\;\, \lite\bigl(
+n \leq 0,\allowbreak\; \con{p}( n ) \teq 1,\allowbreak\; \con{p}( n ) \teq 2 * \con{p}( n - 1 ) \bigr)
+\end{equation}
 This state of affairs is highly unsatisfactory, given the frequence of

### 2015-05-07 tex/main.tex
more

@@
 
-Our empirical evaluation on benchmarks from the IsaPlanner proof planner~\cite{} and Leon verifier~\cite{} provides
+Our empirical evaluation on benchmarks from the IsaPlanner proof planner~\cite\cite{DBLP:conf/itp/JohanssonDB10}
+and the Leon verifier~\cite{suter-et-al-2011} provides
 evidence that this translation improves the effectiveness of the SMT solvers

### 2015-05-07 tex/main.tex
more

--- 2015-05-07 tex/main.tex
@@
 
-Our empirical evaluation on benchmarks from the IsaPlanner proof planner~\cite{} and Leon verifier~\cite{} provides
+Our empirical evaluation on benchmarks from the IsaPlanner proof planner~\cite{DBLP:conf/itp/JohanssonDB10}
+and the Leon verifier~\cite{suter-et-al-2011} provides
 evidence that this translation improves the effectiveness of the SMT solvers

### 2015-05-07 tex/main.tex
Minor, address Leon in related work.

@@
 %The latter ensures that $\I$ satisfies a formula that entails both $\con{s}( \con{c} ) > 100$ and $\varphi_{\con{s}}[\con{c}/x]$.
-Since $\I$ also satisfies both the translated function definition for $\con{s}$ (e.g. the first conjunct of formula~(\ref{eq:ex-after})) 
+Since $\I$ also satisfies both the translated function definition for $\con{s}$, i.e. the first conjunct of formula~(\ref{eq:ex-after}),
 and $\absconstraints( L_0 )$, it must also satisfy the formula $\psi_1$:
@@
 \]
-
-The existential $\exists b : {\fargtype{\con{s}}}.\;\, (\farg{\con{s}}( b ) \teq \con{c}-1)$ above
+The existential $\exists b : {\fargtype{\con{s}}}.\;\, (\farg{\con{s}}( b ) \teq \con{c}-1)$ in the above formula
 ensures that whenever $\I$ satisfies the set $L_1 = L_0 \cup \{ \neg \con{c} \leq 0, \con{s}(\con{c}) \teq \con{c} + \con{s}( \con{c}-1 ) \}$,
@@
 %
-This formula is closed under function expansion,
-since it contains the $\con{s}$ applications corresponding to
-$\con{s}( \con{c}-i )$ for $i = 0, \ldots, n$. The formula also
-entails $\varphi_\con{s}[\con{c}-i/x]$ for $i = 0, \ldots, n$.
+This formula is closed under function expansion
+since it entails $\varphi_\con{s}[\con{c}-i/x]$ for $i = 0, \ldots, n$,
+and contains (only) the $\con{s}$ applications corresponding to $\con{s}( \con{c}-i )$ for $i = 0, \ldots, n$.
 Since $\varphi_\con{s}$ is an admissible definition,
-there must exist a $\Sigma$-interpretation satisfying $\psi \wedge \forallf{\con{s}} x.\;\, \varphi_\con{s}$
+there is a $\Sigma$-interpretation satisfying $\psi \wedge \forallf{\con{s}} x.\;\, \varphi_\con{s}$
 which entails formula~(\ref{eq:ex-before}).\xend
@@
 depth, which is increased iteratively as necessary.
-\jb{Andy: Can you add some more here? You're the Leon pro! E.g. why/when do we think
-that our approach might be better?}
-\ct{About Leon, the reviewers and readers may wonder why we do not do the same, that is, just progressively unfold the recursive definitions until a model is found.
-We should probably say something in this regard.
-}
+The approach described in this paper works in an analagous manner, 
+where instead the SMT solver is invoked only once 
+and where quantifier instantiation amounts to function unfolding.
+It is worth investigating how existing approaches for algorithms based on function unfolding
+can inform approaches for dedicated quantifier instantiation techniques for function definitions,
+and vice versa.
+%\jb{Andy: Can you add some more here? You're the Leon pro! E.g. why/when do we think
+%that our approach might be better?}
+%\ct{About Leon, the reviewers and readers may wonder why we do not do the same, that is, just progressively unfold the recursive definitions until a model is found.
+%We should probably say something in this regard.
+%}
+% It is not a question of "should you instead unfold function definitions?" since that's essential what quantifier instantiation does.  
+% The question is "is quantifier instantiation a good way of unfolding function definitions?" 
+% or "how can we instrument quantifier instantiation to mimic/improve upon existing strategies for function unfolding? --ajr
 

### 2015-05-07 tex/main.tex
more

--- 2015-05-07 tex/main.tex
@@
 \ct{I think we need more conditions here on how a symbol $\con{f}_i$ can occur in $\varphi_j$.}
-\jb{do we? I'd think that the admissibility criterion below is enough.}
+\jb{do we? I thought that the admissibility criterion below is enough.}
 and $\psi$ contains no function definitions.
@@
 it would terminate when seen as a functional program for \relax{some}
-evaluation order. For example,
-$\{\forallf{\con{f}} n : {\typeint}.\;\, \lite\bigl(
-n \leq 0,\allowbreak\; \con{p}( n ) \teq 1,\allowbreak\; \con{p}( n ) \teq 2 * \con{p}( n - 1 ) \bigr)\}$
-can be shown well-founded under an evaluation strategy that consides one of the
-two $\lite$ branches, based on the result of evaluating the condition.
-Logically, such dependencies can be captured by congruence rules. Krauss
-develops these ideas systematically in the more general context of higher-order
-logic \cite[Section 2]{krauss-2009-phd}.
-
-Another interesting fragment is the class of \emph{productive} corecursive
+evaluation order. For example, the definition~(\ref{eq:pii}) from the
+introduction can be shown well-founded under an evaluation strategy that
+consides one of the two $\lite$ branches, based on the result of evaluating the
+condition. Logically, such dependencies can be captured by congruence rules.
+Krauss develops these ideas in the more general context of
+higher-order logic \cite[Section 2]{krauss-2009-phd}.
+
+Another useful fragment is the class of \emph{productive} corecursive
 functions. Corecursive functions are functions to a coalgebraic datatype.
@@
 productive corecursion totally specify the functions it defines. The
-following theorem connects these notion with admissibility.
-
-\begin{theorem}
-Let
-$\DDD = \{ \forallf{\con{f}_1} \vec x.\; \varphi_1, \ldots, \forallf{\con{f}_n}
-\vec x.\; \varphi_n \}
-$ be a set of function definitions such that for every
+following theorem connects totality with admissibility.
+
+\begin{definition}\rm
+Let $\DDD$ be a set of function definitions for $\sfundefs{\Sigma}$. The set
+$\DDD$ is \emph{total} if for every
 $(\stypes{\Sigma},\, \sfuns{\Sigma} - \sfundefs{\Sigma})$-interpretations
@@
 and for which $\I \models \DDD$.
-Then the set\/ $\DDD$ is admissible.
+\end{definition}
+
+\begin{theorem}
+If
+$\DDD = \{ \forallf{\con{f}_1} \vec x.\; \varphi_1, \ldots, \forallf{\con{f}_n}
+\vec x.\; \varphi_n \}
+$ is a total set of function definitions, then it is admissible.
 \end{theorem}
@@
 three-valued logic, inspired by abstract interpretation. It was also the first
-tool of its kind to corecursion and codatatypes
+tool of its kind to corecursion and coalgebraic datatypes
 \cite{blanchette-2013-relational}. The three-valued logic approach extends each
@@
 In future work, it would be interesting to evaluate the approach against other
-model finders, notably Leon \ct{?? Leon is a model finder? Alloy?} and Nitpick, and enrich the benchmark suite with
-problems exercising \cvc's support for codatatypes
+counterexample generators, notably Leon and Nitpick, and enrich the benchmark
+suite with problems exercising \cvc's support for coalgebraic datatypes
 \cite{reynolds-blanchette-2015-codata}. We also plan to integrate \cvc as a
-counterexample generator in proof assistants.
-\ct{Further work could also include identifying simple and/or syntactic sufficient conditions for admissibility. }
+counterexample generator in proof assistants. Finally, further work could also
+include identifying further sufficient conditions for admissibility, thereby
+enlarging the scope of the scheme presented here.
+
+\ct{?? Leon is a model finder? Alloy?}
+\jb{Alloy has no notion of datatypes, so it's pretty useless. Nitpick implements
+datatypes and a more FOL-like logic (actually, HOL) on top of Alloy's backend,
+using the Kuncak \& Jackson idiom for Alloy, so that's as close as we can get
+to evaluating Alloy itself. Leon is, in part, a counterexample generator.
+``Model finder'' is probably a stretch.}
 
@@
 \ackname.}
-We would like to thank Damien Busato-Gaston and Emmanouil Koukoutos for
-providing the initial set of benchmarks used in the evaluation.
+Viktor Kuncak and Stephan Merz have made this work possible. We would also like
+to thank Damien Busato-Gaston and Emmanouil Koukoutos for providing the initial
+set of benchmarks used in the evaluation.
 }

### 2015-05-07 tex/main.tex
Change Leon reference.

@@
 Our empirical evaluation on benchmarks from the IsaPlanner proof planner~\cite{DBLP:conf/itp/JohanssonDB10}
-and the Leon verifier~\cite{suter-et-al-2011} provides
+and the Leon verifier~\cite{blanc2013overview} provides
 evidence that this translation improves the effectiveness of the SMT solvers
@@
 
-The Leon system \cite{suter-et-al-2011} implements a procedure that can both
+The Leon system~\cite{blanc2013overview} implements a procedure that can both
 produces proofs and counterexamples for properties of terminating functions

### 2015-05-07 tex/main.tex
Minor changes/fixes.

@@
 normally uses $i, \ldots, n$ (and keeps $x$ for real numbers and such)?}
+\ajr{To me, $n$ typically refers to some concrete numeral, e.g.
+$14$.
+The reason I changed the instance above is to be consistent with Example 2, which I noticed was using $x$ (and also uses $n$ to refer to a concrete numeral).
+So, I think I'd still prefer $x$ here.}
 The definition of $\const{s}$ specifies that it returns the sum of the
@@
 and each $\fargx{\con{f}}{i}$ as the $i^{th}$ projection on such tuples for each $i = 1, \ldots, n$.
-Note that $\J$ satisfies \emph{every} constraint of the form $\exists a : {\fargtype{\con{f}}}. ( \vecfarg{\con{f}}( a ) \teq \vec t )$,
+Since $\J$ satisfies $\varphi$, it satisfies a set of ground literals $L$ that entail $\varphi$.
+Furthermore, $\J$ satisfies \emph{every} constraint of the form $\exists a : {\fargtype{\con{f}}}. ( \vecfarg{\con{f}}( a ) \teq \vec t )$,
 since by our construction of $\J$ there exists an $a$ such that $a^\J = \vec t^\J$.
@@
 %then $\I$ satisfies $t$ if and only if $\J$ satisfies $t'$.
-Since $\J$ satisfies $\varphi$, it satisfies a set of ground literals $L$ that entail $\varphi$.
 Thus, $\J$ satisfies $L \cup \absconstraints( L )$, and by Lemma~\ref{lem:conv} we conclude $\J$ satisfies $\conv( \varphi )$.
@@
 where $L_0 = \{ \con{s}( \con{c} ) > 100 \}$, 
-where $\absconstraints( L_0 )$ contains the single constraint $\exists a : {\fargtype{\con{s}}}. (\farg{\con{s}}( a ) \teq \con{c})$.
+where $\absconstraints( L_0 )$ is the set consisting of the single constraint $\exists a : {\fargtype{\con{s}}}. (\farg{\con{s}}( a ) \teq \con{c})$.
 %The latter ensures that $\I$ satisfies a formula that entails both $\con{s}( \con{c} ) > 100$ and $\varphi_{\con{s}}[\con{c}/x]$.
@@
 since it entails $\varphi_\con{s}[\con{c}-i/x]$ for $i = 0, \ldots, n$,
-and contains (only) the $\con{s}$ applications corresponding to $\con{s}( \con{c}-i )$ for $i = 0, \ldots, n$.
+and contains only $\con{s}$ applications corresponding to $\con{s}( \con{c}-i )$ for $i = 0, \ldots, n$.
 Since $\varphi_\con{s}$ is an admissible definition,
@@
 \emph{unsat} on a benchmark $\varphi$ and \emph{sat} on its
-corresponding benchmark $\conv( \varphi )$, or vice versa, 
-thereby corroborating Theorem~\ref{thm:equi}.
-\ct{?! \emph{Theorems} do not need corroboration! 
-Maybe we should say "consistently with Theorem~\ref{thm:equi}"?}
+corresponding benchmark $\conv( \varphi )$, or vice versa.
+This is consistent with our expectations and Theorem~\ref{thm:equi}, 
+since these benchmarks contained only sets of function definitions that we expected were admissible.
+%thereby confirming Theorem~\ref{thm:equi}.
+%\ct{?! \emph{Theorems} do not need corroboration! 
+%Maybe we should say "consistently with Theorem~\ref{thm:equi}"?}
 

### 2015-05-07 tex/main.tex
more

--- 2015-05-07 tex/main.tex
@@
 $(\stypes{\Sigma},\, \sfuns{\Sigma} - \sfundefs{\Sigma})$-interpretations
-$\I_0$, there exists a unique
-$\Sigma$-interpretation $\I$ that extends $\I_0$
-and for which $\I \models \DDD$.
+$\J_0$, there exists a unique
+$\Sigma$-interpretation $\J$ that extends $\J_0$
+and for which $\J \models \DDD$.
 \end{definition}
@@
 \begin{proofsketch}
-Let $\varphi$ be a $T$-satisfiable formula closed under function expansion with
-respect to $\DDD$.
-
-We must show that the formula
-$\varphi \wedge \bigwedge \Delta$ is also $T$-satisfiable.
-
-Let $\J$ be a model of $\varphi$.
-
-Let $\I_0$ be the restriction of $\J$ to the function symbols in $\sfuns{\Sigma} -
+Let $\psi$ be a satisfiable formula closed under function expansion with
+respect to $\DDD$. We show that $\psi \wedge \bigwedge \Delta$ is also
+satisfiable. Let $\I$ be a model of $\psi$, and
+let $\J_0$ be the restriction of $\I$ to the function symbols in $\sfuns{\Sigma} -
 \sfundefs{\Sigma}$. By hypothesis, there exists a $\Sigma$-interpretation
-$\I$ such that $\I \models \Delta$.
-
-We must show that $\I$ coincides with $\J$ at the points of the functions
-$\con{f}_i$ that they care about:
-
-I.e. $\I$ is also a model of $\varphi$.
-And by assumoption $I \models \Delta$, hence
+$\J$ such that $\J \models \Delta$.
+
+Since $\psi$ is closed under function expansion, it already constraints the
+functions in $\sfundefs{\Sigma}$ recursively in as much as their interpretation
+is necessary when interpreting $\psi$. Thus, any point $v$ for which
+$\const{f}_i^\I(v)$ is queried when interpreting $\psi$ will have its
+expected value according to its definition and hence coincide with $\J$.
+And since $\psi$ does not depend on the interpretation at the other points,
+$\J$ is, like $\I$, a model of $\psi$. Furthermore, $\J \models \Delta$ by
+assumption. Thus, $\J \models \psi \wedge \bigwedge \Delta$.
 

### 2015-05-07 tex/main.tex
more

--- 2015-05-07 tex/main.tex
@@
 If
-$\DDD = \{ \forallf{\con{f}_1} \vec x.\; \varphi_1, \ldots, \forallf{\con{f}_n}
-\vec x.\; \varphi_n \}
-$ is a total set of function definitions, then it is admissible.
+$\DDD$ is a total set of function definitions for
+$\sfundefs{\Sigma} = \{\con{f}_1,\ldots,\con{f}_n\}$, then it is admissible.
 \end{theorem}
 \begin{proofsketch}
-Let $\varphi$ be a $T$-satisfiable formula closed under function expansion with
-respect to $\DDD$.
-
-We must show that the formula
-$\varphi \wedge \bigwedge \Delta$ is also $T$-satisfiable.
-
-Let $\J$ be a model of $\varphi$.
-
-Let $\I_0$ be the restriction of $\J$ to the function symbols in $\sfuns{\Sigma} -
+Let $\psi$ be a satisfiable formula closed under function expansion with
+respect to $\DDD$. We show that $\psi \wedge \bigwedge \Delta$ is also
+satisfiable. Let $\I$ be a model of $\psi$, and
+let $\J_0$ be the restriction of $\I$ to the function symbols in $\sfuns{\Sigma} -
 \sfundefs{\Sigma}$. By hypothesis, there exists a $\Sigma$-interpretation
-$\I$ such that $\I \models \Delta$.
-
-We must show that $\I$ coincides with $\J$ at the points of the functions
-$\con{f}_i$ that they care about:
-
-I.e. $\I$ is also a model of $\varphi$.
-And by assumoption $I \models \Delta$, hence
+$\J$ such that $\J \models \Delta$.
+
+Since $\psi$ is closed under function expansion, it already constraints the
+functions in $\sfundefs{\Sigma}$ recursively in as much as their interpretation
+is necessary when interpreting $\psi$. Thus, any point $v$ for which
+$\const{f}_i^\I(v)$ is queried when interpreting $\psi$ will have its
+expected value according to its definition and hence coincide with $\J$.
+And since $\psi$ does not depend on the interpretation at the other points,
+$\J$ is, like $\I$, a model of $\psi$. Furthermore, $\J \models \Delta$ by
+assumption. Thus, $\J \models \psi \wedge \bigwedge \Delta$.
 

### 2015-05-07 tex/main.tex
more

@@
 assumption. Thus, $\J \models \psi \wedge \bigwedge \Delta$.
-
 %We have
 %$\varphi \,\models_T\, \bigwedge\nolimits_{\,i=1}^{n} \{ \varphi_i[ \vec t / \vec x ] \mid \con{f}_i( \vec t ) \in \terms( \varphi ) \}$.
-
 \qed
@@
 
-\newpage
-
-For example, if $T$ is the theory of integer arithmetic,
-the set $\{ \forallf{\con{f}} \vec x:\typeint.\;\, \con{f}( x ) \teq 0 \}$ is admissible.
-An admissible set can include definitions of nonterminating functions,
-\ct{There is no notion of termination in FOL. We either define this properly or we just replace "definitions of nonterminating functions, e.g.," with "definitions like"}
-\ajr{I agree. Jasmin, can you fix this?  We need to be careful in the same way when we talk about the "terminating" functions fragment we handle.}
-\jb{Yes, I need to rework and expand this whole part. Today.}
-e.g.,
-$\{ \forallf{\con{f}} x : \typeint.\;\, \con{f}( x ) \teq \con{f}( x - 1 ) \}$ or even $\{ \forallf{\con{f}} x.\allowbreak\;\, \con{f}( x ) \teq \con{f}( x ) \}$ are admissible.
-On the other hand, the set $\{ \forallf{\con{f}} x.\;\, \con{f}( x ) \teq \con{f}( x ) + 1 \}$ is not admissible,
+Beyond totality, an admissible set can include definitions of non-well-founded
+functions---for example,
+$\{ \forallf{\con{f}} x : \typeint.\;\, \con{f}( x ) \teq \con{f}( x - 1 ) \}$
+or even $\{ \forallf{\con{g}} x.\allowbreak\;\, \con{g}( x ) \teq \con{g}( x ) \}$.
+More generally, we conjecture that one can ignore all tail-recursive calls
+(i.e., calls that occur on the right-hand side, possibly in some $\lite$ branch)
+when proving well-foundedness or productivity, without affecting admissibility.
+
+An example of an inadmissible function set is
+$\{ \forallf{\con{f}} x.\;\, \con{f}( x ) \teq \con{f}( x ) + 1 \}$,
 since the formula $\ltrue$ is (trivially) closed under function expansion with respect to this set,
-and there is no model of $T$ satisfying $\forallf{\con{f}} x.\;\, \con{f}( x ) \teq \con{f}( x ) + 1$.
-Consider the set 
-$\{ 
+and there is no model of $T$
+satisfying $\con{f}$'s definition. A more subtle example is given by %the set 
+\[\{ 
  \forallf{\con{f}} x.\;\, \con{f}( x ) \teq \con{f}( x ),\;
@@
    \con{g}( x ) \teq \con{g}( x ) + 1)
-\}$.
+\}\]
 While this set has a model where $\con f$ and $\con g$ are interpreted as the constant $5$, it is not admissible
 since the formula $\con{f}( 0 ) \teq 4$ is closed under function expansion with respect to this set,
-but there exists no interpretation satisfying both $\con{f}( 0 ) \teq 4$ and $\forallf{\con{g}} x.\;\, \lite( \con{f}( x ) \teq 5,\, \con{g}( x ) \teq \con{f}( x ),\, \con{g}( x ) \teq \con{g}( x ) + 1 )$.
-
-\jb{TODO: Jasmin: Identify fragments: terminating recursive functions, etc.}
+but there exists no interpretation satisfying both $\con{f}( 0 ) \teq 4$ and
+$\con{g}$'s definition.
 

### 2015-05-07 tex/main.tex
more

@@
 \}\]
-While this set has a model where $\con f$ and $\con g$ are interpreted as the constant $5$, it is not admissible
-since the formula $\con{f}( 0 ) \teq 4$ is closed under function expansion with respect to this set,
-but there exists no interpretation satisfying both $\con{f}( 0 ) \teq 4$ and
-$\con{g}$'s definition.
+While this set has a model where $\con f$ and $\con g$ are interpreted as the
+constant $5$, it is not admissible since the formula $\con{f}( 0 ) \teq 4$ is
+closed under function expansion with respect to this set and yet there exists
+no interpretation satisfying both $\con{f}( 0 ) \teq 4$ and $\con{g}$'s
+definition.
 

### 2015-05-07 tex/main.tex
more

--- 2015-05-07 tex/main.tex
@@
 \}\]
-While this set has a model where $\con f$ and $\con g$ are interpreted as the constant $5$, it is not admissible
-since the formula $\con{f}( 0 ) \teq 4$ is closed under function expansion with respect to this set,
-but there exists no interpretation satisfying both $\con{f}( 0 ) \teq 4$ and
-$\con{g}$'s definition.
+While this set has a model where $\con f$ and $\con g$ are interpreted as the
+constant function $5$, it is not admissible since %the formula
+$\con{f}( 0 ) \teq 4$ is
+closed under function expansion with respect to this set and yet there exists
+no interpretation satisfying both $\con{f}( 0 ) \teq 4$ and $\con{g}$'s
+definition.
 

### 2015-05-07 tex/main.tex
more

@@
 constructor before entering the nested call. Like terminating recursion,
-productive corecursion totally specify the functions it defines. The
-following theorem connects totality with admissibility.
+productive corecursion totally specify the functions it defines.
+It is even possible to mix recursion and corecursion in the same function
+\cite{blanchette-et-al-2015-fouco}.
+
+Here, we focus on a fragment of all reasonable recursive--corecursive
+specifications, where each function is either well-founded or productive.
+Mutual (co)recursion is not catered for, but it can be encoded in terms of
+simple (co)recursion.
+
+\begin{definition}
+Let $\DDD = \{ \forallf{\con{f}_1} \vec x.\; \varphi_1, \ldots,
+\forallf{\con{f}_n} \vec x.\; \varphi_n \}$ be a set of function definitions.
+This set forms a \emph{layered }(\emph{co})\emph{recursive specification} if
+$\const{f}_i$'s definition may not syntactically refer to $\const{f}_{j}$ for
+all $i < j$ and each $\const{f}_i$ is either well-founded recursive
+or productive corecursive.
+\end{definition}
 
@@
 Let $\DDD$ be a set of function definitions for $\sfundefs{\Sigma}$. The set
-$\DDD$ is \emph{total} if for every
+$\DDD$ is \emph{} if for every
 $(\stypes{\Sigma},\, \sfuns{\Sigma} - \sfundefs{\Sigma})$-interpretations

### 2015-05-07 tex/main.tex
Minor

@@
 essentially uninterpreted fragment does not allow the argument of a
-recursive function to be used deep inside a complex term on the right-hand side,
+recursive function to be used inside a complex term on the right-hand side,
 whereas the finite model finding techniques of Reynolds et al.\ are not
@@
 %
-\ct{I think we need more conditions here on how a symbol $\con{f}_i$ can occur in $\varphi_j$.}
-\jb{do we? I thought that the admissibility criterion below is enough.}
+%\ct{I think we need more conditions here on how a symbol $\con{f}_i$ can occur in $\varphi_j$.}
+%\jb{do we? I thought that the admissibility criterion below is enough.}
+% should be enough, unless I'm missing something? -ajr
 and $\psi$ contains no function definitions.

### 2015-05-07 tex/main.tex
commented out for now

@@
 \[
- \psi \,\models_T\, \bigwedge\nolimits_{\,i=1}^{n} \{ \psi_i[ \vec t / \vec x ] \mid \con{f}_i( \vec t ) \in \terms( \psi ) \} .
+ \psi \,\models_T\, \bigwedge\nolimits_{\,i=1}^{n} \{ \psi_i[ \vec t / \vec x ] \mid \con{f}_i( \vec t ) \in \terms( \psi ) \} %.
 \]
@@
 
-Here, we focus on a fragment of all reasonable recursive--corecursive
-specifications, where each function is either well-founded or productive.
-Mutual (co)recursion is not catered for, but it can be encoded in terms of
-simple (co)recursion.
-
 \begin{definition}
@@
 Let $\DDD$ be a set of function definitions for $\sfundefs{\Sigma}$. The set
-$\DDD$ is \emph{} if for every
+$\DDD$ is \emph{layered-total} if for every
 $(\stypes{\Sigma},\, \sfuns{\Sigma} - \sfundefs{\Sigma})$-interpretations
@@
 Let $\varphi$ be a formula not containing function definitions,
-and let $\I$ be a $\extendsig{\Sigma}$-interpretation.
+and let $\I$ be an $\extendsig{\Sigma}$-interpretation.
 Then, $\I$ satisfies $\conv( \varphi )$ if and only if
@@
 
-\ct{?? Leon is a model finder? Alloy?}
-\jb{Alloy has no notion of datatypes, so it's pretty useless. Nitpick implements
-datatypes and a more FOL-like logic (actually, HOL) on top of Alloy's backend,
-using the Kuncak \& Jackson idiom for Alloy, so that's as close as we can get
-to evaluating Alloy itself. Leon is, in part, a counterexample generator.
-``Model finder'' is probably a stretch.}
+%\ct{?? Leon is a model finder? Alloy?}
+%\jb{Alloy has no notion of datatypes, so it's pretty useless. Nitpick implements
+%datatypes and a more FOL-like logic (actually, HOL) on top of Alloy's backend,
+%using the Kuncak \& Jackson idiom for Alloy, so that's as close as we can get
+%to evaluating Alloy itself. Leon is, in part, a counterexample generator.
+%``Model finder'' is probably a stretch.}
 

### 2015-05-07 tex/main.tex
killed ocmment
killed comment

@@
 %
-\ct{What the rationale for using the letter $y$ in $\stypes{\Sigma}$?}
-\jb{y as in t\textbf{y}pe. changed to `ty'}
+%\ct{What the rationale for using the letter $y$ in $\stypes{\Sigma}$?}
+%\jb{y as in t\textbf{y}pe. changed to `ty'}
 (or sorts) and a set $\sfuns{\Sigma}$ of function symbols over these types.
@@
 %
-\jb{I think $x$ is a great name for a metavariable standing for a variable, but
-why insist on using it for concrete variables of type int, when literature
-normally uses $i, \ldots, n$ (and keeps $x$ for real numbers and such)?}
-\ajr{To me, $n$ typically refers to some concrete numeral, e.g.
-$14$.
-The reason I changed the instance above is to be consistent with Example 2, which I noticed was using $x$ (and also uses $n$ to refer to a concrete numeral).
-So, I think I'd still prefer $x$ here.}
 The definition of $\const{s}$ specifies that it returns the sum of the

### 2015-05-07 tex/main.tex
Minor, change n to x.

@@
 \begin{abstract}
-Model finding in SMT solvers is
-generally limited to universally quantified formulas in some restricted fragment.
+SMT solvers have recently been extended with techniques for finding models
+in presence of universally quantified formulas in some restricted fragments.
 This paper introduces a translation which reduces axioms specifying a large
@@
 \begin{equation}\label{eq:pii}
-\forall n : {\typeint}.\;\, \lite\bigl(
-n \leq 0,\allowbreak\; \con{p}( n ) \teq 1,\allowbreak\; \con{p}( n ) \teq 2 * \con{p}( n - 1 ) \bigr)
+\forall x : {\typeint}.\;\, \lite\bigl(
+x \leq 0,\allowbreak\; \con{p}( x ) \teq 1,\allowbreak\; \con{p}( x ) \teq 2 * \con{p}( x - 1 ) \bigr)
 \end{equation}
@@
 \relax{productive} corecursive functions. It is, however, not met by inconsistent
-definitions such as $\forall n : {\typeint}.\;\, \con{f}(n) \teq \con{f}(n) +
+definitions such as $\forall x : {\typeint}.\;\, \con{f}(x) \teq \con{f}(x) +
 1$, %ct and it is possible to construct consistent examples that do not meet it.
@@
 consists of the set of all $\Sigma$-interpretations.
-We refer to the type and function symbols in this theory as \emph{uninterpreted}.
-\ct{This is both ambiguous and insufficient.
-Since this is SMT we should fix some background theory $T$,
-then extend its signature with additional symbols and call \emph{those} uninterpreted.
-}
+Informally, we refer to the type and function symbols in this theory as \emph{uninterpreted}.
+%\ct{This is both ambiguous and insufficient.
+%Since this is SMT we should fix some background theory $T$,
+%then extend its signature with additional symbols and call \emph{those} uninterpreted.
+%}
 
@@
 \end{lemma}
+\begin{proofsketch}
+By definition of $\conv$ and case analysis on the return values of $\conv_0$.
+\end{proofsketch}
 
@@
 \begin{verbatim}
-(define-fun-rec s ((n Int)) Int (ite (<= n 0) 0 (+ n (s (- n 1)))))
+(define-fun-rec s ((x Int)) Int (ite (<= x 0) 0 (+ x (s (- x 1)))))
 (declare-fun c () Int)
@@
 \begin{verbatim}
-(define-fun-rec h ((n Int)) Int (+ (h n) 1))
+(define-fun-rec h ((x Int)) Int (+ (h x) 1))
 \end{verbatim}

### 2015-05-07 tex/main.tex
more

--- 2015-05-07 tex/main.tex
@@
 This paper introduces a translation which reduces axioms specifying a large
-class of recursive functions, including terminating and
-productive functions to universally quantified formulas for which known
+class of recursive functions, including well-founded (terminating),
+to universally quantified formulas for which known
 techniques are applicable.
@@
 The recursive functions must meet a semantic criterion to be admissible
-(Section~\ref{sec:prelim}). This criterion is met by terminating recursive
-function definitions, but also by
-\relax{productive} corecursive functions. It is, however, not met by inconsistent
+(Section~\ref{sec:prelim}). This criterion is met by well-founded (terminating)
+recursive function definitions. It is not met by inconsistent
 definitions such as $\forall n : {\typeint}.\;\, \con{f}(n) \teq \con{f}(n) +
-1$, %ct and it is possible to construct consistent examples that do not meet it.
-as well as other example of consistent definition.
+1$. %ct and it is possible to construct consistent examples that do not meet it.
+%as well as other examples of consistent definition.
 
@@
 
+\begin{theorem}\ref{thm:adm}
+If
+$\DDD$ is a set of well-founded function definitions for
+$\sfundefs{\Sigma} = \{\con{f}_1,\ldots,\con{f}_n\}$, then it is admissible.
+\end{theorem}
+\begin{proofsketch}
+Let $\psi$ be a satisfiable formula closed under function expansion with
+respect to $\DDD$. We show that $\psi \wedge \bigwedge \Delta$ is also
+satisfiable. Let $\I$ be a model of $\psi$, and
+let $\J_0$ be the restriction of $\I$ to the function symbols in $\sfuns{\Sigma} -
+\sfundefs{\Sigma}$. By hypothesis, there exists a $\Sigma$-interpretation
+$\J$ such that $\J \models \Delta$.
+
+Since $\psi$ is closed under function expansion, it already constraints the
+functions in $\sfundefs{\Sigma}$ recursively in as much as their interpretation
+is necessary when interpreting $\psi$. Thus, any point $v$ for which
+$\const{f}_i^\I(v)$ is queried when interpreting $\psi$ will have its
+expected value according to its definition and hence coincide with $\J$.
+And since $\psi$ does not depend on the interpretation at the other points,
+$\J$ is, like $\I$, a model of $\psi$. Furthermore, $\J \models \Delta$ by
+assumption. Thus, $\J \models \psi \wedge \bigwedge \Delta$.
+%We have
+%$\varphi \,\models_T\, \bigwedge\nolimits_{\,i=1}^{n} \{ \varphi_i[ \vec t / \vec x ] \mid \con{f}_i( \vec t ) \in \terms( \varphi ) \}$.
+\qed
+\end{proofsketch}
+
 Another useful fragment is the class of \emph{productive} corecursive
@@
 It is even possible to mix recursion and corecursion in the same function
-\cite{blanchette-et-al-2015-fouco}.
-
-\begin{definition}
-Let $\DDD = \{ \forallf{\con{f}_1} \vec x.\; \varphi_1, \ldots,
-\forallf{\con{f}_n} \vec x.\; \varphi_n \}$ be a set of function definitions.
-This set forms a \emph{layered }(\emph{co})\emph{recursive specification} if
-$\const{f}_i$'s definition may not syntactically refer to $\const{f}_{j}$ for
-all $i < j$ and each $\const{f}_i$ is either well-founded recursive
-or productive corecursive.
-\end{definition}
-
-\begin{definition}\rm
-Let $\DDD$ be a set of function definitions for $\sfundefs{\Sigma}$. The set
-$\DDD$ is \emph{layered-total} if for every
-$(\stypes{\Sigma},\, \sfuns{\Sigma} - \sfundefs{\Sigma})$-interpretations
-$\J_0$, there exists a unique
-$\Sigma$-interpretation $\J$ that extends $\J_0$
-and for which $\J \models \DDD$.
-\end{definition}
-
-\begin{theorem}
-If
-$\DDD$ is a total set of function definitions for
-$\sfundefs{\Sigma} = \{\con{f}_1,\ldots,\con{f}_n\}$, then it is admissible.
-\end{theorem}
-\begin{proofsketch}
-Let $\psi$ be a satisfiable formula closed under function expansion with
-respect to $\DDD$. We show that $\psi \wedge \bigwedge \Delta$ is also
-satisfiable. Let $\I$ be a model of $\psi$, and
-let $\J_0$ be the restriction of $\I$ to the function symbols in $\sfuns{\Sigma} -
-\sfundefs{\Sigma}$. By hypothesis, there exists a $\Sigma$-interpretation
-$\J$ such that $\J \models \Delta$.
-
-Since $\psi$ is closed under function expansion, it already constraints the
-functions in $\sfundefs{\Sigma}$ recursively in as much as their interpretation
-is necessary when interpreting $\psi$. Thus, any point $v$ for which
-$\const{f}_i^\I(v)$ is queried when interpreting $\psi$ will have its
-expected value according to its definition and hence coincide with $\J$.
-And since $\psi$ does not depend on the interpretation at the other points,
-$\J$ is, like $\I$, a model of $\psi$. Furthermore, $\J \models \Delta$ by
-assumption. Thus, $\J \models \psi \wedge \bigwedge \Delta$.
-%We have
-%$\varphi \,\models_T\, \bigwedge\nolimits_{\,i=1}^{n} \{ \varphi_i[ \vec t / \vec x ] \mid \con{f}_i( \vec t ) \in \terms( \varphi ) \}$.
-\qed
-\end{proofsketch}
+\cite{blanchette-et-al-2015-fouco}. We believe Theorem~\ref{thm:adm} can
+be extended to such specifications.
 

### 2015-05-07 tex/main.tex

@@
 recursive function definitions. It is not met by inconsistent
-definitions such as $\forall n : {\typeint}.\;\, \con{f}(n) \teq \con{f}(n) +
+definitions such as $\forall x : {\typeint}.\;\, \con{f}(x) \teq \con{f}(x) +
 1$. %ct and it is possible to construct consistent examples that do not meet it.
-%as well as other examples of consistent definition.
+%as well as other example of consistent definition.
 

### 2015-05-07 tex/main.tex
more

@@
 
-\begin{theorem}\ref{thm:adm}
+\begin{theorem}\label{thm:adm}
 If
@@
 $\J$ such that $\J \models \Delta$.
-
+%
 Since $\psi$ is closed under function expansion, it already constraints the
-functions in $\sfundefs{\Sigma}$ recursively in as much as their interpretation
-is necessary when interpreting $\psi$. Thus, any point $v$ for which
+functions in $\sfundefs{\Sigma}$ recursively as far as is
+necessary for interpreting $\psi$. Thus, any point $v$ for which
 $\const{f}_i^\I(v)$ is queried when interpreting $\psi$ will have its
 expected value according to its definition and hence coincide with $\J$.
-And since $\psi$ does not depend on the interpretation at the other points,
-$\J$ is, like $\I$, a model of $\psi$. Furthermore, $\J \models \Delta$ by
-assumption. Thus, $\J \models \psi \wedge \bigwedge \Delta$.
+And since the $\psi^\I$ does not depend on the interpretation at the other
+points, $\J$ is, like $\I$, a model of $\psi$. Furthermore, $\J \models \Delta$
+by assumption. Thus, $\J \models \psi \wedge \bigwedge \Delta$.
 %We have
@@
 \cite{blanchette-et-al-2015-fouco}. We believe Theorem~\ref{thm:adm} can
-be extended to such specifications.
-
-Beyond totality, an admissible set can include definitions of non-well-founded
-functions---for example,
-$\{ \forallf{\con{f}} x : \typeint.\;\, \con{f}( x ) \teq \con{f}( x - 1 ) \}$
-or even $\{ \forallf{\con{g}} x.\allowbreak\;\, \con{g}( x ) \teq \con{g}( x ) \}$.
-More generally, we conjecture that one can ignore all tail-recursive calls
-(i.e., calls that occur on the right-hand side, possibly in some $\lite$ branch)
-when proving well-foundedness or productivity, without affecting admissibility.
+be extended to cover such specifications.
+
+Beyond totality, an admissible set can include underspecified functions
+such as $\{ \forallf{\con{f}} x : \typeint.\;\, \con{f}( x )
+\teq \con{f}( x - 1 ) \}$ or $\{ \forallf{\con{g}} x.\allowbreak\;\, \con{g}( x
+) \teq \con{g}( x ) \}$. More generally, it appears that one can ignore all
+tail-recursive calls (i.e., calls that occur on the right-hand side, possibly
+in some $\lite$ branch) when establishing well-foundedness or productivity,
+without affecting admissibility, but this remains to be proved.
 
@@
 and there is no model of $T$
-satisfying $\con{f}$'s definition. A more subtle example is given by %the set 
+satisfying $\con{f}$'s definition. A more subtle specimen is
 \[\{ 

### 2015-05-07 tex/main.tex
fix

@@
 definition before applying the translation described in
-Section~\ref{sec:encoding}. It is interesting it connect it with the more
+Section~\ref{sec:encoding}. It is interesting it connect it with the
 standard notion of \emph{well-founded} function definitions, often called
@@
 let $\J_0$ be the restriction of $\I$ to the function symbols in $\sfuns{\Sigma} -
-\sfundefs{\Sigma}$. By hypothesis, there exists a $\Sigma$-interpretation
-$\J$ such that $\J \models \Delta$.
+\sfundefs{\Sigma}$. Because well-founded definitions uniquely characterize
+the interpretation of the functions they define, there exists a
+$\Sigma$-interpretation $\J$ such that $\J \models \Delta$.
 %
@@
 expected value according to its definition and hence coincide with $\J$.
-And since the $\psi^\I$ does not depend on the interpretation at the other
+And since $\psi^\I$ does not depend on the interpretation at the other
 points, $\J$ is, like $\I$, a model of $\psi$. Furthermore, $\J \models \Delta$

### 2015-05-07 tex/main.tex
more

@@
 \cite{ReyEtAl-1-RR-13,reynolds-et-al-2013}, in the text already.
-Outside the world of SMT, model finding has been
-studied many times.
-
-Some finite model finders are based on a reduction to Boolean satisfiability
-(SAT). They translate the input problem to propositional logic and pass it to a
-SAT solver. The translation is parameterized by upper or exact finite bounds on
+The finite model finding support in the instantiation-based iProver \cite{xxxx}
+is also closery related, given the similarities with SMT.
+
+%Model finding has been studied outside the world of SMT.
+
+Some finite model finders are based on a reduction to a decidable logic,
+typically propositional logic. They translate the input problem to the weaker
+logic and pass it to a solver for that logic.
+The translation is parameterized by upper or exact finite bounds on
 the cardinalities of the atomic types. This procedure was pioneered by McCune
 in the earlier versions of Mace (also capitalized MACE) \cite{mccune-1994}.
-Other SAT-based finders are Paradox \cite{claessen-sorensson-2003}, FMDarwin
-\cite{xxxx}, and iProver \cite{xxxx} for first-order logic with equality;
-\ct{FMDarwin is not SAT-based although it is Mace-like.
-Calling iProver SAT-based is a stretch since it is based 
-on a natively first-order calculus even it uses an auxiliary SAT solver heavily.
-}
-Alloy Analyzer and its back-end Kodkod \cite{torlak-jackson-2007} for
-first-order relational logic; and Refute \cite{weber-2008} and Nitpick
+Other conceptually similar finders are Paradox \cite{claessen-sorensson-2003}
+and FM-Darwin \cite{baumgartner-et-al-2009} for first-order logic with
+equality; the Alloy Analyzer and its back-end Kodkod \cite{torlak-jackson-2007}
+for first-order relational logic; and Refute \cite{weber-2008} and Nitpick
 \cite{blanchette-nipkow-2010} for higher-order logic.
 
-An alternative to translating the problem to propositional logic is to perform
+An alternative to translating the problem is to perform
 an exhaustive model search directly on the original problem. Given fixed
@@
 where instead the SMT solver is invoked only once 
-and where quantifier instantiation amounts to function unfolding.
-It is worth investigating how existing approaches for algorithms based on function unfolding
-can inform approaches for dedicated quantifier instantiation techniques for function definitions,
-and vice versa.
-%\jb{Andy: Can you add some more here? You're the Leon pro! E.g. why/when do we think
-%that our approach might be better?}
-%\ct{About Leon, the reviewers and readers may wonder why we do not do the same, that is, just progressively unfold the recursive definitions until a model is found.
-%We should probably say something in this regard.
-%}
-% It is not a question of "should you instead unfold function definitions?" since that's essential what quantifier instantiation does.  
-% The question is "is quantifier instantiation a good way of unfolding function definitions?" 
-% or "how can we instrument quantifier instantiation to mimic/improve upon existing strategies for function unfolding? --ajr
+and where quantifier instantiation amounts to function unfolding. It is worth
+investigating how existing approaches for algorithms based on function
+unfolding can inform approaches for dedicated quantifier instantiation
+techniques for function definitions, and vice versa.
 
@@
 tools are optimized to work with problems that correspond to total functional
-programs. The QuickCheck tool for Haskell \cite{claessen-hughes-2000} is an early example, based
+programs. The QuickCheck tool for Haskell \cite{claessen-hughes-2000} is an
+early example, based
 on random testing. Bounded exhaustive testing \cite{runciman-et-al-2008} and

### 2015-05-07 tex/main.tex
more

@@
 $\conv(\varphi)$ is equisatisfiable with $\varphi$ for any input $\varphi$.
-\ct{Must define "equisatisfiable".
-What we mean here is a weak form of satisfiability, we could instead say that
-$\conv(\varphi)$ is $T$-satisfiable iff $\varphi$ is $T$-satisfiable.
-} 
+%\ct{Must define "equisatisfiable".
+%What we mean here is a weak form of satisfiability, we could instead say that
+%$\conv(\varphi)$ is $T$-satisfiable iff $\varphi$ is $T$-satisfiable.
+%} 
+%\jb{We do define it, even it is weak form.}
 The intuition is that the

### 2015-05-07 tex/main.tex
more

@@
 
-First-order logic has built-in notion of computation or termination. To ensure
+First-order logic has no built-in notion of computation or termination. To ensure
 that a function specification is well-founded, it is sufficient to require that
-it would terminate when seen as a functional program for \relax{some}
-evaluation order. For example, the definition~(\ref{eq:pii}) from the
+the function would terminate when seen as a functional program, under \emph{some}
+evaluation order. For example, definition~(\ref{eq:pii}) from the
 introduction can be shown well-founded under an evaluation strategy that
@@
 condition. Logically, such dependencies can be captured by congruence rules.
-Krauss develops these ideas in the more general context of
+Krauss developed these ideas in the more general context of
 higher-order logic \cite[Section 2]{krauss-2009-phd}.
@@
 \begin{theorem}\label{thm:adm}
-If
-$\DDD$ is a set of well-founded function definitions for
+If\/
+$\DDD$ is a set of well-founded function definitions for\/
 $\sfundefs{\Sigma} = \{\con{f}_1,\ldots,\con{f}_n\}$, then it is admissible.
@@
 And since $\psi^\I$ does not depend on the interpretation at the other
-points, $\J$ is, like $\I$, a model of $\psi$. Furthermore, $\J \models \Delta$
-by assumption. Thus, $\J \models \psi \wedge \bigwedge \Delta$.
+points, $\J$ is, like $\I$, a model of $\psi$.
+Since $\J \models \Delta$, we have $\J \models \psi \wedge \bigwedge \Delta$,
+as desired.
 %We have

### 2015-05-07 tex/main.tex
more
more refs

@@
 \renewcommand\vec[1]{\overline{#1}}
+
+\let\oldcup=\cup
+\def\cup{\mathrel{\oldcup}}
 
@@
    \itemx[] \return $\varphi'$
+\\[-1.5\baselineskip] %% TYPESETTING
   \end{itemize}
@@
 and let $\I$ be an $\extendsig{\Sigma}$-interpretation.
-Then, $\I$ satisfies $\conv( \varphi )$ if and only if
+Then $\I$ satisfies $\conv( \varphi )$ if and only if
 $\I$ satisfies $L \cup \absconstraints( L )$, where $L$ is a set of ground $\Sigma$-literals that entail $\varphi$.
@@
 which is $\varphi$.
-\ajr{Sketchy referring to "limit" of sequence, if it can be infinite. TODO: clarify/restrict.}
+%\ajr{Sketchy referring to "limit" of sequence, if it can be infinite. TODO: clarify/restrict.}
 \qed
@@
 \begin{quote}
-\small
+%\small -- no small, unless you convince me that these mixtures of font sizes
+% are aesthetic and make sense --jb
 \begin{verbatim}
@@
 \begin{quote}
-\small
+%\small
 \begin{verbatim}
@@
 \cite{ReyEtAl-1-RR-13,reynolds-et-al-2013}, in the text already.
-The finite model finding support in the instantiation-based iProver \cite{xxxx}
-is also closery related, given the similarities with SMT.
+The finite model finding support in the instantiation-based iProver
+\cite{korovin-2013} is also close, given the similarities with SMT.
 

### 2015-05-07 tex/main.tex
reformat
more

@@
  \begin{itemize}
-   \itemx[] $\mathsf{if}$ $\tau = \Bool$ and $t = \boolop(t_1,\ldots,t_n)$ $\mathsf{then}$
+   \itemx[] $\mathsf{if}$ $\tau = \Bool$ $\mathsf{and}$ $t = \boolop(t_1,\ldots,t_n)$ $\mathsf{then}$
     \begin{itemize}
@@
       \itemx[] $\mathsf{let}$ $\chi = \chi_1 \wedge \cdots \wedge \chi_n$ $\mathsf{in}$
-      \itemx[] $\mathsf{if}$ $p = \ppos$ $\mathsf{then}$
-      \begin{itemize}
-        \itemx[] \return \bigtuple{$\boolop(t'_1, \ldots, t'_n) ) \wedge \chi,\, \ltrue$}
-      \end{itemize}
-      \itemx[] $\mathsf{else}$ $\mathsf{if}$ $p = \pneg$ $\mathsf{then}$
-      \begin{itemize}
-        \itemx[] \return \bigtuple{$\boolop(t'_1, \ldots, t'_n) \vee \neg \chi,\, \ltrue$}
-      \end{itemize}
-      \itemx[] $\mathsf{else}$
-      \begin{itemize}
-        \itemx[] \return \bigtuple{$\boolop(t'_1, \ldots, t'_n),\, \chi$}
+      \itemx[] $\mathsf{if}$ $p = \ppos$ $\mathsf{then}$ \bigtuple{$\boolop(t'_1, \ldots, t'_n) ) \wedge \chi,\, \ltrue$}
+      \itemx[] $\mathsf{else}$ $\mathsf{if}$ $p = \pneg$ $\mathsf{then}$ \bigtuple{$\boolop(t'_1, \ldots, t'_n) \vee \neg \chi,\, \ltrue$}
+      \itemx[] $\mathsf{else}$ \bigtuple{$\boolop(t'_1, \ldots, t'_n),\, \chi$}
 \\[-.8\baselineskip]
-      \end{itemize}
     \end{itemize}
@@
     \begin{itemize}
-      \itemx[] $\mathsf{let}$ $( u',\, \chi ) = \conv_0( u,\, p )$ $\mathsf{in}$
-      %\itemx[] $\mathsf{assert}( \chi = \ltrue )$;
-      \itemx[] \return \bigtuple{$\forall a : \fargtype{\con{f}}.\; u' [ \vecfarg{\con{f}}( a ) / \vec x ],\, \ltrue$}
+      \itemx[] $\mathsf{let}$ $( u',\, \chi ) = \conv_0( u,\, p )$ $\mathsf{in}$ \bigtuple{$\forall a : \fargtype{\con{f}}.\; u' [ \vecfarg{\con{f}}( a ) / \vec x ],\, \ltrue$}
 \\[-.8\baselineskip]
@@
     \begin{itemize}
-      \itemx[] $\mathsf{let}$ $( u',\, \chi ) = \conv_0( u,\, p )$ $\mathsf{in}$
-      \itemx[] \return \bigtuple{$\forall \vec x.\; u',\, \forall \vec x.\; \chi$}
+      \itemx[] $\mathsf{let}$ $( u',\, \chi ) = \conv_0( u,\, p )$ $\mathsf{in}$ \bigtuple{$\forall \vec x.\; u',\, \forall \vec x.\; \chi$}
 \\[-.8\baselineskip]
@@
 \itemx[\ ]
-$\conv( \varphi )$ $=$
-\\[-.8\baselineskip]
- \begin{itemize}
-   \itemx[] $\mathsf{let}$ $( \varphi',\, \_ ) = \conv_0( \varphi,\, \ppos )$ $\mathsf{in}$
-   %\itemx[] $\mathsf{assert}( D = \emptyset )$;
-   \itemx[] \return $\varphi'$
+$\conv( \varphi )$ $=$ $\mathsf{let}$ $( \varphi',\, \_ ) = \conv_0( \varphi,\, \ppos )$ $\mathsf{in}$ $\varphi'$
 \\[-1.5\baselineskip] %% TYPESETTING
-  \end{itemize}
 \end{enumerate}
@@
 variety of inductive theorem provers. These benchmarks heavily involve
-recursively-defined functions, and are limited to the combined theory of uninterpreted
+recursive functions and are limited to the combined theory of uninterpreted
 functions and algebraic datatypes. The second source consists of 166
@@
 functions provided by the user are admissible.
-Admissibility must be checked separately \emph{by the user}---for example, using a
+Admissibility must be checked separately \relax{by the user}---for example, using a
 syntactic criterion or a termination prover.

### 2015-05-07 tex/main.tex
more
some edits

@@
 %
-to the above problem, \cvc will still return ``satisfiable"
+to the above problem, \cvc will answer \emph{sat}
 if \texttt{--fmf-fun} is specified.
@@
 quantifiers ranging over datatypes). This idiom was further developed by Dunets
-et al.\ \cite{dunets-et-al-2010}, who showed how to handle primitive recursion.
-Their definedness guards play a similar role to the existential constraints
-generated by the $\conv$ translations.
+et al.\ \cite{dunets-et-al-2010}, who presented a translation scheme
+for primitive recursion. Their definedness guards play a similar role to the
+existential constraints generated by our $\conv$ translation.
 
@@
 three-valued logic, inspired by abstract interpretation. It was also the first
-tool of its kind to corecursion and coalgebraic datatypes
+tool of its kind to support corecursion and coalgebraic datatypes
 \cite{blanchette-2013-relational}. The three-valued logic approach extends each

### 2015-05-07 tex/main.tex
more

@@
 class of recursive functions, including well-founded (terminating),
-to universally quantified formulas for which known
+to universally quantified formulas for which these
 techniques are applicable.

### 2015-05-07 tex/main.tex
more

@@
 
-In the past few years, techniques have been developed to finding models for
+In the past few years, techniques have been developed to find models for
 quantified formulas in SMT.
@@
 
-Unfortunately, neither of these fragments can accommodate the vast majority of
+Unfortunately, none of these fragments can accommodate the vast majority of
 quantified formulas that correspond to recursive function definitions: The
@@
 type. The translation preserves satisfiability and, for admissible definitions,
-unsatisfiability, and makes finite model finding possible in practice
-for problems in this class.
+unsatisfiability, and makes finite model finding possible for problems in this
+class.
 
@@
 
-Our setting is a (monomorphic) many-sorted first-order logic
+Our setting is a monomorphic (or many-sorted) first-order logic
 like the one defined by SMT-LIB \cite{smtlib25}.
@@
 (or sorts) and a set $\sfuns{\Sigma}$ of function symbols over these types.
-We assume that signatures always include a Boolean type $\Bool$ and constants
+We assume that signatures always contain a Boolean type $\Bool$ and constants
 $\ltrue, \lfalse : \Bool$ for truth and falsity,
@@
 $\forallf{\const{f}} \vec x.\; \varphi$, where $\con{f} \in \sfuns{\Sigma}$ is
-an uninterpreted function symbol, 
-whose semantics is the same as standard quantified formulas.
+an uninterpreted function symbol. Their
+semantics is the same as for standard quantified formulas.
 Given $\con{f} : \tau_1 \times \cdots \times \tau_n \rightarrow \tau$,
@@
 definition before applying the translation described in
-Section~\ref{sec:encoding}. It is interesting it connect it with the
+Section~\ref{sec:encoding}. It is interesting to connect it to the
 standard notion of \emph{well-founded} function definitions, often called
@@
 definitions, all recursive calls are decreasing with respect to a well-founded
-relation $r$, which must be supplied by the user or inferred automatically
+relation, which must be supplied by the user or inferred automatically
 using a termination prover. This ensures that the function is uniquely defined
@@
 introduction can be shown well-founded under an evaluation strategy that
-consides one of the two $\lite$ branches, based on the result of evaluating the
-condition. Logically, such dependencies can be captured by congruence rules.
-Krauss developed these ideas in the more general context of
+evaluates the condition of a $\lite$ before evaluating the relevant branch,
+ignoring the other branch. Logically, such dependencies can be captured by
+congruence rules. Krauss developed these ideas in the more general context of
 higher-order logic \cite[Section 2]{krauss-2009-phd}.
@@
 satisfiable. Let $\I$ be a model of $\psi$, and
-let $\J_0$ be the restriction of $\I$ to the function symbols in $\sfuns{\Sigma} -
+let $\I_0$ be the restriction of $\I$ to the function symbols in $\sfuns{\Sigma} -
 \sfundefs{\Sigma}$. Because well-founded definitions uniquely characterize
 the interpretation of the functions they define, there exists a
-$\Sigma$-interpretation $\J$ such that $\J \models \Delta$.
+$\Sigma$-interpretation $\J$ that extends $\I_0$ such that $\J \models \Delta$.
 %
@@
 necessary for interpreting $\psi$. Thus, any point $v$ for which
-$\const{f}_i^\I(v)$ is queried when interpreting $\psi$ will have its
+$\const{f}_i^\I(v)$ is needed for interpreting $\psi$ will have its
 expected value according to its definition and hence coincide with $\J$.
@@
 points, $\J$ is, like $\I$, a model of $\psi$.
-Since $\J \models \Delta$, we have $\J \models \psi \wedge \bigwedge \Delta$,
-as desired.
+Since $\J \models \Delta$ by assumption, we have $\J \models \psi \wedge
+\bigwedge \Delta$, as desired.
 %We have
@@
 
-Another useful fragment is the class of \emph{productive} corecursive
-functions. Corecursive functions are functions to a coalgebraic datatype.
-These functions can call themselves forever, without their being inconsistent.
-Productive corecursive functions are functions that progressively reveal parts
-of their (potentially infinite)
+Another useful fragment of function definitions is the class of
+\emph{productive} corecursive functions. Corecursive functions are functions to
+a coalgebraic datatype. These functions can call themselves forever, without
+their being inconsistent. Productive corecursive functions are functions that
+progressively reveal parts of their potentially infinite
 output \cite{turner-1995,mcbride-productive}.
-For example, given a type of infinite streams constructed by $\con{scons}$,
+For example, given a type of infinite streams constructed by
+$\con{scons} : \ty{int} \times \ty{stream} \to \ty{stream}$,
 the definition
-$\{\forallf{\const{e}} n.\;\, \const{e}(n) \teq \con{scons}(n,\; \const{e}(n + 1))\}$
-falls within this fragment, since each call to $\const{e}$ produces one
+$\{\forallf{\const{e}} x.\;\, \const{e}(x) \teq \con{scons}(x,\; \const{e}(x + 1))\}$
+falls within this fragment: Each call to $\const{e}$ produces one
 constructor before entering the nested call. Like terminating recursion,
@@
 It is even possible to mix recursion and corecursion in the same function
-\cite{blanchette-et-al-2015-fouco}. We believe Theorem~\ref{thm:adm} can
-be extended to cover such specifications.
-
-Beyond totality, an admissible set can include underspecified functions
-such as $\{ \forallf{\con{f}} x : \typeint.\;\, \con{f}( x )
-\teq \con{f}( x - 1 ) \}$ or $\{ \forallf{\con{g}} x.\allowbreak\;\, \con{g}( x
-) \teq \con{g}( x ) \}$. More generally, it appears that one can ignore all
+\cite{blanchette-et-al-2015-fouco}. Theorem~\ref{thm:adm} can be extended to
+cover such specifications, based on the observation that an infinite unfolding
+of a corecursive definition yields a uniquely specified infinite object.
+
+Beyond totality, an admissible set can contain underspecified functions
+such as $\forallf{\con{f}} x : \typeint.\;\, \con{f}( x )
+\teq \con{f}( x - 1 )$ or $\forallf{\con{g}} x.\allowbreak\;\, \con{g}( x
+) \teq \con{g}( x )$. More generally, it appears that one can ignore all
 tail-recursive calls (i.e., calls that occur on the right-hand side, possibly
 in some $\lite$ branch) when establishing well-foundedness or productivity,
-without affecting admissibility, but this remains to be proved.
+without affecting admissibility. % but this remains to be proved.
 
@@
 \[\{ 
- \forallf{\con{f}} x.\;\, \con{f}( x ) \teq \con{f}( x ),\;
- \forallf{\con{g}} x.\;\, \lite\bigl( \con{f}( x ) \teq 5,\;
+ \forallf{\con{f}} x.\;\, \con{f}( x ) \teq \con{f}( x ),\;\,
+ \forallf{\con{g}} x.\;\, \lite( \con{f}( x ) \teq 5,\;
    \con{g}( x ) \teq \con{f}( x ),\,
@@
 For each $\con{f} : \tau_1 \times \cdots \times \tau_n \rightarrow \tau \in \sfundefs{\Sigma}$,
-the extended signature $\extendsig{\Sigma}$ includes
+the extended signature $\extendsig{\Sigma}$ contains
 \begin{itemize}
@@
 are unable to find
-a model for this formula. The signature $\extendsig{\Sigma}$ includes the type
+a model for this formula. The signature $\extendsig{\Sigma}$ contains the type
 $\fargtype{\con{s}}$ and the uninterpreted function $\farg{\con{s}}$ of type $\fargtype{\con{s}}
@@
 we write $\absconstraints( \varphi )$ to denote the set of constraints 
-that force our concretization functions to include the necessary elements in their range
+that force our concretization functions to have the necessary elements in their range
 for determining the satisfiability of $\varphi$ with respect to function definitions in our translation.

### 2015-05-07 tex/main.tex
more

@@
 \teq \con{f}( x - 1 )$ or $\forallf{\con{g}} x.\allowbreak\;\, \con{g}( x
-) \teq \con{g}( x )$. More generally, it appears that one can ignore all
-tail-recursive calls (i.e., calls that occur on the right-hand side, possibly
-in some $\lite$ branch) when establishing well-foundedness or productivity,
-without affecting admissibility. % but this remains to be proved.
+) \teq \con{g}( x )$. More generally, we conjecture that one can ignore all
+tail-recursive calls (i.e., calls that occur on the right-hand side of the
+definition, possibly in some $\lite$ branch) when establishing well-foundedness
+or productivity, without affecting admissibility.
 
@@
 \end{enumerate}
-\caption{\,Definition of function $\conv$}
+\caption{\,Definition of translation $\conv$}
 \label{fig:A}
@@
 %We may translate $\varphi$ into a equisatisfiable formula $\phi'$ for which known model-finding procedures~\cite{GeDeM-CAV-09, ReyEtAl-1-RR-13} are applicable.
-This section presents a method for constructing an extended signature
+We present a method that constructs an extended signature
 $\extendsig{ \Sigma }$ and an $\extendsig{ \Sigma }$-formula $\varphi'$ such that
@@
 $\varphi$ and $\varphi'$ are \emph{equisatisfiable}.
-The intuition behind this translation
+The idea behind this translation
 is to use an uninterpreted type to abstract the set of
-\relax{relevant} tuples for each recursive function $\con{f}$ and restrict the
-quantification of the function definition for $\con{f}$ to a variable of this
+\relax{relevant} tuples for each defined function $\con{f}$ and restrict the
+quantification of $\con{f}$'s definition to a variable of this
 type. Informally, the relevant tuples $\vec t$ of a function $\con{f}$ are the
@@
 
-The function $\conv$ defined in Figure~\ref{fig:A} translates the $\Sigma$-formula
+The translation $\conv$ defined in Figure~\ref{fig:A} translates the $\Sigma$-formula
 $\varphi$ into an $\extendsig{\Sigma}$-formula. It relies
@@
 %
-The definition of $\const{s}$ specifies that it returns the sum of the
+The definition of $\const{s}$ specifies that it returns the sum of all
 positive integers up to $x$. The formula $\varphi$ is in definitional form with
@@
 a model for this formula. The signature $\extendsig{\Sigma}$ contains the type
-$\fargtype{\con{s}}$ and the uninterpreted function $\farg{\con{s}}$ of type $\fargtype{\con{s}}
+$\fargtype{\con{s}}$ and the uninterpreted function $\farg{\con{s}} : \fargtype{\con{s}}
 \rightarrow \typeint$. The result of $\conv( \varphi )$ 
@@
 \\[-\jot]
- & {\wedge}\; \bigl( \con{s}( \con{c} ) > 100 \wedge \exists a : {\fargtype{\con{s}}}.\;\, \farg{\con{s}}( a ) \teq \con{c} \bigr)
+ & {\wedge}\; \con{s}( \con{c} ) > 100 \wedge \exists a : {\fargtype{\con{s}}}.\;\, \farg{\con{s}}( a ) \teq \con{c}
 \end{aligned}
@@
 % Yes, I'm planning to post the 2 versions of the benchmark on the evaluation page (when we add it) --ajr
+% Yes, but how can you be sure that Z3 will run forever? Your benchmarks surely have a timeout! --jb
 \cvc and \ziii run indefinitely on formula~(\ref{eq:ex-before}), 
-while they both produce a model for formula~(\ref{eq:ex-after}) 
+whereas they both produce a model for formula~(\ref{eq:ex-after}) 
 within 100 milliseconds.\xend
@@
 %\ct{"Preserves models" means that it does not drop models, not that it adds models.}
-One model $\I$ for formula~(\ref{eq:ex-after}) in the previous example interprets
+One model $\I$ for formula~(\ref{eq:ex-after}) in the above example interprets
 $\fargtype{\con{s}}$ as a finite set of elements $\{ u_0, \ldots, u_{14} \}$,
@@
 In other words, $\const{s}$ is interpreted as a function mapping $x$ to the sum
-of all positive integers between $0$ and $x$ when $0 \leq x \leq 13$, and $105$
+of all positive integers up to $x$ when $0 \leq x \leq 13$, and $105$
 otherwise.
 The interpretation $\I$, restricted to $\Sigma$, is {not} a model for the original formula~(\ref{eq:ex-before}),
-since $\I$ wrongly interprets $\con{s}( x )$ as $105$ when $x < 0$ or $x > 14$.
+since $\I$ wrongly interprets $\con{s}( n )$ as $105$ when $n < 0$ or $n > 14$.
 
@@
 \varphi )$ implies the existence or nonexistence of a model for
-$\varphi$. Moreover, information in models for $\conv( \varphi )$ contains
+$\varphi$. Moreover, the models of $\conv( \varphi )$ contain
 pertinent information regarding the models of $\varphi$. For example, the model
-$\I$ for formula~(\ref{eq:ex-after}) described above interprets $\con{c}$ as $14$,
-and there exists a model of formula~(\ref{eq:ex-before}) that also interprets $\con{c}$
-as $14$. In general, there exists a model of $\varphi$ that
+$\I$ for formula~(\ref{eq:ex-after}) described above interprets $\con{c}$ as $14$
+and $\con{s}(n) = \sum_{i=1}^n i$ for $0 \le n \le 14$,
+and there exists a model of formula~(\ref{eq:ex-before}) that also interprets
+$\con{c}$ and and $\con{s}(n)$ in the same way (for $0 \le n \le 14$).
+In general, there exists a model of $\varphi$ that
 coincides with each model of $\conv( \varphi )$ on its interpretation of all
 functions in $\sfuns{\Sigma} - \sfundefs{\Sigma}$, i.e., $\{
-\con{s} \}$ in the example.
+\con{s} \}$ in the example. Furthermore, the model of $\conv( \varphi )$ will
+also give correct information for the defined functions at all points belonging
+to the domains of the corresponding abstract types. This can sometimes help
+users debug their specifications.
 
@@
 %user, and hence already has an intended interpretation.
+
+%%% @ANDY: Now I understand. My experience with Nitpick is that users sometimes
+%%% want to see the partial information available about defined functions. I've
+%%% added a couple of sentences above to that effect.
 
@@
 \]
-The following lemma states the central invariant behind our conversion function $\conv$.
+The following lemma states the central invariant behind our translation $\conv$.
 
@@
 the command-line option \texttt{--fmf-fun}), it will replace its list of known
-assertions based on the conversion $\conv$ before checking for satisfiability.
+assertions based on the translation $\conv$ before checking for satisfiability.
 %If the functions provided in the input are admissible,
@@
 for primitive recursion. Their definedness guards play a similar role to the
-existential constraints generated by our $\conv$ translation.
+existential constraints generated by our translation $\conv$.
 

### 2015-05-07 tex/main.tex
more

@@
 
-\newcommand{\absconstraints}{X}
+\newcommand{\absconstraints}{\mathrm{X}}
 
@@
 pertinent information regarding the models of $\varphi$. For example, the model
-$\I$ for formula~(\ref{eq:ex-after}) described above interprets $\con{c}$ as $14$
+$\I$ for formula~(\ref{eq:ex-after}) given above interprets $\con{c}$ as $14$
 and $\con{s}(n) = \sum_{i=1}^n i$ for $0 \le n \le 14$,
@@
 
-In the following, for a ground formula $\varphi$, 
-we write $\absconstraints( \varphi )$ to denote the set of constraints 
-that force our concretization functions to have the necessary elements in their range
-for determining the satisfiability of $\varphi$ with respect to function definitions in our translation.
-Formally, we define $\absconstraints( \varphi )$ as:
-\[
-\absconstraints( \varphi ) = \{ \exists a : {\fargtype{\con{f}}}. ( \vecfarg{\con{f}}( a ) \teq \vec t ) \mid \con{f}( \vec t ) \in \terms( \varphi ), \con{f} \in \sfundefs{\Sigma} \}
-\]
-The following lemma states the central invariant behind our translation $\conv$.
+In the following, for a set of ground literals $L$, 
+we write $\absconstraints( L )$ to denote the set of constraints that
+force the concretization functions to have the necessary elements in their
+range for determining the satisfiability of $L$ with respect to function
+definitions in the translation.
+Formally, we define
+$\absconstraints( L )$
+as
+$\{ \exists a : {\fargtype{\con{f}}}.\;\, \vecfarg{\con{f}}( a ) \teq \vec t
+\mid \con{f}( \vec t ) \in \terms( L ),\: \con{f} \in \sfundefs{\Sigma}
+\}$.
+The following lemma states the central invariant behind the translation $\conv$.
 
@@
 \begin{cor}\label{cor:conv}
-If $\varphi$ is a formula not containing function definitions, then $\conv( \varphi )$ entails $\varphi$.
+If $\varphi$ is a formula not containing function definitions, then $\extendsig{\Sigma} = \Sigma$
+and $\conv( \varphi )$ entails $\varphi$.
 \end{cor}
@@
 $\sfundefs{\Sigma}$
-and the set of function definitions\/ $\DDD$ corresponding to functions in
+and the set of function definitions\/ $\DDD$ corresponding to functions in\/
 $\sfundefs{\Sigma}$ is admissible, then\/ $\varphi$ and\/ $\conv( \varphi )$ are

### 2015-05-07 tex/main.tex
more

@@
 all functions $\con{f} \in \sfuns{\Sigma}$ as $\con{f}^\I$,
-and for each function $\con{f}$ of type $\tau_1 \times \cdots \tau_n \rightarrow \tau$ in $\sfundefs{\Sigma}$,
-interprets $\fargtype{\con{f}}$ as a tuple of values of type $\tau_1^\I \times \cdots \times \tau_n^\I$,
-and each $\fargx{\con{f}}{i}$ as the $i^{th}$ projection on such tuples for each $i = 1, \ldots, n$.
+and for each function $\con{f} : \tau_1 \times \cdots \times \tau_n \rightarrow \tau$ in $\sfundefs{\Sigma}$,
+interprets $\fargtype{\con{f}}$ as a tuple of values belonging to $\tau_1^\I \times \cdots \times \tau_n^\I$,
+and each $\fargx{\con{f}}{i}$ as the $i$\vvthinspace{th} projection on such tuples for $i = 1, \ldots, n$.
 Since $\J$ satisfies $\varphi$, it satisfies a set of ground literals $L$ that entail $\varphi$.
-Furthermore, $\J$ satisfies \emph{every} constraint of the form $\exists a : {\fargtype{\con{f}}}. ( \vecfarg{\con{f}}( a ) \teq \vec t )$,
-since by our construction of $\J$ there exists an $a$ such that $a^\J = \vec t^\J$.
+Furthermore, $\J$ satisfies \relax{every} constraint of the form $\exists a : {\fargtype{\con{f}}}.\;\, \vecfarg{\con{f}}( a ) \teq \vec t$,
+since by our construction of $\J$ there exists a value $v \in {\fargtype{\con{f}}}^\J$ such that $v = \vec t^\J$.
 %By case analysis it follows that $\conv_0$ maintains the invariant that
@@
 First, we define a sequence of $\Sigma$-literals sets $L_0 \subseteq L_1 \subseteq \ldots$
-such that $\I$ satisfies $L_i \cup \absconstraints( L_i )$ for each $i = 0, 1, \ldots$.
+such that $\I$ satisfies $L_i \cup \absconstraints( L_i )$ for $i = 0, 1, \ldots$.
 Since $\I$ satisfies $\conv( \varphi_0 )$,

### 2015-05-07 tex/main.tex
cleaned up limit

@@
 Second, we show that if $\conv( \varphi )$ is satisfied by a
-$\extendsig{\Sigma}$-interpretation $\I$, then $\varphi$ is satisfied by
+$\extendsig{\Sigma}$-interpretation $\I$, then $\varphi$ is satisfied by a
 $\Sigma$-interpretation $\J$. Since $\varphi$ is in definitional form with
-respect to the functions defined by $\DDD$, we know $\varphi$ is of the form
+respect to the functions defined by $\DDD$, it must be of the form
 $\DDD \wedge \varphi_0$. 
@@
 %To construct $\varphi'$, 
-First, we define a sequence of $\Sigma$-literals sets $L_0 \subseteq L_1 \subseteq \ldots$
-such that $\I$ satisfies $L_i \cup \absconstraints( L_i )$ for $i = 0, 1, \ldots$.
+First, we define a sequence of $\Sigma$-literals sets $L_0 \subseteq L_1 \subseteq \cdots$
+such that $\I$ satisfies $L_i \cup \absconstraints( L_i )$ for $i = 0, 1, \ldots\vthinspace$.
 Since $\I$ satisfies $\conv( \varphi_0 )$,
@@
 Let $L_{i+1} = L_0 \cup L$.
-Let $L_n$ be the limit of this sequence,
-and let $\psi$ be the $\Sigma$-formula $\bigwedge L_n$.
+Let $L_\infty$ be the limit of this sequence (i.e., $\ell \in L_\infty$ if and only
+$\ell \in L_i$ for some $i$),
+and let $\psi$ be the $\Sigma$-formula $\bigwedge L_\infty$.
 To show $\psi$ is closed under function expansion with respect to $\DDD$,
-first note that by construction $\psi$ entails $\psi_n$.
+first note that by construction $\psi$ entails $\psi_\infty$.
 For any function $\con{f}$ and terms $\vec t$, since $\varphi_{\con{f}}[ \vec t / \vec x ]$ does not contain function definitions,
@@
 which by definition implies that $\psi$ is closed under function expansion with respect to $\DDD$.
-Furthermore, $\psi$ entails $\varphi_0$ since $L_0 \subseteq L_n$.
+Furthermore, $\psi$ entails $\varphi_0$ since $L_0 \subseteq L_\infty$.
 Since $\psi$ is a $T$-satisfiable formula that is closed under function expansion with respect to $\DDD$ and $\DDD$ is admissible, 

### 2015-05-07 tex/main.tex
more

@@
 
-%We write $\functerms^\Sigma( t )$ to denote the set of $f$-applications in $t$ such that $f \in \sfundefs{\Sigma}$.
+%We write $\functerms^\Sigma( t )$ to denote the set of $f$-applications in $t$ such that $\con{f} \in \sfundefs{\Sigma}$.
 
@@
 For each $i \geq 0$,
-let $\psi_i$ be the formula $\bigwedge \{ \conv( \varphi_{\con{f}}[ \vec t / \vec x ] ) \mid \con{f}( \vec t ) \in \terms( L_i ), f \in \sfundefs{\Sigma} \}$,
-where the function definition of $\con{f}$ in $\DDD$ is $\forallf{f} \vec x. \varphi_{\con{f}}$.
-Since $\I$ satisfies $\conv( \forallf{f} \vec x. \varphi_{\con{f}} )$ and $\absconstraints( L_i )$,
+let $\psi_i$ be the formula $\bigwedge\, \{ \conv( \varphi_{\con{f}}[ \vec t / \vec x ] ) \mid \con{f}( \vec t ) \in \terms( L_i ),\: \con{f} \in \sfundefs{\Sigma} \}$,
+where the function definition of $\con{f}$ in $\DDD$ is $\forallf{\con{f}} \vec x.\; \varphi_{\con{f}}$.
+Since $\I$ satisfies $\conv( \forallf{\con{f}} \vec x.\; \varphi_{\con{f}} )$ and $\absconstraints( L_i )$,
 we know that $\I$ also satisfies $\psi_i$.
@@
 Let $L_{i+1} = L_0 \cup L$.
-Let $L_\infty$ be the limit of this sequence (i.e., $\ell \in L_\infty$ if and only
+Let $L_\infty$ be the limit of this sequence (i.e., $\ell \in L_\infty$ if and only if
 $\ell \in L_i$ for some $i$),
@@
 $\conv( \varphi_{\con{f}}[ \vec t / \vec x ] )$ entails $\varphi_{\con{f}}[ \vec t / \vec x ]$.
-Thus, $\psi$ entails $\{ \varphi_{\con{f}}[ \vec t / \vec x ] \mid \con{f}( \vec t ) \in \terms( \psi ), f \in \sfundefs{\Sigma} \}$,
+Thus, $\psi$ entails $\{ \varphi_{\con{f}}[ \vec t / \vec x ] \mid \con{f}( \vec t ) \in \terms( \psi ), \con{f} \in \sfundefs{\Sigma} \}$,
 which by definition implies that $\psi$ is closed under function expansion with respect to $\DDD$.

### 2015-05-07 tex/main.tex
more tuning

@@
 Second, we show that if $\conv( \varphi )$ is satisfied by a
-$\extendsig{\Sigma}$-interpretation $\I$, then $\varphi$ is satisfied by a
-$\Sigma$-interpretation $\J$. Since $\varphi$ is in definitional form with
+$\extendsig{\Sigma}$-interpretation $\J$, then $\varphi$ is satisfied by a
+$\Sigma$-interpretation $\I$. Since $\varphi$ is in definitional form with
 respect to the functions defined by $\DDD$, it must be of the form
 $\DDD \wedge \varphi_0$. 
-%We argue that $\I$ satisfies a $\Sigma$-formula
+%We argue that $\J$ satisfies a $\Sigma$-formula
 %$\varphi'$ that is closed under function expansion with respect to $\DDD$, and
@@
 First, we define a sequence of $\Sigma$-literals sets $L_0 \subseteq L_1 \subseteq \cdots$
-such that $\I$ satisfies $L_i \cup \absconstraints( L_i )$ for $i = 0, 1, \ldots\vthinspace$.
-Since $\I$ satisfies $\conv( \varphi_0 )$,
+such that $\J$ satisfies $L_i \cup \absconstraints( L_i )$ for $i = 0, 1, \ldots\vthinspace$.
+Since $\J$ satisfies $\conv( \varphi_0 )$,
 by Lemma~\ref{lem:conv}, 
-$\I$ satisfies a set of literals $L \cup \absconstraints( L )$ where $L$ is a set of $\Sigma$-literals that entail $\varphi_0$.
+$\J$ satisfies a set of literals $L \cup \absconstraints( L )$ where $L$ is a set of $\Sigma$-literals that entail $\varphi_0$.
 Let $L_0 = L$.
@@
 where the function definition of $\con{f}$ in $\DDD$ is $\forallf{\con{f}} \vec x.\; \varphi_{\con{f}}$.
-Since $\I$ satisfies $\conv( \forallf{\con{f}} \vec x.\; \varphi_{\con{f}} )$ and $\absconstraints( L_i )$,
-we know that $\I$ also satisfies $\psi_i$.
+Since $\J$ satisfies $\conv( \forallf{\con{f}} \vec x.\; \varphi_{\con{f}} )$ and $\absconstraints( L_i )$,
+we know that $\J$ also satisfies $\psi_i$.
 Thus by Lemma~\ref{lem:conv},
-$\I$ satisfies a set of literals $L \cup \absconstraints( L )$ where $L$ is a set of $\Sigma$-literals that entail $\psi_i$.
+$\J$ satisfies a set of literals $L \cup \absconstraints( L )$ where $L$ is a set of $\Sigma$-literals that entail $\psi_i$.
 Let $L_{i+1} = L_0 \cup L$.
@@
 and let $\psi$ be the $\Sigma$-formula $\bigwedge L_\infty$.
-To show $\psi$ is closed under function expansion with respect to $\DDD$,
-first note that by construction $\psi$ entails $\psi_\infty$.
-For any function $\con{f}$ and terms $\vec t$, since $\varphi_{\con{f}}[ \vec t / \vec x ]$ does not contain function definitions,
+To show that $\psi$ is closed under function expansion with respect to $\DDD$,
+we first note that by construction $\psi$ entails $\psi_\infty$.
+For any function~$\con{f}$ and terms~$\vec t$, since $\varphi_{\con{f}}[ \vec t / \vec x ]$ does not contain function definitions,
 by Corollary~\ref{cor:conv},
@@
 Thus, $\psi$ entails $\{ \varphi_{\con{f}}[ \vec t / \vec x ] \mid \con{f}( \vec t ) \in \terms( \psi ), \con{f} \in \sfundefs{\Sigma} \}$,
-which by definition implies that $\psi$ is closed under function expansion with respect to $\DDD$.
+meaning that $\psi$ is closed under function expansion with respect to $\DDD$.
 Furthermore, $\psi$ entails $\varphi_0$ since $L_0 \subseteq L_\infty$.
 Since $\psi$ is a $T$-satisfiable formula that is closed under function expansion with respect to $\DDD$ and $\DDD$ is admissible, 
-by definition there exists a $\Sigma$-interpretation $\J$ satisfying $\psi \wedge \DDD$, which entails $\DDD \wedge \varphi_0$,
+by definition there exists a $\Sigma$-interpretation $\I$ satisfying $\psi \wedge \DDD$, which entails $\DDD \wedge \varphi_0$,
 which is $\varphi$.

### 2015-05-07 tex/main.tex
more

@@
 we first note that by construction $\psi$ entails $\psi_\infty$.
-For any function~$\con{f}$ and terms~$\vec t$, since $\varphi_{\con{f}}[ \vec t / \vec x ]$ does not contain function definitions,
+For any function symbol~$\con{f}$ and terms~$\vec t$, since $\varphi_{\con{f}}[ \vec t / \vec x ]$ does not contain function definitions,
 by Corollary~\ref{cor:conv},
@@
 by definition there exists a $\Sigma$-interpretation $\I$ satisfying $\psi \wedge \DDD$, which entails $\DDD \wedge \varphi_0$,
-which is $\varphi$.
-%\ajr{Sketchy referring to "limit" of sequence, if it can be infinite. TODO: clarify/restrict.}
+i.e., $\varphi$.
 \qed
@@
 Hence when $\DDD$ is admissible, this implies the existence of a model for $\varphi$.
-For clarity, we demonstrate the consequences of this theorem in the context of a concrete example.
+For clarity, we demonstrate the consequences of this theorem by means of an example.
 
@@
             \con{s}(\con{c}) \teq \con{c} + \con{s}( \con{c}-1 )
-            \wedge \exists b : {\fargtype{\con{s}}}.\;\, (\farg{\con{s}}( b ) \teq \con{c}-1) \bigr) \bigr)
+            \wedge \exists b : {\fargtype{\con{s}}}.\;\, \farg{\con{s}}( b ) \teq \con{c}-1 \bigr)
 \]
 The existential $\exists b : {\fargtype{\con{s}}}.\;\, (\farg{\con{s}}( b ) \teq \con{c}-1)$ in the above formula
-ensures that whenever $\I$ satisfies the set $L_1 = L_0 \cup \{ \neg \con{c} \leq 0, \con{s}(\con{c}) \teq \con{c} + \con{s}( \con{c}-1 ) \}$,
+ensures that whenever $\I$ satisfies the set $L_1 = L_0 \cup \{ \neg \con{c} \leq 0,\; \con{s}(\con{c}) \teq \con{c} + \con{s}( \con{c}-1 ) \}$,
 $\I$ satisfies $\absconstraints( L_1 )$ as well.
@@
 & \con{s}( \con{c} ) > 100 \wedge %\displaystyle
-  \bigwedge\nolimits_{\vthinspace i=0}^{\!n-1} \bigl( \neg (\con{c}-i \leq 0) \wedge \con{s}( \con{c}-i ) \teq (\con{c}-i) + \con{s}( (\con{c}-i)-1 ) \bigr) \\[-\jot]
+  \bigwedge\nolimits_{\vthinspace i=0}^{\!n-1} \bigl( \neg (\con{c}-i \leq 0) \wedge \con{s}( \con{c}-i ) \teq \con{c}-i + \con{s}( \con{c}-i-1 ) \bigr) \\[-\jot]
 {\wedge}\; & \con{c}-n \leq 0 \wedge \con{s}( \con{c}-n ) \teq 0
@@
 This formula is closed under function expansion
-since it entails $\varphi_\con{s}[\con{c}-i/x]$ for $i = 0, \ldots, n$,
-and contains only $\con{s}$ applications corresponding to $\con{s}( \con{c}-i )$ for $i = 0, \ldots, n$.
-Since $\varphi_\con{s}$ is an admissible definition,
-there is a $\Sigma$-interpretation satisfying $\psi \wedge \forallf{\con{s}} x.\;\, \varphi_\con{s}$
+since it entails $\varphi_\con{s}[(\con{c}-i)/x]$ for $i = 0, \ldots, n$,
+and it contains only $\con{s}$ applications corresponding to $\con{s}( \con{c}-i )$ for $i = 0, \ldots, n$.
+Since $\{\forallf{\con{s}} x.\;\, \varphi_\con{s}\}$ is admissible,
+there exists a $\Sigma$-interpretation satisfying $\psi \wedge \forallf{\con{s}} x.\;\, \varphi_\con{s}$
 which entails formula~(\ref{eq:ex-before}).\xend

### 2015-05-07 tex/main.tex
more

@@
 \]
-The existential $\exists b : {\fargtype{\con{s}}}.\;\, (\farg{\con{s}}( b ) \teq \con{c}-1)$ in the above formula
+The existential constraint in the above formula
 ensures that whenever $\I$ satisfies the set $L_1 = L_0 \cup \{ \neg \con{c} \leq 0,\; \con{s}(\con{c}) \teq \con{c} + \con{s}( \con{c}-1 ) \}$,
@@
 %
-\begin{equation} \label{eq:fun-closure}
-\begin{split}
+\[\begin{split}
 & \con{s}( \con{c} ) > 100 \wedge %\displaystyle
@@
 {\wedge}\; & \con{c}-n \leq 0 \wedge \con{s}( \con{c}-n ) \teq 0
-\end{split}
-\end{equation}
+\end{split}\]
 %

### 2015-05-07 tex/main.tex
more

@@
 This formula is unsatisfiable only if
-formula~(\ref{eq:ex-before}) is unsatisfiable since $\fargtype{f}$ can be
-interpreted as $\typeint$, and $\farg{f}$ can be interpreted as the identity
+formula~(\ref{eq:ex-before}) is unsatisfiable since $\fargtype{\con{s}}$ can be
+interpreted as the integers and $\farg{\con{s}}$ as the identity
 function. Conversely, we claim formula~(\ref{eq:ex-after}) is satisfiable

### 2015-05-07 tex/main.tex
more

@@
 %The latter ensures that $\I$ satisfies a formula that entails both $\con{s}( \con{c} ) > 100$ and $\varphi_{\con{s}}[\con{c}/x]$.
-Since $\I$ also satisfies both the translated function definition for $\con{s}$, i.e. the first conjunct of formula~(\ref{eq:ex-after}),
-and $\absconstraints( L_0 )$, it must also satisfy the formula $\psi_1$:
+Since $\I$ also satisfies both the translated function definition for $\con{s}$---i.e.,
+the first conjunct of formula~(\ref{eq:ex-after})---and $\absconstraints( L_0
+)$, it must also satisfy the formula $\psi_1$:
 \[
@@
 $\I$ satisfies $\absconstraints( L_1 )$ as well.
-Hence, by repeated applications of this reasoning, it follows that
+Hence, by repeated application of this reasoning, it follows that
 a model of formula~(\ref{eq:ex-after}) that interprets $\con{c}$ as $n$ must also satisfy the formula $\psi$:
@@
 
-In this section, we evaluate both the overall impact of the encoding introduced
-in the previous section and the performance of individual SMT techniques for
-benchmarks in the encoding.
-The experimental data is publicly available.%
+In this section, we evaluate both the overall impact of the translation
+introduced in the previous section and the performance of individual SMT
+techniques. The experimental data is publicly available.%
 \footnote{\url{http://lara.epfl.ch/~reynolds/SMT2015-recfun/}}
@@
 benchmarks from the Leon repository,\footnote{%Available at
-\url{https://github.com/epfl-lara/leon}} which were constructed from
+\url{https://github.com/epfl-lara/leon/}} which were constructed from
 verification conditions from simple Scala programs. These benchmarks also
-heavily involve recursively-defined functions over algebraic datatypes, 
-but cover a wide variety of theories, including bit vectors, arrays, and linear and
-nonlinear arithmetic. Each benchmark is in definitional form with respect to a
-set of terminating functions.
+heavily involve recursively defined functions over algebraic datatypes, 
+but cover a wide variety of theories, including bit vectors, arrays, and
+both linear and nonlinear arithmetic. Each benchmark is in definitional form
+with respect to a set of well-founded functions.
 %
@@
 %conjecture.
-A majority of these benchmarks are unsatisfiable, although a
-handful of benchmarks in the \leon set are satisfiable.
-
-For each of these 245 benchmarks, we considered (up to) three randomly selected
+Nearly all of these benchmarks are unsatisfiable, except for a
+handful of benchmarks in the \leon set.
+
+For each of these 245 benchmarks, we considered up to three randomly selected
 mutated forms of its conjecture $\psi$. In particular, we considered unique
@@
 and easy-to-find countermodels. In total, we considered 213 mutated forms of
-conjectures from \isa, and 427 mutated forms of conjectures from \leon. We will
+conjectures from \isa and 427 mutated forms of conjectures from \leon. We will
 call these sets \isam and \leonm respectively.
@@
 original 245). We considered these 885 benchmarks both before and after the
-translation $\conv$ as described in Section~\ref{sec:encoding}. For example,
-\isa contains 79 original benchmarks $\varphi$ and 79 translated benchmarks
-$\conv(\varphi)$.
+translation~$\conv$. % as described in Section~\ref{sec:encoding}.
+For example, \isa contains 79 original benchmarks $\varphi$ and 79 translated
+benchmarks $\conv(\varphi)$.
 
@@
 The configuration \cvcf runs heuristic instantiation and the finite model
-finding procedure described in~\cite{ReyEtAl-1-RR-13,reynolds-et-al-2013}. The
-last configuration \cvci incorporates techniques for automating inductive
+finding procedure due to Reynolds et al.\ \cite{ReyEtAl-1-RR-13,reynolds-et-al-2013}. The
+last configuration, \cvci, incorporates techniques for automating inductive
 reasoning in SMT~\cite{reynolds-kuncak-2015}.
@@
 This is consistent with our expectations and Theorem~\ref{thm:equi}, 
-since these benchmarks contained only sets of function definitions that we expected were admissible.
+since these benchmarks contained only sets of function definitions
+that are supposed to be well-founded.
 %thereby confirming Theorem~\ref{thm:equi}.
@@
 Figure~\ref{fig:sat} shows that for untranslated benchmarks (the ``$\varphi$''
-columns), the number of \emph{sat} responses is very low, across all
+columns), the number of \emph{sat} responses is very low across all
 configurations. This confirms the shortcomings of existing SMT techniques for
-finding counterexamples for benchmarks containing recursively-defined functions.
+finding counterexamples for benchmarks containing recursively defined functions.
 %
@@
 We conclude that the translation $\conv$ enables SMT
-solvers to find counterexamples for conjectures involving recursively-defined functions
+solvers to find counterexamples for conjectures involving recursively defined functions
 whose definitions are admissible.
@@
 clear winner, finding 318 total benchmarks to be unsatisfiable. Moreover, the
-translation $\conv$ helps \ziii for \emph{unsat} responses as well, as we
-see \ziii solves a total of 209 whereas it solves only 187 before the
+translation $\conv$ helps \ziii for \emph{unsat} responses as well:
+\ziii solves a total of 209, whereas it solves only 187 before the
 translation.
 %
-This number is 2 fewer than \cvcd, which finds 211 benchmarks unsatisfiable. Thus, in
-most cases the translation $\conv$ does not significantly degrade performance
+In contrast, \cvcd finds 211 benchmarks unsatisfiable with the translation.
+Thus, in most cases the translation $\conv$ does not significantly degrade performance
 for unsatisfiable benchmarks, and in some cases may actually aid the solver for
-determining unsatisfiability. The exception here is \cvci, whose inductive
-arguments are guided by the types of quantified variables.
+determining unsatisfiability. The exception is \cvci, whose inductive
+arguments are guided by the types of the quantified variables.
 
@@
 (version~1.5 prerelease).
-Function definitions $\forallf{\con{f}} \vec x.\; \varphi$ can be written using the $\definefunreccmd$ command,
+Function definitions $\forallf{\con{f}} \vec x.\; \varphi$ can be written using the $\definefunreccmd$ command
 from the upcoming SMT-LIB 2.5 standard \cite{smtlib25}.
@@
 Accordingly, the solver will output the approximation of the interpretation it
-used for recursive function definitions. For our example, it will produce the following output:
+used for recursive function definitions. For our example, it outputs
 %\ajr{prefer having the full output like this here?}
@@
 functions provided by the user are admissible.
-Admissibility must be checked separately \relax{by the user}---for example, using a
+Admissibility must be checked separately \relax{by the user}, perhaps using a
 syntactic criterion or a termination prover.
 If this criterion is not met, \cvc may answer \emph{sat} on
-an unsatisfiable problem. For example, if we add the inconsistent definition
+an unsatisfiable problem. Indeed, if we add the inconsistent definition
 %
@@
 %
-to the above problem, \cvc will answer \emph{sat}
-if \texttt{--fmf-fun} is specified.
+to the above problem and run \cvc with the \texttt{--fmf-fun} option,
+it wrongly answers \emph{sat}.
 

### 2015-05-07 tex/main.tex
more

@@
 
-Most of the above tool cannot cope with algebraic datatypes or other infinite
+Most of the above tools cannot cope with algebraic datatypes or other infinite
 types. Kuncak and Jackson \cite{kuncak-jackson-2005} presented an idiom for
 encoding datatypes and recursive functions in Alloy, by approximating datatypes
-by finite subterm-closed substructures. The approach produces sound (fragments
+by finite subterm-closed substructures. The approach finds sound (fragments
 of) models for formulas in the \relax{existential--bounded-universal} fragment
@@
 
-The higher-order model finder Nitpick for the Isabelle/HOL proof assistant uses
-another variant of Kuncak and Jackson's approach inside a Kleene-style
-three-valued logic, inspired by abstract interpretation. It was also the first
-tool of its kind to support corecursion and coalgebraic datatypes
+The higher-order model finder Nitpick for the Isabelle/HOL proof assistant
+relies on another variant of Kuncak and Jackson's approach inside a
+Kleene-style three-valued logic, inspired by abstract interpretation. It was
+also the first tool of its kind to support corecursion and coalgebraic datatypes
 \cite{blanchette-2013-relational}. The three-valued logic approach extends each
@@
 on Kodkod's relational logic, but our initial experiments with \cvc convinced
-us that it is more efficient to avoid unknown values by adding constraints.
+us that it is more efficient to avoid unknown values by adding existential
+constraints.
 
@@
 quantifiers altogether by unfolding recursive definitions up to a certain
-depth, which is increased iteratively as necessary.
+depth, which is increased on a per-need basis.
 The approach described in this paper works in an analagous manner, 
 where instead the SMT solver is invoked only once 
-and where quantifier instantiation amounts to function unfolding. It is worth
-investigating how existing approaches for algorithms based on function
+and where quantifier instantiation amounts to function unfolding. It would be
+worth investigating how existing approaches for algorithms based on function
 unfolding can inform approaches for dedicated quantifier instantiation
@@
 Model finding is concerned with satisfying arbitrary logical constraints. Some
-tools are optimized to work with problems that correspond to total functional
+tools are tailored for problems that correspond to total functional
 programs. The QuickCheck tool for Haskell \cite{claessen-hughes-2000} is an
@@
 on random testing. Bounded exhaustive testing \cite{runciman-et-al-2008} and
-narrowing \cite{lindblad-2008-testing} are other popular strategies. These tool
-are often much faster than model finders, but they typically cannot cope with
-underspecification and nonexecutable functions.
+narrowing \cite{lindblad-2008-testing} are other successful strategies. These
+tool are often much faster than model finders, but they typically cannot cope
+with underspecification and nonexecutable functions.
 
@@
 \cite{reynolds-blanchette-2015-codata}. We also plan to integrate \cvc as a
-counterexample generator in proof assistants. Finally, further work could also
+counterexample generator in proof assistants. Finally, future work could also
 include identifying further sufficient conditions for admissibility, thereby

### 2015-05-07 tex/main.tex
more

@@
 \end{equation}
-This state of affairs is highly unsatisfactory, given the frequence of
+This state of affairs is highly unsatisfactory, given the frequency of
 recursive definitions in practice and the impending addition of a dedicated
@@
 depth, which is increased on a per-need basis.
-The approach described in this paper works in an analagous manner, 
+The approach described in this paper works in an analogous manner, 
 where instead the SMT solver is invoked only once 

### 2015-05-07 tex/main.tex
moved to EasyChair

@@
-\documentclass[runningheads,a4paper]{llncs}
-\usepackage[T1]{fontenc}
-\usepackage{amsmath}
-\usepackage{amssymb}
-\usepackage[scaled=.82]{beramono}
-\usepackage[scaled=.86]{helvet}
-\usepackage{mathptmx}
-\usepackage{listings}
-\usepackage{colonequals}
-%\usepackage{mathpartir}
-\usepackage{xspace}
-\setcounter{tocdepth}{3}
-\usepackage{graphicx}
-\usepackage{stmaryrd}
-\usepackage{cite}
-\usepackage[
+\PassOptionsToClass{11pt,withtimes}{easychair}
+\PassOptionsToPackage{hyphens}{url}
+\PassOptionsToPackage{scaled=.86}{helvet}
+\PassOptionsToPackage{
    a4paper,
@@
    bookmarksopenlevel=2,
-   bookmarksopen]{hyperref}
-\usepackage{framed}
+   bookmarksopen}{hyperref}
+\documentclass{easychair}
+
+%\documentclass[runningheads,a4paper]{llncs}
+\usepackage[T1]{fontenc}
+\usepackage{amsmath}
+\usepackage{amssymb}
+\usepackage[scaled=.82]{beramono}
+\usepackage{mathptmx}
+\usepackage{listings}
+\usepackage{colonequals}
+%\usepackage{mathpartir}
+\usepackage{xspace}
+\setcounter{tocdepth}{3}
+\usepackage{graphicx}
+\usepackage{stmaryrd}
+\usepackage{cite}
+%\usepackage{framed}
 
@@
 
-\spnewtheorem*{proofsketch}{Proof sketch}{\itshape}{\rmfamily}
+\def\proofname{Proof sketch}
+
+\theoremstyle{plain}
+\newtheorem{theorem}{Theorem}%[section]
+\newtheorem{lemma}{Lemma}%[section]
+\newtheorem{corollary}{Corollary}%[section]
+
+\theoremstyle{definition}
+\newtheorem{definition}{Definition}%[section]
+\newtheorem{example}{Example}%[section]
+
+%\spnewtheorem*{proof}{Proof sketch}{\itshape}{\rmfamily}
 
@@
 \newcommand\xend{{\hfill$\scriptstyle\blacksquare$}}
-\renewcommand\qed{{\hfill$\squareforqed$}}
+%\renewcommand\qed{{\hfill$\squareforqed$}}
+\let\qed=\relax
 
@@
 }
+\titlerunning {Model Finding for Recursive Functions in SMT}
 
@@
 \end{theorem}
-\begin{proofsketch}
+\begin{proof}
 Let $\psi$ be a satisfiable formula closed under function expansion with
@@
 \qed
-\end{proofsketch}
+\end{proof}
 
@@
 \end{lemma}
-\begin{proofsketch}
+\begin{proof}
 By definition of $\conv$ and case analysis on the return values of $\conv_0$.
-\end{proofsketch}
+\end{proof}
 
@@
 \end{theorem}
-\begin{proofsketch}
+\begin{proof}
 First, we show that if $\varphi$ is satisfied by $\Sigma$-interpretation $\I$, then $\conv( \varphi )$ is satisfied by a $\extendsig{\Sigma}$-interpretation $\J$.

### 2015-05-07 tex/main.tex
more

@@
 constraints
-$\exists a : \fargtype{\con{f}}.\;\, (\vecfarg{\con{f}}( a ) \teq \vec t)$; these
+$\exists a : \fargtype{\con{f}}.\;\, \vecfarg{\con{f}}( a ) \teq \vec t$; these
 ensure that the restricted definition for $\con{f}$ covers all relevant tuples
@@
 
-If $t$ is an application of a logical predicate $\boolop
-\in \{{\neg}, {\wedge}, {\vee}, {\teq}, {\lite}\}$,
+If $t$ is an application of a predicate $\boolop$,
+%\in \{{\neg}, {\wedge}, {\vee}, {\teq}, {\lite}\}$,
 $\conv_0$ calls itself recursively on the arguments $t_i$ and polarity $\pol(

### 2015-05-07 tex/main.tex
more

@@
 
-If $t$ is an application of a predicate $\boolop$,
-%\in \{{\neg}, {\wedge}, {\vee}, {\teq}, {\lite}\}$,
+If $t$ is an application of a predicate $\boolop$, including the built-in
+operators ${\neg}$, ${\wedge}$, ${\vee}$, ${\teq}$, and ${\lite}$,
 $\conv_0$ calls itself recursively on the arguments $t_i$ and polarity $\pol(

### 2015-05-07 tex/main.tex
more

@@
 \usepackage{cite}
+\usepackage{amsthm}
 %\usepackage{framed}
@@
 
-\def\proofname{Proof sketch}
-
 \theoremstyle{plain}
 \newtheorem{theorem}{Theorem}%[section]
-\newtheorem{lemma}{Lemma}%[section]
-\newtheorem{corollary}{Corollary}%[section]
+\newtheorem{lemma}[theorem]{Lemma}%[section]
+\newtheorem{corollary}[theorem]{Corollary}%[section]
 
@@
 \newtheorem{example}{Example}%[section]
+
+\theoremstyle{remark}
+\newtheorem*{proofsketch}{Proof sketch}
 
@@
 
+\def\squareforqed{\hbox{\rlap{$\sqcap$}$\sqcup$}}
+
 \newcommand\xend{{\hfill$\scriptstyle\blacksquare$}}
-%\renewcommand\qed{{\hfill$\squareforqed$}}
-\let\qed=\relax
+\renewcommand\qed{{\hfill$\squareforqed$}}
 
@@
 \end{theorem}
-\begin{proof}
+\begin{proofsketch}
 Let $\psi$ be a satisfiable formula closed under function expansion with
@@
 \qed
-\end{proof}
+\end{proofsketch}
 
@@
 \end{lemma}
-\begin{proof}
+\begin{proofsketch}
 By definition of $\conv$ and case analysis on the return values of $\conv_0$.
-\end{proof}
+\end{proofsketch}
 
@@
 \end{theorem}
-\begin{proof}
+\begin{proofsketch}
 First, we show that if $\varphi$ is satisfied by $\Sigma$-interpretation $\I$, then $\conv( \varphi )$ is satisfied by a $\extendsig{\Sigma}$-interpretation $\J$.

### 2015-05-07 tex/main.tex
more

@@
 \begin{abstract}
+\noindent
 SMT solvers have recently been extended with techniques for finding models
@@
 (Section~\ref{sec:evaluation}). The approach is implemented as a preprocessor
-in \cvc (Section~\ref{sec:front-end}). When invoked with a
+in \cvc (Section~\ref{sec:front-end}). When invoked with a specific
 command-line option, the solver assumes that functions introduced using the
@@
 \cite{blanchette-et-al-2015-fouco}. Theorem~\ref{thm:adm} can be extended to
-cover such specifications, based on the observation that an infinite unfolding
-of a corecursive definition yields a uniquely specified infinite object.
+cover such specifications, based on the observation that infinite unfolding
+a corecursive definition computes a unique infinite object.
 
 Beyond totality, an admissible set can contain underspecified functions
-such as $\forallf{\con{f}} x : \typeint.\;\, \con{f}( x )
+such as $\forallf{\con{f}} x : \typeint.\allowbreak\;\, \con{f}( x )
 \teq \con{f}( x - 1 )$ or $\forallf{\con{g}} x.\allowbreak\;\, \con{g}( x
@@
 %%% TYPESETTING: approximated, to get centering
-\newcommand{\itemx}{\itemindent5em\item}
+\newcommand{\itemx}{\itemindent6.35em\item}
 
@@
     \begin{itemize}
-      \itemx[] $\mathsf{let}$ $( t'_i,\, \chi_i ) = \conv_0( t_i,\, \pol( \boolop,\, i,\, p ) )$ $\mathsf{for}$ $i = 1, \ldots, n$ $\mathsf{in}$%\kern5em$\mid$
+      \itemx[] $\mathsf{let}$ $( t'_i,\, \chi_i ) = \conv_0( t_i,\, \pol( \boolop,\, i,\, p ) )$ $\mathsf{for}$ $i = 1, \ldots, n$ $\mathsf{in}$%\kern6.35em$\mid$
       \itemx[] $\mathsf{let}$ $\chi = \chi_1 \wedge \cdots \wedge \chi_n$ $\mathsf{in}$
@@
 $\conv( \varphi )$ $=$ $\mathsf{let}$ $( \varphi',\, \_ ) = \conv_0( \varphi,\, \ppos )$ $\mathsf{in}$ $\varphi'$
-\\[-1.5\baselineskip] %% TYPESETTING
+%\\[-1.25\baselineskip] %% TYPESETTING
 \end{enumerate}

### 2015-05-07 tex/main.tex
more

@@
 \newtheorem{definition}{Definition}%[section]
-\newtheorem{example}{Example}%[section]
 
@@
 \newtheorem*{proofsketch}{Proof sketch}
+\newtheorem{example}{Example}%[section]
 
@@
 \boolop, i, p )$ returns the polarity of the $i$\vvthinspace{th} argument of an
-application of~$\boolop$ having polarity $p$, where $\boolop$ is an interpreted
-predicate.
+application of~$\boolop$ having polarity $p$ if $\boolop$ is an interpreted
+predicate or $\pnone$ otherwise.
 
@@
 $\{ \forallf{\con{s}} x.\;\, \varphi_\con{s} \}$ is admissible,
-where $\varphi_\con{s}$ is the formula $\lite\bigl( x \leq 0, \con{s}(x) \teq 0, \con{s}( x ) \teq x + \con{s}( x-1 ) \bigr)$.
+where $\varphi_\con{s}$ is the formula $\lite\bigl( x \leq 0,\; \con{s}(x) \teq 0,\; \con{s}( x ) \teq x + \con{s}( x-1 ) \bigr)$.
 %\ct{
@@
 introduced in the previous section and the performance of individual SMT
-techniques. The experimental data is publicly available.%
+techniques. The experimental data is publicly available online.%
 \footnote{\url{http://lara.epfl.ch/~reynolds/SMT2015-recfun/}}

### 2015-05-07 tex/main.tex
more

@@
 We presented a translation scheme that makes more SMT problems amenable to
-finite model findings in SMT. In conjunction with the finite model finding
-techniques introduced by Reynolds et al.\
-\cite{ReyEtAl-1-RR-13,reynolds-et-al-2013}, it makes it possible to
+finite model findings in SMT. In conjunction with finite model finding
+techniques,
+%introduced by Reynolds et al.\ \cite{ReyEtAl-1-RR-13,reynolds-et-al-2013},
+%%% Great work, but let's not cite it more than 4 times if it can be avoided. --jb
+it makes it possible to
 find models of quantified formulas over infinite types, such as integers
@@
 include identifying further sufficient conditions for admissibility, thereby
-enlarging the scope of the scheme presented here.
+enlarging the applicability of the translation scheme presented here.
 

### 2015-05-07 tex/main.tex
more

@@
 %
-\[(\forallf{\con{f}_1} \vec x_1.\; \varphi_1) \wedge \cdots \wedge
-(\forallf{\con{f}_n} \vec x_n.\; \varphi_n) \wedge \psi\]
+$(\forallf{\con{f}_1} \vec x_1.\; \varphi_1) \wedge \cdots \wedge
+(\forallf{\con{f}_n} \vec x_n.\; \varphi_n) \wedge \psi$,
 %
@@
 the extended signature $\extendsig{\Sigma}$ contains
-\begin{itemize}
-\item an uninterpreted \emph{abstract type} $\fargtype{\con{f}}$;
-\item $n$ uninterpreted \emph{concretization functions} $\fargx{\con{f}}{1} : \fargtype{\con{f}} \rightarrow \tau_1$, \ldots, $\fargx{\con{f}}{n} : \fargtype{\con{f}} \rightarrow \tau_n$.
-\end{itemize}
+%\begin{itemize}
+%\item
+an uninterpreted \emph{abstract type} $\fargtype{\con{f}}$ and
+%\item
+$n$ uninterpreted \emph{concretization functions} $\fargx{\con{f}}{1} : \fargtype{\con{f}} \rightarrow \tau_1$, \ldots, $\fargx{\con{f}}{n} : \fargtype{\con{f}} \rightarrow \tau_n$.
+%\end{itemize}
 %The interpretation of uninterpreted type $\fargtype{\con{f}}$ will denote the elements (tuples) on which the function $\con{f}$ is applied.
@@
 
-If $t$ is an unannotated quantified formula, $\conv_0$ recursively calls itself
+If $t$ is an unannotated quantified formula, $\conv_0$ calls itself
 on the body with the same polarity. A quantifier is prefixed to the
@@
 One model $\I$ for formula~(\ref{eq:ex-after}) in the above example interprets
-$\fargtype{\con{s}}$ as a finite set of elements $\{ u_0, \ldots, u_{14} \}$,
+$\fargtype{\con{s}}$ as a finite set $\{ u_0, \ldots, u_{14} \}$,
 $\farg{\con{s}}$ as a finite map $u_i \mapsto i$ for $i = 0, \ldots, 14$,
@@
 %
-%% TYPESETTING: \mathbin and to make it fit nicely
-\begin{equation} \label{eq:approx-interp}
-\lambda x : \typeint.\;\, \lite( x \mathbin{\teq} 0,\, 0,\,
-  \lite( x \mathbin{\teq} 1,\, 1,\,
-    \lite( x \mathbin{\teq} 2,\, 3,\,
-      \lite(\ldots,\; \lite( x \mathbin{\teq} 13,\, 91,\, 105 )\negvthinspace \ldots ))))
-\end{equation}
+%\begin{equation} \label{eq:approx-interp}
+\[\lambda x : \typeint.\;\, \lite( x \teq 0,\; 0,\;
+  \lite( x \teq 1,\; 1,\;
+    \lite( x \teq 2,\; 3,\;
+      \lite(\ldots,\; \lite( x \teq 13,\, 91,\, 105 )\negvthinspace \ldots ))))\]
+%\end{equation}
 %
@@
 \begin{proofsketch}
-First, we show that if $\varphi$ is satisfied by $\Sigma$-interpretation $\I$, then $\conv( \varphi )$ is satisfied by a $\extendsig{\Sigma}$-interpretation $\J$.
+First, we show that if $\varphi$ is satisfied by $\Sigma$-interpretation $\I$, then $\conv( \varphi )$ is satisfied by an $\extendsig{\Sigma}$-interpretation $\J$.
 Let $\J$ be the $\extendsig{\Sigma}$-interpretation that interprets all types $\tau \in \stypes{\Sigma}$ as $\tau^\I$,
@@
 Hence when $\DDD$ is admissible, this implies the existence of a model for $\varphi$.
-For clarity, we demonstrate the consequences of this theorem by means of an example.
+%For clarity, we demonstrate the consequences of this theorem by means of an example.
 
@@
 %
-\[\begin{split}
-& \con{s}( \con{c} ) > 100 \wedge %\displaystyle
-  \bigwedge\nolimits_{\vthinspace i=0}^{\!n-1} \bigl( \neg (\con{c}-i \leq 0) \wedge \con{s}( \con{c}-i ) \teq \con{c}-i + \con{s}( \con{c}-i-1 ) \bigr) \\[-\jot]
-{\wedge}\; & \con{c}-n \leq 0 \wedge \con{s}( \con{c}-n ) \teq 0
-\end{split}\]
+\[\con{s}( \con{c} ) > 100 \wedge %\displaystyle
+  \bigwedge\nolimits_{\vthinspace i=0}^{\!n-1} \bigl( \neg (\con{c}-i \leq 0) \wedge \con{s}( \con{c}-i ) \teq \con{c}-i + \con{s}( \con{c}-i-1 ) \bigr)
+\wedge \con{c}-n \leq 0 \wedge \con{s}( \con{c}-n ) \teq 0\]
 %
@@
 introduced in the previous section and the performance of individual SMT
-techniques. The experimental data is publicly available online.%
-\footnote{\url{http://lara.epfl.ch/~reynolds/SMT2015-recfun/}}
-
+techniques.
+%
 We gathered a set of 245 benchmarks from two sources, which we will refer to as
@@
 conjectures from \isa and 427 mutated forms of conjectures from \leon. We will
-call these sets \isam and \leonm respectively.
-
-In total, our benchmark set consisted of 885 benchmarks (640 mutants plus the
+call these sets \isam and \leonm, respectively.
+
+Thus, our benchmark set consist of 885 benchmarks (640 mutants plus the
 original 245). We considered these 885 benchmarks both before and after the
@@
 For solvers, we considered the SMT solver \ziii~\cite{de-moura-bjoerner-2008},
-which runs both heuristic methods for quantifier instantiation~\cite{MouraBjoerner07},
-as well as methods for finding models for quantified formulas~\cite{GeDeM-CAV-09}.
+which runs both heuristic methods for quantifier instantiation~\cite{MouraBjoerner07}
+and methods for finding models for quantified formulas~\cite{GeDeM-CAV-09}.
 We also considered three configurations of \cvc~\cite{barrett-et-al-2011} which
@@
 but does not include techniques for finding models.
-The configuration \cvcf runs heuristic instantiation and the finite model
-finding procedure due to Reynolds et al.\ \cite{ReyEtAl-1-RR-13,reynolds-et-al-2013}. The
-last configuration, \cvci, incorporates techniques for automating inductive
+\cvcf runs heuristic instantiation and the finite model
+finding procedure due to Reynolds et al.\ \cite{ReyEtAl-1-RR-13,reynolds-et-al-2013}.
+\cvci incorporates techniques for automating inductive
 reasoning in SMT~\cite{reynolds-kuncak-2015}.
@@
 
-The results are shown in Figures \ref{fig:sat} and \ref{fig:unsat}.
-They are divided into benchmarks triggering \emph{unsat} and \emph{sat}
+The results are summarized in Figures \ref{fig:sat} and \ref{fig:unsat}.
+The benchmarks and more detailed results are available online.%
+\footnote{\url{http://lara.epfl.ch/~reynolds/SMT2015-recfun/}}
+The figures are divided into benchmarks triggering \emph{unsat} and \emph{sat}
 responses and further into benchmarks before and after the translation $\conv$.
@@
 function definitions.
-For the convenience of end users, it
-has been implemented in the latest development version of \cvc %~\cite{ReyEtAl-1-RR-13}
+For convenience, it
+has been implemented in the development version of \cvc %~\cite{ReyEtAl-1-RR-13}
 (version~1.5 prerelease).
@@
 For example, formula~(\ref{eq:ex-before}) from Example~\ref{ex:translation}
-can be specified as
+can be specified as %follows:
 %
@@
 %\small -- no small, unless you convince me that these mixtures of font sizes
-% are aesthetic and make sense --jb
+% are aesthetic and serve a purpose --jb
 \begin{verbatim}
@@
 Accordingly, the solver will output the approximation of the interpretation it
-used for recursive function definitions. For our example, it outputs
+used for recursive function definitions. For the example above, it outputs
 %\ajr{prefer having the full output like this here?}

### 2015-05-07 tex/main.tex
more

@@
 
-In this paper, we present a method for translating formulas involving recursice function
+In this paper, we present a method for translating formulas involving recursive function
 definitions into formulas where finite model finding techniques can be applied.
@@
 %\jb{Alloy has no notion of datatypes, so it's pretty useless. Nitpick implements
-%datatypes and a more FOL-like logic (actually, HOL) on top of Alloy's backend,
+%datatypes and a more FOL-like logic (actually, HOL) on top of Alloy's back-end,
 %using the Kuncak \& Jackson idiom for Alloy, so that's as close as we can get

### 2015-05-08 tex/main.tex
Minor

@@
 
-%\newcommand{\pnone}{\con{none}}
-%\newcommand{\ppos}{\con{pos}}
-%\newcommand{\pneg}{\con{neg}}
+\newcommand{\pnone}{\con{none}}
+\newcommand{\ppos}{\con{pos}}
+\newcommand{\pneg}{\con{neg}}
 %%% @ANDY: as usual, revert if you don't like
-\newcommand{\pnone}{{\pm}}
-\newcommand{\ppos}{{+}}
-\newcommand{\pneg}{{-}}
+% too confusing, too many different meanings for + and -, with below the latter is a theory symbol, a polarity, and a negation of polarities :)
+%\newcommand{\pnone}{{\pm}}
+%\newcommand{\ppos}{{+}}
+%\newcommand{\pneg}{{-}}
 \newcommand{\pol}{\con{pol}}
@@
 to translate and a \relax{polarity}~$p$, which is either $\ppos$, $\pneg$, or
-$\pnone$ (none). \vthinspace$\conv_0$ returns a pair $( t'\negvthinspace,\, \chi )$, where $t'$ is a term of
+$\pnone$. \vthinspace$\conv_0$ returns a pair $( t'\negvthinspace,\, \chi )$, where $t'$ is a term of
 type $\tau$ and $\chi$ is an additional constraint.
@@
 \end{equation*}
-The negation $\pnegate p$ of a polarity $p$ is $\pneg$ if $p$ is $\ppos$, $\ppos$
+where the negation $\pnegate p$ of a polarity $p$ is $\pneg$ if $p$ is $\ppos$, $\ppos$
 if $p$ is $\pneg$, and $\pnone$ if $p$ is $\pnone$.
-
 If $t$ is positive, $\conv_0$ adds the constraint $\chi$ derived from
@@
 $\chi$ must be $\ltrue$.
-
 If $t$ is an unannotated quantified formula, $\conv_0$ calls itself
@@
 quantified formula and constraint returned by the recursive call.
-
 Otherwise, $t$ is either an application of an uninterpreted predicate or a term
@@
 
-In the following, for a set of ground literals $L$, 
+In the following, we sketch the correctness of our translation $\conv$.
+For a set of ground literals $L$, 
 we write $\absconstraints( L )$ to denote the set of constraints that
@@
 \begin{lemma}\label{lem:conv}
-Let $\varphi$ be a formula not containing function definitions,
+Let $\psi$ be a formula not containing function definitions,
 and let $\I$ be an $\extendsig{\Sigma}$-interpretation.
-Then $\I$ satisfies $\conv( \varphi )$ if and only if
-$\I$ satisfies $L \cup \absconstraints( L )$, where $L$ is a set of ground $\Sigma$-literals that entail $\varphi$.
+Then $\I$ satisfies $\conv( \psi )$ if and only if
+$\I$ satisfies $L \cup \absconstraints( L )$, where $L$ is a set of ground $\Sigma$-literals that entail $\psi$.
 %There exists a set of ground $\Sigma$-literals $L$ such that $L$ entails $\varphi$, 
@@
 By definition of $\conv$ and case analysis on the return values of $\conv_0$.
+\qed
 \end{proofsketch}
@@
 \begin{cor}\label{cor:conv}
-If $\varphi$ is a formula not containing function definitions, then $\extendsig{\Sigma} = \Sigma$
-and $\conv( \varphi )$ entails $\varphi$.
+If $\psi$ is a formula not containing function definitions, then $\conv( \psi )$ entails $\psi$.
 \end{cor}
@@
 First, $\conv( \varphi )$ cannot be unsatisfiable when $\varphi$ is satisfiable
-since any model of $\varphi$ can be extended in a straightforward way to satisfy $\conv( \varphi )$.
+since any $\Sigma$-interpretation that satisfies $\varphi$ can be extended in a straightforward way to 
+a $\extendsig{\Sigma}$-interpretation that satisfies $\conv( \varphi )$.
 Second, if a model is found for $\conv( \varphi )$,
@@
 %any model of $\varphi_{\con{s}}[t/x]$ must also satisfy $\varphi_{\con{s}}[t-1/x]$ when $t > 0$. 
-It is easy to see that any interpretation $\I$ satisfying formula~(\ref{eq:ex-after}) satisfies $L_0 \cup \absconstraints( L_0 )$,
+Clearly any interpretation $\I$ satisfying formula~(\ref{eq:ex-after}) satisfies $L_0 \cup \absconstraints( L_0 )$,
 where $L_0 = \{ \con{s}( \con{c} ) > 100 \}$, 
@@
 and where quantifier instantiation amounts to function unfolding. It would be
-worth investigating how existing approaches for algorithms based on function
+worth investigating how existing approaches for function
 unfolding can inform approaches for dedicated quantifier instantiation

### 2015-05-08 tex/body.tex
renamed main

--- 2015-05-08 tex/main.tex
@@
 \usepackage{booktabs}
+\usepackage{version}
 %\usepackage{float}
@@
 \usepackage{program}
+
+\excludeversion{longv}
+\includeversion{shortv}
 
@@
 \end{theorem}
+\begin{longv}
 \begin{proofsketch}
@@
 \end{proofsketch}
+\end{longv}
 
@@
 \end{lemma}
+\begin{longv}
 \begin{proofsketch}
@@
 \cvc adds the annotated quantified formula
+\begin{longv}
 $$\forallf{\con{s}} x.\;\, \con{s}( x ) \teq \lite\bigl( x \leq 0,\; 0,\; \con{s}( x-1 )\bigr)$$
@@
 $$\forallf{\con{s}} x.\;\, \lite\bigl( x \leq 0,\; \con{s}( x ) \teq 0,\; \con{s}( x ) \teq \con{s}( x-1 ) \bigr)$$
+\end{longv}
+\begin{shortv}
+$\forallf{\con{s}} x.\;\, \con{s}( x ) \teq \lite\bigl( x \leq 0,\; 0,\; \con{s}( x-1 )\bigr)$
+to its list of assertions,
+which after rewriting becomes
+$\forallf{\con{s}} x.\;\, \lite\bigl( x \leq 0,\; \con{s}( x ) \teq 0,\; \con{s}( x ) \teq \con{s}( x-1 ) \bigr)$
+\end{shortv}
 If \cvc's finite model finding mode for recursive functions is enabled (using
@@
 Accordingly, the solver will output the approximation of the interpretation it
-used for recursive function definitions. For the example above, it outputs
+used for recursive function definitions. 
+\begin{longv}
+For the example above, it outputs
 %\ajr{prefer having the full output like this here?}
@@
 \end{quote}
+\end{longv}
+\begin{shortv}
+\ajr{todo}
+\end{shortv}
 

### 2015-05-08 tex/body.tex
setup with multi files

@@
-\PassOptionsToClass{11pt,withtimes}{easychair}
-\PassOptionsToPackage{hyphens}{url}
-\PassOptionsToPackage{scaled=.86}{helvet}
-\PassOptionsToPackage{
-   a4paper,
-   pdftex,
-   pdftitle={Model Finding for Recursive Functions in SMT},
-   pdfauthor={Andrew Reynolds, Jasmin Christian Blanchette, and Cesare Tinelli},
-   pdfkeywords={},
-   pdfborder={0 0 0},
-   draft=false,
-   bookmarksnumbered,
-   bookmarks,
-   bookmarksdepth=2,
-   bookmarksopenlevel=2,
-   bookmarksopen}{hyperref}
-\documentclass{easychair}
-
 %\documentclass[runningheads,a4paper]{llncs}
@@
 \usepackage{program}
-
-\excludeversion{longv}
-\includeversion{shortv}
 

### 2015-05-08 tex/prelude.tex
setup with multi files

--- 2015-05-07 tex/main.tex
@@
 \documentclass{easychair}
-
-%\documentclass[runningheads,a4paper]{llncs}
-\usepackage[T1]{fontenc}
-\usepackage{amsmath}
-\usepackage{amssymb}
-\usepackage[scaled=.82]{beramono}
-\usepackage{mathptmx}
-\usepackage{listings}
-\usepackage{colonequals}
-%\usepackage{mathpartir}
-\usepackage{xspace}
-\setcounter{tocdepth}{3}
-\usepackage{graphicx}
-\usepackage{stmaryrd}
-\usepackage{cite}
-\usepackage{amsthm}
-%\usepackage{framed}
-
-% http://mirror.switch.ch/ftp/mirror/tex/macros/latex/contrib/xcolor/xcolor.pdf
-\usepackage{xcolor}
-
-%\usepackage{amsthm}
-\usepackage{amsmath}    % need for subequations
-\usepackage{booktabs}
-%\usepackage{float}
-%\usepackage{fullpage}
-%\floatstyle{ruled}
-%\newfloat{alg}{ht}{loa}
-%\floatname{alg}{Algorithm}
-
-\theoremstyle{plain}
-\newtheorem{theorem}{Theorem}%[section]
-\newtheorem{lemma}[theorem]{Lemma}%[section]
-\newtheorem{corollary}[theorem]{Corollary}%[section]
-
-\theoremstyle{definition}
-\newtheorem{definition}{Definition}%[section]
-
-\theoremstyle{remark}
-\newtheorem*{proofsketch}{Proof sketch}
-\newtheorem{example}{Example}%[section]
-
-%\spnewtheorem*{proof}{Proof sketch}{\itshape}{\rmfamily}
-
-\usepackage{url}
-\urldef{\mailsa}\path|firstname.lastname@epfl.ch|
-\newcommand{\keywords}[1]{\par\addvspace\baselineskip
-\noindent\keywordname\enspace\ignorespaces#1}
-
-\newcommand\cvcd{{\cvc}d\xspace}
-\newcommand\cvcf{{\cvc}f\xspace}
-\newcommand\cvci{{\cvc}i\xspace}
-\newcommand\ziiib{\ziii}
-
-\renewcommand\models{\mathrel{\vDash}}
-
-%\newcommand{\DDD}{\mathcalx{D}}
-\newcommand\DDD{\Delta}
-
-%\newcommand\return{$\mathsf{return}$\xspace}
-\newcommand\return{}
-
-%\newcommand\bigtuple[1]{$(${#1}$)$}
-\newcommand\bigtuple[1]{$\bigl(${#1}$\bigr)$}
-
-%%% for \mathcalx
-\DeclareFontFamily{OT1}{pzc}{}
-\DeclareFontShape{OT1}{pzc}{m}{it}{<-> s * [1.10] pzcmi7t}{}
-\DeclareMathAlphabet{\mathcalx}{OT1}{pzc}{m}{it}
-
-%%% for correctly sized Greek
-\DeclareSymbolFont{letters}{OML}{txmi}{m}{it}
-
-%\newcommand\pneg{\neg}
-\newcommand\pnegate{-}
-%\newcommand\pnegate{{\sim}}
-
-\newcommand{\con}[1]{\mathsf{#1}}
-\let\const=\con
-
-\renewcommand\vec[1]{\overline{#1}}
-
-\let\oldcup=\cup
-\def\cup{\mathrel{\oldcup}}
-
-\let\oldchi=\chi
-\def\chi{{\vthinspace\oldchi}}
-
-\let\oldSigma=\Sigma
-\def\Sigma{\mathrm{\oldSigma}}
-
-\let\oldDelta=\Delta
-\def\Delta{\mathrm{\oldDelta}}
-
-\def\Chi{\mathrm{X}}
-
-\let\oldneg=\neg
-\def\neg{\oldneg\;}
-
-\let\oldvee=\vee
-\def\vee{\mathrel{\oldvee}}
-\let\oldwedge=\wedge
-\def\wedge{\mathrel{\oldwedge}}
-
-\newcommand\win{\bf}
-%\newcommand\bench{\ttfamily}
-
-\newcommand\isa{Isa\xspace}
-\newcommand\isam{Isa-Mut\xspace}
-\newcommand\leon{Leon\xspace}
-\newcommand\leonm{Leon-Mut\xspace}
-
-%\newcommand{\cvc}{\textsc{cvc}{\small 4}\xspace}
-%\newcommand{\cvciii}{\textsc{cvc}{\small 3}\xspace}
-%\newcommand{\ziii}{\textsc{z}{\small 3}\xspace}
-
-%%% @ANDY, @CESARE: I hope you'll forgive me these. I find the above fonts
-%%% distracting. --JB
-\newcommand\cvc{CVC4\xspace}
-\newcommand\cvciii{CVC3\xspace}
-\newcommand\ziii{Z3\xspace}
-
-\newcommand{\teq}{\approx}
-\newcommand{\cc}[1]{#1^*}
-
-%%% I'm trying to avoid bold whenever it can be avoided (except for headings
-%%% etc.) --JB
-%\newcommand{\terms}{\mathbf{T}}
-\newcommand{\terms}{\mathcalx{T}}
-
-%\newcommand{\functerms}{\mathbf{DFT}}
-\newcommand{\vars}{\mathbf{V}}
-\newcommand{\M}{\mathcalx{M}}
-\newcommand{\I}{\mathcalx{I}}
-\newcommand{\J}{\mathcalx{J}}
-\def\AIF{\qtab\keyword{if}\ }
-\def\THEN{\ \keyword{then}\ }
-\def\AELSE{\untab\qtab\keyword{else}\ }
-\def\FI{\untab}
-\def\RETURN{\keyword{return}\ }
-\def\ENDPROC{\untab}
-%%% @ANDY: I'm trying "A" like "abstraction", again in sync with the abstract--concrete
-%%% terminology from abstract interpretation
-\newcommand{\conv}{\mathcalx{A}}
-
-%\newcommand{\stypes}[1]{#1^\mathrm{s}}
-\newcommand{\stypes}[1]{#1^\mathrm{ty}}
-
-\newcommand{\sfuns}[1]{#1^\mathrm{f}}
-\newcommand{\sfundefs}[1]{#1^\mathrm{dfn}}
-%\newcommand{\sfunndefs}[1]{#1^\mathrm{nf}}
-\newcommand{\typeint}{\ty{Int}}
-
-%\newcommand{\pnone}{\con{none}}
-%\newcommand{\ppos}{\con{pos}}
-%\newcommand{\pneg}{\con{neg}}
-%%% @ANDY: as usual, revert if you don't like
-\newcommand{\pnone}{{\pm}}
-\newcommand{\ppos}{{+}}
-\newcommand{\pneg}{{-}}
-\newcommand{\pol}{\con{pol}}
-
-%\newtheorem{remark}{Remark}
-
-\newcommand\ty[1]{\con{#1}}
-\newcommand{\Bool}{\ty{Bool}}
-\newcommand{\ltrue}{\top}
-\newcommand{\lfalse}{\bot}
-\newcommand{\lite}{\con{ite}}
-
-%\newcommand\concret{\con{a}}
-%\newcommand\concret{\con{c}}
-\newcommand\concret{\gamma} %%% from the abstract interpretation literature
-
-\newcommand{\boolop}{\con{b}}
-\newcommand{\forallf}[1]{\forall_{\!#1\:}}
-\newcommand{\fnull}{\emptyset}
-\newcommand{\vecfarg}[1]{\vec{\concret}_{#1}}
-\newcommand{\farg}[1]{\concret_{#1}}
-\newcommand{\fargx}[2]{\concret_{#1,#2}}
-\newcommand{\fargtype}[1]{\alpha_{#1}}
-
-\newcommand{\Sigmalia}{\Sigma_{\mathcalx{LIA}}}
-\newcommand{\extendsig}[1]{\mathcalx{E}( #1 )}
-
-\newcommand{\rem}[1]{\textcolor{red}{[#1]}}
-\newcommand{\ajr}[1]{\rem{#1 --ajr}}
-\newcommand{\jb}[1]{\rem{#1 --jb}}
-\newcommand{\ct}[1]{\rem{#1 --ct}}
-
-\newcommand{\negvthinspace}{\kern-0.083333em}
-\newcommand{\vthinspace}{\kern+0.083333em}
-\newcommand{\vvthinspace}{\kern+0.0416666em}
-\newcommand{\typ}[1]{^{\vthinspace #1}}
-
-\newcommand{\definefunreccmd}{\con{define}\text{-}\con{fun}\text{-}\con{rec}}
-\newcommand{\definefunsreccmd}{\con{define}\text{-}\con{funs}\text{-}\con{rec}}
-
-\newcommand{\Mo}{{\mathcal{I}}}
-
-%\newcommand{\euf}{\ensuremath{\mathrm{E}}\xspace}
-\newcommand{\euf}{\ensuremath{\mathcalx{UF}}\xspace}
-%\newcommand{\ari}{\ensuremath{\mathrm{A}}\xspace}
-\newcommand{\ari}{\ensuremath{\mathcalx{A}}\xspace}
-
-\newcommand{\absconstraints}{\mathrm{X}}
-
-\newtheorem{cor}{Corollary}
-
-%\input{scalalistings}
-%\input{smtlib2listings}
-\usepackage{program}
-
-\def\squareforqed{\hbox{\rlap{$\sqcap$}$\sqcup$}}
-
-\newcommand\xend{{\hfill$\scriptstyle\blacksquare$}}
-\renewcommand\qed{{\hfill$\squareforqed$}}
-
-\begin{document}
-
-\title{Model Finding for Recursive Functions in SMT\thanks{%
-This research is partially supported by the Inria technological development
-action ``Contre-exemples Utilisables par Isabelle et Coq'' (CUIC).
-}
-}
-\titlerunning {Model Finding for Recursive Functions in SMT}
-
-\author {Andrew Reynolds\inst{1} \and Jasmin Christian Blanchette\inst{2,3} \and Cesare Tinelli \inst{4}}
-\authorrunning {A. Reynolds, J. C. Blanchette, C. Tinelli}
-\institute{
-\'Ecole Polytechnique F\'ed\'erale de Lausanne (EPFL), Switzerland
-\and
-Inria Nancy \& LORIA, Villers-l\`es-Nancy, France
-\and
-Max-Planck-Institut f\"ur Informatik, Saarbr\"ucken, Germany
-\and
-Department of Computer Science, University of Iowa, USA
-}
-
-\maketitle
-
-\begin{abstract}
-\noindent
-SMT solvers have recently been extended with techniques for finding models
-in presence of universally quantified formulas in some restricted fragments.
-This paper introduces a translation which reduces axioms specifying a large
-class of recursive functions, including well-founded (terminating),
-to universally quantified formulas for which these
-techniques are applicable.
-An empirical evaluation confirms that the approach improves 
-the performance of existing solvers on benchmarks from two sources. 
-The translation is implemented as a preprocessor in the solver CVC4.
-\end{abstract}
-
-%% The institutions above shouldn't count as footnotes
-\setcounter{footnote}{0}
-
-\section{Introduction}
-\label{sec:introduction}
-
-%\ct{The paper used "recursive datatype" in some places and "inductive datatype" in others. I have replaced both by "algebraic datatype" which is more accurate for \cvc.
-%("recursive datatype" is a more general notion and "inductive datatype" has a stronger meaning in higher-order logics.)}  
-%% I agree --JB.
-
-Many solvers based on SMT (satisfiability modulo theories) can reason about
-quantified formulas using incomplete instantiation-based methods
-\cite{MouraBjoerner07,ReynoldsTinelliMoura14}.
-These methods work well in the context of proving (i..e, showing
-unsatisfiability), but they are of little help for finding models (i.e.,
-showing satisfiability). Often, a single universal quantifier in one of the
-axioms of a problem is enough to prevent the discovery of models.
-
-In the past few years, techniques have been developed to find models for
-quantified formulas in SMT.
-Ge and de Moura \cite{GeDeM-CAV-09} introduced a complete instantiation-based
-procedure for formulas in the \relax{essentially uninterpreted} fragment.
-This fragment is limited to universally quantified formulas where all
-%ct occurrences of 
-variables occur as direct subterms of uninterpreted
-functions---e.g., $\forall x%\typ{\typeint} --- the type is not important, right?
-.\;\, \con{f}( x )
-\teq \con{g}( x ) + 5$.
-% since all
-% occurrences of $x$ occur beneath uninterpreted functions $\con{f}$ and
-% $\con{g}$.
-Other syntactic criteria extend
-this fragment slightly, including cases when variables occur as arguments of
-arithmetic predicates. Subsequently, Reynolds et al.\
-\cite{ReyEtAl-1-RR-13,reynolds-et-al-2013} introduced techniques for finding finite
-models for quantified
-formulas over uninterpreted types and types having a fixed finite
-interpretation. %, such as fixed-width bit vectors and enumerated datatypes.
-These techniques can
-find a model for a formula such as $\forall x,\, y : \tau.\;\, x \teq
-y \vee \allowbreak \neg \con{f}( x ) \teq \con{f}( y )$, where $\tau$ is an uninterpreted type.
-% where the uninterpreted type $\tau$ is
-% interpreted as a finite set and $\con{f}$ as injective finite map.
-
-Unfortunately, none of these fragments can accommodate the vast majority of
-quantified formulas that correspond to recursive function definitions: The
-essentially uninterpreted fragment does not allow the argument of a
-recursive function to be used inside a complex term on the right-hand side,
-whereas the finite model finding techniques of Reynolds et al.\ are not
-applicable for functions over infinite domains such as the integers or
-algebraic datatypes. A simple example where both approaches fail is
-\begin{equation}\label{eq:pii}
-\forall x : {\typeint}.\;\, \lite\bigl(
-x \leq 0,\allowbreak\; \con{p}( x ) \teq 1,\allowbreak\; \con{p}( x ) \teq 2 * \con{p}( x - 1 ) \bigr)
-\end{equation}
-This state of affairs is highly unsatisfactory, given the frequency of
-recursive definitions in practice and the impending addition of a dedicated
-command, \texttt{define-fun-rec}, to the SMT-LIB standard \cite{smtlib25}.
-
-In this paper, we present a method for translating formulas involving recursice function
-definitions into formulas where finite model finding techniques can be applied.
-The recursive functions must meet a semantic criterion to be admissible
-(Section~\ref{sec:prelim}). This criterion is met by well-founded (terminating)
-recursive function definitions. It is not met by inconsistent
-definitions such as $\forall x : {\typeint}.\;\, \con{f}(x) \teq \con{f}(x) +
-1$. %ct and it is possible to construct consistent examples that do not meet it.
-%as well as other example of consistent definition.
-
-We define a translation for a class of
-formulas involving admissible recursive function definitions
-(Section~\ref{sec:encoding}). The main insight is that a
-recursive definition $\forall x : \tau.\;\, \const{f}(x) \teq t$
-can be translated to
-$\forall a :
-\fargtype{\tau}.\;\, \const{f}(\farg{\con{f}}(a)) \teq t[\concret(a)/x]$, where
-$\fargtype{\tau}$ is an uninterpreted \relax{abstract} type and $\farg{\con{f}} :
-\fargtype{\tau} \to \tau$ converts the abstract type to the concrete
-type. The translation preserves satisfiability and, for admissible definitions,
-unsatisfiability, and makes finite model finding possible for problems in this
-class.
-
-Our empirical evaluation on benchmarks from the IsaPlanner proof planner~\cite{DBLP:conf/itp/JohanssonDB10}
-and the Leon verifier~\cite{blanc2013overview} provides
-evidence that this translation improves the effectiveness of the SMT solvers
-\cvc and \ziii for finding counterexamples to verification conditions
-(Section~\ref{sec:evaluation}). The approach is implemented as a preprocessor
-in \cvc (Section~\ref{sec:front-end}). When invoked with a specific
-command-line option, the solver assumes that functions introduced using the
-\texttt{define-fun-rec} command are admissible and translates them accordingly.
-
-%this option targets users who wish to determine the satisfiability of inputs
-%that involving function definitions that are known to be consistent, whereas it
-%does not target users who wish to check the consistency of a set of function
-%definitions.
-
-\section{Preliminaries}
-\label{sec:prelim}
-
-Our setting is a monomorphic (or many-sorted) first-order logic
-like the one defined by SMT-LIB \cite{smtlib25}.
-A \emph{signature} $\Sigma$ consists of
-a set $\stypes{\Sigma}$ of %ct type symbols
-first-order types
-%
-%\ct{What the rationale for using the letter $y$ in $\stypes{\Sigma}$?}
-%\jb{y as in t\textbf{y}pe. changed to `ty'}
-(or sorts) and a set $\sfuns{\Sigma}$ of function symbols over these types.
-We assume that signatures always contain a Boolean type $\Bool$ and constants
-$\ltrue, \lfalse : \Bool$ for truth and falsity,
-an infix equality predicate ${\teq} : \tau \times \tau \to \Bool$
-for each $\tau \in \stypes{\Sigma}$,
-standard Boolean connectives ($\neg$, $\wedge$, $\vee$, etc.),
-and an if--then--else function symbol
-$\lite : \Bool \times \tau \times \tau \rightarrow \tau$
-for each $\tau \in \stypes{\Sigma}$.
-Formulas are terms of type $\Bool$.
-We write $t\typ{\tau}$ to denote a metavariable ranging over terms of
-type~$\tau$ and $\terms( t )$ to denote the set of subterms in $t$.
-Given a term $u$, we write $u[\vec t/\vec x]$ to denote the result of replacing
-all occurrences of $\vec x$ with $\vec t$ in $u$.
-
-%We write $\functerms^\Sigma( t )$ to denote the set of $f$-applications in $t$ such that $\con{f} \in \sfundefs{\Sigma}$.
-
-A \emph{$\Sigma$-interpretation $\I$} %is a mathematical structure that
-maps each type $\tau \in \stypes{\Sigma}$ to a nonempty set $\tau^\I$,
-the \emph{domain} of~$\tau$ in~$\I$,
-and each function symbol $\con{f} : \tau_1 \times \cdots \times \tau_n \rightarrow \tau$ in
-$\sfuns{\Sigma}$
-to a total function $\con{f}^\I : \tau_1^\I \times \cdots \times \tau_n^\I \rightarrow \tau^\I$.
-A \emph{theory} is a pair $T = (\Sigma, \Mo)$ where
-$\Sigma$ is a signature and $\Mo$ is a class of $\Sigma$-interpretations,
-the \emph{models} of $T$.
-A $\Sigma$-formula $\varphi$ is
-\emph{$T$-satisfiable}
-if it is satisfied by some interpretation in $\Mo$;
-otherwise, it is \emph{$T$-unsatisfiable}.
-A formula $\varphi$ \emph{$T$-entails} $\psi$, written $\varphi \models_T \psi$,
-if all interpretations in $\Mo$ that satisfy $\varphi$ also satisfy $\psi$.
-Two formulas $\varphi$ and $\psi$ are \emph{$T$-equivalent} 
-if each $T$-entails the other.
-Given a signature $\Sigma$,
-the \emph{theory of equality with uninterpreted functions} \euf
-consists of the set of all $\Sigma$-interpretations.
-Informally, we refer to the type and function symbols in this theory as \emph{uninterpreted}.
-%\ct{This is both ambiguous and insufficient.
-%Since this is SMT we should fix some background theory $T$,
-%then extend its signature with additional symbols and call \emph{those} uninterpreted.
-%}
-
-In addition to standard quantified formulas $\forall \vec x.\; \varphi$,
-we consider \emph{annotated quantified formulas} of the form
-$\forallf{\const{f}} \vec x.\; \varphi$, where $\con{f} \in \sfuns{\Sigma}$ is
-an uninterpreted function symbol. Their
-semantics is the same as for standard quantified formulas.
-Given $\con{f} : \tau_1 \times \cdots \times \tau_n \rightarrow \tau$,
-an annotated
-quantified formula $\forallf{\con{f}} \vec x.\; \varphi$ is a \emph{function definition}
-(\,\emph{for $\con{f}$}\vthinspace) if $\vec x$ is a tuple of variables
-$x_1 : \tau_1$, $\ldots,$ $x_n : \tau_n$
-and $\varphi$ is a quantifier-free formula 
-$T$-equivalent to $\con{f}( \vec x ) \teq t$ for some term $t$ of type $\tau$.
-%\ct{Why can't $\varphi$ be existential instead of just quantifier free?} 
-%I don't think it makes much sense to have any quantifiers within function bodies, 
-%since they aren't operational, that's why I put quantifier-free. --ajr
-We write $\exists
-\vec x.\; \varphi$ as a shorthand for $\neg \forall \vec x.\; \neg \varphi$.
-
-\begin{definition}\rm
-A formula $\varphi$ is in \emph{definitional form with respect to}
-$\{ \con{f}_1, \ldots, \con{f}_n \} \subseteq \sfuns{\Sigma}$ if it is of the
-form
-%
-%\kern-\abovedisplayskip
-%\kern+\abovedisplayshortskip
-%
-$(\forallf{\con{f}_1} \vec x_1.\; \varphi_1) \wedge \cdots \wedge
-(\forallf{\con{f}_n} \vec x_n.\; \varphi_n) \wedge \psi$,
-%
-where $\con{f}_1, \ldots, \con{f}_n$ are distinct function symbols,
-%ct added
-$\forallf{\con{f}_i} \vec x_i.\; \varphi_i$ is a function definition
-for $i = 1, \ldots, n$,
-%
-%\ct{I think we need more conditions here on how a symbol $\con{f}_i$ can occur in $\varphi_j$.}
-%\jb{do we? I thought that the admissibility criterion below is enough.}
-% should be enough, unless I'm missing something? -ajr
-and $\psi$ contains no function definitions.
-We call $\psi$ the \emph{conjecture} of $\varphi$.
-%\ct{Don't we need additional restrictions on the $\varphi_i$'s here?
-%(E.g., they are quantifier-free, or perhaps existential or containing only let binders.)}
-% Now fixed above -ajr
-\end{definition}
-
-Given a signature $\Sigma$, we distinguish the subset of \emph{defined}
-function symbols $\sfundefs{\Sigma} \subseteq \sfuns{\Sigma}$.
-We consider $\Sigma$-formulas that are in definitional form with respect to
-$\sfundefs{\Sigma}$.
-
-\begin{definition}\rm
-Given a set of function definitions 
-$\DDD = \{ \forallf{\con{f}_1} \vec x.\; \varphi_1, \ldots, \forallf{\con{f}_n} \vec x.\; \varphi_n \}$, 
-a ground formula $\psi$ 
-%\ct{I suppose $\psi$ is meant to be ground.} yes --ajr
-is \emph{closed under function expansion with respect to $\DDD$} if 
-\[
- \psi \,\models_T\, \bigwedge\nolimits_{\,i=1}^{n} \{ \psi_i[ \vec t / \vec x ] \mid \con{f}_i( \vec t ) \in \terms( \psi ) \} %.
-\]
-The set $\DDD$ is \emph{admissible} if for every $T$-satisfiable formula 
-$\psi$ closed under function expansion with respect to $\DDD$,
-the formula
-$\psi \wedge \bigwedge \Delta$ is also $T$-satisfiable.
-%\ct{I}
-\end{definition}
-
-Admissibility is a semantic criterion that must be satisfied for each function
-definition before applying the translation described in
-Section~\ref{sec:encoding}. It is interesting to connect it to the
-standard notion of \emph{well-founded} function definitions, often called
-\emph{terminating} definitions in a slight abuse of terminology. In such
-definitions, all recursive calls are decreasing with respect to a well-founded
-relation, which must be supplied by the user or inferred automatically
-using a termination prover. This ensures that the function is uniquely defined
-at all points.
-
-First-order logic has no built-in notion of computation or termination. To ensure
-that a function specification is well-founded, it is sufficient to require that
-the function would terminate when seen as a functional program, under \emph{some}
-evaluation order. For example, definition~(\ref{eq:pii}) from the
-introduction can be shown well-founded under an evaluation strategy that
-evaluates the condition of a $\lite$ before evaluating the relevant branch,
-ignoring the other branch. Logically, such dependencies can be captured by
-congruence rules. Krauss developed these ideas in the more general context of
-higher-order logic \cite[Section 2]{krauss-2009-phd}.
-
-\begin{theorem}\label{thm:adm}
-If\/
-$\DDD$ is a set of well-founded function definitions for\/
-$\sfundefs{\Sigma} = \{\con{f}_1,\ldots,\con{f}_n\}$, then it is admissible.
-\end{theorem}
-\begin{proofsketch}
-Let $\psi$ be a satisfiable formula closed under function expansion with
-respect to $\DDD$. We show that $\psi \wedge \bigwedge \Delta$ is also
-satisfiable. Let $\I$ be a model of $\psi$, and
-let $\I_0$ be the restriction of $\I$ to the function symbols in $\sfuns{\Sigma} -
-\sfundefs{\Sigma}$. Because well-founded definitions uniquely characterize
-the interpretation of the functions they define, there exists a
-$\Sigma$-interpretation $\J$ that extends $\I_0$ such that $\J \models \Delta$.
-%
-Since $\psi$ is closed under function expansion, it already constraints the
-functions in $\sfundefs{\Sigma}$ recursively as far as is
-necessary for interpreting $\psi$. Thus, any point $v$ for which
-$\const{f}_i^\I(v)$ is needed for interpreting $\psi$ will have its
-expected value according to its definition and hence coincide with $\J$.
-And since $\psi^\I$ does not depend on the interpretation at the other
-points, $\J$ is, like $\I$, a model of $\psi$.
-Since $\J \models \Delta$ by assumption, we have $\J \models \psi \wedge
-\bigwedge \Delta$, as desired.
-%We have
-%$\varphi \,\models_T\, \bigwedge\nolimits_{\,i=1}^{n} \{ \varphi_i[ \vec t / \vec x ] \mid \con{f}_i( \vec t ) \in \terms( \varphi ) \}$.
-\qed
-\end{proofsketch}
-
-Another useful fragment of function definitions is the class of
-\emph{productive} corecursive functions. Corecursive functions are functions to
-a coalgebraic datatype. These functions can call themselves forever, without
-their being inconsistent. Productive corecursive functions are functions that
-progressively reveal parts of their potentially infinite
-output \cite{turner-1995,mcbride-productive}.
-For example, given a type of infinite streams constructed by
-$\con{scons} : \ty{int} \times \ty{stream} \to \ty{stream}$,
-the definition
-$\{\forallf{\const{e}} x.\;\, \const{e}(x) \teq \con{scons}(x,\; \const{e}(x + 1))\}$
-falls within this fragment: Each call to $\const{e}$ produces one
-constructor before entering the nested call. Like terminating recursion,
-productive corecursion totally specify the functions it defines.
-It is even possible to mix recursion and corecursion in the same function
-\cite{blanchette-et-al-2015-fouco}. Theorem~\ref{thm:adm} can be extended to
-cover such specifications, based on the observation that infinite unfolding
-a corecursive definition computes a unique infinite object.
-
-Beyond totality, an admissible set can contain underspecified functions
-such as $\forallf{\con{f}} x : \typeint.\allowbreak\;\, \con{f}( x )
-\teq \con{f}( x - 1 )$ or $\forallf{\con{g}} x.\allowbreak\;\, \con{g}( x
-) \teq \con{g}( x )$. More generally, we conjecture that one can ignore all
-tail-recursive calls (i.e., calls that occur on the right-hand side of the
-definition, possibly in some $\lite$ branch) when establishing well-foundedness
-or productivity, without affecting admissibility.
-
-An example of an inadmissible function set is
-$\{ \forallf{\con{f}} x.\;\, \con{f}( x ) \teq \con{f}( x ) + 1 \}$,
-since the formula $\ltrue$ is (trivially) closed under function expansion with respect to this set,
-and there is no model of $T$
-satisfying $\con{f}$'s definition. A more subtle specimen is
-\[\{ 
- \forallf{\con{f}} x.\;\, \con{f}( x ) \teq \con{f}( x ),\;\,
- \forallf{\con{g}} x.\;\, \lite( \con{f}( x ) \teq 5,\;
-   \con{g}( x ) \teq \con{f}( x ),\,
-   \con{g}( x ) \teq \con{g}( x ) + 1)
-\}\]
-While this set has a model where $\con f$ and $\con g$ are interpreted as the
-constant function $5$, it is not admissible since %the formula
-$\con{f}( 0 ) \teq 4$ is
-closed under function expansion with respect to this set and yet there exists
-no interpretation satisfying both $\con{f}( 0 ) \teq 4$ and $\con{g}$'s
-definition.
-
-%Terminating function specifications are always admissible, but also some
-%nonterminating
-%An admissible function definition need not specify a terminating function, e.g.
-%$\{ \forallf{\con{f}} x^\typeint.\;\, \con{f}( x ) \teq \con{f}( x - 1 ) \}$ or even $\{ \forallf{\con{f}} x.\;\, \con{f}( x ) \teq \con{f}( x ) \}$ are admissible.
-
-\section{The Translation}
-\label{sec:encoding}
-
-%%% TYPESETTING: approximated, to get centering
-\newcommand{\itemx}{\itemindent6.35em\item}
-
-\begin{figure}[t]
-\normalsize
-\begin{enumerate}
-%\begin{framed}
-\itemx[\ ]
-$\conv_0( t\typ{\tau},\, p )$ $=$
-\\[-.8\baselineskip]
- \begin{itemize}
-   \itemx[] $\mathsf{if}$ $\tau = \Bool$ $\mathsf{and}$ $t = \boolop(t_1,\ldots,t_n)$ $\mathsf{then}$
-    \begin{itemize}
-      \itemx[] $\mathsf{let}$ $( t'_i,\, \chi_i ) = \conv_0( t_i,\, \pol( \boolop,\, i,\, p ) )$ $\mathsf{for}$ $i = 1, \ldots, n$ $\mathsf{in}$%\kern6.35em$\mid$
-      \itemx[] $\mathsf{let}$ $\chi = \chi_1 \wedge \cdots \wedge \chi_n$ $\mathsf{in}$
-      \itemx[] $\mathsf{if}$ $p = \ppos$ $\mathsf{then}$ \bigtuple{$\boolop(t'_1, \ldots, t'_n) ) \wedge \chi,\, \ltrue$}
-      \itemx[] $\mathsf{else}$ $\mathsf{if}$ $p = \pneg$ $\mathsf{then}$ \bigtuple{$\boolop(t'_1, \ldots, t'_n) \vee \neg \chi,\, \ltrue$}
-      \itemx[] $\mathsf{else}$ \bigtuple{$\boolop(t'_1, \ldots, t'_n),\, \chi$}
-\\[-.8\baselineskip]
-    \end{itemize}
-  \itemx[] $\mathsf{else}$ $\mathsf{if}$ $t = \forallf{\con{f}} \vec x.\;\, u$ $\mathsf{then}$
-    \begin{itemize}
-      \itemx[] $\mathsf{let}$ $( u',\, \chi ) = \conv_0( u,\, p )$ $\mathsf{in}$ \bigtuple{$\forall a : \fargtype{\con{f}}.\; u' [ \vecfarg{\con{f}}( a ) / \vec x ],\, \ltrue$}
-\\[-.8\baselineskip]
-    \end{itemize}
-  \itemx[] $\mathsf{else}$ $\mathsf{if}$ $t = \forall \vec x.\;\, u$ $\mathsf{then}$
-    \begin{itemize}
-      \itemx[] $\mathsf{let}$ $( u',\, \chi ) = \conv_0( u,\, p )$ $\mathsf{in}$ \bigtuple{$\forall \vec x.\; u',\, \forall \vec x.\; \chi$}
-\\[-.8\baselineskip]
-    \end{itemize}
-   \itemx[] $\mathsf{else}$
-   \begin{itemize}
-     \itemx[] \return \bigtuple{$t,\; \bigwedge\, \{ \exists a : \fargtype{\con{f}}.\; \vecfarg{\con{f}}( a ) \teq \vec t \mid \con{f}( \vec t ) \in \terms( t ),\, \con{f} \in \sfundefs{\Sigma} \}$}
-   \end{itemize}
- \end{itemize}
-\end{enumerate}
-\begin{enumerate}
-%\begin{framed}
-\itemx[\ ]
-$\conv( \varphi )$ $=$ $\mathsf{let}$ $( \varphi',\, \_ ) = \conv_0( \varphi,\, \ppos )$ $\mathsf{in}$ $\varphi'$
-%\\[-1.25\baselineskip] %% TYPESETTING
-\end{enumerate}
-\caption{\,Definition of translation $\conv$}
-\label{fig:A}
-\end{figure}
-
-We fix a signature $\Sigma$ and a $\Sigma$-formula
-$\varphi$ in definitional form with respect to $\sfundefs{\Sigma}$,
-whose definitions are admissible.
-%We may translate $\varphi$ into a equisatisfiable formula $\phi'$ for which known model-finding procedures~\cite{GeDeM-CAV-09, ReyEtAl-1-RR-13} are applicable.
-We present a method that constructs an extended signature
-$\extendsig{ \Sigma }$ and an $\extendsig{ \Sigma }$-formula $\varphi'$ such that
-$\varphi'$ is satisfiable if and only if $\varphi$ is satisfiable---i.e.,
-$\varphi$ and $\varphi'$ are \emph{equisatisfiable}.
-The idea behind this translation
-is to use an uninterpreted type to abstract the set of
-\relax{relevant} tuples for each defined function $\con{f}$ and restrict the
-quantification of $\con{f}$'s definition to a variable of this
-type. Informally, the relevant tuples $\vec t$ of a function $\con{f}$ are the
-ones for which the interpretation of $\con{f}( \vec t )$ is relevant to the
-satisfiability of $\varphi$.
-
-More precisely,
-suppose our signature $\Sigma$ contains a set of defined function symbols $\sfundefs{\Sigma} \subseteq \sfuns{\Sigma}$.
-For each $\con{f} : \tau_1 \times \cdots \times \tau_n \rightarrow \tau \in \sfundefs{\Sigma}$,
-the extended signature $\extendsig{\Sigma}$ contains
-%\begin{itemize}
-%\item
-an uninterpreted \emph{abstract type} $\fargtype{\con{f}}$ and
-%\item
-$n$ uninterpreted \emph{concretization functions} $\fargx{\con{f}}{1} : \fargtype{\con{f}} \rightarrow \tau_1$, \ldots, $\fargx{\con{f}}{n} : \fargtype{\con{f}} \rightarrow \tau_n$.
-%\end{itemize}
-%The interpretation of uninterpreted type $\fargtype{\con{f}}$ will denote the elements (tuples) on which the function $\con{f}$ is applied.
-%The role of the uninterpreted functions $\farg{\con{f}}^1$, $\ldots$, $\farg{\con{f}}^n$ will be discussed more in the following.
-
-The translation $\conv$ defined in Figure~\ref{fig:A} translates the $\Sigma$-formula
-$\varphi$ into an $\extendsig{\Sigma}$-formula. It relies
-on the auxiliary function $\conv_0$, which takes two arguments:\ the term $t$
-to translate and a \relax{polarity}~$p$, which is either $\ppos$, $\pneg$, or
-$\pnone$ (none). \vthinspace$\conv_0$ returns a pair $( t'\negvthinspace,\, \chi )$, where $t'$ is a term of
-type $\tau$ and $\chi$ is an additional constraint.
-This constraint is needed if the polarity is $\pnone$; otherwise, it is encoded
-directly in $t'$ and $\chi$ is simply $\ltrue$.
-In the auxiliary function $\conv_0$, the function $\pol(
-\boolop, i, p )$ returns the polarity of the $i$\vvthinspace{th} argument of an
-application of~$\boolop$ having polarity $p$ if $\boolop$ is an interpreted
-predicate or $\pnone$ otherwise.
-
-%The role of $D$ is to ensure that
-%the (restricted) function definition for $\con{f}$ includes certain tuples in
-%its domain, which we explain more in the following.
-
-The translation alters $\varphi$ in two main ways. First, it restricts the
-quantification on function definitions for $\con{f}$ to the corresponding
-uninterpreted type $\fargtype{\con{f}}$ and inserts appropriate selectors
-$\fargx{\con{f}}{i}$. Second, it augments $\varphi$ with additional existential
-constraints
-$\exists a : \fargtype{\con{f}}.\;\, \vecfarg{\con{f}}( a ) \teq \vec t$; these
-ensure that the restricted definition for $\con{f}$ covers all relevant tuples
-of terms, namely the ones for which an $\con{f}$ application exists and is
-relevant to the satisfiability of $\varphi$.
-%\jb{what does ``current branch of'' mean? can we drop it / rephrase it / vague
-%it up?}
-
-If $t$ is an application of a predicate $\boolop$, including the built-in
-operators ${\neg}$, ${\wedge}$, ${\vee}$, ${\teq}$, and ${\lite}$,
-$\conv_0$ calls itself recursively on the arguments $t_i$ and polarity $\pol(
-\boolop, i, p )$, where $\pol$ is defined as
-\begin{equation*}
-\pol( \boolop, i, p ) =
-\begin{cases}
-p & \text{if either $\boolop \in \{{\wedge}, {\vee}\}$ or $\boolop = \lite$ and $i \in \{2, 3\}$} \\[-\jot]
-\pnegate p & \text{if $\boolop = {\neg}$} \\[-\jot]
-\pnone & \text{otherwise}
-\end{cases}
-\end{equation*}
-The negation $\pnegate p$ of a polarity $p$ is $\pneg$ if $p$ is $\ppos$, $\ppos$
-if $p$ is $\pneg$, and $\pnone$ if $p$ is $\pnone$.
-
-If $t$ is positive, $\conv_0$ adds the constraint $\chi$ derived from
-the subterms as a conjunction. Dually, if $t$ is negative, it constructs a
-disjunction with the negated constraint. It $t$ has no polarity, the
-constraint $\chi$ is simply returned to the caller.
-
-If $t$ is a function definition of a function in
-$\sfundefs{\Sigma}$, $\conv_0$ constructs a quantified formula over a single variable
-$a$ of type $\fargtype{\con{f}}$ and replaces all occurrences of $\vec x$ in
-$t'_1$ with $\vecfarg{\con{f}}( a )$. Since function definitions
-are a top-level conjunct, %by case analysis on the return values of $\conv$,
-$\chi$ must be $\ltrue$.
-
-If $t$ is an unannotated quantified formula, $\conv_0$ calls itself
-on the body with the same polarity. A quantifier is prefixed to the
-quantified formula and constraint returned by the recursive call.
-
-Otherwise, $t$ is either an application of an uninterpreted predicate or a term
-of a type other than $\Bool$. The returned constraint is a conjunction of
-formulas of the form $\exists a : {\fargtype{\con{f}}}.\;\, \vecfarg{\con{f}}(
-a ) \teq \vec t$ for each subterm $\con{f}( \vec t )$ of $t$ such that $\con{f}
-\in \sfundefs{\Sigma}$. Such constraints, when asserted positively, ensure that
-the tuple of uninterpreted functions $\vecfarg{\con{f}}$ has $\vec t$
-in its range.
-
-%We demonstrate this translation with an example.
-
-\begin{example}
-\label{ex:translation}
-Consider the combined signature of linear arithmetic and uninterpreted
-functions $\Sigma = \Sigmalia \mathrel{\uplus} \Sigma_{\euf}$, where
-%$\stypes{\Sigma_{\euf}} = \{ \typeint \}$ 
-%\ct{shouldn't this be $\stypes{\Sigmalia} = \{ \typeint \}$?}
-%and
-% Something went wrong, let's just remove it --ajr
-$\sfuns{\Sigma_{\euf}} = \{ \con{s} : { \typeint \rightarrow \typeint },\, \const{c} : { \typeint } \}$.
-Let $\varphi$ be the $\Sigma$-formula
-\begin{equation} \label{eq:ex-before}
-\forallf{\con{s}} x : {\typeint}.\;\, \lite\bigl( x \leq 0,\; \con{s}(x) \teq 0,\;
-  \con{s}( x ) \teq x + \con{s}( x - 1 ) \bigr) \wedge \con{s}( \con{c} ) > 100
-\end{equation}
-%
-The definition of $\const{s}$ specifies that it returns the sum of all
-positive integers up to $x$. The formula $\varphi$ is in definitional form with
-respect to $\sfundefs{\Sigma}$ 
-%\jb{correct? It said $\Sigma_u$, which was definitely wrong} 
-and states that the sum of all
-positive numbers up to $\const{c}$ is greater than $100$. This formula is
-satisfiable with a model that interprets $\const{c}$ as $14$ or more.
-Due to the universal quantifier,
-current SMT techniques
-%~\cite{GeDeM-CAV-09, ReyEtAl-1-RR-13,reynolds-et-al-2013}
-%%% We described them in the intro. When I see references like that appear with
-%%% no description, as a reader, I get all paranoid and wonder if it's the old
-%%% stuff from the intro (Ge, Moura, Renoylds et al.) or if it's new stuff ---
-%%% and I cannot remember numbers like [9,18,19] across sections or even
-%%% paragraphs. --jb
-are unable to find
-a model for this formula. The signature $\extendsig{\Sigma}$ contains the type
-$\fargtype{\con{s}}$ and the uninterpreted function $\farg{\con{s}} : \fargtype{\con{s}}
-\rightarrow \typeint$. The result of $\conv( \varphi )$ 
-is $\varphi'$ which (after simplification) is the $\extendsig{\Sigma}$-formula
-%
-\begin{equation} \label{eq:ex-after}
-\!\begin{aligned}[c]
-  & \phantom{{\wedge}\; }\bigl(
-      \forall a : \fargtype{\con{s}}.\; \lite\bigl(
-        \!\begin{aligned}[t]
-         &  \farg{\con{s}}( a ) \leq 0,\;
-          \con{s}(\farg{\con{s}}( a )) \teq 0,\;
-\\[-\jot]
-  & \con{s}(\farg{\con{s}}( a )) \teq \farg{\con{s}}( a ) + \con{s}( \farg{\con{s}}( a )-1 )
-    \wedge \exists b : {\fargtype{\con{s}}}.\;\, \farg{\con{s}}( b ) \teq \farg{\con{s}}( a )-1 \bigr) \bigr)
-\end{aligned}
-\\[-\jot]
- & {\wedge}\; \con{s}( \con{c} ) > 100 \wedge \exists a : {\fargtype{\con{s}}}.\;\, \farg{\con{s}}( a ) \teq \con{c}
-\end{aligned}
-\end{equation}
-%
-The universal quantification in the resulting formula is over an uninterpreted
-type $\fargtype{\con{s}}$, thus enabling the finite model finding
-techniques by Reynolds et al.\ \cite{ReyEtAl-1-RR-13,reynolds-et-al-2013},
-implemented in \cvc, to search for a finite interpretation for $\fargtype{\con{s}}$. 
-Since all occurrences of the quantified variable $a$ in this formula are 
-beneath applications of the uninterpreted function $\farg{\con{s}}$, 
-the formula is in the essentially uninterpreted fragment,
-for which Ge and de Moura \cite{GeDeM-CAV-09} provide 
-a complete instantiation procedure, implemented in \ziii. 
-Both
-%%% Added "apparently". Or are we sure?  
-% Yes, I'm planning to post the 2 versions of the benchmark on the evaluation page (when we add it) --ajr
-% Yes, but how can you be sure that Z3 will run forever? Your benchmarks surely have a timeout! --jb
-\cvc and \ziii run indefinitely on formula~(\ref{eq:ex-before}), 
-whereas they both produce a model for formula~(\ref{eq:ex-after}) 
-within 100 milliseconds.\xend
-\end{example}
-
-The translation $\conv$ results in formulas whose models are different than those of $\varphi$.
-%\ct{"Preserves models" means that it does not drop models, not that it adds models.}
-One model $\I$ for formula~(\ref{eq:ex-after}) in the above example interprets
-$\fargtype{\con{s}}$ as a finite set $\{ u_0, \ldots, u_{14} \}$,
-$\farg{\con{s}}$ as a finite map $u_i \mapsto i$ for $i = 0, \ldots, 14$,
-$\con{c}$ as $14$,
-and $\con{s}$ as the (almost constant) function
-%
-%\begin{equation} \label{eq:approx-interp}
-\[\lambda x : \typeint.\;\, \lite( x \teq 0,\; 0,\;
-  \lite( x \teq 1,\; 1,\;
-    \lite( x \teq 2,\; 3,\;
-      \lite(\ldots,\; \lite( x \teq 13,\, 91,\, 105 )\negvthinspace \ldots ))))\]
-%\end{equation}
-%
-In other words, $\const{s}$ is interpreted as a function mapping $x$ to the sum
-of all positive integers up to $x$ when $0 \leq x \leq 13$, and $105$
-otherwise.
-The interpretation $\I$, restricted to $\Sigma$, is {not} a model for the original formula~(\ref{eq:ex-before}),
-since $\I$ wrongly interprets $\con{s}( n )$ as $105$ when $n < 0$ or $n > 14$.
-
-However, under the assumption that the function definitions in
-$\sfundefs{\Sigma}$ are admissible, %we claim that
-$\conv(\varphi)$ is equisatisfiable with $\varphi$ for any input $\varphi$.
-%\ct{Must define "equisatisfiable".
-%What we mean here is a weak form of satisfiability, we could instead say that
-%$\conv(\varphi)$ is $T$-satisfiable iff $\varphi$ is $T$-satisfiable.
-%} 
-%\jb{We do define it, even it is weak form.}
-The intuition is that the
-interpretation of a term such as $\con{s}( -1 )$ in the previous example is not
-relevant to the satisfiability of formula~(\ref{eq:ex-before}). Thus, a
-satisfiable or unsatisfiable response from an SMT solver on input $\conv(
-\varphi )$ implies the existence or nonexistence of a model for
-$\varphi$. Moreover, the models of $\conv( \varphi )$ contain
-pertinent information regarding the models of $\varphi$. For example, the model
-$\I$ for formula~(\ref{eq:ex-after}) given above interprets $\con{c}$ as $14$
-and $\con{s}(n) = \sum_{i=1}^n i$ for $0 \le n \le 14$,
-and there exists a model of formula~(\ref{eq:ex-before}) that also interprets
-$\con{c}$ and and $\con{s}(n)$ in the same way (for $0 \le n \le 14$).
-In general, there exists a model of $\varphi$ that
-coincides with each model of $\conv( \varphi )$ on its interpretation of all
-functions in $\sfuns{\Sigma} - \sfundefs{\Sigma}$, i.e., $\{
-\con{s} \}$ in the example. Furthermore, the model of $\conv( \varphi )$ will
-also give correct information for the defined functions at all points belonging
-to the domains of the corresponding abstract types. This can sometimes help
-users debug their specifications.
-
-%%% @ANDY: I don't get the following comment --jb
-%   @JASMIN : The point of this comment is that the user is not interested in the SMT solver inferring any new information regarding the interpretation of s,
-%             since s is a "defined" function, the user already has an intended interpretation in their head of what s is (here, the sum of positive integers between 0...x).
-%             Thus, it isn't a big deal that the SMT solver "approximates" the interpretation of s during a check, since we already know what s is.
-%From a practical point of view, this is not an
-%issue, because $s$ is the very function that was explicitly defined by the
-%user, and hence already has an intended interpretation.
-
-%%% @ANDY: Now I understand. My experience with Nitpick is that users sometimes
-%%% want to see the partial information available about defined functions. I've
-%%% added a couple of sentences above to that effect.
-
-In the following, for a set of ground literals $L$, 
-we write $\absconstraints( L )$ to denote the set of constraints that
-force the concretization functions to have the necessary elements in their
-range for determining the satisfiability of $L$ with respect to function
-definitions in the translation.
-Formally, we define
-$\absconstraints( L )$
-as
-$\{ \exists a : {\fargtype{\con{f}}}.\;\, \vecfarg{\con{f}}( a ) \teq \vec t
-\mid \con{f}( \vec t ) \in \terms( L ),\: \con{f} \in \sfundefs{\Sigma}
-\}$.
-The following lemma states the central invariant behind the translation $\conv$.
-
-\begin{lemma}\label{lem:conv}
-Let $\varphi$ be a formula not containing function definitions,
-and let $\I$ be an $\extendsig{\Sigma}$-interpretation.
-Then $\I$ satisfies $\conv( \varphi )$ if and only if
-$\I$ satisfies $L \cup \absconstraints( L )$, where $L$ is a set of ground $\Sigma$-literals that entail $\varphi$.
-%There exists a set of ground $\Sigma$-literals $L$ such that $L$ entails $\varphi$, 
-%and $\I$ satisfies $L \cup \absconstraints( L )$.
-\end{lemma}
-\begin{proofsketch}
-By definition of $\conv$ and case analysis on the return values of $\conv_0$.
-\end{proofsketch}
-
-\begin{cor}\label{cor:conv}
-If $\varphi$ is a formula not containing function definitions, then $\extendsig{\Sigma} = \Sigma$
-and $\conv( \varphi )$ entails $\varphi$.
-\end{cor}
-
-\begin{theorem}\label{thm:equi}
-If $\varphi$ is a\/ $\Sigma$-formula that is in definitional form with respect to\/
-$\sfundefs{\Sigma}$
-and the set of function definitions\/ $\DDD$ corresponding to functions in\/
-$\sfundefs{\Sigma}$ is admissible, then\/ $\varphi$ and\/ $\conv( \varphi )$ are
-equisatisfiable.
-\end{theorem}
-\begin{proofsketch}
-First, we show that if $\varphi$ is satisfied by $\Sigma$-interpretation $\I$, then $\conv( \varphi )$ is satisfied by an $\extendsig{\Sigma}$-interpretation $\J$.
-Let $\J$ be the $\extendsig{\Sigma}$-interpretation that interprets all types $\tau \in \stypes{\Sigma}$ as $\tau^\I$,
-all functions $\con{f} \in \sfuns{\Sigma}$ as $\con{f}^\I$,
-and for each function $\con{f} : \tau_1 \times \cdots \times \tau_n \rightarrow \tau$ in $\sfundefs{\Sigma}$,
-interprets $\fargtype{\con{f}}$ as a tuple of values belonging to $\tau_1^\I \times \cdots \times \tau_n^\I$,
-and each $\fargx{\con{f}}{i}$ as the $i$\vvthinspace{th} projection on such tuples for $i = 1, \ldots, n$.
-Since $\J$ satisfies $\varphi$, it satisfies a set of ground literals $L$ that entail $\varphi$.
-Furthermore, $\J$ satisfies \relax{every} constraint of the form $\exists a : {\fargtype{\con{f}}}.\;\, \vecfarg{\con{f}}( a ) \teq \vec t$,
-since by our construction of $\J$ there exists a value $v \in {\fargtype{\con{f}}}^\J$ such that $v = \vec t^\J$.
-%By case analysis it follows that $\conv_0$ maintains the invariant that
-%if $\conv_0( t\typ{\Bool}, \ppos )$ returns $( {t'}, \chi )$,
-%then $\I$ satisfies $t$ if and only if $\J$ satisfies $t'$.
-Thus, $\J$ satisfies $L \cup \absconstraints( L )$, and by Lemma~\ref{lem:conv} we conclude $\J$ satisfies $\conv( \varphi )$.
-
-Second, we show that if $\conv( \varphi )$ is satisfied by a
-$\extendsig{\Sigma}$-interpretation $\J$, then $\varphi$ is satisfied by a
-$\Sigma$-interpretation $\I$. Since $\varphi$ is in definitional form with
-respect to the functions defined by $\DDD$, it must be of the form
-$\DDD \wedge \varphi_0$. 
-%We argue that $\J$ satisfies a $\Sigma$-formula
-%$\varphi'$ that is closed under function expansion with respect to $\DDD$, and
-%where $\varphi'$ entails $\varphi_0$.
-%To construct $\varphi'$, 
-First, we define a sequence of $\Sigma$-literals sets $L_0 \subseteq L_1 \subseteq \cdots$
-such that $\J$ satisfies $L_i \cup \absconstraints( L_i )$ for $i = 0, 1, \ldots\vthinspace$.
-Since $\J$ satisfies $\conv( \varphi_0 )$,
-by Lemma~\ref{lem:conv}, 
-$\J$ satisfies a set of literals $L \cup \absconstraints( L )$ where $L$ is a set of $\Sigma$-literals that entail $\varphi_0$.
-Let $L_0 = L$.
-For each $i \geq 0$,
-let $\psi_i$ be the formula $\bigwedge\, \{ \conv( \varphi_{\con{f}}[ \vec t / \vec x ] ) \mid \con{f}( \vec t ) \in \terms( L_i ),\: \con{f} \in \sfundefs{\Sigma} \}$,
-where the function definition of $\con{f}$ in $\DDD$ is $\forallf{\con{f}} \vec x.\; \varphi_{\con{f}}$.
-Since $\J$ satisfies $\conv( \forallf{\con{f}} \vec x.\; \varphi_{\con{f}} )$ and $\absconstraints( L_i )$,
-we know that $\J$ also satisfies $\psi_i$.
-Thus by Lemma~\ref{lem:conv},
-$\J$ satisfies a set of literals $L \cup \absconstraints( L )$ where $L$ is a set of $\Sigma$-literals that entail $\psi_i$.
-Let $L_{i+1} = L_0 \cup L$.
-Let $L_\infty$ be the limit of this sequence (i.e., $\ell \in L_\infty$ if and only if
-$\ell \in L_i$ for some $i$),
-and let $\psi$ be the $\Sigma$-formula $\bigwedge L_\infty$.
-To show that $\psi$ is closed under function expansion with respect to $\DDD$,
-we first note that by construction $\psi$ entails $\psi_\infty$.
-For any function symbol~$\con{f}$ and terms~$\vec t$, since $\varphi_{\con{f}}[ \vec t / \vec x ]$ does not contain function definitions,
-by Corollary~\ref{cor:conv},
-$\conv( \varphi_{\con{f}}[ \vec t / \vec x ] )$ entails $\varphi_{\con{f}}[ \vec t / \vec x ]$.
-Thus, $\psi$ entails $\{ \varphi_{\con{f}}[ \vec t / \vec x ] \mid \con{f}( \vec t ) \in \terms( \psi ), \con{f} \in \sfundefs{\Sigma} \}$,
-meaning that $\psi$ is closed under function expansion with respect to $\DDD$.
-Furthermore, $\psi$ entails $\varphi_0$ since $L_0 \subseteq L_\infty$.
-Since $\psi$ is a $T$-satisfiable formula that is closed under function expansion with respect to $\DDD$ and $\DDD$ is admissible, 
-by definition there exists a $\Sigma$-interpretation $\I$ satisfying $\psi \wedge \DDD$, which entails $\DDD \wedge \varphi_0$,
-i.e., $\varphi$.
-\qed
-\end{proofsketch}
-
-The intuition of the above proof is as follows.
-First, $\conv( \varphi )$ cannot be unsatisfiable when $\varphi$ is satisfiable
-since any model of $\varphi$ can be extended in a straightforward way to satisfy $\conv( \varphi )$.
-Second, if a model is found for $\conv( \varphi )$,
-then the constraints of the form $\exists a : \fargtype{\con{f}}.\;\, (\vecfarg{\con{f}}( a ) \teq \vec t)$
-occurring in $\conv( \varphi )$ ensure that this model also satisfies
-a $\Sigma$-formula that is closed under function expansion that entails the conjecture of $\varphi$.
-Hence when $\DDD$ is admissible, this implies the existence of a model for $\varphi$.
-%For clarity, we demonstrate the consequences of this theorem by means of an example.
-
-\begin{example}
-Let us revisit the translated formula~(\ref{eq:ex-after}) from
-Example~\ref{ex:translation}. 
-This formula is unsatisfiable only if
-formula~(\ref{eq:ex-before}) is unsatisfiable since $\fargtype{\con{s}}$ can be
-interpreted as the integers and $\farg{\con{s}}$ as the identity
-function. Conversely, we claim formula~(\ref{eq:ex-after}) is satisfiable
-only if formula~(\ref{eq:ex-before}) is satisfiable, noting that the set
-$\{ \forallf{\con{s}} x.\;\, \varphi_\con{s} \}$ is admissible,
-where $\varphi_\con{s}$ is the formula $\lite\bigl( x \leq 0,\; \con{s}(x) \teq 0,\; \con{s}( x ) \teq x + \con{s}( x-1 ) \bigr)$.
-%\ct{
-%You cannot possibly mean "if and only if" in the two sentences above.
-%Once I show that (1) is unsat iff (2) is unsat, I do not need to argue that
-%(1) is sat iff (2) is sat. That is immediate.
-%}
-% You're right, changed both to "only if" --ajr
-%As stated
-%in the proof of Theorem~\ref{thm:equi}, the reason is that any model of
-%formula~(\ref{eq:ex-after}) must satisfy a formula that is closed under
-%function expansion, hence implying that formula~(\ref{eq:ex-before}) has a
-%model.
-%The rightmost conjunct $\exists a : {\fargtype{\con{s}}}. (\farg{\con{s}}( a ) \teq \con{c})$ 
-%ensures that if $\con{s}( \con{c} ) > 100$ is a constraint that 
-%formula~(\ref{eq:ex-after}) satisfying $\con{s}( \con{c} ) > 100$ also
-%satisfies $\varphi_{\con{s}}[\con{c}/x]$. 
-%The nested existential
-%$\exists b.\;\, \farg{\con{s}}( b ) \teq \farg{\con{s}}( a )-1$ ensures that
-%a similar constraint is enforced recursively: when $\varphi_{\con{s}}[t/x]$ is used in part to evaluate $\con{s}( \con{c} )$,
-%any model of $\varphi_{\con{s}}[t/x]$ must also satisfy $\varphi_{\con{s}}[t-1/x]$ when $t > 0$. 
-It is easy to see that any interpretation $\I$ satisfying formula~(\ref{eq:ex-after}) satisfies $L_0 \cup \absconstraints( L_0 )$,
-where $L_0 = \{ \con{s}( \con{c} ) > 100 \}$, 
-where $\absconstraints( L_0 )$ is the set consisting of the single constraint $\exists a : {\fargtype{\con{s}}}. (\farg{\con{s}}( a ) \teq \con{c})$.
-%The latter ensures that $\I$ satisfies a formula that entails both $\con{s}( \con{c} ) > 100$ and $\varphi_{\con{s}}[\con{c}/x]$.
-Since $\I$ also satisfies both the translated function definition for $\con{s}$---i.e.,
-the first conjunct of formula~(\ref{eq:ex-after})---and $\absconstraints( L_0
-)$, it must also satisfy the formula $\psi_1$:
-\[
-\lite\bigl( \con{c} \leq 0,\; 
-            \con{s}(\con{c}) \teq 0,\;
-            \con{s}(\con{c}) \teq \con{c} + \con{s}( \con{c}-1 )
-            \wedge \exists b : {\fargtype{\con{s}}}.\;\, \farg{\con{s}}( b ) \teq \con{c}-1 \bigr)
-\]
-The existential constraint in the above formula
-ensures that whenever $\I$ satisfies the set $L_1 = L_0 \cup \{ \neg \con{c} \leq 0,\; \con{s}(\con{c}) \teq \con{c} + \con{s}( \con{c}-1 ) \}$,
-$\I$ satisfies $\absconstraints( L_1 )$ as well.
-Hence, by repeated application of this reasoning, it follows that
-a model of formula~(\ref{eq:ex-after}) that interprets $\con{c}$ as $n$ must also satisfy the formula $\psi$:
-%
-\[\con{s}( \con{c} ) > 100 \wedge %\displaystyle
-  \bigwedge\nolimits_{\vthinspace i=0}^{\!n-1} \bigl( \neg (\con{c}-i \leq 0) \wedge \con{s}( \con{c}-i ) \teq \con{c}-i + \con{s}( \con{c}-i-1 ) \bigr)
-\wedge \con{c}-n \leq 0 \wedge \con{s}( \con{c}-n ) \teq 0\]
-%
-This formula is closed under function expansion
-since it entails $\varphi_\con{s}[(\con{c}-i)/x]$ for $i = 0, \ldots, n$,
-and it contains only $\con{s}$ applications corresponding to $\con{s}( \con{c}-i )$ for $i = 0, \ldots, n$.
-Since $\{\forallf{\con{s}} x.\;\, \varphi_\con{s}\}$ is admissible,
-there exists a $\Sigma$-interpretation satisfying $\psi \wedge \forallf{\con{s}} x.\;\, \varphi_\con{s}$
-which entails formula~(\ref{eq:ex-before}).\xend
-\end{example}
-
-\section{Evaluation}
-\label{sec:evaluation}
-
-In this section, we evaluate both the overall impact of the translation
-introduced in the previous section and the performance of individual SMT
-techniques.
-%
-We gathered a set of 245 benchmarks from two sources, which we will refer to as
-\isa and \leon. The first source consists of the 79
-benchmarks from the IsaPlanner
-suite~\cite{DBLP:conf/itp/JohanssonDB10} that do not contain higher-order
-functions. The benchmarks have been used recently as challenge problems for a
-variety of inductive theorem provers. These benchmarks heavily involve
-recursive functions and are limited to the combined theory of uninterpreted
-functions and algebraic datatypes. The second source consists of 166
-benchmarks from the Leon repository,\footnote{%Available at
-\url{https://github.com/epfl-lara/leon/}} which were constructed from
-verification conditions from simple Scala programs. These benchmarks also
-heavily involve recursively defined functions over algebraic datatypes, 
-but cover a wide variety of theories, including bit vectors, arrays, and
-both linear and nonlinear arithmetic. Each benchmark is in definitional form
-with respect to a set of well-founded functions.
-%
-% in particular they each
-%are a conjoined list of function definitions followed by a (negated)
-%conjecture.
-Nearly all of these benchmarks are unsatisfiable, except for a
-handful of benchmarks in the \leon set.
-
-For each of these 245 benchmarks, we considered up to three randomly selected
-mutated forms of its conjecture $\psi$. In particular, we considered unique
-conjectures that are obtained as a result of swapping a subterm of $\psi$ at
-one position with another of the same type at another position.
-Note that benchmarks created in this way have a high likelihood of having small
-and easy-to-find countermodels. In total, we considered 213 mutated forms of
-conjectures from \isa and 427 mutated forms of conjectures from \leon. We will
-call these sets \isam and \leonm, respectively.
-
-Thus, our benchmark set consist of 885 benchmarks (640 mutants plus the
-original 245). We considered these 885 benchmarks both before and after the
-translation~$\conv$. % as described in Section~\ref{sec:encoding}.
-For example, \isa contains 79 original benchmarks $\varphi$ and 79 translated
-benchmarks $\conv(\varphi)$.
-
-For solvers, we considered the SMT solver \ziii~\cite{de-moura-bjoerner-2008},
-which runs both heuristic methods for quantifier instantiation~\cite{MouraBjoerner07}
-and methods for finding models for quantified formulas~\cite{GeDeM-CAV-09}.
-We also considered three configurations of \cvc~\cite{barrett-et-al-2011} which
-we will refer to as \cvcd, \cvcf, and \cvci.
-The default configuration \cvcd runs heuristic and conflict-based techniques for quantifier instantiation~\cite{ReynoldsTinelliMoura14},
-but does not include techniques for finding models.
-\cvcf runs heuristic instantiation and the finite model
-finding procedure due to Reynolds et al.\ \cite{ReyEtAl-1-RR-13,reynolds-et-al-2013}.
-\cvci incorporates techniques for automating inductive
-reasoning in SMT~\cite{reynolds-kuncak-2015}.
-
-\begin{figure}[t]
-\normalsize
-\centering
-\begin{tabular}{l@{\kern1.5em}r@{\kern0.375em}r@{\kern1.5em}r@{\kern0.375em}r@{\kern1.5em}r@{\kern0.375em}r@{\kern1.5em}r@{\kern0.375em}r@{\,\,}}
-  & \multicolumn{2}{c@{\kern1.5em}}{\phantom{0}\ziiib}     & \multicolumn{2}{c@{\kern1.5em}}{\phantom{0}\cvcd}     & \multicolumn{2}{c@{\kern1.5em}}{\phantom{0}\cvcf}     & \multicolumn{2}{c@{\,\,}}{\phantom{0}\cvci}
-\\%[-1pt]
-  & \hfill $\varphi$ \hfill & $\conv(\varphi)$\!\!
-  & \hfill $\varphi$ \hfill & $\conv(\varphi)$\!\!
-  & \hfill $\varphi$ \hfill & $\conv(\varphi)$\!\!
-  & \hfill $\varphi$ \hfill & $\conv(\varphi)$\!\!
-\\
-\midrule
-\isa & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0
-\\
-\leon  & 0 & 2 & 0 & 0 & 0 & {\win 9} & 0 & 0
-\\
-\isam & 0 & 35 & 0 & 0 & 0 & {\win 153} & 0 & 0
-\\
-\leonm  & 11  & 75 & 6 & 6 & 6 & {\win 169} & 6 & 6
-\\[\jot]
-Total & \phantom{0}11  & 112 & \phantom{00}6 & \phantom{00}6 & \phantom{00}6 & {\win 331} & \phantom{00}6 & \phantom{00}6
-\end{tabular}
-\caption{\,Number of \emph{sat} responses on benchmarks without and with translation with $\conv$}
-\label{fig:sat}
-\end{figure}
-
-\begin{figure}[t]
-\normalsize
-\centering
-\begin{tabular}{l@{\kern1.5em}r@{\kern0.375em}r@{\kern1.5em}r@{\kern0.375em}r@{\kern1.5em}r@{\kern0.375em}r@{\kern1.5em}r@{\kern0.375em}r@{\,\,}}
-  & \multicolumn{2}{c@{\kern1.5em}}{\phantom{0}\ziiib}     & \multicolumn{2}{c@{\kern1.5em}}{\phantom{0}\cvcd}     & \multicolumn{2}{c@{\kern1.5em}}{\phantom{0}\cvcf}     & \multicolumn{2}{c@{\,\,}}{\phantom{0}\cvci}
-\\%[-1pt]
-  & \hfill $\varphi$ \hfill & $\conv(\varphi)$\!\!
-  & \hfill $\varphi$ \hfill & $\conv(\varphi)$\!\!
-  & \hfill $\varphi$ \hfill & $\conv(\varphi)$\!\!
-  & \hfill $\varphi$ \hfill & $\conv(\varphi)$\!\!
-\\
-\midrule
-\isa & 14  & 15 & 15  & 15 & 15  & 15 & {\win 61}  & 14
-\\
-\leon  & 73  & 78 & 80  & 80 & 80  & 76 & {\win 96}  & 78
-\\
-\isam & 17  & 18 & 18  & 18 & 18  & 18 & {\win 44}  & 17
-\\
-\leonm  & 83  & 98 & 103 & 98 & 104 & 95 & {\win 117} & 98
-\\[\jot]
-Total & 187 & 209 & 216 & 211 & 217 & 204 & {\win 318} & 217
-\end{tabular}
-\caption{\,Number of \emph{unsat} responses on benchmarks without and with translation with $\conv$}
-\label{fig:unsat}
-\end{figure}
-
-The results are summarized in Figures \ref{fig:sat} and \ref{fig:unsat}.
-The benchmarks and more detailed results are available online.%
-\footnote{\url{http://lara.epfl.ch/~reynolds/SMT2015-recfun/}}
-The figures are divided into benchmarks triggering \emph{unsat} and \emph{sat}
-responses and further into benchmarks before and after the translation $\conv$.
-The raw evaluation data reveals no cases in which a solver answered
-\emph{unsat} on a benchmark $\varphi$ and \emph{sat} on its
-corresponding benchmark $\conv( \varphi )$, or vice versa.
-This is consistent with our expectations and Theorem~\ref{thm:equi}, 
-since these benchmarks contained only sets of function definitions
-that are supposed to be well-founded.
-%thereby confirming Theorem~\ref{thm:equi}.
-%\ct{?! \emph{Theorems} do not need corroboration! 
-%Maybe we should say "consistently with Theorem~\ref{thm:equi}"?}
-
-Figure~\ref{fig:sat} shows that for untranslated benchmarks (the ``$\varphi$''
-columns), the number of \emph{sat} responses is very low across all
-configurations. This confirms the shortcomings of existing SMT techniques for
-finding counterexamples for benchmarks containing recursively defined functions.
-%
-The translation $\conv$ (the ``$\conv(\varphi)$'' columns) has a major
-impact. \cvcf finds 331 of the 885 benchmarks to be satisfiable,
-including 9~benchmarks in the unmutated \leon benchmark set. The
-performance of \ziiib for counterexamples also improves dramatically, as it
-solves 112 satisfiable problems, including 10 that were unsolved by \cvcf.
-%
-We conclude that the translation $\conv$ enables SMT
-solvers to find counterexamples for conjectures involving recursively defined functions
-whose definitions are admissible.
-%Both \ziiib and \cvcf were able to find a
-%large number of counterexamples after the translation, with the latter
-%configuration finding significantly more instances (331 to 112).
-
-In terms of number of \emph{unsat} responses, the configuration \cvci is the
-clear winner, finding 318 total benchmarks to be unsatisfiable. Moreover, the
-translation $\conv$ helps \ziii for \emph{unsat} responses as well:
-\ziii solves a total of 209, whereas it solves only 187 before the
-translation.
-%
-In contrast, \cvcd finds 211 benchmarks unsatisfiable with the translation.
-Thus, in most cases the translation $\conv$ does not significantly degrade performance
-for unsatisfiable benchmarks, and in some cases may actually aid the solver for
-determining unsatisfiability. The exception is \cvci, whose inductive
-arguments are guided by the types of the quantified variables.
-
-\section{Front-End Support in \cvc}
-\label{sec:front-end}
-
-The translation $\conv$ significantly improves the effectiveness of known SMT
-techniques for finding counterexamples for formulas involving recursive
-function definitions.
-For convenience, it
-has been implemented in the development version of \cvc %~\cite{ReyEtAl-1-RR-13}
-(version~1.5 prerelease).
-Function definitions $\forallf{\con{f}} \vec x.\; \varphi$ can be written using the $\definefunreccmd$ command
-from the upcoming SMT-LIB 2.5 standard \cite{smtlib25}.
-For example, formula~(\ref{eq:ex-before}) from Example~\ref{ex:translation}
-can be specified as %follows:
-%
-\begin{quote}
-%\small -- no small, unless you convince me that these mixtures of font sizes
-% are aesthetic and serve a purpose --jb
-\begin{verbatim}
-(define-fun-rec s ((x Int)) Int (ite (<= x 0) 0 (+ x (s (- x 1)))))
-(declare-fun c () Int)
-(assert (> (s c) 100))
-(check-sat)
-\end{verbatim}
-\end{quote}
-%
-When reading this input,
-\cvc adds the annotated quantified formula
-$$\forallf{\con{s}} x.\;\, \con{s}( x ) \teq \lite\bigl( x \leq 0,\; 0,\; \con{s}( x-1 )\bigr)$$
-to its list of assertions,
-which after rewriting becomes
-$$\forallf{\con{s}} x.\;\, \lite\bigl( x \leq 0,\; \con{s}( x ) \teq 0,\; \con{s}( x ) \teq \con{s}( x-1 ) \bigr)$$
-If \cvc's finite model finding mode for recursive functions is enabled (using
-the command-line option \texttt{--fmf-fun}), it will replace its list of known
-assertions based on the translation $\conv$ before checking for satisfiability.
-%If the functions provided in the input are admissible,
-%then a ``satisfiable" response from the solver implies that a model exists for the original formula.
-Accordingly, the solver will output the approximation of the interpretation it
-used for recursive function definitions. For the example above, it outputs
-%\ajr{prefer having the full output like this here?}
-%\jb{I like it the way you did it now, but I took the freedom to reformat the output slightly
-%(to avoid the small font---I generally try to avoid ad hoc font mixtures---and to show
-%the nesting more clearly).}
-%it may output the interpretation~(\ref{eq:approx-interp}) from Section~\ref{sec:encoding} for $s$.
-%
-\begin{quote}
-%\small
-\begin{verbatim}
-(model
-  (define-fun s (($x1 Int)) Int 
-     (ite (= $x1 14) 105 (ite (= $x1 13) 91 (ite (= $x1 12) 78 
-        (ite (= $x1 11) 66 (ite (= $x1 10) 55 (ite (= $x1 4) 10 
-           (ite (= $x1 9) 45 (ite (= $x1 8) 36 (ite (= $x1 7) 28 
-              (ite (= $x1 6) 21 (ite (= $x1 3) 6 (ite (= $x1 5) 15 
-                 (ite (= $x1 2) 3 (ite (= $x1 1) 1 0)))))))))))))))
-  (define-fun c () Int 14)
-)
-\end{verbatim}
-\end{quote}
-
-The \texttt{--fmf-fun} option instructs \cvc to assume that the
-functions provided by the user are admissible.
-Admissibility must be checked separately \relax{by the user}, perhaps using a
-syntactic criterion or a termination prover.
-If this criterion is not met, \cvc may answer \emph{sat} on
-an unsatisfiable problem. Indeed, if we add the inconsistent definition
-%
-\begin{quote}
-%\small
-\begin{verbatim}
-(define-fun-rec h ((x Int)) Int (+ (h x) 1))
-\end{verbatim}
-\end{quote}
-%
-to the above problem and run \cvc with the \texttt{--fmf-fun} option,
-it wrongly answers \emph{sat}.
-
-%Usually not interested in showing logical consistency, there
-%exist other tools that check e.g. termination.
-%Thus, want SMT solver to check satisfiability the assumption that recursive
-%functions are consistent (where this needs to be defined precisely).
-
-\section{Related Work}
-
-We described the most closely related work, by Ge and de Moura
-\cite{GeDeM-CAV-09} and by Reynolds et al.\
-\cite{ReyEtAl-1-RR-13,reynolds-et-al-2013}, in the text already.
-The finite model finding support in the instantiation-based iProver
-\cite{korovin-2013} is also close, given the similarities with SMT.
-
-%Model finding has been studied outside the world of SMT.
-
-Some finite model finders are based on a reduction to a decidable logic,
-typically propositional logic. They translate the input problem to the weaker
-logic and pass it to a solver for that logic.
-The translation is parameterized by upper or exact finite bounds on
-the cardinalities of the atomic types. This procedure was pioneered by McCune
-in the earlier versions of Mace (also capitalized MACE) \cite{mccune-1994}.
-Other conceptually similar finders are Paradox \cite{claessen-sorensson-2003}
-and FM-Darwin \cite{baumgartner-et-al-2009} for first-order logic with
-equality; the Alloy Analyzer and its back-end Kodkod \cite{torlak-jackson-2007}
-for first-order relational logic; and Refute \cite{weber-2008} and Nitpick
-\cite{blanchette-nipkow-2010} for higher-order logic.
-
-An alternative to translating the problem is to perform
-an exhaustive model search directly on the original problem. Given fixed
-cardinalities, the search space is represented in memory as multidimensional
-arrays. The procedure tries different values in the function and predicate
-table entries, checking each time if the problem is satisfied and backtracking
-if necessary. This approach was pioneered by FINDER \cite{slaney-1994} and
-SEM \cite{zhang-zhang-1995} and serves as the basis of many more model finders,
-notably the Alloy Analyzer's precursor \cite{jackson-1996} and the later
-versions of~Mace \cite{mccune-prover9-mace4}.
-
-Most of the above tools cannot cope with algebraic datatypes or other infinite
-types. Kuncak and Jackson \cite{kuncak-jackson-2005} presented an idiom for
-encoding datatypes and recursive functions in Alloy, by approximating datatypes
-by finite subterm-closed substructures. The approach finds sound (fragments
-of) models for formulas in the \relax{existential--bounded-universal} fragment
-(i.e., formulas whose prenex normal forms contain no unbounded universal
-quantifiers ranging over datatypes). This idiom was further developed by Dunets
-et al.\ \cite{dunets-et-al-2010}, who presented a translation scheme
-for primitive recursion. Their definedness guards play a similar role to the
-existential constraints generated by our translation $\conv$.
-
-The higher-order model finder Nitpick for the Isabelle/HOL proof assistant
-relies on another variant of Kuncak and Jackson's approach inside a
-Kleene-style three-valued logic, inspired by abstract interpretation. It was
-also the first tool of its kind to support corecursion and coalgebraic datatypes
-\cite{blanchette-2013-relational}. The three-valued logic approach extends each
-approximated type with an unknown value, which is propagated by function
-application. This approach works reasonably well in Nitpick, because it builds
-on Kodkod's relational logic, but our initial experiments with \cvc convinced
-us that it is more efficient to avoid unknown values by adding existential
-constraints.
-
-The Leon system~\cite{blanc2013overview} implements a procedure that can both
-produces proofs and counterexamples for properties of terminating functions
-written in a subset of Scala. Leon is based on an SMT solver. It avoids
-quantifiers altogether by unfolding recursive definitions up to a certain
-depth, which is increased on a per-need basis.
-The approach described in this paper works in an analogous manner, 
-where instead the SMT solver is invoked only once 
-and where quantifier instantiation amounts to function unfolding. It would be
-worth investigating how existing approaches for algorithms based on function
-unfolding can inform approaches for dedicated quantifier instantiation
-techniques for function definitions, and vice versa.
-
-Model finding is concerned with satisfying arbitrary logical constraints. Some
-tools are tailored for problems that correspond to total functional
-programs. The QuickCheck tool for Haskell \cite{claessen-hughes-2000} is an
-early example, based
-on random testing. Bounded exhaustive testing \cite{runciman-et-al-2008} and
-narrowing \cite{lindblad-2008-testing} are other successful strategies. These
-tool are often much faster than model finders, but they typically cannot cope
-with underspecification and nonexecutable functions.
-
-% others
-%are Agsy for Agda \cite{xxx}, Quickcheck for Isabelle/HOL \cite{xxx}, and
-% QuickChick for Coq \cite{xxx}.
-
-%Unlike an SMT solver, Nitpick cannot rely on a built-in notion of
-%a datatype; it axiomatizes finite subterm-closed substructures.
-
-%\begin{verbatim}
-%Nitpick
-%  * nice: three-valued logic (e.g. local overflow)
-%    * solving a different problem:
-%      * no built-in notion, what's approximated is an idealized notion, no
-%        injections $\concret$
-%  * not so nice:
-%    one abstract domain per type, not per function argument
-%  * guards + one-domain per arg are a good combi, guards + one-single-domain is bad because of
-%    odd schemes, give unsat example
-%    [* show that it works also for example with weird recursion schemes]
-%
-%  * other features:
-%    * inductive / coinductive predicates
-%    * quotient types
-%
-%  * abstract interpretation
-%
-%Leon
-%\end{verbatim}
-
-\section{Conclusion}
-\label{sec:conclusion}
-
-We presented a translation scheme that makes more SMT problems amenable to
-finite model findings in SMT. In conjunction with finite model finding
-techniques,
-%introduced by Reynolds et al.\ \cite{ReyEtAl-1-RR-13,reynolds-et-al-2013},
-%%% Great work, but let's not cite it more than 4 times if it can be avoided. --jb
-it makes it possible to
-find models of quantified formulas over infinite types, such as integers
-and algebraic datatypes.
-
-In future work, it would be interesting to evaluate the approach against other
-counterexample generators, notably Leon and Nitpick, and enrich the benchmark
-suite with problems exercising \cvc's support for coalgebraic datatypes
-\cite{reynolds-blanchette-2015-codata}. We also plan to integrate \cvc as a
-counterexample generator in proof assistants. Finally, future work could also
-include identifying further sufficient conditions for admissibility, thereby
-enlarging the applicability of the translation scheme presented here.
-
-%\ct{?? Leon is a model finder? Alloy?}
-%\jb{Alloy has no notion of datatypes, so it's pretty useless. Nitpick implements
-%datatypes and a more FOL-like logic (actually, HOL) on top of Alloy's backend,
-%using the Kuncak \& Jackson idiom for Alloy, so that's as close as we can get
-%to evaluating Alloy itself. Leon is, in part, a counterexample generator.
-%``Model finder'' is probably a stretch.}
-
-{%\footnotesize
-\def\ackname{Acknowledgment}
-\paragraph{%\footnotesize
-\ackname.}
-Viktor Kuncak and Stephan Merz have made this work possible. We would also like
-to thank Damien Busato-Gaston and Emmanouil Koukoutos for providing the initial
-set of benchmarks used in the evaluation.
-}
-
-{
-\bibliographystyle{abbrv}
-\bibliography{bib}
-}
-
-\end{document}

### 2015-05-08 tex/body.tex
Leave front-end model, exclude text surrounding theorem 2.

@@
 
+\begin{longv}
 In the following, we sketch the correctness of our translation $\conv$.
@@
 \end{lemma}
-\begin{longv}
 \begin{proofsketch}
@@
 \end{proofsketch}
-\end{longv}
 
@@
 \end{cor}
+\end{longv}
 
@@
 %For clarity, we demonstrate the consequences of this theorem by means of an example.
+\end{longv}
 
@@
 used for recursive function definitions. 
-\begin{longv}
 For the example above, it outputs
@@
 \end{quote}
-\end{longv}
-\begin{shortv}
-\ajr{todo}
-\end{shortv}
 
@@
 If this criterion is not met, \cvc may answer \emph{sat} on
-an unsatisfiable problem. Indeed, if we add the inconsistent definition
+an unsatisfiable problem. 
+\begin{longv}
+Indeed, if we add the inconsistent definition
 %
@@
 it wrongly answers \emph{sat}.
+\end{longv}
 

### 2015-05-08 tex/body.tex
more

--- 2015-05-08 tex/body.tex
@@
 algebraic datatypes. A simple example where both approaches fail is
-\begin{equation}\label{eq:pii}
-\forall x : {\typeint}.\;\, \lite\bigl(
-x \leq 0,\allowbreak\; \con{p}( x ) \teq 1,\allowbreak\; \con{p}( x ) \teq 2 * \con{p}( x - 1 ) \bigr)
-\end{equation}
-This state of affairs is highly unsatisfactory, given the frequency of
+$\forall x : {\typeint}.\;\, \lite\bigl(
+x \leq 0,\allowbreak\; \con{p}( x ) \teq 1,\allowbreak\; \con{p}( x ) \teq 2 * \con{p}( x - 1 ) \bigr)$.
+This state of affairs is unsatisfactory, given the frequency of
 recursive definitions in practice and the impending addition of a dedicated
@@
 (Section~\ref{sec:prelim}). This criterion is met by well-founded (terminating)
-recursive function definitions. It is not met by inconsistent
+recursive function definitions. \begin{longv}It is not met by inconsistent
 definitions such as $\forall x : {\typeint}.\;\, \con{f}(x) \teq \con{f}(x) +
-1$. %ct and it is possible to construct consistent examples that do not meet it.
+1$.\end{longv} %ct and it is possible to construct consistent examples that do not meet it.
 %as well as other example of consistent definition.
@@
 \fargtype{\tau}.\;\, \const{f}(\farg{\con{f}}(a)) \teq t[\concret(a)/x]$, where
-$\fargtype{\tau}$ is an uninterpreted \relax{abstract} type and $\farg{\con{f}} :
-\fargtype{\tau} \to \tau$ converts the abstract type to the concrete
-type. The translation preserves satisfiability and, for admissible definitions,
+$\fargtype{\tau}$ is an uninterpreted \relax{abstract} type and $\farg{\con{f}} \begin{longv}:
+\fargtype{\tau} \to \tau\end{longv}$ converts the abstract type to the concrete
+type. The translation preserves satisfiability and\begin{longv}, for admissible definitions,\end{longv}
 unsatisfiability, and makes finite model finding possible for problems in this
@@
 (Section~\ref{sec:evaluation}). The approach is implemented as a preprocessor
-in \cvc (Section~\ref{sec:front-end}). When invoked with a specific
+in \cvc (Section~\ref{sec:front-end}).\begin{longv} When invoked with a specific
 command-line option, the solver assumes that functions introduced using the
-\texttt{define-fun-rec} command are admissible and translates them accordingly.
+\texttt{define-fun-rec} command are admissible and translates them accordingly.\end{longv}
 
@@
 \emph{$T$-satisfiable}
-if it is satisfied by some interpretation in $\Mo$;
-otherwise, it is \emph{$T$-unsatisfiable}.
+if it is satisfied by some interpretation in $\Mo$\begin{longv};
+otherwise, it is \emph{$T$-unsatisfiable}\end{longv}.
 A formula $\varphi$ \emph{$T$-entails} $\psi$, written $\varphi \models_T \psi$,
@@
 %
-where $\con{f}_1, \ldots, \con{f}_n$ are distinct function symbols,
+where $\con{f}_1, \ldots, \con{f}_n$ are distinct\begin{longv} function symbols\end{longv},
 %ct added
@@
 the function would terminate when seen as a functional program, under \emph{some}
-evaluation order. For example, definition~(\ref{eq:pii}) from the
-introduction can be shown well-founded under an evaluation strategy that
+evaluation order. For example, the definition
+$\forall x : {\typeint}.\;\, \lite\bigl(
+x \leq 0,\allowbreak\; \con{p}( x ) \teq 1,\allowbreak\; \con{p}( x ) \teq 2 * \con{p}( x - 1 ) \bigr)$
+can be shown well-founded under an evaluation strategy that
 evaluates the condition of a $\lite$ before evaluating the relevant branch,
@@
 \begin{theorem}\label{thm:adm}
-If\/
-$\DDD$ is a set of well-founded function definitions for\/
-$\sfundefs{\Sigma} = \{\con{f}_1,\ldots,\con{f}_n\}$, then it is admissible.
+If $\DDD$ is a set of well-founded function definitions for\/
+$\sfundefs{\Sigma} \begin{longv} = \{\con{f}_1,\ldots,\con{f}_n\}\end{longv}$, then it is admissible.
 \end{theorem}

### 2015-05-08 tex/body.tex
Minor

@@
 call these sets \isam and \leonm, respectively.
-
 Thus, our benchmark set consist of 885 benchmarks (640 mutants plus the
@@
 When reading this input,
+\begin{longv}
 \cvc adds the annotated quantified formula
-\begin{longv}
 $$\forallf{\con{s}} x.\;\, \con{s}( x ) \teq \lite\bigl( x \leq 0,\; 0,\; \con{s}( x-1 )\bigr)$$
@@
 \begin{shortv}
+\cvc adds
 $\forallf{\con{s}} x.\;\, \con{s}( x ) \teq \lite\bigl( x \leq 0,\; 0,\; \con{s}( x-1 )\bigr)$
@@
 which after rewriting becomes
-$\forallf{\con{s}} x.\;\, \lite\bigl( x \leq 0,\; \con{s}( x ) \teq 0,\; \con{s}( x ) \teq \con{s}( x-1 ) \bigr)$
-\end{shortv}
+$\forallf{\con{s}} x.\;\, \lite\bigl( x \leq 0,\; \con{s}( x ) \teq 0,\; \con{s}( x ) \teq \con{s}( x-1 ) \bigr)$.\end{shortv}
 If \cvc's finite model finding mode for recursive functions is enabled (using

### 2015-05-08 tex/body.tex
shorten

--- 2015-05-08 tex/body.tex
@@
 
+\begin{longv}
 Another useful fragment of function definitions is the class of
@@
 or productivity, without affecting admissibility.
+\end{longv}
+
+\newcommand\badassex{%
+ \{\forallf{\con{f}} x.\;\, \con{f}( x ) \teq \con{f}( x ),\;\,
+ \forallf{\con{g}} x.\;\, \lite( \con{f}( x ) \teq 5,\;
+   \con{g}( x ) \teq \con{f}( x ),\;
+   \con{g}( x ) \teq \con{g}( x ) + 1)\}}
 
@@
 satisfying $\con{f}$'s definition. A more subtle specimen is
-\[\{ 
- \forallf{\con{f}} x.\;\, \con{f}( x ) \teq \con{f}( x ),\;\,
- \forallf{\con{g}} x.\;\, \lite( \con{f}( x ) \teq 5,\;
-   \con{g}( x ) \teq \con{f}( x ),\,
-   \con{g}( x ) \teq \con{g}( x ) + 1)
-\}\]
+\begin{shortv}$\badassex$.\end{shortv}%
+\begin{longv}\[\badassex\]\end{longv}
 While this set has a model where $\con f$ and $\con g$ are interpreted as the
@@
 $\con{f}( 0 ) \teq 4$ is
-closed under function expansion with respect to this set and yet there exists
+closed under function expansion \begin{longv}with respect to this set\end{longv}
+and yet there exists
 no interpretation satisfying both $\con{f}( 0 ) \teq 4$ and $\con{g}$'s

### 2015-05-08 tex/body.tex
shorten

@@
 
-\renewcommand\models{\mathrel{\vDash}}
+\renewcommand\models{\mathrel{\vthinspace\vDash}}
 
@@
 
+\newcommand\closurefmla{\psi \models_T \bigwedge\nolimits_{\,i=1}^{n} \{ \psi_i[ \vec t / \vec x ] \mid \con{f}_i( \vec t ) \in \terms( \psi ) \}}
+
 \begin{definition}\rm
@@
 is \emph{closed under function expansion with respect to $\DDD$} if
-\[
- \psi \,\models_T\, \bigwedge\nolimits_{\,i=1}^{n} \{ \psi_i[ \vec t / \vec x ] \mid \con{f}_i( \vec t ) \in \terms( \psi ) \} %.
-\]
+\begin{shortv}$\closurefmla$.\end{shortv}%
+\begin{longv}\[\closurefmla\]\end{longv}
 The set $\DDD$ is \emph{admissible} if for every $T$-satisfiable formula 

### 2015-05-08 tex/body.tex
more shortening

@@
 \boolop, i, p )$, where $\pol$ is defined as
-\begin{equation*}
-\pol( \boolop, i, p ) =
+\[\pol( \boolop, i, p ) =
 \begin{cases}
@@
 \pnone & \text{otherwise}
-\end{cases}
-\end{equation*}
+\end{cases}\]
 where the negation $\pnegate p$ of a polarity $p$ is $\pneg$ if $p$ is $\ppos$, $\ppos$
@@
 If $t$ is an unannotated quantified formula, $\conv_0$ calls itself
-on the body with the same polarity. A quantifier is prefixed to the
-quantified formula and constraint returned by the recursive call.
+on the body with the same polarity. \begin{longv}A quantifier is prefixed to the
+quantified formula and constraint returned by the recursive call.\end{longv}
 Otherwise, $t$ is either an application of an uninterpreted predicate or a term
@@
 
+\newcommand\badlambda{\lambda x : \typeint.\allowbreak\;\, \lite( x \teq 0,\allowbreak\; 0,\allowbreak\;
+  \lite( x \teq 1,\allowbreak\; 1,\allowbreak\;
+    \lite( x \teq 2,\allowbreak\; 3,\allowbreak\;
+      \lite(\ldots,\allowbreak\; \lite( x \teq 13,\allowbreak\, 91,\allowbreak\, 105 )\negvthinspace \ldots ))))}
+
 The translation $\conv$ results in formulas whose models are different than those of $\varphi$.
@@
 %\begin{equation} \label{eq:approx-interp}
-\[\lambda x : \typeint.\;\, \lite( x \teq 0,\; 0,\;
-  \lite( x \teq 1,\; 1,\;
-    \lite( x \teq 2,\; 3,\;
-      \lite(\ldots,\; \lite( x \teq 13,\, 91,\, 105 )\negvthinspace \ldots ))))\]
+\begin{shortv}$\badlambda$.\end{shortv}%
+\begin{longv}\[\badlambda\]\end{longv}
 %\end{equation}
@@
 $\sfundefs{\Sigma}$
-and the set of function definitions\/ $\DDD$ corresponding to functions in\/
+and the set of function definitions\/ $\DDD$ corresponding to\/
 $\sfundefs{\Sigma}$ is admissible, then\/ $\varphi$ and\/ $\conv( \varphi )$ are
@@
 Hence, by repeated application of this reasoning, it follows that
-a model of formula~(\ref{eq:ex-after}) that interprets $\con{c}$ as $n$ must also satisfy the formula $\psi$:
+a model of formula~(\ref{eq:ex-after}) that interprets $\con{c}$ as $n$ must also satisfy% the formula
+~$\psi$:
 %
@@
 Some finite model finders are based on a reduction to a decidable logic,
-typically propositional logic. They translate the input problem to the weaker
-logic and pass it to a solver for that logic.
+typically propositional logic. \begin{longv}They translate the input problem to the weaker
+logic and pass it to a solver for that logic.\end{longv}
 The translation is parameterized by upper or exact finite bounds on
 the cardinalities of the atomic types. This procedure was pioneered by McCune
-in the earlier versions of Mace (also capitalized MACE) \cite{mccune-1994}.
+in the earlier versions of Mace (or MACE) \cite{mccune-1994}.
 Other conceptually similar finders are Paradox \cite{claessen-sorensson-2003}
@@
 if necessary. This approach was pioneered by FINDER \cite{slaney-1994} and
-SEM \cite{zhang-zhang-1995} and serves as the basis of many more model finders,
+SEM \cite{zhang-zhang-1995}\begin{longv} and serves as the basis of many more model finders,
 notably the Alloy Analyzer's precursor \cite{jackson-1996} and the later
-versions of~Mace \cite{mccune-prover9-mace4}.
-
+versions of~Mace \cite{mccune-prover9-mace4}\end{longv}.
+
+\begin{longv}
 Most of the above tools cannot cope with algebraic datatypes or other infinite
-types. Kuncak and Jackson \cite{kuncak-jackson-2005} presented an idiom for
+types. \end{longv}%
+Kuncak and Jackson \cite{kuncak-jackson-2005} presented an idiom for
 encoding datatypes and recursive functions in Alloy, by approximating datatypes
@@
 of) models for formulas in the \relax{existential--bounded-universal} fragment
-(i.e., formulas whose prenex normal forms contain no unbounded universal
-quantifiers ranging over datatypes). This idiom was further developed by Dunets
+\begin{longv}(i.e., formulas whose prenex normal forms contain no unbounded universal
+quantifiers ranging over datatypes)\end{longv}. This idiom was further developed by Dunets
 et al.\ \cite{dunets-et-al-2010}, who presented a translation scheme
@@
 
-The higher-order model finder Nitpick for the Isabelle/HOL proof assistant
+The higher-order model finder Nitpick \cite{blanchette-nipkow-2010}
+for the Isabelle/HOL proof assistant
 relies on another variant of Kuncak and Jackson's approach inside a
-Kleene-style three-valued logic, inspired by abstract interpretation. It was
-also the first tool of its kind to support corecursion and coalgebraic datatypes
-\cite{blanchette-2013-relational}. The three-valued logic approach extends each
+Kleene-style three-valued logic\begin{longv}, inspired by abstract interpretation\end{longv}.
+\begin{longv}It was also the first tool of its kind to support corecursion and
+coalgebraic datatypes \cite{blanchette-2013-relational}.\end{longv}
+The three-valued logic approach extends each
 approximated type with an unknown value, which is propagated by function
 application. This approach works reasonably well in Nitpick, because it builds
-on Kodkod's relational logic, but our initial experiments with \cvc convinced
-us that it is more efficient to avoid unknown values by adding existential
+on a relational logic, but our initial experiments with \cvc suggest
+that it is more efficient to avoid unknown values by adding existential
 constraints.
@@
 where instead the SMT solver is invoked only once 
-and where quantifier instantiation amounts to function unfolding. It would be
+and where quantifier instantiation amounts to function unfolding.
+\begin{longv}It would be
 worth investigating how existing approaches for function
 unfolding can inform approaches for dedicated quantifier instantiation
-techniques for function definitions, and vice versa.
+techniques for function definitions, and vice versa.\end{longv}
 

### 2015-05-08 tex/body.tex
more
Add necessary intro to Ex 2.

@@
 unsatisfiability, and makes finite model finding possible for problems in this
-class.
+class. Detailed proofs are included in a technical report \cite{our-report}.
 
@@
 
+\begin{shortv}
+In following, we give an intuition of the above theorem in the context of an example.
+For a set of ground $\Sigma$-literals $L$, let $\absconstraints( L )$ be the set
+$\{ \exists a : {\fargtype{\con{f}}}.\;\, \vecfarg{\con{f}}( a ) \teq \vec t
+\mid \con{f}( \vec t ) \in \terms( L ),\: \con{f} \in \sfundefs{\Sigma}
+\}$.
+\end{shortv}
+
 \begin{example}
@@
 %any model of $\varphi_{\con{s}}[t/x]$ must also satisfy $\varphi_{\con{s}}[t-1/x]$ when $t > 0$. 
-Clearly any interpretation $\I$ satisfying formula~(\ref{eq:ex-after}) satisfies $L_0 \cup \absconstraints( L_0 )$,
+Clearly any interpretation $\I$ satisfying formula~(\ref{eq:ex-after}) satisfies 
+$L_0 \cup \absconstraints( L_0 )$,
 where $L_0 = \{ \con{s}( \con{c} ) > 100 \}$, 

### 2015-05-08 tex/body.tex
Remove some text from Sec 3

@@
 $\pnone$. \vthinspace$\conv_0$ returns a pair $( t'\negvthinspace,\, \chi )$, where $t'$ is a term of
-type $\tau$ and $\chi$ is an additional constraint.
-This constraint is needed if the polarity is $\pnone$; otherwise, it is encoded
-directly in $t'$ and $\chi$ is simply $\ltrue$.
-In the auxiliary function $\conv_0$, the function $\pol(
-\boolop, i, p )$ returns the polarity of the $i$\vvthinspace{th} argument of an
-application of~$\boolop$ having polarity $p$ if $\boolop$ is an interpreted
-predicate or $\pnone$ otherwise.
+the same type as $t$, and $\chi$ is a $\extendsig{\Sigma}$-formula.
+%an additional constraint.
+%This constraint is needed if the polarity is $\pnone$; otherwise, it is encoded
+%directly in $t'$ and $\chi$ is simply $\ltrue$.
+%In the auxiliary function $\conv_0$, the function $\pol(
+%\boolop, i, p )$ returns the polarity of the $i$\vvthinspace{th} argument of an
+%application of~$\boolop$ having polarity $p$ if $\boolop$ is an interpreted
+%predicate or $\pnone$ otherwise.
 
@@
 the first conjunct of formula~(\ref{eq:ex-after})---and $\absconstraints( L_0
-)$, it must also satisfy the formula $\psi_1$:
+)$, it must also satisfy\begin{longv} the formula $\psi_1$\end{longv}:
 \[

### 2015-05-08 tex/body.tex
Minor
minor

@@
 translation~$\conv$. % as described in Section~\ref{sec:encoding}.
-For example, \isa contains 79 original benchmarks $\varphi$ and 79 translated
+For example, \isa contains 79 original benchmarks $\varphi$ and 213 translated
 benchmarks $\conv(\varphi)$.
@@
 impact. \cvcf finds 331 of the 885 benchmarks to be satisfiable,
-including 9~benchmarks in the unmutated \leon benchmark set. The
+including 9~benchmarks in the nonmutated \leon benchmark set. The
 performance of \ziiib for counterexamples also improves dramatically, as it

### 2015-05-08 tex/body.tex
minor

@@
 x \leq 0,\allowbreak\; \con{p}( x ) \teq 1,\allowbreak\; \con{p}( x ) \teq 2 * \con{p}( x - 1 ) \bigr)$
-can be shown well-founded under an evaluation strategy that
+can be shown well-founded under an strategy that
 evaluates the condition of a $\lite$ before evaluating the relevant branch,

### 2015-05-08 tex/body.tex
minor

@@
 $\I$ for formula~(\ref{eq:ex-after}) given above interprets $\con{c}$ as $14$
-and $\con{s}(n) = \sum_{i=1}^n i$ for $0 \le n \le 14$,
+and $\con{s}(n)$ as $\sum_{i=1}^n i$ for $0 \le n \le 14$,
 and there exists a model of formula~(\ref{eq:ex-before}) that also interprets

### 2015-05-08 tex/body.tex
Revert fix.
More minor.

--- 2015-05-08 tex/body.tex
@@
 x \leq 0,\allowbreak\; \con{p}( x ) \teq 1,\allowbreak\; \con{p}( x ) \teq 2 * \con{p}( x - 1 ) \bigr)$
-can be shown well-founded under an strategy that
+can be shown well-founded under a strategy that
 evaluates the condition of a $\lite$ before evaluating the relevant branch,
@@
 translation~$\conv$. % as described in Section~\ref{sec:encoding}.
-For example, \isa contains 79 original benchmarks $\varphi$ and 213 translated
+For example, \isa contains 79 original benchmarks $\varphi$ and 79 translated
 benchmarks $\conv(\varphi)$.
@@
 narrowing \cite{lindblad-2008-testing} are other successful strategies. These
-tool are often much faster than model finders, but they typically cannot cope
+tools are often much faster than model finders, but they typically cannot cope
 with underspecification and nonexecutable functions.

### 2015-05-08 tex/body.tex
tuning

--- 2015-05-08 tex/body.tex
@@
 unsatisfiability, and makes finite model finding possible for problems in this
-class. Detailed proofs are included in a technical report \cite{our-report}.
+class.\begin{shortv} Detailed proofs of correctness are included in a technical
+report \cite{our-report}.\end{shortv}
 
@@
 
-In addition to standard quantified formulas $\forall \vec x.\; \varphi$,
+In addition to $\forall \vec x.\; \varphi$,
 we consider \emph{annotated quantified formulas} of the form
 $\forallf{\const{f}} \vec x.\; \varphi$, where $\con{f} \in \sfuns{\Sigma}$ is
-an uninterpreted function symbol. Their
+uninterpreted. Their
 semantics is the same as for standard quantified formulas.
@@
 %since they aren't operational, that's why I put quantifier-free. --ajr
+%%\jb{Actually, in Isabelle we can have them and we use them sometimes. It certainly
+%% makes sense. I can give you real examples if you like.}
 We write $\exists
-\vec x.\; \varphi$ as a shorthand for $\neg \forall \vec x.\; \neg \varphi$.
+\vec x.\; \varphi$ as an abbreviation for $\neg \forall \vec x.\; \neg \varphi$.
 
@@
 evaluates the condition of a $\lite$ before evaluating the relevant branch,
-ignoring the other branch. Logically, such dependencies can be captured by
-congruence rules. Krauss developed these ideas in the more general context of
+ignoring the other branch. \begin{longv}Logically, such dependencies can be captured by
+congruence rules. \end{longv}Krauss developed these ideas in the more general context of
 higher-order logic \cite[Section 2]{krauss-2009-phd}.
@@
  \{\forallf{\con{f}} x.\;\, \con{f}( x ) \teq \con{f}( x ),\;\,
- \forallf{\con{g}} x.\;\, \lite( \con{f}( x ) \teq 5,\;
-   \con{g}( x ) \teq \con{f}( x ),\;
+ \forallf{\con{g}} x.\;\, \lite( \con{f}( x ) \teq 5,\allowbreak\;
+   \con{g}( x ) \teq \con{f}( x ),\allowbreak\;
    \con{g}( x ) \teq \con{g}( x ) + 1)\}}
@@
      \itemx[] \return \bigtuple{$t,\; \bigwedge\, \{ \exists a : \fargtype{\con{f}}.\; \vecfarg{\con{f}}( a ) \teq \vec t \mid \con{f}( \vec t ) \in \terms( t ),\, \con{f} \in \sfundefs{\Sigma} \}$}
+\\[-.8\baselineskip]
    \end{itemize}
@@
 constraint $\chi$ is simply returned to the caller.
-
+%
+\begin{longv}\par\end{longv}
+%
 If $t$ is a function definition of a function in
@@
 $\I$ for formula~(\ref{eq:ex-after}) given above interprets $\con{c}$ as $14$
-and $\con{s}(n) = \sum_{i=1}^n i$ for $0 \le n \le 14$,
+and $\con{s}(n)$ as $\sum_{i=1}^n i$ for $0 \le n \le 14$,
 and there exists a model of formula~(\ref{eq:ex-before}) that also interprets
-$\con{c}$ and and $\con{s}(n)$ in the same way (for $0 \le n \le 14$).
-In general, there exists a model of $\varphi$ that
-coincides with each model of $\conv( \varphi )$ on its interpretation of all
-functions in $\sfuns{\Sigma} - \sfundefs{\Sigma}$, i.e., $\{
-\con{s} \}$ in the example. Furthermore, the model of $\conv( \varphi )$ will
+$\con{c}$ and $\con{s}(n)$ in the same way (for $0 \le n \le 14$).
+In general, for every model of $\conv( \varphi )$,
+there exists a model of $\varphi$ that
+coincides with it on its interpretation of all
+function symbols in $\sfuns{\Sigma} - \sfundefs{\Sigma}$ ($\{
+\con{s} \}$ in the example). Furthermore, the model of $\conv( \varphi )$ will
 also give correct information for the defined functions at all points belonging
@@
 the first conjunct of formula~(\ref{eq:ex-after})---and $\absconstraints( L_0
-)$, it must also satisfy\begin{longv} the formula $\psi_1$\end{longv}:
+)$, it must also satisfy\begin{longv} the formula $\psi_1$\end{longv}
 \[
@@
                  (ite (= $x1 2) 3 (ite (= $x1 1) 1 0)))))))))))))))
-  (define-fun c () Int 14)
-)
+  (define-fun c () Int 14))
 \end{verbatim}
@@
 functions provided by the user are admissible.
-Admissibility must be checked separately \relax{by the user}, perhaps using a
+Admissibility must be checked \relax{by the user}, perhaps using a
 syntactic criterion or a termination prover.
-If this criterion is not met, \cvc may answer \emph{sat} on
+If it is not met, \cvc may answer \emph{sat} for
 an unsatisfiable problem. 
@@
 the cardinalities of the atomic types. This procedure was pioneered by McCune
-in the earlier versions of Mace (or MACE) \cite{mccune-1994}.
+in the earlier versions of
+\begin{shortv}MACE\end{shortv}\begin{longv}Mace (or MACE)\end{longv}
+\cite{mccune-1994}.
 Other conceptually similar finders are Paradox \cite{claessen-sorensson-2003}
@@
 
-An alternative to translating the problem is to perform
+An alternative is to perform
 an exhaustive model search directly on the original problem. Given fixed
-cardinalities, the search space is represented in memory as multidimensional
-arrays. The procedure tries different values in the function and predicate
-table entries, checking each time if the problem is satisfied and backtracking
-if necessary. This approach was pioneered by FINDER \cite{slaney-1994} and
-SEM \cite{zhang-zhang-1995}\begin{longv} and serves as the basis of many more model finders,
-notably the Alloy Analyzer's precursor \cite{jackson-1996} and the later
-versions of~Mace \cite{mccune-prover9-mace4}\end{longv}.
+cardinalities, the search space is represented as multidimensional
+tables. The procedure tries different values in the function and predicate
+tables, checking each time if the problem is satisfied.
+This approach was pioneered by FINDER
+\cite{slaney-1994} and SEM \cite{zhang-zhang-1995}\begin{longv} and serves as
+the basis of many more model finders, notably the Alloy Analyzer's precursor
+\cite{jackson-1996} and the later versions of~Mace
+\cite{mccune-prover9-mace4}\end{longv}.
 
@@
 programs. The QuickCheck tool for Haskell \cite{claessen-hughes-2000} is an
-early example, based
-on random testing. Bounded exhaustive testing \cite{runciman-et-al-2008} and
-narrowing \cite{lindblad-2008-testing} are other successful strategies. These
-tool are often much faster than model finders, but they typically cannot cope
-with underspecification and nonexecutable functions.
+early example, based on random testing. Bounded exhaustive
+testing\begin{longv} \cite{runciman-et-al-2008}\end{longv} and
+narrowing\begin{longv} \cite{lindblad-2008-testing}\end{longv} are other
+successful strategies. These tool are often much faster than model finders, but
+they typically cannot cope with underspecification and nonexecutable functions.
 

### 2015-05-08 tex/body.tex

@@
 programs. The QuickCheck tool for Haskell \cite{claessen-hughes-2000} is an
-early example, based on random testing. Bounded exhaustive
-testing\begin{longv} \cite{runciman-et-al-2008}\end{longv} and
-narrowing\begin{longv} \cite{lindblad-2008-testing}\end{longv} are other
-successful strategies. These tool are often much faster than model finders, but
-they typically cannot cope with underspecification and nonexecutable functions.
+early example, based on random testing. Bounded exhaustive testing\begin{longv}
+\cite{runciman-et-al-2008}\end{longv} and narrowing\begin{longv}
+\cite{lindblad-2008-testing}\end{longv} are other successful strategies. These
+tools are often much faster than model finders, but they typically cannot cope
+with underspecification and nonexecutable functions.
 

### 2015-05-08 tex/body.tex
more

@@
 the first conjunct of formula~(\ref{eq:ex-after})---and $\absconstraints( L_0
-)$, it must also satisfy\begin{longv} the formula $\psi_1$\end{longv}
+)$, it must also satisfy\begin{longv} the formula $\psi_1$\end{longv}:
 \[

### 2015-05-08 tex/body.tex
tuned

--- 2015-05-08 tex/body.tex
@@
 the first conjunct of formula~(\ref{eq:ex-after})---and $\absconstraints( L_0
-)$, it must also satisfy\begin{longv} the formula $\psi_1$\end{longv}
+)$, it must also satisfy\begin{longv} the formula $\psi_1$:\end{longv}
 \[

### 2015-05-08 tex/body.tex
hopefully that works
edits

--- 2015-05-08 tex/body.tex
@@
 \newcommand{\absconstraints}{\mathrm{X}}
-
-\newtheorem{cor}{Corollary}
 
@@
 recursive function to be used inside a complex term on the right-hand side,
-whereas the finite model finding techniques of Reynolds et al.\ are not
-applicable for functions over infinite domains such as the integers or
+whereas the finite model finding techniques %of Reynolds et al.\
+are not applicable for functions over infinite domains such as the integers or
 algebraic datatypes. A simple example where both approaches fail is
@@
 recursive definitions in practice and the impending addition of a dedicated
-command, \texttt{define-fun-rec}, to the SMT-LIB standard \cite{smtlib25}.
-
-In this paper, we present a method for translating formulas involving recursive function
+command for introducing them, \texttt{define-fun-rec}, to the SMT-LIB standard \cite{smtlib25}.
+
+%In this paper,
+We present a method for translating formulas involving recursive function
 definitions into formulas where finite model finding techniques can be applied.
@@
 is \emph{closed under function expansion with respect to $\DDD$} if
-\begin{shortv}$\closurefmla$.\end{shortv}%
-\begin{longv}\[\closurefmla\]\end{longv}
+\begin{shortv}$\closurefmla$. \end{shortv}%
+\begin{longv}\[\closurefmla\]\end{longv}%
 The set $\DDD$ is \emph{admissible} if for every $T$-satisfiable formula 
@@
 %
-Since $\psi$ is closed under function expansion, it already constraints the
+Since $\psi$ is closed under function expansion, it already constrains the
 functions in $\sfundefs{\Sigma}$ recursively as far as is
@@
 \emph{productive} corecursive functions. Corecursive functions are functions to
-a coalgebraic datatype. These functions can call themselves forever, without
+a coalgebraic datatype. These functions can be non-well-founded, without
 their being inconsistent. Productive corecursive functions are functions that
@@
 the definition
-$\{\forallf{\const{e}} x.\;\, \const{e}(x) \teq \con{scons}(x,\; \const{e}(x + 1))\}$
+$\forallf{\const{e}} x.\;\, \const{e}(x) \teq \con{scons}(x,\; \const{e}(x + 1))$
 falls within this fragment: Each call to $\const{e}$ produces one
@@
 \cite{blanchette-et-al-2015-fouco}. Theorem~\ref{thm:adm} can be extended to
-cover such specifications, based on the observation that infinite unfolding
-a corecursive definition computes a unique infinite object.
+cover such specifications, based on the observation that unfolding
+a corecursive definition infinitely computes a unique infinite object.
 
@@
 \teq \con{f}( x - 1 )$ or $\forallf{\con{g}} x.\allowbreak\;\, \con{g}( x
-) \teq \con{g}( x )$. More generally, we conjecture that one can ignore all
+) \teq \con{g}( x )$. We conjecture that one can ignore all
 tail-recursive calls (i.e., calls that occur on the right-hand side of the
@@
 satisfying $\con{f}$'s definition. A more subtle specimen is
-\begin{shortv}$\badassex$.\end{shortv}%
-\begin{longv}\[\badassex\]\end{longv}
+\begin{shortv}$\badassex$. \end{shortv}%
+\begin{longv}\[\badassex\]\end{longv}%
 While this set has a model where $\con f$ and $\con g$ are interpreted as the
@@
 $\conv_0( t\typ{\tau},\, p )$ $=$
-\\[-.8\baselineskip]
+%\\[-.8\baselineskip]
  \begin{itemize}
@@
       \itemx[] $\mathsf{else}$ \bigtuple{$\boolop(t'_1, \ldots, t'_n),\, \chi$}
-\\[-.8\baselineskip]
+%\\[-.8\baselineskip]
     \end{itemize}
@@
       \itemx[] $\mathsf{let}$ $( u',\, \chi ) = \conv_0( u,\, p )$ $\mathsf{in}$ \bigtuple{$\forall a : \fargtype{\con{f}}.\; u' [ \vecfarg{\con{f}}( a ) / \vec x ],\, \ltrue$}
-\\[-.8\baselineskip]
+%\\[-.8\baselineskip]
     \end{itemize}
@@
       \itemx[] $\mathsf{let}$ $( u',\, \chi ) = \conv_0( u,\, p )$ $\mathsf{in}$ \bigtuple{$\forall \vec x.\; u',\, \forall \vec x.\; \chi$}
-\\[-.8\baselineskip]
+%\\[-.8\baselineskip]
     \end{itemize}
@@
      \itemx[] \return \bigtuple{$t,\; \bigwedge\, \{ \exists a : \fargtype{\con{f}}.\; \vecfarg{\con{f}}( a ) \teq \vec t \mid \con{f}( \vec t ) \in \terms( t ),\, \con{f} \in \sfundefs{\Sigma} \}$}
-\\[-.8\baselineskip]
+%\\[-.8\baselineskip]
    \end{itemize}
@@
 $\pnone$. \vthinspace$\conv_0$ returns a pair $( t'\negvthinspace,\, \chi )$, where $t'$ is a term of
-the same type as $t$, and $\chi$ is a $\extendsig{\Sigma}$-formula.
+the same type as $t$ and $\chi$ is a $\extendsig{\Sigma}$-formula.
 %an additional constraint.
@@
 If $t$ is an application of a predicate $\boolop$, including the built-in
-operators ${\neg}$, ${\wedge}$, ${\vee}$, ${\teq}$, and ${\lite}$,
+operators ${\oldneg}$, ${\wedge}$, ${\vee}$, ${\teq}$, and ${\lite}$,
 $\conv_0$ calls itself recursively on the arguments $t_i$ and polarity $\pol(
@@
 the subterms as a conjunction. Dually, if $t$ is negative, it constructs a
-disjunction with the negated constraint. It $t$ has no polarity, the
-constraint $\chi$ is simply returned to the caller.
+disjunction with the negated constraint, to achieve the same overall effect as
+in the positive case. It $t$ has no polarity, the constraint $\chi$ is
+returned to the caller.
 %
@@
 %
-If $t$ is a function definition of a function in
-$\sfundefs{\Sigma}$, $\conv_0$ constructs a quantified formula over a single variable
+If $t$ is a function definition, % of a function in $\sfundefs{\Sigma}$,
+$\conv_0$ constructs a quantified formula over a single variable
 $a$ of type $\fargtype{\con{f}}$ and replaces all occurrences of $\vec x$ in
@@
 \rightarrow \typeint$. The result of $\conv( \varphi )$ 
-is $\varphi'$ which (after simplification) is the $\extendsig{\Sigma}$-formula
+is $\varphi'$, which (after simplification) is the $\extendsig{\Sigma}$-formula
 %
@@
 implemented in \cvc, to search for a finite interpretation for $\fargtype{\con{s}}$. 
-Since all occurrences of the quantified variable $a$ in this formula are 
+Since all occurrences of the quantified variable~$a$ in this formula are 
 beneath applications of the uninterpreted function $\farg{\con{s}}$, 
@@
 a complete instantiation procedure, implemented in \ziii. 
-Both
+Accordingly,
 %%% Added "apparently". Or are we sure?  
@@
 \cvc and \ziii run indefinitely on formula~(\ref{eq:ex-before}), 
-whereas they both produce a model for formula~(\ref{eq:ex-after}) 
+whereas they produce a model for formula~(\ref{eq:ex-after}) 
 within 100 milliseconds.\xend
@@
 $\con{c}$ as $14$,
-and $\con{s}$ as the (almost constant) function
+and $\con{s}$ as the almost constant function
 %
 %\begin{equation} \label{eq:approx-interp}
-\begin{shortv}$\badlambda$.\end{shortv}%
-\begin{longv}\[\badlambda\]\end{longv}
+\begin{shortv}$\badlambda$. \end{shortv}%
+\begin{longv}\[\badlambda\]\end{longv}%
 %\end{equation}
@@
 \begin{longv}
-In the following, we sketch the correctness of our translation $\conv$.
-For a set of ground literals $L$, 
+%In the following,
+We sketch the correctness of translation~$\conv$.
+For a set of ground literals~$L$, 
 we write $\absconstraints( L )$ to denote the set of constraints that
 force the concretization functions to have the necessary elements in their
-range for determining the satisfiability of $L$ with respect to function
+range for determining the satisfiability of $L$ with respect to the function
 definitions in the translation.
@@
 
-\begin{cor}\label{cor:conv}
+\begin{corollary}\label{cor:conv}
 If $\psi$ is a formula not containing function definitions, then $\conv( \psi )$ entails $\psi$.
-\end{cor}
+\end{corollary}
 \end{longv}
@@
 \begin{proofsketch}
-First, we show that if $\varphi$ is satisfied by $\Sigma$-interpretation $\I$, then $\conv( \varphi )$ is satisfied by an $\extendsig{\Sigma}$-interpretation $\J$.
-Let $\J$ be the $\extendsig{\Sigma}$-interpretation that interprets all types $\tau \in \stypes{\Sigma}$ as $\tau^\I$,
+First, we show that if $\varphi$ is satisfied by an
+$\Sigma$-interpretation $\I$, then $\conv( \varphi )$ is satisfied by an $\extendsig{\Sigma}$-interpretation $\J$.
+Let $\J$ be the $\extendsig{\Sigma}$-interpretation that interprets all types $\tau \in \stypes{\Sigma}$ as~$\tau^\I$,
 all functions $\con{f} \in \sfuns{\Sigma}$ as $\con{f}^\I$,
 and for each function $\con{f} : \tau_1 \times \cdots \times \tau_n \rightarrow \tau$ in $\sfundefs{\Sigma}$,
-interprets $\fargtype{\con{f}}$ as a tuple of values belonging to $\tau_1^\I \times \cdots \times \tau_n^\I$,
+interprets $\fargtype{\con{f}}$ as $\tau_1^\I \times \cdots \times \tau_n^\I$
 and each $\fargx{\con{f}}{i}$ as the $i$\vvthinspace{th} projection on such tuples for $i = 1, \ldots, n$.
@@
 %then $\I$ satisfies $t$ if and only if $\J$ satisfies $t'$.
-Thus, $\J$ satisfies $L \cup \absconstraints( L )$, and by Lemma~\ref{lem:conv} we conclude $\J$ satisfies $\conv( \varphi )$.
+Thus, $\J$ satisfies $L \cup \absconstraints( L )$, and by Lemma~\ref{lem:conv}
+we conclude $\J$ satisfies $\conv( \varphi )$.
 
@@
 let $\psi_i$ be the formula $\bigwedge\, \{ \conv( \varphi_{\con{f}}[ \vec t / \vec x ] ) \mid \con{f}( \vec t ) \in \terms( L_i ),\: \con{f} \in \sfundefs{\Sigma} \}$,
-where the function definition of $\con{f}$ in $\DDD$ is $\forallf{\con{f}} \vec x.\; \varphi_{\con{f}}$.
+where $\forallf{\con{f}} \vec x.\; \varphi_{\con{f}} \in \DDD$.
 Since $\J$ satisfies $\conv( \forallf{\con{f}} \vec x.\; \varphi_{\con{f}} )$ and $\absconstraints( L_i )$,
@@
 Second, if a model is found for $\conv( \varphi )$,
-then the constraints of the form $\exists a : \fargtype{\con{f}}.\;\, (\vecfarg{\con{f}}( a ) \teq \vec t)$
+then the constraints $\exists a : \fargtype{\con{f}}.\;\, \vecfarg{\con{f}}( a ) \teq \vec t$
 occurring in $\conv( \varphi )$ ensure that this model also satisfies
-a $\Sigma$-formula that is closed under function expansion that entails the conjecture of $\varphi$.
-Hence when $\DDD$ is admissible, this implies the existence of a model for $\varphi$.
+a $\Sigma$-formula that is closed under function expansion and
+that entails the conjecture of $\varphi$.
+This implies the existence of a model for $\varphi$,
+assuming $\DDD$ is admissible.
 %For clarity, we demonstrate the consequences of this theorem by means of an example.
@@
 This formula is unsatisfiable only if
-formula~(\ref{eq:ex-before}) is unsatisfiable since $\fargtype{\con{s}}$ can be
+formula~(\ref{eq:ex-before}) is unsatisfiable, since $\fargtype{\con{s}}$ can be
 interpreted as the integers and $\farg{\con{s}}$ as the identity
@@
 $L_0 \cup \absconstraints( L_0 )$,
-where $L_0 = \{ \con{s}( \con{c} ) > 100 \}$, 
-where $\absconstraints( L_0 )$ is the set consisting of the single constraint $\exists a : {\fargtype{\con{s}}}. (\farg{\con{s}}( a ) \teq \con{c})$.
+where $L_0 = \{ \con{s}( \con{c} ) > 100 \}$
+and $\absconstraints( L_0 )$ is the set consisting of the single constraint
+$\exists a : {\fargtype{\con{s}}}.\;\, \farg{\con{s}}( a ) \teq \con{c}$.
 %The latter ensures that $\I$ satisfies a formula that entails both $\con{s}( \con{c} ) > 100$ and $\varphi_{\con{s}}[\con{c}/x]$.
-Since $\I$ also satisfies both the translated function definition for $\con{s}$---i.e.,
-the first conjunct of formula~(\ref{eq:ex-after})---and $\absconstraints( L_0
-)$, it must also satisfy\begin{longv} the formula $\psi_1$\end{longv}
+Since $\I$ also satisfies both the translated function definition for $\con{s}$
+(the first conjunct of formula~(\ref{eq:ex-after})) and $\absconstraints( L_0
+)$, it must also satisfy\begin{longv} the formula $\psi_1$:\end{longv}
 \[
@@
 \[\con{s}( \con{c} ) > 100 \wedge %\displaystyle
-  \bigwedge\nolimits_{\vthinspace i=0}^{\!n-1} \bigl( \neg (\con{c}-i \leq 0) \wedge \con{s}( \con{c}-i ) \teq \con{c}-i + \con{s}( \con{c}-i-1 ) \bigr)
+  \smash{\bigwedge\nolimits_{\vthinspace i=0}^{\!n-1}} \bigl( \neg (\con{c}-i \leq 0) \wedge \con{s}( \con{c}-i ) \teq \con{c}-i + \con{s}( \con{c}-i-1 ) \bigr)
 \wedge \con{c}-n \leq 0 \wedge \con{s}( \con{c}-n ) \teq 0\]
@@
 Since $\{\forallf{\con{s}} x.\;\, \varphi_\con{s}\}$ is admissible,
-there exists a $\Sigma$-interpretation satisfying $\psi \wedge \forallf{\con{s}} x.\;\, \varphi_\con{s}$
+there exists a $\Sigma$-interpretation satisfying $\psi \wedge \forallf{\con{s}} x.\;\, \varphi_\con{s}$,
 which entails formula~(\ref{eq:ex-before}).\xend
@@
 
+{\looseness=-1
 In this section, we evaluate both the overall impact of the translation
@@
 %
-We gathered a set of 245 benchmarks from two sources, which we will refer to as
+We gathered 245 benchmarks from two sources, which we will refer to as
 \isa and \leon. The first source consists of the 79
@@
 suite~\cite{DBLP:conf/itp/JohanssonDB10} that do not contain higher-order
-functions. The benchmarks have been used recently as challenge problems for a
-variety of inductive theorem provers. These benchmarks heavily involve
+functions. These benchmarks have been used recently as challenge problems for a
+variety of inductive theorem provers. They heavily involve
 recursive functions and are limited to the combined theory of uninterpreted
@@
 \url{https://github.com/epfl-lara/leon/}} which were constructed from
-verification conditions from simple Scala programs. These benchmarks also
+verification conditions about simple Scala programs. These benchmarks also
 heavily involve recursively defined functions over algebraic datatypes, 
 but cover a wide variety of theories, including bit vectors, arrays, and
-both linear and nonlinear arithmetic. Each benchmark is in definitional form
+both linear and nonlinear arithmetic. All benchmarks are in definitional form
 with respect to a set of well-founded functions.
@@
 %conjecture.
-Nearly all of these benchmarks are unsatisfiable, except for a
-handful of benchmarks in the \leon set.
+Nearly all of them are unsatisfiable, except for a
+handful in the \leon set.
+
+}
 
@@
 one position with another of the same type at another position.
-Note that benchmarks created in this way have a high likelihood of having small
-and easy-to-find countermodels. In total, we considered 213 mutated forms of
+Note that benchmarks created in this way have a high likelihood of having
+small, easy-to-find countermodels. In total, we considered 213 mutated forms of
 conjectures from \isa and 427 mutated forms of conjectures from \leon. We will
@@
 The translation $\conv$ significantly improves the effectiveness of known SMT
-techniques for finding counterexamples for formulas involving recursive
+techniques for finding counterexamples to formulas involving recursive
 function definitions.
@@
 When reading this input,
-\begin{longv}
+\begin{longv}%
 \cvc adds the annotated quantified formula
@@
 $$\forallf{\con{s}} x.\;\, \lite\bigl( x \leq 0,\; \con{s}( x ) \teq 0,\; \con{s}( x ) \teq \con{s}( x-1 ) \bigr)$$
-\end{longv}
-\begin{shortv}
+\end{longv}%
+\begin{shortv}%
 \cvc adds
@@
 which after rewriting becomes
-$\forallf{\con{s}} x.\;\, \lite\bigl( x \leq 0,\; \con{s}( x ) \teq 0,\; \con{s}( x ) \teq \con{s}( x-1 ) \bigr)$.\end{shortv}
+$\forallf{\con{s}} x.\;\, \lite\bigl( x \leq 0,\; \con{s}( x ) \teq 0,\; \con{s}( x ) \teq \con{s}( x-1 ) \bigr)$. \end{shortv}%
 If \cvc's finite model finding mode for recursive functions is enabled (using
@@
 
-The \texttt{--fmf-fun} option instructs \cvc to assume that the
-functions provided by the user are admissible.
-Admissibility must be checked \relax{by the user}, perhaps using a
+The \texttt{--fmf-fun} option tells \cvc to assume that
+functions introduced using \texttt{define-\allowbreak fun-\allowbreak rec} are admissible.
+Admissibility must be checked \begin{longv}separately \end{longv}\relax{by the user}---e.g., using a
 syntactic criterion or a termination prover.
@@
 \begin{verbatim}
-(define-fun-rec h ((x Int)) Int (+ (h x) 1))
+(define-fun-rec h ((x Int)) Int (+ (h x) x))
 \end{verbatim}
@@
 approximated type with an unknown value, which is propagated by function
-application. This approach works reasonably well in Nitpick, because it builds
+application. This scheme works reasonably well in Nitpick, because it builds
 on a relational logic, but our initial experiments with \cvc suggest
-that it is more efficient to avoid unknown values by adding existential
+that it is more efficient to avoid unknowns by adding existential
 constraints.
@@
 The Leon system~\cite{blanc2013overview} implements a procedure that can both
-produces proofs and counterexamples for properties of terminating functions
+produces proofs and counter\-examples for properties of terminating functions
 written in a subset of Scala. Leon is based on an SMT solver. It avoids
@@
 depth, which is increased on a per-need basis.
-The approach described in this paper works in an analogous manner, 
+Our translation~$\conv$ works in an analogous manner, 
 where instead the SMT solver is invoked only once 

### 2015-05-08 tex/body.tex
fixes

@@
 This is consistent with our expectations and Theorem~\ref{thm:equi}, 
-since these benchmarks contained only sets of function definitions
-that are supposed to be well-founded.
+since these benchmarks contain only well-founded function definitions.
 %thereby confirming Theorem~\ref{thm:equi}.
@@
 
-The Leon system~\cite{blanc2013overview} implements a procedure that can both
-produces proofs and counter\-examples for properties of terminating functions
+The Leon system~\cite{blanc2013overview} implements a procedure that can
+produce both proofs and counter\-examples for properties of terminating functions
 written in a subset of Scala. Leon is based on an SMT solver. It avoids

### 2015-05-09 tex/body.tex
simpler

@@
 
-\newcommand\badassex{%
- \{\forallf{\con{f}} x.\;\, \con{f}( x ) \teq \con{f}( x ),\;\,
- \forallf{\con{g}} x.\;\, \lite( \con{f}( x ) \teq 5,\allowbreak\;
-   \con{g}( x ) \teq \con{f}( x ),\allowbreak\;
-   \con{g}( x ) \teq \con{g}( x ) + 1)\}}
-
-An example of an inadmissible function set is
-$\{ \forallf{\con{f}} x.\;\, \con{f}( x ) \teq \con{f}( x ) + 1 \}$,
+%\newcommand\badassex{
+% \{\forallf{\con{f}} x : \typeint.\;\, \con{f}( x ) \teq \con{f}( x ),\;\,
+% \forallf{\con{g}} x : \typeint.\;\, \lite( \con{f}( x ) \teq 5,\allowbreak\;
+%   \con{g}( x ) \teq \con{f}( x ),\allowbreak\;
+%   \con{g}( x ) \teq \con{g}( x ) + 1)\}}
+\newcommand\badassex{
+ \{\forallf{\con{f}} x : \typeint.\;\, \con{f}( x ) \teq \con{f}( x ),\;\,
+ \forallf{\con{g}} x : \typeint.\;\, \con{g}( x ) \teq \con{g}( x ) + \nobreak\con{f}( x )\}}
+
+An example of an inadmissible set is
+$\{ \forallf{\con{f}} x : \typeint.\;\, \con{f}( x ) \teq \con{f}( x ) + 1 \}$,
 since the formula $\ltrue$ is (trivially) closed under function expansion with respect to this set,
@@
 While this set has a model where $\con f$ and $\con g$ are interpreted as the
-constant function $5$, it is not admissible since %the formula
-$\con{f}( 0 ) \teq 4$ is
+constant function $0$, it is not admissible since %the formula
+$\con{f}( 0 ) \teq 1$ is
 closed under function expansion \begin{longv}with respect to this set\end{longv}
 and yet there exists
-no interpretation satisfying both $\con{f}( 0 ) \teq 4$ and $\con{g}$'s
+no interpretation satisfying both $\con{f}( 0 ) \teq 1$ and $\con{g}$'s
 definition.

### 2015-05-09 tex/body.tex
comma
tail rec def

@@
 Since $\J \models \Delta$ by assumption, we have $\J \models \psi \wedge
-\bigwedge \Delta$, as desired.
+\bigwedge \Delta$ as desired.
 %We have
@@
 ) \teq \con{g}( x )$. We conjecture that one can ignore all
-tail-recursive calls (i.e., calls that occur on the right-hand side of the
-definition, possibly in some $\lite$ branch) when establishing well-foundedness
+tail-recursive calls (i.e., calls that occupy the right-hand side of the
+definition, potentially under some $\lite$ branch) when establishing well-foundedness
 or productivity, without affecting admissibility.

### 2015-05-13 tex/body.tex
more
added sentence to long version
tiny version

@@
 
+%% The institutions above shouldn't count as footnotes
+\setcounter{footnote}{0}
+
+\begin{nottinyv}
 \begin{abstract}
@@
 
-%% The institutions above shouldn't count as footnotes
-\setcounter{footnote}{0}
-
 \section{Introduction}
 \label{sec:introduction}
+\end{nottinyv}
 
@@
 definitions into formulas where finite model finding techniques can be applied.
-The recursive functions must meet a semantic criterion to be admissible
-(Section~\ref{sec:prelim}). This criterion is met by well-founded (terminating)
+The recursive functions must meet a semantic criterion to be admissible\begin{nottinyv}
+(Section~\ref{sec:prelim})\end{nottinyv}. This criterion is met by well-founded (terminating)
 recursive function definitions. \begin{longv}It is not met by inconsistent
@@
 We define a translation for a class of
-formulas involving admissible recursive function definitions
-(Section~\ref{sec:encoding}). The main insight is that a
+formulas involving admissible recursive function definitions\begin{nottinyv}
+(Section~\ref{sec:encoding})\end{nottinyv}. The main insight is that a
 recursive definition $\forall x : \tau.\;\, \const{f}(x) \teq t$
@@
 evidence that this translation improves the effectiveness of the SMT solvers
-\cvc and \ziii for finding counterexamples to verification conditions
-(Section~\ref{sec:evaluation}). The approach is implemented as a preprocessor
-in \cvc (Section~\ref{sec:front-end}).\begin{longv} When invoked with a specific
+\cvc and \ziii for finding counterexamples to verification conditions\begin{nottinyv}
+(Section~\ref{sec:evaluation})\end{nottinyv}. The approach is implemented as a preprocessor
+in \cvc \begin{nottinyv}(Section~\ref{sec:front-end})\end{nottinyv}. \begin{longv}%
+When invoked with a specific
 command-line option, the solver assumes that functions introduced using the
@@
 
+\begin{nottinyv}
 \section{Preliminaries}
@@
 all occurrences of $\vec x$ with $\vec t$ in $u$.
+\begin{longv}When applied to terms, the symbol $=$ denotes syntactic equality.\end{longv}
 
@@
 and algebraic datatypes.
+\end{nottinyv}
 
@@
 
+\begin{nottinyv}
 {%\footnotesize
@@
 }
+\end{nottinyv}
 

### 2015-05-29 tex/body.tex
tiny
tuning

@@
 
+\hyphenation{counter-example counter-examples Isa-belle}
+
 \begin{document}
@@
 This research is partially supported by the Inria technological development
-action ``Contre-exemples Utilisables par Isabelle et Coq'' (CUIC).
+action ``Contre-exemples utilisables par Isabelle et Coq'' (CUIC).
 }
@@
 
-\begin{nottinyv}
+%\begin{nottinyv}
 \begin{abstract}
@@
 \label{sec:introduction}
-\end{nottinyv}
+%\end{nottinyv}
 
@@
 $\forall a :
-\fargtype{\tau}.\;\, \const{f}(\farg{\con{f}}(a)) \teq t[\concret(a)/x]$, where
+\fargtype{\tau}.\;\, \const{f}(\farg{\con{f}}(a)) \teq t[\farg{\con{f}}(a)/x]$, where
 $\fargtype{\tau}$ is an uninterpreted \relax{abstract} type and $\farg{\con{f}} \begin{longv}:
@@
 
-\begin{nottinyv}
+%\begin{nottinyv}
 \section{Preliminaries}
@@
 \end{tabular}
-\caption{\,Number of \emph{sat} responses on benchmarks without and with translation with $\conv$}
+\caption{\,Number of \emph{sat} responses on benchmarks without and with $\conv$ translation}
 \label{fig:sat}
@@
 \end{tabular}
-\caption{\,Number of \emph{unsat} responses on benchmarks without and with translation with $\conv$}
+\caption{\,Number of \emph{unsat} responses on benchmarks without and with $\conv$ translation}
 \label{fig:unsat}
@@
 and algebraic datatypes.
-\end{nottinyv}
+%\end{nottinyv}
 
@@
 
-\begin{nottinyv}
+\begin{tinyv}
+This abstract is based on a regular submission to the SMT 2015 workshop,
+which itself is based on a longer technical report. Both are available online.%
+\footnote{\url{http://lara.epfl.ch/~reynolds/SMT2015-recfun/}}
+\end{tinyv}
+
+%\begin{nottinyv}
 {%\footnotesize
@@
 }
-\end{nottinyv}
+%\end{nottinyv}
 

### 2015-05-29 tex/body.tex
more

@@
 This paper introduces a translation which reduces axioms specifying a large
-class of recursive functions, including well-founded (terminating),
+class of recursive functions, including well-founded (terminating) functions,
 to universally quantified formulas for which these
@@
 \fargtype{\tau} \to \tau\end{longv}$ converts the abstract type to the concrete
-type. The translation preserves satisfiability and\begin{longv}, for admissible definitions,\end{longv}
+type. Additional constraints ensure that the abstract values that are relevant
+to the formula's satisfiability
+exist. The translation preserves satisfiability and\begin{longv}, for admissible definitions,\end{longv}
 unsatisfiability, and makes finite model finding possible for problems in this

### 2015-06-01 tex/camera_ready/paper.tex
added camera-ready and make fit on 11 pages

--- 2015-05-29 tex/body.tex
@@
+\PassOptionsToClass{11pt,withtimes}{easychair}
+\PassOptionsToPackage{hyphens}{url}
+\PassOptionsToPackage{scaled=.86}{helvet}
+\PassOptionsToPackage{
+   a4paper,
+   pdftex,
+   pdftitle={Model Finding for Recursive Functions in SMT},
+   pdfauthor={Andrew Reynolds, Jasmin Christian Blanchette, and Cesare Tinelli},
+   pdfkeywords={},
+   pdfborder={0 0 0},
+   draft=false,
+   bookmarksnumbered,
+   bookmarks,
+   bookmarksdepth=2,
+   bookmarksopenlevel=2,
+   bookmarksopen}{hyperref}
+\documentclass{easychair}
+
+%envcountsect,
+\usepackage{version}
+
+\excludeversion{tinyv}
+\includeversion{nottinyv}
+\includeversion{shortv}
+\excludeversion{longv}
+
 %\documentclass[runningheads,a4paper]{llncs}
@@
 
-\newcommand\closurefmla{\psi \models_T \bigwedge\nolimits_{\,i=1}^{n} \{ \psi_i[ \vec t / \vec x ] \mid \con{f}_i( \vec t ) \in \terms( \psi ) \}}
+\newcommand\closurefmla{\psi \models_T \bigwedge\nolimits_{\,i=1}^{n} \{ \varphi_i[ \vec t / \vec x ] \mid \con{f}_i( \vec t ) \in \terms( \psi ) \}}
 
@@
 quantifiers altogether by unfolding recursive definitions up to a certain
-depth, which is increased on a per-need basis.
+depth. %, which is increased on a per-need basis.
 Our translation~$\conv$ works in an analogous manner, 
@@
 to thank Damien Busato-Gaston and Emmanouil Koukoutos for providing the initial
-set of benchmarks used in the evaluation.
+set of benchmarks used in the evaluation, and the anonymous reviewers for their
+suggestions and comments.
 }
@@
 \bibliographystyle{abbrv}
-\bibliography{bib}
+%\bibliography{bib}
+
+\begin{thebibliography}{10}
+
+\bibitem{barrett-et-al-2011}
+C.~Barrett, C.~L. Conway, M.~Deters, L.~Hadarean, D.~Jovanovic, T.~King,
+  A.~Reynolds, and C.~Tinelli.
+\newblock {CVC4}.
+\newblock In G.~Gopalakrishnan and S.~Qadeer, editors, {\em CAV 2011}, volume
+  6806 of {\em LNCS}, pages 171--177. Springer, 2011.
+
+\bibitem{smtlib25}
+C.~Barrett, P.~Fontaine, and C.~Tinelli.
+\newblock The {SMT-LIB} standard---{V}ersion 2.5.
+\newblock \url{http://smt-lib.org/language.shtml}, to appear.
+
+\bibitem{baumgartner-et-al-2009}
+P.~Baumgartner, A.~Fuchs, H.~de~Nivelle, and C.~Tinelli.
+\newblock Computing finite models by reduction to function-free clause logic.
+\newblock {\em J. Applied Logic}, 7(1):58--74, 2009.
+
+\bibitem{blanc2013overview}
+R.~Blanc, V.~Kuncak, E.~Kneuss, and P.~Suter.
+\newblock An overview of the {Leon} verification system---{V}erification by
+  translation to recursive functions.
+\newblock In {\em Scala {'}13}. ACM, 2013.
+
+\bibitem{blanchette-nipkow-2010}
+J.~C. Blanchette and T.~Nipkow.
+\newblock Nitpick: A counterexample generator for higher-order logic based on a
+  relational model finder.
+\newblock In M.~Kaufmann and L.~C. Paulson, editors, {\em ITP 2010}, volume
+  6172 of {\em LNCS}, pages 131--146. Springer, 2010.
+
+\bibitem{claessen-hughes-2000}
+K.~Claessen and J.~Hughes.
+\newblock Quick{C}heck: A lightweight tool for random testing of {H}askell
+  programs.
+\newblock In {\em ICFP {'}00}, pages 268--279. ACM, 2000.
+
+\bibitem{claessen-sorensson-2003}
+K.~Claessen and N.~S{\"o}rensson.
+\newblock New techniques that improve {MACE}-style model finding.
+\newblock In {\em MODEL}, 2003.
+
+\bibitem{MouraBjoerner07}
+L.~de~Moura and N.~Bj{\o}rner.
+\newblock Efficient {E}-matching for {SMT} solvers.
+\newblock In F.~Pfenning, editor, {\em CADE-21}, volume 4603 of {\em LNCS},
+  pages 183--198. Springer, 2007.
+
+\bibitem{de-moura-bjoerner-2008}
+L.~de~Moura and N.~Bj{\o}rner.
+\newblock {Z3}: {An} efficient {SMT} solver.
+\newblock In C.~R. Ramakrishnan and J.~Rehof, editors, {\em TACAS 2008}, volume
+  4963 of {\em LNCS}, pages 337--340. Springer, 2008.
+
+\bibitem{dunets-et-al-2010}
+A.~Dunets, G.~Schellhorn, and W.~Reif.
+\newblock Automated flaw detection in algebraic specifications.
+\newblock {\em J.~Autom.\ Reasoning}, 45(4):359--395, 2010.
+
+\bibitem{GeDeM-CAV-09}
+Y.~Ge and L.~de~Moura.
+\newblock Complete instantiation for quantified formulas in satisfiability
+  modulo theories.
+\newblock In {\em CAV '09}, volume 5643 of {\em LNCS}, pages 306--320.
+  Springer, 2009.
+
+\bibitem{DBLP:conf/itp/JohanssonDB10}
+M.~Johansson, L.~Dixon, and A.~Bundy.
+\newblock Case-analysis for rippling and inductive proof.
+\newblock In {\em {ITP} 2010}, pages 291--306, 2010.
+
+\bibitem{korovin-2013}
+K.~Korovin.
+\newblock Non-cyclic sorts for first-order satisfiability.
+\newblock In P.~Fontaine, C.~Ringeissen, and R.~A. Schmidt, editors, {\em
+  FroCoS 2013}, volume 8152 of {\em LNCS}, pages 214--228. Springer, 2013.
+
+\bibitem{krauss-2009-phd}
+A.~Krauss.
+\newblock {\em Automating Recursive Definitions and Termination Proofs in
+  Higher-Order Logic}.
+\newblock {Ph.D.}\ thesis, Technische Universit\"at M\"unchen, 2009.
+
+\bibitem{kuncak-jackson-2005}
+V.~Kuncak and D.~Jackson.
+\newblock Relational analysis of algebraic datatypes.
+\newblock In M.~Wermelinger and H.~Gall, editors, {\em ESEC/FSE 2005}. ACM,
+  2005.
+
+\bibitem{mccune-1994}
+W.~McCune.
+\newblock A {D}avis--{P}utnam program and its application to finite first-order
+  model search: Quasigroup existence problems.
+\newblock Technical report, Argonne National Laboratory, 1994.
+
+\bibitem{reynolds-blanchette-2015-codata}
+A.~Reynolds and J.~C. Blanchette.
+\newblock A decision procedure for (co)datatypes in {SMT} solvers.
+\newblock In A.~Felty and A.~Middeldorp, editors, {\em CADE-25}, LNCS.
+  Springer, 2015.
+
+\bibitem{our-report}
+A.~Reynolds, J.~C. Blanchette, and C.~Tinelli.
+\newblock Model finding for recursive functions in {SMT}.
+\newblock Tech.\ report, \url{http://lara.epfl.ch/~reynolds/SMT2015-recfun/},
+  2015.
+
+\bibitem{reynolds-kuncak-2015}
+A.~Reynolds and V.~Kuncak.
+\newblock Induction for {SMT} solvers.
+\newblock In D.~D'Souza, A.~Lal, and K.~G. Larsen, editors, {\em {VMCAI} 2015},
+  volume 8931 of {\em LNCS}, pages 80--98. Springer, 2014.
+
+\bibitem{ReynoldsTinelliMoura14}
+A.~Reynolds, C.~Tinelli, and L.~de~Moura.
+\newblock Finding conflicting instances of quantified formulas in {SMT}.
+\newblock In {\em FMCAD 2014}, pages 195--202. {IEEE}, 2014.
+
+\bibitem{ReyEtAl-1-RR-13}
+A.~Reynolds, C.~Tinelli, A.~Goel, and S.~Krsti\'c.
+\newblock Finite model finding in {SMT}.
+\newblock In N.~Sharygina and H.~Veith, editors, {\em CAV 2013}, volume 8044 of
+  {\em LNCS}, pages 640--655. Springer, 2013.
+
+\bibitem{reynolds-et-al-2013}
+A.~Reynolds, C.~Tinelli, A.~Goel, S.~Krsti{\'c}, M.~Deters, and C.~Barrett.
+\newblock Quantifier instantiation techniques for finite model finding in
+  {SMT}.
+\newblock In M.~P. Bonacina, editor, {\em CADE-24}, volume 7898 of {\em LNCS},
+  pages 377--391. Springer, 2013.
+
+\bibitem{slaney-1994}
+J.~K. Slaney.
+\newblock {FINDER}: Finite domain enumerator system description.
+\newblock In A.~Bundy, editor, {\em CADE-12}, volume 814 of {\em LNCS}, pages
+  798--801. Springer, 1994.
+
+\bibitem{torlak-jackson-2007}
+E.~Torlak and D.~Jackson.
+\newblock Kodkod: A relational model finder.
+\newblock In O.~Grumberg and M.~Huth, editors, {\em TACAS 2007}, volume 4424 of
+  {\em {LNCS}}, pages 632--647. Springer, 2007.
+
+\bibitem{weber-2008}
+T.~Weber.
+\newblock {\em SAT-Based Finite Model Generation for Higher-Order Logic}.
+\newblock {Ph.D.}\ thesis, Technische Universit\"at M\"unchen, 2008.
+
+\bibitem{zhang-zhang-1995}
+J.~Zhang and H.~Zhang.
+\newblock {SEM}: A system for enumerating models.
+\newblock In C.~S. Mellish, editor, {\em IJCAI-95}, volume~1, pages 298--303.
+  Morgan Kaufmann, 1995.
+
+\end{thebibliography}
 }

### 2015-06-04 tex/camera_ready/paper.tex

@@
 
+\newcommand{\svars}[1]{#1^\mathrm{v}}
+
 \newcommand{\sfuns}[1]{#1^\mathrm{f}}
@@
 \and
-Department of Computer Science, University of Iowa, USA
+Department of Computer Science, The University of Iowa, USA
 }
@@
 for each $\tau \in \stypes{\Sigma}$.
+For each $\tau \in \stypes{\Sigma}$,
+we fix an infinite set $\svars{\Sigma}_\tau$ of \emph{variables of type $\tau$} and
+define $\svars{\Sigma}$ as $\bigcup_{\tau \in \stypes{\Sigma}} \svars{\Sigma}_\tau$.
+$\Sigma$-terms are built as usual over functions symbols in $\Sigma$ and variables in $\svars{\Sigma}$. 
 Formulas are terms of type $\Bool$.
@@
 the \emph{domain} of~$\tau$ in~$\I$,
-and each function symbol $\con{f} : \tau_1 \times \cdots \times \tau_n \rightarrow \tau$ in
-$\sfuns{\Sigma}$
-to a total function $\con{f}^\I : \tau_1^\I \times \cdots \times \tau_n^\I \rightarrow \tau^\I$.
+ each function symbol $\con{f} : \tau_1 \times \cdots \times \tau_n \rightarrow \tau$ in $\sfuns{\Sigma}$
+to a total function $\con{f}^\I : \tau_1^\I \times \cdots \times \tau_n^\I \rightarrow \tau^\I$
+and each variable $x:\tau$ of $\svars{\Sigma}$ as an element of $\tau^\I$.
 A \emph{theory} is a pair $T = (\Sigma, \Mo)$ where
 $\Sigma$ is a signature and $\Mo$ is a class of $\Sigma$-interpretations,
-the \emph{models} of $T$.
-A $\Sigma$-formula $\varphi$ is
-\emph{$T$-satisfiable}
+the \emph{models} of $T$, closed under variable reassignment
+(that is, for every $I \in \Mo$, every $\Sigma$-interpretation that differs
+from $I$ only on the variables of $\svars \Sigma$ is also in $\Mo$).
+A $\Sigma$-formula $\varphi$ is \emph{$T$-satisfiable}
 if it is satisfied by some interpretation in $\Mo$\begin{longv};
@@
 if each $T$-entails the other.
-Given a signature $\Sigma$,
-the \emph{theory of equality with uninterpreted functions} \euf
-consists of the set of all $\Sigma$-interpretations.
-Informally, we refer to the type and function symbols in this theory as \emph{uninterpreted}.
-%\ct{This is both ambiguous and insufficient.
-%Since this is SMT we should fix some background theory $T$,
-%then extend its signature with additional symbols and call \emph{those} uninterpreted.
-%}
-
-In addition to $\forall \vec x.\; \varphi$,
-we consider \emph{annotated quantified formulas} of the form
+%ct changed the following because it is not adequate for our purposes
+%
+%Given a signature $\Sigma$,
+%the \emph{theory of equality with uninterpreted functions} \euf
+%consists of the set of all $\Sigma$-interpretations.
+%Informally, we refer to the type and function symbols in this theory as \emph{uninterpreted}.
+%%\ct{This is both ambiguous and insufficient.
+%%Since this is SMT we should fix some background theory $T$,
+%%then extend its signature with additional symbols and call \emph{those} uninterpreted.
+%%}
+If $T_1 = (\Sigma_1, \Mo_1)$ is a theory and $\Sigma_2$ is a signature 
+with $\sfuns{\Sigma_1} \cap \sfuns{\Sigma_2} = \emptyset$,
+the \emph{extension of $T_1$ to $\Sigma_2$} is the theory $T = (\Sigma, \Mo)$ where 
+$\sfuns{\Sigma} = \sfuns{\Sigma_1} \cup \sfuns{\Sigma_2}$,
+$\stypes{\Sigma} = \stypes{\Sigma_1} \cup \stypes{\Sigma_2}$,
+and $\Mo$ is the set of all $\Sigma$-interpretations $\I$
+whose $\Sigma_1$-reduct is a model of $T_1$.
+We refer to the symbols of $\Sigma_2$ that are not is $\Sigma_1$ as \emph{uninterpreted}.
+For the rest of the paper \emph{we fix a theory $T = (\Sigma, \Mo)$ 
+with uninterpreted symbols} constructed as above.
+
+We will consider \emph{annotated quantified formulas} of the form
 $\forallf{\const{f}} \vec x.\; \varphi$, where $\con{f} \in \sfuns{\Sigma}$ is
 uninterpreted. Their
-semantics is the same as for standard quantified formulas.
+semantics is the same as for standard quantified formulas
+$\forall \vec x.\; \varphi$.
 Given $\con{f} : \tau_1 \times \cdots \times \tau_n \rightarrow \tau$,
@@
 
-Given a signature $\Sigma$, we distinguish the subset of \emph{defined}
-function symbols $\sfundefs{\Sigma} \subseteq \sfuns{\Sigma}$.
+In the signature $\Sigma$, we distinguish a subset $\sfundefs{\Sigma} \subseteq \sfuns{\Sigma}$ of \emph{defined} uninterpreted function symbols.
 We consider $\Sigma$-formulas that are in definitional form with respect to
@@
 x \leq 0,\allowbreak\; \con{p}( x ) \teq 1,\allowbreak\; \con{p}( x ) \teq 2 * \con{p}( x - 1 ) \bigr)$
-can be shown well-founded under a strategy that
+where $T$ is integer arithmetic extended with the uninterpreted symbol 
+$\con p:\typeint \rightarrow \Bool$ can be shown well-founded under a strategy that
 evaluates the condition of a $\lite$ before evaluating the relevant branch,
@@
 An example of an inadmissible set is
-$\{ \forallf{\con{f}} x : \typeint.\;\, \con{f}( x ) \teq \con{f}( x ) + 1 \}$,
-since the formula $\ltrue$ is (trivially) closed under function expansion with respect to this set,
+$\{ \forallf{\con{f}} x : \typeint.\;\, \con{f}( x ) \teq \con{f}( x ) + 1 \}$
+where $T$ is integer arithmetic extended to a set of uninterpreted symbols 
+$\{\con f:\typeint \rightarrow \typeint,\, \con g:\typeint \rightarrow \typeint,\, \ldots \}$.
+The reason is that the formula $\ltrue$ is (trivially) closed under function expansion with respect to this set,
 and there is no model of $T$
-satisfying $\con{f}$'s definition. A more subtle specimen is
+satisfying $\con{f}$'s definition. A more subtle %ct specimen 
+example 
+is
 \begin{shortv}$\badassex$. \end{shortv}%
@@
 \label{ex:translation}
-Consider the combined signature of linear arithmetic and uninterpreted
-functions $\Sigma = \Sigmalia \mathrel{\uplus} \Sigma_{\euf}$, where
-%$\stypes{\Sigma_{\euf}} = \{ \typeint \}$ 
-%\ct{shouldn't this be $\stypes{\Sigmalia} = \{ \typeint \}$?}
-%and
-% Something went wrong, let's just remove it --ajr
-$\sfuns{\Sigma_{\euf}} = \{ \con{s} : { \typeint \rightarrow \typeint },\, \const{c} : { \typeint } \}$.
+Let $T$ be linear arithmetic with the uninterpreted symbols  $\{ \con{s} : { \typeint \rightarrow \typeint },\, \const{c} : { \typeint } \}$.
 Let $\varphi$ be the $\Sigma$-formula

### 2015-06-04 tex/camera_ready/paper.tex
shortened

@@
 used for recursive function definitions. 
-For the example above, it outputs
-%\ajr{prefer having the full output like this here?}
-%\jb{I like it the way you did it now, but I took the freedom to reformat the output slightly
-%(to avoid the small font---I generally try to avoid ad hoc font mixtures---and to show
-%the nesting more clearly).}
-%it may output the interpretation~(\ref{eq:approx-interp}) from Section~\ref{sec:encoding} for $s$.
-%
-\begin{quote}
-%\small
-\begin{verbatim}
-(model
-  (define-fun s (($x1 Int)) Int 
-    (ite (= $x1 14) 105 (ite (= $x1 13) 91 (ite (= $x1 12) 78 
-      (ite (= $x1 11) 66 (ite (= $x1 10) 55 (ite (= $x1 4) 10 
-        (ite (= $x1 9) 45 (ite (= $x1 8) 36 (ite (= $x1 7) 28 
-          (ite (= $x1 6) 21 (ite (= $x1 3) 6 (ite (= $x1 5) 15 
-            (ite (= $x1 2) 3 (ite (= $x1 1) 1 0)))))))))))))))
-  (define-fun c () Int 14))
-\end{verbatim}
-\end{quote}
+For the example above, it outputs a partial model of~$\con{s}$, where only
+the values of $\con{s}$ for $x = 0,\ldots,14$ are given.
 
@@
 encoding datatypes and recursive functions in Alloy, by approximating datatypes
-by finite subterm-closed substructures. The approach finds sound (fragments
-of) models for formulas in the \relax{existential--bounded-universal} fragment
-\begin{longv}(i.e., formulas whose prenex normal forms contain no unbounded universal
-quantifiers ranging over datatypes)\end{longv}. This idiom was further developed by Dunets
+by finite subterm-closed substructures.
+%The approach finds sound (fragments
+%of) models for formulas in the \relax{existential--bounded-universal} fragment
+%\begin{longv}(i.e., formulas whose prenex normal forms contain no unbounded universal
+%quantifiers ranging over datatypes)\end{longv}.
+This idiom was extended by Dunets
 et al.\ \cite{dunets-et-al-2010}, who presented a translation scheme
-for primitive recursion. Their definedness guards play a similar role to the
-existential constraints generated by our translation $\conv$.
+for primitive recursion. Their definedness guards play a similar role to our
+existential constraints.
 
@@
 tools are tailored for problems that correspond to total functional
-programs. The QuickCheck tool for Haskell \cite{claessen-hughes-2000} is an
+programs. The QuickCheck tool \cite{claessen-hughes-2000} for Haskell is an
 early example, based on random testing. Bounded exhaustive testing\begin{longv}

### 2015-06-04 tex/camera_ready/paper.tex
minor edits

@@
  each function symbol $\con{f} : \tau_1 \times \cdots \times \tau_n \rightarrow \tau$ in $\sfuns{\Sigma}$
-to a total function $\con{f}^\I : \tau_1^\I \times \cdots \times \tau_n^\I \rightarrow \tau^\I$
+to a total function $\con{f}^\I : \tau_1^\I \times \cdots \times \tau_n^\I \rightarrow \tau^\I$,
 and each variable $x:\tau$ of $\svars{\Sigma}$ as an element of $\tau^\I$.
@@
 the \emph{models} of $T$, closed under variable reassignment
-(that is, for every $I \in \Mo$, every $\Sigma$-interpretation that differs
+(i.e., for every $I \in \Mo$, every $\Sigma$-interpretation that differs
 from $I$ only on the variables of $\svars \Sigma$ is also in $\Mo$).
@@
 whose $\Sigma_1$-reduct is a model of $T_1$.
-We refer to the symbols of $\Sigma_2$ that are not is $\Sigma_1$ as \emph{uninterpreted}.
-For the rest of the paper \emph{we fix a theory $T = (\Sigma, \Mo)$ 
+We refer to the symbols of $\Sigma_2$ that are not in $\Sigma_1$ as \emph{uninterpreted}.
+For the rest of the paper, \relax{we fix a theory $T = (\Sigma, \Mo)$ 
 with uninterpreted symbols} constructed as above.
 
-We will consider \emph{annotated quantified formulas} of the form
+We consider \emph{annotated quantified formulas} of the form
 $\forallf{\const{f}} \vec x.\; \varphi$, where $\con{f} \in \sfuns{\Sigma}$ is
@@
 used for recursive function definitions. 
-For the example above, it outputs a partial model of~$\con{s}$, where only
-the values of $\con{s}$ for $x = 0,\ldots,14$ are given.
+For the example above, it outputs a partial model of~$\con{s}$ where only
+the values of $\con{s}(x)$ for $x = 0,\ldots,14$ are given.
 

### 2015-06-04 tex/camera_ready/paper.tex
minor

@@
 
-We consider \emph{annotated quantified formulas} of the form
+Unconventionally, we consider \emph{annotated quantified formulas} of the form
 $\forallf{\const{f}} \vec x.\; \varphi$, where $\con{f} \in \sfuns{\Sigma}$ is

### 2015-06-04 tex/body.tex
fixed typo
more

@@
 
+\newcommand{\svars}[1]{#1^\mathrm{v}}
 \newcommand{\sfuns}[1]{#1^\mathrm{f}}
@@
 for each $\tau \in \stypes{\Sigma}$.
+%
+For each $\tau \in \stypes{\Sigma}$,
+we fix an infinite set $\svars{\Sigma}_\tau$ of \emph{variables of type $\tau$} and
+define $\svars{\Sigma}$ as $\bigcup_{\tau \in \stypes{\Sigma}} \svars{\Sigma}_\tau$.
+$\Sigma$-terms are built as usual over functions symbols in $\Sigma$ and variables in $\svars{\Sigma}$.
+%
 Formulas are terms of type $\Bool$.
@@
 the \emph{domain} of~$\tau$ in~$\I$,
-and each function symbol $\con{f} : \tau_1 \times \cdots \times \tau_n \rightarrow \tau$ in
-$\sfuns{\Sigma}$
-to a total function $\con{f}^\I : \tau_1^\I \times \cdots \times \tau_n^\I \rightarrow \tau^\I$.
+each function symbol $\con{f} : \tau_1 \times \cdots \times \tau_n \rightarrow \tau$ in $\sfuns{\Sigma}$
+to a total function $\con{f}^\I : \tau_1^\I \times \cdots \times \tau_n^\I \rightarrow \tau^\I$,
+and each variable $x:\tau$ of $\svars{\Sigma}$ to an element of $\tau^\I$.
 A \emph{theory} is a pair $T = (\Sigma, \Mo)$ where
@@
 quantifiers altogether by unfolding recursive definitions up to a certain
-depth, which is increased on a per-need basis.
+depth\begin{longv}, which is increased on a per-need basis\end{longv}.
 Our translation~$\conv$ works in an analogous manner, 

### 2015-06-04 tex/camera_ready/paper.tex
more

@@
 to a total function $\con{f}^\I : \tau_1^\I \times \cdots \times \tau_n^\I \rightarrow \tau^\I$,
-and each variable $x:\tau$ of $\svars{\Sigma}$ as an element of $\tau^\I$.
+and each variable $x:\tau$ of $\svars{\Sigma}$ to an element of $\tau^\I$.
 A \emph{theory} is a pair $T = (\Sigma, \Mo)$ where

### 2015-06-04 tex/body.tex
minor fixes + backports

@@
 $\Sigma$ is a signature and $\Mo$ is a class of $\Sigma$-interpretations,
-the \emph{models} of $T$.
-A $\Sigma$-formula $\varphi$ is
-\emph{$T$-satisfiable}
+the \emph{models} of $T$, closed under variable reassignment
+(i.e., for every $I \in \Mo$, every $\Sigma$-interpretation that differs
+from $I$ only on the variables of $\svars \Sigma$ is also in $\Mo$).
+A $\Sigma$-formula $\varphi$ is \emph{$T$-satisfiable}
 if it is satisfied by some interpretation in $\Mo$\begin{longv};
@@
 if each $T$-entails the other.
-Given a signature $\Sigma$,
-the \emph{theory of equality with uninterpreted functions} \euf
-consists of the set of all $\Sigma$-interpretations.
-Informally, we refer to the type and function symbols in this theory as \emph{uninterpreted}.
-%\ct{This is both ambiguous and insufficient.
-%Since this is SMT we should fix some background theory $T$,
-%then extend its signature with additional symbols and call \emph{those} uninterpreted.
-%}
-
-In addition to $\forall \vec x.\; \varphi$,
+If $T_1 = (\Sigma_1, \Mo_1)$ is a theory and $\Sigma_2$ is a signature 
+with $\sfuns{\Sigma_1} \cap \sfuns{\Sigma_2} = \emptyset$,
+the \emph{extension of $T_1$ to $\Sigma_2$} is the theory $T = (\Sigma, \Mo)$ where 
+$\sfuns{\Sigma} = \sfuns{\Sigma_1} \cup \sfuns{\Sigma_2}$,
+$\stypes{\Sigma} = \stypes{\Sigma_1} \cup \stypes{\Sigma_2}$,
+and $\Mo$ is the set of all $\Sigma$-interpretations $\I$
+whose $\Sigma_1$-reduct is a model of $T_1$.
+We refer to the symbols of $\Sigma_2$ that are not in $\Sigma_1$ as \emph{uninterpreted}.
+For the rest of the paper, \relax{we fix a theory $T = (\Sigma, \Mo)$ 
+with uninterpreted symbols} constructed as above.
+
+Unconventionally,
 we consider \emph{annotated quantified formulas} of the form
@@
 uninterpreted. Their
-semantics is the same as for standard quantified formulas.
+semantics is the same as for standard quantified formulas $\forall \vec x.\; \varphi$.
 Given $\con{f} : \tau_1 \times \cdots \times \tau_n \rightarrow \tau$,
@@
 
-Given a signature $\Sigma$, we distinguish the subset of \emph{defined}
-function symbols $\sfundefs{\Sigma} \subseteq \sfuns{\Sigma}$.
+In the signature $\Sigma$, we distinguish a subset $\sfundefs{\Sigma}
+\subseteq \sfuns{\Sigma}$ of \emph{defined} uninterpreted function symbols.
 We consider $\Sigma$-formulas that are in definitional form with respect to
@@
 x \leq 0,\allowbreak\; \con{p}( x ) \teq 1,\allowbreak\; \con{p}( x ) \teq 2 * \con{p}( x - 1 ) \bigr)$
-can be shown well-founded under a strategy that
-evaluates the condition of a $\lite$ before evaluating the relevant branch,
+where $T$ is integer arithmetic extended with the uninterpreted symbol 
+$\con p:\typeint \rightarrow \Bool$, can be shown well-founded under a strategy that
+evaluates the condition of an $\lite$ before evaluating the relevant branch,
 ignoring the other branch. \begin{longv}Logically, such dependencies can be captured by
@@
 $\{ \forallf{\con{f}} x : \typeint.\;\, \con{f}( x ) \teq \con{f}( x ) + 1 \}$,
-since the formula $\ltrue$ is (trivially) closed under function expansion with respect to this set,
+where $T$ is integer arithmetic extended to a set of uninterpreted symbols 
+$\{\con f:\typeint \rightarrow \typeint,\, \con g:\typeint \rightarrow \typeint,\, \ldots \}$.
+The reason is that the formula $\ltrue$ is (trivially) closed under function expansion with respect to this set,
 and there is no model of $T$
-satisfying $\con{f}$'s definition. A more subtle specimen is
+satisfying $\con{f}$'s definition. A more subtle example is
 \begin{shortv}$\badassex$. \end{shortv}%
@@
 \label{ex:translation}
-Consider the combined signature of linear arithmetic and uninterpreted
-functions $\Sigma = \Sigmalia \mathrel{\uplus} \Sigma_{\euf}$, where
-%$\stypes{\Sigma_{\euf}} = \{ \typeint \}$ 
-%\ct{shouldn't this be $\stypes{\Sigmalia} = \{ \typeint \}$?}
-%and
-% Something went wrong, let's just remove it --ajr
-$\sfuns{\Sigma_{\euf}} = \{ \con{s} : { \typeint \rightarrow \typeint },\, \const{c} : { \typeint } \}$.
+Let $T$ be linear arithmetic with the uninterpreted symbols $\{ \con{s} : {
+\typeint \rightarrow \typeint },\, \const{c} : { \typeint } \}$.
 Let $\varphi$ be the $\Sigma$-formula
@@
 used for recursive function definitions. 
-For the example above, it outputs
-%\ajr{prefer having the full output like this here?}
-%\jb{I like it the way you did it now, but I took the freedom to reformat the output slightly
-%(to avoid the small font---I generally try to avoid ad hoc font mixtures---and to show
-%the nesting more clearly).}
-%it may output the interpretation~(\ref{eq:approx-interp}) from Section~\ref{sec:encoding} for $s$.
-%
+For the example above, it outputs a partial model of~$\con{s}$ where only
+the values of $\con{s}(x)$ for $x = 0,\ldots,14$ are correctly given:
 \begin{quote}

### 2015-06-04 tex/camera_ready/paper.tex
minor fixes + backports

@@
 where $T$ is integer arithmetic extended with the uninterpreted symbol 
-$\con p:\typeint \rightarrow \Bool$ can be shown well-founded under a strategy that
-evaluates the condition of a $\lite$ before evaluating the relevant branch,
+$\con p:\typeint \rightarrow \Bool$, can be shown well-founded under a strategy that
+evaluates the condition of an $\lite$ before evaluating the relevant branch,
 ignoring the other branch. \begin{longv}Logically, such dependencies can be captured by
@@
 An example of an inadmissible set is
-$\{ \forallf{\con{f}} x : \typeint.\;\, \con{f}( x ) \teq \con{f}( x ) + 1 \}$
+$\{ \forallf{\con{f}} x : \typeint.\;\, \con{f}( x ) \teq \con{f}( x ) + 1 \}$,
 where $T$ is integer arithmetic extended to a set of uninterpreted symbols 

### 2015-06-05 tex/camera_ready/paper.tex
Fixed a number of errors, generally tightened the text, added some more explanatory text

@@
 Formulas are terms of type $\Bool$.
-We write $t\typ{\tau}$ to denote a metavariable ranging over terms of
-type~$\tau$ and $\terms( t )$ to denote the set of subterms in $t$.
+We write $t\typ{\tau}$ to denote %ct a metavariable ranging over 
+terms of type~$\tau$ and $\terms( t )$ to denote the set of subterms in $t$.
 Given a term $u$, we write $u[\vec t/\vec x]$ to denote the result of replacing
@@
 where $T$ is integer arithmetic extended with the uninterpreted symbol 
-$\con p:\typeint \rightarrow \Bool$, can be shown well-founded under a strategy that
+$\con p:\typeint \rightarrow \typeint$, can be shown well-founded under a strategy that
 evaluates the condition of an $\lite$ before evaluating the relevant branch,
@@
    \begin{itemize}
-     \itemx[] \return \bigtuple{$t,\; \bigwedge\, \{ \exists a : \fargtype{\con{f}}.\; \vecfarg{\con{f}}( a ) \teq \vec t \mid \con{f}( \vec t ) \in \terms( t ),\, \con{f} \in \sfundefs{\Sigma} \}$}
+     \itemx[] \return \bigtuple{$t,\; \bigwedge\, \{ \exists a : \fargtype{\con{f}}.\; \vecfarg{\con{f}}( a ) \teq \vec s \mid \con{f}( \vec s ) \in \terms( t ),\, \con{f} \in \sfundefs{\Sigma} \}$}
 %\\[-.8\baselineskip]
@@
 \end{enumerate}
-\caption{\,Definition of translation $\conv$}
+\caption{\,Definition of translation $\conv$.
+}
 \label{fig:A}
@@
 
-We fix a signature $\Sigma$ and a $\Sigma$-formula
-$\varphi$ in definitional form with respect to $\sfundefs{\Sigma}$,
+For the rest of the section, let $\varphi$ be a $\Sigma$-formula
+in definitional form with respect to $\sfundefs{\Sigma}$,
 whose definitions are admissible.
@@
 $\extendsig{ \Sigma }$ and an $\extendsig{ \Sigma }$-formula $\varphi'$ such that
-$\varphi'$ is satisfiable if and only if $\varphi$ is satisfiable---i.e.,
-$\varphi$ and $\varphi'$ are \emph{equisatisfiable}.
+$\varphi'$ is $T$-satisfiable if and only if $\varphi$ is $T$-satisfiable---i.e.,
+$\varphi$ and $\varphi'$ are \emph{equisatisfiable (in $T$)}.
 The idea behind this translation
 is to use an uninterpreted type to abstract the set of
-\relax{relevant} tuples for each defined function $\con{f}$ and restrict the
+\emph{relevant} tuples for each defined function $\con{f}$ and restrict the
 quantification of $\con{f}$'s definition to a variable of this
-type. Informally, the relevant tuples $\vec t$ of a function $\con{f}$ are the
+type. 
+Informally, the relevant tuples $\vec t$ of a function $\con{f}$ are the
 ones for which the interpretation of $\con{f}( \vec t )$ is relevant to the
 satisfiability of $\varphi$.
-
 More precisely,
-suppose our signature $\Sigma$ contains a set of defined function symbols $\sfundefs{\Sigma} \subseteq \sfuns{\Sigma}$.
-For each $\con{f} : \tau_1 \times \cdots \times \tau_n \rightarrow \tau \in \sfundefs{\Sigma}$,
+%suppose our signature $\Sigma$ contains a set of defined function symbols $\sfundefs{\Sigma} \subseteq \sfuns{\Sigma}$.
+for each $\con{f} : \tau_1 \times \cdots \times \tau_n \rightarrow \tau \in \sfundefs{\Sigma}$,
 the extended signature $\extendsig{\Sigma}$ contains
@@
 The translation $\conv$ defined in Figure~\ref{fig:A} translates the $\Sigma$-formula
-$\varphi$ into an $\extendsig{\Sigma}$-formula. It relies
+$\varphi$ into the $\extendsig{\Sigma}$-formula $\varphi'$. It relies
 on the auxiliary function $\conv_0$, which takes two arguments:\ the term $t$
-to translate and a \relax{polarity}~$p$, which is either $\ppos$, $\pneg$, or
-$\pnone$. \vthinspace$\conv_0$ returns a pair $( t'\negvthinspace,\, \chi )$, where $t'$ is a term of
+to translate and a \relax{polarity}~$p$ for $t$, which is either $\ppos$, $\pneg$, or
+$\pnone$. 
+\vthinspace$\conv_0$ returns a pair $( t'\negvthinspace,\, \chi )$, where $t'$ is a term of
 the same type as $t$ and $\chi$ is a $\extendsig{\Sigma}$-formula.
@@
 
-The translation alters $\varphi$ in two main ways. First, it restricts the
+The translation alters $\varphi$ in two ways. First, it restricts the
 quantification on function definitions for $\con{f}$ to the corresponding
-uninterpreted type $\fargtype{\con{f}}$ and inserts appropriate selectors
-$\fargx{\con{f}}{i}$. Second, it augments $\varphi$ with additional existential
-constraints
-$\exists a : \fargtype{\con{f}}.\;\, \vecfarg{\con{f}}( a ) \teq \vec t$; these
+uninterpreted type $\fargtype{\con{f}}$, inserting applications of the concretization functions $\fargx{\con{f}}{i}$ as needed. 
+Second, it augments $\varphi$ with additional constraints of the form
+$\exists a : \fargtype{\con{f}}.\;\, \vecfarg{\con{f}}( a ) \teq \vec s$
+where, if $\vec s = (s_1,\ldots,s_m)$, $\vecfarg{\con{f}}( a ) \teq \vec s$ abbreviates the formula
+$\bigwedge_{i=1,\ldots,m} \fargx{\con{f}}{i}(a) \teq s_i$.
+These constraints
 ensure that the restricted definition for $\con{f}$ covers all relevant tuples
-of terms, namely the ones for which an $\con{f}$ application exists and is
-relevant to the satisfiability of $\varphi$.
+of terms, namely those occurring in applications of $\con{f}$ 
+that are relevant to the satisfiability of $\varphi$.
 %\jb{what does ``current branch of'' mean? can we drop it / rephrase it / vague
@@
 
-If $t$ is an application of a predicate $\boolop$, including the built-in
+If $t$ is an application of a predicate symbol $\boolop$, including the
 operators ${\oldneg}$, ${\wedge}$, ${\vee}$, ${\teq}$, and ${\lite}$,
 $\conv_0$ calls itself recursively on the arguments $t_i$ and polarity $\pol(
-\boolop, i, p )$, where $\pol$ is defined as
+\boolop, i, p )$, with $\pol$ defined as
 \[\pol( \boolop, i, p ) =
@@
 \end{cases}\]
-where the negation $\pnegate p$ of a polarity $p$ is $\pneg$ if $p$ is $\ppos$, $\ppos$
+where $\pnegate p$ is $\pneg$ if $p$ is $\ppos$, $\ppos$
 if $p$ is $\pneg$, and $\pnone$ if $p$ is $\pnone$.
-If $t$ is positive, $\conv_0$ adds the constraint $\chi$ derived from
-the subterms as a conjunction. Dually, if $t$ is negative, it constructs a
-disjunction with the negated constraint, to achieve the same overall effect as
-in the positive case. It $t$ has no polarity, the constraint $\chi$ is
+The term $t$ is then reconstructed as $\boolop(t_1',\ldots,t_n')$ 
+where each $t_i'$ is the result of the recursive call with argument $t_i$. 
+If the polarity $p$ associated with $t$ is $\ppos$, 
+$\conv_0$ conjunctively adds to $\boolop(t_1',\ldots,t_n')$ the constraint $\chi$ derived from
+the subterms. Dually, if $p$ is $\pneg$, it adds a
+disjunction with the negated constraint, to achieve the same overall effect.
+%ct  as in the positive case. 
+It $p$ is $\pnone$, the constraint $\chi$ is
 returned to the caller.
@@
 $a$ of type $\fargtype{\con{f}}$ and replaces all occurrences of $\vec x$ in
-$t'_1$ with $\vecfarg{\con{f}}( a )$. Since function definitions
-are a top-level conjunct, %by case analysis on the return values of $\conv$,
-$\chi$ must be $\ltrue$.
+the body of that formula with $\vecfarg{\con{f}}( a )$. 
+%ct removed this for consistency because we give no rationale for the other cases when we return true
+%Since function definitions are a top-level conjunct, %by case analysis on the return values of $\conv$,
+%$\chi$ must be $\ltrue$.
+%
 If $t$ is an unannotated quantified formula, $\conv_0$ calls itself
@@
 quantified formula and constraint returned by the recursive call.\end{longv}
-Otherwise, $t$ is either an application of an uninterpreted predicate or a term
-of a type other than $\Bool$. The returned constraint is a conjunction of
+Otherwise, $t$ is either an application of an uninterpreted predicate symbol or a term
+of a type other than $\Bool$. Then, the returned constraint is a conjunction of
 formulas of the form $\exists a : {\fargtype{\con{f}}}.\;\, \vecfarg{\con{f}}(
-a ) \teq \vec t$ for each subterm $\con{f}( \vec t )$ of $t$ such that $\con{f}
+a ) \teq \vec s$ for each subterm $\con{f}( \vec s )$ of $t$ such that $\con{f}
 \in \sfundefs{\Sigma}$. Such constraints, when asserted positively, ensure that
-the tuple of uninterpreted functions $\vecfarg{\con{f}}$ has $\vec t$
-in its range.
+%ct 
+% the tuple of uninterpreted functions $\vecfarg{\con{f}}$ has $\vec s$ in its range.
+%ct I find the following reformulation clearer
+the input tuple $\vec s$ is the preimage of some element in the abstract domain 
+$\fargtype{\con f}$.
 
@@
 and states that the sum of all
-positive numbers up to $\const{c}$ is greater than $100$. This formula is
+positive numbers up to some value $\const{c}$ is greater than $100$. It is
 satisfiable with a model that interprets $\const{c}$ as $14$ or more.
@@
 are unable to find
-a model for this formula. The signature $\extendsig{\Sigma}$ contains the type
-$\fargtype{\con{s}}$ and the uninterpreted function $\farg{\con{s}} : \fargtype{\con{s}}
-\rightarrow \typeint$. The result of $\conv( \varphi )$ 
-is $\varphi'$, which (after simplification) is the $\extendsig{\Sigma}$-formula
+a model for $\varphi$. The signature $\extendsig{\Sigma}$ extends $\Sigma$ with the type
+$\fargtype{\con{s}}$ and the uninterpreted function symbol $\farg{\con{s}} : \fargtype{\con{s}}
+\rightarrow \typeint$. The result of $\conv( \varphi )$, after simplification, 
+is the $\extendsig{\Sigma}$-formula
 %
@@
 %
-The universal quantification in the resulting formula is over an uninterpreted
-type $\fargtype{\con{s}}$, thus enabling the finite model finding
+The universal quantification in the resulting formula~(\ref{eq:ex-after}) is over an uninterpreted
+type $\fargtype{\con{s}}$, making~(\ref{eq:ex-after}) amenable to the finite model finding
 techniques by Reynolds et al.\ \cite{ReyEtAl-1-RR-13,reynolds-et-al-2013},
-implemented in \cvc, to search for a finite interpretation for $\fargtype{\con{s}}$. 
-Since all occurrences of the quantified variable~$a$ in this formula are 
+implemented in \cvc, which search for a finite interpretation for $\fargtype{\con{s}}$. 
+Furthermore, since all occurrences of the quantified variable~$a$ in~(\ref{eq:ex-after}) are 
 beneath applications of the uninterpreted function $\farg{\con{s}}$, 
@@
 a complete instantiation procedure, implemented in \ziii. 
-Accordingly,
+Fittingly,
 %%% Added "apparently". Or are we sure?  
@@
 \cvc and \ziii run indefinitely on formula~(\ref{eq:ex-before}), 
-whereas they produce a model for formula~(\ref{eq:ex-after}) 
+whereas they produce a model for~(\ref{eq:ex-after}) 
 within 100 milliseconds.\xend
@@
 
-The translation $\conv$ results in formulas whose models are different than those of $\varphi$.
+Note that the translation $\conv$ results in formulas whose models
+(i.e., satisfying interpretations) are different in general than those of $\varphi$.
 %\ct{"Preserves models" means that it does not drop models, not that it adds models.}
@@
 otherwise.
-The interpretation $\I$, restricted to $\Sigma$, is {not} a model for the original formula~(\ref{eq:ex-before}),
-since $\I$ wrongly interprets $\con{s}( n )$ as $105$ when $n < 0$ or $n > 14$.
+The $\Sigma$-reduct of $\I$ is not a model of the original formula~(\ref{eq:ex-before}),
+since $\I$ 
+%ct wrongly in what sense?
+%wrongly 
+interprets $\con{s}( n )$ as $105$ when $n < 0$ or $n > 14$.
 
@@
 $\sfundefs{\Sigma}$ are admissible, %we claim that
-$\conv(\varphi)$ is equisatisfiable with $\varphi$ for any input $\varphi$.
-%\ct{Must define "equisatisfiable".
-%What we mean here is a weak form of satisfiability, we could instead say that
-%$\conv(\varphi)$ is $T$-satisfiable iff $\varphi$ is $T$-satisfiable.
-%} 
-%\jb{We do define it, even it is weak form.}
-The intuition is that the
-interpretation of a term such as $\con{s}( -1 )$ in the previous example is not
-relevant to the satisfiability of formula~(\ref{eq:ex-before}). Thus, a
-satisfiable or unsatisfiable response from an SMT solver on input $\conv(
-\varphi )$ implies the existence or nonexistence of a model for
-$\varphi$. Moreover, the models of $\conv( \varphi )$ contain
-pertinent information regarding the models of $\varphi$. For example, the model
+$\conv(\varphi)$ is equisatisfiable with $\varphi$ for any input $\varphi$.\footnote{
+%ct moved in a footnote to avoid breaking the flow
+For the previous example, the intuition is that the
+interpretation of a term such as $\con{s}( -1 )$, say, is not
+relevant to the satisfiability of formula~(\ref{eq:ex-before}). 
+}
+%ct removed because obvious
+%Thus, a ``satisfiable'' or ``unsatisfiable'' response from an SMT solver on input $\conv(
+%\varphi )$ implies the existence or nonexistence of a model for
+%$\varphi$. 
+Moreover, the models of $\conv( \varphi )$ contain
+pertinent information about the models of $\varphi$. For example, the model
 $\I$ for formula~(\ref{eq:ex-after}) given above interprets $\con{c}$ as $14$
@@
 coincides with it on its interpretation of all
-function symbols in $\sfuns{\Sigma} - \sfundefs{\Sigma}$ ($\{
-\con{s} \}$ in the example). Furthermore, the model of $\conv( \varphi )$ will
+function symbols in $\sfuns{\Sigma} - \sfundefs{\Sigma}$ 
+($\{\con{s} \}$ in the example
+\ct{This is not right. To start, the set difference should contain all the arithmetic symbols here. Moreover, isn't $\con s$ in $\sfundefs{\Sigma}$?}). 
+Furthermore, the model of $\conv( \varphi )$ will
 also give correct information for the defined functions at all points belonging
@@
 \begin{theorem}\label{thm:equi}
-If $\varphi$ is a\/ $\Sigma$-formula that is in definitional form with respect to\/
+If $\varphi$ is a\/ $\Sigma$-formula in definitional form with respect to\/
 $\sfundefs{\Sigma}$
@@
 $\sfundefs{\Sigma}$ is admissible, then\/ $\varphi$ and\/ $\conv( \varphi )$ are
-equisatisfiable.
+equisatisfiable in $T$.
 \end{theorem}
@@
 \begin{shortv}
-In following, we give an intuition of the above theorem in the context of an example.
+We give an intuition of the above theorem in the context of an example.
 For a set of ground $\Sigma$-literals $L$, let $\absconstraints( L )$ be the set
@@
 \begin{example}
-Let us revisit the translated formula~(\ref{eq:ex-after}) from
-Example~\ref{ex:translation}. 
-This formula is unsatisfiable only if
-formula~(\ref{eq:ex-before}) is unsatisfiable, since $\fargtype{\con{s}}$ can be
-interpreted as the integers and $\farg{\con{s}}$ as the identity
-function. Conversely, we claim formula~(\ref{eq:ex-after}) is satisfiable
-only if formula~(\ref{eq:ex-before}) is satisfiable, noting that the set
+Let us revisit the formulas in Example~\ref{ex:translation}. 
+%ct rephrased in more direct terms.
+%This formula is unsatisfiable only if
+%formula~(\ref{eq:ex-before}) is unsatisfiable, since $\fargtype{\con{s}}$ can be
+%interpreted as the integers and $\farg{\con{s}}$ as the identity
+%function. 
+If the original formula~(\ref{eq:ex-before}) is $T$-satisfiable then the translated
+formula~(\ref{eq:ex-after}) is clearly also $T$-satisfiable since $\fargtype{\con{s}}$ can be
+interpreted as the integers and $\farg{\con{s}}$ as the identity function. 
+%
+Conversely, we claim that~(\ref{eq:ex-after}) is $T$-satisfiable
+only if~(\ref{eq:ex-before}) is $T$-satisfiable, noting that the set
 $\{ \forallf{\con{s}} x.\;\, \varphi_\con{s} \}$ is admissible,
@@
 where $L_0 = \{ \con{s}( \con{c} ) > 100 \}$
-and $\absconstraints( L_0 )$ is the set consisting of the single constraint
+and $\absconstraints( L_0 )$ consists of the single constraint
 $\exists a : {\fargtype{\con{s}}}.\;\, \farg{\con{s}}( a ) \teq \con{c}$.
@@
 Since $\I$ also satisfies both the translated function definition for $\con{s}$
-(the first conjunct of formula~(\ref{eq:ex-after})) and $\absconstraints( L_0
+(the first conjunct of~(\ref{eq:ex-after})) and $\absconstraints( L_0
 )$, it must also satisfy\begin{longv} the formula $\psi_1$:\end{longv}
@@
 In this section, we evaluate both the overall impact of the translation
-introduced in the previous section and the performance of individual SMT
+introduced in Section~\ref{sec:encoding} and the performance of individual SMT
 techniques.
@@
 variety of inductive theorem provers. They heavily involve
-recursive functions and are limited to the combined theory of uninterpreted
-functions and algebraic datatypes. The second source consists of 166
+recursive functions and are limited to a theory of algebraic datatypes
+with a signature that contains uninterpreted function symbols over these datatypes. 
+The second source consists of 166
 benchmarks from the Leon repository,\footnote{%Available at
@@
 heavily involve recursively defined functions over algebraic datatypes, 
-but cover a wide variety of theories, including bit vectors, arrays, and
+but cover a wide variety of additional theories, including bit vectors, arrays, and
 both linear and nonlinear arithmetic. All benchmarks are in definitional form
@@
 For solvers, we considered the SMT solver \ziii~\cite{de-moura-bjoerner-2008},
-which runs both heuristic methods for quantifier instantiation~\cite{MouraBjoerner07}
-and methods for finding models for quantified formulas~\cite{GeDeM-CAV-09}.
+which runs heuristic methods for quantifier instantiation~\cite{MouraBjoerner07}
+as well as methods for finding models for quantified formulas~\cite{GeDeM-CAV-09}.
 We also considered three configurations of \cvc~\cite{barrett-et-al-2011} which
@@
 \begin{figure}[t]
-\normalsize
+%ct I would not use the normal size in figure makes them look out of proportion
+%\normalsize
+\small
 \centering
@@
 \begin{figure}[t]
-\normalsize
+%\normalsize
+\small
 \centering
@@
 since these benchmarks contain only well-founded function definitions.
-%thereby confirming Theorem~\ref{thm:equi}.
-%\ct{?! \emph{Theorems} do not need corroboration! 
-%Maybe we should say "consistently with Theorem~\ref{thm:equi}"?}
 
@@
 configurations. This confirms the shortcomings of existing SMT techniques for
-finding counterexamples for benchmarks containing recursively defined functions.
+finding models for benchmarks containing recursively defined functions.
 %
@@
 performance of \ziiib for counterexamples also improves dramatically, as it
-solves 112 satisfiable problems, including 10 that were unsolved by \cvcf.
+finds 101 more benchmarks to be satisfiable, including 10 that are not solved by \cvcf.
 %
 We conclude that the translation $\conv$ enables SMT
-solvers to find counterexamples for conjectures involving recursively defined functions
+solvers to find models for conjectures involving recursively defined functions
 whose definitions are admissible.
@@
 
-In terms of number of \emph{unsat} responses, the configuration \cvci is the
+Over \emph{unsat} benchmarks, configuration \cvci is the
 clear winner, finding 318 total benchmarks to be unsatisfiable. Moreover, the
 translation $\conv$ helps \ziii for \emph{unsat} responses as well:
-\ziii solves a total of 209, whereas it solves only 187 before the
+\ziii solves a total of 209 after the translation, whereas it solves only 187 before the
 translation.
 %
-In contrast, \cvcd finds 211 benchmarks unsatisfiable with the translation.
-Thus, in most cases the translation $\conv$ does not significantly degrade performance
-for unsatisfiable benchmarks, and in some cases may actually aid the solver for
-determining unsatisfiability. The exception is \cvci, whose inductive
-arguments are guided by the types of the quantified variables.
+In contrast, \cvcd finds 211 benchmarks unsatisfiable after the translation.
+In general, the translation $\conv$ does not significantly degrade performance
+over unsatisfiable benchmarks, and in some cases it actually aids the solver 
+in determining unsatisfiability. The exception is \cvci whose inductive
+arguments, however, are guided by the types of the quantified variables.
 
@@
 
-The translation $\conv$ significantly improves the effectiveness of known SMT
-techniques for finding counterexamples to formulas involving recursive
+The translation $\conv$ significantly improves the effectiveness of current SMT
+techniques for finding models of formulas involving recursive
 function definitions.
-For convenience, it
-has been implemented in the development version of \cvc %~\cite{ReyEtAl-1-RR-13}
+We have implemented in the development version of \cvc %~\cite{ReyEtAl-1-RR-13}
 (version~1.5 prerelease).
 Function definitions $\forallf{\con{f}} \vec x.\; \varphi$ can be written using the $\definefunreccmd$ command
-from the upcoming SMT-LIB 2.5 standard \cite{smtlib25}.
+from the new version (2.5) of the SMT-LIB standard~\cite{smtlib25}.
 For example, formula~(\ref{eq:ex-before}) from Example~\ref{ex:translation}
@@
 functions introduced using \texttt{define-\allowbreak fun-\allowbreak rec} are admissible.
-Admissibility must be checked \begin{longv}separately \end{longv}\relax{by the user}---e.g., using a
+We stress that admissibility must be verified \begin{longv}separately \end{longv}\relax{by the user}---e.g., using a
 syntactic criterion or a termination prover.
@@
 where instead the SMT solver is invoked only once 
-and where quantifier instantiation amounts to function unfolding.
+and quantifier instantiation is used in lieu of function unfolding.
 \begin{longv}It would be
@@
 
-We presented a translation scheme that makes more SMT problems amenable to
-finite model findings in SMT. In conjunction with finite model finding
-techniques,
-%introduced by Reynolds et al.\ \cite{ReyEtAl-1-RR-13,reynolds-et-al-2013},
-%%% Great work, but let's not cite it more than 4 times if it can be avoided. --jb
-it makes it possible to
-find models of quantified formulas over infinite types, such as integers
-and algebraic datatypes.
+We presented a translation scheme that extends the scope of finite model finding 
+techniques in SMT, allowing one to use them to find models of quantified formulas 
+over \emph{infinite} types, such as integers and algebraic datatypes.
 %\end{nottinyv}
@@
 \cite{reynolds-blanchette-2015-codata}. We also plan to integrate \cvc as a
-counterexample generator in proof assistants. Finally, future work could also
-include identifying further sufficient conditions for admissibility, thereby
+counterexample generator in proof assistants. Further work would also
+include identifying additional sufficient conditions for admissibility, thereby
 enlarging the applicability of the translation scheme presented here.
@@
 C.~Barrett, P.~Fontaine, and C.~Tinelli.
-\newblock The {SMT-LIB} standard---{V}ersion 2.5.
-\newblock \url{http://smt-lib.org/language.shtml}, to appear.
+\newblock {The SMT-LIB Standard: Version 2.5}.
+\newblock Technical report, The University of Iowa, 2015.
+\newblock (Available at \url{http://smt-lib.org}).
 

### 2015-06-05 tex/camera_ready/paper.tex
edits

@@
 evidence that this translation improves the effectiveness of the SMT solvers
-\cvc and \ziii for finding counterexamples to verification conditions\begin{nottinyv}
+\cvc and \ziii for finding countermodels to verification conditions\begin{nottinyv}
 (Section~\ref{sec:evaluation})\end{nottinyv}. The approach is implemented as a preprocessor
@@
 $\forall x : {\typeint}.\;\, \lite\bigl(
-x \leq 0,\allowbreak\; \con{p}( x ) \teq 1,\allowbreak\; \con{p}( x ) \teq 2 * \con{p}( x - 1 ) \bigr)$
+x \leq 0,\allowbreak\; \con{p}( x ) \teq 1,\allowbreak\; \con{p}( x ) \teq 2 * \con{p}( x - 1 ) \bigr)$,
 where $T$ is integer arithmetic extended with the uninterpreted symbol 
@@
 \itemx[\ ]
-$\conv( \varphi )$ $=$ $\mathsf{let}$ $( \varphi',\, \_ ) = \conv_0( \varphi,\, \ppos )$ $\mathsf{in}$ $\varphi'$
+$\conv( \varphi )$ $=$ $\mathsf{let}$ $( \varphi',\, \chi ) = \conv_0( \varphi,\, \ppos )$ $\mathsf{in}$ $\varphi'$
 %\\[-1.25\baselineskip] %% TYPESETTING
 \end{enumerate}
-\caption{\,Definition of translation $\conv$.
-}
+\caption{\,Definition of translation $\conv$}
 \label{fig:A}
@@
 For the rest of the section, let $\varphi$ be a $\Sigma$-formula
-in definitional form with respect to $\sfundefs{\Sigma}$,
+in definitional form with respect to $\sfundefs{\Sigma}$
 whose definitions are admissible.
@@
 $\varphi'$ is $T$-satisfiable if and only if $\varphi$ is $T$-satisfiable---i.e.,
-$\varphi$ and $\varphi'$ are \emph{equisatisfiable (in $T$)}.
+$\varphi$ and $\varphi'$ are \emph{equisatisfiable} (\emph{in $T$}).
 The idea behind this translation
@@
 \vthinspace$\conv_0$ returns a pair $( t'\negvthinspace,\, \chi )$, where $t'$ is a term of
-the same type as $t$ and $\chi$ is a $\extendsig{\Sigma}$-formula.
+the same type as $t$ and $\chi$ is an $\extendsig{\Sigma}$-formula.
 %an additional constraint.
@@
 
-The translation alters $\varphi$ in two ways. First, it restricts the
+The translation alters the formula $\varphi$ in two ways. First, it restricts the
 quantification on function definitions for $\con{f}$ to the corresponding
@@
 Second, it augments $\varphi$ with additional constraints of the form
-$\exists a : \fargtype{\con{f}}.\;\, \vecfarg{\con{f}}( a ) \teq \vec s$
-where, if $\vec s = (s_1,\ldots,s_m)$, $\vecfarg{\con{f}}( a ) \teq \vec s$ abbreviates the formula
-$\bigwedge_{i=1,\ldots,m} \fargx{\con{f}}{i}(a) \teq s_i$.
-These constraints
+$\exists a :\nobreak \fargtype{\con{f}}.\;\, \vecfarg{\con{f}}( a ) \teq \vec s$,
+where $\vecfarg{\con{f}}( a ) \teq \vec s$ abbreviates the formula
+$\bigwedge_{i=1}^m\, \fargx{\con{f}}{i}(a) \teq s_i$
+with $\vec s = (s_1,\ldots,s_m)$.
+These existential %%% for typesetting
+constraints
 ensure that the restricted definition for $\con{f}$ covers all relevant tuples
@@
 that are relevant to the satisfiability of $\varphi$.
-%\jb{what does ``current branch of'' mean? can we drop it / rephrase it / vague
-%it up?}
 
@@
 %ct removed this for consistency because we give no rationale for the other cases when we return true
-%Since function definitions are a top-level conjunct, %by case analysis on the return values of $\conv$,
-%$\chi$ must be $\ltrue$.
+(Since function definitions are top-level conjuncts, %by case analysis on the return values of $\conv$,
+$\chi$ must be $\ltrue$ and can be ignored.)
 %
@@
 formulas of the form $\exists a : {\fargtype{\con{f}}}.\;\, \vecfarg{\con{f}}(
-a ) \teq \vec s$ for each subterm $\con{f}( \vec s )$ of $t$ such that $\con{f}
+a ) \teq \vec s$ for each subterm $\con{f}( \vec s )$ of~$t$ such that $\con{f}
 \in \sfundefs{\Sigma}$. Such constraints, when asserted positively, ensure that
+the tuple of uninterpreted functions $\vecfarg{\con{f}}$ has the argument
+tuple $\vec s$ in its range.
 %ct 
-% the tuple of uninterpreted functions $\vecfarg{\con{f}}$ has $\vec s$ in its range.
+% 
 %ct I find the following reformulation clearer
-the input tuple $\vec s$ is the preimage of some element in the abstract domain 
-$\fargtype{\con f}$.
+%the input tuple $\vec s$ is the preimage of some element in the abstract domain 
+%$\fargtype{\con f}$.
 
@@
 and states that the sum of all
-positive numbers up to some value $\const{c}$ is greater than $100$. It is
+positive numbers up to some constant $\const{c}$ is greater than $100$. It is
 satisfiable with a model that interprets $\const{c}$ as $14$ or more.
@@
 current SMT techniques
-%~\cite{GeDeM-CAV-09, ReyEtAl-1-RR-13,reynolds-et-al-2013}
-%%% We described them in the intro. When I see references like that appear with
-%%% no description, as a reader, I get all paranoid and wonder if it's the old
-%%% stuff from the intro (Ge, Moura, Renoylds et al.) or if it's new stuff ---
-%%% and I cannot remember numbers like [9,18,19] across sections or even
-%%% paragraphs. --jb
 are unable to find
@@
 %
-The universal quantification in the resulting formula~(\ref{eq:ex-after}) is over an uninterpreted
-type $\fargtype{\con{s}}$, making~(\ref{eq:ex-after}) amenable to the finite model finding
+The universal quantification in formula~(\ref{eq:ex-after}) is over an uninterpreted
+type $\fargtype{\con{s}}$, making it amenable to the finite model finding
 techniques by Reynolds et al.\ \cite{ReyEtAl-1-RR-13,reynolds-et-al-2013},
 implemented in \cvc, which search for a finite interpretation for $\fargtype{\con{s}}$. 
-Furthermore, since all occurrences of the quantified variable~$a$ in~(\ref{eq:ex-after}) are 
+Furthermore, since all occurrences of the quantified variable~$a$ are 
 beneath applications of the uninterpreted function $\farg{\con{s}}$, 
@@
 a complete instantiation procedure, implemented in \ziii. 
-Fittingly,
-%%% Added "apparently". Or are we sure?  
+As expected,
+%%% Are we sure?  
 % Yes, I'm planning to post the 2 versions of the benchmark on the evaluation page (when we add it) --ajr
@@
 Note that the translation $\conv$ results in formulas whose models
-(i.e., satisfying interpretations) are different in general than those of $\varphi$.
+(i.e., satisfying interpretations) are generally different from those of $\varphi$.
 %\ct{"Preserves models" means that it does not drop models, not that it adds models.}
@@
 $\sfundefs{\Sigma}$ are admissible, %we claim that
-$\conv(\varphi)$ is equisatisfiable with $\varphi$ for any input $\varphi$.\footnote{
+$\conv(\varphi)$ is equisatisfiable with $\varphi$ for any input $\varphi$.
+%\footnote{
 %ct moved in a footnote to avoid breaking the flow
-For the previous example, the intuition is that the
-interpretation of a term such as $\con{s}( -1 )$, say, is not
-relevant to the satisfiability of formula~(\ref{eq:ex-before}). 
-}
+%jb killed altogether
+%For the previous example, the intuition is that the
+%interpretation of a term such as $\con{s}( -1 )$ is not
+%relevant to the satisfiability of formula~(\ref{eq:ex-before}). 
+%}
 %ct removed because obvious
@@
 coincides with it on its interpretation of all
-function symbols in $\sfuns{\Sigma} - \sfundefs{\Sigma}$ 
-($\{\con{s} \}$ in the example
-\ct{This is not right. To start, the set difference should contain all the arithmetic symbols here. Moreover, isn't $\con s$ in $\sfundefs{\Sigma}$?}). 
+function symbols in $\sfuns{\Sigma} - \sfundefs{\Sigma}$.
+%($\{\con{s} \}$ in the example
+%\ct{This is not right. To start, the set difference should contain all the arithmetic symbols here. Moreover, isn't $\con s$ in $\sfundefs{\Sigma}$?})
 Furthermore, the model of $\conv( \varphi )$ will
@@
 since any $\Sigma$-interpretation that satisfies $\varphi$ can be extended in a straightforward way to 
-a $\extendsig{\Sigma}$-interpretation that satisfies $\conv( \varphi )$.
+an $\extendsig{\Sigma}$-interpretation that satisfies $\conv( \varphi )$.
 Second, if a model is found for $\conv( \varphi )$,
@@
 For a set of ground $\Sigma$-literals $L$, let $\absconstraints( L )$ be the set
-$\{ \exists a : {\fargtype{\con{f}}}.\;\, \vecfarg{\con{f}}( a ) \teq \vec t
-\mid \con{f}( \vec t ) \in \terms( L ),\: \con{f} \in \sfundefs{\Sigma}
+$\{ \exists a : {\fargtype{\con{f}}}.\;\, \vecfarg{\con{f}}( a ) \teq \vec s
+\mid \con{f}( \vec s ) \in \terms( L ),\: \con{f} \in \sfundefs{\Sigma}
 \}$.
@@
 %function. 
-If the original formula~(\ref{eq:ex-before}) is $T$-satisfiable then the translated
+If the original formula~(\ref{eq:ex-before}) is $T$-satisfiable, the translated
 formula~(\ref{eq:ex-after}) is clearly also $T$-satisfiable since $\fargtype{\con{s}}$ can be
@@
 $\{ \forallf{\con{s}} x.\;\, \varphi_\con{s} \}$ is admissible,
-where $\varphi_\con{s}$ is the formula $\lite\bigl( x \leq 0,\; \con{s}(x) \teq 0,\; \con{s}( x ) \teq x + \con{s}( x-1 ) \bigr)$.
+where $\varphi_\con{s}$ is the formula $\lite\bigl( x \leq 0,\allowbreak\; \con{s}(x) \teq\nobreak 0,\allowbreak\; \con{s}( x ) \teq x + \con{s}( x-1 ) \bigr)$.
 %\ct{
@@
 %any model of $\varphi_{\con{s}}[t/x]$ must also satisfy $\varphi_{\con{s}}[t-1/x]$ when $t > 0$. 
-Clearly any interpretation $\I$ satisfying formula~(\ref{eq:ex-after}) satisfies 
+Clearly, any interpretation $\I$ satisfying formula~(\ref{eq:ex-after}) satisfies 
 $L_0 \cup \absconstraints( L_0 )$,
@@
 %\normalsize
-\small
+%\small
 \centering
@@
 including 9~benchmarks in the nonmutated \leon benchmark set. The
-performance of \ziiib for counterexamples also improves dramatically, as it
+performance of \ziiib for countermodels also improves dramatically, as it
 finds 101 more benchmarks to be satisfiable, including 10 that are not solved by \cvcf.
@@
 We conclude that the translation $\conv$ enables SMT
-solvers to find models for conjectures involving recursively defined functions
+solvers to find countermodels for conjectures involving recursively defined functions
 whose definitions are admissible.
 %Both \ziiib and \cvcf were able to find a
-%large number of counterexamples after the translation, with the latter
+%large number of countermodels after the translation, with the latter
 %configuration finding significantly more instances (331 to 112).
 
-Over \emph{unsat} benchmarks, configuration \cvci is the
+Over \emph{unsat} benchmarks, \cvci is the
 clear winner, finding 318 total benchmarks to be unsatisfiable. Moreover, the
 translation $\conv$ helps \ziii for \emph{unsat} responses as well:
-\ziii solves a total of 209 after the translation, whereas it solves only 187 before the
-translation.
-%
-In contrast, \cvcd finds 211 benchmarks unsatisfiable after the translation.
+\ziii solves a total of 209 with the translation, whereas it solves only 187 without it.
+%
+In contrast, \cvcd finds 211 benchmarks unsatisfiable with the translation.
 In general, the translation $\conv$ does not significantly degrade performance
 over unsatisfiable benchmarks, and in some cases it actually aids the solver 
-in determining unsatisfiability. The exception is \cvci whose inductive
-arguments, however, are guided by the types of the quantified variables.
+in determining unsatisfiability. The exception is \cvci, whose inductive
+arguments are guided by the types of the quantified variables.
 
@@
 function definitions.
-We have implemented in the development version of \cvc %~\cite{ReyEtAl-1-RR-13}
+We have implemented it in the development version of \cvc %~\cite{ReyEtAl-1-RR-13}
 (version~1.5 prerelease).
@@
 from the new version (2.5) of the SMT-LIB standard~\cite{smtlib25}.
-For example, formula~(\ref{eq:ex-before}) from Example~\ref{ex:translation}
+Thus, formula~(\ref{eq:ex-before}) from Example~\ref{ex:translation}
 can be specified as %follows:
@@
 functions introduced using \texttt{define-\allowbreak fun-\allowbreak rec} are admissible.
-We stress that admissibility must be verified \begin{longv}separately \end{longv}\relax{by the user}---e.g., using a
+We stress that admissibility must be discharged separately by the user---e.g., using a
 syntactic criterion or a termination prover.
-If it is not met, \cvc may answer \emph{sat} for
+If some function definitions are not admissible, \cvc may answer \emph{sat} for
 an unsatisfiable problem. 
@@
 techniques in SMT, allowing one to use them to find models of quantified formulas 
-over \emph{infinite} types, such as integers and algebraic datatypes.
-%\end{nottinyv}
-
+over \relax{infinite} types, such as integers and algebraic datatypes.
+%
 In future work, it would be interesting to evaluate the approach against other
@@
 \newblock Technical report, The University of Iowa, 2015.
-\newblock (Available at \url{http://smt-lib.org}).
+\newblock Available at \url{http://smt-lib.org/}.
 

### 2015-06-05 tex/camera_ready/paper.tex
more

@@
 \in \sfundefs{\Sigma}$. Such constraints, when asserted positively, ensure that
-the tuple of uninterpreted functions $\vecfarg{\con{f}}$ has the argument
-tuple $\vec s$ in its range.
+some element in the abstract domain $\fargtype{\con f}$ is the preimage of
+the argument tuple $\vec s$. % under $\vecfarg{\con{f}}$.
+
+%the tuple of uninterpreted functions $\vecfarg{\con{f}}$ has the argument
+%tuple $\vec s$ in its range.
 %ct 
@@
 \newblock Technical report, The University of Iowa, 2015.
-\newblock Available at \url{http://smt-lib.org/}.
+\newblock 
+Available at \url{http://smt-lib.org/}.
 

### 2015-06-05 tex/camera_ready/paper.tex
minor

@@
 %
-The universal quantification in formula~(\ref{eq:ex-after}) is over an uninterpreted
+The universal quantification in formula~(\ref{eq:ex-after}) ranges over an uninterpreted
 type $\fargtype{\con{s}}$, making it amenable to the finite model finding

### 2015-06-05 tex/camera_ready/paper.tex
tuning

--- 2015-06-05 tex/camera_ready/paper.tex
@@
 %
-The universal quantification in formula~(\ref{eq:ex-after}) is over an uninterpreted
+The universal quantifier in formula~(\ref{eq:ex-after}) ranges over an uninterpreted
 type $\fargtype{\con{s}}$, making it amenable to the finite model finding

### 2015-06-05 tex/camera_ready/paper.tex
tuning

@@
 \newcommand\badassex{
- \{\forallf{\con{f}} x : \typeint.\;\, \con{f}( x ) \teq \con{f}( x ),\;\,
+ \{\forallf{\con{f}} x : \typeint.\;\, \con{f}( x ) \teq \con{f}( x ),\allowbreak\;\,
  \forallf{\con{g}} x : \typeint.\;\, \con{g}( x ) \teq \con{g}( x ) + \nobreak\con{f}( x )\}}
@@
 where $T$ is integer arithmetic extended to a set of uninterpreted symbols 
-$\{\con f:\typeint \rightarrow \typeint,\, \con g:\typeint \rightarrow \typeint,\, \ldots \}$.
+$\{\con f, \con g :\typeint \rightarrow \typeint,\, \ldots \}$.
 The reason is that the formula $\ltrue$ is (trivially) closed under function expansion with respect to this set,
@@
 where $\vecfarg{\con{f}}( a ) \teq \vec s$ abbreviates the formula
-$\bigwedge_{i=1}^m\, \fargx{\con{f}}{i}(a) \teq s_i$
-with $\vec s = (s_1,\ldots,s_m)$.
+$\bigwedge_{i=1}^n\, \fargx{\con{f}}{i}(a) \teq s_i$
+with $\vec s = (s_1,\ldots,s_n)$.
 These existential %%% for typesetting

### 2015-06-05 tex/camera_ready/paper.tex
small is back

@@
 \label{ex:translation}
-Let $T$ be linear arithmetic with the uninterpreted symbols  $\{ \con{s} : { \typeint \rightarrow \typeint },\, \const{c} : { \typeint } \}$.
+Let $T$ be linear arithmetic with the uninterpreted symbols  $\{ \const{c} : { \typeint },\> \con{s} : { \typeint \rightarrow \typeint } \}$.
 Let $\varphi$ be the $\Sigma$-formula
@@
 %\small
+\small
 \centering
@@
 \begin{figure}[t]
+\small
 %\normalsize

### 2015-06-05 tex/body.tex
more

@@
 $\forall x : {\typeint}.\;\, \lite\bigl(
-x \leq 0,\allowbreak\; \con{p}( x ) \teq 1,\allowbreak\; \con{p}( x ) \teq 2 * \con{p}( x - 1 ) \bigr)$
+x \leq 0,\allowbreak\; \con{p}( x ) \teq 1,\allowbreak\; \con{p}( x ) \teq 2 * \con{p}( x - 1 ) \bigr)$,
 where $T$ is integer arithmetic extended with the uninterpreted symbol 
-$\con p:\typeint \rightarrow \Bool$, can be shown well-founded under a strategy that
+$\con p:\typeint \rightarrow \typeint$, can be shown well-founded under a strategy that
 evaluates the condition of an $\lite$ before evaluating the relevant branch,
@@
 \itemx[\ ]
-$\conv( \varphi )$ $=$ $\mathsf{let}$ $( \varphi',\, \_ ) = \conv_0( \varphi,\, \ppos )$ $\mathsf{in}$ $\varphi'$
+$\conv( \varphi )$ $=$ $\mathsf{let}$ $( \varphi',\, \chi ) = \conv_0( \varphi,\, \ppos )$ $\mathsf{in}$ $\varphi'$
 %\\[-1.25\baselineskip] %% TYPESETTING
@@
 
-We fix a signature $\Sigma$ and a $\Sigma$-formula
-$\varphi$ in definitional form with respect to $\sfundefs{\Sigma}$,
+For the rest of the section, let $\varphi$ be a $\Sigma$-formula
+in definitional form with respect to $\sfundefs{\Sigma}$
 whose definitions are admissible.
@@
 $\extendsig{ \Sigma }$ and an $\extendsig{ \Sigma }$-formula $\varphi'$ such that
-$\varphi'$ is satisfiable if and only if $\varphi$ is satisfiable---i.e.,
-$\varphi$ and $\varphi'$ are \emph{equisatisfiable}.
+$\varphi'$ is $T$-satisfiable if and only if $\varphi$ is $T$-satisfiable---i.e.,
+$\varphi$ and $\varphi'$ are \emph{equisatisfiable} (\emph{in $T$}).
 The idea behind this translation
 is to use an uninterpreted type to abstract the set of
-\relax{relevant} tuples for each defined function $\con{f}$ and restrict the
+\emph{relevant} tuples for each defined function $\con{f}$ and restrict the
 quantification of $\con{f}$'s definition to a variable of this
-type. Informally, the relevant tuples $\vec t$ of a function $\con{f}$ are the
+type. 
+Informally, the relevant tuples $\vec t$ of a function $\con{f}$ are the
 ones for which the interpretation of $\con{f}( \vec t )$ is relevant to the
 satisfiability of $\varphi$.
-
 More precisely,
-suppose our signature $\Sigma$ contains a set of defined function symbols $\sfundefs{\Sigma} \subseteq \sfuns{\Sigma}$.
-For each $\con{f} : \tau_1 \times \cdots \times \tau_n \rightarrow \tau \in \sfundefs{\Sigma}$,
+%suppose our signature $\Sigma$ contains a set of defined function symbols $\sfundefs{\Sigma} \subseteq \sfuns{\Sigma}$.
+for each $\con{f} : \tau_1 \times \cdots \times \tau_n \rightarrow \tau \in \sfundefs{\Sigma}$,
 the extended signature $\extendsig{\Sigma}$ contains
@@
 The translation $\conv$ defined in Figure~\ref{fig:A} translates the $\Sigma$-formula
-$\varphi$ into an $\extendsig{\Sigma}$-formula. It relies
+$\varphi$ into the $\extendsig{\Sigma}$-formula $\varphi'$. It relies
 on the auxiliary function $\conv_0$, which takes two arguments:\ the term $t$
-to translate and a \relax{polarity}~$p$, which is either $\ppos$, $\pneg$, or
-$\pnone$. \vthinspace$\conv_0$ returns a pair $( t'\negvthinspace,\, \chi )$, where $t'$ is a term of
-the same type as $t$ and $\chi$ is a $\extendsig{\Sigma}$-formula.
+to translate and a \relax{polarity}~$p$ for $t$, which is either $\ppos$, $\pneg$, or
+$\pnone$. 
+\vthinspace$\conv_0$ returns a pair $( t'\negvthinspace,\, \chi )$, where $t'$ is a term of
+the same type as $t$ and $\chi$ is an $\extendsig{\Sigma}$-formula.
 %an additional constraint.
@@
 
-The translation alters $\varphi$ in two main ways. First, it restricts the
+The translation alters the formula $\varphi$ in two ways. First, it restricts the
 quantification on function definitions for $\con{f}$ to the corresponding
-uninterpreted type $\fargtype{\con{f}}$ and inserts appropriate selectors
-$\fargx{\con{f}}{i}$. Second, it augments $\varphi$ with additional existential
+uninterpreted type $\fargtype{\con{f}}$, inserting applications of the concretization functions $\fargx{\con{f}}{i}$ as needed. 
+Second, it augments $\varphi$ with additional constraints of the form
+$\exists a :\nobreak \fargtype{\con{f}}.\;\, \vecfarg{\con{f}}( a ) \teq \vec s$,
+where $\vecfarg{\con{f}}( a ) \teq \vec s$ abbreviates the formula
+$\bigwedge_{i=1}^n\, \fargx{\con{f}}{i}(a) \teq s_i$
+with $\vec s = (s_1,\ldots,s_n)$.
+These existential %%% for typesetting
 constraints
-$\exists a : \fargtype{\con{f}}.\;\, \vecfarg{\con{f}}( a ) \teq \vec t$; these
 ensure that the restricted definition for $\con{f}$ covers all relevant tuples
-of terms, namely the ones for which an $\con{f}$ application exists and is
-relevant to the satisfiability of $\varphi$.
-%\jb{what does ``current branch of'' mean? can we drop it / rephrase it / vague
-%it up?}
-
-If $t$ is an application of a predicate $\boolop$, including the built-in
+of terms, namely those occurring in applications of $\con{f}$ 
+that are relevant to the satisfiability of $\varphi$.
+
+If $t$ is an application of a predicate symbol $\boolop$, including the
 operators ${\oldneg}$, ${\wedge}$, ${\vee}$, ${\teq}$, and ${\lite}$,
 $\conv_0$ calls itself recursively on the arguments $t_i$ and polarity $\pol(
-\boolop, i, p )$, where $\pol$ is defined as
+\boolop, i, p )$, with $\pol$ defined as
 \[\pol( \boolop, i, p ) =
@@
 $a$ of type $\fargtype{\con{f}}$ and replaces all occurrences of $\vec x$ in
-$t'_1$ with $\vecfarg{\con{f}}( a )$. Since function definitions
-are a top-level conjunct, %by case analysis on the return values of $\conv$,
-$\chi$ must be $\ltrue$.
+the body of that formula with $\vecfarg{\con{f}}( a )$. 
+(Since function definitions are top-level conjuncts, %by case analysis on the return values of $\conv$,
+$\chi$ must be $\ltrue$ and can be ignored.)
+%
 If $t$ is an unannotated quantified formula, $\conv_0$ calls itself
@@
 quantified formula and constraint returned by the recursive call.\end{longv}
-Otherwise, $t$ is either an application of an uninterpreted predicate or a term
-of a type other than $\Bool$. The returned constraint is a conjunction of
+Otherwise, $t$ is either an application of an uninterpreted predicate symbol or a term
+of a type other than $\Bool$. Then, the returned constraint is a conjunction of
 formulas of the form $\exists a : {\fargtype{\con{f}}}.\;\, \vecfarg{\con{f}}(
-a ) \teq \vec t$ for each subterm $\con{f}( \vec t )$ of $t$ such that $\con{f}
+a ) \teq \vec s$ for each subterm $\con{f}( \vec s )$ of~$t$ such that $\con{f}
 \in \sfundefs{\Sigma}$. Such constraints, when asserted positively, ensure that
-the tuple of uninterpreted functions $\vecfarg{\con{f}}$ has $\vec t$
-in its range.
+some element in the abstract domain $\fargtype{\con f}$ is the preimage of
+the argument tuple $\vec s$. % under $\vecfarg{\con{f}}$.
 
@@
 \label{ex:translation}
-Let $T$ be linear arithmetic with the uninterpreted symbols $\{ \con{s} : {
-\typeint \rightarrow \typeint },\, \const{c} : { \typeint } \}$.
+Let $T$ be linear arithmetic with the uninterpreted symbols  $\{ \const{c} : { \typeint },\> \con{s} : { \typeint \rightarrow \typeint } \}$.
 Let $\varphi$ be the $\Sigma$-formula
@@
 and states that the sum of all
-positive numbers up to $\const{c}$ is greater than $100$. This formula is
+positive numbers up to some constant $\const{c}$ is greater than $100$. It is
 satisfiable with a model that interprets $\const{c}$ as $14$ or more.
@@
 current SMT techniques
-%~\cite{GeDeM-CAV-09, ReyEtAl-1-RR-13,reynolds-et-al-2013}
-%%% We described them in the intro. When I see references like that appear with
-%%% no description, as a reader, I get all paranoid and wonder if it's the old
-%%% stuff from the intro (Ge, Moura, Renoylds et al.) or if it's new stuff ---
-%%% and I cannot remember numbers like [9,18,19] across sections or even
-%%% paragraphs. --jb
 are unable to find
-a model for this formula. The signature $\extendsig{\Sigma}$ contains the type
-$\fargtype{\con{s}}$ and the uninterpreted function $\farg{\con{s}} : \fargtype{\con{s}}
-\rightarrow \typeint$. The result of $\conv( \varphi )$ 
-is $\varphi'$, which (after simplification) is the $\extendsig{\Sigma}$-formula
+a model for $\varphi$. The signature $\extendsig{\Sigma}$ extends $\Sigma$ with the type
+$\fargtype{\con{s}}$ and the uninterpreted function symbol $\farg{\con{s}} : \fargtype{\con{s}}
+\rightarrow \typeint$. The result of $\conv( \varphi )$, after simplification, 
+is the $\extendsig{\Sigma}$-formula
 %
@@
 %
-The universal quantification in the resulting formula is over an uninterpreted
-type $\fargtype{\con{s}}$, thus enabling the finite model finding
+The universal quantifier in formula~(\ref{eq:ex-after}) ranges over an uninterpreted
+type $\fargtype{\con{s}}$, making it amenable to the finite model finding
 techniques by Reynolds et al.\ \cite{ReyEtAl-1-RR-13,reynolds-et-al-2013},
-implemented in \cvc, to search for a finite interpretation for $\fargtype{\con{s}}$. 
-Since all occurrences of the quantified variable~$a$ in this formula are 
+implemented in \cvc, which search for a finite interpretation for $\fargtype{\con{s}}$. 
+Furthermore, since all occurrences of the quantified variable~$a$ are 
 beneath applications of the uninterpreted function $\farg{\con{s}}$, 
@@
 a complete instantiation procedure, implemented in \ziii. 
-Accordingly,
-%%% Added "apparently". Or are we sure?  
+As expected,
+%%% Are we sure?  
 % Yes, I'm planning to post the 2 versions of the benchmark on the evaluation page (when we add it) --ajr
@@
 
-The translation $\conv$ results in formulas whose models are different than those of $\varphi$.
+Note that the translation $\conv$ results in formulas whose models
+(i.e., satisfying interpretations) are generally different from those of $\varphi$.
 %\ct{"Preserves models" means that it does not drop models, not that it adds models.}
@@
 $\conv(\varphi)$ is equisatisfiable with $\varphi$ for any input $\varphi$.
-%\ct{Must define "equisatisfiable".
-%What we mean here is a weak form of satisfiability, we could instead say that
-%$\conv(\varphi)$ is $T$-satisfiable iff $\varphi$ is $T$-satisfiable.
+%\footnote{
+%ct moved in a footnote to avoid breaking the flow
+%jb killed altogether
+%For the previous example, the intuition is that the
+%interpretation of a term such as $\con{s}( -1 )$ is not
+%relevant to the satisfiability of formula~(\ref{eq:ex-before}). 
 %}
-%\jb{We do define it, even it is weak form.}
-The intuition is that the
-interpretation of a term such as $\con{s}( -1 )$ in the previous example is not
-relevant to the satisfiability of formula~(\ref{eq:ex-before}). Thus, a
-satisfiable or unsatisfiable response from an SMT solver on input $\conv(
-\varphi )$ implies the existence or nonexistence of a model for
-$\varphi$. Moreover, the models of $\conv( \varphi )$ contain
-pertinent information regarding the models of $\varphi$. For example, the model
+%ct removed because obvious
+%Thus, a ``satisfiable'' or ``unsatisfiable'' response from an SMT solver on input $\conv(
+%\varphi )$ implies the existence or nonexistence of a model for
+%$\varphi$. 
+Moreover, the models of $\conv( \varphi )$ contain
+pertinent information about the models of $\varphi$. For example, the model
 $\I$ for formula~(\ref{eq:ex-after}) given above interprets $\con{c}$ as $14$
@@
 coincides with it on its interpretation of all
-function symbols in $\sfuns{\Sigma} - \sfundefs{\Sigma}$ ($\{
-\con{s} \}$ in the example). Furthermore, the model of $\conv( \varphi )$ will
+function symbols in $\sfuns{\Sigma} - \sfundefs{\Sigma}$.
+%($\{\con{s} \}$ in the example
+%\ct{This is not right. To start, the set difference should contain all the arithmetic symbols here. Moreover, isn't $\con s$ in $\sfundefs{\Sigma}$?})
+Furthermore, the model of $\conv( \varphi )$ will
 also give correct information for the defined functions at all points belonging
@@
 \begin{shortv}
-In following, we give an intuition of the above theorem in the context of an example.
+We give an intuition of the above theorem in the context of an example.
 For a set of ground $\Sigma$-literals $L$, let $\absconstraints( L )$ be the set
-$\{ \exists a : {\fargtype{\con{f}}}.\;\, \vecfarg{\con{f}}( a ) \teq \vec t
-\mid \con{f}( \vec t ) \in \terms( L ),\: \con{f} \in \sfundefs{\Sigma}
+$\{ \exists a : {\fargtype{\con{f}}}.\;\, \vecfarg{\con{f}}( a ) \teq \vec s
+\mid \con{f}( \vec s ) \in \terms( L ),\: \con{f} \in \sfundefs{\Sigma}
 \}$.
@@
 \begin{example}
-Let us revisit the translated formula~(\ref{eq:ex-after}) from
-Example~\ref{ex:translation}. 
-This formula is unsatisfiable only if
-formula~(\ref{eq:ex-before}) is unsatisfiable, since $\fargtype{\con{s}}$ can be
-interpreted as the integers and $\farg{\con{s}}$ as the identity
-function. Conversely, we claim formula~(\ref{eq:ex-after}) is satisfiable
-only if formula~(\ref{eq:ex-before}) is satisfiable, noting that the set
+Let us revisit the formulas in Example~\ref{ex:translation}. 
+%ct rephrased in more direct terms.
+%This formula is unsatisfiable only if
+%formula~(\ref{eq:ex-before}) is unsatisfiable, since $\fargtype{\con{s}}$ can be
+%interpreted as the integers and $\farg{\con{s}}$ as the identity
+%function. 
+If the original formula~(\ref{eq:ex-before}) is $T$-satisfiable, the translated
+formula~(\ref{eq:ex-after}) is clearly also $T$-satisfiable since $\fargtype{\con{s}}$ can be
+interpreted as the integers and $\farg{\con{s}}$ as the identity function. 
+%
+Conversely, we claim that~(\ref{eq:ex-after}) is $T$-satisfiable
+only if~(\ref{eq:ex-before}) is $T$-satisfiable, noting that the set
 $\{ \forallf{\con{s}} x.\;\, \varphi_\con{s} \}$ is admissible,
-where $\varphi_\con{s}$ is the formula $\lite\bigl( x \leq 0,\; \con{s}(x) \teq 0,\; \con{s}( x ) \teq x + \con{s}( x-1 ) \bigr)$.
+where $\varphi_\con{s}$ is the formula $\lite\bigl( x \leq 0,\allowbreak\; \con{s}(x) \teq\nobreak 0,\allowbreak\; \con{s}( x ) \teq x + \con{s}( x-1 ) \bigr)$.
 %\ct{
@@
 %any model of $\varphi_{\con{s}}[t/x]$ must also satisfy $\varphi_{\con{s}}[t-1/x]$ when $t > 0$. 
-Clearly any interpretation $\I$ satisfying formula~(\ref{eq:ex-after}) satisfies 
+Clearly, any interpretation $\I$ satisfying formula~(\ref{eq:ex-after}) satisfies 
 $L_0 \cup \absconstraints( L_0 )$,
 where $L_0 = \{ \con{s}( \con{c} ) > 100 \}$
-and $\absconstraints( L_0 )$ is the set consisting of the single constraint
+and $\absconstraints( L_0 )$ consists of the single constraint
 $\exists a : {\fargtype{\con{s}}}.\;\, \farg{\con{s}}( a ) \teq \con{c}$.
@@
 \begin{figure}[t]
-\normalsize
+%ct I would not use the normal size in figure makes them look out of proportion
+%\normalsize
+%\small
+\small
 \centering
@@
 \begin{figure}[t]
-\normalsize
+\small
+%\normalsize
+%\small
 \centering
@@
 including 9~benchmarks in the nonmutated \leon benchmark set. The
-performance of \ziiib for counterexamples also improves dramatically, as it
-solves 112 satisfiable problems, including 10 that were unsolved by \cvcf.
+performance of \ziiib for countermodels also improves dramatically, as it
+finds 101 more benchmarks to be satisfiable, including 10 that are not solved by \cvcf.
 %
 We conclude that the translation $\conv$ enables SMT
-solvers to find counterexamples for conjectures involving recursively defined functions
+solvers to find countermodels for conjectures involving recursively defined functions
 whose definitions are admissible.
 %Both \ziiib and \cvcf were able to find a
-%large number of counterexamples after the translation, with the latter
+%large number of countermodels after the translation, with the latter
 %configuration finding significantly more instances (331 to 112).
 
-In terms of number of \emph{unsat} responses, the configuration \cvci is the
+Over \emph{unsat} benchmarks, \cvci is the
 clear winner, finding 318 total benchmarks to be unsatisfiable. Moreover, the
 translation $\conv$ helps \ziii for \emph{unsat} responses as well:
-\ziii solves a total of 209, whereas it solves only 187 before the
-translation.
+\ziii solves a total of 209 with the translation, whereas it solves only 187 without it.
 %
 In contrast, \cvcd finds 211 benchmarks unsatisfiable with the translation.
-Thus, in most cases the translation $\conv$ does not significantly degrade performance
-for unsatisfiable benchmarks, and in some cases may actually aid the solver for
-determining unsatisfiability. The exception is \cvci, whose inductive
+In general, the translation $\conv$ does not significantly degrade performance
+over unsatisfiable benchmarks, and in some cases it actually aids the solver 
+in determining unsatisfiability. The exception is \cvci, whose inductive
 arguments are guided by the types of the quantified variables.
@@
 
-The translation $\conv$ significantly improves the effectiveness of known SMT
-techniques for finding counterexamples to formulas involving recursive
+The translation $\conv$ significantly improves the effectiveness of current SMT
+techniques for finding models of formulas involving recursive
 function definitions.
-For convenience, it
-has been implemented in the development version of \cvc %~\cite{ReyEtAl-1-RR-13}
+We have implemented it in the development version of \cvc %~\cite{ReyEtAl-1-RR-13}
 (version~1.5 prerelease).
 Function definitions $\forallf{\con{f}} \vec x.\; \varphi$ can be written using the $\definefunreccmd$ command
-from the upcoming SMT-LIB 2.5 standard \cite{smtlib25}.
-For example, formula~(\ref{eq:ex-before}) from Example~\ref{ex:translation}
+from the new version (2.5) of the SMT-LIB standard~\cite{smtlib25}.
+Thus, formula~(\ref{eq:ex-before}) from Example~\ref{ex:translation}
 can be specified as %follows:
@@
 functions introduced using \texttt{define-\allowbreak fun-\allowbreak rec} are admissible.
-Admissibility must be checked \begin{longv}separately \end{longv}\relax{by the user}---e.g., using a
+We stress that admissibility must be discharged separately by the user---e.g., using a
 syntactic criterion or a termination prover.
-If it is not met, \cvc may answer \emph{sat} for
+If some function definitions are not admissible, \cvc may answer \emph{sat} for
 an unsatisfiable problem. 
@@
 tools are tailored for problems that correspond to total functional
-programs. The QuickCheck tool \cite{claessen-hughes-2000} for Haskell is an
+programs. QuickCheck \cite{claessen-hughes-2000} for Haskell is an
 early example, based on random testing. Bounded exhaustive testing\begin{longv}
@@
 
-We presented a translation scheme that makes more SMT problems amenable to
-finite model findings in SMT. In conjunction with finite model finding
-techniques,
-%introduced by Reynolds et al.\ \cite{ReyEtAl-1-RR-13,reynolds-et-al-2013},
-%%% Great work, but let's not cite it more than 4 times if it can be avoided. --jb
-it makes it possible to
-find models of quantified formulas over infinite types, such as integers
-and algebraic datatypes.
-%\end{nottinyv}
-
+We presented a translation scheme that extends the scope of finite model finding 
+techniques in SMT, allowing one to use them to find models of quantified formulas 
+over \relax{infinite} types, such as integers and algebraic datatypes.
+%
 In future work, it would be interesting to evaluate the approach against other

### 2015-06-06 tex/body.tex
Fix option name

@@
 If \cvc's finite model finding mode for recursive functions is enabled (using
-the command-line option \texttt{--fmf-fun}), it will replace its list of known
+the command-line option \texttt{-}\texttt{-fmf-fun}), it will replace its list of known
 assertions based on the translation $\conv$ before checking for satisfiability.
@@
 
-The \texttt{--fmf-fun} option tells \cvc to assume that
+The \texttt{-}\texttt{-fmf-fun} option tells \cvc to assume that
 functions introduced using \texttt{define-\allowbreak fun-\allowbreak rec} are admissible.
@@
 %
-to the above problem and run \cvc with the \texttt{--fmf-fun} option,
+to the above problem and run \cvc with the \texttt{-}\texttt{-fmf-fun} option,
 it wrongly answers \emph{sat}.

### 2015-11-24 smt2015/body.tex
removed, tuned

--- 2015-06-06 tex/body.tex
@@
 by finite subterm-closed substructures. The approach finds sound (fragments
-of) models for formulas in the \relax{existential--bounded-universal} fragment
-\begin{longv}(i.e., formulas whose prenex normal forms contain no unbounded universal
+of) models for formulas in the \relax{existential--bounded-universal} fragment%
+\begin{longv} (i.e., formulas whose prenex normal forms contain no unbounded universal
 quantifiers ranging over datatypes)\end{longv}. This idiom was further developed by Dunets

### 2015-11-24 ijcar2016/body.tex
working on IJCAR 2016 setup

--- 2015-11-24 smt2015/body.tex
@@
-%\documentclass[runningheads,a4paper]{llncs}
-\usepackage[T1]{fontenc}
-\usepackage{amsmath}
-\usepackage{amssymb}
-\usepackage[scaled=.82]{beramono}
-\usepackage{mathptmx}
-\usepackage{listings}
-\usepackage{colonequals}
-%\usepackage{mathpartir}
-\usepackage{xspace}
-\setcounter{tocdepth}{3}
-\usepackage{graphicx}
-\usepackage{stmaryrd}
-\usepackage{cite}
-\usepackage{amsthm}
-%\usepackage{framed}
-
-% http://mirror.switch.ch/ftp/mirror/tex/macros/latex/contrib/xcolor/xcolor.pdf
-\usepackage{xcolor}
-
-%\usepackage{amsthm}
-\usepackage{amsmath}    % need for subequations
-\usepackage{booktabs}
-\usepackage{version}
-%\usepackage{float}
-%\usepackage{fullpage}
-%\floatstyle{ruled}
-%\newfloat{alg}{ht}{loa}
-%\floatname{alg}{Algorithm}
-
-\theoremstyle{plain}
-\newtheorem{theorem}{Theorem}%[section]
-\newtheorem{lemma}[theorem]{Lemma}%[section]
-\newtheorem{corollary}[theorem]{Corollary}%[section]
-
-\theoremstyle{definition}
-\newtheorem{definition}{Definition}%[section]
-
-\theoremstyle{remark}
-\newtheorem*{proofsketch}{Proof sketch}
-\newtheorem{example}{Example}%[section]
-
-%\spnewtheorem*{proof}{Proof sketch}{\itshape}{\rmfamily}
+%\theoremstyle{plain}
+%\newtheorem{theorem}{Theorem}%[section]
+%\newtheorem{lemma}[theorem]{Lemma}%[section]
+%\newtheorem{corollary}[theorem]{Corollary}%[section]
+
+%\theoremstyle{definition}
+%\newtheorem{definition}{Definition}%[section]
+
+%\theoremstyle{remark}
+%\newtheorem*{proofsketch}{Proof sketch}
+%\newtheorem{example}{Example}%[section]
+
+\spnewtheorem*{proofsketch}{Proof sketch}{\itshape}{\rmfamily}
 
@@
 
-\author {Andrew Reynolds\inst{1} \and Jasmin Christian Blanchette\inst{2,3} \and Cesare Tinelli \inst{4}}
+\author {Andrew Reynolds\inst{1} \and Jasmin Christian Blanchette\inst{2,3} \and \\ Simon~Cruanes\inst{2} \and Cesare~Tinelli \inst{1}}
 \authorrunning {A. Reynolds, J. C. Blanchette, C. Tinelli}
 \institute{
-\'Ecole Polytechnique F\'ed\'erale de Lausanne (EPFL), Switzerland
+Department of Computer Science, The University of Iowa, USA
 \and
-Inria Nancy \& LORIA, Villers-l\`es-Nancy, France
+Inria Nancy -- Grand-Est, Villers-l\`es-Nancy, France
 \and
 Max-Planck-Institut f\"ur Informatik, Saarbr\"ucken, Germany
-\and
-Department of Computer Science, The University of Iowa, USA
+%\and
+%\'Ecole Polytechnique F\'ed\'erale de Lausanne (EPFL), Switzerland
 }
@@
 
-%\begin{nottinyv}
 \begin{abstract}
@@
 \label{sec:introduction}
-%\end{nottinyv}
 
@@
 definitions into formulas where finite model finding techniques can be applied.
-The recursive functions must meet a semantic criterion to be admissible\begin{nottinyv}
-(Section~\ref{sec:prelim})\end{nottinyv}. This criterion is met by well-founded (terminating)
-recursive function definitions. \begin{longv}It is not met by inconsistent
+The recursive functions must meet a semantic criterion to be admissible
+(Section~\ref{sec:prelim}). This criterion is met by well-founded (terminating)
+recursive function definitions. \begin{rep}It is not met by inconsistent
 definitions such as $\forall x : {\typeint}.\;\, \con{f}(x) \teq \con{f}(x) +
-1$.\end{longv} %ct and it is possible to construct consistent examples that do not meet it.
+1$.\end{rep} %ct and it is possible to construct consistent examples that do not meet it.
 %as well as other example of consistent definition.
@@
 We define a translation for a class of
-formulas involving admissible recursive function definitions\begin{nottinyv}
-(Section~\ref{sec:encoding})\end{nottinyv}. The main insight is that a
+formulas involving admissible recursive function definitions
+(Section~\ref{sec:encoding}). The main insight is that a
 recursive definition $\forall x : \tau.\;\, \const{f}(x) \teq t$
@@
 \fargtype{\tau}.\;\, \const{f}(\farg{\con{f}}(a)) \teq t[\farg{\con{f}}(a)/x]$, where
-$\fargtype{\tau}$ is an uninterpreted \relax{abstract} type and $\farg{\con{f}} \begin{longv}:
-\fargtype{\tau} \to \tau\end{longv}$ converts the abstract type to the concrete
+$\fargtype{\tau}$ is an uninterpreted \relax{abstract} type and $\farg{\con{f}} \begin{rep}:
+\fargtype{\tau} \to \tau\end{rep}$ converts the abstract type to the concrete
 type. Additional constraints ensure that the abstract values that are relevant
 to the formula's satisfiability
-exist. The translation preserves satisfiability and\begin{longv}, for admissible definitions,\end{longv}
+exist. The translation preserves satisfiability and\begin{rep}, for admissible definitions,\end{rep}
 unsatisfiability, and makes finite model finding possible for problems in this
-class.\begin{shortv} Detailed proofs of correctness are included in a technical
-report \cite{our-report}.\end{shortv}
+class.\begin{conf} Detailed proofs of correctness are included in a technical
+report \cite{our-report}.\end{conf}
 
@@
 evidence that this translation improves the effectiveness of the SMT solvers
-\cvc and \ziii for finding countermodels to verification conditions\begin{nottinyv}
-(Section~\ref{sec:evaluation})\end{nottinyv}. The approach is implemented as a preprocessor
-in \cvc \begin{nottinyv}(Section~\ref{sec:front-end})\end{nottinyv}. \begin{longv}%
+\cvc and \ziii for finding countermodels to verification conditions
+(Section~\ref{sec:evaluation}). The approach is implemented as a preprocessor
+in \cvc (Section~\ref{sec:front-end}). \begin{rep}%
 When invoked with a specific
 command-line option, the solver assumes that functions introduced using the
-\texttt{define-fun-rec} command are admissible and translates them accordingly.\end{longv}
+\texttt{define-fun-rec} command are admissible and translates them accordingly.\end{rep}
 
@@
 
-%\begin{nottinyv}
+%
 \section{Preliminaries}
@@
 all occurrences of $\vec x$ with $\vec t$ in $u$.
-\begin{longv}When applied to terms, the symbol $=$ denotes syntactic equality.\end{longv}
+\begin{rep}When applied to terms, the symbol $=$ denotes syntactic equality.\end{rep}
 
@@
 A $\Sigma$-formula $\varphi$ is \emph{$T$-satisfiable}
-if it is satisfied by some interpretation in $\Mo$\begin{longv};
-otherwise, it is \emph{$T$-unsatisfiable}\end{longv}.
+if it is satisfied by some interpretation in $\Mo$\begin{rep};
+otherwise, it is \emph{$T$-unsatisfiable}\end{rep}.
 A formula $\varphi$ \emph{$T$-entails} $\psi$, written $\varphi \models_T \psi$,
@@
 %
-where $\con{f}_1, \ldots, \con{f}_n$ are distinct\begin{longv} function symbols\end{longv},
+where $\con{f}_1, \ldots, \con{f}_n$ are distinct\begin{rep} function symbols\end{rep},
 %ct added
@@
 is \emph{closed under function expansion with respect to $\DDD$} if
-\begin{shortv}$\closurefmla$. \end{shortv}%
-\begin{longv}\[\closurefmla\]\end{longv}%
+\begin{conf}$\closurefmla$. \end{conf}%
+\begin{rep}\[\closurefmla\]\end{rep}%
 The set $\DDD$ is \emph{admissible} if for every $T$-satisfiable formula 
@@
 evaluates the condition of an $\lite$ before evaluating the relevant branch,
-ignoring the other branch. \begin{longv}Logically, such dependencies can be captured by
-congruence rules. \end{longv}Krauss developed these ideas in the more general context of
+ignoring the other branch. \begin{rep}Logically, such dependencies can be captured by
+congruence rules. \end{rep}Krauss developed these ideas in the more general context of
 higher-order logic \cite[Section 2]{krauss-2009-phd}.
@@
 If $\DDD$ is a set of well-founded function definitions for\/
-$\sfundefs{\Sigma} \begin{longv} = \{\con{f}_1,\ldots,\con{f}_n\}\end{longv}$, then it is admissible.
+$\sfundefs{\Sigma} \begin{rep} = \{\con{f}_1,\ldots,\con{f}_n\}\end{rep}$, then it is admissible.
 \end{theorem}
-\begin{longv}
+\begin{rep}
 \begin{proofsketch}
@@
 \end{proofsketch}
-\end{longv}
-
-\begin{longv}
+\end{rep}
+
+\begin{rep}
 Another useful fragment of function definitions is the class of
@@
 or productivity, without affecting admissibility.
-\end{longv}
+\end{rep}
 
@@
 satisfying $\con{f}$'s definition. A more subtle example is
-\begin{shortv}$\badassex$. \end{shortv}%
-\begin{longv}\[\badassex\]\end{longv}%
+\begin{conf}$\badassex$. \end{conf}%
+\begin{rep}\[\badassex\]\end{rep}%
 While this set has a model where $\con f$ and $\con g$ are interpreted as the
@@
 $\con{f}( 0 ) \teq 1$ is
-closed under function expansion \begin{longv}with respect to this set\end{longv}
+closed under function expansion \begin{rep}with respect to this set\end{rep}
 and yet there exists
@@
 %
-\begin{longv}\par\end{longv}
+\begin{rep}\par\end{rep}
 %
@@
 If $t$ is an unannotated quantified formula, $\conv_0$ calls itself
-on the body with the same polarity. \begin{longv}A quantifier is prefixed to the
-quantified formula and constraint returned by the recursive call.\end{longv}
+on the body with the same polarity. \begin{rep}A quantifier is prefixed to the
+quantified formula and constraint returned by the recursive call.\end{rep}
 Otherwise, $t$ is either an application of an uninterpreted predicate symbol or a term
@@
 %\begin{equation} \label{eq:approx-interp}
-\begin{shortv}$\badlambda$. \end{shortv}%
-\begin{longv}\[\badlambda\]\end{longv}%
+\begin{conf}$\badlambda$. \end{conf}%
+\begin{rep}\[\badlambda\]\end{rep}%
 %\end{equation}
@@
 
-\begin{longv}
+\begin{rep}
 %In the following,
@@
 \end{corollary}
-\end{longv}
+\end{rep}
 
@@
 \end{theorem}
-\begin{longv}
+\begin{rep}
 \begin{proofsketch}
@@
 %For clarity, we demonstrate the consequences of this theorem by means of an example.
-\end{longv}
-
-\begin{shortv}
+\end{rep}
+
+\begin{conf}
 We give an intuition of the above theorem in the context of an example.
@@
 \}$.
-\end{shortv}
+\end{conf}
 
@@
 (the first conjunct of~(\ref{eq:ex-after})) and $\absconstraints( L_0
-)$, it must also satisfy\begin{longv} the formula $\psi_1$:\end{longv}
+)$, it must also satisfy\begin{rep} the formula $\psi_1$:\end{rep}
 \[
@@
 When reading this input,
-\begin{longv}%
+\begin{rep}%
 \cvc adds the annotated quantified formula
@@
 $$\forallf{\con{s}} x.\;\, \lite\bigl( x \leq 0,\; \con{s}( x ) \teq 0,\; \con{s}( x ) \teq \con{s}( x-1 ) \bigr)$$
-\end{longv}%
-\begin{shortv}%
+\end{rep}%
+\begin{conf}%
 \cvc adds
@@
 which after rewriting becomes
-$\forallf{\con{s}} x.\;\, \lite\bigl( x \leq 0,\; \con{s}( x ) \teq 0,\; \con{s}( x ) \teq \con{s}( x-1 ) \bigr)$. \end{shortv}%
+$\forallf{\con{s}} x.\;\, \lite\bigl( x \leq 0,\; \con{s}( x ) \teq 0,\; \con{s}( x ) \teq \con{s}( x-1 ) \bigr)$. \end{conf}%
 If \cvc's finite model finding mode for recursive functions is enabled (using
@@
 an unsatisfiable problem. 
-\begin{longv}
+\begin{rep}
 Indeed, if we add the inconsistent definition
@@
 it wrongly answers \emph{sat}.
-\end{longv}
+\end{rep}
 
@@
 Some finite model finders are based on a reduction to a decidable logic,
-typically propositional logic. \begin{longv}They translate the input problem to the weaker
-logic and pass it to a solver for that logic.\end{longv}
+typically propositional logic. \begin{rep}They translate the input problem to the weaker
+logic and pass it to a solver for that logic.\end{rep}
 The translation is parameterized by upper or exact finite bounds on
@@
 in the earlier versions of
-\begin{shortv}MACE\end{shortv}\begin{longv}Mace (or MACE)\end{longv}
+\begin{conf}MACE\end{conf}\begin{rep}Mace (or MACE)\end{rep}
 \cite{mccune-1994}.
@@
 This approach was pioneered by FINDER
-\cite{slaney-1994} and SEM \cite{zhang-zhang-1995}\begin{longv} and serves as
+\cite{slaney-1994} and SEM \cite{zhang-zhang-1995}\begin{rep} and serves as
 the basis of many more model finders, notably the Alloy Analyzer's precursor
 \cite{jackson-1996} and the later versions of~Mace
-\cite{mccune-prover9-mace4}\end{longv}.
-
-\begin{longv}
+\cite{mccune-prover9-mace4}\end{rep}.
+
+\begin{rep}
 Most of the above tools cannot cope with algebraic datatypes or other infinite
-types. \end{longv}%
+types. \end{rep}%
 Kuncak and Jackson \cite{kuncak-jackson-2005} presented an idiom for
@@
 of) models for formulas in the \relax{existential--bounded-universal} fragment%
-\begin{longv} (i.e., formulas whose prenex normal forms contain no unbounded universal
-quantifiers ranging over datatypes)\end{longv}. This idiom was further developed by Dunets
+\begin{rep} (i.e., formulas whose prenex normal forms contain no unbounded universal
+quantifiers ranging over datatypes)\end{rep}. This idiom was further developed by Dunets
 et al.\ \cite{dunets-et-al-2010}, who presented a translation scheme
@@
 relies on another variant of Kuncak and Jackson's approach inside a
-Kleene-style three-valued logic\begin{longv}, inspired by abstract interpretation\end{longv}.
-\begin{longv}It was also the first tool of its kind to support corecursion and
-coalgebraic datatypes \cite{blanchette-2013-relational}.\end{longv}
+Kleene-style three-valued logic\begin{rep}, inspired by abstract interpretation\end{rep}.
+\begin{rep}It was also the first tool of its kind to support corecursion and
+coalgebraic datatypes \cite{blanchette-2013-relational}.\end{rep}
 The three-valued logic approach extends each
@@
 quantifiers altogether by unfolding recursive definitions up to a certain
-depth\begin{longv}, which is increased on a per-need basis\end{longv}.
+depth\begin{rep}, which is increased on a per-need basis\end{rep}.
 Our translation~$\conv$ works in an analogous manner, 
@@
 and quantifier instantiation is used in lieu of function unfolding.
-\begin{longv}It would be
+\begin{rep}It would be
 worth investigating how existing approaches for function
 unfolding can inform approaches for dedicated quantifier instantiation
-techniques for function definitions, and vice versa.\end{longv}
+techniques for function definitions, and vice versa.\end{rep}
 
@@
 programs. QuickCheck \cite{claessen-hughes-2000} for Haskell is an
-early example, based on random testing. Bounded exhaustive testing\begin{longv}
-\cite{runciman-et-al-2008}\end{longv} and narrowing\begin{longv}
-\cite{lindblad-2008-testing}\end{longv} are other successful strategies. These
+early example, based on random testing. Bounded exhaustive testing\begin{rep}
+\cite{runciman-et-al-2008}\end{rep} and narrowing\begin{rep}
+\cite{lindblad-2008-testing}\end{rep} are other successful strategies. These
 tools are often much faster than model finders, but they typically cannot cope
@@
 
-\begin{tinyv}
-This abstract is based on a regular submission to the SMT 2015 workshop,
-which itself is based on a longer technical report. Both are available online.%
-\footnote{\url{http://lara.epfl.ch/~reynolds/SMT2015-recfun/}}
-\end{tinyv}
-
-%\begin{nottinyv}
 {%\footnotesize
@@
 }
-%\end{nottinyv}
+%
 

### 2016-01-15 ijcar2016/body.tex
tuned

@@
+%%% BEGIN Times font
+\usepackage{mathptmx}
+\usepackage[scaled=.82]{beramono}
+\usepackage[scaled=.86]{helvet}
+\DeclareSymbolFont{letters}{OML}{txmi}{m}{it}
+%%% END Times font
+
 %\theoremstyle{plain}
@@
 (Section~\ref{sec:evaluation}). The approach is implemented as a preprocessor
-in \cvc (Section~\ref{sec:front-end}). \begin{rep}%
+in \cvc and in the Nunchaku model finder
+(Section~\ref{sec:implementations}). \begin{rep}%
 When invoked with a specific
@@
 
+\section{Implementations}
+\label{sec:implementations}
+
+The translation $\conv$ significantly improves the effectiveness of current SMT
+techniques for finding models of formulas involving recursive
+function definitions.
+We have implemented it in the development version of \cvc %~\cite{ReyEtAl-1-RR-13}
+(version~1.5 prerelease).
+
+\subsection{CVC4}
+\label{ssec:cvc4}
+
+In \cvc, function definitions $\forallf{\con{f}} \vec x.\; \varphi$ can be written using the $\definefunreccmd$ command
+from the SMT-LIB 2.5 standard~\cite{smtlib25}.
+Thus, formula~(\ref{eq:ex-before}) from Example~\ref{ex:translation}
+can be specified as %follows:
+%
+\begin{quote}
+%\small -- no small, unless you convince me that these mixtures of font sizes
+% are aesthetic and serve a purpose --jb
+\begin{verbatim}
+(define-fun-rec s ((x Int)) Int (ite (<= x 0) 0 (+ x (s (- x 1)))))
+(declare-fun c () Int)
+(assert (> (s c) 100))
+(check-sat)
+\end{verbatim}
+\end{quote}
+%
+When reading this input,
+\begin{rep}%
+\cvc adds the annotated quantified formula
+$$\forallf{\con{s}} x.\;\, \con{s}( x ) \teq \lite\bigl( x \leq 0,\; 0,\; \con{s}( x-1 )\bigr)$$
+to its list of assertions,
+which after rewriting becomes
+$$\forallf{\con{s}} x.\;\, \lite\bigl( x \leq 0,\; \con{s}( x ) \teq 0,\; \con{s}( x ) \teq \con{s}( x-1 ) \bigr)$$
+\end{rep}%
+\begin{conf}%
+\cvc adds
+$\forallf{\con{s}} x.\;\, \con{s}( x ) \teq \lite\bigl( x \leq 0,\; 0,\; \con{s}( x-1 )\bigr)$
+to its list of assertions,
+which after rewriting becomes
+$\forallf{\con{s}} x.\;\, \lite\bigl( x \leq 0,\; \con{s}( x ) \teq 0,\; \con{s}( x ) \teq \con{s}( x-1 ) \bigr)$. \end{conf}%
+If \cvc's finite model finding mode for recursive functions is enabled (using
+the command-line option \texttt{-}\texttt{-fmf-fun}), it will replace its list of known
+assertions based on the translation $\conv$ before checking for satisfiability.
+%If the functions provided in the input are admissible,
+%then a ``satisfiable" response from the solver implies that a model exists for the original formula.
+Accordingly, the solver will output the approximation of the interpretation it
+used for recursive function definitions. 
+For the example above, it outputs a %partial
+model of~$\con{s}$ where only
+the values of $\con{s}(x)$ for $x = 0,\ldots,14$ are correctly given:
+\begin{quote}
+%\small
+\begin{verbatim}
+(model
+  (define-fun s (($x1 Int)) Int 
+    (ite (= $x1 14) 105 (ite (= $x1 13) 91 (ite (= $x1 12) 78 
+      (ite (= $x1 11) 66 (ite (= $x1 10) 55 (ite (= $x1 4) 10 
+        (ite (= $x1 9) 45 (ite (= $x1 8) 36 (ite (= $x1 7) 28 
+          (ite (= $x1 6) 21 (ite (= $x1 3) 6 (ite (= $x1 5) 15 
+            (ite (= $x1 2) 3 (ite (= $x1 1) 1 0)))))))))))))))
+  (define-fun c () Int 14))
+\end{verbatim}
+\end{quote}
+
+The \texttt{-}\texttt{-fmf-fun} option tells \cvc to assume that
+functions introduced using \texttt{define-\allowbreak fun-\allowbreak rec} are admissible.
+We stress that admissibility must be discharged separately by the user---e.g., using a
+syntactic criterion or a termination prover.
+If some function definitions are not admissible, \cvc may answer \emph{sat} for
+an unsatisfiable problem. 
+\begin{rep}
+Indeed, if we add the inconsistent definition
+%
+\begin{quote}
+%\small
+\begin{verbatim}
+(define-fun-rec h ((x Int)) Int (+ (h x) x))
+\end{verbatim}
+\end{quote}
+%
+to the above problem and run \cvc with the \texttt{-}\texttt{-fmf-fun} option,
+it wrongly answers \emph{sat}.
+\end{rep}
+
+%Usually not interested in showing logical consistency, there
+%exist other tools that check e.g. termination.
+%Thus, want SMT solver to check satisfiability the assumption that recursive
+%functions are consistent (where this needs to be defined precisely).
+
+\subsection{Nunchaku}
+\label{ssec:nunchaku}
+
+  * Nunchaku
+    * intended as a higher-order model finder with a rich input language,
+      including (co)datatypes, (co)recursive functions, and
+      (co)inductive predicates
+  * first version, 0.1, came out in January 2016, without support for
+    higher-order constructs yet
+  * standalone tool; have a frontend for Isabelle and are envisioning
+    frontends for other proof assistants, notably Coq and the
+    TLA\textsuperscript{+} proof system
+
+  * flavor of input language
+    * case expressions
+
+  * architecture: pipeline of transformations, each of which consists of a
+    pair of functions $(translate, reconstruct)$:
+    * translates transforms a problem over a signature $\Sigma$ in a logic $\mathcal{L}$
+      to a problem over a signature $\Sigma'$ in a logic $\mathcal{L}'$
+    * reconstruct transforms a model back from $\Sigma'$ to $\Sigma$
+    * similar to institutions \cite{goguen-meseguer}
+
+  * table with the current pipeline
+
+  * rec transformation is just one operation in pipeline
+    * reconstruct function: transforms a total model of the encoded problem
+      into a partial model of the original model, trimming out junk and adding
+      "..." to represent missing entries
+
+  * higher-order support will be done as a natural extension of the encoding
+    presented in this paper
+    * explain a bit
+    * two o
+
+\section{Examples}
+\label{sec:examples}
+
+\subsection{ANDY?}
+
+\subsection{Lists and Lazy Lists}
+
+\subsection{AVL Trees?}
+
+\subsection{Huffman's Algorithm}
+
 \section{Evaluation}
@@
 
-\section{Front-End Support in \cvc}
-\label{sec:front-end}
-
-The translation $\conv$ significantly improves the effectiveness of current SMT
-techniques for finding models of formulas involving recursive
-function definitions.
-We have implemented it in the development version of \cvc %~\cite{ReyEtAl-1-RR-13}
-(version~1.5 prerelease).
-Function definitions $\forallf{\con{f}} \vec x.\; \varphi$ can be written using the $\definefunreccmd$ command
-from the new version (2.5) of the SMT-LIB standard~\cite{smtlib25}.
-Thus, formula~(\ref{eq:ex-before}) from Example~\ref{ex:translation}
-can be specified as %follows:
-%
-\begin{quote}
-%\small -- no small, unless you convince me that these mixtures of font sizes
-% are aesthetic and serve a purpose --jb
-\begin{verbatim}
-(define-fun-rec s ((x Int)) Int (ite (<= x 0) 0 (+ x (s (- x 1)))))
-(declare-fun c () Int)
-(assert (> (s c) 100))
-(check-sat)
-\end{verbatim}
-\end{quote}
-%
-When reading this input,
-\begin{rep}%
-\cvc adds the annotated quantified formula
-$$\forallf{\con{s}} x.\;\, \con{s}( x ) \teq \lite\bigl( x \leq 0,\; 0,\; \con{s}( x-1 )\bigr)$$
-to its list of assertions,
-which after rewriting becomes
-$$\forallf{\con{s}} x.\;\, \lite\bigl( x \leq 0,\; \con{s}( x ) \teq 0,\; \con{s}( x ) \teq \con{s}( x-1 ) \bigr)$$
-\end{rep}%
-\begin{conf}%
-\cvc adds
-$\forallf{\con{s}} x.\;\, \con{s}( x ) \teq \lite\bigl( x \leq 0,\; 0,\; \con{s}( x-1 )\bigr)$
-to its list of assertions,
-which after rewriting becomes
-$\forallf{\con{s}} x.\;\, \lite\bigl( x \leq 0,\; \con{s}( x ) \teq 0,\; \con{s}( x ) \teq \con{s}( x-1 ) \bigr)$. \end{conf}%
-If \cvc's finite model finding mode for recursive functions is enabled (using
-the command-line option \texttt{-}\texttt{-fmf-fun}), it will replace its list of known
-assertions based on the translation $\conv$ before checking for satisfiability.
-%If the functions provided in the input are admissible,
-%then a ``satisfiable" response from the solver implies that a model exists for the original formula.
-Accordingly, the solver will output the approximation of the interpretation it
-used for recursive function definitions. 
-For the example above, it outputs a partial model of~$\con{s}$ where only
-the values of $\con{s}(x)$ for $x = 0,\ldots,14$ are correctly given:
-\begin{quote}
-%\small
-\begin{verbatim}
-(model
-  (define-fun s (($x1 Int)) Int 
-    (ite (= $x1 14) 105 (ite (= $x1 13) 91 (ite (= $x1 12) 78 
-      (ite (= $x1 11) 66 (ite (= $x1 10) 55 (ite (= $x1 4) 10 
-        (ite (= $x1 9) 45 (ite (= $x1 8) 36 (ite (= $x1 7) 28 
-          (ite (= $x1 6) 21 (ite (= $x1 3) 6 (ite (= $x1 5) 15 
-            (ite (= $x1 2) 3 (ite (= $x1 1) 1 0)))))))))))))))
-  (define-fun c () Int 14))
-\end{verbatim}
-\end{quote}
-
-The \texttt{-}\texttt{-fmf-fun} option tells \cvc to assume that
-functions introduced using \texttt{define-\allowbreak fun-\allowbreak rec} are admissible.
-We stress that admissibility must be discharged separately by the user---e.g., using a
-syntactic criterion or a termination prover.
-If some function definitions are not admissible, \cvc may answer \emph{sat} for
-an unsatisfiable problem. 
-\begin{rep}
-Indeed, if we add the inconsistent definition
-%
-\begin{quote}
-%\small
-\begin{verbatim}
-(define-fun-rec h ((x Int)) Int (+ (h x) x))
-\end{verbatim}
-\end{quote}
-%
-to the above problem and run \cvc with the \texttt{-}\texttt{-fmf-fun} option,
-it wrongly answers \emph{sat}.
-\end{rep}
-
-%Usually not interested in showing logical consistency, there
-%exist other tools that check e.g. termination.
-%Thus, want SMT solver to check satisfiability the assumption that recursive
-%functions are consistent (where this needs to be defined precisely).
-
 \section{Related Work}

### 2016-01-15 ijcar2016/body.tex
more

@@
 \and
-Inria Nancy -- Grand-Est, Villers-l\`es-Nancy, France
+Inria Nancy -- Grand Est \& LORIA, Villers-l\`es-Nancy, France
 \and
@@
 the performance of existing solvers on benchmarks from two sources. 
-The translation is implemented as a preprocessor in the solver CVC4.
+The translation is implemented as a preprocessor in the CVC4 solver
+and in a new higher-order model finder called Nunchaku.
 \end{abstract}
@@
 satisfying $\con{f}$'s definition. A more subtle example is
-\begin{conf}$\badassex$. \end{conf}%
-\begin{rep}\[\badassex\]\end{rep}%
+\[\badassex\]
 While this set has a model where $\con f$ and $\con g$ are interpreted as the
@@
     TLA\textsuperscript{+} proof system
+  * displace Nitpick, based on Kodkod
+    * CVC4 has built-in (co)datatypes and efficient arithmetic
+    * also, Nitpick is intimately tied to Isabelle
 

### 2016-01-20 ijcar2016/body.tex
more

@@
 \end{rep}
+
+TODO: Say something about admissibility for tail-recursion: Watch out if the
+argument always changes...
 
@@
 
+\textbf{TODO:} Mention monotonicity analysis and zero sorts.
+
 %Usually not interested in showing logical consistency, there
@@
 
+  * advantage over Nitpick:
+    * generates a single problem, not parameterized by cardinalities
+      * pushes all the work to the SMT solver
+        * we are just starting with that -- and see lots of potential for
+          optimizations and tuning
+
   * flavor of input language
@@
 
-\section{Examples}
-\label{sec:examples}
-
-\subsection{ANDY?}
-
-\subsection{Lists and Lazy Lists}
-
-\subsection{AVL Trees?}
-
-\subsection{Huffman's Algorithm}
+\section{Case Studies}
+\label{sec:case-studies}
+
+\subsection{Lazy Lists}
+
+XXX
+
+\subsection{Huffman Coding}
+
+XXX
 

### 2016-01-21 ijcar2016/body.tex
start writing section about Nunchaku

@@
 \DeclareSymbolFont{letters}{OML}{txmi}{m}{it}
+
 %%% END Times font
@@
 
-\newcommand\closurefmla{\psi \models_T \bigwedge\nolimits_{\,i=1}^{n} \{ \varphi_i[ \vec t / \vec x ] \mid \con{f}_i( \vec t ) \in \terms( \psi ) \}}
+\newcommand\closurefmla{
+  \ensuremath{\psi \models_T \bigwedge\nolimits_{\,i=1}^{n} \{ \varphi_i[ \vec t / \vec x ] \mid \con{f}_i( \vec t ) \in \terms( \psi ) \}} }
 
@@
 
-  * Nunchaku
-    * intended as a higher-order model finder with a rich input language,
-      including (co)datatypes, (co)recursive functions, and
-      (co)inductive predicates
-  * first version, 0.1, came out in January 2016, without support for
-    higher-order constructs yet
-  * standalone tool; have a frontend for Isabelle and are envisioning
-    frontends for other proof assistants, notably Coq and the
-    TLA\textsuperscript{+} proof system
-  * displace Nitpick, based on Kodkod
-    * CVC4 has built-in (co)datatypes and efficient arithmetic
-    * also, Nitpick is intimately tied to Isabelle
-
-  * advantage over Nitpick:
-    * generates a single problem, not parameterized by cardinalities
-      * pushes all the work to the SMT solver
-        * we are just starting with that -- and see lots of potential for
-          optimizations and tuning
-
-  * flavor of input language
-    * case expressions
-
-  * architecture: pipeline of transformations, each of which consists of a
-    pair of functions $(translate, reconstruct)$:
-    * translates transforms a problem over a signature $\Sigma$ in a logic $\mathcal{L}$
-      to a problem over a signature $\Sigma'$ in a logic $\mathcal{L}'$
-    * reconstruct transforms a model back from $\Sigma'$ to $\Sigma$
-    * similar to institutions \cite{goguen-meseguer}
-
-  * table with the current pipeline
-
-  * rec transformation is just one operation in pipeline
-    * reconstruct function: transforms a total model of the encoded problem
-      into a partial model of the original model, trimming out junk and adding
-      "..." to represent missing entries
-
-  * higher-order support will be done as a natural extension of the encoding
-    presented in this paper
-    * explain a bit
-    * two o
+Nunchaku is a new model finder, intended
+to be used for finding counter-examples in several proof assistants. The
+first version, 0.1, was released in January 2016 with support for (co)datatypes,
+(co)recursive functions and (co)inductive predicates.
+Support for higher-order
+functions is planned in the next release.
+We intend for Nunchaku to be supported at least in Isabelle (where basic
+support for it already exists), Coq, and the TLA\textsuperscript{+} proof
+system.
+
+Nunchaku is the spiritual successor to Nitpick~\cite{blanchette-nipkow-2010},
+but is developed as a standalone OCaml program (with its own input language) so
+as to be usable independently from Isabelle.
+Whereas Nitpick relies on propositional checking with Kodkod,
+generating a succession of problems where cardinalities of finite types
+grow at each step, Nunchaku
+translates its input to one first-order logic formula that targets
+the finite-model finding fragment of CVC4, including the (co)datatypes support.
+Using CVC4 will also allow Nunchaku to provide efficient arithmetic and makes
+equality reasoning straightforward.
+
+The following problem is expressed in Nunchaku's input syntax; it starts with
+a datatype declaration, then two mutually recursive coinductive predicates
+are defined, and finally a goal (``find a non-zero natural number $m$ that
+is even'').
+\begin{quote}
+\begin{verbatim}
+data nat := zero | Suc nat.
+
+copred even : nat -> prop :=
+  even zero;
+  forall (n : nat). odd n => even (Suc n)
+and odd : nat -> prop :=
+  forall (n : nat). even n => odd (Suc n).
+
+goal exists m. (even m && ~ (m = zero)).
+\end{verbatim}
+\end{quote}
+
+Nunchaku parses and types the input problem before applying a sequence
+of translations, each reducing the distance to the target fragment. Here,
+for instance, the coinductive predicates
+\verb!even! and \verb!odd! are \emph{polarized} (specialized into
+\verb!even+!, used only in positive positions, and \verb!even-! used in negative
+positions),
+$m$ is Skolemized into a fresh constants,
+then then translated into admissible functions
+before another pass uses the encoding from Section~\ref{sec:encoding}.
+If a model is found, it is translated back to the input language.
+Conceptually, the sequence of transformation is a \emph{two ways pipeline}
+built by composing pairs \textsf{(encode,decode)}.
+For each such pair, \textsf{encode}
+transforms a problem over a signature $\Sigma$ in a logic
+$\mathcal{L}$ to a problem over a signature $\Sigma'$ in a slightly different
+logic $\mathcal{L}'$, and \textsf{decode} translates a model of \mathcal{L'}
+over $\Sigma'$ into a model of $\mathcal{L}$ over $\Sigma$. The table
+from Figure~\ref{fig:pipeline} presents the current pipeline implemented in Nunchaku.
+
+\begin{figure}[tb]
+%{{{
+\begin{tabular}{|l|p{0.6\textwidth}|}
+  \hline
+  component & description \\
+  \hline
+  \hline
+  Type inference
+    & Infer types and checks definitions \\
+  \hline
+  Type skolemization
+    & Replaces $\exists \alpha. p[\alpha]$
+      with $p[\tau]$ where $\tau$ is a fresh type \\
+  \hline
+  Monomorphization
+    & specializes polymorphic functions on their type arguments \\
+  \hline
+  Elimination of Equations
+    & Translates multiple-equations definitions of functions into
+      a single nested pattern matching \\
+  \hline
+  Polarization
+    & specialize some predicates into a version used in positive position
+      and one used in negative position \\
+  \hline
+  Unrolling
+    & Add a decreasing argument to non-well-founded predicates \\
+  \hline
+  Skolemization
+    & Introduce Skolem symbols for term variables \\
+  \hline
+  Elimination of Predicates
+    & Changes a multi-clauses (co)inductive predicate definition
+      into a single recursive function \\
+  \hline
+  Recursion Elimination
+    & The encoding from Section~\ref{sec:encoding} \\
+  \hline
+  Elimination of Match
+    & Transforms pattern-matching expressions into tests and selectors \\
+  \hline
+  CVC4 Invocation 
+    & Call CVC4 to obtain a model \\
+  \hline
+\end{tabular}
+%}}}
+\caption{Pipeline of Transformations in Nunchaku}
+\label{fig:pipeline}
+\end{figure}
+
+TODO: similar to institutions \cite{goguen-meseguer} (?)
+
+\subsection{Decoding of Recursive Functions}
+
+Nunchaku is designed for finding counter-models to formulas a user thinks
+are theorems. Decoding models from CVC4 to hide any trace of the successive
+encodings is therefore vital for readability. Each step
+of the pipeline only has to concern itself with removing its changes to
+the signature and logic; our encoding of recursive functions is no exceptions.
+
+TODO explain how:
+- domains of proj. functions are computed
+- domains of original function is deduced
+- decision tree for original function is built
+- note how junk is removed in step 3 (but undefined "..." as last case)
+
+\subsection{Support for Higher-Order Functions}
+
+TODO:
+- currying with explicit application symbol --> function arguments are now
+  just constants
+- apply encoding on application symbol
+- gotcha: one function can be used with several arities.
+
+- optim: specialization (similar to monomorphization)
 

### 2016-01-21 ijcar2016/body.tex
use a description rather than a table

@@
 logic $\mathcal{L}'$, and \textsf{decode} translates a model of \mathcal{L'}
-over $\Sigma'$ into a model of $\mathcal{L}$ over $\Sigma$. The table
-from Figure~\ref{fig:pipeline} presents the current pipeline implemented in Nunchaku.
-
-\begin{figure}[tb]
+over $\Sigma'$ into a model of $\mathcal{L}$ over $\Sigma$. The pipeline
+currently implemented in Nunchaku is described below.
+
+\begin{description}
 %{{{
-\begin{tabular}{|l|p{0.6\textwidth}|}
-  \hline
-  component & description \\
-  \hline
-  \hline
-  Type inference
-    & Infer types and checks definitions \\
-  \hline
-  Type skolemization
-    & Replaces $\exists \alpha. p[\alpha]$
-      with $p[\tau]$ where $\tau$ is a fresh type \\
-  \hline
-  Monomorphization
-    & specializes polymorphic functions on their type arguments \\
-  \hline
-  Elimination of Equations
-    & Translates multiple-equations definitions of functions into
-      a single nested pattern matching \\
-  \hline
-  Polarization
-    & specialize some predicates into a version used in positive position
-      and one used in negative position \\
-  \hline
-  Unrolling
-    & Add a decreasing argument to non-well-founded predicates \\
-  \hline
-  Skolemization
-    & Introduce Skolem symbols for term variables \\
-  \hline
-  Elimination of Predicates
-    & Changes a multi-clauses (co)inductive predicate definition
-      into a single recursive function \\
-  \hline
-  Recursion Elimination
-    & The encoding from Section~\ref{sec:encoding} \\
-  \hline
-  Elimination of Match
-    & Transforms pattern-matching expressions into tests and selectors \\
-  \hline
-  CVC4 Invocation 
-    & Call CVC4 to obtain a model \\
-  \hline
-\end{tabular}
+  \item[Type inference] infers types and checks definitions;
+  \item[Type skolemization]
+    replaces $\exists \alpha. p[\alpha]$
+      with $p[\tau]$ where $\tau$ is a fresh type;
+  \item[Monomorphization]
+    specializes polymorphic functions on their type arguments;
+  \item[Elimination of Equations]
+    translates multiple-equations definitions of functions into
+      a single nested pattern matching;
+  \item[Polarization]
+    specializes some predicates into a version used in positive position
+      and one used in negative position;
+  \item[Unrolling]
+    adds a decreasing argument to non-well-founded predicates;
+  \item[Skolemization]
+    introduces Skolem symbols for term variables;
+  \item[Elimination of Predicates]
+    changes a multi-clauses (co)inductive predicate definition
+      into a single recursive function;
+  \item[Recursion Elimination]
+    performs the encoding from Section~\ref{sec:encoding};
+  \item[Elimination of Match]
+    transforms pattern-matching expressions into tests and selectors;
+  \item[CVC4 Invocation] calls CVC4 to obtain a model.
 %}}}
-\caption{Pipeline of Transformations in Nunchaku}
-\label{fig:pipeline}
-\end{figure}
+\end{description}
 

### 2016-01-21 ijcar2016/body.tex
compile

@@
 $\mathcal{L}$ to a problem over a signature $\Sigma'$ in a slightly different
-logic $\mathcal{L}'$, and \textsf{decode} translates a model of \mathcal{L'}
+logic $\mathcal{L}'$, and \textsf{decode} translates a model of $\mathcal{L'}$
 over $\Sigma'$ into a model of $\mathcal{L}$ over $\Sigma$. The pipeline

### 2016-01-24 ijcar2016/body.tex
Start revising evaluation section.

@@
 
-\newcommand\cvcd{{\cvc}d\xspace}
+\newcommand\cvcd{{\cvc}e\xspace}
+\newcommand\cvci{{\cvc}ei\xspace}
 \newcommand\cvcf{{\cvc}f\xspace}
-\newcommand\cvci{{\cvc}i\xspace}
+\newcommand\cvcfe{{\cvc}fe\xspace}
+\newcommand\cvcfm{{\cvc}fm\xspace}
 \newcommand\ziiib{\ziii}
@@
 
-\newcommand\isa{Isa\xspace}
-\newcommand\isam{Isa-Mut\xspace}
+\newcommand\isanun{Isa\xspace}
+\newcommand\isa{Ip\xspace}
+\newcommand\isam{Ip-Mut\xspace}
 \newcommand\leon{Leon\xspace}
@@
 %
-We gathered 245 benchmarks from two sources, which we will refer to as
-\isa and \leon. The first source consists of the 79
+We gathered 602 benchmarks from three sources, which we will refer to as
+\isanun, \isa and \leon. 
+The first source consists of 357 benchmarks coming from\ajr{where Isa-Nun benchmarks come from, what they represent, are they expected to be unsat/sat}.
+The second source consists of the 79
 benchmarks from the IsaPlanner
@@
 with a signature that contains uninterpreted function symbols over these datatypes. 
-The second source consists of 166
+The third source consists of 166
 benchmarks from the Leon repository,\footnote{%Available at
@@
 with respect to a set of well-founded functions.
-%
-% in particular they each
-%are a conjoined list of function definitions followed by a (negated)
-%conjecture.
-Nearly all of them are unsatisfiable, except for a
-handful in the \leon set.
-
+A majority of the benchmarks in the latter two sets are unsatisfiable.
 }
 
-For each of these 245 benchmarks, we considered up to three randomly selected
+For each of the 245 benchmarks in the \isa and \leon sets, 
+we considered up to three randomly selected
 mutated forms of its conjecture $\psi$. In particular, we considered unique
@@
 call these sets \isam and \leonm, respectively.
-Thus, our benchmark set consist of 885 benchmarks (640 mutants plus the
-original 245). We considered these 885 benchmarks both before and after the
-translation~$\conv$. % as described in Section~\ref{sec:encoding}.
+Thus, our benchmark set consists of 1242 benchmarks (640 mutants plus the
+original 602). 
+Each of the 1242 benchmarks were generated in SMT lib version 2 format~\cite{}.
+For \isanun, \ajr{describe generation of smt2 benchmarks without encoding}
+The others were translated to SMT lib version 2 by the Leon tool.
+
+We additionally considered these 1242 benchmarks both before and after the translation~$\conv$.
 For example, \isa contains 79 original benchmarks $\varphi$ and 79 translated
 benchmarks $\conv(\varphi)$.
+For the \isanun, \ajr{describe generation of smt2 benchmarks with encoding}.
+All others were translated by \cvc 
+by outputting the result of running the preprocessor from Section~\ref{ssec:cvc4} on each benchmark.
 
@@
 as well as methods for finding models for quantified formulas~\cite{GeDeM-CAV-09}.
-We also considered three configurations of \cvc~\cite{barrett-et-al-2011} which
-we will refer to as \cvcd, \cvcf, and \cvci.
-The default configuration \cvcd runs heuristic and conflict-based techniques for quantifier instantiation~\cite{ReynoldsTinelliMoura14},
+We also considered various configurations of \cvc~\cite{barrett-et-al-2011},
+which we will refer to as \cvcd, \cvci, \cvcf, \cvcfe, and \cvcfm.
+The default configuration \cvcd runs heuristic 
+and conflict-based techniques for quantifier instantiation~\cite{ReynoldsTinelliMoura14},
 but does not include techniques for finding models.
-\cvcf runs heuristic instantiation and the finite model
+The configuration \cvci additionally incorporates techniques for automating inductive
+reasoning in SMT~\cite{reynolds-kuncak-2015}.
+All other configurations run the finite model
 finding procedure due to Reynolds et al.\ \cite{ReyEtAl-1-RR-13,reynolds-et-al-2013}.
-\cvci incorporates techniques for automating inductive
-reasoning in SMT~\cite{reynolds-kuncak-2015}.
+The configuration \cvcfe additionally incorporates heuristic quantifier instantiation,
+and \cvcfm incorporates the fairness scheme for monotonic sorts as described in Section~\ref{ssec:cvc4}.
 
@@
 \centering
-\begin{tabular}{l@{\kern1.5em}r@{\kern0.375em}r@{\kern1.5em}r@{\kern0.375em}r@{\kern1.5em}r@{\kern0.375em}r@{\kern1.5em}r@{\kern0.375em}r@{\,\,}}
-  & \multicolumn{2}{c@{\kern1.5em}}{\phantom{0}\ziiib}     & \multicolumn{2}{c@{\kern1.5em}}{\phantom{0}\cvcd}     & \multicolumn{2}{c@{\kern1.5em}}{\phantom{0}\cvcf}     & \multicolumn{2}{c@{\,\,}}{\phantom{0}\cvci}
+\begin{tabular}{l@{\kern1.5em}r@{\kern0.375em}r@{\kern1.5em}r@{\kern0.375em}r@{\kern1.5em}r@{\kern0.375em}r@{\kern1.5em}r@{\kern0.375em}r@{\kern1.5em}r@{\kern0.375em}r@{\kern1.5em}r@{\kern0.375em}r@{\,\,}}
+  & \multicolumn{2}{c@{\kern1.5em}}{\phantom{0}\ziiib}     & \multicolumn{2}{c@{\kern1.5em}}{\phantom{0}\cvcd}     
+  & \multicolumn{2}{c@{\kern1.5em}}{\phantom{0}\cvcf}      & \multicolumn{2}{c@{\kern1.5em}}{\cvcfe} 
+  & \multicolumn{2}{c@{\kern1.5em}}{\cvcfm}     & \multicolumn{2}{c@{\,\,}}{\cvci}
 \\%[-1pt]
+  & \hfill $\varphi$ \hfill & $\conv(\varphi)$\!\!
+  & \hfill $\varphi$ \hfill & $\conv(\varphi)$\!\!
   & \hfill $\varphi$ \hfill & $\conv(\varphi)$\!\!
@@
 \midrule
-\isa & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0
+
+\isanun & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0
 \\
-\leon  & 0 & 2 & 0 & 0 & 0 & {\win 9} & 0 & 0
+\isa & 0 & 43 & 0 & 0 & 0 & {\win 161} & 0 & {\win 161} & 0 & {\win 161} & 0 & 0
 \\
-\isam & 0 & 35 & 0 & 0 & 0 & {\win 153} & 0 & 0
+\leon  & 8 & 30 & 6 & 6 & 6 & {\win 54} & 6 & 53 & 6 & {\win 54} & 6 & 6
 \\
-\leonm  & 11  & 75 & 6 & 6 & 6 & {\win 169} & 6 & 6
+\isam & 6 & 61 & 0 & 0 & 2 & 180 & 2 & {\win 181} & 2 & {\win 181} & 0 & 0
+\\
+\leonm  & 0 & 14 & 0 & 0 & 0 & 155 & 0 & {\win 156} & 0 & 155 & 0 & 0
 \\[\jot]
-Total & \phantom{0}11  & 112 & \phantom{00}6 & \phantom{00}6 & \phantom{00}6 & {\win 331} & \phantom{00}6 & \phantom{00}6
+Total & 14 & 148 & 6 & 6 & 8 & 550 & 8 & {\win 551} & 8 & {\win 551} & 6 & 6
 \end{tabular}
@@
 \midrule
-\isa & 14  & 15 & 15  & 15 & 15  & 15 & {\win 61}  & 14
+\isanun & {\win 41} & 40 & 28 & 34 & 26 & 39 & 26 & 39 & 26 & 39 & 38 & 31
 \\
-\leon  & 73  & 78 & 80  & 80 & 80  & 76 & {\win 96}  & 78
+\isanun & 14 & 15 & 15 & 15 & 1 & 15 & 15 & 15 & 1 & 15 & {\win 64} & 14
 \\
-\isam & 17  & 18 & 18  & 18 & 18  & 18 & {\win 44}  & 17
+\leon & 54 & 55 & 56 & 56 & 13 & 55 & 56 & 55 & 13 & 55 & {\win 93} & 53
 \\
-\leonm  & 83  & 98 & 103 & 98 & 104 & 95 & {\win 117} & 98
+\isam & 48 & 52 & 58 & 52 & 12 & 53 & {\win 58} & 50 & 12 & 53 & {\win 58} & 50
+\\
+\leonm & 94 & 107 & 106 & 107 & 39 & 108 & 106 & 108 & 39 & 108 & {\win 123} & 106
 \\[\jot]
-Total & 187 & 209 & 216 & 211 & 217 & 204 & {\win 318} & 217
+Total & 251 & 269 & 263 & 264 & 91 & 270 & 261 & 267 & 91 & 270 & {\win 376} & 254
 \end{tabular}
@@
 \end{figure}
+
+\ajr{To revise}
 
@@
 Viktor Kuncak and Stephan Merz have made this work possible. We would also like
-to thank Damien Busato-Gaston and Emmanouil Koukoutos for providing the initial
-set of benchmarks used in the evaluation, and the anonymous reviewers for their
+to thank Damien Busato-Gaston and Emmanouil Koukoutos for providing the
+set of Leon benchmarks used in the evaluation, and the anonymous reviewers for their
 suggestions and comments.

### 2016-01-24 ijcar2016/body.tex
Minor to evaluation, incomplete.

@@
 
-\isanun & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0
+\isanun  & 0 & 14 & 0 & 0 & 0 & 155 & 0 & {\win 156} & 0 & 155 & 0 & 0
 \\
-\isa & 0 & 43 & 0 & 0 & 0 & {\win 161} & 0 & {\win 161} & 0 & {\win 161} & 0 & 0
+\isa & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0
+\\
+\isam & 0 & 43 & 0 & 0 & 0 & {\win 161} & 0 & {\win 161} & 0 & {\win 161} & 0 & 0
 \\
@@
 \\
-\isam & 6 & 61 & 0 & 0 & 2 & 180 & 2 & {\win 181} & 2 & {\win 181} & 0 & 0
-\\
-\leonm  & 0 & 14 & 0 & 0 & 0 & 155 & 0 & {\win 156} & 0 & 155 & 0 & 0
+\leonm & 6 & 61 & 0 & 0 & 2 & 180 & 2 & {\win 181} & 2 & {\win 181} & 0 & 0
 \\[\jot]
@@
 The benchmarks and more detailed results are available online.%
-\footnote{\url{http://lara.epfl.ch/~reynolds/SMT2015-recfun/}}
+\footnote{\url{http://lara.epfl.ch/~reynolds/IJCAR2016-recfun/}}
 The figures are divided into benchmarks triggering \emph{unsat} and \emph{sat}
@@
 The translation $\conv$ (the ``$\conv(\varphi)$'' columns) has a major
-impact. \cvcf finds 331 of the 885 benchmarks to be satisfiable,
-including 9~benchmarks in the nonmutated \leon benchmark set. The
-performance of \ziiib for countermodels also improves dramatically, as it
-finds 101 more benchmarks to be satisfiable, including 10 that are not solved by \cvcf.
+impact. \cvcf finds 550 of the 1242 benchmarks to be satisfiable,
+including 6~benchmarks in the nonmutated \leon benchmark set. 
+The two optimizations for finite model finding in \cvc (configurations \cvcfe and \cvcfm) 
+led to a net gain of 1 satisfiable benchmark each.
+The performance of \ziiib for countermodels also improved dramatically, as it
+finds 134 more benchmarks to be satisfiable, including 5 that are not solved by \cvcf.
 %
@@
 Over \emph{unsat} benchmarks, \cvci is the
-clear winner, finding 318 total benchmarks to be unsatisfiable. Moreover, the
-translation $\conv$ helps \ziii for \emph{unsat} responses as well:
-\ziii solves a total of 209 with the translation, whereas it solves only 187 without it.
-%
-In contrast, \cvcd finds 211 benchmarks unsatisfiable with the translation.
-In general, the translation $\conv$ does not significantly degrade performance
-over unsatisfiable benchmarks, and in some cases it actually aids the solver 
-in determining unsatisfiability. The exception is \cvci, whose inductive
-arguments are guided by the types of the quantified variables.
+clear winner, finding 376 total benchmarks to be unsatisfiable. 
+Apart from this configuration, 
+the translation $\conv$ helps all configurations for \emph{unsat} responses as well.
+In particular,
+\ziii solves a total of 269 with the translation, whereas it solves only 251 without it.
+The configuration \cvcf improved even more significantly, 
+solving 270 unsat benchmarks overall, which is more than both \cvcd and \ziii.
+
+\ajr{Importance of the translation helping in practice for Isabelle/Nunchaku, if any}
 

### 2016-01-24 ijcar2016/body.tex
Update beginning of implementation section

@@
 
-The translation $\conv$ significantly improves the effectiveness of current SMT
-techniques for finding models of formulas involving recursive
-function definitions.
-We have implemented it in the development version of \cvc %~\cite{ReyEtAl-1-RR-13}
-(version~1.5 prerelease).
+We have implemented the translation $\conv$ in two separate systems, 
+as a preprocessor in the development version of \cvc %~\cite{ReyEtAl-1-RR-13} (version~1.5 prerelease),
+and in the higher-order model finder Nunchaku.
+This section describes how the translation is implemented in both these systems,
+as well as optimizations used by \cvc for finding models for the translated benchmarks.
+
+%The translation $\conv$ significantly improves the effectiveness of current SMT
+%techniques for finding models of formulas involving recursive
+%function definitions.
+%We have implemented it in the development version of \cvc %~\cite{ReyEtAl-1-RR-13}
+%(version~1.5 prerelease).
 
@@
 from the SMT-LIB 2.5 standard~\cite{smtlib25}.
-Thus, formula~(\ref{eq:ex-before}) from Example~\ref{ex:translation}
+Formula~(\ref{eq:ex-before}) from Example~\ref{ex:translation}
 can be specified as %follows:
@@
 
+\par
+
 \textbf{TODO:} Mention monotonicity analysis and zero sorts.
@@
 original 602). 
-Each of the 1242 benchmarks were generated in SMT lib version 2 format~\cite{}.
+Each of the 1242 benchmarks were generated in SMT lib version 2.5 format~\cite{smtlib25}. %ajr : although these don't require 2.5 features
 For \isanun, \ajr{describe generation of smt2 benchmarks without encoding}
@@
 as well as methods for finding models for quantified formulas~\cite{GeDeM-CAV-09}.
-We also considered various configurations of \cvc~\cite{barrett-et-al-2011},
+We also considered five configurations of \cvc~\cite{barrett-et-al-2011},
 which we will refer to as \cvcd, \cvci, \cvcf, \cvcfe, and \cvcfm.
@@
 finding procedure due to Reynolds et al.\ \cite{ReyEtAl-1-RR-13,reynolds-et-al-2013}.
-The configuration \cvcfe additionally incorporates heuristic quantifier instantiation,
+The configuration \cvcfe additionally incorporates heuristic quantifier instantiation as described in Section 2.3 of~\cite{reynolds-et-al-2013},
 and \cvcfm incorporates the fairness scheme for monotonic sorts as described in Section~\ref{ssec:cvc4}.

### 2016-01-24 ijcar2016/body.tex
Minor

@@
 
-\par
-
 \textbf{TODO:} Mention monotonicity analysis and zero sorts.
@@
 The two optimizations for finite model finding in \cvc (configurations \cvcfe and \cvcfm) 
-led to a net gain of 1 satisfiable benchmark each.
+led to a net gain of 1 satisfiable benchmark each with respect to \cvcf.
 The performance of \ziiib for countermodels also improved dramatically, as it
@@
 whose definitions are admissible.
-%Both \ziiib and \cvcf were able to find a
-%large number of countermodels after the translation, with the latter
-%configuration finding significantly more instances (331 to 112).
 
@@
 solving 270 unsat benchmarks overall, which is more than both \cvcd and \ziii.
-
-\ajr{Importance of the translation helping in practice for Isabelle/Nunchaku, if any}
+\ajr{Importance of the translation helping unsat in practice for Isabelle/Nunchaku, if any}
 

### 2016-01-25 ijcar2016/body.tex
Remove induction, minor.

@@
 
-\newcommand\cvcd{{\cvc}e\xspace}
-\newcommand\cvci{{\cvc}ei\xspace}
+\newcommand\cvcd{{\cvc}h\xspace}
 \newcommand\cvcf{{\cvc}f\xspace}
-\newcommand\cvcfe{{\cvc}fe\xspace}
+\newcommand\cvcfe{{\cvc}fh\xspace}
 \newcommand\cvcfm{{\cvc}fm\xspace}
@@
 We have implemented the translation $\conv$ in two separate systems, 
-as a preprocessor in the development version of \cvc %~\cite{ReyEtAl-1-RR-13} (version~1.5 prerelease),
+as a preprocessor in the development version of \cvc (version~1.5 prerelease),
 and in the higher-order model finder Nunchaku.
@@
 We also considered five configurations of \cvc~\cite{barrett-et-al-2011},
-which we will refer to as \cvcd, \cvci, \cvcf, \cvcfe, and \cvcfm.
-The default configuration \cvcd runs heuristic 
+which we will refer to as \cvcd, \cvcf, \cvcfe, and \cvcfm.
+The configuration \cvcd runs heuristic 
 and conflict-based techniques for quantifier instantiation~\cite{ReynoldsTinelliMoura14},
 but does not include techniques for finding models.
-The configuration \cvci additionally incorporates techniques for automating inductive
-reasoning in SMT~\cite{reynolds-kuncak-2015}.
-All other configurations run the finite model
+The other configurations run the finite model
 finding procedure due to Reynolds et al.\ \cite{ReyEtAl-1-RR-13,reynolds-et-al-2013}.
@@
 \centering
-\begin{tabular}{l@{\kern1.5em}r@{\kern0.375em}r@{\kern1.5em}r@{\kern0.375em}r@{\kern1.5em}r@{\kern0.375em}r@{\kern1.5em}r@{\kern0.375em}r@{\kern1.5em}r@{\kern0.375em}r@{\kern1.5em}r@{\kern0.375em}r@{\,\,}}
+\begin{tabular}{l@{\kern1.5em}r@{\kern0.375em}r@{\kern1.5em}r@{\kern0.375em}r@{\kern1.5em}r@{\kern0.375em}r@{\kern1.5em}r@{\kern0.375em}r@{\kern1.5em}r@{\kern0.375em}r@{\,\,}}
   & \multicolumn{2}{c@{\kern1.5em}}{\phantom{0}\ziiib}     & \multicolumn{2}{c@{\kern1.5em}}{\phantom{0}\cvcd}     
   & \multicolumn{2}{c@{\kern1.5em}}{\phantom{0}\cvcf}      & \multicolumn{2}{c@{\kern1.5em}}{\cvcfe} 
-  & \multicolumn{2}{c@{\kern1.5em}}{\cvcfm}     & \multicolumn{2}{c@{\,\,}}{\cvci}
+  & \multicolumn{2}{c@{\,\,}}{\cvcfm}
 \\%[-1pt]
-  & \hfill $\varphi$ \hfill & $\conv(\varphi)$\!\!
   & \hfill $\varphi$ \hfill & $\conv(\varphi)$\!\!
@@
 
-\isanun  & 0 & 14 & 0 & 0 & 0 & 155 & 0 & {\win 156} & 0 & 155 & 0 & 0
+\isanun  & 0 & 14 & 0 & 0 & 0 & 155 & 0 & {\win 156} & 0 & 155
 \\
-\isa & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0
+\isa & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0
 \\
-\isam & 0 & 43 & 0 & 0 & 0 & {\win 161} & 0 & {\win 161} & 0 & {\win 161} & 0 & 0
+\isam & 0 & 43 & 0 & 0 & 0 & {\win 161} & 0 & {\win 161} & 0 & {\win 161}
 \\
-\leon  & 8 & 30 & 6 & 6 & 6 & {\win 54} & 6 & 53 & 6 & {\win 54} & 6 & 6
+\leon  & 8 & 30 & 6 & 6 & 6 & {\win 54} & 6 & 53 & 6 & {\win 54}
 \\
-\leonm & 6 & 61 & 0 & 0 & 2 & 180 & 2 & {\win 181} & 2 & {\win 181} & 0 & 0
+\leonm & 6 & 61 & 0 & 0 & 2 & 180 & 2 & {\win 181} & 2 & {\win 181}
 \\[\jot]
-Total & 14 & 148 & 6 & 6 & 8 & 550 & 8 & {\win 551} & 8 & {\win 551} & 6 & 6
+Total & 14 & 148 & 6 & 6 & 8 & 550 & 8 & {\win 551} & 8 & {\win 551}
 \end{tabular}
@@
 \midrule
-\isanun & {\win 41} & 40 & 28 & 34 & 26 & 39 & 26 & 39 & 26 & 39 & 38 & 31
+\isanun & {\win 41} & 40 & 28 & 34 & 26 & 39 & 26 & 39 & 26 & 39
 \\
-\isanun & 14 & 15 & 15 & 15 & 1 & 15 & 15 & 15 & 1 & 15 & {\win 64} & 14
+\isa & 14 & {\win 15} & {\win 15} & {\win 15} & 1 & {\win 15} & {\win 15} & {\win 15} & 1 & {\win 15}
 \\
-\leon & 54 & 55 & 56 & 56 & 13 & 55 & 56 & 55 & 13 & 55 & {\win 93} & 53
+\leon & 54 & 55 & {\win 56} & {\win 56} & 13 & 55 & {\win 56} & 55 & 13 & 55
 \\
-\isam & 48 & 52 & 58 & 52 & 12 & 53 & {\win 58} & 50 & 12 & 53 & {\win 58} & 50
+\isam & 48 & 52 & {\win 58} & 52 & 12 & 53 & {\win 58} & 50 & 12 & 53
 \\
-\leonm & 94 & 107 & 106 & 107 & 39 & 108 & 106 & 108 & 39 & 108 & {\win 123} & 106
+\leonm & 94 & 107 & 106 & 107 & 39 & {\win 108} & 106 & {\win 108} & 39 & {\win 108}
 \\[\jot]
-Total & 251 & 269 & 263 & 264 & 91 & 270 & 261 & 267 & 91 & 270 & {\win 376} & 254
+Total & 251 & 269 & 263 & 264 & 91 & {\win 270} & 261 & 267 & 91 & {\win 270}
 \end{tabular}
@@
 \end{figure}
-
-\ajr{To revise}
 
@@
 
-Over \emph{unsat} benchmarks, \cvci is the
-clear winner, finding 376 total benchmarks to be unsatisfiable. 
-Apart from this configuration, 
+Moreover,
 the translation $\conv$ helps all configurations for \emph{unsat} responses as well.
-In particular,
 \ziii solves a total of 269 with the translation, whereas it solves only 251 without it.
-The configuration \cvcf improved even more significantly, 
-solving 270 unsat benchmarks overall, which is more than both \cvcd and \ziii.
+Surprisingly,
+the configuration \cvcf, which is not tailored for handling unsatisfiable benchmarks,
+solves 270 unsat benchmarks overall, which is more than both \cvcd and \ziii.
+These results suggest that the translation do not degrade the performance 
+of SMT solvers for unsatisfiable problems involving recursive functions, 
+and instead often improve their performance.
 \ajr{Importance of the translation helping unsat in practice for Isabelle/Nunchaku, if any}
@@
 \cite{reynolds-blanchette-2015-codata}. We also plan to integrate \cvc as a
-counterexample generator in proof assistants. Further work would also
+counterexample generator in proof assistants.\ajr{Revise: this should say plans to develop Nunchaku.} Further work would also
 include identifying additional sufficient conditions for admissibility, thereby

### 2016-01-25 ijcar2016/body.tex
Implementation in cvc4 section.

@@
 
-\textbf{TODO:} Mention monotonicity analysis and zero sorts.
+\paragraph{Optimizations for Finding Finite Models of $\conv(\varphi)$}
+Like other systems,
+the finite model finding capability of \cvc incrementally fixes bounds on the cardinalities of uninterpreted types,
+and increases these bounds until it encounters a model.
+When multiple sorts are present, it uses a fairness scheme which
+bounds the sum of cardinalities of all uninterpreted types~\cite{reynolds2013finite}.
+For instance, if a signature has two uninterpreted types $\tau_1$ and $\tau_2$,
+it will first search for models where 
+the cardinality of $\tau_1$ plus the cardinality of $\tau_2$ is at most $2$,
+then $3$, $4$, and so on.
+To accelerate the search for models,
+we use an optimization based on statically inferring \emph{monotone} types.
+A monotone type is one in which models can always be extended with an additional element of that type.
+Inference of monotone types has been show in previous work~\cite{claessen2011mono,korovin-2013}
+to be an important component of finite model finding systems.
+Moreover, it is easy to see that types $\fargtype{\con{f}}$ introduced by our translation $\conv$ are monotone.
+The finite model finding capability of \cvc may take advantage of this fact by fixing the bounds for all 
+monotone types simultaneously.
+That is, if $\tau_1$ and $\tau_2$ are both monotone types (either introduced by our translation, or are otherwise inferred to be),
+then we fix the bound for \emph{both} of these types to be $1$, then $2$, and so on.
+This scheme allows the solver greater flexibility with respect to the default scheme,
+and comes with no loss of generality with respect to models, since monotone sorts can always be extended to have equal cardinalities.
+
+By default, 
+\cvc uses techniques to minimize the number of literals it considers when constructing propositional satisfying assignments for formulas~\cite{relevancy2007}.
+However, we have found such techniques degrade performance for finite model finding on
+benchmarks having recursive functions that are defined by cases.
+For this reason, we prefer disabling such techniques for benchmarks coming from our translation.
+
+\ajr{More?}
 

### 2016-01-25 ijcar2016/body.tex
Add related work
edits up to mid of Sect 2

@@
 
-%%% I'm trying to avoid bold whenever it can be avoided (except for headings
-%%% etc.) --JB
-%\newcommand{\terms}{\mathbf{T}}
 \newcommand{\terms}{\mathcalx{T}}
@@
 \def\ENDPROC{\untab}
-%%% @ANDY: I'm trying "A" like "abstraction", again in sync with the abstract--concrete
-%%% terminology from abstract interpretation
 \newcommand{\conv}{\mathcalx{A}}
@@
 An empirical evaluation confirms that the approach improves 
-the performance of existing solvers on benchmarks from two sources. 
+the performance of existing solvers on benchmarks from three sources. 
 The translation is implemented as a preprocessor in the CVC4 solver
@@
 \cite{MouraBjoerner07,ReynoldsTinelliMoura14}.
-These methods work well in the context of proving (i..e, showing
+These methods work well in the context of proving (i.e., showing
 unsatisfiability), but they are of little help for finding models (i.e.,
@@
 Unfortunately, none of these fragments can accommodate the vast majority of
-quantified formulas that correspond to recursive function definitions: The
+quantified formulas that correspond to recursive function definitions. The
 essentially uninterpreted fragment does not allow the argument of a
@@
 algebraic datatypes. A simple example where both approaches fail is
-$\forall x : {\typeint}.\;\, \lite\bigl(
-x \leq 0,\allowbreak\; \con{p}( x ) \teq 1,\allowbreak\; \con{p}( x ) \teq 2 * \con{p}( x - 1 ) \bigr)$.
+$$\forall x : {\typeint}.\;\,
+\con{p}( x ) \teq \lite\bigl( x \leq 0,\allowbreak\;  1,\allowbreak\;
+  2 * \con{p}( x - 1 ) \bigr)$$ %.
 This state of affairs is unsatisfactory, given the frequency of
-recursive definitions in practice and the impending addition of a dedicated
-command for introducing them, \texttt{define-fun-rec}, to the SMT-LIB standard \cite{smtlib25}.
+recursive definitions in practice and the %impending
+addition of a %dedicated
+command for introducing them, \texttt{define-fun-rec},
+to the SMT-LIB standard \cite{smtlib25}.
 
@@
 (Section~\ref{sec:prelim}). This criterion is met by well-founded (terminating)
-recursive function definitions. \begin{rep}It is not met by inconsistent
-definitions such as $\forall x : {\typeint}.\;\, \con{f}(x) \teq \con{f}(x) +
-1$.\end{rep} %ct and it is possible to construct consistent examples that do not meet it.
-%as well as other example of consistent definition.
+recursive function definitions. It is not met by inconsistent
+definitions such as \,$\forall x : {\typeint}.\;\, \con{f}(x) \teq \con{f}(x) +
+1$.
 
@@
 formulas involving admissible recursive function definitions
-(Section~\ref{sec:encoding}). The main insight is that a
-recursive definition $\forall x : \tau.\;\, \const{f}(x) \teq t$
-can be translated to
+(Section~\ref{sec:encoding}). A
+recursive definition $\forall x : \tau.\;\, \const{f}(x) \teq \mathit{rhs}$
+is translated to
 $\forall a :
-\fargtype{\tau}.\;\, \const{f}(\farg{\con{f}}(a)) \teq t[\farg{\con{f}}(a)/x]$, where
-$\fargtype{\tau}$ is an uninterpreted \relax{abstract} type and $\farg{\con{f}} \begin{rep}:
-\fargtype{\tau} \to \tau\end{rep}$ converts the abstract type to the concrete
+\fargtype{\tau}.\;\, \const{f}(\farg{\con{f}}(a)) \teq \mathit{rhs}[\farg{\con{f}}(a)/x]$, where
+$\fargtype{\tau}$ is an uninterpreted \relax{abstract} type and $\farg{\con{f}} :
+\fargtype{\tau} \to \tau$ converts the abstract type to the concrete
 type. Additional constraints ensure that the abstract values that are relevant
 to the formula's satisfiability
-exist. The translation preserves satisfiability and\begin{rep}, for admissible definitions,\end{rep}
+exist. The translation preserves satisfiability and, for admissible definitions,
 unsatisfiability, and makes finite model finding possible for problems in this
@@
 
-Our empirical evaluation on benchmarks from the IsaPlanner proof planner~\cite{DBLP:conf/itp/JohanssonDB10}
-and the Leon verifier~\cite{blanc2013overview} provides
-evidence that this translation improves the effectiveness of the SMT solvers
-\cvc and \ziii for finding countermodels to verification conditions
-(Section~\ref{sec:evaluation}). The approach is implemented as a preprocessor
-in \cvc and in the Nunchaku model finder
-(Section~\ref{sec:implementations}). \begin{rep}%
-When invoked with a specific
-command-line option, the solver assumes that functions introduced using the
-\texttt{define-fun-rec} command are admissible and translates them accordingly.\end{rep}
+The approach is implemented as a preprocessor in \cvc and in the Nunchaku
+model finder (Section~\ref{sec:implementations}). As a case study, we employed
+Nunchaku on Isabelle/HOL~\cite{nipkow-et-al-2002} formalizations of binary
+trees and coinductive lists (Section~\ref{sec:case-studies}).
+%
+We evaluated the two evaluation on benchmarks from
+IsaPlanner~\cite{DBLP:conf/itp/JohanssonDB10}, Leon~\cite{blanc2013overview},
+and Isabelle/HOL, to demonstrate that this translation improves the
+effectiveness of the SMT solvers \cvc and \ziii for finding countermodels to
+verification conditions (Section~\ref{sec:evaluation}). Unlike earlier work,
+our approach can be combined with off-the-shelf SMT solvers
+(Section~\ref{sec:related-work}).
 
@@
 
-%
 \section{Preliminaries}
@@
 %
-For each $\tau \in \stypes{\Sigma}$,
-we fix an infinite set $\svars{\Sigma}_\tau$ of \emph{variables of type $\tau$} and
+We fix an infinite set $\svars{\Sigma}_\tau$ of \emph{variables of type $\tau$}
+for each $\tau \in \stypes{\Sigma}$ and
 define $\svars{\Sigma}$ as $\bigcup_{\tau \in \stypes{\Sigma}} \svars{\Sigma}_\tau$.
@@
 maps each type $\tau \in \stypes{\Sigma}$ to a nonempty set $\tau^\I$,
-the \emph{domain} of~$\tau$ in~$\I$,
+the \emph{domain} of~$\tau$ in~$\I,$
 each function symbol $\con{f} : \tau_1 \times \cdots \times \tau_n \rightarrow \tau$ in $\sfuns{\Sigma}$
@@
 %
-where $\con{f}_1, \ldots, \con{f}_n$ are distinct\begin{rep} function symbols\end{rep},
+where $\con{f}_1, \ldots, \con{f}_n$ are distinct function symbols,
 %ct added
@@
 and $\psi$ contains no function definitions.
-We call $\psi$ the \emph{conjecture} of $\varphi$.
+We call $\psi$ the \emph{goal} of $\varphi$.
 %\ct{Don't we need additional restrictions on the $\varphi_i$'s here?
@@
 a $\Sigma$-formula that is closed under function expansion and
-that entails the conjecture of $\varphi$.
+that entails the goal of $\varphi$.
 This implies the existence of a model for $\varphi$,
@@
 
+TODO:
+When invoked with a specific
+command-line option, the solver assumes that functions introduced using the
+\texttt{define-fun-rec} command are admissible and translates them accordingly.
+
 In \cvc, function definitions $\forallf{\con{f}} \vec x.\; \varphi$ can be written using the $\definefunreccmd$ command
@@
 mutated forms of its conjecture $\psi$. In particular, we considered unique
-conjectures that are obtained as a result of swapping a subterm of $\psi$ at
+formulas that are obtained as a result of swapping a subterm of $\psi$ at
 one position with another of the same type at another position.
@@
 small, easy-to-find countermodels. In total, we considered 213 mutated forms of
-conjectures from \isa and 427 mutated forms of conjectures from \leon. We will
+theorems from \isa and 427 mutated forms of theorems from \leon. We will
 call these sets \isam and \leonm, respectively.
@@
 We conclude that the translation $\conv$ enables SMT
-solvers to find countermodels for conjectures involving recursively defined functions
-whose definitions are admissible.
+solvers to find countermodels for conjectures involving recursively defined
+functions whose definitions are admissible.
 
@@
 \section{Related Work}
+\label{sec:related-work}
 
@@
 existential constraints generated by our translation $\conv$.
+An approach related in scope to ours is given in~\cite{baumgartner2013},
+which establishes the satisfiability of formulas in the presence of admissible
+axioms over infinite domains by proving their negation is entailed.
 

### 2016-01-25 smt2015/body.tex
edits up to mid of Sect 2

@@
 \usepackage[scaled=.82]{beramono}
-\usepackage{mathptmx}
+%\usepackage{mathptmx}
 \usepackage{listings}

### 2016-01-25 ijcar2016/body.tex
well founded and a bit more

@@
 
-\newcommand\closurefmla{
-  \ensuremath{\psi \models_T \bigwedge\nolimits_{\,i=1}^{n} \{ \varphi_i[ \vec t / \vec x ] \mid \con{f}_i( \vec t ) \in \terms( \psi ) \}} }
-
 \begin{definition}\rm
@@
 is \emph{closed under function expansion with respect to $\DDD$} if
-\begin{conf}$\closurefmla$. \end{conf}%
-\begin{rep}\[\closurefmla\]\end{rep}%
+\[\psi \models_T \smash{\bigwedge\nolimits_{\,i=1}^{n} \{ \varphi_i[ \vec t / \vec x ] \mid \con{f}_i( \vec t ) \in \terms( \psi ) \}}\]
 The set $\DDD$ is \emph{admissible} if for every $T$-satisfiable formula 
@@
 definition before applying the translation described in
-Section~\ref{sec:encoding}. It is interesting to connect it to the
+Section~\ref{sec:encoding}. It is useful to connect it to the
 standard notion of \emph{well-founded} function definitions, often called
@@
 First-order logic has no built-in notion of computation or termination. To ensure
-that a function specification is well-founded, it is sufficient to require that
+that a function specification is well founded, it is sufficient to require that
 the function would terminate when seen as a functional program, under \emph{some}
@@
 where $T$ is integer arithmetic extended with the uninterpreted symbol 
-$\con p:\typeint \rightarrow \typeint$, can be shown well-founded under a strategy that
+$\con p:\typeint \rightarrow \typeint$, can be shown well founded under a strategy that
 evaluates the condition of an $\lite$ before evaluating the relevant branch,
@@
 \emph{productive} corecursive functions. Corecursive functions are functions to
-a coalgebraic datatype. These functions can be non-well-founded, without
+a coalgebraic datatype. These functions can be ill founded, without
 their being inconsistent. Productive corecursive functions are functions that
@@
   \item[Unrolling]
-    adds a decreasing argument to non-well-founded predicates;
+    adds a decreasing argument to possibly ill-founded predicates;
   \item[Skolemization]

### 2016-01-25 ijcar2016/body.tex
more edits

@@
+%%% TODO: mention that admissible does not mean finite model
+%%% e.g. ite inside or outside, or custom Bool operators
+
 %%% BEGIN Times font
@@
 \DeclareSymbolFont{letters}{OML}{txmi}{m}{it}
-
 %%% END Times font
@@
 \let\const=\con
+
+\newcommand\innerkw[1]{\textrm{#1}}
 
@@
 
+\let\labelitemi=\labelitemii %% CHEAT!
+\makeatletter
+\ps@myheadings
+\makeatother
+
 \begin{document}
@@
 \author {Andrew Reynolds\inst{1} \and Jasmin Christian Blanchette\inst{2,3} \and \\ Simon~Cruanes\inst{2} \and Cesare~Tinelli \inst{1}}
-\authorrunning {A. Reynolds, J. C. Blanchette, C. Tinelli}
+\authorrunning {A. Reynolds et al.} % , J. C. Blanchette, S. Cruanes, C. Tinelli}
 \institute{
@@
 unsatisfiability, and makes finite model finding possible for problems in this
-class.\begin{conf} Detailed proofs of correctness are included in a technical
-report \cite{our-report}.\end{conf}
+class.
+%\begin{conf} Detailed proofs of correctness are included in a technical
+%report \cite{our-report}.\end{conf}
 
@@
 standard Boolean connectives ($\neg$, $\wedge$, $\vee$, etc.),
-and an if--then--else function symbol
+and an \innerkw{if}--\innerkw{then}--\innerkw{else} function symbol
 $\lite : \Bool \times \tau \times \tau \rightarrow \tau$
@@
 and $\psi$ contains no function definitions.
-We call $\psi$ the \emph{goal} of $\varphi$.
+We call $\psi$ the \emph{main formula} of $\varphi$.
 %\ct{Don't we need additional restrictions on the $\varphi_i$'s here?
@@
 evaluation order. For example, the definition
-$\forall x : {\typeint}.\;\, \lite\bigl(
-x \leq 0,\allowbreak\; \con{p}( x ) \teq 1,\allowbreak\; \con{p}( x ) \teq 2 * \con{p}( x - 1 ) \bigr)$,
+$\forall x : {\typeint}.\;\, 
+\con{p}( x ) \teq \lite\bigl( x \leq 0,\allowbreak\;  1,\allowbreak\;
+  2 * \con{p}( x - 1 ) \bigr)$,
 where $T$ is integer arithmetic extended with the uninterpreted symbol 
@@
 evaluates the condition of an $\lite$ before evaluating the relevant branch,
-ignoring the other branch. \begin{rep}Logically, such dependencies can be captured by
-congruence rules. \end{rep}Krauss developed these ideas in the more general context of
+ignoring the other branch. Logically, such dependencies can be captured by
+congruence rules. Krauss developed these ideas in the more general context of
 higher-order logic \cite[Section 2]{krauss-2009-phd}.
@@
 If $\DDD$ is a set of well-founded function definitions for\/
-$\sfundefs{\Sigma} \begin{rep} = \{\con{f}_1,\ldots,\con{f}_n\}\end{rep}$, then it is admissible.
+$\sfundefs{\Sigma} = \{\con{f}_1,\ldots,\con{f}_n\}$, then it is admissible.
 \end{theorem}
-\begin{rep}
 \begin{proofsketch}
@@
 \end{proofsketch}
-\end{rep}
-
-TODO: Say something about admissibility for tail-recursion: Watch out if the
-argument always changes...
-
-\begin{rep}
+
+% TODO: Say something about admissibility for tail-recursion: Watch out if the
+% argument always changes...
+
 Another useful fragment of function definitions is the class of
@@
 a corecursive definition infinitely computes a unique infinite object.
-
+%We conjecture that a large fragment of mixed recursive--corecursive
+%specifications are
+
+\begin{rep}
 Beyond totality, an admissible set can contain underspecified functions
@@
  \begin{itemize}
-   \itemx[] $\mathsf{if}$ $\tau = \Bool$ $\mathsf{and}$ $t = \boolop(t_1,\ldots,t_n)$ $\mathsf{then}$
+   \itemx[] $\innerkw{if}$ $\tau = \Bool$ $\mathsf{and}$ $t = \boolop(t_1,\ldots,t_n)$ $\innerkw{then}$
     \begin{itemize}
-      \itemx[] $\mathsf{let}$ $( t'_i,\, \chi_i ) = \conv_0( t_i,\, \pol( \boolop,\, i,\, p ) )$ $\mathsf{for}$ $i = 1, \ldots, n$ $\mathsf{in}$%\kern6.35em$\mid$
-      \itemx[] $\mathsf{let}$ $\chi = \chi_1 \wedge \cdots \wedge \chi_n$ $\mathsf{in}$
-      \itemx[] $\mathsf{if}$ $p = \ppos$ $\mathsf{then}$ \bigtuple{$\boolop(t'_1, \ldots, t'_n) ) \wedge \chi,\, \ltrue$}
-      \itemx[] $\mathsf{else}$ $\mathsf{if}$ $p = \pneg$ $\mathsf{then}$ \bigtuple{$\boolop(t'_1, \ldots, t'_n) \vee \neg \chi,\, \ltrue$}
-      \itemx[] $\mathsf{else}$ \bigtuple{$\boolop(t'_1, \ldots, t'_n),\, \chi$}
+      \itemx[] $\innerkw{let}$ $( t'_i,\, \chi_i ) = \conv_0( t_i,\, \pol( \boolop,\, i,\, p ) )$ $\mathsf{for}$ $i = 1, \ldots, n$ $\innerkw{in}$%\kern6.35em$\mid$
+      \itemx[] $\innerkw{let}$ $\chi = \chi_1 \wedge \cdots \wedge \chi_n$ $\innerkw{in}$
+      \itemx[] $\innerkw{if}$ $p = \ppos$ $\innerkw{then}$ \bigtuple{$\boolop(t'_1, \ldots, t'_n) ) \wedge \chi,\, \ltrue$}
+      \itemx[] $\innerkw{else}$ $\innerkw{if}$ $p = \pneg$ $\innerkw{then}$ \bigtuple{$\boolop(t'_1, \ldots, t'_n) \vee \neg \chi,\, \ltrue$}
+      \itemx[] $\innerkw{else}$ \bigtuple{$\boolop(t'_1, \ldots, t'_n),\, \chi$}
 %\\[-.8\baselineskip]
     \end{itemize}
-  \itemx[] $\mathsf{else}$ $\mathsf{if}$ $t = \forallf{\con{f}} \vec x.\;\, u$ $\mathsf{then}$
+  \itemx[] $\innerkw{else}$ $\innerkw{if}$ $t = \forallf{\con{f}} \vec x.\;\, u$ $\innerkw{then}$
     \begin{itemize}
-      \itemx[] $\mathsf{let}$ $( u',\, \chi ) = \conv_0( u,\, p )$ $\mathsf{in}$ \bigtuple{$\forall a : \fargtype{\con{f}}.\; u' [ \vecfarg{\con{f}}( a ) / \vec x ],\, \ltrue$}
+      \itemx[] $\innerkw{let}$ $( u',\, \chi ) = \conv_0( u,\, p )$ $\innerkw{in}$ \bigtuple{$\forall a : \fargtype{\con{f}}.\; u' [ \vecfarg{\con{f}}( a ) / \vec x ],\, \ltrue$}
 %\\[-.8\baselineskip]
     \end{itemize}
-  \itemx[] $\mathsf{else}$ $\mathsf{if}$ $t = \forall \vec x.\;\, u$ $\mathsf{then}$
+  \itemx[] $\innerkw{else}$ $\innerkw{if}$ $t = \forall \vec x.\;\, u$ $\innerkw{then}$
     \begin{itemize}
-      \itemx[] $\mathsf{let}$ $( u',\, \chi ) = \conv_0( u,\, p )$ $\mathsf{in}$ \bigtuple{$\forall \vec x.\; u',\, \forall \vec x.\; \chi$}
+      \itemx[] $\innerkw{let}$ $( u',\, \chi ) = \conv_0( u,\, p )$ $\innerkw{in}$ \bigtuple{$\forall \vec x.\; u',\, \forall \vec x.\; \chi$}
 %\\[-.8\baselineskip]
     \end{itemize}
-   \itemx[] $\mathsf{else}$
+   \itemx[] $\innerkw{else}$
    \begin{itemize}
@@
 \itemx[\ ]
-$\conv( \varphi )$ $=$ $\mathsf{let}$ $( \varphi',\, \chi ) = \conv_0( \varphi,\, \ppos )$ $\mathsf{in}$ $\varphi'$
-%\\[-1.25\baselineskip] %% TYPESETTING
+$\conv( \varphi )$ $=$ $\innerkw{let}$ $( \varphi',\, \chi ) = \conv_0( \varphi,\, \ppos )$ $\innerkw{in}$ $\varphi'$
+\\[-1.5\baselineskip] %% TYPESETTING
 \end{enumerate}
@@
 of terms, namely those occurring in applications of $\con{f}$ 
-that are relevant to the satisfiability of $\varphi$.
+that are relevant to the satisfiability of $\varphi$. The constraints are
+generated as deep in the formula as possible, based on our knowledge of the
+polarities of logical connectives, to allow models in which the domains
+interpreting the $\fargtype{\con{f}}$ types are as small as possible.
 
@@
 $\conv_0$ calls itself recursively on the arguments $t_i$ and polarity $\pol(
-\boolop, i, p )$, with $\pol$ defined as
+\boolop, i, p )$, with $\pol$ d\rlap{efined as}
 \[\pol( \boolop, i, p ) =
@@
 The term $t$ is then reconstructed as $\boolop(t_1',\ldots,t_n')$ 
-where each $t_i'$ is the result of the recursive call with argument $t_i$. 
+where each $t_i'$ is the result of the recursive call with argument~$t_i$. 
 If the polarity $p$ associated with $t$ is $\ppos$, 
 $\conv_0$ conjunctively adds to $\boolop(t_1',\ldots,t_n')$ the constraint $\chi$ derived from
-the subterms. Dually, if $p$ is $\pneg$, it adds a
+the subterms and returns $\top$ as the constraint.
+Dually, if $p$ is $\pneg$, it adds a
 disjunction with the negated constraint, to achieve the same overall effect.
@@
 returned to the caller.
-%
-\begin{rep}\par\end{rep}
-%
+
 If $t$ is a function definition, % of a function in $\sfundefs{\Sigma}$,
@@
 If $t$ is an unannotated quantified formula, $\conv_0$ calls itself
-on the body with the same polarity. \begin{rep}A quantifier is prefixed to the
-quantified formula and constraint returned by the recursive call.\end{rep}
+on the body with the same polarity; a quantifier is prefixed to the
+quantified formula and constraint returned by the recursive call.
 Otherwise, $t$ is either an application of an uninterpreted predicate symbol or a term
@@
 \label{ex:translation}
-Let $T$ be linear arithmetic with the uninterpreted symbols  $\{ \const{c} : { \typeint },\> \con{s} : { \typeint \rightarrow \typeint } \}$.
+Let $T$ be linear integer arithmetic
+with the uninterpreted symbols  $\{ \const{c} : { \typeint },\allowbreak\> \con{s} : { \typeint \rightarrow \typeint } \}$.
 Let $\varphi$ be the $\Sigma$-formula
@@
 %\begin{equation} \label{eq:approx-interp}
-\begin{conf}$\badlambda$. \end{conf}%
-\begin{rep}\[\badlambda\]\end{rep}%
+\[\badlambda\]
 %\end{equation}
@@
 also give correct information for the defined functions at all points belonging
-to the domains of the corresponding abstract types. This can sometimes help
-users debug their specifications.
+to the domains of the corresponding abstract types $\fargtype{\con{f}}$.
+This can sometimes help users debug their specifications.
 
@@
 
-\begin{rep}
 %In the following,
@@
 definitions in the translation.
-Formally, we define
-$\absconstraints( L )$
-as
-$\{ \exists a : {\fargtype{\con{f}}}.\;\, \vecfarg{\con{f}}( a ) \teq \vec t
-\mid \con{f}( \vec t ) \in \terms( L ),\: \con{f} \in \sfundefs{\Sigma}
-\}$.
+Formally,
+\[\absconstraints( L ) \vthinspace=\vthinspace \smash{\{ \exists a : {\fargtype{\con{f}}}.\;\, \vecfarg{\con{f}}( a ) \teq \nobreak \vec t
+\nobreak\mid\allowbreak \con{f}( \vec t ) \in \terms( L ),\: \con{f} \in \sfundefs{\Sigma}
+\}}\]
 The following lemma states the central invariant behind the translation $\conv$.
@@
 \end{corollary}
-\end{rep}
 
@@
 \end{theorem}
-\begin{rep}
 \begin{proofsketch}
@@
 since any $\Sigma$-interpretation that satisfies $\varphi$ can be extended in a straightforward way to 
-an $\extendsig{\Sigma}$-interpretation that satisfies $\conv( \varphi )$.
+an $\extendsig{\Sigma}$-interpretation that satisfies $\conv( \varphi )$, by
+interpreting the abstract types in the same way as the cartesian products they
+abstract, thereby satisfying all existential constraints introduced by \conv.
 Second, if a model is found for $\conv( \varphi )$,
@@
 a $\Sigma$-formula that is closed under function expansion and
-that entails the goal of $\varphi$.
+that entails the main formula of $\varphi$.
 This implies the existence of a model for $\varphi$,
@@
 %For clarity, we demonstrate the consequences of this theorem by means of an example.
-\end{rep}
 
@@
 a datatype declaration, then two mutually recursive coinductive predicates
-are defined, and finally a goal (``find a non-zero natural number $m$ that
-is even'').
+are defined, and finally a distinguished axiom called the goal
+(``find a non-zero natural number $m$ that is even'').
 \begin{quote}
@@
 the cardinalities of the atomic types. This procedure was pioneered by McCune
-in the earlier versions of
-\begin{conf}MACE\end{conf}\begin{rep}Mace (or MACE)\end{rep}
-\cite{mccune-1994}.
+in the earlier versions of Mace (or MACE) \cite{mccune-1994}.
 Other conceptually similar finders are Paradox \cite{claessen-sorensson-2003}
@@
 This approach was pioneered by FINDER
-\cite{slaney-1994} and SEM \cite{zhang-zhang-1995}\begin{rep} and serves as
+\cite{slaney-1994} and SEM \cite{zhang-zhang-1995} and serves as
 the basis of many more model finders, notably the Alloy Analyzer's precursor
 \cite{jackson-1996} and the later versions of~Mace
-\cite{mccune-prover9-mace4}\end{rep}.
-
-\begin{rep}
+\cite{mccune-prover9-mace4}.
+
 Most of the above tools cannot cope with algebraic datatypes or other infinite
-types. \end{rep}%
+types.
 Kuncak and Jackson \cite{kuncak-jackson-2005} presented an idiom for
@@
 of) models for formulas in the \relax{existential--bounded-universal} fragment%
-\begin{rep} (i.e., formulas whose prenex normal forms contain no unbounded universal
-quantifiers ranging over datatypes)\end{rep}. This idiom was further developed by Dunets
+(i.e., formulas whose prenex normal forms contain no unbounded universal
+quantifiers ranging over datatypes). This idiom was further developed by Dunets
 et al.\ \cite{dunets-et-al-2010}, who presented a translation scheme
@@
 relies on another variant of Kuncak and Jackson's approach inside a
-Kleene-style three-valued logic\begin{rep}, inspired by abstract interpretation\end{rep}.
-\begin{rep}It was also the first tool of its kind to support corecursion and
-coalgebraic datatypes \cite{blanchette-2013-relational}.\end{rep}
+Kleene-style three-valued logic, inspired by abstract interpretation.
+It was also the first tool of its kind to support corecursion and
+coalgebraic datatypes \cite{blanchette-2013-relational}.
 The three-valued logic approach extends each
@@
 quantifiers altogether by unfolding recursive definitions up to a certain
-depth\begin{rep}, which is increased on a per-need basis\end{rep}.
+depth, which is increased on a per-need basis.
 Our translation~$\conv$ works in an analogous manner, 
@@
 and quantifier instantiation is used in lieu of function unfolding.
-\begin{rep}It would be
+It would be
 worth investigating how existing approaches for function
 unfolding can inform approaches for dedicated quantifier instantiation
-techniques for function definitions, and vice versa.\end{rep}
+techniques for function definitions, and vice versa.
 
@@
 programs. QuickCheck \cite{claessen-hughes-2000} for Haskell is an
-early example, based on random testing. Bounded exhaustive testing\begin{rep}
-\cite{runciman-et-al-2008}\end{rep} and narrowing\begin{rep}
-\cite{lindblad-2008-testing}\end{rep} are other successful strategies. These
+early example, based on random testing. Bounded exhaustive testing
+\cite{runciman-et-al-2008} and narrowing
+\cite{lindblad-2008-testing} are other successful strategies. These
 tools are often much faster than model finders, but they typically cannot cope

### 2016-01-25 ijcar2016/body.tex
compile

@@
 interpreting the abstract types in the same way as the cartesian products they
-abstract, thereby satisfying all existential constraints introduced by \conv.
+abstract, thereby satisfying all existential constraints introduced by $\conv$.
 Second, if a model is found for $\conv( \varphi )$,

### 2016-01-25 ijcar2016/body.tex
tuned terminology

@@
 and $\psi$ contains no function definitions.
-We call $\psi$ the \emph{main formula} of $\varphi$.
+We call $\psi$ the \emph{goal} of $\varphi$.
 %\ct{Don't we need additional restrictions on the $\varphi_i$'s here?
@@
 a $\Sigma$-formula that is closed under function expansion and
-that entails the main formula of $\varphi$.
+that entails the goal of $\varphi$.
 This implies the existence of a model for $\varphi$,

### 2016-01-25 ijcar2016/body.tex
tuned for space

@@
 Furthermore, $\psi$ entails $\varphi_0$ since $L_0 \subseteq L_\infty$.
-Since $\psi$ is a $T$-satisfiable formula that is closed under function expansion with respect to $\DDD$ and $\DDD$ is admissible, 
+Since $\psi$ is a $T$-satisfiable formula that is closed under function expansion\begin{rep} with respect to $\DDD$\end{rep} and $\DDD$ is admissible, 
 by definition there exists a $\Sigma$-interpretation $\I$ satisfying $\psi \wedge \DDD$, which entails $\DDD \wedge \varphi_0$,
@@
 abstract, thereby satisfying all existential constraints introduced by $\conv$.
-Second, if a model is found for $\conv( \varphi )$,
-then the constraints $\exists a : \fargtype{\con{f}}.\;\, \vecfarg{\con{f}}( a ) \teq \vec t$
-occurring in $\conv( \varphi )$ ensure that this model also satisfies
+Conversely, if a model is found for $\conv( \varphi )$,
+existential constraints
+%the constraints $\exists a : \fargtype{\con{f}}.\;\, \vecfarg{\con{f}}( a ) \teq \vec t$
+introduced by $\conv$ ensure that this model also satisfies
 a $\Sigma$-formula that is closed under function expansion and
@@
 (the first conjunct of~(\ref{eq:ex-after})) and $\absconstraints( L_0
-)$, it must also satisfy\begin{rep} the formula $\psi_1$:\end{rep}
+)$, it must also satis\rlap{fy}
 \[
@@
 Hence, by repeated application of this reasoning, it follows that
-a model of formula~(\ref{eq:ex-after}) that interprets $\con{c}$ as $n$ must also satisfy% the formula
-~$\psi$:
+a model of formula~(\ref{eq:ex-after}) that interprets $\con{c}$ as $n$ must also satisfy~$\psi$:
 %

### 2016-01-26 ijcar2016/body.tex
break
Fix results rows
some edits (mostly on 4.1)

@@
 
-\title{Model Finding for Recursive Functions in SMT\thanks{%
-This research is partially supported by the Inria technological development
-action ``Contre-exemples utilisables par Isabelle et Coq'' (CUIC).
-}
-}
+\title{Model Finding for Recursive Functions in SMT}
 \titlerunning {Model Finding for Recursive Functions in SMT}
@@
 Hence, by repeated application of this reasoning, it follows that
-a model of formula~(\ref{eq:ex-after}) that interprets $\con{c}$ as $n$ must also satisfy~$\psi$:
-%
-\[\con{s}( \con{c} ) > 100 \wedge %\displaystyle
-  \smash{\bigwedge\nolimits_{\vthinspace i=0}^{\!n-1}} \bigl( \neg (\con{c}-i \leq 0) \wedge \con{s}( \con{c}-i ) \teq \con{c}-i + \con{s}( \con{c}-i-1 ) \bigr)
-\wedge \con{c}-n \leq 0 \wedge \con{s}( \con{c}-n ) \teq 0\]
+a model of formula~(\ref{eq:ex-after}) that interprets $\con{c}$ as~$n$ must also satisfy~$\psi$:
+%
+\[\begin{aligned}[t]
+\con{s}( \con{c} ) > 100 %\displaystyle
+& {} \wedge \relax{\bigwedge\nolimits_{\vthinspace i=0}^{\!\smash{n-1}}} \bigl( \neg (\con{c}-i \leq 0) \wedge \con{s}( \con{c}-i ) \teq \con{c}-i + \con{s}( \con{c}-i-1 ) \bigr) \\[-\jot]
+& {} \wedge \con{c}-n \leq 0 \wedge \con{s}( \con{c}-n ) \teq 0
+\end{aligned}\]
 %
@@
 We have implemented the translation $\conv$ in two separate systems, 
-as a preprocessor in the development version of \cvc (version~1.5 prerelease),
-and in the higher-order model finder Nunchaku.
-This section describes how the translation is implemented in both these systems,
-as well as optimizations used by \cvc for finding models for the translated benchmarks.
+as a preprocessor in \cvc (version~1.5 prerelease)
+and in the \cvc-based higher-order model finder Nunchaku.
+This section describes how the translation is implemented in these systems,
+as well as optimizations used by \cvc for finding models of translated
+problems. % for the translated benchmarks.
 
@@
 
-TODO:
-When invoked with a specific
-command-line option, the solver assumes that functions introduced using the
-\texttt{define-fun-rec} command are admissible and translates them accordingly.
+% When invoked with a specific
+% command-line option, the solver assumes that functions introduced using the
+% \texttt{define-fun-rec} command are admissible and translates them accordingly.
 
 In \cvc, function definitions $\forallf{\con{f}} \vec x.\; \varphi$ can be written using the $\definefunreccmd$ command
-from the SMT-LIB 2.5 standard~\cite{smtlib25}.
+from SMT-LIB 2.5~\cite{smtlib25}.
 Formula~(\ref{eq:ex-before}) from Example~\ref{ex:translation}
@@
 % are aesthetic and serve a purpose --jb
-\begin{verbatim}
-(define-fun-rec s ((x Int)) Int (ite (<= x 0) 0 (+ x (s (- x 1)))))
-(declare-fun c () Int)
-(assert (> (s c) 100))
-(check-sat)
-\end{verbatim}
+\verb|(define-fun-rec s ((x Int)) Int (ite (<= x 0) 0 (+ x (s (- x 1)))))|\kern-200mm \\
+\verb|(declare-fun c () Int)| \\
+\verb|(assert (> (s c) 100))| \\
+\verb|(check-sat)|
 \end{quote}
@@
 $\forallf{\con{s}} x.\;\, \lite\bigl( x \leq 0,\; \con{s}( x ) \teq 0,\; \con{s}( x ) \teq \con{s}( x-1 ) \bigr)$. \end{conf}%
-If \cvc's finite model finding mode for recursive functions is enabled (using
-the command-line option \texttt{-}\texttt{-fmf-fun}), it will replace its list of known
-assertions based on the translation $\conv$ before checking for satisfiability.
+By specifying the command-line option \texttt{-}\texttt{-fmf-fun}, users
+can enable \cvc's finite model finding mode for recursive functions.
+In this mode, \cvc will replace its list of known
+assertions based on the $\conv$ translation before checking for satisfiability.
 %If the functions provided in the input are admissible,
@@
 
-The \texttt{-}\texttt{-fmf-fun} option tells \cvc to assume that
+With the \texttt{-}\texttt{-fmf-fun} option enabled, \cvc assumes that
 functions introduced using \texttt{define-\allowbreak fun-\allowbreak rec} are admissible.
-We stress that admissibility must be discharged separately by the user---e.g., using a
+Admissibility must be discharged separately by the user---e.g., using a
 syntactic criterion or a termination prover.
@@
 
-\paragraph{Optimizations for Finding Finite Models of $\conv(\varphi)$}
+\cvc implements a few optimizations designed to help finding finite models of
+$\conv(\varphi)$.
 Like other systems,
@@
 \\
-\leon & 54 & 55 & {\win 56} & {\win 56} & 13 & 55 & {\win 56} & 55 & 13 & 55
+\isam & 54 & 55 & {\win 56} & {\win 56} & 13 & 55 & {\win 56} & 55 & 13 & 55
 \\
-\isam & 48 & 52 & {\win 58} & 52 & 12 & 53 & {\win 58} & 50 & 12 & 53
+\leon & 48 & 52 & {\win 58} & 52 & 12 & 53 & {\win 58} & 50 & 12 & 53
 \\
@@
 set of Leon benchmarks used in the evaluation, and the anonymous reviewers for their
-suggestions and comments.
+suggestions and comments. Cruanes is supported by the Inria technological
+development action ``Contre-exemples utilisables par Isabelle et Coq'' (CUIC).
 }

### 2016-01-26 ijcar2016/body.tex
tuned 4.1

@@
 Like other systems,
-the finite model finding capability of \cvc incrementally fixes bounds on the cardinalities of uninterpreted types,
+the finite model finding capability of \cvc incrementally fixes bounds on the cardinalities of uninterpreted types
 and increases these bounds until it encounters a model.
-When multiple sorts are present, it uses a fairness scheme which
+When multiple types are present, it uses a fairness scheme that
 bounds the sum of cardinalities of all uninterpreted types~\cite{reynolds2013finite}.
-For instance, if a signature has two uninterpreted types $\tau_1$ and $\tau_2$,
+For example, if a signature has two uninterpreted types $\tau_1$ and $\tau_2$,
 it will first search for models where 
-the cardinality of $\tau_1$ plus the cardinality of $\tau_2$ is at most $2$,
+$\left\vert\tau_1\right\vert + \left\vert\tau_2\right\vert$ is at most $2$,
 then $3$, $4$, and so on.
 To accelerate the search for models,
-we use an optimization based on statically inferring \emph{monotone} types.
-A monotone type is one in which models can always be extended with an additional element of that type.
-Inference of monotone types has been show in previous work~\cite{claessen2011mono,korovin-2013}
-to be an important component of finite model finding systems.
-Moreover, it is easy to see that types $\fargtype{\con{f}}$ introduced by our translation $\conv$ are monotone.
-The finite model finding capability of \cvc may take advantage of this fact by fixing the bounds for all 
-monotone types simultaneously.
-That is, if $\tau_1$ and $\tau_2$ are both monotone types (either introduced by our translation, or are otherwise inferred to be),
-then we fix the bound for \emph{both} of these types to be $1$, then $2$, and so on.
-This scheme allows the solver greater flexibility with respect to the default scheme,
-and comes with no loss of generality with respect to models, since monotone sorts can always be extended to have equal cardinalities.
-
-By default, 
-\cvc uses techniques to minimize the number of literals it considers when constructing propositional satisfying assignments for formulas~\cite{relevancy2007}.
-However, we have found such techniques degrade performance for finite model finding on
-benchmarks having recursive functions that are defined by cases.
-For this reason, we prefer disabling such techniques for benchmarks coming from our translation.
-
-\ajr{More?}
+we implemented an optimization based on statically inferring \emph{monotonic} types.
+A monotonic type is one in which models can always be extended with additional
+elements of that type \cite{blanchette-krauss-2011,claessen-et-al-2011}.
+%%% I find this part more confusing than helping, since we use monotonicity to merge
+%%% sorts (or at least their cardinality bounds), whereas Paradox and iProver use
+%%% it to separate sorts! Our use of monotonicity is more related to Nitpick's. --jb
+%%%
+% korovin-2013
+%Inference of monotonic types has been show in previous work~
+%to be an important component of finite model finding systems.
+Types $\fargtype{\con{f}}$ introduced by our translation $\conv$ are monotonic,
+because $\teq$ is never used directly on such types \cite{claessen-et-al-2011}.
+%
+%The finite model finding capability of
+\cvc takes advantage of this by fixing the bounds for all 
+monotonic types simultaneously.
+That is, if $\tau_1$ and $\tau_2$ are inferred monotonic (whether they are
+introduced by our translation or are present in the original problem),
+the solver fixes the bound for \relax{both} types to be $1$, then $2$, and so on.
+This scheme allows the solver greater flexibility compared with the default scheme,
+and comes with no loss of generality with respect to models,
+since monotonic types can always be extended to have equal cardinalities.
+
+By default, \cvc uses techniques to minimize the number of literals it
+considers when constructing propositional satisfying assignments for
+formulas~\cite{relevancy2007}. However, we have found such techniques degrade
+performance for finite model finding on problems having recursive functions
+that are defined by cases. For this reason, we prefer disabling such
+techniques for problems coming from our translation.
+
+%%% \ajr{More?} -- Probably no space. Otherwise, you could mention zero sorts. --jb
 
@@
 The configuration \cvcfe additionally incorporates heuristic quantifier instantiation as described in Section 2.3 of~\cite{reynolds-et-al-2013},
-and \cvcfm incorporates the fairness scheme for monotonic sorts as described in Section~\ref{ssec:cvc4}.
+and \cvcfm incorporates the fairness scheme for monotonic types as described in Section~\ref{ssec:cvc4}.
 

### 2016-01-26 ijcar2016/body.tex
tuned

@@
 That is, if $\tau_1$ and $\tau_2$ are inferred monotonic (whether they are
-introduced by our translation or are present in the original problem),
+present in the original problem or introduced by our translation),
 the solver fixes the bound for \relax{both} types to be $1$, then $2$, and so on.
@@
 performance for finite model finding on problems having recursive functions
-that are defined by cases. For this reason, we prefer disabling such
-techniques for problems coming from our translation.
+that are defined by cases. For this reason, we disable the techniques for
+problems produced from our translation.
 

### 2016-01-26 ijcar2016/body.tex
more on Nun

@@
 
-\hyphenation{counter-example counter-examples Isa-belle}
+\hyphenation{counter-example counter-examples Isa-belle data-type data-types
+  co-data-type co-data-types}
 
@@
 
-Nunchaku is a new model finder, intended
-to be used for finding counter-examples in several proof assistants. The
-first version, 0.1, was released in January 2016 with support for (co)datatypes,
-(co)recursive functions and (co)inductive predicates.
-Support for higher-order
-functions is planned in the next release.
-We intend for Nunchaku to be supported at least in Isabelle (where basic
-support for it already exists), Coq, and the TLA\textsuperscript{+} proof
-system.
-
-Nunchaku is the spiritual successor to Nitpick~\cite{blanchette-nipkow-2010},
-but is developed as a standalone OCaml program (with its own input language) so
-as to be usable independently from Isabelle.
-Whereas Nitpick relies on propositional checking with Kodkod,
-generating a succession of problems where cardinalities of finite types
-grow at each step, Nunchaku
-translates its input to one first-order logic formula that targets
-the finite-model finding fragment of CVC4, including the (co)datatypes support.
-Using CVC4 will also allow Nunchaku to provide efficient arithmetic and makes
-equality reasoning straightforward.
-
-The following problem is expressed in Nunchaku's input syntax; it starts with
-a datatype declaration, then two mutually recursive coinductive predicates
-are defined, and finally a distinguished axiom called the goal
-(``find a non-zero natural number $m$ that is even'').
+Nunchaku is a new higher-order model finder designed to be integrated with
+several proof assistants. The first version, 0.1, was released in January 2016
+with support for (co)datatypes, (co)recursive functions and (co)inductive
+predicates. Support for higher-order functions is planned for the next
+release. We have developed a preliminary Isabelle frontend and are planning
+further frontends for Coq, the TLA\textsuperscript{+} Proof System, and
+possibly other systems.
+
+Nunchaku is the spiritual successor to Nitpick for
+Isabelle/HOL~\cite{blanchette-nipkow-2010}, but is developed as a standalone
+OCaml program, with its own input language. Whereas Nitpick generates a
+succession of problems where cardinalities of finite types grow at each step,
+Nunchaku translates its input to one first-order logic program that targets
+the finite model finding fragment of CVC4, including interpreted
+(co)datatypes \cite{reynolds-blanchette-2015-codata}.
+Using CVC4 also allows Nunchaku to provide efficient arithmetic
+reasoning and to detect unsatisfiability in addition to satisfiability.
+% and makes
+% equality reasoning straightforward.
+
+The input syntax was inspired by that other systems based on higher-order
+logic (e.g., Isabelle/HOL) and by functional programming languages (e.g.,
+OCaml). The following simple problem gives a taste of the syntax:
+%
 \begin{quote}
-\begin{verbatim}
-data nat := zero | Suc nat.
-
-copred even : nat -> prop :=
-  even zero;
-  forall (n : nat). odd n => even (Suc n)
-and odd : nat -> prop :=
-  forall (n : nat). even n => odd (Suc n).
-
-goal exists m. (even m && ~ (m = zero)).
-\end{verbatim}
+\verb$data nat := Zero | Suc nat.$ \\[2\jot]
+\verb$pred even : nat -> prop :=$ \\
+\verb$  even Zero;$ \\
+\verb$  forall n. odd n => even (Suc n)$ \\
+\verb$and odd : nat -> prop :=$ \\
+\verb$  forall n. even n => odd (Suc n).$ \\[2\jot]
+\verb$val m : nat.$ \\
+\verb$goal even m && ~ (m = Zero).$
 \end{quote}
+%
+The problem defines a datatype (\texttt{nat}) and two mutually recursive
+inductive predicates (\texttt{even} and \texttt{odd}), it declares a constant
+\texttt{m}, and specifies a goal to satisfy (``$m$ is even and nonzero'').
+Nunchaku quickly finds the following partial model:
+%
+\begin{quote}
+\verb$val m := Suc (Suc Zero).$ \\
+\verb$val odd := fun x. if x = Suc Zero then true else ?__.$ \\
+\verb$val even := fun x. if x = Suc (Suc Zero) || x = Zero then true$ \\
+\verb$                   else ?__.$
+\end{quote}
+%
+The partial model gives sufficient information to the user to evaluate the
+goal: ``2 is even if 1 is odd, 1 is odd if 0 is even, and 0 is even.''
 
 Nunchaku parses and types the input problem before applying a sequence
-of translations, each reducing the distance to the target fragment. Here,
-for instance, the coinductive predicates
-\verb!even! and \verb!odd! are \emph{polarized} (specialized into
-\verb!even+!, used only in positive positions, and \verb!even-! used in negative
-positions),
-$m$ is Skolemized into a fresh constants,
-then then translated into admissible functions
-before another pass uses the encoding from Section~\ref{sec:encoding}.
-If a model is found, it is translated back to the input language.
-Conceptually, the sequence of transformation is a \emph{two ways pipeline}
-built by composing pairs \textsf{(encode,decode)}.
-For each such pair, \textsf{encode}
-transforms a problem over a signature $\Sigma$ in a logic
-$\mathcal{L}$ to a problem over a signature $\Sigma'$ in a slightly different
-logic $\mathcal{L}'$, and \textsf{decode} translates a model of $\mathcal{L'}$
-over $\Sigma'$ into a model of $\mathcal{L}$ over $\Sigma$. The pipeline
-currently implemented in Nunchaku is described below.
+of translations, each reducing the distance to the target fragment.
+In our example, the predicates
+\verb!even! and \verb!odd! are \emph{polarized} (specialized into a pair of
+predicates such that one is used in positive positions and the other in
+negative positions),
+$m$ is skolemized into a fresh constants,
+then translated into admissible recursive functions,
+before another pass applies the encoding described in this paper.
+If a model is found, it is translated back to the input language,
+with \verb$?__$ placeholders as necessary.
+
+Conceptually, the sequence of transformation is a \relax{two-way pipeline}
+built by composing pairs $(\textsf{encode}, \textsf{decode})$ of
+transformations. For each such pair, \textsf{encode} transforms a problem over
+a signature $\Sigma$ in a logic $\mathcal{L}$ to a problem over a signature
+$\Sigma'$ in a logic $\mathcal{L}'$, and \textsf{decode} translates a model in
+$\mathcal{L'}$ over $\Sigma'$ into a model in $\mathcal{L}$ over $\Sigma$,
+in the spirit of institution theory \cite{goguen-burstall-1992}. The
+pipeline currently implemented in Nunchaku consists of the following phases:
 
@@
   \item[Type skolemization]
-    replaces $\exists \alpha. p[\alpha]$
-      with $p[\tau]$ where $\tau$ is a fresh type;
+    replaces $\exists \alpha.\; \varphi[\alpha]$
+      with $\varphi[\tau]$, where $\tau$ is a fresh type;
   \item[Monomorphization]
-    specializes polymorphic functions on their type arguments;
-  \item[Elimination of Equations]
-    translates multiple-equations definitions of functions into
+    specializes polymorphic definitions on their type arguments and removes
+    unused definitions;
+  \item[Elimination of equations]
+    translates multiple-equation definitions of recursive functions into
       a single nested pattern matching;
   \item[Polarization]
-    specializes some predicates into a version used in positive position
-      and one used in negative position;
+    specializes predicates into a version used in positive positions
+      and a version used in negative positions;
   \item[Unrolling]
-    adds a decreasing argument to possibly ill-founded predicates;
+    adds a decreasing argument to possibly ill-founded predicates to make
+    them well founded;
   \item[Skolemization]
@@
   \item[Elimination of Predicates]
-    changes a multi-clauses (co)inductive predicate definition
+    changes a multiple-clauses (co)inductive predicate definition
       into a single recursive function;
-  \item[Recursion Elimination]
+  \item[Recursion elimination]
     performs the encoding from Section~\ref{sec:encoding};
-  \item[Elimination of Match]
-    transforms pattern-matching expressions into tests and selectors;
-  \item[CVC4 Invocation] calls CVC4 to obtain a model.
+  \item[Elimination of pattern matching]
+    rewrites pattern-matching expressions using datatype
+    discriminators and selectors;
+  \item[CVC4 invocation] runs CVC4 to obtain a model.
 %}}}
@@
 
-TODO: similar to institutions \cite{goguen-meseguer} (?)
-
-\subsection{Decoding of Recursive Functions}
-
-Nunchaku is designed for finding counter-models to formulas a user thinks
-are theorems. Decoding models from CVC4 to hide any trace of the successive
-encodings is therefore vital for readability. Each step
-of the pipeline only has to concern itself with removing its changes to
-the signature and logic; our encoding of recursive functions is no exceptions.
-
-TODO explain how:
-- domains of proj. functions are computed
-- domains of original function is deduced
-- decision tree for original function is built
-- note how junk is removed in step 3 (but undefined "..." as last case)
-
-\subsection{Support for Higher-Order Functions}
-
-TODO:
-- currying with explicit application symbol --> function arguments are now
-  just constants
-- apply encoding on application symbol
-- gotcha: one function can be used with several arities.
-
-- optim: specialization (similar to monomorphization)
+% \subsection{Decoding of Recursive Functions}
+% 
+% Nunchaku is designed for finding counter-models to formulas a user thinks
+% are theorems. Decoding models from CVC4 to hide any trace of the successive
+% encodings is therefore vital for readability. Each step
+% of the pipeline only has to concern itself with removing its changes to
+% the signature and logic; our encoding of recursive functions is no exceptions.
+% 
+% TODO explain how:
+% - domains of proj. functions are computed
+% - domains of original function is deduced
+% - decision tree for original function is built
+% - note how junk is removed in step 3 (but undefined "..." as last case)
+% 
+% \subsection{Support for Higher-Order Functions}
+% 
+% TODO:
+% - currying with explicit application symbol --> function arguments are now
+%   just constants
+% - apply encoding on application symbol
+% - gotcha: one function can be used with several arities.
+% 
+% - optim: specialization (similar to monomorphization)
 

### 2016-01-26 ijcar2016/body.tex
more edits

@@
 If a model is found, it is translated back to the input language,
-with \verb$?__$ placeholders as necessary.
+with \verb$?__$ placeholders indicating unknown values.
 
 Conceptually, the sequence of transformation is a \relax{two-way pipeline}
-built by composing pairs $(\textsf{encode}, \textsf{decode})$ of
+built by composing pairs $(\textsf{encode},\vthinspace \textsf{decode})$ of
 transformations. For each such pair, \textsf{encode} transforms a problem over
-a signature $\Sigma$ in a logic $\mathcal{L}$ to a problem over a signature
-$\Sigma'$ in a logic $\mathcal{L}'$, and \textsf{decode} translates a model in
-$\mathcal{L'}$ over $\Sigma'$ into a model in $\mathcal{L}$ over $\Sigma$,
+a signature $\Sigma$ in a logic~$\mathcal{L}$ to a problem over a signature
+$\Sigma'$ in a logic~$\mathcal{L}'$, and \textsf{decode} translates a model
+in~$\mathcal{L}'$ over $\Sigma'$ into a model in~$\mathcal{L}$ over $\Sigma$,
 in the spirit of institution theory \cite{goguen-burstall-1992}. The
-pipeline currently implemented in Nunchaku consists of the following phases:
+pipeline currently consists of the following phases:
 
@@
     introduces Skolem symbols for term variables;
-  \item[Elimination of Predicates]
-    changes a multiple-clauses (co)inductive predicate definition
-      into a single recursive function;
+  \item[Elimination of (co)inductive predicates]
+    recasts a multiple-clause (co)inductive predicate specification
+      into a recursive equation;
   \item[Recursion elimination]

### 2016-01-26 ijcar2016/body.tex
some work on eval

@@
 
-\newcommand\isanun{Isa\xspace}
-\newcommand\isa{Ip\xspace}
-\newcommand\isam{Ip-Mut\xspace}
+\newcommand\isanun{Nunchaku-Mut\xspace}
+\newcommand\ip{IsaPlanner\xspace}
+\newcommand\ipm{\ip-Mut\xspace}
 \newcommand\leon{Leon\xspace}
-\newcommand\leonm{Leon-Mut\xspace}
+\newcommand\leonm{\leon-Mut\xspace}
 
@@
 \label{sec:introduction}
-
-%\ct{The paper used "recursive datatype" in some places and "inductive datatype" in others. I have replaced both by "algebraic datatype" which is more accurate for \cvc.
-%("recursive datatype" is a more general notion and "inductive datatype" has a stronger meaning in higher-order logics.)}  
-%% I agree --JB.
 
@@
 several proof assistants. The first version, 0.1, was released in January 2016
-with support for (co)datatypes, (co)recursive functions and (co)inductive
-predicates. Support for higher-order functions is planned for the next
-release. We have developed a preliminary Isabelle frontend and are planning
-further frontends for Coq, the TLA\textsuperscript{+} Proof System, and
-possibly other systems.
+with support for (co)algebraic datatypes, (co)recursive functions and
+(co)inductive predicates. Support for higher-order functions is planned for
+the next release. We have developed a preliminary Isabelle frontend and are
+planning further frontends for Coq, the TLA\textsuperscript{+} Proof System,
+and possibly other systems.
 
@@
 Nunchaku translates its input to one first-order logic program that targets
-the finite model finding fragment of CVC4, including interpreted
-(co)datatypes \cite{reynolds-blanchette-2015-codata}.
+the finite model finding fragment of CVC4, including
+(co)algebraic datatypes \cite{reynolds-blanchette-2015-codata}.
 Using CVC4 also allows Nunchaku to provide efficient arithmetic
@@
 
-{\looseness=-1
 In this section, we evaluate both the overall impact of the translation
@@
 We gathered 602 benchmarks from three sources, which we will refer to as
-\isanun, \isa and \leon. 
-The first source consists of 357 benchmarks coming from\ajr{where Isa-Nun benchmarks come from, what they represent, are they expected to be unsat/sat}.
-The second source consists of the 79
-benchmarks from the IsaPlanner
+\ip, \leon, and \isanun:
+\begin{itemize}
+\item
+\ip{} consists of the 79 benchmarks from the IsaPlanner
 suite~\cite{DBLP:conf/itp/JohanssonDB10} that do not contain higher-order
@@
 with a signature that contains uninterpreted function symbols over these datatypes.
-The third source consists of 166
+
+\smallskip
+
+\item
+\leon{} consists of 166
 benchmarks from the Leon repository,\footnote{%Available at
@@
 but cover a wide variety of additional theories, including bit vectors, arrays, and
-both linear and nonlinear arithmetic. All benchmarks are in definitional form
-with respect to a set of well-founded functions.
-A majority of the benchmarks in the latter two sets are unsatisfiable.
-}
-
-For each of the 245 benchmarks in the \isa and \leon sets, 
+both linear and nonlinear arithmetic.
+
+\smallskip
+
+\item
+\isanun{} consists of 357 benchmarks originating from Isabelle/HOL. They
+involve (co)recursively defined functions over (co)algebraic datatypes and
+uninterpreted functions but no other theories. They were obtained by mutation
+of negated Isabelle theorems, as was done for evaluating
+Nitpick~\cite{blanchette-nipkow-2010}. Benchmarks created by mutation have a
+high likelihood of having small, easy-to-find models.
+\end{itemize}
+
+The \ip{} and \leon{} benchmarks are expressed in SMT-LIB 2.5 and are in
+definitional form with respect to a set of well-founded functions.
+The Leon tool was used to generate SMT-LIB files.
+A majority of these benchmarks are unsatisfiable.
+For each of the 245 benchmarks,
 we considered up to three randomly selected
-mutated forms of its conjecture $\psi$. In particular, we considered unique
-formulas that are obtained as a result of swapping a subterm of $\psi$ at
+mutated forms of its goal $\psi$. In particular, we considered unique
+formulas that are obtained as a result of exchanging a subterm of $\psi$ at
 one position with another of the same type at another position.
-Note that benchmarks created in this way have a high likelihood of having
-small, easy-to-find countermodels. In total, we considered 213 mutated forms of
-theorems from \isa and 427 mutated forms of theorems from \leon. We will
-call these sets \isam and \leonm, respectively.
-Thus, our benchmark set consists of 1242 benchmarks (640 mutants plus the
-original 602). 
-Each of the 1242 benchmarks were generated in SMT lib version 2.5 format~\cite{smtlib25}. %ajr : although these don't require 2.5 features
-For \isanun, \ajr{describe generation of smt2 benchmarks without encoding}
-The others were translated to SMT lib version 2 by the Leon tool.
-
-We additionally considered these 1242 benchmarks both before and after the translation~$\conv$.
-For example, \isa contains 79 original benchmarks $\varphi$ and 79 translated
-benchmarks $\conv(\varphi)$.
-For the \isanun, \ajr{describe generation of smt2 benchmarks with encoding}.
-All others were translated by \cvc 
-by outputting the result of running the preprocessor from Section~\ref{ssec:cvc4} on each benchmark.
-
-For solvers, we considered the SMT solver \ziii~\cite{de-moura-bjoerner-2008},
-which runs heuristic methods for quantifier instantiation~\cite{MouraBjoerner07}
+In total, we considered 213 mutated forms of
+theorems from \ip and 427 mutated forms of theorems from \leon. We will
+call these sets \ipm and \leonm, respectively. Each of these benchmarks exists
+in two versions:\ without and with the $\conv$~translation. Problems with
+$\conv$ were produced by running \cvc's preprocessor % (Section~\ref{ssec:cvc4})
+on each benchmark.
+
+For \isanun, the Isabelle Nunchaku frontend was used to generate thousands of
+Nunchaku problems from Isabelle/HOL theory files involving lists, trees, and
+other functional data structures. Nunchaku was then used to generate SMT-LIB
+files, again in two versions:\ without and with the $\conv$~translation.
+Problems requiring higher-order logic were discarted, since Nunchaku does not
+yet support them, leaving 357 problems.
+
+% Our entire benchmark set consists of 1242 benchmarks (640 mutants plus the
+% original 602).
+
+%Each of the 1242 benchmarks were generated in SMT lib version 2.5 format~\cite{smtlib25}. %ajr : although these don't require 2.5 features
+
+%We additionally considered these 1242 benchmarks both before and after the translation~$\conv$.
+%For example, \ip contains 79 original benchmarks $\varphi$ and 79 translated
+%benchmarks $\conv(\varphi)$.
+%For the \isanun, \ajr{describe generation of smt2 benchmarks with encoding}.
+
+Among SMT solvers, we considered \ziii~\cite{de-moura-bjoerner-2008} and
+\cvc~\cite{barrett-et-al-2011}.
+\ziii runs heuristic methods for quantifier instantiation~\cite{MouraBjoerner07}
 as well as methods for finding models for quantified formulas~\cite{GeDeM-CAV-09}.
-We also considered five configurations of \cvc~\cite{barrett-et-al-2011},
-which we will refer to as \cvcd, \cvcf, \cvcfe, and \cvcfm.
+For \cvc, we considered four configurations, referred
+to as \cvcd, \cvcf, \cvcfe, and \cvcfm.
 The configuration \cvcd runs heuristic 
@@
 
-\isanun  & 0 & 14 & 0 & 0 & 0 & 155 & 0 & {\win 156} & 0 & 155
+\ip & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0
 \\
-\isa & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0
-\\
-\isam & 0 & 43 & 0 & 0 & 0 & {\win 161} & 0 & {\win 161} & 0 & {\win 161}
+\ipm & 0 & 43 & 0 & 0 & 0 & {\win 161} & 0 & {\win 161} & 0 & {\win 161}
 \\
@@
 \leonm & 6 & 61 & 0 & 0 & 2 & 180 & 2 & {\win 181} & 2 & {\win 181}
+\\
+\isanun  & 0 & 14 & 0 & 0 & 0 & 155 & 0 & {\win 156} & 0 & 155
 \\[\jot]
@@
 \midrule
-\isanun & {\win 41} & 40 & 28 & 34 & 26 & 39 & 26 & 39 & 26 & 39
+\ip & 14 & {\win 15} & {\win 15} & {\win 15} & 1 & {\win 15} & {\win 15} & {\win 15} & 1 & {\win 15}
 \\
-\isa & 14 & {\win 15} & {\win 15} & {\win 15} & 1 & {\win 15} & {\win 15} & {\win 15} & 1 & {\win 15}
-\\
-\isam & 54 & 55 & {\win 56} & {\win 56} & 13 & 55 & {\win 56} & 55 & 13 & 55
+\ipm & 54 & 55 & {\win 56} & {\win 56} & 13 & 55 & {\win 56} & 55 & 13 & 55
 \\
@@
 \leonm & 94 & 107 & 106 & 107 & 39 & {\win 108} & 106 & {\win 108} & 39 & {\win 108}
+\\
+\isanun & {\win 41} & 40 & 28 & 34 & 26 & 39 & 26 & 39 & 26 & 39
 \\[\jot]
@@
 The results are summarized in Figures \ref{fig:sat} and \ref{fig:unsat}.
+Bold indicates the maximum of a row.
 The benchmarks and more detailed results are available online.%
@@
 The two optimizations for finite model finding in \cvc (configurations \cvcfe and \cvcfm) 
-led to a net gain of 1 satisfiable benchmark each with respect to \cvcf.
+led to a net gain of one~satisfiable benchmark each with respect to \cvcf.
 The performance of \ziiib for countermodels also improved dramatically, as it
@@
 the configuration \cvcf, which is not tailored for handling unsatisfiable benchmarks,
-solves 270 unsat benchmarks overall, which is more than both \cvcd and \ziii.
+solves 270 \emph{unsat} benchmarks overall, which is more than both \cvcd and \ziii.
 These results suggest that the translation do not degrade the performance 
 of SMT solvers for unsatisfiable problems involving recursive functions, 
-and instead often improve their performance.
-\ajr{Importance of the translation helping unsat in practice for Isabelle/Nunchaku, if any}
+and instead often improve their performance. It would be interesting to try this
+out in Sledgehammer \cite{blanchette-et-al-2013-smt} and to try Nunchaku as a proof tool.
 

### 2016-01-26 ijcar2016/body.tex
more

@@
 the cardinalities of the atomic types. This procedure was pioneered by McCune
-in the earlier versions of Mace (or MACE) \cite{mccune-1994}.
+in the earlier versions of Mace (originally styled MACE) \cite{mccune-1994}.
 Other conceptually similar finders are Paradox \cite{claessen-sorensson-2003}
@@
 In future work, it would be interesting to evaluate the approach against other
-counterexample generators, notably Leon and Nitpick, and enrich the benchmark
-suite with problems exercising \cvc's support for coalgebraic datatypes
-\cite{reynolds-blanchette-2015-codata}. We also plan to integrate \cvc as a
-counterexample generator in proof assistants.\ajr{Revise: this should say plans to develop Nunchaku.} Further work would also
-include identifying additional sufficient conditions for admissibility, thereby
-enlarging the applicability of the translation scheme presented here.
+counterexample generators, notably Leon, Nitpick, and Quickcheck,
+and enrich the benchmark suite with more problems exercising \cvc's support for
+coalgebraic datatypes \cite{reynolds-blanchette-2015-codata}.
+We are also working on an encoding of higher-order functions in SMT-LIB, as a
+generalization to the current translation scheme, for Nunchaku. Further work
+would also include identifying additional sufficient conditions for
+admissibility, thereby enlarging the applicability of the translation scheme
+presented here.
 
@@
 
-{%\footnotesize
+{\footnotesize
+
 \def\ackname{Acknowledgment}
@@
 development action ``Contre-exemples utilisables par Isabelle et Coq'' (CUIC).
+
 }

### 2016-01-26 ijcar2016/body.tex
Fixed numbers from misclassified benchmarks
removed refs to case studies

@@
 \newcommand\innerkw[1]{\textrm{#1}}
+\newcommand\keyw[1]{\texttt{#1}}
 
@@
 \newcommand{\J}{\mathcalx{J}}
-\def\AIF{\qtab\keyword{if}\ }
-\def\THEN{\ \keyword{then}\ }
-\def\AELSE{\untab\qtab\keyword{else}\ }
-\def\FI{\untab}
-\def\RETURN{\keyword{return}\ }
-\def\ENDPROC{\untab}
 \newcommand{\conv}{\mathcalx{A}}
@@
 The approach is implemented as a preprocessor in \cvc and in the Nunchaku
-model finder (Section~\ref{sec:implementations}). As a case study, we employed
-Nunchaku on Isabelle/HOL~\cite{nipkow-et-al-2002} formalizations of binary
-trees and coinductive lists (Section~\ref{sec:case-studies}).
+model finder (Section~\ref{sec:implementations}).
+%As a case study, we employed
+%Nunchaku on Isabelle/HOL~\cite{nipkow-et-al-2002} formalizations of
+%coinductive lists and binary trees (Section~\ref{sec:case-studies}).
 %
@@
 (Section~\ref{sec:related-work}).
+
+An earlier version of this paper was presented at the SMT 2015 workshop in San
+Francisco \cite{reynolds-et-al-2015-smt}. This paper extends the workshop
+paper with proof sketches, an expanded implementation section covering
+Nunchaku and relevant \cvc optimizations, %the case studies,
+and the evaluation on Isabelle benchmarks produced by Nunchaku.
 
@@
 planning further frontends for Coq, the TLA\textsuperscript{+} Proof System,
-and possibly other systems.
+and possibly other proof assistants.
 
@@
 
-\section{Case Studies}
-\label{sec:case-studies}
-
-\subsection{Lazy Lists}
-
-XXX
-
-\subsection{Huffman Coding}
-
-XXX
+%\section{Case Studies}
+%\label{sec:case-studies}
+%
+%We show Nunchaku in action on two small Isabelle formalizations:\ coinductive
+%lists and a binary data structure called AA trees. These two examples were
+%used before to demonstrate Nitpick \cite{}.
+%
+%\subsection{Coinductive Lists}
+%
+%TODO: Adapt!
+%
+%\newcommand\List[1]{[#1]}
+%\newcommand\Nil{\List{\vthinspace}}
+%\newcommand\bool{\mathit{bool}}
+%
+%The Isabelle codatatype $\alpha\>\mathit{llist}$ of lazy lists is generated by the
+%constructors $\const{LNil} : {\alpha\>\mathit{llist}}$ and $\const{LCons} : {\alpha\to
+%\alpha\>\mathit{llist} \to \alpha\>\mathit{llist}}$. It is of particular interest to countermodel finding
+%because many %basic
+%properties of finite lists do not carry over to infinite
+%lists, often in baffling ways. To illustrate this, we conjecture that
+%appending $\textit{ys}$ to $\textit{xs}$ yields $\textit{xs}$ iff
+%$\textit{ys}$ is $\const{LNil}$:
+%\[\textit{xs} \mathbin{@} \textit{ys} = \textit{xs} \,\longleftrightarrow\,
+%  \textit{ys} = \const{LNil}\]
+%The operator $\mathbin{@}$ is defined by primitive corecursion by the equation
+%\[\textit{xs} \mathbin{@} \textit{ys} =
+%  (\innerkw{case}~\textit{xs}~\innerkw{of}~\const{LNil} \Rightarrow \textit{ys}
+%  \mid \const{LCons}\;x\;\textit{xs}' \Rightarrow \const{LCons}\;x\;(\textit{xs}' \mathbin{@} \textit{ys}))\]
+%Nunchaku immediately
+%finds the countermodel $\textit{xs} = \textit{ys} =
+%\List{0, 0, \ldots}$, in which a cardinality of 1 is sufficient for
+%$\alpha$ and $\alpha\>\mathit{llist}$, and the bisimilarity predicate $\sim$ is unrolled only
+%once. Indeed, appending $\List{0, 0, \ldots} \not= \Nil{}$ to $\List{0, 0, \ldots}$ leaves
+%$\List{0, 0, \ldots}$ unchanged. Many other counterexamples are possible---for example,
+%$\textit{xs} = \List{0, 0, \ldots}$ and $\textit{ys} = \List{1}$---but Nitpick
+%tends to reuse the objects that are part of a subterm-closed substructure to
+%keep cardinalities low. Although very simple, the counterexample is beyond
+%Quickcheck's and Refute's reach, since they do not support codatatypes.
+%
+%The next example requires the following lexicographic order predicate:
+%%
+%\begin{quote}
+%$\keyw{coinductive}~\vthinspace\smash{{\preceq}\typ{\mathit{nat}\>\mathit{llist}\to\mathit{nat}\>\mathit{llist}\to\bool}}~\keyw{where}$ \\
+%$\const{LNil} \preceq \textit{xs}$ \\
+%$x \le y \,\Longrightarrow\, \const{LCons}~x~\textit{xs} \preceq \const{LCons}~y~\textit{ys}$ \\
+%$\textit{xs} \preceq \textit{ys} \,\Longrightarrow\, \const{LCons}~x~\textit{xs} \preceq \const{LCons}~x~\textit{ys}$
+%\end{quote}
+%%
+%The intention is to define a linear order on lazy lists of
+%natural numbers, and hence the following properties should hold:
+%%
+%\begin{align*}
+%\textsc{Refl}{:}\kern.75em
+%  & \textit{xs} \preceq \textit{xs}
+%& \textsc{Antisym}{:}\kern.75em
+%  & \textit{xs} \preceq \textit{ys} \mathrel{\land} \textit{ys} \preceq \textit{xs} \longrightarrow \textit{xs} = \textit{ys} \\
+%\textsc{Linear}{:}\kern.75em
+%  & \textit{xs} \preceq \textit{ys} \mathrel{\lor} \textit{ys} \preceq \textit{xs}
+%& \textsc{Trans}{:}\kern.75em
+%  & \textit{xs} \preceq \textit{ys} \mathrel{\land} \textit{ys} \preceq \textit{zs} \longrightarrow \textit{xs} \preceq \textit{zs}%.
+%\end{align*}
+%%
+%However, Nitpick finds a counterexample for \textsc{Antisym}: $\textit{xs} = \List{1,
+%1}$ and $\textit{ys} = \List{1}$. On closer inspection, the assumption $x \leq y$ of
+%the second introduction rule for $\preceq$ should have been $x < y$; otherwise,
+%any two lists \textit{xs}, \textit{ys} with the same head satisfy
+%$\textit{xs} \preceq \textit{ys}$. Once we repair the specification, no more
+%counterexamples are found for the four properties up to cardinality 6
+%for $\textit{nat}$ and $\mathit{nat}\>\mathit{llist}$ within the time limit of 30~seconds.
+%%This is a strong indication that the properties hold.
+%Andreas Lochbihler used Isabelle
+%to prove all four properties \cite{lochbihler-2010-coinduction}.
+%
+%We could continue like this and sketch a full-blown theory of lazy lists. Once the
+%definitions and main theorems %are in place and
+%have been thoroughly tested using
+%Nitpick, we could start working on the proofs. Developing theories this way
+%can save a lot of time, because faulty theorems and definitions are discovered
+%early. %in the process.
+%
+%\subsection{AA Trees}
 

### 2016-01-26 ijcar2016/body.tex
Minor fix

@@
+%!TEX root =  pap.tex
+
 %%% TODO: mention that admissible does not mean finite model
@@
 \newcommand{\M}{\mathcalx{M}}
-\newcommand{\I}{\mathcalx{I}}
-\newcommand{\J}{\mathcalx{J}}
+%CT somehow this compiles into a Roman font, removed for now
+%\newcommand{\I}{\mathcalx{I}}
+%\newcommand{\J}{\mathcalx{J}}
+\newcommand{\I}{\mathcal{I}}
+\newcommand{\J}{{\mathcal{I}'}}
 \newcommand{\conv}{\mathcalx{A}}
@@
 \hyphenation{counter-example counter-examples Isa-belle data-type data-types
-  co-data-type co-data-types}
+  co-data-type co-data-types bench-mark bench-marks}
 
@@
 SMT solvers have recently been extended with techniques for finding models
-in presence of universally quantified formulas in some restricted fragments.
+in the presence of universally quantified formulas in some restricted fragments
+of first-order logic.
 This paper introduces a translation which reduces axioms specifying a large
@@
 \cite{MouraBjoerner07,ReynoldsTinelliMoura14}.
-These methods work well in the context of proving (i.e., showing
-unsatisfiability), but they are of little help for finding models (i.e.,
-showing satisfiability). Often, a single universal quantifier in one of the
+These methods work well for proving the
+unsatisfiability of an input set of formulas, 
+but they are of little help for finding models of them
+when they are satisfiable. Often, a single universal quantifier in one of the
 axioms of a problem is enough to prevent the discovery of models.
 
-In the past few years, techniques have been developed to find models for
+In the past few years, a few techniques have been developed to find models for
 quantified formulas in SMT.
 Ge and de Moura \cite{GeDeM-CAV-09} introduced a complete instantiation-based
-procedure for formulas in the \relax{essentially uninterpreted} fragment.
+procedure for formulas in the \emph{essentially uninterpreted} fragment.
 This fragment is limited to universally quantified formulas where all
@@
 variables occur as direct subterms of uninterpreted
-functions---e.g., $\forall x%\typ{\typeint} --- the type is not important, right?
-.\;\, \con{f}( x )
-\teq \con{g}( x ) + 5$.
+functions, as in $\forall x : \typeint.\; \con{f}( x ) \teq \con{g}( x ) + 5$.
 % since all
@@
 this fragment slightly, including cases when variables occur as arguments of
-arithmetic predicates. Subsequently, Reynolds et al.\
+arithmetic predicate symbols. Subsequently, Reynolds et al.\
 \cite{ReyEtAl-1-RR-13,reynolds-et-al-2013} introduced techniques for finding finite
@@
 essentially uninterpreted fragment does not allow the argument of a
-recursive function to be used inside a complex term on the right-hand side,
+recursive function to be used inside a complex term on the right-hand side
+of the definition,
 whereas the finite model finding techniques %of Reynolds et al.\
@@
   2 * \con{p}( x - 1 ) \bigr)$$ %.
+where $\lite$ is the `if--then--else' operator.
 This state of affairs is unsatisfactory, given the frequency of
-recursive definitions in practice and the %impending
+recursive definitions in practice, underscored by the recent
 addition of a %dedicated
@@
 definitions into formulas where finite model finding techniques can be applied.
-The recursive functions must meet a semantic criterion to be admissible
-(Section~\ref{sec:prelim}). This criterion is met by well-founded (terminating)
-recursive function definitions. It is not met by inconsistent
-definitions such as \,$\forall x : {\typeint}.\;\, \con{f}(x) \teq \con{f}(x) +
+The definitions must meet a semantic criterion to be admissible 
+(Section~\ref{sec:prelim}). 
+This criterion is general enough to include well-founded (terminating)
+recursive function definitions and restrictive enough to exclude inconsistent
+equations such as \,$\forall x : {\typeint}.\;\, \con{f}(x) \teq \con{f}(x) +
 1$.
@@
 (Section~\ref{sec:encoding}). A
-recursive definition $\forall x : \tau.\;\, \const{f}(x) \teq \mathit{rhs}$
+recursive equation $\forall x : \tau.\;\, \const{f}(x) \teq t$
 is translated to
-$\forall a :
-\fargtype{\tau}.\;\, \const{f}(\farg{\con{f}}(a)) \teq \mathit{rhs}[\farg{\con{f}}(a)/x]$, where
-$\fargtype{\tau}$ is an uninterpreted \relax{abstract} type and $\farg{\con{f}} :
-\fargtype{\tau} \to \tau$ converts the abstract type to the concrete
-type. Additional constraints ensure that the abstract values that are relevant
+$\forall a :\nobreak
+\fargtype{\tau}.\allowbreak\;\, \const{f}(\farg{\con{f}}(a)) \teq t[\farg{\con{f}}(a)/x]$, where
+%%% To me, "abstract" and "concrete" are just convenient labels to refer to the
+%%% thing. I don't want readers to start wondering, "What is an abstract type?"
+%%% So I find it important to leave "uninterpreted" outside of parentheses. As
+%%% for italics, I went for quotes instead, since this is an informal explanation
+%%% anyway. Later, we define \emph{abstract type} with italics. --JB
+$\fargtype{\tau}$ is an ``abstract'' uninterpreted type and $\farg{\con{f}} :
+\fargtype{\tau} \to \tau$ is an uninterpreted function
+from $\fargtype{\tau}$ to the corresponding concrete type $\tau$. 
+Additional constraints ensure that the abstract values that are relevant
 to the formula's satisfiability
@@
 %
-We evaluated the two evaluation on benchmarks from
+We evaluated the two implementations on benchmarks from
 IsaPlanner~\cite{DBLP:conf/itp/JohanssonDB10}, Leon~\cite{blanc2013overview},
 and Isabelle/HOL, to demonstrate that this translation improves the
-effectiveness of the SMT solvers \cvc and \ziii for finding countermodels to
+effectiveness of the SMT solvers \cvc and \ziii in finding countermodels to
 verification conditions (Section~\ref{sec:evaluation}). Unlike earlier work,
-our approach can be combined with off-the-shelf SMT solvers
+our approach relies on off-the-shelf SMT solvers
 (Section~\ref{sec:related-work}).
@@
 define $\svars{\Sigma}$ as $\bigcup_{\tau \in \stypes{\Sigma}} \svars{\Sigma}_\tau$.
-$\Sigma$-terms are built as usual over functions symbols in $\Sigma$ and variables in $\svars{\Sigma}$.
+(Well-typed) $\Sigma$-terms are built as usual over functions symbols in $\Sigma$ and variables in $\svars{\Sigma}$.
 %
@@
 terms of type~$\tau$ and $\terms( t )$ to denote the set of subterms in $t$.
-Given a term $u$, we write $u[\vec t/\vec x]$ to denote the result of replacing
-all occurrences of $\vec x$ with $\vec t$ in $u$.
+Given a term $u$, a variable tuple $\vec x = (x_1^{\tau_1},\dotsc,x_n^{\tau_n})$ and 
+a term tuple $\vec t = (t_{\smash{1}\vphantom{n}}^{\tau_1},\dotsc,t_n^{\tau_n})$, 
+we write $u[\vec t/\vec x]$ to denote the result of simultaneously replacing
+all occurrences of $x_i$ with $t_i$ in $u$, for each $i=1,\dotsc,n$.
 \begin{rep}When applied to terms, the symbol $=$ denotes syntactic equality.\end{rep}
@@
 each function symbol $\con{f} : \tau_1 \times \cdots \times \tau_n \rightarrow \tau$ in $\sfuns{\Sigma}$
-to a total function $\con{f}^\I : \tau_1^\I \times \cdots \times \tau_n^\I \rightarrow \tau^\I$,
+to a total function $\con{f}^\I : \tau_{\smash{1}}^\I \times \cdots \times \tau_n^\I \rightarrow \tau^\I$,
 and each variable $x:\tau$ of $\svars{\Sigma}$ to an element of $\tau^\I$.
@@
 (\,\emph{for $\con{f}$}\vthinspace) if $\vec x$ is a tuple of variables
-$x_1 : \tau_1$, $\ldots,$ $x_n : \tau_n$
+$x_1^{\tau_1}$, $\dotsc,$ $x_n^{\tau_n}$
 and $\varphi$ is a quantifier-free formula
@@
 A formula $\varphi$ is in \emph{definitional form with respect to}
-$\{ \con{f}_1, \ldots, \con{f}_n \} \subseteq \sfuns{\Sigma}$ if it is of the
+$\{ \con{f}_1, \dotsc, \con{f}_n \} \subseteq \sfuns{\Sigma}$ if it is of the
 form
@@
 %
-where $\con{f}_1, \ldots, \con{f}_n$ are distinct function symbols,
+where $\con{f}_1, \dotsc, \con{f}_n$ are distinct function symbols,
 %ct added
 $\forallf{\con{f}_i} \vec x_i.\; \varphi_i$ is a function definition
-for $i = 1, \ldots, n$,
+for $i = 1, \dotsc, n$,
 %
@@
 Given a set of function definitions
-$\DDD = \{ \forallf{\con{f}_1} \vec x.\; \varphi_1, \ldots, \forallf{\con{f}_n} \vec x.\; \varphi_n \}$, 
+$\DDD = \{ \forallf{\con{f}_1} \vec x_1.\; \varphi_1, \dotsc, \forallf{\con{f}_n} \vec x_n.\; \varphi_n \}$,
 a ground formula $\psi$
@@
 is \emph{closed under function expansion with respect to $\DDD$} if
-\[\psi \models_T \smash{\bigwedge\nolimits_{\,i=1}^{n} \{ \varphi_i[ \vec t / \vec x ] \mid \con{f}_i( \vec t ) \in \terms( \psi ) \}}\]
+\[\psi \models_T \smash{\bigwedge\nolimits_{\,i=1}^{n} \{ \varphi_i[ \vec t / \vec x_i ] \mid \con{f}_i( \vec t ) \in \terms( \psi ) \}}\]
 The set $\DDD$ is \emph{admissible} if for every $T$-satisfiable formula
@@
 Admissibility is a semantic criterion that must be satisfied for each function
-definition before applying the translation described in
+definition before applying our translation, described in
 Section~\ref{sec:encoding}. It is useful to connect it to the
@@
 that a function specification is well founded, it is sufficient to require that
-the function would terminate when seen as a functional program, under \emph{some}
+the defined function be terminating when seen as a functional program, under \emph{some}
 evaluation order. For example, the definition
@@
   2 * \con{p}( x - 1 ) \bigr)$,
-where $T$ is integer arithmetic extended with the uninterpreted symbol 
-$\con p:\typeint \rightarrow \typeint$, can be shown well founded under a strategy that
+where the theory $T$ is integer arithmetic extended with the uninterpreted symbol
+$\con p:\typeint \rightarrow \typeint$, can be shown to be well founded under a strategy that
 evaluates the condition of an $\lite$ before evaluating the relevant branch,
@@
 If $\DDD$ is a set of well-founded function definitions for\/
-$\sfundefs{\Sigma} = \{\con{f}_1,\ldots,\con{f}_n\}$, then it is admissible.
+$\sfundefs{\Sigma} = \{\con{f}_1,\dotsc,\con{f}_n\}$, then it is admissible.
 \end{theorem}
+
 \begin{proofsketch}
@@
 satisfiable. Let $\I$ be a model of $\psi$, and
-let $\I_0$ be the restriction of $\I$ to the function symbols in $\sfuns{\Sigma} -
+let $\I_0$ be the reduct of $\I$ to the function symbols in $\sfuns{\Sigma} \setminus
 \sfundefs{\Sigma}$. Because well-founded definitions uniquely characterize
@@
 
-Another useful fragment of function definitions is the class of
+Another useful class of function definitions is that of
 \emph{productive} corecursive functions. Corecursive functions are functions to
 a coalgebraic datatype. These functions can be ill founded, without
-their being inconsistent. Productive corecursive functions are functions that
+their being inconsistent. 
+Intuitively, productive corecursive functions are functions that
 progressively reveal parts of their potentially infinite
@@
 $\con{scons} : \ty{int} \times \ty{stream} \to \ty{stream}$,
-the definition
+the function defined by
 $\forallf{\const{e}} x.\;\, \const{e}(x) \teq \con{scons}(x,\; \const{e}(x + 1))$
-falls within this fragment: Each call to $\const{e}$ produces one
+%%% If what follows ':' is a complete sentence, the standard Am.E. convention
+%%% is to capitalize it. --JB
+falls within this class: Each call to $\const{e}$ produces one
 constructor before entering the nested call. Like terminating recursion,
-productive corecursion totally specify the functions it defines.
+productive corecursion totally specifies the functions it defines.
 It is even possible to mix recursion and corecursion in the same function
@@
  \begin{itemize}
-   \itemx[] $\innerkw{if}$ $\tau = \Bool$ $\mathsf{and}$ $t = \boolop(t_1,\ldots,t_n)$ $\innerkw{then}$
+   \itemx[] $\innerkw{if}$ $\tau = \Bool$ $\mathsf{and}$ $t = \boolop(t_1,\dotsc,t_n)$ $\innerkw{then}$
     \begin{itemize}
-      \itemx[] $\innerkw{let}$ $( t'_i,\, \chi_i ) = \conv_0( t_i,\, \pol( \boolop,\, i,\, p ) )$ $\mathsf{for}$ $i = 1, \ldots, n$ $\innerkw{in}$%\kern6.35em$\mid$
+      \itemx[] $\innerkw{let}$ $( t'_i,\, \chi_i ) = \conv_0( t_i,\, \pol( \boolop,\, i,\, p ) )$ $\mathsf{for}$ $i = 1, \dotsc, n$ $\innerkw{in}$%\kern6.35em$\mid$
       \itemx[] $\innerkw{let}$ $\chi = \chi_1 \wedge \cdots \wedge \chi_n$ $\innerkw{in}$
-      \itemx[] $\innerkw{if}$ $p = \ppos$ $\innerkw{then}$ \bigtuple{$\boolop(t'_1, \ldots, t'_n) ) \wedge \chi,\, \ltrue$}
-      \itemx[] $\innerkw{else}$ $\innerkw{if}$ $p = \pneg$ $\innerkw{then}$ \bigtuple{$\boolop(t'_1, \ldots, t'_n) \vee \neg \chi,\, \ltrue$}
-      \itemx[] $\innerkw{else}$ \bigtuple{$\boolop(t'_1, \ldots, t'_n),\, \chi$}
+      \itemx[] $\innerkw{if}$ $p = \ppos$ $\innerkw{then}$ \bigtuple{$\boolop(t'_1, \dotsc, t'_n) ) \wedge \chi,\, \ltrue$}
+      \itemx[] $\innerkw{else}$ $\innerkw{if}$ $p = \pneg$ $\innerkw{then}$ \bigtuple{$\boolop(t'_1, \dotsc, t'_n) \vee \neg \chi,\, \ltrue$}
+      \itemx[] $\innerkw{else}$ \bigtuple{$\boolop(t'_1, \dotsc, t'_n),\, \chi$}
 %\\[-.8\baselineskip]
@@
 $\varphi$ and $\varphi'$ are \emph{equisatisfiable} (\emph{in $T$}).
-The idea behind this translation
-is to use an uninterpreted type to abstract the set of
-\emph{relevant} tuples for each defined function $\con{f}$ and restrict the
-quantification of $\con{f}$'s definition to a variable of this
-type. 
-Informally, the relevant tuples $\vec t$ of a function $\con{f}$ are the
+The idea behind this translation is to use an uninterpreted type $\alpha_\con{f}$
+to abstract the set of
+\emph{relevant} input tuples for each defined function $\con{f}$ and restrict the
+quantification of $\con{f}$'s definition to a single variable of this type.
+Informally, the relevant input tuples $\vec t$ of a function $\con{f}$ are the
 ones for which the interpretation of $\con{f}( \vec t )$ is relevant to the
@@
 %\item
-an uninterpreted \emph{abstract type} $\fargtype{\con{f}}$ and
+an uninterpreted \emph{abstract type} $\fargtype{\con{f}}$ 
+abstracting the Cartesian product $\tau_1 \times \cdots \times \tau_n$
+and
 %\item
-$n$ uninterpreted \emph{concretization functions} $\fargx{\con{f}}{1} : \fargtype{\con{f}} \rightarrow \tau_1$, \ldots, $\fargx{\con{f}}{n} : \fargtype{\con{f}} \rightarrow \tau_n$.
+$n$ uninterpreted \emph{concretization functions} $\fargx{\con{f}}{1} : \fargtype{\con{f}} \rightarrow \tau_1$, $\dotsc,$ $\fargx{\con{f}}{n} : \fargtype{\con{f}} \rightarrow \tau_n$.
 %\end{itemize}
@@
 $\bigwedge_{i=1}^n\, \fargx{\con{f}}{i}(a) \teq s_i$
-with $\vec s = (s_1,\ldots,s_n)$.
+with $\vec s = (s_1,\dotsc,s_n)$.
 These existential %%% for typesetting
@@
 that are relevant to the satisfiability of $\varphi$. The constraints are
-generated as deep in the formula as possible, based on our knowledge of the
-polarities of logical connectives, to allow models in which the domains
-interpreting the $\fargtype{\con{f}}$ types are as small as possible.
+generated as deep in the formula as possible, based on the
+polarities of Boolean connectives, to allow models where the sets
+denoted by the $\fargtype{\con{f}}$ types are as small as possible.
 
@@
 if $p$ is $\pneg$, and $\pnone$ if $p$ is $\pnone$.
-The term $t$ is then reconstructed as $\boolop(t_1',\ldots,t_n')$ 
+The term $t$ is then reconstructed as $\boolop(t_1',\dotsc,t_n')$
 where each $t_i'$ is the result of the recursive call with argument~$t_i$.
 If the polarity $p$ associated with $t$ is $\ppos$,
-$\conv_0$ conjunctively adds to $\boolop(t_1',\ldots,t_n')$ the constraint $\chi$ derived from
+$\conv_0$ conjunctively adds to $\boolop(t_1',\dotsc,t_n')$ the constraint $\chi$ derived from
 the subterms and returns $\top$ as the constraint.
@@
 a complete instantiation procedure, implemented in \ziii.
-As expected,
-%%% Are we sure?  
-% Yes, I'm planning to post the 2 versions of the benchmark on the evaluation page (when we add it) --ajr
-% Yes, but how can you be sure that Z3 will run forever? Your benchmarks surely have a timeout! --jb
-\cvc and \ziii run indefinitely on formula~(\ref{eq:ex-before}), 
-whereas they produce a model for~(\ref{eq:ex-after}) 
+As expected, \cvc and \ziii run indefinitely on formula~(\ref{eq:ex-before}).
+However, they produce a model for~(\ref{eq:ex-after})
 within 100 milliseconds.\xend
@@
     \lite( x \teq 2,\allowbreak\; 3,\allowbreak\;
-      \lite(\ldots,\allowbreak\; \lite( x \teq 13,\allowbreak\, 91,\allowbreak\, 105 )\negvthinspace \ldots ))))}
+      \lite(\dotsc,\allowbreak\; \lite( x \teq 13,\allowbreak\, 91,\allowbreak\, 105 )\negvthinspace \ldots ))))}
 
@@
 One model $\I$ for formula~(\ref{eq:ex-after}) in the above example interprets
-$\fargtype{\con{s}}$ as a finite set $\{ u_0, \ldots, u_{14} \}$,
-$\farg{\con{s}}$ as a finite map $u_i \mapsto i$ for $i = 0, \ldots, 14$,
+$\fargtype{\con{s}}$ as a finite set $\{ u_0, \dotsc, u_{14} \}$,
+$\farg{\con{s}}$ as a finite map $u_i \mapsto i$ for $i = 0, \dotsc, 14$,
 $\con{c}$ as $14$,
@@
 $\sfundefs{\Sigma}$ are admissible, %we claim that
-$\conv(\varphi)$ is equisatisfiable with $\varphi$ for any input $\varphi$.
+$\conv(\varphi)$ is equisatisfiable with $\varphi$ for any $\varphi$.
 %\footnote{
@@
 coincides with it on its interpretation of all
-function symbols in $\sfuns{\Sigma} - \sfundefs{\Sigma}$.
+function symbols in $\sfuns{\Sigma} \setminus \sfundefs{\Sigma}$.
 %($\{\con{s} \}$ in the example
@@
 to the domains of the corresponding abstract types $\fargtype{\con{f}}$.
-This can sometimes help users debug their specifications.
+This can sometimes help users debug their
+function definitions.
 
@@
 Formally,
-\[\absconstraints( L ) \vthinspace=\vthinspace \smash{\{ \exists a : {\fargtype{\con{f}}}.\;\, \vecfarg{\con{f}}( a ) \teq \nobreak \vec t
+\begin{equation} \label{eq:xl}
+\absconstraints( L ) \vthinspace=\vthinspace \smash{\{ \exists a : {\fargtype{\con{f}}}.\;\, \vecfarg{\con{f}}( a ) \teq \nobreak \vec t
 \nobreak\mid\allowbreak \con{f}( \vec t ) \in \terms( L ),\: \con{f} \in \sfundefs{\Sigma}
-\}}\]
+\}}
+\end{equation}
+
 The following lemma states the central invariant behind the translation $\conv$.
@@
 Let $\psi$ be a formula not containing function definitions,
-and let $\I$ be an $\extendsig{\Sigma}$-interpretation.
+and let $\I$ be an $\extendsig{\Sigma}$-interpretation whose $\Sigma$-reduct
+is a model of $T$.
 Then $\I$ satisfies $\conv( \psi )$ if and only if
-$\I$ satisfies $L \cup \absconstraints( L )$, where $L$ is a set of ground $\Sigma$-literals that entail $\psi$.
+it satisfies $L \cup \absconstraints( L )$, where $L$ is a set of ground $\Sigma$-literals that $T$-entail $\psi$.
 %There exists a set of ground $\Sigma$-literals $L$ such that $L$ entails $\varphi$,
@@
 
-\begin{corollary}\label{cor:conv}
-If $\psi$ is a formula not containing function definitions, then $\conv( \psi )$ entails $\psi$.
-\end{corollary}
+\begin{lemma}\label{lem:conv}
+If $\psi$ is a formula not containing function definitions, 
+then $\conv( \psi ) \models_T\psi$.
+\end{lemma}
 
@@
 interprets $\fargtype{\con{f}}$ as $\tau_1^\I \times \cdots \times \tau_n^\I$
-and each $\fargx{\con{f}}{i}$ as the $i$\vvthinspace{th} projection on such tuples for $i = 1, \ldots, n$.
+and each $\fargx{\con{f}}{i}$ as the $i$\vvthinspace{th} projection on such tuples for $i = 1, \dotsc, n$.
 Since $\J$ satisfies $\varphi$, it satisfies a set of ground literals $L$ that entail $\varphi$.
@@
 For any function symbol~$\con{f}$ and terms~$\vec t$, since $\varphi_{\con{f}}[ \vec t / \vec x ]$ does not contain function definitions,
-by Corollary~\ref{cor:conv},
+by Lemma~\ref{lem:conv},
 $\conv( \varphi_{\con{f}}[ \vec t / \vec x ] )$ entails $\varphi_{\con{f}}[ \vec t / \vec x ]$.
@@
 an $\extendsig{\Sigma}$-interpretation that satisfies $\conv( \varphi )$, by
-interpreting the abstract types in the same way as the cartesian products they
+interpreting the abstract types in the same way as the Cartesian products they
 abstract, thereby satisfying all existential constraints introduced by $\conv$.
@@
 that entails the goal of $\varphi$.
-This implies the existence of a model for $\varphi$,
-assuming $\DDD$ is admissible.
+This implies the existence of a model for $\varphi$
+provided that $\DDD$ is admissible.
 %For clarity, we demonstrate the consequences of this theorem by means of an example.
 
-\begin{conf}
-We give an intuition of the above theorem in the context of an example.
-For a set of ground $\Sigma$-literals $L$, let $\absconstraints( L )$ be the set
-$\{ \exists a : {\fargtype{\con{f}}}.\;\, \vecfarg{\con{f}}( a ) \teq \vec s
-\mid \con{f}( \vec s ) \in \terms( L ),\: \con{f} \in \sfundefs{\Sigma}
-\}$.
-\end{conf}
-
+%\begin{conf}
+We give an intuition of Theorem~\ref{thm:equi} in the context of an example.
+%For a set of ground $\Sigma$-literals $L$, let $\absconstraints( L )$ be the set
+%$\{ \exists a : {\fargtype{\con{f}}}.\;\, \vecfarg{\con{f}}( a ) \teq \vec s
+%\mid \con{f}( \vec s ) \in \terms( L ),\: \con{f} \in \sfundefs{\Sigma}
+%\}$.
+%\end{conf}
+%
 \begin{example}
@@
 where $L_0 = \{ \con{s}( \con{c} ) > 100 \}$
-and $\absconstraints( L_0 )$ consists of the single constraint
+and $\absconstraints( L_0 )$, defined by equation (\ref{eq:xl}), consists of the single constraint
 $\exists a : {\fargtype{\con{s}}}.\;\, \farg{\con{s}}( a ) \teq \con{c}$.
@@
 This formula is closed under function expansion
-since it entails $\varphi_\con{s}[(\con{c}-i)/x]$ for $i = 0, \ldots, n$,
-and it contains only $\con{s}$ applications corresponding to $\con{s}( \con{c}-i )$ for $i = 0, \ldots, n$.
+since it entails $\varphi_\con{s}[(\con{c}-i)/x]$ for $i = 0, \dotsc, n$
+and contains only $\con{s}$ applications corresponding to $\con{s}( \con{c}-i )$ for $i = 0, \dotsc, n$.
 Since $\{\forallf{\con{s}} x.\;\, \varphi_\con{s}\}$ is admissible,
@@
 and in the \cvc-based higher-order model finder Nunchaku.
-This section describes how the translation is implemented in these systems,
-as well as optimizations used by \cvc for finding models of translated
+This section describes how the translation is implemented in each system,
+as well as optimizations used by \cvc to find models of translated
 problems. % for the translated benchmarks.
@@
 When reading this input,
-\begin{rep}%
 \cvc adds the annotated quantified formula
@@
 $$\forallf{\con{s}} x.\;\, \lite\bigl( x \leq 0,\; \con{s}( x ) \teq 0,\; \con{s}( x ) \teq \con{s}( x-1 ) \bigr)$$
-\end{rep}%
-\begin{conf}%
-\cvc adds
-$\forallf{\con{s}} x.\;\, \con{s}( x ) \teq \lite\bigl( x \leq 0,\; 0,\; \con{s}( x-1 )\bigr)$
-to its list of assertions,
-which after rewriting becomes
-$\forallf{\con{s}} x.\;\, \lite\bigl( x \leq 0,\; \con{s}( x ) \teq 0,\; \con{s}( x ) \teq \con{s}( x-1 ) \bigr)$. \end{conf}%
 By specifying the command-line option \texttt{-}\texttt{-fmf-fun}, users
@@
 model of~$\con{s}$ where only
-the values of $\con{s}(x)$ for $x = 0,\ldots,14$ are correctly given:
+the values of $\con{s}(x)$ for $x = 0,\dotsc,14$ are correctly given:
 \begin{quote}
@@
 functions introduced using \texttt{define-\allowbreak fun-\allowbreak rec} are admissible.
-Admissibility must be discharged separately by the user---e.g., using a
-syntactic criterion or a termination prover.
+Admissibility must be proved externally by the user---e.g., manually, using a
+syntactic criterion, or with the help of a termination prover.
 If some function definitions are not admissible, \cvc may answer \emph{sat} for
-an unsatisfiable problem. 
-\begin{rep}
-Indeed, if we add the inconsistent definition
+an unsatisfiable problem. For example, if we add the inconsistent definition
 %
@@
 %
-to the above problem and run \cvc with the \texttt{-}\texttt{-fmf-fun} option,
-it wrongly answers \emph{sat}.
-\end{rep}
+to the above problem and run \cvc with \texttt{-}\texttt{-fmf-fun},
+it wrongly answers \emph{sat.}
 
@@
 monotonic types simultaneously.
-That is, if $\tau_1$ and $\tau_2$ are inferred monotonic (whether they are
-present in the original problem or introduced by our translation),
+That is, if $\tau_1$ and $\tau_2$ are inferred to be monotonic (regardless
+of whether they are present in the original problem or introduced by our translation),
 the solver fixes the bound for \relax{both} types to be $1$, then $2$, and so on.
@@
 negative positions),
-$m$ is skolemized into a fresh constants,
+%$m$ is skolemized into a fresh constant,
 then translated into admissible recursive functions,
@@
 \item
-\ip{} consists of the 79 benchmarks from the IsaPlanner
+The \ip{} set consists of the 79 benchmarks from the IsaPlanner
 suite~\cite{DBLP:conf/itp/JohanssonDB10} that do not contain higher-order
@@
 \item
-\leon{} consists of 166
+The \leon{} set consists of 166
 benchmarks from the Leon repository,\footnote{%Available at
 \url{https://github.com/epfl-lara/leon/}} which were constructed from
-verification conditions about simple Scala programs. These benchmarks also
+verification conditions on simple Scala programs. These benchmarks also
 heavily involve recursively defined functions over algebraic datatypes,
@@
 \item
-\isanun{} consists of 357 benchmarks originating from Isabelle/HOL. They
+The \isanun{} set consists of 357 benchmarks originating from Isabelle/HOL. They
 involve (co)recursively defined functions over (co)algebraic datatypes and
@@
 of negated Isabelle theorems, as was done for evaluating
-Nitpick~\cite{blanchette-nipkow-2010}. Benchmarks created by mutation have a
+Nitpick~\cite{blanchette-nipkow-2010}. 
+In our experience, benchmarks created by mutation have a
 high likelihood of having small, easy-to-find models.
@@
 call these sets \ipm and \leonm, respectively. Each of these benchmarks exists
-in two versions:\ without and with the $\conv$~translation. Problems with
+in two versions:\ with and without the $\conv$~translation. Problems with
 $\conv$ were produced by running \cvc's preprocessor % (Section~\ref{ssec:cvc4})
@@
 other functional data structures. Nunchaku was then used to generate SMT-LIB
-files, again in two versions:\ without and with the $\conv$~translation.
-Problems requiring higher-order logic were discarted, since Nunchaku does not
+files, again in two versions:\ with and without the $\conv$~translation.
+Problems requiring higher-order logic were discarded, since Nunchaku does not
 yet support them, leaving 357 problems.
@@
 For \cvc, we considered four configurations, referred
-to as \cvcd, \cvcf, \cvcfe, and \cvcfm.
-The configuration \cvcd runs heuristic 
+to as \cvcd, \cvcf, \cvcfe, and \cvcfm here.
+Configuration \cvcd runs heuristic
 and conflict-based techniques for quantifier instantiation~\cite{ReynoldsTinelliMoura14},
@@
 finding procedure due to Reynolds et al.\ \cite{ReyEtAl-1-RR-13,reynolds-et-al-2013}.
-The configuration \cvcfe additionally incorporates heuristic quantifier instantiation as described in Section 2.3 of~\cite{reynolds-et-al-2013},
+Configuration \cvcfe additionally incorporates heuristic quantifier instantiation as described in Section 2.3 of~\cite{reynolds-et-al-2013},
 and \cvcfm incorporates the fairness scheme for monotonic types as described in Section~\ref{ssec:cvc4}.
@@
 \\
-\ipm & 0 & 43 & 0 & 0 & 0 & {\win 161} & 0 & {\win 161} & 0 & {\win 161}
+\ipm & 0 & 41 & 0 & 0 & 0 & {\win 153} & 0 & {\win 153} & 0 & {\win 153}
 \\
-\leon  & 8 & 30 & 6 & 6 & 6 & {\win 54} & 6 & 53 & 6 & {\win 54}
+\leon  & 0 & 2 & 0 & 0 & 0 & 9 & 0 & 9 & 0 & {\win 10}
 \\
-\leonm & 6 & 61 & 0 & 0 & 2 & 180 & 2 & {\win 181} & 2 & {\win 181}
+\leonm & 11 & 78 & 6 & 6 & 6 & {\win 189} & 6 & {\win 189} & 6 & {\win 189}
 \\
-\isanun  & 0 & 14 & 0 & 0 & 0 & 155 & 0 & {\win 156} & 0 & 155
+\isanun  & 3 & 27 & 0 & 0 & 3 & 199 & 2 & {\win 200} & 2 & 199
 \\[\jot]
@@
 \\
-\ipm & 54 & 55 & {\win 56} & {\win 56} & 13 & 55 & {\win 56} & 55 & 13 & 55
+\ipm & {\win 18} & {\win 18} & {\win 18} & {\win 18} & 4 & {\win 18} & {\win 18} & {\win 18} & 4 & {\win 18}
 \\
-\leon & 48 & 52 & {\win 58} & 52 & 12 & 53 & {\win 58} & 50 & 12 & 53
+\leon & 74 & 79 & {\win 80} & {\win 80} & 17 & 78 & {\win 80} & 77 & 17 & 78
 \\
-\leonm & 94 & 107 & 106 & 107 & 39 & {\win 108} & 106 & {\win 108} & 39 & {\win 108}
+\leonm & 84 & 98 & {\win 104} & 98 & 24 & 100 & {\win 104} & 98 & 24 & 100
 \\
-\isanun & {\win 41} & 40 & 28 & 34 & 26 & 39 & 26 & 39 & 26 & 39
+\isanun & {\win 61} & 59 & 46 & 53 & 45 & 59 & 44 & 59 & 45 & 59
 \\[\jot]
@@
 The benchmarks and more detailed results are available online.%
-\footnote{\url{http://lara.epfl.ch/~reynolds/IJCAR2016-recfun/}}
+\footnote{\url{http://cs.uiowa.edu/~ajreynol/IJCAR2016-recfun}}
 The figures are divided into benchmarks triggering \emph{unsat} and \emph{sat}
@@
 impact. \cvcf finds 550 of the 1242 benchmarks to be satisfiable,
-including 6~benchmarks in the nonmutated \leon benchmark set. 
+including 9~benchmarks in the nonmutated \leon benchmark set. 
 The two optimizations for finite model finding in \cvc (configurations \cvcfe and \cvcfm) 
@@
 
-Moreover,
+Interestingly,
 the translation $\conv$ helps all configurations for \emph{unsat} responses as well.
@@
 solves 270 \emph{unsat} benchmarks overall, which is more than both \cvcd and \ziii.
-These results suggest that the translation do not degrade the performance 
+These results suggest that the translation does not degrade the performance
 of SMT solvers for unsatisfiable problems involving recursive functions,
-and instead often improve their performance. It would be interesting to try this
-out in Sledgehammer \cite{blanchette-et-al-2013-smt} and to try Nunchaku as a proof tool.
+and instead often improve their performance. 
+These results suggest that it would be interesting to try this translation
+in Sledgehammer \cite{blanchette-et-al-2013-smt} and to try Nunchaku also as a
+proof tool.
 
@@
 
-We described the most closely related work, by Ge and de Moura
+We have already described the most closely related work, by Ge and de Moura
 \cite{GeDeM-CAV-09} and by Reynolds et al.\
-\cite{ReyEtAl-1-RR-13,reynolds-et-al-2013}, in the text already.
+\cite{ReyEtAl-1-RR-13,reynolds-et-al-2013}, earlier in this paper.
 The finite model finding support in the instantiation-based iProver
@@
 by finite subterm-closed substructures. The approach finds sound (fragments
-of) models for formulas in the \relax{existential--bounded-universal} fragment%
+of) models for formulas in the \relax{existential--bounded-universal} fragment 
 (i.e., formulas whose prenex normal forms contain no unbounded universal
@@
 Our translation~$\conv$ works in an analogous manner,
-where instead the SMT solver is invoked only once 
+but the SMT solver is invoked only once
 and quantifier instantiation is used in lieu of function unfolding.
@@
 
-\def\ackname{Acknowledgment}
+\def\ackname{Acknowledgments}
 \paragraph{%\footnotesize
@@
 to thank Damien Busato-Gaston and Emmanouil Koukoutos for providing the
-set of Leon benchmarks used in the evaluation, and the anonymous reviewers for their
-suggestions and comments. Cruanes is supported by the Inria technological
+set of Leon benchmarks used in the evaluation, 
+and
+Mark Summerfield for suggesting
+several textual improvements. %, and the anonymous reviewers for their suggestions and comments. 
+Cruanes is supported by the Inria technological
 development action ``Contre-exemples utilisables par Isabelle et Coq'' (CUIC).

### 2016-01-28 ijcar2016/body.tex
Mark Summerfield's feedback

--- 2016-01-26 ijcar2016/body.tex
@@
 SMT solvers have recently been extended with techniques for finding models
-in presence of universally quantified formulas in some restricted fragments.
+in the presence of universally quantified formulas in some restricted fragments.
 This paper introduces a translation which reduces axioms specifying a large
@@
 %
-We evaluated the two evaluation on benchmarks from
+We evaluated the two implementations on benchmarks from
 IsaPlanner~\cite{DBLP:conf/itp/JohanssonDB10}, Leon~\cite{blanc2013overview},
@@
 constructor before entering the nested call. Like terminating recursion,
-productive corecursion totally specify the functions it defines.
+productive corecursion totally specifies the functions it defines.
 It is even possible to mix recursion and corecursion in the same function
@@
 files, again in two versions:\ without and with the $\conv$~translation.
-Problems requiring higher-order logic were discarted, since Nunchaku does not
+Problems requiring higher-order logic were discarded, since Nunchaku does not
 yet support them, leaving 357 problems.
@@
 solves 270 \emph{unsat} benchmarks overall, which is more than both \cvcd and \ziii.
-These results suggest that the translation do not degrade the performance 
+These results suggest that the translation does not degrade the performance
 of SMT solvers for unsatisfiable problems involving recursive functions,
@@
 to thank Damien Busato-Gaston and Emmanouil Koukoutos for providing the
-set of Leon benchmarks used in the evaluation, and the anonymous reviewers for their
+set of Leon benchmarks used in the evaluation, Mark Summerfield for suggesting
+several textual improvements, and the anonymous reviewers for their
 suggestions and comments. Cruanes is supported by the Inria technological

### 2016-01-28 ijcar2016/body.tex
tuning

@@
 If some function definitions are not admissible, \cvc may answer \emph{sat} for
-an unsatisfiable problem.
-\begin{rep}
-Indeed, if we add the inconsistent definition
+an unsatisfiable problem. For example, if we add the inconsistent definition
 %

### 2016-01-28 ijcar2016/body.tex
Several fixes and edits, all minor or cosmetic

--- 2016-01-26 ijcar2016/body.tex
@@
 \hyphenation{counter-example counter-examples Isa-belle data-type data-types
-  co-data-type co-data-types bench-mark bench-marks}
+  co-data-type co-data-types}
 
@@
 \cite{MouraBjoerner07,ReynoldsTinelliMoura14}.
-These methods work well for proving the
+These methods work well in the context of proving the
 unsatisfiability of an input set of formulas, 
-but they are of little help for finding models of them
+but they are of little help for finding models for them
 when they are satisfiable. Often, a single universal quantifier in one of the
@@
 variables occur as direct subterms of uninterpreted
-functions, as in $\forall x : \typeint.\; \con{f}( x ) \teq \con{g}( x ) + 5$.
+functions, as in $\forall x%\typ{\typeint} --- the type is not important, right?
+.\;\, \con{f}( x )
+\teq \con{g}( x ) + 5$.
 % since all
@@
   2 * \con{p}( x - 1 ) \bigr)$$ %.
-where $\lite$ is the `if--then--else' operator.
+where $\lite$ is the if-then-else operator.
 This state of affairs is unsatisfactory, given the frequency of
@@
 definitions into formulas where finite model finding techniques can be applied.
-The definitions must meet a semantic criterion to be admissible 
+The definitons must meet a semantic criterion to be admissible 
 (Section~\ref{sec:prelim}). 
@@
 recursive function definitions and restrictive enough to exclude inconsistent
-equations such as \,$\forall x : {\typeint}.\;\, \con{f}(x) \teq \con{f}(x) +
+definitions such as \,$\forall x : {\typeint}.\;\, \con{f}(x) \teq \con{f}(x) +
 1$.
@@
 (Section~\ref{sec:encoding}). A
-recursive equation $\forall x : \tau.\;\, \const{f}(x) \teq t$
+recursive definition $\forall x : \tau.\;\, \const{f}(x) \teq \mathit{rhs}$
 is translated to
-$\forall a :\nobreak
-\fargtype{\tau}.\allowbreak\;\, \const{f}(\farg{\con{f}}(a)) \teq t[\farg{\con{f}}(a)/x]$, where
-%%% To me, "abstract" and "concrete" are just convenient labels to refer to the
-%%% thing. I don't want readers to start wondering, "What is an abstract type?"
-%%% So I find it important to leave "uninterpreted" outside of parentheses. As
-%%% for italics, I went for quotes instead, since this is an informal explanation
-%%% anyway. Later, we define \emph{abstract type} with italics. --JB
-$\fargtype{\tau}$ is an ``abstract'' uninterpreted type and $\farg{\con{f}} :
+$\forall a :
+\fargtype{\tau}.\;\, \const{f}(\farg{\con{f}}(a)) \teq \mathit{rhs}[\farg{\con{f}}(a)/x]$, where
+$\fargtype{\tau}$ is an \emph{abstract} (uninterpreted) type and $\farg{\con{f}} :
 \fargtype{\tau} \to \tau$ is an uninterpreted function
@@
 terms of type~$\tau$ and $\terms( t )$ to denote the set of subterms in $t$.
-Given a term $u$, a variable tuple $\vec x = (x_1^{\tau_1},\dotsc,x_n^{\tau_n})$ and 
-a term tuple $\vec t = (t_{\smash{1}\vphantom{n}}^{\tau_1},\dotsc,t_n^{\tau_n})$, 
+Given a term $u$, a variable tuple $\vec x = (x_1^{\tau_1},\ldots,x_n^{\tau_n})$ and 
+a term tuple $\vec t = (t_1^{\tau_1},\ldots,t_n^{\tau_n})$, 
 we write $u[\vec t/\vec x]$ to denote the result of simultaneously replacing
-all occurrences of $x_i$ with $t_i$ in $u$, for each $i=1,\dotsc,n$.
+all occurrences of $x_i$ with $t_i$ in $u$, for each $i=1,\ldots,n$.
 \begin{rep}When applied to terms, the symbol $=$ denotes syntactic equality.\end{rep}
@@
 each function symbol $\con{f} : \tau_1 \times \cdots \times \tau_n \rightarrow \tau$ in $\sfuns{\Sigma}$
-to a total function $\con{f}^\I : \tau_{\smash{1}}^\I \times \cdots \times \tau_n^\I \rightarrow \tau^\I$,
+to a total function $\con{f}^\I : \tau_1^\I \times \cdots \times \tau_n^\I \rightarrow \tau^\I$,
 and each variable $x:\tau$ of $\svars{\Sigma}$ to an element of $\tau^\I$.
@@
 (\,\emph{for $\con{f}$}\vthinspace) if $\vec x$ is a tuple of variables
-$x_1^{\tau_1}$, $\dotsc,$ $x_n^{\tau_n}$
+$x_1^{\tau_1}$, $\ldots,$ $x_n^{\tau_n}$
 and $\varphi$ is a quantifier-free formula
@@
 A formula $\varphi$ is in \emph{definitional form with respect to}
-$\{ \con{f}_1, \dotsc, \con{f}_n \} \subseteq \sfuns{\Sigma}$ if it is of the
+$\{ \con{f}_1, \ldots, \con{f}_n \} \subseteq \sfuns{\Sigma}$ if it is of the
 form
@@
 %
-where $\con{f}_1, \dotsc, \con{f}_n$ are distinct function symbols,
+where $\con{f}_1, \ldots, \con{f}_n$ are distinct function symbols,
 %ct added
 $\forallf{\con{f}_i} \vec x_i.\; \varphi_i$ is a function definition
-for $i = 1, \dotsc, n$,
+for $i = 1, \ldots, n$,
 %
@@
 Given a set of function definitions
-$\DDD = \{ \forallf{\con{f}_1} \vec x_1.\; \varphi_1, \dotsc, \forallf{\con{f}_n} \vec x_n.\; \varphi_n \}$,
+$\DDD = \{ \forallf{\con{f}_1} \vec x_1.\; \varphi_1, \ldots, \forallf{\con{f}_n} \vec x_n.\; \varphi_n \}$,
 a ground formula $\psi$
@@
 If $\DDD$ is a set of well-founded function definitions for\/
-$\sfundefs{\Sigma} = \{\con{f}_1,\dotsc,\con{f}_n\}$, then it is admissible.
+$\sfundefs{\Sigma} = \{\con{f}_1,\ldots,\con{f}_n\}$, then it is admissible.
 \end{theorem}
@@
 $\forallf{\const{e}} x.\;\, \const{e}(x) \teq \con{scons}(x,\; \const{e}(x + 1))$
-%%% If what follows ':' is a complete sentence, the standard Am.E. convention
-%%% is to capitalize it. --JB
-falls within this class: Each call to $\const{e}$ produces one
+falls within this class: each call to $\const{e}$ produces one
 constructor before entering the nested call. Like terminating recursion,
@@
  \begin{itemize}
-   \itemx[] $\innerkw{if}$ $\tau = \Bool$ $\mathsf{and}$ $t = \boolop(t_1,\dotsc,t_n)$ $\innerkw{then}$
+   \itemx[] $\innerkw{if}$ $\tau = \Bool$ $\mathsf{and}$ $t = \boolop(t_1,\ldots,t_n)$ $\innerkw{then}$
     \begin{itemize}
-      \itemx[] $\innerkw{let}$ $( t'_i,\, \chi_i ) = \conv_0( t_i,\, \pol( \boolop,\, i,\, p ) )$ $\mathsf{for}$ $i = 1, \dotsc, n$ $\innerkw{in}$%\kern6.35em$\mid$
+      \itemx[] $\innerkw{let}$ $( t'_i,\, \chi_i ) = \conv_0( t_i,\, \pol( \boolop,\, i,\, p ) )$ $\mathsf{for}$ $i = 1, \ldots, n$ $\innerkw{in}$%\kern6.35em$\mid$
       \itemx[] $\innerkw{let}$ $\chi = \chi_1 \wedge \cdots \wedge \chi_n$ $\innerkw{in}$
-      \itemx[] $\innerkw{if}$ $p = \ppos$ $\innerkw{then}$ \bigtuple{$\boolop(t'_1, \dotsc, t'_n) ) \wedge \chi,\, \ltrue$}
-      \itemx[] $\innerkw{else}$ $\innerkw{if}$ $p = \pneg$ $\innerkw{then}$ \bigtuple{$\boolop(t'_1, \dotsc, t'_n) \vee \neg \chi,\, \ltrue$}
-      \itemx[] $\innerkw{else}$ \bigtuple{$\boolop(t'_1, \dotsc, t'_n),\, \chi$}
+      \itemx[] $\innerkw{if}$ $p = \ppos$ $\innerkw{then}$ \bigtuple{$\boolop(t'_1, \ldots, t'_n) ) \wedge \chi,\, \ltrue$}
+      \itemx[] $\innerkw{else}$ $\innerkw{if}$ $p = \pneg$ $\innerkw{then}$ \bigtuple{$\boolop(t'_1, \ldots, t'_n) \vee \neg \chi,\, \ltrue$}
+      \itemx[] $\innerkw{else}$ \bigtuple{$\boolop(t'_1, \ldots, t'_n),\, \chi$}
 %\\[-.8\baselineskip]
@@
 \emph{relevant} input tuples for each defined function $\con{f}$ and restrict the
-quantification of $\con{f}$'s definition to a single variable of this type.
+quantification of $\con{f}$'s definition to a (single) variable of this type.
 Informally, the relevant input tuples $\vec t$ of a function $\con{f}$ are the
@@
 an uninterpreted \emph{abstract type} $\fargtype{\con{f}}$ 
-abstracting the Cartesian product $\tau_1 \times \cdots \times \tau_n$
+(abstracting the cross product $\tau_1 \times \cdots \times \tau_n$)
 and
 %\item
-$n$ uninterpreted \emph{concretization functions} $\fargx{\con{f}}{1} : \fargtype{\con{f}} \rightarrow \tau_1$, $\dotsc,$ $\fargx{\con{f}}{n} : \fargtype{\con{f}} \rightarrow \tau_n$.
+$n$ uninterpreted \emph{concretization functions} $\fargx{\con{f}}{1} : \fargtype{\con{f}} \rightarrow \tau_1$, \ldots, $\fargx{\con{f}}{n} : \fargtype{\con{f}} \rightarrow \tau_n$.
 %\end{itemize}
@@
 $\bigwedge_{i=1}^n\, \fargx{\con{f}}{i}(a) \teq s_i$
-with $\vec s = (s_1,\dotsc,s_n)$.
+with $\vec s = (s_1,\ldots,s_n)$.
 These existential %%% for typesetting
@@
 that are relevant to the satisfiability of $\varphi$. The constraints are
-generated as deep in the formula as possible, based on the
-polarities of Boolean connectives, to allow models where the sets
+generated as deep in the formula as possible, based on our knowledge of the
+polarities of logical connectives, to allow models where the sets
 denoted by the $\fargtype{\con{f}}$ types are as small as possible.
@@
 if $p$ is $\pneg$, and $\pnone$ if $p$ is $\pnone$.
-The term $t$ is then reconstructed as $\boolop(t_1',\dotsc,t_n')$
+The term $t$ is then reconstructed as $\boolop(t_1',\ldots,t_n')$
 where each $t_i'$ is the result of the recursive call with argument~$t_i$.
 If the polarity $p$ associated with $t$ is $\ppos$,
-$\conv_0$ conjunctively adds to $\boolop(t_1',\dotsc,t_n')$ the constraint $\chi$ derived from
+$\conv_0$ conjunctively adds to $\boolop(t_1',\ldots,t_n')$ the constraint $\chi$ derived from
 the subterms and returns $\top$ as the constraint.
@@
 a complete instantiation procedure, implemented in \ziii.
-As expected, \cvc and \ziii run indefinitely on formula~(\ref{eq:ex-before}).
-However, they produce a model for~(\ref{eq:ex-after})
+As expected,
+%%% Are we sure?
+% Yes, I'm planning to post the 2 versions of the benchmark on the evaluation page (when we add it) --ajr
+% Yes, but how can you be sure that Z3 will run forever? Your benchmarks surely have a timeout! --jb
+\cvc and \ziii run indefinitely on formula~(\ref{eq:ex-before}).
+Howevever, they produce a model for~(\ref{eq:ex-after})
 within 100 milliseconds.\xend
@@
     \lite( x \teq 2,\allowbreak\; 3,\allowbreak\;
-      \lite(\dotsc,\allowbreak\; \lite( x \teq 13,\allowbreak\, 91,\allowbreak\, 105 )\negvthinspace \ldots ))))}
+      \lite(\ldots,\allowbreak\; \lite( x \teq 13,\allowbreak\, 91,\allowbreak\, 105 )\negvthinspace \ldots ))))}
 
@@
 One model $\I$ for formula~(\ref{eq:ex-after}) in the above example interprets
-$\fargtype{\con{s}}$ as a finite set $\{ u_0, \dotsc, u_{14} \}$,
-$\farg{\con{s}}$ as a finite map $u_i \mapsto i$ for $i = 0, \dotsc, 14$,
+$\fargtype{\con{s}}$ as a finite set $\{ u_0, \ldots, u_{14} \}$,
+$\farg{\con{s}}$ as a finite map $u_i \mapsto i$ for $i = 0, \ldots, 14$,
 $\con{c}$ as $14$,
@@
 to the domains of the corresponding abstract types $\fargtype{\con{f}}$.
-This can sometimes help users debug their
+This can sometimes help users debug their %specifications.
 function definitions.
@@
 interprets $\fargtype{\con{f}}$ as $\tau_1^\I \times \cdots \times \tau_n^\I$
-and each $\fargx{\con{f}}{i}$ as the $i$\vvthinspace{th} projection on such tuples for $i = 1, \dotsc, n$.
+and each $\fargx{\con{f}}{i}$ as the $i$\vvthinspace{th} projection on such tuples for $i = 1, \ldots, n$.
 Since $\J$ satisfies $\varphi$, it satisfies a set of ground literals $L$ that entail $\varphi$.
@@
 an $\extendsig{\Sigma}$-interpretation that satisfies $\conv( \varphi )$, by
-interpreting the abstract types in the same way as the Cartesian products they
+interpreting the abstract types in the same way as the cartesian products they
 abstract, thereby satisfying all existential constraints introduced by $\conv$.
@@
 where $L_0 = \{ \con{s}( \con{c} ) > 100 \}$
-and $\absconstraints( L_0 )$, defined by equation (\ref{eq:xl}), consists of the single constraint
+and $\absconstraints( L_0 )$ (see~(\ref{eq:xl})) consists of the single constraint
 $\exists a : {\fargtype{\con{s}}}.\;\, \farg{\con{s}}( a ) \teq \con{c}$.
@@
 This formula is closed under function expansion
-since it entails $\varphi_\con{s}[(\con{c}-i)/x]$ for $i = 0, \dotsc, n$
-and contains only $\con{s}$ applications corresponding to $\con{s}( \con{c}-i )$ for $i = 0, \dotsc, n$.
+since it entails $\varphi_\con{s}[(\con{c}-i)/x]$ for $i = 0, \ldots, n$,
+and contains only $\con{s}$ applications corresponding to $\con{s}( \con{c}-i )$ for $i = 0, \ldots, n$.
 Since $\{\forallf{\con{s}} x.\;\, \varphi_\con{s}\}$ is admissible,
@@
 model of~$\con{s}$ where only
-the values of $\con{s}(x)$ for $x = 0,\dotsc,14$ are correctly given:
+the values of $\con{s}(x)$ for $x = 0,\ldots,14$ are correctly given:
 \begin{quote}
@@
 functions introduced using \texttt{define-\allowbreak fun-\allowbreak rec} are admissible.
-Admissibility must be proved externally by the user---e.g., manually, using a
+Admissibility must be proven externally by the user---e.g., manually, using a
 syntactic criterion, or with the help of a termination prover.
@@
 \url{https://github.com/epfl-lara/leon/}} which were constructed from
-verification conditions on simple Scala programs. These benchmarks also
+verification conditions on simple Scala programs. These bench\-marks also
 heavily involve recursively defined functions over algebraic datatypes,
@@
 impact. \cvcf finds 550 of the 1242 benchmarks to be satisfiable,
-including 9~benchmarks in the nonmutated \leon benchmark set. 
+including 6~benchmarks in the nonmutated \leon benchmark set.
 The two optimizations for finite model finding in \cvc (configurations \cvcfe and \cvcfm)
-led to a net gain of one~satisfiable benchmark each with respect to \cvcf.
-The performance of \ziiib for countermodels also improved dramatically, as it
+leads to a net gain of one~satisfiable benchmark each with respect to \cvcf.
+The performance of \ziiib for countermodels also improves dramatically, as it
 finds 134 more benchmarks to be satisfiable, including 5 that are not solved by \cvcf.
@@
 These results suggest that it would be interesting to try this translation
-in Sledgehammer \cite{blanchette-et-al-2013-smt} and to try Nunchaku also as a
-proof tool.
+in Sledgehammer \cite{blanchette-et-al-2013-smt} and 
+that to try to use Nunchaku also as a proof tool.
 
@@
 Our translation~$\conv$ works in an analogous manner,
-but the SMT solver is invoked only once
+where, however, the SMT solver is invoked only once
 and quantifier instantiation is used in lieu of function unfolding.

### 2016-01-28 ijcar2016/body.tex
tuning Cesare's changes

@@
 \cite{MouraBjoerner07,ReynoldsTinelliMoura14}.
-These methods work well in the context of proving the
+These methods work well for proving the
 unsatisfiability of an input set of formulas, 
-but they are of little help for finding models for them
+but they are of little help for finding models of them
 when they are satisfiable. Often, a single universal quantifier in one of the
@@
 variables occur as direct subterms of uninterpreted
-functions, as in $\forall x%\typ{\typeint} --- the type is not important, right?
-.\;\, \con{f}( x )
-\teq \con{g}( x ) + 5$.
+functions, as in $\forall x : \typeint.\; \con{f}( x ) \teq \con{g}( x ) + 5$.
 % since all
@@
   2 * \con{p}( x - 1 ) \bigr)$$ %.
-where $\lite$ is the if-then-else operator.
+where $\lite$ is the `if--then--else' operator.
 This state of affairs is unsatisfactory, given the frequency of
@@
 definitions into formulas where finite model finding techniques can be applied.
-The definitons must meet a semantic criterion to be admissible 
+The definitions must meet a semantic criterion to be admissible 
 (Section~\ref{sec:prelim}). 
@@
 recursive function definitions and restrictive enough to exclude inconsistent
-definitions such as \,$\forall x : {\typeint}.\;\, \con{f}(x) \teq \con{f}(x) +
+equations such as \,$\forall x : {\typeint}.\;\, \con{f}(x) \teq \con{f}(x) +
 1$.
@@
 (Section~\ref{sec:encoding}). A
-recursive definition $\forall x : \tau.\;\, \const{f}(x) \teq \mathit{rhs}$
+recursive equation $\forall x : \tau.\;\, \const{f}(x) \teq t$
 is translated to
-$\forall a :
-\fargtype{\tau}.\;\, \const{f}(\farg{\con{f}}(a)) \teq \mathit{rhs}[\farg{\con{f}}(a)/x]$, where
-$\fargtype{\tau}$ is an \emph{abstract} (uninterpreted) type and $\farg{\con{f}} :
+$\forall a :\nobreak
+\fargtype{\tau}.\allowbreak\;\, \const{f}(\farg{\con{f}}(a)) \teq t[\farg{\con{f}}(a)/x]$, where
+%%% To me, "abstract" and "concrete" are just convenient labels to refer to the
+%%% thing. I don't want readers to start wondering, "What is an abstract type?"
+%%% So I find it important to leave "uninterpreted" outside of parentheses. As
+%%% for italics, I went for quotes instead, since this is an informal explanation
+%%% anyway. Later, we define \emph{abstract type} with italics. --JB
+$\fargtype{\tau}$ is an ``abstract'' uninterpreted type and $\farg{\con{f}} :
 \fargtype{\tau} \to \tau$ is an uninterpreted function
@@
 terms of type~$\tau$ and $\terms( t )$ to denote the set of subterms in $t$.
-Given a term $u$, a variable tuple $\vec x = (x_1^{\tau_1},\ldots,x_n^{\tau_n})$ and 
-a term tuple $\vec t = (t_1^{\tau_1},\ldots,t_n^{\tau_n})$, 
+Given a term $u$, a variable tuple $\vec x = (x_1^{\tau_1},\dotsc,x_n^{\tau_n})$ and 
+a term tuple $\vec t = (t_{\smash{1}\vphantom{n}}^{\tau_1},\dotsc,t_n^{\tau_n})$, 
 we write $u[\vec t/\vec x]$ to denote the result of simultaneously replacing
-all occurrences of $x_i$ with $t_i$ in $u$, for each $i=1,\ldots,n$.
+all occurrences of $x_i$ with $t_i$ in $u$, for each $i=1,\dotsc,n$.
 \begin{rep}When applied to terms, the symbol $=$ denotes syntactic equality.\end{rep}
@@
 (\,\emph{for $\con{f}$}\vthinspace) if $\vec x$ is a tuple of variables
-$x_1^{\tau_1}$, $\ldots,$ $x_n^{\tau_n}$
+$x_1^{\tau_1}$, $\dotsc,$ $x_n^{\tau_n}$
 and $\varphi$ is a quantifier-free formula
@@
 Given a set of function definitions
-$\DDD = \{ \forallf{\con{f}_1} \vec x_1.\; \varphi_1, \ldots, \forallf{\con{f}_n} \vec x_n.\; \varphi_n \}$,
+$\DDD = \{ \forallf{\con{f}_1} \vec x_1.\; \varphi_1, \dotsc, \forallf{\con{f}_n} \vec x_n.\; \varphi_n \}$,
 a ground formula $\psi$
@@
 If $\DDD$ is a set of well-founded function definitions for\/
-$\sfundefs{\Sigma} = \{\con{f}_1,\ldots,\con{f}_n\}$, then it is admissible.
+$\sfundefs{\Sigma} = \{\con{f}_1,\dotsc,\con{f}_n\}$, then it is admissible.
 \end{theorem}
@@
 $\forallf{\const{e}} x.\;\, \const{e}(x) \teq \con{scons}(x,\; \const{e}(x + 1))$
-falls within this class: each call to $\const{e}$ produces one
+%%% If what follows ':' is a complete sentence, the standard Am.E. convention
+%%% is to capitalize it. --JB
+falls within this class: Each call to $\const{e}$ produces one
 constructor before entering the nested call. Like terminating recursion,
@@
 \emph{relevant} input tuples for each defined function $\con{f}$ and restrict the
-quantification of $\con{f}$'s definition to a (single) variable of this type.
+quantification of $\con{f}$'s definition to a single variable of this type.
 Informally, the relevant input tuples $\vec t$ of a function $\con{f}$ are the
@@
 an uninterpreted \emph{abstract type} $\fargtype{\con{f}}$ 
-(abstracting the cross product $\tau_1 \times \cdots \times \tau_n$)
+abstracting the Cartesian product $\tau_1 \times \cdots \times \tau_n$
 and
 %\item
-$n$ uninterpreted \emph{concretization functions} $\fargx{\con{f}}{1} : \fargtype{\con{f}} \rightarrow \tau_1$, \ldots, $\fargx{\con{f}}{n} : \fargtype{\con{f}} \rightarrow \tau_n$.
+$n$ uninterpreted \emph{concretization functions} $\fargx{\con{f}}{1} : \fargtype{\con{f}} \rightarrow \tau_1$, $\dotsc,$ $\fargx{\con{f}}{n} : \fargtype{\con{f}} \rightarrow \tau_n$.
 %\end{itemize}
@@
 that are relevant to the satisfiability of $\varphi$. The constraints are
-generated as deep in the formula as possible, based on our knowledge of the
-polarities of logical connectives, to allow models where the sets
+generated as deep in the formula as possible, based on the
+polarities of Boolean connectives, to allow models where the sets
 denoted by the $\fargtype{\con{f}}$ types are as small as possible.
@@
 if $p$ is $\pneg$, and $\pnone$ if $p$ is $\pnone$.
-The term $t$ is then reconstructed as $\boolop(t_1',\ldots,t_n')$
+The term $t$ is then reconstructed as $\boolop(t_1',\dotsc,t_n')$
 where each $t_i'$ is the result of the recursive call with argument~$t_i$.
 If the polarity $p$ associated with $t$ is $\ppos$,
-$\conv_0$ conjunctively adds to $\boolop(t_1',\ldots,t_n')$ the constraint $\chi$ derived from
+$\conv_0$ conjunctively adds to $\boolop(t_1',\dotsc,t_n')$ the constraint $\chi$ derived from
 the subterms and returns $\top$ as the constraint.
@@
 a complete instantiation procedure, implemented in \ziii.
-As expected,
-%%% Are we sure?
-% Yes, I'm planning to post the 2 versions of the benchmark on the evaluation page (when we add it) --ajr
-% Yes, but how can you be sure that Z3 will run forever? Your benchmarks surely have a timeout! --jb
-\cvc and \ziii run indefinitely on formula~(\ref{eq:ex-before}).
-Howevever, they produce a model for~(\ref{eq:ex-after})
+As expected, \cvc and \ziii run indefinitely on formula~(\ref{eq:ex-before}).
+However, they produce a model for~(\ref{eq:ex-after})
 within 100 milliseconds.\xend
@@
 to the domains of the corresponding abstract types $\fargtype{\con{f}}$.
-This can sometimes help users debug their %specifications.
+This can sometimes help users debug their
 function definitions.
@@
 where $L_0 = \{ \con{s}( \con{c} ) > 100 \}$
-and $\absconstraints( L_0 )$ (see~(\ref{eq:xl})) consists of the single constraint
+and $\absconstraints( L_0 )$, defined by equation (\ref{eq:xl}), consists of the single constraint
 $\exists a : {\fargtype{\con{s}}}.\;\, \farg{\con{s}}( a ) \teq \con{c}$.
@@
 This formula is closed under function expansion
-since it entails $\varphi_\con{s}[(\con{c}-i)/x]$ for $i = 0, \ldots, n$,
-and contains only $\con{s}$ applications corresponding to $\con{s}( \con{c}-i )$ for $i = 0, \ldots, n$.
+since it entails $\varphi_\con{s}[(\con{c}-i)/x]$ for $i = 0, \dotsc, n$
+and contains only $\con{s}$ applications corresponding to $\con{s}( \con{c}-i )$ for $i = 0, \dotsc, n$.
 Since $\{\forallf{\con{s}} x.\;\, \varphi_\con{s}\}$ is admissible,
@@
 functions introduced using \texttt{define-\allowbreak fun-\allowbreak rec} are admissible.
-Admissibility must be proven externally by the user---e.g., manually, using a
+Admissibility must be proved externally by the user---e.g., manually, using a
 syntactic criterion, or with the help of a termination prover.
@@
 \url{https://github.com/epfl-lara/leon/}} which were constructed from
-verification conditions on simple Scala programs. These bench\-marks also
+verification conditions on simple Scala programs. These benchmarks also
 heavily involve recursively defined functions over algebraic datatypes,
@@
 The two optimizations for finite model finding in \cvc (configurations \cvcfe and \cvcfm)
-leads to a net gain of one~satisfiable benchmark each with respect to \cvcf.
+lead to a net gain of one~satisfiable benchmark each with respect to \cvcf.
 The performance of \ziiib for countermodels also improves dramatically, as it
@@
 These results suggest that it would be interesting to try this translation
-in Sledgehammer \cite{blanchette-et-al-2013-smt} and 
-that to try to use Nunchaku also as a proof tool.
+in Sledgehammer \cite{blanchette-et-al-2013-smt} and to try Nunchaku also as a
+proof tool.
 
@@
 Our translation~$\conv$ works in an analogous manner,
-where, however, the SMT solver is invoked only once
+but the SMT solver is invoked only once
 and quantifier instantiation is used in lieu of function unfolding.

### 2016-03-30 ijcar2016/body.tex
Fix merge
TODOs
did some TODOs

@@
 \hyphenation{counter-example counter-examples Isa-belle data-type data-types
-  co-data-type co-data-types bench-mark bench-marks}
+  co-data-type co-data-types bench-mark bench-marks counter-model counter-models}
 
@@
 
-In the past few years, a few techniques have been developed to find models for
+In the past few years, %a few
+techniques have been developed to find models for
 quantified formulas in SMT.
@@
   2 * \con{p}( x - 1 ) \bigr)$,
-where the theory $T$ is integer arithmetic extended with the uninterpreted symbol
-$\con p:\typeint \rightarrow \typeint$, can be shown to be well founded under a strategy that
+where the theory $T$ is integer arithmetic extended with the %uninterpreted
+symbol $\con p:\typeint \rightarrow \typeint$, can be shown to be well founded
+under a strategy that
 evaluates the condition of an $\lite$ before evaluating the relevant branch,
@@
 $\{\con f, \con g :\typeint \rightarrow \typeint,\, \ldots \}$.
-The reason is that the formula $\ltrue$ is (trivially) closed under function expansion with respect to this set,
-and there is no model of $T$
+The set is inadmissible because the formula $\ltrue$ is
+closed under function expansion with respect to this set
+(since $\con{f}$ does not occur in $\ltrue$),
+and yet there is no model of $T$
 satisfying $\con{f}$'s definition. A more subtle example is
-\[\badassex\]
+$\badassex$.
 While this set has a model where $\con f$ and $\con g$ are interpreted as the
@@
 $\extendsig{ \Sigma }$ and an $\extendsig{ \Sigma }$-formula $\varphi'$ such that
-$\varphi'$ is $T$-satisfiable if and only if $\varphi$ is $T$-satisfiable---i.e.,
-$\varphi$ and $\varphi'$ are \emph{equisatisfiable} (\emph{in $T$}).
+$\varphi'$ is $T$-satisfiable if and only if the $\Sigma$-formula $\varphi$ is
+$T$-satisfiable---i.e., $\varphi$ and $\varphi'$ are \emph{equisatisfiable} (\emph{in $T$}).
 The idea behind this translation is to use an uninterpreted type $\alpha_\con{f}$
@@
 with support for (co)algebraic datatypes, (co)recursive functions and
-(co)inductive predicates. Support for higher-order functions is planned for
-the next release. We have developed a preliminary Isabelle frontend and are
-planning further frontends for Coq, the TLA\textsuperscript{+} Proof System,
-and possibly other proof assistants.
+(co)inductive predicates. Support for higher-order functions is in the works.
+We have developed a preliminary Isabelle frontend and are
+planning further frontends for Coq, HOL4, HOL Light,
+the TLA\textsuperscript{+} Proof System, and possibly other proof assistants.
 
@@
 \begin{quote}
-\verb$data nat := Zero | Suc nat.$ \\[2\jot]
+\verb$data nat := Z | Suc nat.$ \\[2\jot]
 \verb$pred even : nat -> prop :=$ \\
-\verb$  even Zero;$ \\
+\verb$  even Z;$ \\
 \verb$  forall n. odd n => even (Suc n)$ \\
@@
 \verb$val m : nat.$ \\
-\verb$goal even m && ~ (m = Zero).$
+\verb$goal even m && ~ (m = Z).$
 \end{quote}
@@
 inductive predicates (\texttt{even} and \texttt{odd}), it declares a constant
-\texttt{m}, and specifies a goal to satisfy (``$m$ is even and nonzero'').
+\texttt{m}, and specifies a goal to satisfy (``\texttt{m} is even and nonzero'').
 Nunchaku quickly finds the following partial model:
@@
 \begin{quote}
-\verb$val m := Suc (Suc Zero).$ \\
-\verb$val odd := fun x. if x = Suc Zero then true else ?__.$ \\
-\verb$val even := fun x. if x = Suc (Suc Zero) || x = Zero then true$ \\
-\verb$                   else ?__.$
+\verb$val m := Suc (Suc Z).$ \\
+\verb$val odd := fun x. if x = Suc Z then true else ?__.$ \\
+\verb$val even := fun x. if x = Suc (Suc Z) || x = Z then true else ?__.$\kern-200mm
 \end{quote}
@@
 goal: ``2 is even if 1 is odd, 1 is odd if 0 is even, and 0 is even.''
-
-Nunchaku parses and types the input problem before applying a sequence
+Our experience with the Nitpick~\cite{blanchette-nipkow-2010} model finder
+is that users are mostly interested in the values assigned to uninterpreted
+constants (e.g., \texttt{m}). Occasionally, the models of recursive functions
+are instructive in the face of underspecification. A typical example is
+the \texttt{head} function that returns the first element of a nonempty list:
+%
+\begin{quote}
+\verb$data list A := Nil | Cons A (list A).$ \\[2\jot]
+\verb$rec head : pi A. list A -> A :=$ \\
+\verb$  forall y ys. head (Cons y ys) = y.$ \\[2\jot]
+\verb$goal ~ (head Nil = 0).$
+\end{quote}
+%
+Nunchaku transforms the specification of \verb$head$ into
+%
+\begin{quote}
+\verb$head xs = match xs with Nil -> head xs | Cons y ys -> y end$
+\end{quote}
+%
+where the unspecified \verb$Nil$ case is expressed via nonterminating
+recursion (\verb$head$ \verb$xs$ \verb$=$ \verb$head$ \verb$xs$).
+The tool exhibits a model in which \verb$head$ \verb$Nil$ is interpreted as
+a nonzero value.
+
+%(e.g., the function \verb$f$ specified by
+%\verb$forall$ \verb$x.$\ \verb$f$ \verb$x$ \verb$=$ \verb$f$ \verb$x$).
+%(e.g., $\forall x.\;\, \con f(x) \teq \con f(x)$).
+
+Internally, Nunchaku parses and types the input problem before applying a sequence
 of translations, each reducing the distance to the target fragment.
@@
 in the spirit of institution theory \cite{goguen-burstall-1992}. The
-pipeline currently consists of the following phases:
+pipeline includes the following phases:
 
@@
       a single nested pattern matching;
+  \item[Specialization]
+    creates instances of functions with static arguments (i.e., an argument
+    that is passed unchanged to all recursive calls);
   \item[Polarization]
@@
   \item[Unrolling]
-    adds a decreasing argument to possibly ill-founded predicates to make
-    them well founded;
+    adds a decreasing argument to possibly ill-founded predicates; % to make them well founded;
   \item[Skolemization]
@@
       into a recursive equation;
-  \item[Recursion elimination]
+  \item[Elimination of higher-order constructs]
+    eliminates $\lambda$-abstractions and substitutes arrays for
+    higher-order functions;
+  \item[Elimination of recursion]
     performs the encoding from Section~\ref{sec:encoding};
@@
 impact. \cvcf finds 550 of the 1242 benchmarks to be satisfiable,
-including 6~benchmarks in the nonmutated \leon benchmark set.
+including 9~benchmarks in the nonmutated \leon benchmark set. 
 The two optimizations for finite model finding in \cvc (configurations \cvcfe and \cvcfm) 
@@
 solvers to find countermodels for conjectures involving recursively defined
-functions whose definitions are admissible.
+functions.  % whose definitions are admissible.
 
@@
 tools are often much faster than model finders, but they typically cannot cope
-with underspecification and nonexecutable functions.
+with unspecified or underspecified functions (e.g., the \texttt{head} function from
+Section~\ref{ssec:nunchaku}).
+%(even on finite domains).
 
@@
 \def\ackname{Acknowledgments}
-\paragraph{%\footnotesize
+\paragraph{\footnotesize
 \ackname.}

### 2016-03-30 ijcar2016/body.tex
more edits

@@
 \begin{quote}
-\verb$data nat := Z | Suc nat.$ \\[2\jot]
+\verb$data nat := 0 | Suc nat.$ \\[2\jot]
 \verb$pred even : nat -> prop :=$ \\
-\verb$  even Z;$ \\
+\verb$  even 0;$ \\
 \verb$  forall n. odd n => even (Suc n)$ \\
@@
 \verb$val m : nat.$ \\
-\verb$goal even m && ~ (m = Z).$
+\verb$goal even m && ~ (m = 0).$
 \end{quote}
@@
 \begin{quote}
-\verb$val m := Suc (Suc Z).$ \\
-\verb$val odd := fun x. if x = Suc Z then true else ?__.$ \\
-\verb$val even := fun x. if x = Suc (Suc Z) || x = Z then true else ?__.$\kern-200mm
+\verb$val m := Suc (Suc 0).$ \\
+\verb$val odd := fun x. if x = Suc 0 then true else ?__.$ \\
+\verb$val even := fun x. if x = Suc (Suc 0) || x = 0 then true else ?__.$\kern-200mm
 \end{quote}
@@
 Nitpick~\cite{blanchette-nipkow-2010}. 
-In our experience, benchmarks created by mutation have a
-high likelihood of having small, easy-to-find models.
+%In our experience,
+Benchmarks created by mutation have a high likelihood of having small, easy-to-find models.
 \end{itemize}
@@
 in two versions:\ with and without the $\conv$~translation. Problems with
-$\conv$ were produced by running \cvc's preprocessor % (Section~\ref{ssec:cvc4})
-on each benchmark.
+$\conv$ were produced by running \cvc's preprocessor. % (Section~\ref{ssec:cvc4})
+%on each benchmark.
 
@@
 and FM-Darwin \cite{baumgartner-et-al-2009} for first-order logic with
-equality; the Alloy Analyzer and its back-end Kodkod \cite{torlak-jackson-2007}
+equality; the Alloy Analyzer and its backend Kodkod \cite{torlak-jackson-2007}
 for first-order relational logic; and Refute \cite{weber-2008} and Nitpick
 \cite{blanchette-nipkow-2010} for higher-order logic.
-
+%
 An alternative is to perform
@@
 \cite{slaney-1994} and SEM \cite{zhang-zhang-1995} and serves as
-the basis of many more model finders, notably the Alloy Analyzer's precursor
-\cite{jackson-1996} and the later versions of~Mace
+the basis of the Alloy Analyzer's precursor
+\cite{jackson-1996} and later versions of~Mace
 \cite{mccune-prover9-mace4}.
 
-Most of the above tools cannot cope with algebraic datatypes or other infinite
-types.
+Most of the above tools cannot cope with infinite types.
 Kuncak and Jackson \cite{kuncak-jackson-2005} presented an idiom for
-encoding datatypes and recursive functions in Alloy, by approximating datatypes
-by finite subterm-closed substructures. The approach finds sound (fragments
-of) models for formulas in the \relax{existential--bounded-universal} fragment 
+encoding algebraic datatypes and recursive functions in Alloy, by
+approximating datatypes by finite subterm-closed substructures.
+The approach finds sound (fragments of) models for formulas in the
+\relax{existential--bounded-universal} fragment 
 (i.e., formulas whose prenex normal forms contain no unbounded universal
-quantifiers ranging over datatypes). This idiom was further developed by Dunets
+quantifiers ranging over datatypes). This idiom was refined by Dunets
 et al.\ \cite{dunets-et-al-2010}, who presented a translation scheme
@@
 existential constraints generated by our translation $\conv$.
-An approach related in scope to ours is given in~\cite{baumgartner2013},
-which establishes the satisfiability of formulas in the presence of admissible
-axioms over infinite domains by proving their negation is entailed.
 
@@
 approximated type with an unknown value, which is propagated by function
-application. This scheme works reasonably well in Nitpick, because it builds
-on a relational logic, but our initial experiments with \cvc suggest
-that it is more efficient to avoid unknowns by adding existential
-constraints.
+application. This scheme works reasonably well in Nitpick,
+%because it builds on a relational logic,
+but experiments with \cvc suggest that it is more efficient to avoid
+unknowns by adding existential constraints.
 
@@
 Section~\ref{ssec:nunchaku}).
+Another approach, which also fails in the face of underspecification,
+is to take the conjecture as an axiom and to attempt to derive a contradiction
+using an automatic theorem prover \cite{baumgartner2013}. If the other axioms
+are consistent (which can be checked syntactically in some cases),
+a contradiction imples the existence of countermodels.
+
 %(even on finite domains).
@@
 %\jb{Alloy has no notion of datatypes, so it's pretty useless. Nitpick implements
-%datatypes and a more FOL-like logic (actually, HOL) on top of Alloy's back-end,
+%datatypes and a more FOL-like logic (actually, HOL) on top of Alloy's backend,
 %using the Kuncak \& Jackson idiom for Alloy, so that's as close as we can get

### 2016-03-30 ijcar2016/body.tex
one less phase
more TOdid

@@
 \renewcommand\vec[1]{\overline{#1}}
+
+%% TYPESETTING HACKS
+\newcommand\medrightarrow{\mathrel{{{\color{black}\relbar}\kern-0.9ex\rlap{\color{white}\ensuremath{\blacksquare}}\kern-0.9ex}\joinrel{\color{black}\rightarrow}}}
+\newcommand\Medrightarrow{\mathrel{{{\color{black}\Relbar}\kern-0.9ex\rlap{\color{white}\ensuremath{\blacksquare}}\kern-0.9ex}\joinrel{\color{black}\Rightarrow}}}
+\newcommand\medleftrightarrow{\mathrel{\leftarrow\kern-1.685ex\rightarrow}}
+\newcommand\Medleftrightarrow{\mathrel{\Leftarrow\kern-1.685ex\Rightarrow}}
 
@@
 where each $t_i'$ is the result of the recursive call with argument~$t_i$.
-If the polarity $p$ associated with $t$ is $\ppos$,
-$\conv_0$ conjunctively adds to $\boolop(t_1',\dotsc,t_n')$ the constraint $\chi$ derived from
+If the polarity $p$ of $t$ is $\ppos$,
+$\conv_0$ conjunctively adds to $\boolop(t_1',\dotsc,t_n')$ the constraint~$\chi$ derived from
 the subterms and returns $\top$ as the constraint.
 Dually, if $p$ is $\pneg$, it adds a
-disjunction with the negated constraint, to achieve the same overall effect.
+disjunction with the negated constraint to produce the same net effect
+(since $\lnot\;(\phi \mathrel\lor \lnot\;\chi) \Medleftrightarrow \lnot\;\phi \mathrel\land \chi$).
 %ct  as in the positive case.
@@
 satisfiable with a model that interprets $\const{c}$ as $14$ or more.
-Due to the universal quantifier,
-current SMT techniques
-are unable to find
+Due to the universal quantifier, SMT solvers
+cannot find
 a model for $\varphi$. The signature $\extendsig{\Sigma}$ extends $\Sigma$ with the type
-$\fargtype{\con{s}}$ and the uninterpreted function symbol $\farg{\con{s}} : \fargtype{\con{s}}
+$\fargtype{\con{s}}$ and the %uninterpreted
+function symbol $\farg{\con{s}} : \fargtype{\con{s}}
 \rightarrow \typeint$. The result of $\conv( \varphi )$, after simplification,
@@
 We have developed a preliminary Isabelle frontend and are
-planning further frontends for Coq, HOL4, HOL Light,
+planning further frontends for Coq, %HOL4, HOL Light,
 the TLA\textsuperscript{+} Proof System, and possibly other proof assistants.
@@
 reasoning and to detect unsatisfiability in addition to satisfiability.
+We plan to integrate other tools as backends, to exploit the
+strengths of competing approaches.
 % and makes
@@
   \item[Type inference] infers types and checks definitions;
-  \item[Type skolemization]
-    replaces $\exists \alpha.\; \varphi[\alpha]$
-      with $\varphi[\tau]$, where $\tau$ is a fresh type;
+%  \item[Type skolemization]
+%    replaces $\exists \alpha.\; \varphi[\alpha]$
+%      with $\varphi[\tau]$, where $\tau$ is a fresh type;
   \item[Monomorphization]
@@
 are consistent (which can be checked syntactically in some cases),
-a contradiction imples the existence of countermodels.
+a contradiction imples the existence of countermodels. Compared with these
+approaches, the main advantage of our approach is that it can cope with
+underspecification and that it exploits the SMT solver (and its SAT solver) to
+enumerate candidate models efficiently.
 

### 2016-03-30 ijcar2016/body.tex
TOdid some more

@@
 standard notion of \emph{well-founded} function definitions, often called
-\emph{terminating} definitions in a slight abuse of terminology. In such
-definitions, all recursive calls are decreasing with respect to a well-founded
+\emph{terminating} definitions. % in a slight abuse of terminology.
+In such definitions, all recursive calls are decreasing with respect to a well-founded
 relation, which must be supplied by the user or inferred automatically
@@
   2 * \con{p}( x - 1 ) \bigr)$,
-where the theory $T$ is integer arithmetic extended with the %uninterpreted
-symbol $\con p:\typeint \rightarrow \typeint$, can be shown to be well founded
+where the theory~$T$ is integer arithmetic extended with %uninterpreted
+$\con p:\typeint \rightarrow \typeint$, can be shown to be well founded
 under a strategy that
@@
 ignoring the other branch. Logically, such dependencies can be captured by
-congruence rules. Krauss developed these ideas in the more general context of
-higher-order logic \cite[Section 2]{krauss-2009-phd}.
+congruence rules. Krauss developed these ideas in the general context of
+higher-order logic \cite[Section 2]{krauss-2009-phd}, where theories such as
+integer arithmetic can be axiomatized.
 
@@
 output \cite{turner-1995,mcbride-productive}.
-For example, given a type of infinite streams constructed by
+Given a type of infinite streams constructed by
 $\con{scons} : \ty{int} \times \ty{stream} \to \ty{stream}$,
@@
 falls within this class: Each call to $\const{e}$ produces one
-constructor before entering the nested call. Like terminating recursion,
-productive corecursion totally specifies the functions it defines.
-It is even possible to mix recursion and corecursion in the same function
-\cite{blanchette-et-al-2015-fouco}. Theorem~\ref{thm:adm} can be extended to
-cover such specifications, based on the observation that unfolding
-a corecursive definition infinitely computes a unique infinite object.
+constructor before entering the next call. Like terminating recursion,
+productive corecursion totally specifies the functions it defines,
+and the proof of Theorem~\ref{thm:adm} can be adapted to cover that case:
+
+\begin{theorem}\label{thm:adm-co}
+If $\DDD$ is a set of productive function definitions for\/
+$\sfundefs{\Sigma}$, then it is admissible.
+\end{theorem}
+
+%Theorem~\ref{thm:adm} can be extended to
+%cover such specifications, based on the observation that unfolding
+%a corecursive definition infinitely computes a unique infinite object.
 %We conjecture that a large fragment of mixed recursive--corecursive
 %specifications are
-
-\begin{rep}
-Beyond totality, an admissible set can contain underspecified functions
-such as $\forallf{\con{f}} x : \typeint.\allowbreak\;\, \con{f}( x )
-\teq \con{f}( x - 1 )$ or $\forallf{\con{g}} x.\allowbreak\;\, \con{g}( x
-) \teq \con{g}( x )$. We conjecture that one can ignore all
-tail-recursive calls (i.e., calls that occupy the right-hand side of the
-definition, potentially under some $\lite$ branch) when establishing well-foundedness
-or productivity, without affecting admissibility.
-\end{rep}
 
@@
 
+It is even possible to mix recursion and corecursion in the same total function
+\cite{blanchette-et-al-2015-fouco}.
+Beyond totality, an admissible set can contain underspecified functions
+such as $\forallf{\con{f}} x : \typeint.\allowbreak\;\, \con{f}( x )
+\teq \con{f}( x )$ or $\forallf{\con{g}} x.\allowbreak\;\, \con{g}( x
+) \teq \con{g}( x + 1 )$. The latter is problematic operationally, because
+in general the closure of a formula~$\psi$ that depends on some term
+$\con{f}(\con{a})$ is an infinite set
+$\{\psi\} \cup \{\con {g}(\con{a}+k) \teq \con {g}(\con{a}+k+1) \mid k \geq 0\}$.
+A similar issue arise with corecursive definitions specifying infinite
+\emph{acyclic} objects, such as the $\con{e}$ function introduced above.
+Nonetheless, admissibility is still useful if the problem does not refer to
+$\con{g}$ or $\con{e}$, because it tells us that we can safely ignore these
+functions altogether.
+%
+We conjecture that it is safe to ignore all tail-recursive calls (i.e., calls that
+occupy the right-hand side of the definition, potentially under some $\lite$
+branch) when establishing well-foundedness or productivity, without affecting
+admissibility.
+
 An example of an inadmissible set is
 $\{ \forallf{\con{f}} x : \typeint.\;\, \con{f}( x ) \teq \con{f}( x ) + 1 \}$,
-where $T$ is integer arithmetic extended to a set of uninterpreted symbols
+where~$T$ is integer arithmetic extended to a set of uninterpreted symbols
 $\{\con f, \con g :\typeint \rightarrow \typeint,\, \ldots \}$.
@@
 %
-Nunchaku transforms the specification of \verb$head$ into
+Nunchaku transforms the definition of \verb$head$ into
 %
@@
   \item[Elimination of (co)inductive predicates]
-    recasts a multiple-clause (co)inductive predicate specification
+    recasts a multiple-clause (co)inductive predicate definition
       into a recursive equation;

### 2016-03-30 ijcar2016/body.tex
shortened a bit

@@
 
+\let\oldcap=\cap
+\def\cap{\mathrel{\oldcap}}
+
 \let\oldchi=\chi
@@
 %\newcommand{\J}{\mathcalx{J}}
-\newcommand{\I}{\mathcal{I}}
+\newcommand{\I}{\mathcal{I\!}}
 \newcommand{\J}{{\mathcal{I}'}}
@@
 SMT solvers have recently been extended with techniques for finding models
-in the presence of universally quantified formulas in some restricted fragments
+of universally quantified formulas in some restricted fragments
 of first-order logic.
-This paper introduces a translation which reduces axioms specifying a large
-class of recursive functions, including well-founded (terminating) functions,
+This paper introduces a translation that reduces axioms specifying a large
+class of recursive functions, including terminating functions,
 to universally quantified formulas for which these
 techniques are applicable.
-An empirical evaluation confirms that the approach improves
+An evaluation confirms that the approach improves
 the performance of existing solvers on benchmarks from three sources.
@@
 This state of affairs is unsatisfactory, given the frequency of
-recursive definitions in practice, underscored by the recent
-addition of a %dedicated
-command for introducing them, \texttt{define-fun-rec},
-to the SMT-LIB standard \cite{smtlib25}.
+recursive definitions in practice.
+%, underscored by the recent
+%addition of a %dedicated
+%command for introducing them, \texttt{define-fun-rec},
+%to the SMT-LIB standard \cite{smtlib25}.
+
+\pagebreak[2]
 
@@
 
-The approach is implemented as a preprocessor in \cvc and in the Nunchaku
-model finder (Section~\ref{sec:implementations}).
+The approach is implemented as a preprocessor in the SMT solver
+\cvc and in a new higher-order model finder called Nunchaku
+(Section~\ref{sec:implementations}).
 %As a case study, we employed
@@
 each function symbol $\con{f} : \tau_1 \times \cdots \times \tau_n \rightarrow \tau$ in $\sfuns{\Sigma}$
-to a total function $\con{f}^\I : \tau_{\smash{1}}^\I \times \cdots \times \tau_n^\I \rightarrow \tau^\I$,
+to a total function $\con{f}^\I : \tau_{\smash{1}}^\I \times \cdots \times\allowbreak \tau_n^\I \rightarrow\allowbreak \tau^\I$,
 and each variable $x:\tau$ of $\svars{\Sigma}$ to an element of $\tau^\I$.
@@
 uninterpreted. Their
-semantics is the same as for standard quantified formulas $\forall \vec x.\; \varphi$.
+semantics is as for standard quantified formulas $\forall \vec x.\; \varphi$.
 Given $\con{f} : \tau_1 \times \cdots \times \tau_n \rightarrow \tau$,
-an annotated
-quantified formula $\forallf{\con{f}} \vec x.\; \varphi$ is a \emph{function definition}
+a formula $\forallf{\con{f}} \vec x.\; \varphi$ is a \emph{function definition}
 (\,\emph{for $\con{f}$}\vthinspace) if $\vec x$ is a tuple of variables
@@
 \emph{productive} corecursive functions. Corecursive functions are functions to
-a coalgebraic datatype. These functions can be ill founded, without
+a coalgebraic datatype. These functions can be ill founded without
 their being inconsistent. 
@@
 output \cite{turner-1995,mcbride-productive}.
-Given a type of infinite streams constructed by
-$\con{scons} : \ty{int} \times \ty{stream} \to \ty{stream}$,
+Given a type of infinite streams of integers constructed by
+$\con{scons} : \ty{int} \times \ty{stream} \to\allowbreak \ty{stream}$,
 the function defined by
@@
 
-It is even possible to mix recursion and corecursion in the same total function
-\cite{blanchette-et-al-2015-fouco}.
+It is even possible to mix recursion and corecursion in the same function
+\cite{blanchette-et-al-2015-fouco} while preserving totality and admissibility.
 Beyond totality, an admissible set can contain underspecified functions
@@
 A similar issue arise with corecursive definitions specifying infinite
-\emph{acyclic} objects, such as the $\con{e}$ function introduced above.
+\emph{acyclic} objects, such as the $\con{e}$ stream introduced above.
 Nonetheless, admissibility is still useful if the problem does not refer to
-$\con{g}$ or $\con{e}$, because it tells us that we can safely ignore these
-functions altogether.
+$\con{g}$ or $\con{e}$, because it tells us that we can safely ignore their
+definitions.
 %

### 2016-03-30 ijcar2016/body.tex
tuned; shortened

@@
 Nunchaku is a new higher-order model finder designed to be integrated with
-several proof assistants. The first version, 0.1, was released in January 2016
-with support for (co)algebraic datatypes, (co)recursive functions and
+several proof assistants. The first version was released in January 2016
+with support for (co)al\-gebraic datatypes, (co)recursive functions, and
 (co)inductive predicates. Support for higher-order functions is in the works.
-We have developed a preliminary Isabelle frontend and are
+We have developed an Isabelle frontend and are
 planning further frontends for Coq, %HOL4, HOL Light,
-the TLA\textsuperscript{+} Proof System, and possibly other proof assistants.
-
-Nunchaku is the spiritual successor to Nitpick for
-Isabelle/HOL~\cite{blanchette-nipkow-2010}, but is developed as a standalone
+the TLA\textsuperscript{+} Proof System, and other proof assistants.
+
+Nunchaku is a spiritual successor to Nitpick \cite{blanchette-nipkow-2010} for
+Isabelle/HOL, but is developed as a standalone
 OCaml program, with its own input language. Whereas Nitpick generates a
@@
 The problem defines a datatype (\texttt{nat}) and two mutually recursive
-inductive predicates (\texttt{even} and \texttt{odd}), it declares a constant
+inductive predicates (\texttt{even} and \texttt{odd}), declares a constant
 \texttt{m}, and specifies a goal to satisfy (``\texttt{m} is even and nonzero'').
@@
 goal: ``2 is even if 1 is odd, 1 is odd if 0 is even, and 0 is even.''
-Our experience with the Nitpick~\cite{blanchette-nipkow-2010} model finder
+Our experience with Nitpick
 is that users are mostly interested in the values assigned to uninterpreted
-constants (e.g., \texttt{m}). Occasionally, the models of recursive functions
-are instructive in the face of underspecification. A typical example is
+constants (e.g.,~\texttt{m}). Occasionally, the models of underspecified
+recursive functions are instructive. A typical example is
 the \texttt{head} function that returns the first element of a nonempty list:
@@
 
-Conceptually, the sequence of transformation is a \relax{two-way pipeline}
-built by composing pairs $(\textsf{encode},\vthinspace \textsf{decode})$ of
-transformations. For each such pair, \textsf{encode} transforms a problem over
-a signature $\Sigma$ in a logic~$\mathcal{L}$ to a problem over a signature
-$\Sigma'$ in a logic~$\mathcal{L}'$, and \textsf{decode} translates a model
+Conceptually, the sequence of transformation is a \relax{bidirectional pipeline}
+built by composing pairs $(\textit{Encode},\vthinspace \textit{Decode})$ of
+transformations. For each such pair, \textit{Encode} translates a
+$\Sigma$-problem in a logic~$\mathcal{L}$ to a $\Sigma'$-problem
+in a logic~$\mathcal{L}'$, and \textit{Decode} translates a model
 in~$\mathcal{L}'$ over $\Sigma'$ into a model in~$\mathcal{L}$ over $\Sigma$,

### 2016-04-10 ijcar2016/body.tex
Add paragraph to clarify Def 1/2. Address other minor reviewer comments. Minor fitting.

@@
 quantified formulas in SMT.
-Ge and de Moura \cite{GeDeM-CAV-09} introduced a complete instantiation-based
+Ge and de Moura \cite{GeDeM-CAV-09} introduced an instantiation-based
 procedure for formulas in the \emph{essentially uninterpreted} fragment.
@@
 $T$-equivalent to $\con{f}( \vec x ) \teq t$ for some term $t$ of type $\tau$.
+We will consider only annotated quantified formulas that are function definitions.
 %\ct{Why can't $\varphi$ be existential instead of just quantifier free?}
@@
 $(\forallf{\con{f}_1} \vec x_1.\; \varphi_1) \wedge \cdots \wedge
-(\forallf{\con{f}_n} \vec x_n.\; \varphi_n) \wedge \psi$,
+(\forallf{\con{f}_n} \vec x_n.\; \varphi_n) \wedge \phi$,
 %
@@
 % should be enough, unless I'm missing something? -ajr
-and $\psi$ contains no function definitions.
-We call $\psi$ the \emph{goal} of $\varphi$.
+and $\phi$ contains no function definitions.
+We call $\phi$ the \emph{goal} of $\varphi$.
 %\ct{Don't we need additional restrictions on the $\varphi_i$'s here?
@@
 \end{definition}
+
+In Definition 1, notice that $\phi$ is a formula possibly containing quantifiers.
+Given an admissible set of function definitions $\DDD$, 
+we may establish a model exists for $\DDD \wedge \phi$ 
+if we are able to find a ground formula $\psi_0$ that entails $\phi$,
+and subsequently extend $\psi_0$ to a $T$-satisfiable ground formula $\psi$ that is closed under function expansion with respect to $\DDD$.
+We may obtain $\psi$ from $\psi_0$ by conjoining to it ground formulas 
+(typically, conjunctions of ground literals) that entail instances of function definitions from $\DDD$ until the criterion in Definition 2 is met.
 
@@
 for which Ge and de Moura \cite{GeDeM-CAV-09} provide
-a complete instantiation procedure, implemented in \ziii.
+an instantiation procedure, implemented in \ziii.
 As expected, \cvc and \ziii run indefinitely on formula~(\ref{eq:ex-before}).
@@
 When reading this input,
-\cvc adds the annotated quantified formula
-$$\forallf{\con{s}} x.\;\, \con{s}( x ) \teq \lite\bigl( x \leq 0,\; 0,\; \con{s}( x-1 )\bigr)$$
+\cvc adds 
+$\forallf{\con{s}} x.\;\, \con{s}( x ) \teq \lite\bigl( x \leq 0,\; 0,\; \con{s}( x-1 )\bigr)$
 to its list of assertions,
 which after rewriting becomes
-$$\forallf{\con{s}} x.\;\, \lite\bigl( x \leq 0,\; \con{s}( x ) \teq 0,\; \con{s}( x ) \teq \con{s}( x-1 ) \bigr)$$
+$\forallf{\con{s}} x.\;\, \lite\bigl( x \leq 0,\; \con{s}( x ) \teq 0,\; \con{s}( x ) \teq \con{s}( x-1 ) \bigr)$.
 By specifying the command-line option \texttt{-}\texttt{-fmf-fun}, users
@@
 
-By default, \cvc uses techniques to minimize the number of literals it
-considers when constructing propositional satisfying assignments for
-formulas~\cite{relevancy2007}. However, we have found such techniques degrade
-performance for finite model finding on problems having recursive functions
-that are defined by cases. For this reason, we disable the techniques for
-problems produced from our translation.
+%By default, \cvc uses techniques to minimize the number of literals it
+%considers when constructing propositional satisfying assignments for
+%formulas~\cite{relevancy2007}. However, we have found such techniques degrade
+%performance for finite model finding on problems having recursive functions
+%that are defined by cases. For this reason, we disable the techniques for
+%problems produced from our translation.
 

### 2016-04-10 ijcar2016/body.tex
Add back paragraph

@@
 
-%By default, \cvc uses techniques to minimize the number of literals it
-%considers when constructing propositional satisfying assignments for
-%formulas~\cite{relevancy2007}. However, we have found such techniques degrade
-%performance for finite model finding on problems having recursive functions
-%that are defined by cases. For this reason, we disable the techniques for
-%problems produced from our translation.
+By default, \cvc uses techniques to minimize the number of literals it
+considers when constructing propositional satisfying assignments for
+formulas~\cite{relevancy2007}. However, we have found such techniques degrade
+performance for finite model finding on problems having recursive functions
+that are defined by cases. For this reason, we disable the techniques for
+problems produced from our translation.
 

### 2016-04-11 ijcar2016/body.tex
Fix in Section 3 about the E(\SIgma) theory T

@@
 
-\newcommand{\Mo}{{\mathcal{I}}}
+%\newcommand{\Mo}{{\mathcal{I}}}
+\newcommand{\Mo}{{\mathbf{I}}}
 
@@
 the \emph{models} of $T$, closed under variable reassignment
-(i.e., for every $I \in \Mo$, every $\Sigma$-interpretation that differs
-from $I$ only on the variables of $\svars \Sigma$ is also in $\Mo$).
+(i.e., for every $\I \in \Mo$, every $\Sigma$-interpretation that differs
+from $\I$ only on the variables of $\svars \Sigma$ is also in $\Mo$).
 A $\Sigma$-formula $\varphi$ is \emph{$T$-satisfiable}
@@
 if each $T$-entails the other.
-If $T_1 = (\Sigma_1, \Mo_1)$ is a theory and $\Sigma_2$ is a signature
-with $\sfuns{\Sigma_1} \cap \sfuns{\Sigma_2} = \emptyset$,
-the \emph{extension of $T_1$ to $\Sigma_2$} is the theory $T = (\Sigma, \Mo)$ where
-$\sfuns{\Sigma} = \sfuns{\Sigma_1} \cup \sfuns{\Sigma_2}$,
-$\stypes{\Sigma} = \stypes{\Sigma_1} \cup \stypes{\Sigma_2}$,
-and $\Mo$ is the set of all $\Sigma$-interpretations $\I$
-whose $\Sigma_1$-reduct is a model of $T_1$.
-We refer to the symbols of $\Sigma_2$ that are not in $\Sigma_1$ as \emph{uninterpreted}.
+If $T_0 = (\Sigma_0, \Mo_0)$ is a theory and $\Sigma$ is a signature
+with $\sfuns{\Sigma_0} \subseteq \sfuns{\Sigma}$ 
+and $\stypes{\Sigma_0} \subseteq \stypes{\Sigma}$,
+the \emph{extension of $T_0$ to $\Sigma$} is the theory $T = (\Sigma, \Mo)$ 
+where $\Mo$ is the set of all $\Sigma$-interpretations $\I$
+whose $\Sigma_0$-reduct is a model of $T_0$.
+Note that $T$ is a \emph{conservative} extension of $T_0$, in the sense that
+a $\Sigma_0$-formula is $T_0$-satisfiable if and only it is $T$-satisfiable.
+We refer to the symbols of $\Sigma$ that are not in $\Sigma_0$ as \emph{uninterpreted}.
 For the rest of the paper, \relax{we fix a theory $T = (\Sigma, \Mo)$
@@
 We present a method that constructs an extended signature
-$\extendsig{ \Sigma }$ and an $\extendsig{ \Sigma }$-formula $\varphi'$ such that
-$\varphi'$ is $T$-satisfiable if and only if the $\Sigma$-formula $\varphi$ is
-$T$-satisfiable---i.e., $\varphi$ and $\varphi'$ are \emph{equisatisfiable} (\emph{in $T$}).
+$\extendsig{ \Sigma }$ and an $\extendsig{ \Sigma }$-formula $\varphi'$ 
+that is $T'$-satisfiable if and only if $\varphi$ is $T$-satisfiable
+where $T'$ is the extension of $T$ to $\extendsig{ \Sigma }$---i.e., $\varphi$ and $\varphi'$ 
+are \emph{equisatisfiable} (\emph{in $T'$}).
+Since $T'$ is a conservative extension of $T$, for notational simplicity
+we will refer to it also as $T$ from now on.
 The idea behind this translation is to use an uninterpreted type $\alpha_\con{f}$

### 2016-04-11 ijcar2016/body.tex
minor edits to Section 1 and 2

@@
 Other syntactic criteria extend
-this fragment slightly, including cases when variables occur as arguments of
+this fragment slightly, including some cases when variables occur as arguments of
 arithmetic predicate symbols. Subsequently, Reynolds et al.\
@@
 whereas the finite model finding techniques %of Reynolds et al.\
-are not applicable for functions over infinite domains such as the integers or
+are not applicable in the presence of functions over infinite domains such as the integers or
 algebraic datatypes. A simple example where both approaches fail is
@@
 We present a method for translating formulas involving recursive function
-definitions into formulas where finite model finding techniques can be applied.
+definitions to formulas where finite model finding techniques can be applied.
 The definitions must meet a semantic criterion to be admissible 
@@
 effectiveness of the SMT solvers \cvc and \ziii in finding countermodels to
-verification conditions (Section~\ref{sec:evaluation}). Unlike earlier work,
-our approach relies on off-the-shelf SMT solvers
-(Section~\ref{sec:related-work}).
+verification conditions (Section~\ref{sec:evaluation}). 
+Note that, unlike earlier work (Section~\ref{sec:related-work}),
+our approach relies on off-the-shelf SMT solvers.
 
@@
 We write $t\typ{\tau}$ to denote %ct a metavariable ranging over
-terms of type~$\tau$ and $\terms( t )$ to denote the set of subterms in $t$.
+terms of type~$\tau$ and $\terms( t )$ to denote the set of subterms of $t$.
 Given a term $u$, a variable tuple $\vec x = (x_1^{\tau_1},\dotsc,x_n^{\tau_n})$ and 
@@
 \begin{rep}When applied to terms, the symbol $=$ denotes syntactic equality.\end{rep}
-
+When convenient, we will treat a finite set of formulas as the conjunction of its elements.
 %We write $\functerms^\Sigma( t )$ to denote the set of $f$-applications in $t$ such that $\con{f} \in \sfundefs{\Sigma}$.
@@
 A $\Sigma$-formula $\varphi$ is \emph{$T$-satisfiable}
-if it is satisfied by some interpretation in $\Mo$\begin{rep};
+if it is satisfied by some interpretation in $\Mo$,
+which we call a \emph{$T$-model of $\varphi$}\begin{rep},
 otherwise, it is \emph{$T$-unsatisfiable}\end{rep}.
@@
 We refer to the symbols of $\Sigma$ that are not in $\Sigma_0$ as \emph{uninterpreted}.
+
 For the rest of the paper, \relax{we fix a theory $T = (\Sigma, \Mo)$
-with uninterpreted symbols} constructed as above.
+with uninterpreted symbols}, constructed as above.
 
@@
 uninterpreted. Their
-semantics is as for standard quantified formulas $\forall \vec x.\; \varphi$.
+semantics is the same as for standard quantified formulas $\forall \vec x.\; \varphi$.
 Given $\con{f} : \tau_1 \times \cdots \times \tau_n \rightarrow \tau$,
@@
 the formula
-$\psi \wedge \bigwedge \Delta$ is also $T$-satisfiable.
+$\psi \wedge \Delta$ is also $T$-satisfiable.
 %\ct{I}
@@
 
-In Definition 1, notice that $\phi$ is a formula possibly containing quantifiers.
+In Definition 1, notice that the goal $\phi$ is a formula possibly containing quantifiers.
 Given an admissible set of function definitions $\DDD$, 
@@
 \begin{proofsketch}
-Let $\psi$ be a satisfiable formula closed under function expansion with
-respect to $\DDD$. We show that $\psi \wedge \bigwedge \Delta$ is also
-satisfiable. Let $\I$ be a model of $\psi$, and
+Let $\psi$ be a $T$-satisfiable formula closed under function expansion with
+respect to $\DDD$. We show that $\psi \wedge \Delta$ is also
+$T$-satisfiable. Let $\I$ be a $T$-model of $\psi$, and
 let $\I_0$ be the reduct of $\I$ to the function symbols in $\sfuns{\Sigma} \setminus
@@
 the interpretation of the functions they define, there exists a
-$\Sigma$-interpretation $\J$ that extends $\I_0$ such that $\J \models \Delta$.
+model $\J$ of $T$ that extends $\I_0$ such that $\J \models \Delta$.
 %
@@
 And since $\psi^\I$ does not depend on the interpretation at the other
-points, $\J$ is, like $\I$, a model of $\psi$.
-Since $\J \models \Delta$ by assumption, we have $\J \models \psi \wedge
+points, $\J$ is, like $\I$, a $T$-model of $\psi$.
+Since $\J \models \Delta$ by assumption, we have $\J \models \psi
 \bigwedge \Delta$ as desired.
@@
 \emph{productive} corecursive functions. Corecursive functions are functions to
-a coalgebraic datatype. These functions can be ill founded without
-their being inconsistent. 
+a coalgebraic datatype. These functions can be ill founded without being inconsistent. 
 Intuitively, productive corecursive functions are functions that
@@
 output \cite{turner-1995,mcbride-productive}.
-Given a type of infinite streams of integers constructed by
+For instance, given a type of infinite streams of integers constructed by
 $\con{scons} : \ty{int} \times \ty{stream} \to\allowbreak \ty{stream}$,
@@
 productive corecursion totally specifies the functions it defines,
-and the proof of Theorem~\ref{thm:adm} can be adapted to cover that case:
+and the proof of Theorem~\ref{thm:adm} can be adapted to cover that case.
 
@@
 in general the closure of a formula~$\psi$ that depends on some term
-$\con{f}(\con{a})$ is an infinite set
+$\con{f}(\con{a})$ is an infinite set:
 $\{\psi\} \cup \{\con {g}(\con{a}+k) \teq \con {g}(\con{a}+k+1) \mid k \geq 0\}$.
-A similar issue arise with corecursive definitions specifying infinite
+A similar issue arises with corecursive definitions specifying infinite
 \emph{acyclic} objects, such as the $\con{e}$ stream introduced above.
-Nonetheless, admissibility is still useful if the problem does not refer to
+Nonetheless, admissibility is still useful if a goal formula does not refer to
 $\con{g}$ or $\con{e}$, because it tells us that we can safely ignore their
@@
 closed under function expansion with respect to this set
-(since $\con{f}$ does not occur in $\ltrue$),
+(trivially, since $\con{f}$ does not occur in $\ltrue$),
 and yet there is no model of $T$
@@
 closed under function expansion \begin{rep}with respect to this set\end{rep}
-and yet there exists
-no interpretation satisfying both $\con{f}( 0 ) \teq 1$ and $\con{g}$'s
+but there is no interpretation satisfying both $\con{f}( 0 ) \teq 1$ and $\con{g}$'s
 definition.
@@
 development action ``Contre-exemples utilisables par Isabelle et Coq'' (CUIC).
-
+Reynolds and Tinelli are partially supported by grant 1228765 from the National Science Foundation.
 }

### 2016-04-12 ijcar2016/body.tex
A few more minor edits

@@
 $\extendsig{ \Sigma }$ and an $\extendsig{ \Sigma }$-formula $\varphi'$ 
-that is $T'$-satisfiable if and only if $\varphi$ is $T$-satisfiable
+that is $T'$-satisfiable if and only if $\varphi$ is $T$-satisfiable,
 where $T'$ is the extension of $T$ to $\extendsig{ \Sigma }$---i.e., $\varphi$ and $\varphi'$ 
 are \emph{equisatisfiable} (\emph{in $T'$}).
-Since $T'$ is a conservative extension of $T$, for notational simplicity
+Since $T'$ is a conservative extension of $T$, for simplicity
 we will refer to it also as $T$ from now on.
-The idea behind this translation is to use an uninterpreted type $\alpha_\con{f}$
+The idea behind this construction is to use an uninterpreted type $\alpha_\con{f}$
 to abstract the set of
-\emph{relevant} input tuples for each defined function $\con{f}$ and restrict the
+\emph{relevant} input tuples for each defined function $\con{f}$, and restrict the
 quantification of $\con{f}$'s definition to a single variable of this type.
@@
 satisfiability of $\varphi$.
-More precisely,
+
 %suppose our signature $\Sigma$ contains a set of defined function symbols $\sfundefs{\Sigma} \subseteq \sfuns{\Sigma}$.
+We construct the signature $\extendsig{\Sigma}$ so that,
 for each $\con{f} : \tau_1 \times \cdots \times \tau_n \rightarrow \tau \in \sfundefs{\Sigma}$,
-the extended signature $\extendsig{\Sigma}$ contains
+it contains
 %\begin{itemize}
 %\item
-an uninterpreted \emph{abstract type} $\fargtype{\con{f}}$ 
-abstracting the Cartesian product $\tau_1 \times \cdots \times \tau_n$
+an uninterpreted \emph{abstract type} $\fargtype{\con{f}}$,
+abstracting the Cartesian product $\tau_1 \times \cdots \times \tau_n$,
 and
@@
 
-The translation $\conv$ defined in Figure~\ref{fig:A} translates the $\Sigma$-formula
+The translation $\conv$ defined in Figure~\ref{fig:A} converts the $\Sigma$-formula
 $\varphi$ into the $\extendsig{\Sigma}$-formula $\varphi'$. It relies
 on the auxiliary function $\conv_0$, which takes two arguments:\ the term $t$
-to translate and a \relax{polarity}~$p$ for $t$, which is either $\ppos$, $\pneg$, or
+to translate and a \relax{polarity}~$p$ for $t$, which can be $\ppos$, $\pneg$, or
 $\pnone$.
@@
 
-If $t$ is an application of a predicate symbol $\boolop$, including the
+In the call $\conv_0(t, p)$, if $t$'s top symbol is a predicate symbol $\boolop$, including the
 operators ${\oldneg}$, ${\wedge}$, ${\vee}$, ${\teq}$, and ${\lite}$,
-$\conv_0$ calls itself recursively on the arguments $t_i$ and polarity $\pol(
-\boolop, i, p )$, with $\pol$ d\rlap{efined as}
+$\conv_0$ calls itself recursively on the arguments $t_i$ of $b$ and polarity 
+$\pol(\boolop, i, p )$ with $\pol$ defined as
 \[\pol( \boolop, i, p ) =
 \begin{cases}
-p & \text{if either $\boolop \in \{{\wedge}, {\vee}\}$ or $\boolop = \lite$ and $i \in \{2, 3\}$} \\[-\jot]
+p & \text{if $\boolop \in \{{\wedge}, {\vee}\}$, or $\boolop = \lite$ and $i \in \{2, 3\}$} \\[-\jot]
 \pnegate p & \text{if $\boolop = {\neg}$} \\[-\jot]
@@
 where $\pnegate p$ is $\pneg$ if $p$ is $\ppos$, $\ppos$
-if $p$ is $\pneg$, and $\pnone$ if $p$ is $\pnone$.
+if $p$ is $\pneg$, and $\pnone$ otherwise.
 The term $t$ is then reconstructed as $\boolop(t_1',\dotsc,t_n')$
@@
 $\conv_0$ conjunctively adds to $\boolop(t_1',\dotsc,t_n')$ the constraint~$\chi$ derived from
-the subterms and returns $\top$ as the constraint.
+the subterms, and returns $\top$ as the constraint.
 Dually, if $p$ is $\pneg$, it adds a
@@
 %ct  as in the positive case.
-It $p$ is $\pnone$, the constraint $\chi$ is
-returned to the caller.
-
-If $t$ is a function definition, % of a function in $\sfundefs{\Sigma}$,
-$\conv_0$ constructs a quantified formula over a single variable
-$a$ of type $\fargtype{\con{f}}$ and replaces all occurrences of $\vec x$ in
-the body of that formula with $\vecfarg{\con{f}}( a )$.
+It $p$ is $\pnone$, it returns the constraint $\chi$.
+
+If $t$ is a function definition $\forallf{\con{f}} \vec x.\; u$, % of a function in $\sfundefs{\Sigma}$,
+then $\conv_0$ recursively constructs a formula $u'$ from $u$,
+replaces all occurrences of $\vec x$ in $u'$ with $\vecfarg{\con{f}}( a )$
+where $a$ is single variable of type $\fargtype{\con{f}}$,
+and then quantifies $a$.
 (Since function definitions are top-level conjuncts, %by case analysis on the return values of $\conv$,
@@
 %
-If $t$ is an unannotated quantified formula, $\conv_0$ calls itself
-on the body with the same polarity; a quantifier is prefixed to the
-quantified formula and constraint returned by the recursive call.
-Otherwise, $t$ is either an application of an uninterpreted predicate symbol or a term
-of a type other than $\Bool$. Then, the returned constraint is a conjunction of
-formulas of the form $\exists a : {\fargtype{\con{f}}}.\;\, \vecfarg{\con{f}}(
+If $t$ is an unannotated quantified formula $\forall \vec x.\; u$, then $\conv_0$ calls itself
+on $u$ with the same polarity $p$ and returns the quantification over $\vec x$ 
+of the formula $u'$ and of the constraint $\chi$ returned by the recursive call.
+Finally, if $t$ is an application of an uninterpreted predicate symbol or a term
+of a type other than $\Bool$, $\conv_0$ returns $t$ together with a conjunction of
+constraints of the form $\exists a : {\fargtype{\con{f}}}.\;\, \vecfarg{\con{f}}(
 a ) \teq \vec s$ for each subterm $\con{f}( \vec s )$ of~$t$ such that $\con{f}
@@
 The definition of $\const{s}$ specifies that it returns the sum of all
-positive integers up to $x$. The formula $\varphi$ is in definitional form with
-respect to $\sfundefs{\Sigma}$
+positive integers up to $x$. The formula $\varphi$, which is in definitional form with
+respect to $\sfundefs{\Sigma}$,
 %\jb{correct? It said $\Sigma_u$, which was definitely wrong}
-and states that the sum of all
+states that the sum of all
 positive numbers up to some constant $\const{c}$ is greater than $100$. It is
-satisfiable with a model that interprets $\const{c}$ as $14$ or more.
+satisfied in a model of $T$ that interprets $\const{c}$ as $14$ or more.
 Due to the universal quantifier, SMT solvers
@@
 an instantiation procedure, implemented in \ziii.
-As expected, \cvc and \ziii run indefinitely on formula~(\ref{eq:ex-before}).
-However, they produce a model for~(\ref{eq:ex-after})
-within 100 milliseconds.\xend
+Both \cvc and \ziii run indefinitely on formula~(\ref{eq:ex-before}), as expected.
+However, they both produce a model for~(\ref{eq:ex-after}) within 100 milliseconds.\xend
 \end{example}
@@
 %
-In other words, $\const{s}$ is interpreted as a function mapping $x$ to the sum
+In other words, $\I$ interprets $\const{s}$ as a function mapping $x$ to the sum
 of all positive integers up to $x$ when $0 \leq x \leq 13$, and $105$
@@
 we write $\absconstraints( L )$ to denote the set of constraints that
-force the concretization functions to have the necessary elements in their
-range for determining the satisfiability of $L$ with respect to the function
+force the concretization functions to have enough elements in their range 
+to determine the satisfiability of $L$ with respect to the function
 definitions in the translation.
@@
 \begin{lemma}\label{lem:conv}
-Let $\psi$ be a formula not containing function definitions,
-and let $\I$ be an $\extendsig{\Sigma}$-interpretation whose $\Sigma$-reduct
-is a model of $T$.
-Then $\I$ satisfies $\conv( \psi )$ if and only if
-it satisfies $L \cup \absconstraints( L )$, where $L$ is a set of ground $\Sigma$-literals that $T$-entail $\psi$.
+Let $\psi$ be a $\Sigma$-formula not containing function definitions,
+and let $\I$ be an $\extendsig{\Sigma}$-model of $T$.
+Then, $\I$ satisfies $\conv( \psi )$ if and only if
+it satisfies $L \cup \absconstraints( L )$
+for some set $L$ of ground $\Sigma$-literals such that $L \models_T \psi$.
 %There exists a set of ground $\Sigma$-literals $L$ such that $L$ entails $\varphi$,
@@
 \begin{lemma}\label{lem:conv}
-If $\psi$ is a formula not containing function definitions, 
+If $\psi$ is a $\Sigma$-formula not containing function definitions, 
 then $\conv( \psi ) \models_T\psi$.
@@
 \begin{proofsketch}
-First, we show that if $\varphi$ is satisfied by an
-$\Sigma$-interpretation $\I$, then $\conv( \varphi )$ is satisfied by an $\extendsig{\Sigma}$-interpretation $\J$.
-Let $\J$ be the $\extendsig{\Sigma}$-interpretation that interprets all types $\tau \in \stypes{\Sigma}$ as~$\tau^\I$,
-all functions $\con{f} \in \sfuns{\Sigma}$ as $\con{f}^\I$,
-and for each function $\con{f} : \tau_1 \times \cdots \times \tau_n \rightarrow \tau$ in $\sfundefs{\Sigma}$,
+First, we show that if $\varphi$ is satisfied by a
+$\Sigma$-model $\I$ of $T$, then $\conv( \varphi )$ is satisfied by an $\extendsig{\Sigma}$-model $\J$.
+Given such a model $\I$, let $\J$ be the $\extendsig{\Sigma}$-interpretation that interprets all types $\tau \in \stypes{\Sigma}$ as~$\tau^\I$,
+all function symbols $\con{f} \in \sfuns{\Sigma}$ as $\con{f}^\I$,
+and for each $\con{f} : \tau_1 \times \cdots \times \tau_n \rightarrow \tau$ in $\sfundefs{\Sigma}$,
 interprets $\fargtype{\con{f}}$ as $\tau_1^\I \times \cdots \times \tau_n^\I$
@@
 First, we define a sequence of $\Sigma$-literals sets $L_0 \subseteq L_1 \subseteq \cdots$
-such that $\J$ satisfies $L_i \cup \absconstraints( L_i )$ for $i = 0, 1, \ldots\vthinspace$.
+such that $\J$ satisfies $L_i \cup \absconstraints( L_i )$ for $i \geq 0$.
 Since $\J$ satisfies $\conv( \varphi_0 )$,

### 2016-04-12 ijcar2016/body.tex
Fix lemma refs.

@@
 
-\begin{lemma}\label{lem:conv}
+\begin{lemma}\label{lem:conv2}
 If $\psi$ is a $\Sigma$-formula not containing function definitions, 
@@
 For any function symbol~$\con{f}$ and terms~$\vec t$, since $\varphi_{\con{f}}[ \vec t / \vec x ]$ does not contain function definitions,
-by Lemma~\ref{lem:conv},
+by Lemma~\ref{lem:conv2},
 $\conv( \varphi_{\con{f}}[ \vec t / \vec x ] )$ entails $\varphi_{\con{f}}[ \vec t / \vec x ]$.

### 2016-04-12 ijcar2016/body.tex
Restricted statement of Th. 8 and slightly changed notation for goals (for consistency )

@@
 $(\forallf{\con{f}_1} \vec x_1.\; \varphi_1) \wedge \cdots \wedge
-(\forallf{\con{f}_n} \vec x_n.\; \varphi_n) \wedge \phi$,
+(\forallf{\con{f}_n} \vec x_n.\; \varphi_n) \wedge \varphi_0$,
 %
@@
 % should be enough, unless I'm missing something? -ajr
-and $\phi$ contains no function definitions.
-We call $\phi$ the \emph{goal} of $\varphi$.
+and $\varphi_0$ contains no function definitions.
+We call $\varphi_0$ the \emph{goal} of $\varphi$.
 %\ct{Don't we need additional restrictions on the $\varphi_i$'s here?
@@
 
-In Definition 1, notice that the goal $\phi$ is a formula possibly containing quantifiers.
+In Definition 1, notice that the goal $\varphi_0$ is a formula possibly containing quantifiers.
 Given an admissible set of function definitions $\DDD$, 
-we may establish a model exists for $\DDD \wedge \phi$ 
-if we are able to find a ground formula $\psi_0$ that entails $\phi$,
+we may establish a model exists for $\DDD \wedge \varphi_0$ 
+if we are able to find a ground formula $\psi_0$ that entails $\varphi_0$,
 and subsequently extend $\psi_0$ to a $T$-satisfiable ground formula $\psi$ that is closed under function expansion with respect to $\DDD$.
@@
 If $\varphi$ is a\/ $\Sigma$-formula in definitional form with respect to\/
-$\sfundefs{\Sigma}$
-and the set of function definitions\/ $\DDD$ corresponding to\/
-$\sfundefs{\Sigma}$ is admissible, then\/ $\varphi$ and\/ $\conv( \varphi )$ are
+$\sfundefs{\Sigma}$,
+the set of function definitions\/ $\DDD$ corresponding to\/
+$\sfundefs{\Sigma}$ is admissible, 
+and the goal formula $\varphi_0$ of $\varphi$ is ground,
+then\/ $\varphi$ and\/ $\conv( \varphi )$ are
 equisatisfiable in $T$.

### 2016-04-12 ijcar2016/body.tex
Introduced restriction to "Herbrand" theories

@@
 if each $T$-entails the other.
+We call a theory a \emph{Herbrand} theory if 
+for every quantifier-free $\Sigma$-formula $\psi$ over the variables $\vec x$,
+$\{\psi[\vec t/\vec x] \mid t \text{ ground $\Sigma$-term}\} \models_T \forall \vec x.\; \psi$.
+
 If $T_0 = (\Sigma_0, \Mo_0)$ is a theory and $\Sigma$ is a signature
@@
 For the rest of the paper, \relax{we fix a theory $T = (\Sigma, \Mo)$
-with uninterpreted symbols}, constructed as above.
+with uninterpreted symbols}, constructed as above, and assume it is a Herbrand theory.
+While this is an actual restriction, it can be shown that the theories typically considered 
+in SMT are Herbrand.
 

### 2016-04-12 ijcar2016/body.tex
Last set of edits

@@
 The approach is implemented as a preprocessor in the SMT solver
-\cvc and in a new higher-order model finder called Nunchaku
+\cvc~\cite{barrett-et-al-2011} and in a new higher-order model finder called Nunchaku
 (Section~\ref{sec:implementations}).
@@
 and Isabelle/HOL, to demonstrate that this translation improves the
-effectiveness of the SMT solvers \cvc and \ziii in finding countermodels to
+effectiveness of two SMT solvers, \cvc  and 
+\ziii~\cite{de-moura-bjoerner-2008}, in finding countermodels to
 verification conditions (Section~\ref{sec:evaluation}). 
-Note that, unlike earlier work (Section~\ref{sec:related-work}),
+Unlike earlier work (Section~\ref{sec:related-work}),
 our approach relies on off-the-shelf SMT solvers.
@@
 if each $T$-entails the other.
-We call a theory a \emph{Herbrand} theory if 
+We call $T$ a \emph{Herbrand} theory if 
 for every quantifier-free $\Sigma$-formula $\psi$ over the variables $\vec x$,
@@
 and each $\fargx{\con{f}}{i}$ as the $i$\vvthinspace{th} projection on such tuples for $i = 1, \dotsc, n$.
-Since $\J$ satisfies $\varphi$, it satisfies a set of ground literals $L$ that entail $\varphi$.
+Since $\J$ satisfies $\varphi$ and $T$ is Herbrand, $\J$ satisfies a set of ground literals $L$ that entail $\varphi$.
 Furthermore, $\J$ satisfies \relax{every} constraint of the form $\exists a : {\fargtype{\con{f}}}.\;\, \vecfarg{\con{f}}( a ) \teq \vec t$,
-since by our construction of $\J$ there exists a value $v \in {\fargtype{\con{f}}}^\J$ such that $v = \vec t^\J$.
+since by its construction there is a value $v \in {\fargtype{\con{f}}}^\J$ such that $v = \vec t^\J$.
 %By case analysis it follows that $\conv_0$ maintains the invariant that
@@
 \verb|(declare-fun c () Int)| \\
-\verb|(assert (> (s c) 100))| \\
-\verb|(check-sat)|
+\verb|(assert (> (s c) 100))| % \\
+%\verb|(check-sat)|
 \end{quote}
 %
-When reading this input,
-\cvc adds 
+When asked to check the satisfiability of the assertions above,
+\cvc adds the formula
 $\forallf{\con{s}} x.\;\, \con{s}( x ) \teq \lite\bigl( x \leq 0,\; 0,\; \con{s}( x-1 )\bigr)$
@@
 can enable \cvc's finite model finding mode for recursive functions.
-In this mode, \cvc will replace its list of known
-assertions based on the $\conv$ translation before checking for satisfiability.
+In this mode, \cvc will rewrite the asserted formulas according to the $\conv$ translation 
+before checking for satisfiability.
 %If the functions provided in the input are admissible,
 %then a ``satisfiable" response from the solver implies that a model exists for the original formula.
-Accordingly, the solver will output the approximation of the interpretation it
+Accordingly, it will output the approximation of the interpretation it
 used for recursive function definitions.
-For the example above, it outputs a %partial
+For the example above, \cvc outputs a %partial
 model of~$\con{s}$ where only
@@
 $\conv(\varphi)$.
-Like other systems,
+As in other systems,
 the finite model finding capability of \cvc incrementally fixes bounds on the cardinalities of uninterpreted types
@@
 we implemented an optimization based on statically inferring \emph{monotonic} types.
-A monotonic type is one in which models can always be extended with additional
-elements of that type \cite{blanchette-krauss-2011,claessen-et-al-2011}.
+Intuitively, a type of a theory $T$ is monotonic if every model of $T$ can be extended with additional
+elements of that type and remain a model of $T$~\cite{blanchette-krauss-2011,claessen-et-al-2011}.
 %%% I find this part more confusing than helping, since we use monotonicity to merge
@@
 considers when constructing propositional satisfying assignments for
-formulas~\cite{relevancy2007}. However, we have found such techniques degrade
-performance for finite model finding on problems having recursive functions
-that are defined by cases. For this reason, we disable the techniques for
+formulas~\cite{relevancy2007}. However, we have found that such techniques degrade
+performance for finite model finding on problems with recursive functions
+defined by cases. For this reason, we disable the techniques for
 problems produced from our translation.
@@
 OCaml program, with its own input language. Whereas Nitpick generates a
-succession of problems where cardinalities of finite types grow at each step,
+succession of problems where the cardinalities of finite types grow at each step,
 Nunchaku translates its input to one first-order logic program that targets
@@
 
-The input syntax was inspired by that other systems based on higher-order
+The input syntax was inspired by that of other systems based on higher-order
 logic (e.g., Isabelle/HOL) and by functional programming languages (e.g.,
-OCaml). The following simple problem gives a taste of the syntax:
+OCaml). The following simple problem gives a taste of it:
 %
@@
 
-Among SMT solvers, we considered \ziii~\cite{de-moura-bjoerner-2008} and
-\cvc~\cite{barrett-et-al-2011}.
+Among SMT solvers, we considered \ziii and \cvc.
 \ziii runs heuristic methods for quantifier instantiation~\cite{MouraBjoerner07}
@@
 The results are summarized in Figures \ref{fig:sat} and \ref{fig:unsat}.
-Bold indicates the maximum of a row.
-The benchmarks and more detailed results are available online.%
+The bold font indicates the maximum value of a row.
+All the benchmarks and more detailed results are available online.%
 \footnote{\url{http://cs.uiowa.edu/~ajreynol/IJCAR2016-recfun}}
@@
 lead to a net gain of one~satisfiable benchmark each with respect to \cvcf.
-The performance of \ziiib for countermodels also improves dramatically, as it
+The performance of \ziiib for countermodels also improves dramatically, as \ziiib
 finds 134 more benchmarks to be satisfiable, including 5 that are not solved by \cvcf.
@@
 the configuration \cvcf, which is not tailored for handling unsatisfiable benchmarks,
-solves 270 \emph{unsat} benchmarks overall, which is more than both \cvcd and \ziii.
+solves 270 \emph{unsat} benchmarks overall, which is more than either \cvcd or \ziii.
 These results suggest that the translation does not degrade the performance
 of SMT solvers for unsatisfiable problems involving recursive functions,
-and instead often improve their performance. 
-These results suggest that it would be interesting to try this translation
+and in fact it often improves it. 
+They also suggest that it might be interesting to use this translation
 in Sledgehammer \cite{blanchette-et-al-2013-smt} and to try Nunchaku also as a
